### 2018-01-17 Chapters/Chapter01.tex
Classical thesis

%************************************************
%\chapter{An alternative structure to a blockchain: Confirmations without blocks}
%************************************************

\subimport{../blockchain-tangle/paper/}{content}

### 2018-01-17 Chapters/Chapter02.tex
Classical thesis

%*****************************************
%\chapter{Examples}\label{ch:examples}
%*****************************************

\subimport{../sdm-framework-paper/}{content}

### 2018-01-17 Chapters/Chapter03.tex
Classical thesis

%************************************************
%\chapter{Math Test Chapter}\label{ch:mathtest} % $\mathbb{ZNR}$
%************************************************

\subimport{../bass-extension/}{content2}

### 2018-01-17 FrontBackmatter/Bibliography.tex
Classical thesis

%********************************************************************
% Bibliography
%*******************************************************
% work-around to have small caps also here in the headline
\manualmark
\markboth{\spacedlowsmallcaps{\bibname}}{\spacedlowsmallcaps{\bibname}} % work-around to have small caps also
%\phantomsection
\refstepcounter{dummy}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{chapter}{\tocEntry{\bibname}}
\bibliographystyle{plainnat}
\label{app:bibliography}
\bibliography{Bibliography,../blockchain-tangle/paper/mybib,../sdm-framework-paper/mybib,../bass-extension/mybib}

### 2018-01-17 FrontBackmatter/Foreword.tex
Classical thesis

%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Prelude}{prelude}
\chapter*{Prelude, introduction, and discussion of a thesis proposal}
\bigskip

\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.

Though most fields existed prior to the war, after the war they were funded by the public as strategic pieces of the major nations arsenal against future conflagrations. One of the fields in question was that of Management Science (also called Operations Research in military circles, as researchers filled the ranks of planners of war operations). Management Science had started as an industrial field, in movements stemming from Taylor and the origin of the production line by Henry Ford.  That was the first moment in industry in which operations were systematically subject to some of the tools of science: measurement, experimentation, hypothesis-testing, statistics, mathematical optimization, etc.

This is from almost 100 years ago. Today the field has advanced to a great number of nations, and the amount of applications has grown explosively.  Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power at the hands of individuals.  Whereas, during the war, computations were mostly done by hand, it is not outlandish to say that this original field can be referred to, today, as computational management science.

Applied mathematics and computer science serve simultaneously as a theoretical foundation and the major tool available to the field.  Though this is a thesis proposal concerning business, in this document one should expect to find the language and nomenclature of mathematical modeling and computer science as our primary and most natural language.

This proposal will present a number of different topics for exploration. Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of a future thesis brought by three self-contained scientific papers. %---of the kind and quality that is expected in the major scientific journals of the field.

%Though work has already begun in all of these studies, at this moment, for the proposal, each study is offered under its own  level of maturity, ranging from quasi-ready for submission for publication to research at its embryonic phase.

The first paper lies at the intersection of computer science, record-keeping, banking \& finance, and economic inclusion. Our concern will be over a Bitcoin-like system that departs from the blockchain architecture. It has been recently proposed that Directed Acyclic Graphs could certify transactions in a blockchain-like manner.  We investigate some issues regarding this possibility, namely: (i) cryptographic security and game-theoretical attacks; (ii) scalability; (iii) self-governance of the system; (iv) appropriate incentive system to all participants.

%This first paper is the most immature one.  It is work being conducted.  The research questions are still under consideration. Some original theorems (yet to be proved) are presented.  
This first paper is work being conducted. The research questions are still under consideration. Some original theorems (yet to be proved) are presented.

%The second paper lies at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence.  Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics. We implement a RB-Complete\footnote{Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM. The maturity of this work is medium-high: The computer code has been used in a previous Ph.D. Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).  What remains to be done is a full description of the systems' architecture and a performance comparison between different setups.
The second paper lies at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence.  Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics. We implement a RB-Complete\footnote{Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM. The computer code has been used in a previous Ph.D. Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).  What remains to be done is a full description of the systems' architecture and a performance comparison between different setups.

The third paper lies at the intersection of Marketing, Diffusion of Technological Innovation, and modeling. The classic Bass model of diffusion of innovation is extended, in order to account for users who, after adopting the innovation for a while, decide to reject it later on (possibly bringing down the number of active users---which is impossible in Bass' original model). Four alternative mathematical models are presented and discussed with the Facebook users dataset. %This is the work that has received most energy so far, and is the most mature of the papers presented herein.  I believe it is almost ready for submission for publication.

%A fourth paper is still highly immature.  It concerns monetary inflation, and the demonstrations that an approach through fixed effects cannot correct for the real cumulative inflation rates---and may be disastrous when referring to regions undergoing high inflation.  Though part of the mathematical analysis is mature, the work itself is lacking the necessary references to show its relevance:  I still need to review the literature and point out the rather large body of it that is committing a serious error. Because I have not decided whether this work should be part of the thesis, it will be presented in appendix 1.

A note on the overall format of this Thesis Proposal:  many journals, such as Nature, have a series of papers departing from the standard layout of (i) introduction, (ii) literature review, (iii) methodology, (iv) experiments, (v) discussion and (vi) conclusion.  Instead, these journals will generally present (i) introduction, (ii) discussion and conclusion, and---sometimes in smaller fonts---, the materials and methods and results section. Here, content is brought to focus, in detriment to form.  As I believe that the content of the studies should be the major focus of analysis for this proposal, I have also collapsed the introductory and concluding remarks into these few comments.  I hope readers will agree that, at this point, it is of overwhelmingly greater importance to query the scientific content of the proposal, even if in detriment to its overall form.

It is my hope that readers of this proposal will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---the proposed body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.

### 2018-01-17 Chapters/Chapter01.tex
Copy required tex content

@@
 
-\subimport{../blockchain-tangle/paper/}{content}
+\chapter{Introduction}
+
+Bitcoin (BTC) is the first digital currency, also known as digital money, internet money and cryptocurrency.
+It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank.
+Bitcoin is distributed since its ledger is public and is stored in thousands of computers.
+It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community.
+The security of Bitcoin relies on digital signature technology, so the owner of some BTC tokens is, in fact, the owner of a pair of private and public keys.
+One may either use the private key to transfer some tokens to another, or use the public key to a proof-of-ownership.
+If one looses their private key, unfortunately, access to their coins will be lost forever.
+
+Since 2009, Bitcoin has been growing and becoming widely used all around the world.
+It started as an experiment based in a seminal work from \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a \$36 billion market capitalization \citep{coinmarketcapbtc}.
+There are hundreds of companies investing in different uses of the technology, from exchanges to debit cards, and billions of dollars being invested in the new markets created by Bitcoin.
+
+The main problem when one is trying to create a digital money is how to prevent double spending.
+As the money is digital, what would prevent counterfeiting?
+That is, what would prevent users from sending copies of the same money to two (or more) people?
+That is exactly the problem solved by Bitcoin and its underlying blockchain technology.
+The current solution behind fiat money is having a single issuer, a central bank, and trusting the financial institutions.
+
+Despite Bitcoin's huge success, there are still many challenges to be overcome.
+The network must scale to support hundreds of transactions per second, but today its capacity is close to 8 transactions per second.
+Another important challenge is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
+
+Bitcoin has the most decentralized network between the cryptocurrencies, even so there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
+Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behaviour, and it is seen as a problem to be solved.
+
+According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending.
+The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the e-cash literature.
+
+The number of ideas and publications focusing in improving Bitcoin's design is increasing every day.
+Many of these proposals are organized into BIPs (Bitcoin Improvement Proposals) which are discussed and implemented by the community; while others come in the form of whitepapers and alternative software forks (which would include the need of a protocol upgrade).
+Other scaling proposals are published in blogs and forums, describing new cryptocurrencies.
+Bitcoin's community hardly ever publishes their ideas in academic journals.
+
+% Spam problem.
+
+%The DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015} has a different design which has neither mining nor confirmation blocks and transaction fees.
+Each transaction has its own proof-of-work, and is used to confirm other transactions in a directed acyclic graph.
+Although this design has been used by a few cryptocurrencies, none of them has enough users and would strengthen that the problems have been solved.
+
+The DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015} has different design which has neither mining nor confirmation blocks and transaction fees.
+Each transaction has its own proof-of-work, and is used to confirm other transactions in a directed acyclic graph.
+The present work intends to analyze whether the DAG model may be a viable solution to both Bitcoin's scaling problem and Bitcoin's oversized influence of a few miners and mining pools.
+
+\chapter{Bitcoin \& Blockchain}
+
+In \citeauthor{nakamoto2008bitcoin}'s (2009) seminal paper, there is no distinction between bitcoin and blockchain.
+They are just one thing which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money in the internet, in which all users can agree as to whom is entitled to which funds.
+
+But, in practice, it is interesting to separate these concepts.
+Bitcoin uses the blockchain technology to create a distributed ledger, while the blockchain is a technology which allows information to be stored in an immutable and distributed way.
+
+The blockchain technology works through the creation of new blocks.
+Each new block confirms that all the previous blocks are valid and have not been tampered with.
+The mechanism that assures the immutability is proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network.
+
+The proof-of-work is a mathematical problem with the following characteristics: (i) it is hard to find a solution; (ii) this difficulty level may be adjusted; and (iii) it is fast to check whether the proposed solution is correct.
+
+Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given number.
+As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}.
+Thus, if the given number is $2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero).
+But if the given number is $2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero).
+Hence, finding a solution is a hard problem which difficulty depends on the given number.
+The lesser the given number, the higher the difficulty.
+On the other side, checking whether a solution is correct is fast, one just has to apply the SHA-256 twice and compare.
+
+By design, Bitcoin's blockchain proof-of-work difficulty is dynamically adjusted every 2016 blocks to keep an average pace of 10 minutes between block creation.
+Thus, the goal is to adjust the difficulty every 14 days.
+If it takes less than 14 days to find 2016 blocks, that means the network's hash power has increased, thus the difficulty is increased.
+If it takes more than 14 days to find 2016 blocks, it means the network's hash power has decreased, thus the difficulty is decreased.
+
+When miners are finding a solution to a new block, they are mining or working in the new block.
+A block is found when a solution to the proof-of-work is found.
+Two miners may find blocks in a small interval of time.
+In this case, both will propagate their blocks and the network will randomly choose one of them as the next block.
+This phenomenon is called a fork.
+Thus, when the next block is found, one of those blocks will be confirmed and the other will be ignored and referred to as an orphan block.
+The network agrees that work should be done in the block at the longest chain in the blockchain.
+When a new block is found, it indirectly confirms all the previous blocks in the chain and their transactions.
+
+Newly propagated blocks are validated by the Bitcoin's network.
+If the solution of the proof-of-work is incorrect or if any transaction included in the block has any issue, then the block is discarded.
+In order to work properly, the whole network must agree in what is allowed and what is not.
+Should one think that something should be allowed and accept it in their blocks, the remaining of the network will discard their newly propagated blocks.
+That is why Bitcoin's network is distributed and decentralized.
+Everything depends on the agreement of the network, or, precisely, the agreement of the owners of at least 50\% of the hash power.
+Even if the remaining 49\% disagrees, the 50\% or more who agree will generate, on average, more blocks than the remaining of the network and their rules will prevail on the longest chain.
+If a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork (in general, a hard-fork relaxes the constraints, while the latter hardens them).
+
+For instance, no group with more than 50\% of the network's hash power agreed into increasing the maximum block size to increase the number of transactions confirmed by a block, and thus increasing the network's capacity.
+Thus, the capacity remains the same and the community has been discussing the issue in search of a consensus.
+
+Bitcoin uses the blockchain technology to create a distributed ledger.
+It allows every new block to generate new bitcoins and also to collect the fees from the confirmed transactions within the block.
+Bitcoin's transactions have two main parts: (i) inputs, and (ii) outputs.
+Each transaction sends bitcoins from one or more input addresses to one or more output addresses.
+In order to prove that one is the owner of the input bitcoins, one must digitally sign the transaction proving such ownership.
+
+Bitcoins (BTCs) owned by someone are, in fact, unspent outputs in one or more transactions.
+For instance, one may have 6 BTCs spread between three transactions' unspent output: the first with 1 BTC, the second with 2 BTCs, and the third with 3 BTCs.
+
+In a transaction, the inputs are pointers to other transactions' outputs (which they are spending).
+A transaction output may only be spent once and thus may not be partially spent.
+For instance, when one has 3 BTCs in one transaction output and would like to send 1 BTC to a friend, they have to create a transaction with one input spending the 3 BTCs and two outputs, one for the friend with 1 BTC and one's change with 2 BTCs.
+This enables $O(1)$ unspent transaction lookup.
+
+Each transaction's output has a script that is executed by the miners to check whether one has or has not permission to spend that output.
+In other words, whether one has the ownership of that output.
+In order to execute these scripts, the miners also need some data.
+This data is given by the transaction which is spending the output.
+
+The output's scripts usually checks whether the public key is valid and whether the digital signature was signed by the private key associated to that public key.
+Although there are only 3 commonly used scripts, one may create a custom script using the Bitcoin's script language \footnote{Your courageous author once tried to make a transaction with custom script to try to double spend a deposit to an exchange, only to learn through this intrepid adventure that Bitcoin allows only 3 script patterns and the others are treated as invalid.}.
+
+The input contains the data which prove that the sender is the owner of the referred outputs, i.e., the input which must be accepted by the output scripts being spent.
+Usually, each input has the public key of the sender and a digital signature.
+
+Those users accustomed to block explorers may have been misled by the transaction information that these websites provide.
+For example, suppose a miner receives a transaction with ``one input from address A1''.
+This ``one input'' actually consists of a pointer to a previous unspent output (i.e., there is no ``input'' address, as is displayed, but only a pointer).
+This pointer reference allows lookup to be executed in $O(1)$ time.
+
+After lookup, the miner knows how many BTC tokens are available at that unspent output.
+But, in order to certify ownership of that output, the miner receives instructions in the form of a script with the rules that lead to the desired unspent output address (and this one is displayed as the input address by those websites).
+Because this process requires a digital signature, only the holder of the corresponding private key is able to sign such a transaction.
+
+\chapter{DAG model}
+
+DAG model proposes a whole different approach to confirmations.
+It proposes that there is no need for a block to confirm transactions, as transactions can confirm themselves.
+Here, each transaction has its own proof-of-work, named weight, and they must confirm two other, previous, transactions.
+Hence, each transaction has an accumulated weight, which is the accumulated proof-of-work that has confirmed it so far.
+In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph, as in Fig.
+\ref{fig-tangle-example}.
+
+Like the Blockchain, the DAG model is another technology to store immutable data and may be the underlying technology to different applications, such as cryptocurrencies, digital contracts (Ethereum-like), digital notaries, and so forth.
+The main question which this work proposes to answer is: Is it reliable to use DAG model?
+In which conditions?
+
+The transactions may be either confirmed or unconfirmed.
+The confirmed transactions have been already confirmed by at least one more transaction.
+It does not mean they are already irreversible and protected against a double spend attack --- it just means at least one transaction has done some work to confirm it.
+The unconfirmed transactions are called tips and they are eager to be confirmed.
+Usually a new transaction selects two tips to confirm, but this rule may not be followed.
+
+Transactions may have any format, including Bitcoin-like transactions---with inputs, outputs, and scripts.
+But it also may be completely different, just like in the cryptocurrency Iota \citep{tangle2016}.
+This work will not discuss details of the transactions' format, because we understand that it does not affect exactly the network scalability and security.
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{./images/fig-tangle-example.pdf}
+\caption{White nodes represent transactions that have been confirmed at least once.
+Green circles represent unconfirmed transactions (tips).
+Gray and dashed nodes are the transactions currently solving the proof-of-work in order to be propagated.\label{fig-tangle-example}}
+\end{figure}
+
+The accumulated weight of a transaction A is the sum of all weights of the transactions which confirm A, including A itself, i.e., $w_A + \sum_{A \leadsto P} w_P$.
+For example, in Fig.
+\ref{fig-tangle-example}, the accumulated weight of transaction 3 is the sum of the weights of the transactions 5, 6, 7, and 8.
+The accumulated weight may be interpreted as how hard it is to rollback a transaction.
+It is analogous to the number of confirmations of a block in Bitcoin.
+
+The score of a transaction A is the sum of all weights of the transactions which are being confirmed by A, including A itself, i.e., $w_A + \sum_{P \leadsto A} w_P$.
+For example, in Fig.
+\ref{fig-tangle-example}, the score of transaction 3 is the sum of the weights of the transactions 1 and 2.
+It is a measure of how much proof-of-work has been done before confirming this transaction.
+It may indicate whether the confirmed transactions have received enough attention (and hashpower) or whether it may have received tangential attention.
+
+The height of a transaction A is the length of the longest path from transaction A to the genesis transaction.
+For example, in Fig.
+\ref{fig-tangle-example}, the height of transaction 5 is four ($5 \rightarrow 3 \rightarrow 2 \rightarrow 1 \rightarrow$ genesis).
+It may be interpreted as the “`age'' of the transaction.
+The lower the height, the older the transaction.
+
+The depth of a transaction A is the length of the longest path in the inverted graph from transaction A to any unconfirmed transaction (tip).
+For example, in Fig.
+\ref{fig-tangle-example}, the depth of transaction 2 is three ($2 \rightarrow 3 \rightarrow 5 \rightarrow 8$).
+It is the opposite of the height.
+It may be interpreted as the youth of the transaction.
+The lower the depth, the younger the transaction.
+When a new transaction is confirming two transactions with high depth, it is referred as lazy transaction.
+
+An important factor of DAG model is how the new transactions choose which transactions they will confirm.
+There are several possible approaches, such as randomly selecting two of the unconfirmed transactions (tips).
+In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7.
+It may be on purpose, or maybe transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction.
+The selection algorithm seems to be important to protect the network against double spend attacks.
+
+The higher the volume of new transactions, the more unconfirmed transactions will appear.
+In Fig.
+\ref{fig-tangle-swarm}, the reader can notice that the number of new transactions was increased for a while, and then decreased back to the original value.
+The DAG has behaved well when exposed to a high load scenario, since it reduced the number of tips to only three.
+It is like a moving swarm which gets wider when the number of new transactions increases and gets thinner when the number of new transactions decreases.
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{./images/fig-tangle-swarm.pdf}
+\caption{Suddenly the number of transactions per second increases and the width of the swarm grows.
+After a while, the number of transactions per second decreases and the width of the swarm shrinks.\label{fig-tangle-swarm}}
+\end{figure}
+
+Conflicting transactions may happen when two or more transactions try to spend the ``same money'' --- or, in the Bitcoin's transaction format, try to spend the same output.
+In this case, the network must choose which of the transactions will be accepted and the other one will be invalidated, even when both have already been confirmed.
+In fact, when one transaction is invalidated, the whole sub-DAG which confirms it is also invalidated.
+In this case, it may happen to reverse some transactions.
+
+Intuitively, when there is a conflict, the network should accept the transaction which has greater accumulated weight, invalidating the others (see Fig.
+\ref{fig-tangle-conflict}).
+But it may be not enough to prevent a nuclear submarine attack.
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{./images/fig-tangle-conflict.pdf}
+\caption{The red nodes are transactions which had some conflict with previous transaction and were invalidated by the network. Notice that none of them have been confirmed.\label{fig-tangle-conflict}}
+\end{figure}
+
+\section{Nuclear submarine attack}
+
+The nuclear submarine attack, also known as the parasite chain attack, is when the attacker generates a separate DAG (or a side DAG), with many transactions and a lot of proof-of-work.
+This side DAG is off the network, i.e., its transactions have not been propagated.
+Then, at a convenient moment, the attacker suddenly propagates these transactions.
+The whole network needs to decide how to handle these transactions.
+
+If the transactions have no conflict with any transaction of the main DAG, i.e., there is no transaction spending the ``same money'', then it seems easy to handle the transactions.
+But, as it is an attack, probably there will be some conflicts, and it is not easy to choose which transaction should be invalidated.
+As the attacker has been generating a separate DAG, the conflicting transaction may have an accumulated weight similar or greater than the transaction in the main DAG.
+Thus, using only the accumulated weight may not be enough to prevent this attack.
+
+For example, the attacker may generate a transaction which transfers all their funds to another address.
+Then, they start to generate many new transactions which confirms themselves and even confirms some of the transactions in the main DAG.
+But none of these transactions are propagated to the network.
+Then, the attacker buys something in the real world, pays with cryptocurrency, and wait until the payment gets the accumulated weight demanded by the merchant.
+Finally, the attacker suddenly propagates all the transactions to the network in a small window of time.
+This may cause the network to accept the attacker's original transaction instead of the one used to pay the merchant.
+Hence, the merchant transaction is invalidated, and the double spend attack has succeeded.
+
+\section{Proposal: Questions to be explored}
+
+Given these preliminaries, these are some questions with which we will be concerned.
+
+In this paper, we will analyze the network scaling and security.
+How a DAG network scales, simulating different loads and measuring the bandwidth, computational effort, and storage space necessary to handle all the transactions in time.
+What is the minimum transaction's aggregated weight which it may be considered unlikely to be reversed?
+
+We are interested in how the rate of new transactions affects how long it takes to a new transaction to be confirmed for the first time.
+We had already run some simulations under normal load (Fig.
+\ref{fig-tangle-hist}) and high load (Fig.
+\ref{fig-tangle-hist-2}).
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{./images/fig-tangle-hist.png}
+\caption{Histogram of how long has been a transaction waiting until its first confirmation.
+It was a simulation of 15 minutes with new transactions rate changing between 1 and 15 tx/s.\label{fig-tangle-hist}}
+\end{figure}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{./images/fig-tangle-hist-2.png}
+\caption{Histogram of how long has been a transaction waiting until its first confirmation.
+It was a simulation of 5 minutes with new transactions rate of 50 tx/s, i.e., very high load.\label{fig-tangle-hist-2}}
+\end{figure}
+
+Besides the time to the first confirmation, it is also important to measure how long it takes to a new transaction reach a specific accumulated weight.
+It is useful for exchanges and merchants to set their minimum requirements.
+Bitcoin's exchanges and merchants usually requires a minimum of 6 confirmations blocks.
+
+As the DAG network is distributed, its users may use different parameters --- they cannot violate the network's rules, but they may use different tip selection algorithms, for instance.
+We would like to explore how different users' parameters may affect the network, how much is necessary to cause either a hard-fork or a soft-fork.
+
+\cite{tangle2016} suggests that constraining transactions' weight in a range would both prevent spam, because it would be necessary a minimum work to propagate a new transaction, and attacks, because an attacker would not be allowed to propagate a transaction with very high weight.
+We agree with the spam argument, and would like how a minimum weight would affect mobile devices's new transactions.
+We partially agree with the upper bound, because an attacker would be able to generate many transaction with lower weight.
+We will analyze if constraining transactions' weight would be effective against nuclear submarine attacks.
+
+We have another suggestion to prevent nuclear submarine attacks: set a maximum depth for the transactions confirmed by a new transaction.
+So, new transactions would have to confirming newer transactions instead of old transactions, and a nuclear submarine attack would be controlled because the attacker would be not be able to create a large separate DAG.
+But the maximum depth rule would possibly also affect transaction created by low (hash)power devices, because it would take a longer time to solve the proof-of-work and, if the confirmations are going fast, the transaction would possibly be invalidated.
+This raises another important question: what would be an optimal maximum depth allowed?
+
+If we would like to include fee transactions, how would it be distributed between confirmations?
+As transactions may have multiple direct confirmations, it is not obvious who would receive the transaction fee.
+An intuitive suggestion would be to give the fee to the first confirmation with a minimum weight, but it might have some propagation problems, because it takes a while to propagate a transaction in the network and there would be conflict.
+
+In order to solve the fee distribution problem, we will simulate how the network would behave if we include blocks.
+These blocks would be like Bitcoin's, with an adjustable proof-of-work according to the network's hashpower.
+They would receive the fees from all confirmed transactions which had not been confirmed by any block before, and they would be able to generate new coins.
+These blocks would also help solving other problems.
+
+%\begin{enumerate}
+%\item Which parameters turns the network more or less vulnerable to attacks?
+%\item Is there an optimal strategy to attack?
+%\item Could new coins be generated demanding a proof-of-work with dynamic difficulty, like in Bitcoin?
+%\item What is the equation of number of tips over time?
+What is the equation of how long it takes to a transaction to be confirmed for the first time?
+What is the equation of the accumulated weight of a transaction over time?
+Some equations have already been proposed by \citet{tangle2016}.
+%\end{enumerate}
+
+\chapter{Methodology}
+
+The methodology we have been using is computer simulation.
+Through the simulation of many scenarios of DAG model with different parameters, we will understand how the network behaves in complex scenarios, mainly when the load is suddenly increased, and when the network is under attack.
+We will first test our hypothesis through simulation and than prove them mathematically.
+
+The simulator has been developed using an event-based design which is capable of running hours of simulation in just a few minutes.
+It creates agents who decide to make a transaction, then they select which transactions will be confirmed, then they spend some time working in the proof-of-work, and, finally, they propagate the transaction to the network.
+The other agents receive the transaction and may accept or deny it.
+The agents may use different parameters among themselves.
+
+When a new transaction is added to the DAG, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
+The topological sort is used to calculate the longest path from all other transactions to the new transaction, and this longest path is used to update the depth of the whole DAG.
+If the new transaction confirms only transactions which has already been confirmed, the depth update is skipped thanks to Theorem \ref{theorem-new-tx-not-tip}.
+
+The necessary time to generate the proof-of-work is estimated by the function $k \cdot 2^{weight}$, where $k$ is a constant related to the computing power of the agent and weight is randomly chosen.
+
+The time between two new transactions is sampled from the exponential distribution with parameter $\lambda$.
+The value of $\lambda$ is changed over time, causing changes in the load of the network.
+We consider 1 tx/s to be a low load, and 15 tx/s to be a high load.
+Even though we have already tested up to 100 tx/s.
+
+The simulator will output different reports: (i) snapshots of the DAG at interesting moments, such as Fig.
+\ref{fig-tangle-example}, \ref{fig-tangle-swarm}, and \ref{fig-tangle-conflict}; (ii) histogram of confirmation time, such as Fig.
+\ref{fig-tangle-hist} and \ref{fig-tangle-hist-2}; and many others.
+
+\section{Proposal: Theorems to be proved}
+
+The following theorems will be mathematically proved. They have already been used in the implementation of the simulator.
+
+\begin{theorem}
+\label{theorem-new-tx-not-tip}
+When a new transaction confirms an already confirmed transaction, the depth of all transactions remains the same, i.e., the depth of the transactions are only changed when a new transaction confirm an unconfirmed transaction.
+\end{theorem}
+
+\begin{theorem}
+The height of a transaction is equal to the maximum height of its parents plus one.
+\end{theorem}
+
+\begin{theorem}
+If new transactions choose randomly the unconfirmed transactions to be confirmed, then no unconfirmed transaction will be left behind, i.e., all transactions will be confirmed in due time.
+\end{theorem}
+
+\chapter{Conclusion}
+
+Bitcoin's underlying technology blockchain has been called by many as a major invention comparable to the invention of the internet.
+But it is unlikely that Bitcoin and blockchain have achieved the final or most optimal design for a secure and scalable electronic transaction system.
+In this work, we will investigate whether DAG model is suitable to be a real alternative to blockchain.
+
+Preliminary results obtained from our simulator have shown that, using a random tip selection strategy, the network seems to support up to 50 transactions per second, with no tip abandoned (Fig.
+\ref{fig-tangle-swarm}) or taking too long to have at least one confirmation (Fig.
+\ref{fig-tangle-hist-2}).
+Today, Bitcoin network can barely handle 8 transactions per second without increasing the unconfirmed transaction list to hundreds of thousands --- several transactions take days to be confirmed.
+
+We have not yet tested the security of the network against double spend attacks.
+It will be one of the next steps and we expect to block a nuclear submarine attack including a rule for invalidate new conflicting transactions which confirm transactions with depth greater than a given threshold.
+We are unsure whether this rule would invalidate legitimate transactions on a high load scenario.
+
+We also plan to improve the analysis of the confirmation time.
+So far, we have the histograms of how long it takes to a transaction to be confirmed for the first time (Fig.
+\ref{fig-tangle-hist} and \ref{fig-tangle-hist-2}).
+We will also generate the histograms of how long it takes to a transaction to have an accumulated weight of at least a given number.
+It is an important histogram because it helps merchants and users to decide how long they should wait to consider the transaction almost irreversible.
+
+We will also work on the mathematical modeling of DAG networks, and the equations will help us to understand the limits of these networks.

### 2018-01-17 Chapters/Chapter02.tex
Copy required tex content

@@
-%*****************************************
-%\chapter{Examples}\label{ch:examples}
-%*****************************************
+\chapter{Introduction}
 
-\subimport{../sdm-framework-paper/}{content}
+Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+It has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
+Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
+Thus, even though great results have been achieved using SDM, it requires great effort of researchers to improve someone's work.
+
+It is our belief that this tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python.
+Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
+This new tool provides a ready to use system in which experiments can be executed almost as soon as they are designed.
+
+Our motivation was our own effort in order to run our models.
+As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug free.
+Thus, we had to deviate from our main goal --- which was to test our models --- and to focus in the implementation itself.
+Furthermore, new members in our research group had to go through different source codes developed by former members.
+
+The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has automatic tests, (iv) is cross-platform, and (v) is easily extended to fulfill other research models.
+Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier to readers to replicate any previous results and easily understand the source-code of the model.
+
+\chapter{Sparse Distributed Memory}
+
+\input{sdm2.tex}
+
+\chapter{Framework Architecture}
+
+The framework implements the basic operations in a Sparse Distributed Memory which may be used to create more complex operations.
+It is developed in C language --- which may be loaded in many platforms and programming languages --- with a wrapper in Python.
+The Python module makes it easy to create and execute operations in a Sparse Distributed Memory.
+
+We split the SDM memory in two parts: the hard-location address space and the hard-location counters.
+Thus, the addresses (bitstrings) of the hard-locations are stored in one array, while their counters in another.
+This may be useful to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be same in all of them.
+As the slowest part of reading and writing operations is scanning the address space, the performance benefits are high.
+
+Each part may be stored either in the RAM memory or in a file.
+The RAM memory is interesting for quick experiments, automated tests, and others scenarios in which the SDM may be lost.
+The file is interesting for a long-term SDM --- one may create an SDM file with 10,000 random writes, which will be copied over and over to run multiple experiments.
+The file may also be sent to another researcher or may be published with the paper to let others run their own checks and verify the results.
+In summary, the framework fits many different uses and necessities.
+
+There are some other advantages splitting the address space and the counters.
+When the scanner is run in the CPU, the accessed addresses in the RAM are continuous which improves considerably performance.
+When the scanner is run in the GPU, only the address space must be uploaded to the GPU, reducing the required memory.
+
+Let a SDM memory with $N$ dimensions and $H$ hard-locations.
+Then, in a 64-bit computer, the array of hard-location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard-location counters.
+For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard-locations, using 16-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 1.9 GB of memory.
+
+Basic operations were grouped in four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs).
+Operations include creating new bitstrings, flipping bits, generating bitstring with a specific distance from a given bitstring, scan the address space using different algorithms, write a bitstring to a counter, write in an SDM, read from an SDM, and iteratively read from an SDM until convergence.
+
+The reading and writing operations are executed in two steps: first, the address space is swept looking for the activated addresses; then, the operation is performed in the counter.
+Reading operation assemblies the bitstring according to the counters of the activated addresses, while the writing operation changes the counters.
+
+The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner.
+The linear scanner runs in a single core and is the slowest one.
+It was developed only for testing purposes.
+The thread scanner runs at the CPU in multiple threads (and our recommendation is to use the number of threads equals to twice the number of cores).
+The OpenCL scanner runs at the GPUs and support multiple devices.
+The speed of a scan depends on the processors and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark test.
+
+The hamming distance between two bitstrings is calculated counting the number of ones in the XOR of the bitstrings, i.e., $d(x, y) = \text{number of ones in} x \oplus y$.
+In order to count the number of ones of the XOR, the framework uses a pre-calculated table of distances --- by default, it is used a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM.
+Finally, to calculate the distance between to bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$.
+As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$).
+
+Internally, the address space is stored in an array of addresses, which, in turn, is stored in an array of 64-bit integers (Fig.
+\ref{tab:hl-addresses-detail}).
+As the addresses and their 64-bit integers are contiguous in RAM or file, there is a performance gain since they are loaded in chunks.
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[
+mycell/.style={draw, minimum size=7mm},
+matrixA/.style={matrix of nodes,
+    nodes={mycell, anchor=center},
+    column sep=-\pgflinewidth, 
+    row sep=-\pgflinewidth, 
+    },
+matrixB/.style={matrix of nodes,
+    nodes={mycell, anchor=center},
+    column sep=-\pgflinewidth, 
+    row sep=-\pgflinewidth, 
+}]
+
+\matrix[matrixA] (A) { addr$_1$ & addr$_2$ & addr$_3$ & $\cdots$ & addr$_H$ \\ };
+
+\matrix[matrixB, below=of A] (B) {
+addr$_{k, 1}$ & addr$_{k, 2}$ & addr$_{k, 3}$ & $\cdots$ & addr$_{k, 8 \cdot \lceil N/64 \rceil}$ \\
+};
+
+\draw[dashed] (A-1-1.south west)--(B-1-1.north west);
+\draw[dashed] (A-1-1.south east)--(B-1-5.north east);
+\draw [
+	thick,
+    decoration={
+        brace,
+        mirror,
+		amplitude=0.2cm,
+        raise=0.2cm
+    },
+    decorate
+] (B-1-1.south west) -- (B-1-5.south east)
+node [pos=0.5,anchor=north,yshift=-0.5cm] {N bits};
+
+\end{tikzpicture}
+
+\caption{Hard-location addresses are stored in a contiguous array.
+In a 64-bit computer, each hard-location address is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil N/64 \rceil$.\label{tab:hl-addresses-detail}}
+\end{figure}
+
+\section{Proposal:  Framework Validation}
+
+In order to check whether the framework works as expected and according to the model proposed by \citep{Kanerva1988}, we run simulations which were compared to the theoretical analysis conducted some decades ago.
+
+The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
+This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve.
+
+One particular analysis of interest is that of the distance read at a point $\alpha$.
+Suppose an SDM is trying to read an item written at $\alpha$, but the cues received so far lead to a point of distance $d$ from $\alpha$.
+As one reads at $\alpha+d$, a new bitstring $\beta$ is obtained, leading to our question: what is the new distance from $\alpha$ to $\beta$?
+Is it smaller or larger than $d$?
+That, of course, depends on the ratio between $d$ and the number of dimensions of the memory.
+As we have found out, there are some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
+
+\subsection{Some initial anomalous results}
+
+As we ran the simulations reflecting some of Kanerva's graphs, one in particular struck our attention:  The new distances obtained after a read operation were not perfectly predicted by the theoretical model, and we propose that this is due to interaction effects between different attractors.
+
+\citet{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Fig.
+\ref{kanerva-table-7-2}).
+The original prediction considered that the read distance would decline when inside the critical distance in increase afterwards, converging to a \textasciitilde 500-bit distance.
+At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance point.
+
+\begin{figure}[h]
+\centering\includegraphics[width=0.5\textwidth]{images/kanerva-table-7-2-original.png}
+\caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
+\label{kanerva-table-7-2}}
+\end{figure}
+
+Our preliminary results show that the theoretical prediction is not accurate.
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Fig.
+\ref{sdm-10000w-table-7-2}).
+Our results were obtained using a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, the same used by Kanerva.
+
+\begin{figure}[h]
+\centering\includegraphics[width=0.5\textwidth]{images/sdm-10000w-table-7-2.png}
+\caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
+Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
+\label{sdm-10000w-table-7-2}}
+\end{figure}
+
+But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Fig.
+\ref{sdm-100w-table-7-2}).
+This result strengthen our theory that the disparities in the computational results are due to the interaction effect of different attractors.
+
+\begin{figure}[h]
+\centering\includegraphics[width=0.5\textwidth]{images/sdm-100w-table-7-2.png}
+\caption{Results generated by the framework similar from Kanerva's original Table 7.2. It was a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
+\label{sdm-100w-table-7-2}}
+\end{figure}
+
+To obtain the results from Fig.
+\ref{sdm-10000w-table-7-2} and \ref{sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
+Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
+Thereby, in order to show the interaction effects more clearly, we wrote a handmade bitstring to the SDM which had all bits inverted in relation to the origin bitstring --- their hamming distance was equal to 1,000.
+Our handmade bitstring was acting as an opposite attractor, and one can see the accelerating effects towards convergence to both attractors: the origin and the handmade bitstrings (Fig.
+\ref{sdm-10000w-notX-table-7-2}).
+Here we had the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
+
+\begin{figure}[h]
+\centering\includegraphics[width=0.5\textwidth]{images/sdm-10000w-notX-table-7-2.png}
+\caption{This graph shows the interaction effects more clearly.
+As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and the opposite.
+Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
+\label{sdm-10000w-notX-table-7-2}}
+\end{figure}
+
+Obviously, these small deviations from Kanerva's original theoretical predictions deserve a qualification.
+Kanerva was working in the 1980s and the 1990s, and had no access to the immense computational power that we do today.
+It is no surprise that some small interaction effects should exist as machines allow us to explore the ideas of his monumental work.
+
+\section{Proposal: Performance Results}
+
+Our intention is to provide comparative performance metrics under different computation engines (CPU, GPU, etc) and different operating systems (Linux, MacOs, Windows, etc).
+Performance can be measured as the average number of scans of all hard locations per second, reads per second, writes per second, etc.
+
+Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+We also intend to test on machines such as the iMac with dedicated GPU, MacPro with dedicated GPU, and personal computers under Linux with dedicated GPUs.
+
+Beyond that, we are running as state-of-the-art devices: (i) an Amazon EC2 p3.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.8xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and 8x NVIDIA K80 GPU.
+
+\chapter{Conclusion}
+Sparse Distributed Memory is a viable model of human memory, yet it does require researchers to (re-)implement a number of parallel algorithms in different architectures.
+
+We propose to provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypothesis and test them computationally through minimal effort.
+Though the framework is not well-documented for public release at this time, it has already served as the backbone of Chada's Ph.D.
+thesis.
+The single-line command ``pip install sdm'' will install the framework on posix-like systems, and single-line commands will let users test the framework, generate some of the figures from Kanerva's theoretical predictions in their own machines, and --- if interested enough ---, test their own theories and improve the framework in an open-source fashion.
+It is our belief that such work is a necessary component towards accelerating research in this promising field.
+
+% TODO Include in the to do list simulations based on the suggestion of Murilinho.

### 2018-01-17 Chapters/Chapter03.tex
Copy required tex content

@@
-%************************************************
-%\chapter{Math Test Chapter}\label{ch:mathtest} % $\mathbb{ZNR}$
-%************************************************
+% TODO Model 4: Linear analysis
+% TODO Facebook vs Princeton
 
-\subimport{../bass-extension/}{content2}
+\chapter{Introduction}
+
+The way innovations diffuse in the market is an important and useful topic in marketing, which was made popular through Rogers' work \citep{rogers1962diffusion} and has influenced many marketing researchers.
+Within innovation diffusion, \citet{bass1969} proposed a model which forecasts how many people will have adopted new products or technologies by a given point in time.
+INFORMS members have voted this model as one of the Top 10 Most Influential Papers published in the 50-year history of Management Science in connection with the 50th anniversary of the journal \citep{bass2004comments}.
+
+The \citet{bass1969} model was designed to forecast only innovation adoption, which is the first time one consumes the innovation.
+In the model, either the consumer has or has not consumed the innovation by a given point in time.
+Thus, recurrent customers are considered only once, because they have already consumed the innovation before.
+The model has only three parameters, which are estimated through the number of adoptors of the innovation.
+Though very simple, it is considered very robust.
+
+Although it is clear that the marketing investment, the product prices, the economy itself, and many other variables affect the diffusion process, the \citet{bass1969} model does not contemplate these variables and yet it is still able to describe the empirical adoption curve of a large number of new products and technological innovations.
+In order to explain the robustness of the model, \citet{bass1994bass} developed a general model including these variables, and they showed that this general model reduces to the Bass model as a special case.
+They also showed that the shape of the diffusion of innovation process is always the same, an S-curve.
+\citet{norton1987diffusion} analyzed the diffusion of innovation for product substitution, explaining some unexpected changes in innovation adoption which were still unclear.
+
+The motivation behind the present work is that there may be people who reject a particular innovation.
+Such people do not recommend the innovation, on the contrary, they may publicly complain about it and bad-mouth it.
+This negative word-of-mouth effect of rejection has always existed \citep{richins1983negative, bone1995word, smith1995effects, buttle1998word} but it is becoming more and more important as information can spread more easily and faster among people through the usage of new technologies \citep{jansen2009twitter, godes2004using, babic2015effect}.
+Nowadays, before making a decision, people may search the internet for reviews and feedbacks about the innovation, and what they find affects their decision \citep{chen2004impact, dellarocas2003digitization, duan2008online, dellarocas2007exploring, lee2009electronic, pfeffer2014understanding}.
+A number of firms have already perceived this change caused in large part by the social media and have adapted to this new condition, like Starbucks \citep{gallaugher2010social}, for example.
+%Quali east2008measuring,
+
+There is an extensive literature on innovation diffusion.
+Numerous extensions to the \citet{bass1969} model have been proposed (for reviews, see \citet{meade2006modelling} and \citet{peres2010innovation}), among which Mahajan and colleagues were the first to propose a model that includes the negative word-of-mouth \citep{mahajan1984introduction}.
+The latter and other extensions that include negative word-of-mouth are much more complex than the \citet{bass1969} model, and their parameters must be estimated using the number of people who have already adopted the innovation at a given time.
+A problem is that most recent innovations, like Facebook, Twitter, and Netflix do not disclose their total number of users.
+Actually, they deem this number confidential.
+They only disclose the number of active users, not including the number of users who have rejected them.
+
+In this work, I propose an extension to the Bass model which (i) includes the negative effect of the rejections, (ii) is as simple as the \citet{bass1969} model, and (iii) its parameters can be estimated using the number of people who have adopted and have not posteriorly rejected the innovation, i.e., the number of \textit{active adopters}.
+An important difference between the number of \textit{active adopters} and the \textit{total adopters} is that first may decrease over time, while the latter cannot.
+First, I discuss the \citet{bass1969} model and its parameters; then I describe the model extension and analyze four models of rejection; next I detail the estimation method; after that I estimate the models by using Facebook's active users dataset; and finally, I discuss the results and conclude.
+
+The main contribution of this work is that the proposed extension is the first to include the effect of the rejections that can be estimated using the number of \textit{active adopters} instead of the number of \textit{total adopters}.
+The model can be applied to forecast the number of active adopters of these companies in the next quarters, which becomes an important tool for analysts, investors, and the companies themselves.
+As the number of active adopters is related to the market cap of these companies, these forecasts may be useful to estimate the future value of the firms.
+At the time of this writing, Facebook's market cap is \$222.69B, Twitter's is \$23.18B, and Netflix's is \$20.49B\footnote{Data obtained from Yahoo!
+Finance website at December 1\textsuperscript{st}, 2016.}.
+
+\chapter{The Bass model}
+
+The \citet{bass1969} model is a simplification of the diffusion of innovation process proposed by \citet{rogers1962diffusion}.
+The Rogers' classification of adopters has five classes: (i) innovators; (ii) early adopters; (iii) early majority; (iv) late majority; and (v) laggards.
+Bass simplified them to only two classes: innovators and imitators.
+The innovators are the ones who start using an innovation regardless of who else and how many other people are already using it.
+The imitators are the ones who concern themselves about who is using the innovation and, as long as many other people are using it, they are more inclined to adopt it.
+Thus, at the beginning of the diffusion of innovation, the majority of adopters are innovators.
+As more and more people adopt the innovation, the majority of new adopters shift to imitators.
+
+Mathematically, the model presents itself with the following set of equations:
+
+\begin{align}
+S(t) &= mf(t) \\
+Y(t) &= mF(t) \\
+\frac{f(t)}{1-F(t)} &= p + qF(t), F(0)=0 \label{eq:bass:ode}
+\end{align}
+
+Both $S(t)$ and $Y(t)$ are related to the absolute amount of adopters; while $S(t)$ is the number of new adopters at time $t$, $Y(t)$ is the number of people who had already adopted the innovation by time $t$, thus $S(t) = \frac{d}{dt} Y(t)$.
+The model also has three positive parameters: (i) the potential market size $m$; (ii) the innovators parameter $p$; and (iii) the imitators parameter $q$.
+There are also $f(t)$ and $F(t)$ which are related to the percentage of the potential market: $f(t)$ is the percentage of the potential market which is adopting the innovation at time $t$, while $F(t)$ is the percentage of the potential market which had already adopted it at time $t$, thus $f(t) = \frac{d}{dt} F(t)$.
+
+Therefore, $S(t)$ and $f(t)$ are related to the adoption rate at time $t$, while $Y(t)$ and $F(t)$ are related to the accumulated adopters at time $t$.
+In the beginning of the diffusion of innovation, there are no adopters at all, thus $Y(0) = F(0) = 0$.
+As the number of adopters can only increase over time, both $Y(t)$ and $F(t)$ are monotonically increasing functions, and both $S(t)$ and $f(t)$ are always greater or equal to zero.
+
+The non-linear ordinary differential equation \ref{eq:bass:ode} is the main equation in the Bass model.
+Its left side is known as the hazard function and it expresses the probability of someone adopting the innovation, provided that he/she has not chosen to adopt it yet, i.e., the rate of adoption.
+Its right side means that this probability is at least $p$ and increases linearly with the percentage of people who have already adopted the innovation, i.e., $F(t)$.
+
+Solving the differential equation, Bass found a closed formula for the diffusion, which is $F(t) = \frac{ 1 - e^{-(p+q)t} }{ 1 + \frac{q}{p} e^{-(p+q)t} }$.
+This closed formula always has the famous shape of the S-curve, regardless of the values of $p > 0$ and $q > 0$.
+From the closed formula of $F(t)$, it is trivial to obtain the equations of $S(t)$, $Y(t)$, and $f(t)$.
+
+The potential market size $m$ is the unknown number of people who will have adopted the innovation after a very long time.
+It is not exactly the target market of the innovation, but a subset of it, as no product difuses over its entire target market.
+If a company estimates and updates the model more than once for their product, the change in $m$ is a change in the potential market and could help the company to understand whether their decisions in the meantime have increased or decreased the number of future adopters.
+As, $\lim_{t \rightarrow \infty} Y(t) = m$, the whole potential market will have adopted the innovation at some point.
+
+The innovator parameter $p$ is related to the proportion of people in the potential market who adopt the innovation regardless the others.
+In other words, their decision to adopt is not influenced by the social system, but by other external factors.
+The bigger the $p$, the larger the number of innovators, thus the faster the diffusion at the beginning.
+
+The imitator parameter $q$ is related to the influence of those actually using the innovation on those who are not using it yet.
+This is why this parameter multiplies $F(t)$, which is the proportion of the market which had already adopted it.
+This influence is mainly understood as a result of the word-of-mouth recommendation.
+In other words, the more people use the innovation, the more other people will adopt it.
+The bigger the $q$, the larger the imitator effect, thus, the faster the diffusion.
+
+Practitioners have been using this model to forecast future demand.
+First, they measure the number of adopters over time.
+Then, they estimate the parameters $m$, $p$, and $q$.
+Finally, they extrapolate $S(t)$ out of the measured time window and use its value as the forecast demand.
+They also calculate $m - Y(t)$ as a forecast of how many people have not adopted the innovation yet.
+
+\chapter{The extended model}
+
+I propose adding a new term in the differential equation \ref{eq:bass:ode} in order to include the effect of rejection, as in equation \ref{eq:mymodel}.
+Hereafter I will refer to: (i) the people who have adopted the innovation as \textit{total users}; (ii) the people who have adopted the innovation and remain using it as \textit{active users}; and (iii) the people who rejected the innovation as \textit{inactive users}.
+Clearly, the function $Y(t)$ is the number of \textit{total users}, which is equal to the sum of the number of \textit{active users} with the number of \textit{inactive users}.
+
+\begin{align}
+\frac{f(t)}{1-F(t)} &= p + qF(t) - wR(t), F(0)=0, R(0)=0 \label{eq:mymodel}
+\end{align}
+
+The function $R(t)$ is the percentage of accumulated \textit{inactive users} at time $t$.
+Thus, $A(t) = F(t) - R(t)$ is the percentage of accumulated \textit{active users} at time $t$.
+Multiplying by $m$, $mA(t) = Y(t) - mR(t)$ is the number of \textit{active users} at time $t$.
+Since $A(t) \ge 0$, $F(t) \ge R(t)$, which makes sense because it is not possible to have more \textit{inactive users} than \textit{total users}.
+
+The negative word-of-mouth parameter $w$ is related to how much the \textit{inactive users} really affect the new adopters decision in the diffusion process.
+The bigger the $w$, the greater the negative influence of these \textit{inactive users} on the new adopters.
+Another possible understanding is that $w$ is related to how much the \textit{inactive users} are bad-mouthing the innovation.
+The bigger the $w$, the more they bad-mouth the innovation.
+
+Equation \ref{eq:mymodel} could be rewritten as $\frac{f(t)}{1-F(t)} = p + qA(t) + (q-w)R(t)$.
+This form is useful in order to understand the impact of \textit{active users} and \textit{inactive users} on the rate of adoption at time $t$.
+
+If $w = q$, then the rate of adoption increases linearly with the \textit{active users}, since $\frac{f(t)}{1-F(t)} = p + qA(t)$.
+In other words, the positive word-of-mouth has exactly the same influence as the negative word-of-mouth on the new adopters.
+It also means that the rate of adoption is always greater than zero, thus the whole potential market will have adopted the innovation sooner or later.
+
+If $w < q$, then the new adopters are more influenced by the number of \textit{total users} than by the number of \textit{inactive users}.
+It is just as if the \textit{inactive users} do not bad-mouth the innovation so much, or at all.
+Again, the rate of adoption is always greater than zero, thus the whole potential market will have adopted the innovation sooner or later.
+
+If $w > q$, then the proposed extension really differs from the Bass model.
+In this case, the influence of the \textit{inactive users} is greater than the influence of \textit{total users}.
+So, if the rate of adoption were equal to zero ($wR(t) = p + qF(t)$), the innovation might not be adopted by the whole potential market, i.e., $\lim_{t \rightarrow \infty} F(t) < 1$.
+
+The main contribution of this work is to have the choice to use the number of \textit{active users}, which is the information that most of the companies disclose, in order to estimate the parameters of the proposed extension model.
+With that, both the number of \textit{inactive users} and \textit{total users} could be forecast.
+It is important to notice that the number of \textit{active users} ($mA(t)$) could decrease over time.
+In fact, forecasting when this is going to happen may be crucial for corporations.
+
+\chapter{Models for $R(t)$}
+
+The proposed extended model already includes the effect of rejection through the $R(t)$ function and the $w$ parameter.
+In order to complete the model, $R(t)$ has to be well defined.
+Let $r(t) = \frac{d}{dt} R(t)$ be the rate of new \textit{inactive users} at time $t$.
+The proposed differential equations for $R(t)$ are the following:
+
+% XXX
+% Model 1: r(t) = \nu [f(t) - r(t)] ???
+% Model 2: r(t)/[1-R(t)] = \nu [f(t) - r(t)] ???
+% I guess solving both equations is easy and the conclusions would be the same.
+
+\begin{align}
+	\text{Model 1: }& r(t) = \nu f(t) \label{eq:model1} \\
+	\text{Model 2: }& \frac{r(t)}{1-R(t)} = \nu f(t) \label{eq:model2} \\
+	\text{Model 3: }& r(t) = \nu [F(t) - R(t)] \label{eq:model3} \\
+	\text{Model 4: }& \frac{r(t)}{1-R(t)} = \nu [F(t) - R(t)] \label{eq:model4}
+\end{align}
+
+% TODO Discuss why it does not make sense to have a constant in the models.
+% TODO Include the state diagram (1-F(t)), A(t), and R(t) with the transitions.
+
+These four models can be grouped in two families, one for the equations \ref{eq:model1} and \ref{eq:model2}, and another for the equations \ref{eq:model3} and \ref{eq:model4}.
+The former relates the rejection to the rate of new people adopting the innovation, as if people decide whether they will use or reject the innovation when they try it.
+Then, they do not change their position anymore.
+The latter assumption relates the rejection to the number of \textit{active users}, as if the \textit{active users} first adopt the innovation and then they continuously reject it.
+
+The rejection parameter $\nu$ has a different interpretation in each family.
+In equations \ref{eq:model1} and \ref{eq:model2}, it is the proportion of new adopters who will reject the innovation.
+In equations \ref{eq:model3} and \ref{eq:model4}, it is the proportion of active users who are continuously rejecting the innovation.
+
+Therefore, for all these models of rejection, the complete diffusion of innovation model has five parameters to be estimated, namely $m$, $p$, $q$, $w$, and $\nu$.
+
+\section{Model 1}
+
+In this model, the rate of new \textit{inactive users} at time $t$ is proportional to the percentage of people adopting the innovation at time $t$, as if people decide whether they will use or reject the innovation when they are adopting it.
+
+The differential equation \ref{eq:model1} can be easily solved integrating both sides.
+Thus, $R(t) = \nu F(t)$, $A(t) = (1 - \nu) F(t)$, and $f(t) / [1 - F(t)] = p + (q - w \nu) F(t)$.
+As the imitator coefficient must be positive, we must have $w \nu < q$.
+
+The solution shows that this model has exactly the same explanatory power as the Bass model, neither better nor worse.
+This happens because the model's solution has exactly the same equation after the linear transformation $q^* = q - w \nu$.
+
+As $\lim_{t \rightarrow \infty} F(t) = 1$, thus $\lim_{t \rightarrow \infty} R(t) = \nu$. Hence, the proportion of \textit{inactive users} is exactly equal to the rejection parameter.
+
+Solving the differential equation for $F(t)$, it gets $F(t) = (1 - e^{-(p+q-w\nu)t})/(1 + \frac{(q-w\nu)}{p} e^{-(p+q-w\nu)t})$.
+Finally, as $R(t) = \nu F(t)$, thus $R(t) = \nu (1 - e^{-(p+q-w\nu)t})/(1 + \frac{(q-w\nu)}{p} e^{-(p+q-w\nu)t})$.
+
+Unfortunately, it is not possible to estimate this model.
+The problem is that $\forall \nu \in \mathbb{R}^+$, $\exists \hat{\nu} \in \mathbb{R}^+$ such as the set of parameters $(m, p, q, w, \nu)$ and $(m, p, q, w\nu/\hat{\nu}, \hat{\nu})$ have exactly the same residuals when estimated.
+That is, the model can be estimated for any value arbitrarily set for $\nu$.
+Intuitively, as both $F(t)$, $A(t)$, and $R(t)$ have the same shape, the parameters can be estimated with an empirical \textit{active users} dataset and then you can slide up or down $F(t)$ just changing the values of $\nu$ and $w$.
+
+This result is interesting because it shows that the Bass model can already explain the diffusion of innovations which follows this model of rejection.
+Hence, it just confirms the robustness of the Bass model.
+
+\section{Model 2}
+
+The right side of the differential equation \ref{eq:model2} is the rate of rejection, i.e., the probability of someone who rejecting the innovation, provided that he/she has not rejected it yet.
+Thus, in this model, the rate of rejection is proportional to the percentage of people adopting the innovation at time $t$, i.e., the more people adopt the innovation, the more they reject it.
+But if no one is adopting, there would be no rejection also, which would hold the number of \textit{active users} the same.
+
+The differential equation \ref{eq:model2} can also be solved for $R(t)$ algebrically.
+Using the fact that $-\frac{d}{dt} \log[1 - R(t)] = r(t) / [1 - R(t)] = \nu f(t)$, and integrating both sides of this equation yields:
+
+\begin{align}
+	- \frac{d}{dt} \int_0^t \log[1 - R(\tau)] d\tau &= \nu \int_0^t f(\tau) d\tau \\
+	- \log[1 - R(t)] &= \nu F(t) \\
+	1 - R(t) &= e^{-\nu F(t)} \\
+	R(t) &= 1 - e^{-\nu F(t)}
+\end{align}
+
+Finally, $f(t) / [1 - F(t)] = p + qF(t) - w[1 - e^{-\nu F(t)}]$ and $\lim_{t \rightarrow \infty} R(t) = 1 - e^{-\nu}$.
+
+As $0 \leq F(t) \leq 1 \Rightarrow 0 \leq \nu F(t) \leq \nu$, we can do a good approximation of $e^{-\nu F(t)}$ using a Taylor series around the point $\nu/2$ for small values of $\nu$.
+
+From the Taylor series centered around $\nu/2$, we know that $e^{-x} \approx e^{-\nu/2} (1 + \nu/2 - x)$.
+Thus, we have $1 - e^{-\nu F(t)} \approx 1 - e^{-\nu/2} (1 + \nu/2) + e^{-\nu/2} \nu F(t)$, and, finally, $f(t) / [1 - F(t)] = [p - w - w e^{-\nu/2} (1 + \nu/2)] + (q - w e^{-\nu/2} \nu)F(t)$.
+
+Therefore, for small values of $\nu$, this model has approximately the same explanation power as the Bass model and we can write $f(t) = p^* + q^*F(t)$, where $p^* = p-w-we^{-\nu/2}(1+\nu/2)$ and $q^* = q - we^{-\nu/2}\nu$.
+
+In contrast to model 1, no parameter could be arbitrarily set in this model, thus it can be estimated using an empirical \textit{active users} dataset.
+
+%This result is interesting because, although it is possible to estimate the parameters using the number \textit{active users}, it has approximately the same explanation power as
+
+% XXX TODO Discuss the inequalities that make sure f(t) >= 0.
+
+\section{Model 3}
+
+In this model, the rate of new \textit{inactive users} increases linearly with the number of \textit{active users}, since $A(t) = F(t) - R(t)$.
+Thus, while there are \textit{active users}, a fraction $\nu$ of them will be rejecting the innovation.
+Hence, everyone will have reject the innovation sooner or later.
+
+The differential equation \ref{eq:model3} can be rewritten as the following first order linear differential equation, which has to be solved:
+
+\begin{align}
+	\frac{d}{dt}R(t) + \nu R(t) = \nu F(t)
+\end{align}
+
+The solution to this differential equation is:
+
+\begin{align}
+	R(t) &= \nu e^{-\nu t} \int_0^t e^{\nu \tau} F(\tau) d\tau \label{eq:model3:R-1} \\
+	     &= \nu \left[ F(u) * e^{-\nu u} \right](t)
+\end{align}
+
+Or, as $\frac{d}{dt} \left( e^{\nu t} F(t) \right) = \nu e^{\nu t} F(t) + e^{\nu t} f(t)$, $R(t)$ can be rewritten as:
+
+\begin{align}
+	R(t) &= F(t) - e^{-\nu t} \int_0^t e^{\nu \tau} f(\tau) d\tau \label{eq:model3:R-2} \\
+	     &= F(t) - \left[ f(u) * e^{-\nu u} \right](t)
+\end{align}
+
+As $F(t) = \int_0^t f(\tau) d\tau$ and $e^{\nu \tau} \geq 1$, we have that:
+
+\begin{align}
+	\int_0^t e^{\nu \tau} f(\tau) d\tau &\geq F(t) \\
+	-e^{-\nu t} \int_0^t e^{\nu \tau} f(\tau) d\tau &\leq -e^{-\nu t} F(t) \\
+	F(t)-e^{-\nu t} \int_0^t e^{\nu \tau} f(\tau) d\tau &\leq F(t) - e^{-\nu t} F(t)
+\end{align}
+
+From equation \ref{eq:model3:R-2}:
+
+\begin{align}
+	R(t) &\leq F(t) - e^{-\nu t} F(t) \\
+	R(t) &\leq F(t)(1 - e^{-\nu t})
+\end{align}
+
+Finally, $R(t) < F(t)$, $r(t) > 0$, and $\lim_{t \rightarrow \infty} R(t) \leq 1$.
+
+I did not manage to prove that $\lim_{t \rightarrow \infty} R(t) = 1$, but this result appeared in all performed simulations.
+If that is true, then all people will reject the innovation at some point in time - a fact that makes sense.
+
+Unfortunately, it seems that there is no closed formula for $F(t)$. Using equation \ref{eq:model3:R-1}, the final differential equation is:
+
+\begin{align}
+\frac{f(t)}{1-F(t)} = p + qF(t) - w\nu e^{-\nu t} \int_0^t e^{\nu \tau} F(\tau) d\tau
+\end{align}
+
+Or, using \ref{eq:model3:R-2}, it becomes:
+
+\begin{align}
+\frac{f(t)}{1-F(t)} = p + (q-w)F(t) + w e^{-\nu t} \int_0^t e^{\nu \tau} f(\tau) d\tau
+\end{align}
+
+The condition $w \leq p+q$ is sufficient to ensure $f(t) \geq 0$, since $w \leq p+q \Rightarrow wR(t) \leq pR(t) + qR(t) \leq p + qR(t) \leq p + qF(t) \Rightarrow p + qF(t) - wR(t) = f(t) / [1 - F(t)] \geq 0 \Rightarrow f(t) \geq 0$.
+Assuming that $\lim_{t \rightarrow \infty} R(t) = 1$, then it is easy to prove that this condition is also necessary.
+
+\section{Model 4}
+
+In this model, the rate of rejection increases linearly with the number of \textit{active users}.
+Thus, while there are \textit{active users}, the rate of rejection will be greater than zero.
+Hence, everyone will have reject the innovation sooner or later.
+
+Although equation \ref{eq:model4} is a Ricatti equation \citep{bittanti1991riccati}, none of the available techniques could solve the differential equation and it seems that there is no closed formula for $F(t)$.
+Hence, the equation will be analyzed through a linearization around the fixed points.
+
+\begin{align}
+\begin{cases}
+f(F, R) = (p + qF - wR)(1 - F) \\
+r(F, R) = \nu (F-R) (1-R)
+\end{cases}
+\end{align}
+
+Solving the system $f(F, R) = r(F, R) = 0$, the following solutions are found:
+
+\begin{align}
+u_1^* &= (1, 1) \\
+u_2^* &= \left( \frac{p}{w-q}, \frac{p}{w-q} \right) \\
+u_3^* &= \left( \frac{w-p}{q}, 1 \right)
+\end{align}
+
+The only valid solutions are $u_1^*$ and $u_2^*$.
+The solution $u_3^*$ is not valid because $p+q-w > 0 \Rightarrow (w-p)/q < 1 \Rightarrow F < R$ which is not possible because it would imply a negative number of \textit{active users}.
+
+Finally, the linearization around $u^*$ is:
+
+\begin{align}
+\begin{bmatrix}
+	f(R, T) - f(u^*) \\
+	r(R, T) - r(u^*)
+\end{bmatrix} = J|_{u^*} \left( \begin{bmatrix} R \\ T \end{bmatrix}-u^* \right)
+\end{align}
+
+\begin{align}
+J =
+\begin{bmatrix}
+	q(1-F)-(p+qF-wR) & -w(1-F) \\
+	\nu (1-R) & -\nu [(1-R) + (F-R)]
+\end{bmatrix}
+\end{align}
+
+Now, let's analyze the jacobian matrices and their eigenvalue for each valid solution.
+
+\begin{align}
+J|_{u_1^*} =
+\begin{bmatrix}
+	-(p+q-w) & 0 \\
+	0 & 0
+\end{bmatrix}
+\end{align}
+
+As the eigenvalues of $J|_{u_1^*}$ are $0$ and $-(p+q-w) < 0$, the point $u_1^* = (1, 1)$ is a sink, i.e., the neighborhood converges to $u_1^*$ when $t \rightarrow \infty$.
+It may be interpreted that all users will have rejected the innovation after a long time.
+
+%\begin{align}
+%J|_{u_2^*} =
+%\begin{bmatrix}
+%	\frac{q(w-p-q)}{w-q} & \frac{-w(w-p-q)}{w-q} \\
+%	\frac{\nu(w-p-q)}{w-q} & \frac{-\nu(w-p-q)}{w-q}
+%\end{bmatrix}
+%\end{align}
+
+%The eigenvalues of $J|_{u_2^*}$ are $\left( \frac{w-p-q}{2(w-q)} \right) \left( q-\nu \pm \sqrt{(\nu+q)^2 - 4w\nu} \right)$.
+
+% XXX TODO Discuss the inequalities that make sure f(t) >= 0.
+
+\chapter{Estimation method}
+
+The parameters of the model should be estimated using empirical data in order to check the explanation power of the model.
+\citet{bass1969} used a discrete version of his differential equation with ordinary least squares.
+While it worked well for him, it has not in the present work.
+There are several well known problems in the estimation of parameters, most of them related to approximation of derivatives and instability of the estimators.
+These problems have been noticed by many authors \citep{schmittlein1982maximum, srinivasan1986technical, xie1997kalman}.
+
+In the present work, the parameters of the models were estimated using a maximum likelihood function on the residuals between $A(t) = F(t) - R(t)$ from the model and the empirical values from the dataset.
+The residuals were assumed to be normally distributed with $\mu=0$, which leads to the same results as the ordinary least square method.
+In order to calculate the residuals, $F(t)$ and $R(t)$ were calculated based on their differential equations using the 4th order Runge Kutta (RK4) numerical method \citep{butcher1987numerical} with $\Delta t = 0.01$.
+
+The empirical data format was $(t_i, x_i)$, where $t_i$ is the time and $x_i$ is the value, and the measurements were not equally spaced over time. The following Log Likelihood equation was used:
+
+\begin{align}
+\text{residual}(t_i, x_i \mid m, p, q, w, \nu) &= x_i - m \cdot \left[ F(t_i \mid p, q, w) - R(t_i \mid \nu) \right] \\
+\text{Log Likelihood}(\vec{t}, \vec{x} \mid m, p, q, w, \nu) &= -\sum_{i=1}^{N} [\text{residual}(t_i, x_i \mid m, p, q, w, \nu)]^2
+\end{align}
+
+During the evaluation of $F(t)$ and $R(t)$ using the Runge Kutta numerical method, sometimes the exact value of $t_i$ was not reached because of the chosen $\Delta t$.
+In these cases, the value of $x_i$ was calculated using a linear approximation with the nearest points.
+Let $(\hat{t}_k, \hat{x}_k)$ and $(\hat{t}_{k+1}, \hat{x}_{k+1})$ be values calculated from the Runge Kutta method, such that $\hat{t}_k < t_i < \hat{t}_{k+1}$.
+Then, the calculated value at $t_i$ was $\hat{x}_k + \left( \frac{\hat{x}_{k+1} - \hat{x}_k}{\hat{t}_{k+1} - \hat{t}_k} \right) \cdot (t_i - \hat{t}_k)$.
+
+Since there is no explicit solution for the parameters $m$, $p$, $q$, $w$, and $\nu$ which maximize the LogLikelihood function, the parameters were estimated using the Truncated Newton Constrained (TNC) method \citep{dembo1983truncated, nash2000survey, nash1984newton} from the SciPy Python Library \citep{scipy}.
+The constrains were $m > 0$, $p > 0$, $q > 0$, $w \geq 0$, and $\nu \geq 0$ for all models.
+The initial guess for the TNC method was the same for all models estimation.
+Sometimes, the method did not converge and another initial guess had to be used.
+
+Even though it has not been done yet, the confidence interval, estimator average and estimator variance for each parameter will be calculated using the bootstrap method.
+
+% XXX TODO Make explicit the inequalities specific to each model.
+
+\chapter{Preliminary results}
+% Data source: http://news.yahoo.com/number-active-users-facebook-over-230449748.html
+% https://www.statista.com/statistics/264810/number-of-monthly-active-facebook-users-worldwide/
+
+The models parameters were estimated using the number of Facebook's \textit{active users} from December 2004 to March 2013 \citep{facebookdata}.
+In the dataset, $x_i$ was the number of Facebook's \textit{active users} at time $t_i$.
+The dataset had 23 non-equally time spaced measures.
+The users who have accessed Facebook at least once in each month were counted in the number of \textit{active users} in that month.
+
+The model 1 has not been estimated because it is not possible to estimate it.
+
+% TODO Discuss about the validity of Facebook being an "innovation".
+
+The estimated models can be seen at figures \ref{fig:model2fit}, \ref{fig:model3fit}, and \ref{fig:model4fit}.
+These figures also have forecasts for the number of \textit{active users}, \textit{inactive users}, and \textit{total users} of Facebook for the next 4 years (from $t=100$ to $t=140$).
+
+In spite of the favorable goodness of fit using model 2 (see figure \ref{fig:model2fit}), the model does not seem to provide a plausible forecast, because it would mean that Facebook is reaching a stable number of \textit{active users} and the rejections are near the end.
+
+Models 3 and 4 have very similar outcomes (see figures \ref{fig:model3fit} and \ref{fig:model4fit}).
+Their Bayesian Information Criterion (BIC) are also close, but model 4 has a better fit with the data.
+Their forecast makes more sense than the forecast of model 2.
+It predicts that Facebook is very close to the peak of \textit{active users} and, in approximately 3 years, it is going to decline.
+It is also interesting to notice that, according to these outcomes, Facebook may not reach its total potential market.
+
+The difference between the outcomes of model 2 and models 3 and 4 could be explained by the fact that they have different rationales behind their models of rejection.
+While differential equation of the model 2 uses the rate of new \textit{total users}, the differential equations of the models 2 and 3 use the proportion of \textit{active users}.
+
+\begin{figure}
+	\centering
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images/fb-model2-1.eps}}
+	\caption{Fit of Model 2 with Facebook's active users dataset.\label{fig:model2fit}
+		{$mF(t)$ is the total users, $mR(t)$ is the inactive users, and $mA(t)$ is the active users.
+The unit of these functions are thousands of people.
+The parameters are $m=1,497.50$, $p=0.000331$, $q=0.100088$, $w=0.140595$, and $\nu=0.187188$.
+The goodness of fit are $R^2=99.84\%$ and BIC=10,566.52.}}
+\end{figure}
+
+\begin{figure}
+	\centering
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images/fb-model3-1.eps}}
+	\caption{Fit of Model 3 with Facebook's active users dataset.\label{fig:model3fit}
+		{$mF(t)$ is the total users, $mR(t)$ is the inactive users, and $mA(t)$ is the active users.
+The unit of these functions are thousands of people.
+The parameters are $m=1,967.64$, $p=0.000184$, $q=0.097867$, $w=0.330511$, and $\nu=0.006912$.
+The goodness of fit are $R^2=99.83\%$ and BIC=11,485.68}}
+\end{figure}
+
+\begin{figure}
+	\centering
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images/fb-model4-1.eps}}
+	\caption{Fit of Model 4 with Facebook's active users dataset.\label{fig:model4fit}
+		{$mF(t)$ is the total users, $mR(t)$ is the inactive users, and $mA(t)$ is the active users.
+The unit of these functions are thousands of people.
+The parameters are $m=1,854.85$, $p=0.000183$, $q=0.099738$, $w=0.334454$, and $\nu=0.007007$.
+The goodness of fit are $R^2=99.84\%$ and BIC=10,724.55}}
+\end{figure}
+
+\chapter{Conclusion}
+
+The main contribution of this work is the parameter estimation through the empirical number of \textit{active users} dataset, forecasting the number of \textit{total users}, the number of \textit{active users}, and the number of \textit{inactive users}.
+
+If the adopters who have rejected the innovation follow the equations of model 1 and 2, then the proposed extended model is transformed into the Bass model through a linear transformation of the parameters.
+This confirms the Bass model robustness.
+
+Model 2 does not seem to be a good model of rejection, since the number of \textit{active users} never decreases which does not seem to be plausible.
+
+Models 3 and 4 had very similar results when fitting the Facebook dataset.
+The lack of analytical solutions for them, however, is a barrier to better understand their behavior, and to know whether they will always have similar outcomes or they will diverge depending on the data.
+Model 3 seems to be more analytically manageable.
+
+It is important to notice that, in models 3 and 4, the innovation may not be adopted by the whole potential market, but it would in the Bass model.
+Whether it will be adopted by the whole potential market or not depends on the parameters $w$ and $\nu$.
+For instance, it seems that $mF(t)$ is not converging to $m$ at figures \ref{fig:model3fit} and \ref{fig:model4fit}.
+It is a major difference between the proposed extended model and the Bass model.
+
+As this is a working paper, it is also intended to include the analyses of other datasets, like either Twitter's number of \textit{active users}, or WhatsApp's, or Netflix's, or Reddit's, or Dropbox's, or Waze's.
+This would enhance the proposed model power of forecasting.
+
+It is also intended to run a back test with the available data.
+First, the parameters of the model is estimated using a subset of the dataset.
+Then, through extrapolation, the number of \textit{active uses} is forecast.
+Finally, it is compared to this part of the dataset - which must not have been used in the estimation.
+
+The main limitation of this work is that it has no theory to support which of the models of rejection best fit with empirical data.
+Although there is an extensive literature on negative word-of-mouth, this literature does not predict which model would be the best.
+But, if any of these companies discloses the number of \textit{total users} and \textit{active users}, the estimation method could be adapted to estimate the parameters using both pieces of information at the same time, which would make possible to verify which of the models of rejection is the best.
+
+Future work could explore other models of rejection and also other estimation methods, like nonlinear least squares \citep{srinivasan1986technical} and Kalman filter \citep{xie1997kalman}.

### 2018-01-17 Chapters/Chapter01.tex
Fix images path

@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-example.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-example.pdf}
 \caption{White nodes represent transactions that have been confirmed at least once.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-swarm.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-swarm.pdf}
 \caption{Suddenly the number of transactions per second increases and the width of the swarm grows.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-conflict.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-conflict.pdf}
 \caption{The red nodes are transactions which had some conflict with previous transaction and were invalidated by the network. Notice that none of them have been confirmed.\label{fig-tangle-conflict}}
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-hist.png}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist.png}
 \caption{Histogram of how long has been a transaction waiting until its first confirmation.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-hist-2.png}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist-2.png}
 \caption{Histogram of how long has been a transaction waiting until its first confirmation.

### 2018-01-17 Chapters/Chapter02.tex
Fix images path

@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images/kanerva-table-7-2-original.png}
+\centering\includegraphics[width=0.5\textwidth]{images02/kanerva-table-7-2-original.png}
 \caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images/sdm-10000w-table-7-2.png}
+\centering\includegraphics[width=0.5\textwidth]{images02/sdm-10000w-table-7-2.png}
 \caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images/sdm-100w-table-7-2.png}
+\centering\includegraphics[width=0.5\textwidth]{images02/sdm-100w-table-7-2.png}
 \caption{Results generated by the framework similar from Kanerva's original Table 7.2. It was a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images/sdm-10000w-notX-table-7-2.png}
+\centering\includegraphics[width=0.5\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
 \caption{This graph shows the interaction effects more clearly.

### 2018-01-17 Chapters/Chapter03.tex
Fix images path

@@
 	\centering
-	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images/fb-model2-1.eps}}
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images03/fb-model2-1.eps}}
 	\caption{Fit of Model 2 with Facebook's active users dataset.\label{fig:model2fit}
@@
 	\centering
-	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images/fb-model3-1.eps}}
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images03/fb-model3-1.eps}}
 	\caption{Fit of Model 3 with Facebook's active users dataset.\label{fig:model3fit}
@@
 	\centering
-	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images/fb-model4-1.eps}}
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images03/fb-model4-1.eps}}
 	\caption{Fit of Model 4 with Facebook's active users dataset.\label{fig:model4fit}

### 2018-01-17 Chapters/sdm2.tex
Add missing file sdm2.tex

Sparse Distributed Memory (SDM) is a mathematical model for cognitive memory published by \citet{Kanerva1988}. It introduces many interesting mathematical properties of $n$-dimensional binary space that, in a memory model, are psychologically plausible.  Most notable among these are the tip-of-the-tongue phenomenon, conformity to Miller's magic number \citep{Linhares2011} and robustness against loss of neurons.

Unlike traditional memory used by computers, SDM performs read and write operations in a multitude of addresses, also called neurons.  That is, the data is not written, or it is not read in a single address spot, but in many addresses. These are called activated addresses, or activated neurons.

The activation of addresses takes place according to their distances from the datum. Suppose one is writing datum $\eta$ at address $\xi$, then all addresses inside a circle with center $\xi$ and radius $r$ are activated. So, $\eta$ will be stored in all these activated addresses, which are around address $\xi$, such as in Figure \ref{fig-addresses-inside-access-radius}.  An address $\xi'$ is inside the circle if its hamming distance to the center $\xi$ is less than or equal to the radius $r$, i.e. $distance(\xi,\xi')\leq r$.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p_circle_r.pdf}

\caption{Activated addresses inside access \protect \\
radius $r$ around center address.\label{fig-addresses-inside-access-radius}}
\end{figure}

The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings. The distance between two bitstrings is calculated using the Hamming distance. It is defined for two bitstrings of equal length as the number of positions at which the bits are different. For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2. One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties that Euclidean distance does.

Every time you are writing or reading in SDM memory, a number of addresses are activated according to their distance to the center address and your data is written in these activated addresses or read from them.  These issues will be addressed in due detail further on, but a major difference from a traditional computer memory is that your data are always stored and retrieved in a multitude of addresses. This way SDM memory has robustness against loss of addresses.

In traditional memory each datum is stored in an address and every time you are looking for a specific datum you have to search through the memory to find out whether you have it. In spite of computer scientists having developed many algorithms to perform fast searches, almost all of them do a precise search. That is, if you have an imprecise clue of what you need, these algorithms will simply fail.

In SDM, the data space is the same as the address space, which amounts to a vectorial, binary space, that is, a $\{0,1\}^{n}$ space. This way, the addresses where the data will be written are the same as the data themselves. For example, the datum $\eta=00101_{b}\in\{0,1\}^{5}$ will be written to the address $\xi=\eta=00101_{b}$. If one chooses a radius of 1, the SDM will activate all addresses one bit away or less from the center address. So, the datum $00101_{b}$ will be written to the addresses $00101_{b}$, $10101_{b}$, $01101_{b}$, $00001_{b}$, $00111_{b}$ and $00100_{b}$. In this case, when one needs to retrieve the data, one could have an imprecise cue at most one bit away from $\eta$, since all addresses one bit away have $\eta$ stored in themselves.  Extending this train of thought for larger dimensions and radius, much more addresses are activated and one can see why SDM is a distributed memory.

When reading a cue $\eta_{x}$ that is $x$ bits away of $\eta$, the cue shares many addresses with $\eta$. The number of shared addresses decreases as the cue's distance to $\eta$ increases, in other words, as $x$ increases. This is shown in Figure \ref{fig-shared-addresses}.  The target datum $\eta$ was written in all shared addresses, thus they will bias the read output in the direction of $\eta$. If the cue is sufficiently near the target datum $\eta$, the read output will be closer to $\eta$ than $\eta_{x}$ was. Repeating the read operation increasingly gets results closer to $\eta$, until it is exactly the same. So, it may be necessary to perform more than one read operation in order to converge to the target data $\eta$.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p1_inter_p2.pdf}

\caption{Shared addresses between the \protect \\
target datum $\eta$ and the cue $\eta_{x}$. \label{fig-shared-addresses}}
\end{figure}

The addresses of the $\{0,1\}^{n}$ space grows exponentially with the number of dimensions $n$, i.e. $N=2^{n}$. For $n=100$ we have $N\approx10^{30}$, which is incredibly big when related to a computer memory. Furthermore, \citet{Kanerva1988} suggests $n$ between 100 and 10,000. Recently he has postulated 10,000 as a desirable minimum $N$ (personal communication). To solve the feasibility problem of implementing this memory, Kanerva made a random sample of $\{0,1\}^{n}$, in his work, having $N'$ elements. All these addresses in the sample are called hard-locations. Other elements of $\{0,1\}^{n}$, not in $N'$, are called virtual neurons. This is represented in Figure \ref{fig-hardlocations}.  All properties of read and write operations presented before remain valid, but limited to hard-locations. Kanerva suggests taking a sample of about one million hard-locations.

Using this sample of binary space, our data space does not exist completely.  That is, the binary space has $2^{n}$ addresses, but the memory is far away from having these addresses available. In fact, only a fraction of this vectorial space is actually instantiated. Following Kanerva's suggestion of one million hard-locations, for $n=100$, only $100\cdot10^{6}/2^{100}=7\cdot10^{-23}$ percent of the whole space exists, and for $n=1,000$ only $100\cdot10^{6}/2^{1000}=7\cdot10^{-294}$ percent.

Kanerva also suggests the selection of a radius that will activate, on average, one one thousandth of the sample, which is 1,000 hard-locations for a sample of one million addresses. In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/hardlocations.pdf}

\caption{Hard-locations randomly sampled from binary space.\label{fig-hardlocations}}
\end{figure}

Since a cue $\eta_{x}$ near the target bitstring $\eta$ shares many hard-locations with $\eta$, SDM can retrieve data from imprecise cues. Despite this feature, it is very important to know how imprecise this cue could be while still giving accurate results. What is the maximum distance from our cue to the original data that still retrieves the right answer? An interesting approach is to perform a read operation with a cue $\eta_{x}$, that is $x$ bits away from the target $\eta$.  Then measure the distance from the read output and $\eta$. If this distance is smaller than $x$ we are converging. Convergence is simple to handle, just read again and again, until it converges to the target $\eta$. If this distance is greater than $x$ we are diverging. Finally, if this distance equals $x$ we are in a tip-of-the-tongue process.  A tip-of-the-tongue psychologically happens when you know that you know, but you can't say what exactly it is. In SDM mathematical model, a tip-of-the-tongue process takes infinite time to converge. \citet{Kanerva1988} called this $x$ distance, where the read's output averages $x$, the critical distance. Intuitively, it is the distance from which smaller distances converge and greater distances diverge. In Figure \ref{fig-p1-p2-iterative-read}, the circle has radius equal to the critical distance and every $\eta_{x}$ inside the circle should converge.  The figure also shows a convergence in four readings.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p1_p2_iter_read.pdf}

\caption{In this example, four iterative readings were\protect \\
required to converge from $\eta_{x}$ to $\eta$.\label{fig-p1-p2-iterative-read}}
\end{figure}

The $\{0,1\}^{n}$ space has $N=2^{n}$ locations from which we instantiate $N'$ samples. Each location in our sample is called a hard-location.  On these hard-locations we do operations of read and write. One of the insights of SDM is exactly the way we read and write: using data as addresses in a distributed fashion. Each datum $\eta$ is written in every activated hard-location inside the access radius centered on the address, that equals datum, $\xi=\eta$. Kanerva suggested using an access radius $r$ having about one one thousandth of $N'$.  As an imprecise cue $\eta_{x}$ shares hard-locations with the target bitstring $\eta,$ it is possible to retrieve $\eta$ correctly. (Actually, probably more than one read is necessary to retrieve exactly $\eta.)$.  Moreover, if some neurons are lost, only a fraction of the datum is lost and it is possible that the memory can still retrieve the right datum.

A random bitstring is generated with equal probability of $0$'s and $1$'s in each bit. One can readily see that the average distance between two random bitstrings has binomial distribution with mean $n/2$ and standard deviation $\sqrt{n/4}$. For a large $n$, most of the space lies close to the mean and has fewer shared hard-locations.  As two bitstrings with distance far from $n/2$ are very improbable, \citet{Kanerva1988} defined that two bitstrings are orthogonal when their distance is $n/2$.

The write operation needs to store, for each dimension bit which happened more ($0$'s or $1$'s). This way, each hard-location has $n$ counters, one for each dimension. The counter is incremented for each bit $1$ and decremented for each bit $0$. Thus, if the counter is positive, there have been more $1$'s than $0$'s, if the counter is negative, there have been more $0$'s than $1$'s, and if the counter is zero, there have been an equal number of $1$'s and $0$'s. Table \ref{tab:write operation} shows an example of a write operation being performed in a 7-dimensional memory.

\begin{table}
\begin{tabular}{c|c|c|c|c|c|c|c|}
\cline{2-8} 
$\eta$ & 0 & 1 & 1 & 0 & 1 & 0 & 0\tabularnewline
\cline{2-8} 
$\xi_{\textit{before}}$ & 6 & -3 & 12 & -1 & 0 & 2 & 4\tabularnewline
\cline{2-8} 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}}\tabularnewline
\cline{2-8} 
$\xi_{\textit{\textit{after}}}$ & \textbf{5} & \textbf{-2} & \textbf{13} & \textbf{-2} & \textbf{1} & \textbf{1} & \textbf{3}\tabularnewline
\cline{2-8} 
\end{tabular}

\caption{Write operation example in a 7-dimensional memory of data $\eta$
being written to $\xi$, one of the activated addresses.\label{tab:write operation}}

\end{table}

The read is performed polling each activated hard-location and statistically choosing the most written bit for each dimension. It consists of adding all $n$ counters from the activated hard-locations and, for each bit, choosing bit 1 if the counter is positive, choose bit 0 if the counter if negative, and randomly choose bit 0 or 1 if the counter is zero.

\section{Neurons as pointers}

One interesting view is that neurons in SDM work like pointers. As we write bitstrings in memory, the hard-locations' counters are updated and some bits are flipped. Thus, the activated hard-locations do not necessarily point individually to the bitstring that activated it, but together they point correctly. In other words, the read operation depends on many hard-locations to be successful. This effect is represented in Figure \ref{fig-p1-pointers}: where all hard-locations inside the circle are activated and they, individually, don't point to $\eta$.  But, like vectors, adding them up points to $\eta$. If another datum $\nu$ is written into the memory near $\eta$, the shared hard-locations will have information from both of them and wouldn't point to either.  All hard-locations outside of the circle are also pointing somewhere (possibly other data points). This is not shown, however, in order to keep the picture clean and easily understandable.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p1_after_write.pdf}

\caption{Hard-locations pointing, approximately, to the target bitstring.\label{fig-p1-pointers}}
\end{figure}

\section{Applications}

SDM memory or an extension of it has been used in a lot of applications, helping to solve specific problems. \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.  \citet{Rajesh1998} used SDM in an autonomous robot. \citet{Meng2009} modified SDM to clean patterns from noisy inputs.

\citet{Linhares2011} showed that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}. This is the only work, to the author's knowledge, that explores some of intrinsic properties of SDM.

\section{Concepts}

Although Kanerva does not mention concepts directly in his book \citep{Kanerva1988}, the author's interpretation is that each bitstring may be mapped to a concept. Thus, unrelated concepts are orthogonal and concepts could be linked through a bitstring near both of them. For example, ``beauty'' and ``woman'' have distance $n/2$, but a bitstring that means ``beautiful woman'' could have distance $n/4$ to both of them. As a bitstring with distance $n/4$ is very improbable, it is linking those concepts together. \citet{Linhares2011} approached this concept via ``chunking through averaging''.

Due to the distribution of hard-locations between two random bitstrings, the vast majority of concepts is orthogonal to all others. Consider a non-scientific survey during a cognitive science seminar, where students asked to mention ideas unrelated to the course brought up terms like birthdays, boots, dinosaurs, fever, executive order, x-rays, and so on. Not only are the items unrelated to cognitive science, the topic of the seminar, but they are also unrelated to each other.

For any two memory items, one can readily find a stream of thought relating two such items (``Darwin gave dinosaurs the boot''; ``she ran a fever on her birthday''; ``isn't it time for the Supreme Court to x-ray that executive order?'', ... and so forth). Robert French presents an intriguing example in which one suddenly creates a representation linking the otherwise unrelated concepts of ``coffee cups'' and ``old elephants'' \citep{French1997}.

This mapping from concepts to bitstrings brings us two main questions: (i) Suppose we have a bitstring that is linking two major concepts.  How do we know which concepts are linked together? (ii) From a concept bitstring how can we list all concepts that are somehow linked to it? This second question is called the problem of spreading activation.

\section{Read operation}

In his work, Kanerva proposed and analyzed a read algorithm called here Kanerva's read. His read takes all activated hard-locations counters and sum them. The resulting bitstring has bit $1$ where the result is positive, bit $0$ where the result is negative, and a random bit where the result is zero. In a word, each bit is chosen according to all written bitstrings in all hard-locations, being equal to the bit more appeared. Table \ref{tab:kanerva-read} shows an example of Kanerva's read result bitstring.

Daniel Chada, one member of our research group, proposed another way to read in SDM, in this work called Chada's read. Instead of summing all hard-location counters, each hard-location evaluates its resulting bitstring individually. Then, all resulting bitstring are summed again, and the same rule as Kanerva applies. Table \ref{tab:chada-read} shows an example of Chada's read result bitstring. The counter's values are normalized to 1, for positive ones, or -1, for negative ones, and the original values are the same as in Table \ref{tab:kanerva-read}.

The main change between Kanerva's read and Chada's read is that, in the former, a hard-location that has more bitstrings written has a greater weight in the decision of each bit. In the latter, all hard-locations have the same weight, because they can contribute to the sum with only one bitstring.

\begin{table}
\begin{minipage}[t]{0.5\columnwidth}%
\subfloat[Kanerva's read example\label{tab:kanerva-read}]{%
\begin{tabular}{c|c|c|c|c|c|}
\cline{2-6} 
$\xi_{1}$ & -2 & 12 & 4 & 0 & -3\tabularnewline
\cline{2-6} 
$\xi_{2}$ & -5 & -4 & 2 & 8 & -2\tabularnewline
\cline{2-6} 
$\xi_{3}$ & -1 & 0 & -1 & -2 & -1\tabularnewline
\cline{2-6} 
$\xi_{4}$ & 3 & 2 & -1 & 3 & 1\tabularnewline
\hline 
\multicolumn{1}{|c|}{\textbf{$\sum$}} & \textbf{-5} & \textbf{10} & \textbf{4} & \textbf{3} & \textbf{-5}\tabularnewline
\hline 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}}\tabularnewline
\cline{2-6} 
 & 0 & 1 & 1 & 1 & 0\tabularnewline
\cline{2-6} 
\end{tabular}

}%
\end{minipage}%
\begin{minipage}[t]{0.5\columnwidth}%
\subfloat[Chada's read example\label{tab:chada-read}]{%
\begin{tabular}{c|c|c|c|c|c|}
\cline{2-6} 
$\xi_{1}$ & -1 & 1 & 1 & \cellcolor{lightgray}1 & -3\tabularnewline
\cline{2-6} 
$\xi_{2}$ & -1 & -1 & 1 & 1 & -1\tabularnewline
\cline{2-6} 
$\xi_{3}$ & -1 & \cellcolor{lightgray}1 & -1 & -1 & -1\tabularnewline
\cline{2-6} 
$\xi_{4}$ & 1 & 1 & -1 & -1 & 1\tabularnewline
\hline 
\multicolumn{1}{|c|}{\textbf{$\sum$}} & \textbf{-2} & \textbf{1} & \textbf{0} & \textbf{0} & \textbf{-2}\tabularnewline
\hline 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$}}\tabularnewline
\cline{2-6} 
 & 0 & 1 & \cellcolor{lightgray}1 & \cellcolor{lightgray}1 & 0\tabularnewline
\cline{2-6} 
\end{tabular}

}%
\end{minipage}\caption{Comparison of Kanerva's read and Chada's read. Each $\xi_{i}$ is
an activated hard-location and the values come from their counters.
Gray cells' value is obtained randomly with probability 50\%.\label{tab:read-operation}}
\end{table}

\section{Critical Distance}

Kanerva describes the critical distance as the threshold of convergence of a sequence of read words. It is ``the distance beyond which divergence is more likely than convergence''\citep{Kanerva1988}. Furthermore, Kanerva explains that ``a very good estimate of the critical distance can be obtained by finding the distance at which the arithmetic mean of the new distance to the target equals the old distance to the target''\citep{Kanerva1988}.  In other words, the critical distance can be equated as the edge to our memory, the limit of human recollection.

In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 10 million hard-locations, an access-radius of 451 (within 1000 hard-locations in each circle) and 10 thousand writes of random bitstrings in the memory. As computer resources were very poor those days, Kanerva couldn't make a more generic analysis.

Starting from the premise of SDM as a faithful model of human short-term memory, a better understanding of the critical distance may shed light on our understanding of the thresholds that bind our own memory.

### 2018-01-17 FrontBackmatter/Introduction.tex
Introduction

--- 2018-01-17 FrontBackmatter/Foreword.tex
@@
-%*******************************************************
-% Publications
-%*******************************************************
-\pdfbookmark[1]{Prelude}{prelude}
-\chapter*{Prelude, introduction, and discussion of a thesis proposal}
-\bigskip
+\documentclass[12pt]{article}
 
-\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
+\usepackage{natbib}
+
+%% Language and font encodings
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
+
+%% Useful packages
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+
+\title{Essays in \\Computational Management Science}
+\author{Marcelo Salhab Brogliato}
+
+\begin{document}
+\maketitle
+
+\section{Introduction}
+
+If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
 
@@
 Applied mathematics and computer science serve simultaneously as a theoretical foundation and the major tool available to the field.
-Though this is a thesis proposal concerning business, in this document one should expect to find the language and nomenclature of mathematical modeling and computer science as our primary and most natural language.
+Though this is a doctoral thesis concerning business, in this document one should expect to find the language and nomenclature of mathematical modeling and computer science as our primary and most natural language.
 
-This proposal will present a number of different topics for exploration.
-Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of a future thesis brought by three self-contained scientific papers.
-%---of the kind and quality that is expected in the major scientific journals of the field.
+This thesis will present a number of different topics for exploration.
+Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three self-contained scientific papers.
+The first of which will study the possibility of distributed financial ledgers.
 
-%Though work has already begun in all of these studies, at this moment, for the proposal, each study is offered under its own  level of maturity, ranging from quasi-ready for submission for publication to research at its embryonic phase.
+\subsection{Distributed financial ledgers}
+
+United Nations World Food Programme, `Blockchain Against Hunger: Harnessing Technology In Support Of Syrian Refugees', 30 May 2017, https://www.wfp.org/news/news-release/blockchain-against-hunger-harnessing-technology-support-syrian-refugees, accessed in January 17, 2018.
+
+United Nations World Food Programme, 2016, The Year in Review Report, p.
 
 The first paper lies at the intersection of computer science, record-keeping, banking \& finance, and economic inclusion.
-Our concern will be over a Bitcoin-like system that departs from the blockchain architecture.
-It has been recently proposed that Directed Acyclic Graphs could certify transactions in a blockchain-like manner.
+The seminal work of \citet{nakamoto2008bitcoin} described the architecture of Bitcoin, a peer-to-peer electronic cash system, also known as cryptocurrency.
+Bitcoin's currency ledger is public and stored in a blockchain across thousands of computers.
+Even so, no one is able to spend either somebody else's funds nor to double spend their own funds.
+In order to be confirmed, each transaction must be both digitally signed by the owner of the money and the funds verified in the blockchain by Bitcoin's miners.
+One of the interesting parts of Bitcoin are the incentives.
+On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick, and the currency issuance rate is well known.
+On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.
+
+The impact of Bitcoin in society -- and hence in the companies and the government -- has been increasing every day.
+People are increasingly using Bitcoin to exchange money and transfer money overseas.
+Companies are looking into Bitcoin as an alternative to reduce banking fees.
+The poor may be included in the finance system through Bitcoin.
+People may hedge their assets against their governments' money issuance and inflation---as in the sad case of Venezuela.
+
+Bitcoin is the first and most famous cryptocurrency, used worldwide and has a highly volatile market cap, as of this writing, of \$ 224 mi.
+Even so, it faces serious scalability challenges; such as serious quality of service and network congestion when the number of transactions per second is high, and an increase in the transaction fees and uncertain delays in transactions' confirmations.
+
+Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'' developers, which believe that is it risky to increase the the blocksize (in which all transactions are stored).
+It is not known whether a blocksize, say, of 1Gb, would be feasible to sustain the decentralization of the network.
+
+Iota is another cryptocurrency, but, instead of using a blockchain, it proposes the use a ``tangle' architecture', which is a different way to store the currency ledger across thousands of computers.
+Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
+As we will see, the problem here is exactly the opposite of Bitcoin's.
+It needs a minimum of transactions per seconds in order to work properly.
+
+Our analysis suggests an architecture for a distributed currency which is inspired in both Bitcoin's blockchain and Iota's tangle in order to solve the scalability problems.
+While Bitcoin's network saturates when it hits a certain number of transactions per second, Iota's does not work properly with less than a certain number of transactions per second.
+Our proposed architecture works in both scenarios: low and high number of transactions per second.
+
 We investigate some issues regarding this possibility, namely: (i) cryptographic security and game-theoretical attacks; (ii) scalability; (iii) self-governance of the system; (iv) appropriate incentive system to all participants.
 
-%This first paper is the most immature one.  It is work being conducted.  The research questions are still under consideration. Some original theorems (yet to be proved) are presented.  
-This first paper is work being conducted. The research questions are still under consideration. Some original theorems (yet to be proved) are presented.
+\subsection{Sparse Distributed Memory}
 
@@
 It is my hope that readers of this proposal will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---the proposed body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
+
+\bibliographystyle{apa}
+\bibliography{sample.bib}
+\end{document}

### 2018-01-18 Chapters/Introduction.tex
Update on Overleaf.

--- 2018-01-17 FrontBackmatter/Introduction.tex
@@
-\documentclass[12pt]{article}
+\chapter{Introduction}
+\bigskip
 
-\usepackage{natbib}
-
-%% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-
-%% Useful packages
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-
-\title{Essays in \\Computational Management Science}
-\author{Marcelo Salhab Brogliato}
-
-\begin{document}
-\maketitle
-
-\section{Introduction}
-
-If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
+\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
 
@@
 
-\subsection{Distributed financial ledgers}
+\section{Distributed financial ledgers}
 
@@
 
-Iota is another cryptocurrency, but, instead of using a blockchain, it proposes the use a ``tangle' architecture', which is a different way to store the currency ledger across thousands of computers.
+Iota is another cryptocurrency, but, instead of using a blockchain, it proposes the use a ``tangle architecture'', which is a different way to store the currency ledger across thousands of computers.
 Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
@@
 
-\subsection{Sparse Distributed Memory}
+\section{Artificial Intelligence}
 
@@
 It is my hope that readers of this proposal will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---the proposed body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
-
-\bibliographystyle{apa}
-\bibliography{sample.bib}
-\end{document}

### 2018-01-18 Chapters/Introduction.tex
Update introduction

@@
 
-\section{Distributed financial ledgers}
+\subsection{Distributed financial ledgers}
 
@@
 
-United Nations World Food Programme, 2016, The Year in Review Report, p.
+United Nations World Food Programme, 2016, The Year in Review, Report.
 
-The first paper lies at the intersection of computer science, record-keeping, banking \& finance, and economic inclusion.
+Woyke, E., 2017, `How blockchain can bring financial services to the poor', MIT Technology Review, April 18.
+
+The World Bank estimates that there are two billion people without access to financial services.
+As banks are unable to sustain operations in numerous poverty-stricken areas, services such as money transfers, access to credit, digital/distant payments, inflation protection, etc., remain beyond reach for large numbers of people.
+This seems to be one of the factors that perpetuate poverty, and the Bill and Melinda Gates Foundation chose as focus of its ``Level-One project'': to provide basic financial services through cell phones.
+Another initiative, the United Nations World Food Programme has began, in 2017, an experiment in Jordan, in which the organization provides funds for thousands of people towards its goal of food relief.
+An interesting aspect of this program has been the format of the funds distributed:  they have been all on the ethereum blockchain.
+
+The possibility of having a completely digital financial system without the overheads of traditional banking systems has appeared with the release of Bitcoin and similar blockchain technologies.
+This field questions numerous traditional assumptions in computer science, record-keeping, banking \& finance, and economic inclusion.
 The seminal work of \citet{nakamoto2008bitcoin} described the architecture of Bitcoin, a peer-to-peer electronic cash system, also known as cryptocurrency.
@@
 In order to be confirmed, each transaction must be both digitally signed by the owner of the money and the funds verified in the blockchain by Bitcoin's miners.
+
+The question, however, of whether Bitcoin (or related works) can scale to billions of people is far from settled.
+
 One of the interesting parts of Bitcoin are the incentives.
@@
 The poor may be included in the finance system through Bitcoin.
-People may hedge their assets against their governments' money issuance and inflation---as in the sad case of Venezuela.
+People may hedge their assets against their governments' money issuance and inflation---as in the case of Venezuela.
 
-Bitcoin is the first and most famous cryptocurrency, used worldwide and has a highly volatile market cap, as of this writing, of \$ 224 mi.
+Bitcoin is the first and most famous cryptocurrency, used worldwide and has a highly volatile market cap, as of this writing, of \$ 192 bi.
 Even so, it faces serious scalability challenges; such as serious quality of service and network congestion when the number of transactions per second is high, and an increase in the transaction fees and uncertain delays in transactions' confirmations.
@@
 
-Iota is another cryptocurrency, but, instead of using a blockchain, it proposes the use a ``tangle architecture'', which is a different way to store the currency ledger across thousands of computers.
+Iota is another cryptocurrency, but, instead of using a blockchain, it proposes the use a ``tangle' architecture', which is a different way to store the currency ledger across thousands of computers.
 Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
@@
 While Bitcoin's network saturates when it hits a certain number of transactions per second, Iota's does not work properly with less than a certain number of transactions per second.
-Our proposed architecture works in both scenarios: low and high number of transactions per second.
+Our proposed architecture seems to work in both scenarios: low and high number of transactions per second.
 
-We investigate some issues regarding this possibility, namely: (i) cryptographic security and game-theoretical attacks; (ii) scalability; (iii) self-governance of the system; (iv) appropriate incentive system to all participants.
+In this first study we will investigate some issues regarding this possibility, namely: (i) cryptographic security and game-theoretical attacks; (ii) scalability; (iii) self-governance of the system; (iv) appropriate incentive system to all participants.
 
-\section{Artificial Intelligence}
+A second topic that may have an outsized influence ion business and that we will be taking a closer look is a model of artificial intelligence.
 
-%The second paper lies at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence.
-Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics.
-We implement a RB-Complete\footnote{Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
-The maturity of this work is medium-high: The computer code has been used in a previous Ph.D.
-Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).
-What remains to be done is a full description of the systems' architecture and a performance comparison between different setups.
+\subsection{Artificial intelligence}
+% Technology > Computer Science > Artificial Intelligence > Pattern Recognition > SDM.
+
+Technology has been one of the underlying engines behind economic growth.
+It has been changing the whole society -- people, companies, and governments.
+Cities and houses had to be rethought when cars became popular.
+Trains allowed distant places to exchange high volume of goods.
+Airplanes and boats allowed countries to do business overseas.
+And, finally, internet has changed everything -- the way we communicate, behave, do business, do shopping, share ideas, and so forth.
+
+One are of technology that has been redefining business is computer science.
+Together with the internet, computer science has been one of the most important tools to scale a business model -- and create many others which were impossible before.
+More and more expensive human labor has been replace by algorithms.
+Managers are able to make better decisions because they receive real time information.
+The supply chain has incredibly evolved thanks to advances in logistics supported by routing algorithms, storage algorithms, and many others optimization algorithms.
+
+Artificial intelligence has been disrupting many businesses.
+Uber is able to handle hundred of thousands of requests.
+Amazon optimizes the location of each product based on demand.
+Netflix increases the quality of their services offering movies specific to the taste of each customer.
+Spotify learns which kind of music you like most and suggests playlists and radios.
+Banks prevent fraud classifying which patterns are strange to their customers.
+It is not possible to imagine a world with no artificial intelligence.
+Behind many artificial intelligence systems, there is the pattern recognition.
+The capacity to match information from new data with information which has already been seen and is stored in memory.
+It may be used to classification, face recognition, character recognition, and so forth.
+
 The second paper lies at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence.
-Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics.
+Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics of a human memory.
+SDM has already been used to different patter recognition problems, like noise reduction, handwriting recognition, robot automation, and so forth.
+
 We implement a RB-Complete\footnote{Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
@@
 What remains to be done is a full description of the systems' architecture and a performance comparison between different setups.
+
+\section{Innovation adoption}
 
@@
 
-A note on the overall format of this Thesis Proposal:  many journals, such as Nature, have a series of papers departing from the standard layout of (i) introduction, (ii) literature review, (iii) methodology, (iv) experiments, (v) discussion and (vi) conclusion.
-Instead, these journals will generally present (i) introduction, (ii) discussion and conclusion, and---sometimes in smaller fonts---, the materials and methods and results section.
-Here, content is brought to focus, in detriment to form.
-As I believe that the content of the studies should be the major focus of analysis for this proposal, I have also collapsed the introductory and concluding remarks into these few comments.
-I hope readers will agree that, at this point, it is of overwhelmingly greater importance to query the scientific content of the proposal, even if in detriment to its overall form.
-
 It is my hope that readers of this proposal will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---the proposed body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.

### 2018-01-18 Chapters/Introduction.tex
Update on Overleaf.

@@
 
-The question, however, of whether Bitcoin (or related works) can scale to billions of people is far from settled.
+The question of whether Bitcoin (or related works) can scale to billions of people is, however, far from settled.
 
@@
 
-Iota is another cryptocurrency, but, instead of using a blockchain, it proposes the use a ``tangle architecture'', which is a different way to store the currency ledger across thousands of computers.
+\emph{Iota} is a second cryptocurrency, which, instead of using a blockchain, proposes the use of a ``tangle' architecture': a different way to register the currency ledger across thousands of computers.
 Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
 As we will see, the problem here is exactly the opposite of Bitcoin's.
-It needs a minimum of transactions per seconds in order to work properly.
+Iota needs a minimum of transactions per seconds in order to work properly.
 
@@
 
-A second topic that may have an outsized influence ion business and that we will be taking a closer look is a model of artificial intelligence.
+A second topic that may have an outsized influence on business and that we will be taking a closer look is a model of artificial intelligence.
 
-\subsection{Artificial intelligence}
+\section{Artificial intelligence}
 % Technology > Computer Science > Artificial Intelligence > Pattern Recognition > SDM.
@@
 
-\section{Innovation adoption}
+\section{Diffusion of innovation}
 
-The third paper lies at the intersection of Marketing, Diffusion of Technological Innovation, and modeling.
-The classic Bass model of diffusion of innovation is extended, in order to account for users who, after adopting the innovation for a while, decide to reject it later on (possibly bringing down the number of active users---which is impossible in Bass' original model).
-Four alternative mathematical models are presented and discussed with the Facebook users dataset.
-%This is the work that has received most energy so far, and is the most mature of the papers presented herein.
-I believe it is almost ready for submission for publication.
+In 2014, a group of Princeton's researchers predicted that Facebook's users would abandon the platform by 2017 \citep{cannarella2014epidemiological}.
+The forecast was done applying a disease spreading model which has correctly predicted the abandonment of ``MySpace''.
+Facebook replied applying exactly the same methodology as Princeton's researchers.
+In their own words: ``Using the same robust methodology featured in the paper, we attempted to find out more about this `Princeton University' -- and you won't believe what we found!''.
+Then, they conclude: ``This trend suggests that Princeton will have only half its current enrollment by 2018, and by 2021 it will have no students at all, agreeing with the previous graph of scholarly scholarliness.
+Based on our robust scientific analysis, future generations will only be able to imagine this now-rubble institution that once walked this earth''.
 
-%A fourth paper is still highly immature.
-It concerns monetary inflation, and the demonstrations that an approach through fixed effects cannot correct for the real cumulative inflation rates---and may be disastrous when referring to regions undergoing high inflation.
-Though part of the mathematical analysis is mature, the work itself is lacking the necessary references to show its relevance:  I still need to review the literature and point out the rather large body of it that is committing a serious error.
-Because I have not decided whether this work should be part of the thesis, it will be presented in appendix 1.
+Whilst this brouhaha reminds one of the dangers of extrapolation, our third paper will revisit the prospects of our esteemed colleagues in Facebook.
+Lying at the intersection of Marketing, Diffusion of Technological Innovation, and modeling, the Bass model of diffusion of innovation will be extended, in order to account for users who, after adopting the innovation for a while, decide to reject it later on (possibly bringing down the number of active users---which is impossible in Bass' original model).
+Four alternative mathematical models are presented and discussed with the Facebook's users dataset.
 
-It is my hope that readers of this proposal will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---the proposed body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
+\section{The fine print...}
+
+Before embarking on the technical topics, small qualifications must be asked from my readers.
+First, as stated above, though these problems have immense and urgent importance to the fields of study in business, the language in which we will approach them and discuss them most naturally will be that of mathematics and computer science.
+There will not be surveys, interviews, questionnaires, or such methods typically used in the social sciences: This is basically a work of modeling.
+
+A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---this body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
+Please bear with me as we venture into the world of computational management science.

### 2018-01-18 Chapters/Chapter01.tex
Update blockchain

--- 2018-01-17 Chapters/Chapter01.tex
@@
 
+\chapter{Analysis of Bitcoin}
+
+\section{Hash function}
+
+A hash function $h: \{0, 1\}^\infty \rightarrow \{0, 1\}^{n}$ has the following properties:
+
+\begin{enumerate}
+	\item $x = y \Rightarrow h(x) = h(y)$
+	\item $h(x) \sim \mathcal{U}(0, 2^{n}-1)$, where $\mathcal{U}$ is the uniform distribution, i.e., $\forall a \in [0, 2^{n}-1], \mathbf{P}(h(x)=a) = \frac{1}{2^{n}}$
+\end{enumerate}
+
+In other words, when two inputs are the same, they have the same output.
+But, when the inputs are different, their outputs are uniformly distributed.
+Clearly, the hash functions are surjective but not injective.
+They are not injective because the image of $h$ has only $2^n$ elements and the domain has infinite elements.
+When $x \ne y$ and $h(x) = h(y)$, we say that $x$ and $y$ are a collision.
+A hash function is considered to be safe when it is unknown how to find a collision of a given hash.
+
+\section{Bitcoin mining}
+
+Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin hash function.
+The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given number.
+The lesser the $A$, the harder to find a new block.
+In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.
+
+Hence, in order to find a new block, one must try different inputs ($x_1, x_2, \dots, x_k$) until they find a solution, i.e., all attempts will fail ($h(x_i) \geq A$ for $i < k$) but the last ($h(x_k) < A$).
+The probability of finding a solution exactly in the $k^{th}$ attempt follows a geometric distribution.
+Let $X$ be the number of attempts until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \frac{A}{2^{256}}$.
+Also, we have $\mathbf{P}(X \leq k) = 1 - (1-p)^k$.
+The average number of attempts is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = \frac{1-p}{p^2}$.
+
+%As one may fail many times until they find a solution, the probability of finding a solution in the $K$ trial follows a geometric distribution.
+Let $X$ be the number of trials until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \mathbf{P}(a < A)$, i.e., there would be $k-1$ failures until a success in the $k-th$ trial.
+The average number of trials is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = (1-p)/p^2$.
+
+In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average.
+Suppose that the Bitcoin's network is able to calculate $H$ hashes per second -- $H$ is the total hash rate of the network.
+The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block.
+So, the rule of finding a new block every 10 minutes (600 seconds) -- on average -- leads to the following equation: $\mathbf{E}(T) = 600$.
+So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = 600 \Rightarrow p = \frac{1}{600H}$.
+Finally, $\mathbf{E}(X) = 600H$, $\mathbf{E}(T) = 600$, $\mathbf{V}(X) = (600H)^2 - 600H$, and $\mathbf{V}(T) = 600^2 - 600/H$.
+
+The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{600H} \right)^{tH}$.
+But, as the Bitcoin's network hash rate is really big, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{600}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{600}$.
+
+\begin{theorem}
+	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{600}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{600}}$.
+\end{theorem}
+\begin{proof}
+\begin{align*}
+	\mathbf{P}(T \leq t) &= 1 - (1-p)^{tH} \\
+		     &= 1 - \left( 1 - \frac{1}{600H} \right)^{tH} \\
+	\text{Replacing $u = 600H$,} \\
+	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{600}} \\
+				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{600}} \\
+				&= 1 - \left( 1/e \right)^{\frac{t}{600}} \\
+				&= 1 - e^{-\frac{t}{600}}
+\end{align*}
+\end{proof}
+
+% A partir de qual valor de H é razoável dizer que T segue uma distribuição exponencial?
+\begin{theorem}
+	$x > M \Rightarrow |(1+1/x)^x - e| < e/M$.
+\end{theorem}
+\begin{proof}
+	Let's use the classical inequality $\frac{x}{1+x} < \log(1+x) < x$ for $x > -1$.
+So, $\frac{1/x}{1+1/x} < \log(1+x) < 1/x$.
+Simplifying, $\frac{1/x}{1+1/x} = 1/(1+x)$.
+Thus, $1/(1+x) < \log(1+1/x) < 1/x \Rightarrow x/(1+x) < x \log(1 + 1/x) < 1$.
+
+	As $\log(1 + \frac{1}{M}) > 0$, $1 < 1 + \log(1 + \frac{1}{M})$.
+
+	$x > M \Rightarrow 1/x < 1/M \Rightarrow 1+1/x < 1+1/M \Rightarrow 1/(1+1/x) > 1/(1+1/M) \Rightarrow x/(1+x) > M/(1+M)$.
+
+	Again, $\log(1+x) < x \Rightarrow \log(1-1/M) < -1/M \Rightarrow 1 + \log(1-1/M) < (M-1)/M < M/(1+M)$, since $(x-1)/x < x/(x+1)$.
+
+	Hence, $1 + \log(1-1/M) < M/(1+M) < x/(1+x) < x\log(1+1/x)$, and $x\log(1+1/x) < 1 < 1 + \log(1 + \frac{1}{M})$.
+
+	Finally,
+
+	\begin{align*}
+		1 + \log(1-1/M) &< x\log(1+1/x) < 1 + \log(1 + \frac{1}{M}) \\
+		e^{1 + \log(1-1/M)} &< e^{x\log(1+1/x)} < e^{1 + \log(1 + \frac{1}{M})} \\
+		e \cdot e^{\log(1-1/M)} &< e^{\log((1+1/x)^x)} < e \cdot e^{\log(1 + \frac{1}{M})} \\
+		e(1-1/M) &< (1+1/x)^x < e(1 + \frac{1}{M}) \\
+		e-e/M &< (1+1/x)^x < e + e/M \\
+		-e/M &< (1+1/x)^x - e < e/M \\
+	\end{align*}
+	Therefore, $\left| \left( 1+1/x \right)^x - e \right| < e/M$.
+\end{proof}
+
+We may consider $H$ big enough to say that $T$ follows an exponential distribution when $e/H < \epsilon$, where $\epsilon$ is the maximum approximation error.
+When $\epsilon = 10^{-6} \Rightarrow H > e \cdot 10^6$.
+So, when $H > 2.6 \text{Mh/s}$, our approximation is good enough.
+
+Let's analyze the probability of finding a new block for a miner who has $\alpha$ percent of the network's total hash rate.
+Let $T_\alpha = \frac{X}{\alpha H}$ be the time required for the miner to find a new block.
+As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{600}$.
+Hence, we confirm the intuition that the miner with $\alpha$ percent of the network's total hash power will find $\alpha$ percent of the blocks.
+
+\begin{theorem}
+	When the miner with $\alpha$ percent of the network's total hash rate is part of the mining network, $\mathbf{P}(\text{next block is from $T_\alpha$}) = \alpha.$
+\end{theorem}
+\begin{proof}
+\begin{align*}
+	\mathbf{P}(\text{next block is from $T_\alpha$}) &= P \left( T_\alpha = \min\{T_\alpha, T_{1-\alpha}\} \right) \\
+		&= \frac{\lambda_\alpha}{\lambda_\alpha + \lambda_{1-\alpha}} \\
+		&= \frac{\alpha/600}{\alpha/600 + (1-\alpha)/600} \\
+		&= \frac{\alpha}{\alpha + 1 - \alpha} \\
+		&= \alpha.
+\end{align*}
+\end{proof}
+
+\begin{theorem}
+	When one miner with $\alpha$ percent of the network's total hash rate multiplies their hash rate by $m$, the probability of finding the next block is multiplied by $\frac{m}{m \alpha + 1 - \alpha}$.
+	\label{thm-miner-multiply}
+\end{theorem}
+\begin{proof}
+	When miners increase their hash rate, they also increase the network's total hash rate.
+Let $H$ be the network's hash rate before the increase.
+Thus, the network's total hash rate after the increase is $H + (m-1) \alpha H = (1 - \alpha + m \alpha) H$.
+So,
+\begin{align*}
+	\mathbf{P}(\text{next block is from $T_{m \alpha}$}) &= P \left( T_{m \alpha} = \min\{T_{m \alpha}, T_{1-\alpha}\} \right) \\
+		&= \frac{\lambda_{m \alpha}}{\lambda_{m \alpha} + \lambda_{1-\alpha}} \\
+		&= \frac{m \alpha/600}{m \alpha/600 + (1-\alpha)/600} \\
+		&= \frac{m \alpha}{m \alpha + 1-\alpha} \\
+		&= \alpha \left( \frac{m}{m \alpha + 1 - \alpha} \right).
+\end{align*}
+\end{proof}
+
+\begin{cor}
+	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximatelly multiplies their probability of finding the next block by $m$.
+\end{cor}
+\begin{proof}
+	$$\lim_{\alpha \rightarrow 0} \mathbf{P}(\text{next block is from $T_{m \alpha}$}) = \lim_{\alpha \rightarrow 0} \frac{m}{m \alpha + 1 - \alpha} = m.$$
+\end{proof}
+
+That way, it is not exacly correct to say that when one doubles their hash rate, their probability will double as well. It is only true for small miners.
+
+%Then, $\mathbf{P}(t_0 < T < t_1) = \mathbf{P}(t_0 < X/H < t_1) = \mathbf{P}(t_0 H < X < t_1 H) = [1 - (1-p)^{t_1 H}] - [1 - (1-p)^{t_0 H}] = (1-p)^{t_0 H} - (1-p)^{t_1 H}$.
+
+The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, i.e., $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, where $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$.
+These conditions give the following equations: $1 - e^{-t_0/600} = \alpha/2$, and $e^{-t_1/600} = \alpha/2$.
+Solving these equations, we have $t_0 = -600 \ln{(1 - \alpha/2)}$, and $t_1 = -600 \ln{(\alpha/2)}$.
+
+For $\alpha = 10\%$, $t_0 = 30.77$ and $t_1 = 1797.44$, or $[0.51, 30.76]$ in minutes.
+
+Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on.
+Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i \sim \text{Erlang}(n, \frac{1}{600})$.
+Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.
+
+Many exchanges use the rule of 6 confirmations.
+So, for $n=6$, we have: $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$.
+The symmetrical confidence interval with level $\alpha=10\%$ is $[27, 105]$ in minutes.
+
+% Probabilidade de orfão
+An orphan block would be created if a new block is found during the propagation time of a new block.
+Let $\alpha$ be the percentage of the total hash rate of the node which is outdated, and $\Delta t$ the propagation time in seconds.
+Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{600}}$.
+For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, than $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.
+
+\section{Analysis of network's hash rate change}
+% Análise: Mudança no network's hash rate antes de atualizar o $A$.
+
+The given number $A$ is adjusted every 2016 blocks.
+So, let's analyze what happens when the network's hash rate changes before the adjustment.
+First, $\mathbf{P}(13 \text{ days} < Y_{2016} < 15 \text{ days}) = \mathbf{P}(13 \cdot 24 \cdot 3600 < Y_{2016} < 14 \cdot 24 \cdot 3600) = 0.9986$, i.e., it will mostly happen between 13 and 15 days, assuming that the network's hash rate remains constant.
+But, as the network's total hash rate remained constant, the adjusted $A$ will not be changed.
+
+%Let's assume that the network's total hash rate will change only once, from $H$ to $\alpha H$ -- so, if $\alpha = 1$, there would be no change, if $\alpha > 1$, there would be an increase, and if $\alpha < 1$, there would be a decrease.
+As the time required to find a new block would be $T_\alpha = \frac{X}{\alpha H}$, we got the same random variable as we have gotten when analyzing a miner with $\alpha$ percent of the network's total hash rate.
+But, the analysis will be different because here $\alpha$ may be greater than one.
+
+Let $H \cdot u(t)$ be the network's total hash rate over time.
+So, the number of hashes calculated in $t$ seconds is $H \int_0^t u(t) dt$.
+Hence, $\mathbf{P}(T \leq t) = \mathbf{P}(X \leq H \int_0^t u(t)dt)$.
+When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{600} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{600} \cdot e^{-\frac{1}{600} \int_0^t u(t)dt}$.
+
+Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
+So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ follows an exponential distribution with $\lambda = \frac{\alpha}{600}$.
+Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha$ follows an Erlang distribution.
+The time required to find 2016 blocks will be divided by $\alpha$, and the variance will be divided by $\alpha^2$.
+When the network's hash rate increases ($\alpha > 1$), the variance decreases.
+When the network's hash rate decreases ($\alpha < 1$), the variance increases.
+
+For example, if $\alpha = 2$, $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved.
+If $\alpha = 0.5$, $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled.
+So, even if half of the network stops mining, it will take at most 29 days to get back to normal.
+
+% TODO Step-wise analysis
+
+\subsection{Hash rate smoothly changing}
+
+Let $u(t) = \frac{1+abx}{1+bx}$.
+It is an useful function because $u(0) = 1$ and $\lim_{t \rightarrow \infty} u(t) = a$.
+The bigger the $b$, the faster $u(t) \rightarrow a$.
+For example, if $a=2$, it means $H$ would be smoothly doubling.
+If $a=0.5$, it means $H$ would be smoothly halving.
+
+It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a \Rightarrow \int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
+$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
+$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
+
+Assuming that $n = \frac{\lambda(a-1)}{b}$ is integer, we have:
+$$F_T(t) = 1 - (1+bt)^n e^{-\lambda at}$$
+
+Thus,
+\begin{align*}
+\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
+	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
+	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
+	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
+	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
+\end{align*}
+
+Hence, as $\mathcal{L}\{f_T(t)\} = s \mathcal{L}\{F_T(t)\}$,
+
+$$\mathcal{L}\{f_T(t)\} = 1 - \sum_{k=0}^n \binom{n}{k} \frac{s b^k k!}{(s+\lambda a)^{k+1}}$$
+
+Then,
+
+\begin{align*}
+\frac{d}{ds} \mathcal{L}\{f_T(t)\} 
+	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{d}{ds} \frac{s}{(s+\lambda a)^{k+1}} \\
+	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \left[ \frac{1}{(s+a\lambda)^{k+1}} - \frac{s(k+1)}{(s+a \lambda)^{k+1}} \right] \\
+\frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0} 
+	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{1}{(\lambda a)^{k+1}} \\
+	&= - \frac{1}{a\lambda} \sum_{k=0}^n \binom{n}{k} k! \left( \frac{b}{\lambda a} \right)^k \\
+	&= - \frac{1}{a\lambda} \sum_{k=0}^n \frac{n!}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \\
+	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \right] \\
+	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{n-k} \right] \tag{$k \rightarrow n-k$} \\
+	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{-k} \right] \\
+	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right]
+\end{align*}
+
+Finally, as $\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0}$,
+$$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n!
+\left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right] \text{, where $n = \frac{\lambda(a-1)}{b}$}$$
+
+Let's check this equation for already known scenarios.
+When $a=1$, then $n=0$ and $\mathbf{E}[T] = 1/\lambda$.
+When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied.
+In fact, $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.
+
+\begin{theorem}
+	$$a > 1, x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
+\end{theorem}
+\begin{proof}
+$x > M \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM}$.
+As $1-a<0$, $\frac{1-a}{1+bx} > \frac{1-a}{1+bM}$.
+Thus, $\frac{1-a}{1+bM} < \frac{1-a}{1+bx} + a - a = \frac{1+abx}{1+bx} - a < 0 < \frac{a-1}{1+bM}$.
+Hence, $-\frac{a-1}{1+bM} < \frac{1+abx}{1+bx} - a < \frac{a-1}{1+bM}$.
+\end{proof}
+
+For instance, if we would like to know the impact of smoothly double the hash rate in the next week, $\lambda = 1/600$, $a=2$ and $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$.
+$b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01$.
+So, $b > 0.000163690$ and $n<10.1818$.
+So, for $n=10$, $b=0.000166666$, and $\epsilon = 0.009823 < 0.01$, as expected.
+Finally, $\mathbf{E}[T] = 557.65$.
+If the hash rate had suddenly doubled, the average time between blocks would be 300.
+
+%$$\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0} = \frac{b^n n! 600^{n+1}}{a^{n+1}} e^{\frac{a}{600b}} \text{, where $n = \frac{a-1}{600b}$}$$
+
+%$a > 1, x > M \Rightarrow 1 + bx > 1 + bM \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM} \Rightarrow \frac{1-a}{1+bx} + a > \frac{1-a}{1+bM} + a \Rightarrow \frac{1+abx}{1+bx} - a > \frac{1-a}{1+bM}$.
+
+\subsection{Hash rate changing linearly}
+
+Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(t) = 1 + a^2t$.
+Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{t + a^2t^2/2}{600}}$.
+
+In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.
+
+Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.
+
+\begin{align*}
+	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
+		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{600}} \right) dx \\
+		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{600}} dx \\
+		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{600}} dx \\
+\end{align*}
+
+Let $b=\frac{a}{20\sqrt{3}}$ and $c = \frac{10\sqrt{3}}{a} \left( s + \frac{1}{600} \right)$, then $sx + \frac{x+a^2x^2/2}{600} = (bx + c)^2 - c^2$.
+Let $y = bx+c \Rightarrow dy = bdx$.
+
+\begin{align*}
+	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{600}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
+		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
+		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
+		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
+\end{align*}
+
+$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
+
+Hence,
+$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
+
+As $\frac{d}{ds}c = \frac{10\sqrt{3}}{a}$,
+
+\begin{align*}
+\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{20\sqrt{3\pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{10\sqrt{3\pi}s}{2ab} \\
+\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{1200a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right) \\
+\mathbf{E}[T] &= \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{1200a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right)
+\end{align*}
+
+It is easy to notice that $lim_{a \rightarrow 0} \mathbf{E}[T] = 600$, and $lim_{a \rightarrow \infty} \mathbf{E}[T] = 0$.
+
+$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.
+
+\textcolor{red}{I do not know how to interpret this.}
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{images/bitcoin-H-increasing.png}
+\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
+\end{figure}
+
+\section{Attack in the Bitcoin network}
+% Ataque submarino
+% Probabilidade de "correr por fora"
+
+Attackers are not part of the mining network, because they are trying to generate a different longest chain.
+
+Let $\beta H$ be the hash rate of the attackers, and $\gamma H$ be the network's hash rate without the attackers.
+Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{600}$ and $\lambda_{\text{network}} = \frac{\gamma}{600}$, respectively.
+
+As \cite{nakamoto2008bitcoin} have done, we will also model the attack using the Gambler's Ruin.
+In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$.
+The rounds are independent.
+The gambler starts with \$$k$ plays continuously until he either accumulates a target amount of \$$m$, or loses all his money.
+Let $\rho = \frac{1-p}{p}$, then the probability of losing his fortune is:
+
+$$
+\mathbf{P}(\text{losing his fortune}) =
+\begin{cases}
+	\frac{\rho^k - \rho^m}{1-\rho^m} \text{, if $\rho \ne 1$,} \\
+	\frac{m-k}{m} \text{, if $\rho = 1$.}
+\end{cases}
+$$
+
+When $m \rightarrow +\infty$,
+
+$$
+\mathbf{P}(\text{losing his fortune}) =
+\begin{cases}
+	\rho^k \text{, if $\rho < 1$,} \\
+	1 \text{, if $\rho \geq 1$.}
+\end{cases}
+$$
+
+% TODO Expected time to absorption.
+
+The gambler wins \$1 is the same as the network find a new block, the gambler losses \$1 is the same as the attacker find a new block.
+The initial \$$k$ is the same as the number of blocks the attacker is behind the network.
+Thus, the gambler losses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.
+
+In our case, $p = \frac{\lambda_{\text{network}}}{\lambda_{\text{network}} + \lambda_{\text{attacker}}} = \frac{\gamma}{\beta + \gamma}$, thus $\rho = \frac{\beta}{\gamma}$.
+Hence, $\rho < 1 \Leftrightarrow \beta < \gamma$.
+
+Suppose that the attacker is mining with the network.
+Suddently, he stops mining with the network and starts attacking, i.e., starts to mine in another chain.
+In this scenario, since the attacker's hash rate is not mining with the network anymore, $\gamma = 1 - \beta$.
+Thus, $\beta < \gamma \Rightarrow \beta < 0.5 \Leftrightarrow \rho < 1$.
+Here comes the conclusion that, if the attacker has 50\% or more of the network's hash rate, then his attack will be certainly successful.
+We got exactly the same equations and conclusions as \cite{nakamoto2008bitcoin}.
+
+But this scenario seems not to be the optimal attack, because the attacker has waited the $k$ confirmations before starting the attack.
+A better approach would be to start attacking just after propagating the transaction.
+In this case, our previous model is not good, because even if the attacker have found more blocks than the network, he cannot propagate those blocks before the network has found $k$ confirmations.
+So, we have to model the probabilities before the network has found the $k$ block.
+Then, if the attacker has more blocks than the network, he has successfully attacked.
+Otherwise, we got back to the previous model, in which the attacker must still find more blocks.
+
+\begin{theorem}
+	Let $a_{ij}$ be the probability of moving from state $(i, j)$ to state $(s, k)$ in the Markov chain.
+	$$a_{ij} = \binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}.$$
+\end{theorem}
+\begin{proof}
+
+In order to go from $(i, j)$ to $(s, k)$, the attacker must find exacly $i-s$ blocks, and the network must find exactly $k-j$ blocks.
+It is as they would be walking the grid from the point $(i, j)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}.
+There are many possible paths and all of them occurs with probability $(1-p)^{s-i} p^{k-j}$.
+
+The walking ends when $(\cdot, k$) is reach, i.e., when the network finds $k$ blocks, no matter how many blocks the attacker has found.
+So, it is not allowed to walk through the line $(\cdot, k)$.
+Thus, the number of paths between $(i, j)$ and $(s, k)$ going only up or right, without going through the line $(\cdot, k)$ is exactly the number of paths between $(i, j)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s-i$ right movements ($r$) and $k-j-1$ up movements ($u$).
+The number of permutations is $\frac{(k-j-1+s-i)!}{(s-i)!(k-j-1)!} = \binom{(k-j-1)+(s-i)}{s-i}$ because there are $s-i$ repetitions of the element $r$ and $k-j-1$ repetitions of the element $u$.
+
+Finally, the probability is $\binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}$.
+
+\begin{figure}[ht]
+  \centering
+  \begin{tikzpicture}
+    \coordinate (Origin)   at (0,0);
+    \coordinate (XAxisMin) at (-1,0);
+    \coordinate (XAxisMax) at (5,0);
+    \coordinate (YAxisMin) at (0,-1);
+    \coordinate (YAxisMax) at (0,5);
+    %\draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
+    %\draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis
+
+    \clip (-0.5,-0.5) rectangle (10.5cm,7cm); % Clips the picture...
+    %\pgftransformcm{1}{0}{0}{1}{\pgfpoint{0cm}{0cm}}
+          % This is actually the transformation matrix entries that
+          % gives the slanted unit vectors. You might check it on
+           % MATLAB etc. . I got it by guessing.
+    \draw[style=help lines,dashed] (0,0) grid[step=1cm] (14,6);
+          % Draws a grid in the new coordinates.
+          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
+              % Puts the shaded rectangle
+    \foreach \x in {0,1,...,10}{% Two indices running over each
+      \foreach \y in {0,1,...,6}{% node on the grid we have drawn 
+        \node[draw,circle,inner sep=1pt,fill] at (\x,\y) {};
+            % Places a dot at those points
+      }
+    }
+
+    \draw [ultra thick,-latex,red] (0,0) -- (0,1) node [midway, left] {};
+    \draw [ultra thick,-latex,red] (0,1) -- (0,2) node [midway, left] {};
+    \draw [ultra thick,-latex,red] (0,2) -- (1,2) node [midway, above] {};
+    \draw [ultra thick,-latex,red] (1,2) -- (1,3) node [midway, right] {};
+    \draw [ultra thick,-latex,red] (1,3) -- (2,3) node [midway, right] {};
+    \draw [ultra thick,-latex,red] (2,3) -- (2,4) node [midway, right] {};
+    \draw [ultra thick,-latex,red] (2,4) -- (2,5) node [midway, right] {};
+    \draw [ultra thick,-latex,red] (2,5) -- (3,5) node [midway, right] {};
+	\draw [ultra thick,-latex,red] (3,5) -- (3,6) node [above] {(3, 6)};
+
+    \draw [ultra thick,-latex,blue] (0,0) -- (1,0) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (1,0) -- (2,0) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (2,0) -- (2,1) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (2,1) -- (2,2) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (2,2) -- (2,3) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (2,3) -- (3,3) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (3,3) -- (4,3) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (4,3) -- (5,3) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (5,3) -- (5,4) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (5,4) -- (6,4) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (6,4) -- (6,5) node [midway, left] {};
+    \draw [ultra thick,-latex,blue] (6,5) -- (7,5) node [midway, left] {};
+	\draw [ultra thick,-latex,blue] (7,5) -- (7,6) node [above] {(7,6)};
+  \end{tikzpicture}
+  \caption{Both the attacker and the network are mining.
+Each step up is a new block found by the network with probability $p$.
+Each step right is a new block found by the attacker with probability $1-p$.
+It ends when the network finds $k$ blocks -- in this example, $k=6$.
+The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$.
+Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network.
+In the red path, the attacker still have to catch up 3 blocks to have a successfull attack, which happens with probability $\rho^3$, if $p < 0.5$.}
+  \label{figure:bitcoin-attack-paths}
+\end{figure}
+\end{proof}
+
+\begin{cor}
+	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
+\end{cor}
+\begin{proof}
+	This probability $\mathbf{P}(S = s) = a_{00}$.
+\end{proof}
+
+Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.
+
+\begin{theorem}
+	$$\mathbf{P}(S \geq k) = 1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} (1-p)^s p^k.$$
+\end{theorem}
+\begin{proof}
+	Let's use the following identity:
+	$$\frac{1}{(1-z)^{a+1}} = \sum_{i=0}^{\infty} \binom{i+a}{i} z^i \text{, for $|z|<1$}$$
+
+	Thus, replacing $z=1-p$, $i=s$, and $a=k-1$, we have:
+
+	$$\frac{1}{p^{k}} = \sum_{s=0}^{\infty} \binom{s+k-1}{s} (1-p)^s$$
+	$$1 = \sum_{s=0}^{\infty} \binom{s+k-1}{s} (1-p)^s p^k.$$
+
+	Now, just split $\sum_{s=0}^{\infty} = \sum_{s=0}^{k-1} + \sum_{s=k}^{\infty}$ and it is done.
+\end{proof}
+
+Using this last theorem, we moved from an infinity sum to a finity sum.
+
+\begin{theorem}
+	Let $p = \frac{\gamma}{\beta + \gamma}$.
+$$
+\mathbf{P}(\text{successful attack}) =
+\begin{cases}
+%1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} (1-p)^s p^k \left(1 - \left( \frac{1-p}{p} \right)^{k-s} \right)
+	1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} \left( (1-p)^s p^k - (1-p)^k p^s \right) \text{, $p \geq 0.5$} \\
+	1 \text{, $p < 0.5$}.\\
+\end{cases}
+$$
+\end{theorem}
+\begin{proof}
+$$
+\mathbf{P}(\text{successful attack}) = \mathbf{P}(S \geq k) + \sum_{i=0}^{k-1} \mathbf{P}(s=i) \rho^{k-i}
+$$
+\end{proof}
+
+For $k=6$, $p=0.9$, $\mathbf{P}(\text{successful attack}) = 0.0005914121600000266$.
+
+For $k=6$, $p=0.7$, $\mathbf{P}(\text{successful attack}) = 0.15644958192000014$.
+
+\begin{figure}[ht]
+\centering\includegraphics[width=\textwidth]{images/fig-bitcoin-attack.png}
+\caption{Probability of a successful attack according to the network's hash rate of the attacker ($\beta$).\label{fig-bitcoin-attack}}
+\end{figure}
+
+%The number of blocks found in a given time interval $\Delta t$ is given by the Poisson distribution.
+So, the number of blocks found by the attackers $N_{\text{attacker}}$ follows a Poisson distribution with parameter $\lambda_{\text{attacker}} \Delta t = \frac{\beta \Delta t}{600}$, and the number of blocks found by the network $N_{\text{network}}$ follows a Poission distribution with parameter $\lambda_{\text{network}} \Delta t = \frac{\gamma \Delta t}{600}$.
+We would like to calculate $\mathbf{P}(N_{\text{attackers}} - N_{\text{network}} \geq k)$.
+
+%$\mathbf{P}(Y_{n+k}^{\text{attacker}} < Y_{n}^{\text{network}})$
+
+\section{Confirmation time and network capacity}
+% Tempo na fila para confirmar transação
+
+Let's say that when a new transaction is propagated it is enqueued in the unconfirmed transaction queue.
+Then, when a new block is found, some of these transactions in the queue are confirmed.
+We are interested in some measures of the queue, like the expected time to confirm a transaction and the queue's length.
+
+Let's assume that all transactions have exactly the same size $S$ and pays exactly the same fee.
+If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transaction in each block.
+Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single zero of $z^s(1+m(1-z))-1$ with $|z_s|>1$.
+The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$.
+Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$.
+The probabilities form a simple geometric series, and so they are exponentially decreasing.
+We may interpret it as a stable system.
+In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/600$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
+Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
+When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.
+
+The average waiting time of a transaction to be confirmed, when $m < s$, is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.
+
+When $m \ll s$, $z_s \rightarrow 1 + 1/m$.
+So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = 600 \text{ seconds}$.
+In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.
+
+When $\lambda_{\text{TX}} \rightarrow s$, $z_s \rightarrow 1$. So, $\mathbf{E}(\pi) \rightarrow +\infty$.
+
+Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/600$ transactions per second, where $s$ is the average number of transactions per block.
+
+\section{Fork analysis}
+
 \chapter{DAG model}
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-example.pdf}
+\centering\includegraphics[width=\textwidth]{images/fig-tangle-example.pdf}
 \caption{White nodes represent transactions that have been confirmed at least once.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-swarm.pdf}
+\centering\includegraphics[width=\textwidth]{images/fig-tangle-swarm.pdf}
 \caption{Suddenly the number of transactions per second increases and the width of the swarm grows.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-conflict.pdf}
+\centering\includegraphics[width=\textwidth]{images/fig-tangle-conflict.pdf}
 \caption{The red nodes are transactions which had some conflict with previous transaction and were invalidated by the network. Notice that none of them have been confirmed.\label{fig-tangle-conflict}}
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-hist.png}
+\centering\includegraphics[width=\textwidth]{images/fig-tangle-hist.png}
 \caption{Histogram of how long has been a transaction waiting until its first confirmation.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images/fig-tangle-hist-2.png}
+\centering\includegraphics[width=\textwidth]{images/fig-tangle-hist-2.png}
 \caption{Histogram of how long has been a transaction waiting until its first confirmation.
@@
 
-\section{Proposal: Theorems to be proved}
+\chapter{Analysis of DAG Model}
+
+In order to a further understanding of the DAG Model, we will analyze the mathematical model behind it.
+
+The DAG Model's mining starts just like Bitcoin's, i.e., $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.
+Then, it follows that solving a transaction proof-of-work follows a geometric distribution.
+Let $X$ be the number of attempts until a success, then $\mathbf{P}(X=k) = (1-p)^{k-1}p$, where $p=\frac{A}{2^{256}}$.
+Also, $\mathbf{P}(X \leq k)=1-(1-p)^k$.
+The average number of attempts is $\mathbf{E}(X)=1/p$.
+
+Let $w$ be the transaction's weight, i.e., $\mathbf{E}(X)=w \Rightarrow w = 1/p$.
+Let $T = X/H$ be the time required to solve the transaction's proof-of-work, where $H$ is the hash rate of the transaction's owner.
+Then, $\mathbf{P}(T < t) = \mathbf{P}(X < tH) = 1 - (1 - p)^{tH} = 1 - \left( 1 - \frac{1}{w} \right)^{tH} = 1 - e^{tH \log\left(1 - \frac{1}{w} \right)}$, i.e., $T$ follows an exponential distribution with $\lambda = -H \log\left(1 - \frac{1}{w} \right)$.
+
+% https://math.stackexchange.com/questions/635443/sum-of-exponential-random-variable-with-different-means
+
+The number of transactions solving the proof-of-work may be modeled by a Birth and Death process, since if we have $k$ transactions solving the proof-of-work only two things may happen: either a new transaction will be created or one the transactions will finish solving the proof-of-work.
+Let the time between new transactions follows an exponential distribution with $\mu$ parameter, then the probability of a new transaction emerges before any transaction finishes its proof-of-work is $q_k = \mathbf{P}(T_{\text{new tx}} = \min\{T_{\text{new tx}}, T_1, T_2, \dots, T_k\}) = \frac{\mu}{\mu + \sum_{i=1}^k \lambda_i}$.
+Hence, the probability of any transaction finishes its proof-of-work before a new transaction emerges is $p_k = 1 - q_k = \frac{\sum_{i=1}^k \lambda_i}{\mu + \sum_{i=1}^k \lambda_i}$.
+
+Every time a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work.
+When it finishes solving the proof-of-work, it is propagated and becomes a tip.
+So, two new transactions may choose the same tips to confirm.
+If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions -- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
 

### 2018-01-18 Chapters/Chapter01.tex
Update on Overleaf.

@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/bitcoin-H-increasing.png}
+\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
 \caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/fig-bitcoin-attack.png}
+\centering\includegraphics[width=\textwidth]{./images01/fig-bitcoin-attack.png}
 \caption{Probability of a successful attack according to the network's hash rate of the attacker ($\beta$).\label{fig-bitcoin-attack}}
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/fig-tangle-example.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-example.pdf}
 \caption{White nodes represent transactions that have been confirmed at least once.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/fig-tangle-swarm.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-swarm.pdf}
 \caption{Suddenly the number of transactions per second increases and the width of the swarm grows.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/fig-tangle-conflict.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-conflict.pdf}
 \caption{The red nodes are transactions which had some conflict with previous transaction and were invalidated by the network. Notice that none of them have been confirmed.\label{fig-tangle-conflict}}
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/fig-tangle-hist.png}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist.png}
 \caption{Histogram of how long has been a transaction waiting until its first confirmation.
@@
 \begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{images/fig-tangle-hist-2.png}
+\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist-2.png}
 \caption{Histogram of how long has been a transaction waiting until its first confirmation.

### 2018-01-19 Chapters/Chapter01.tex
Update on Overleaf.

@@
 \chapter{Introduction}
-
-Bitcoin (BTC) is the first digital currency, also known as digital money, internet money and cryptocurrency.
-It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank.
-Bitcoin is distributed since its ledger is public and is stored in thousands of computers.
-It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community.
-The security of Bitcoin relies on digital signature technology, so the owner of some BTC tokens is, in fact, the owner of a pair of private and public keys.
-One may either use the private key to transfer some tokens to another, or use the public key to a proof-of-ownership.
-If one looses their private key, unfortunately, access to their coins will be lost forever.
-
-Since 2009, Bitcoin has been growing and becoming widely used all around the world.
-It started as an experiment based in a seminal work from \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a \$36 billion market capitalization \citep{coinmarketcapbtc}.
-There are hundreds of companies investing in different uses of the technology, from exchanges to debit cards, and billions of dollars being invested in the new markets created by Bitcoin.
 
@@
 
+Bitcoin (BTC) is the first digital currency, also known as digital money, internet money and cryptocurrency.
+It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank.
+Bitcoin is distributed since its ledger is public and is stored in thousands of computers.
+It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community.
+The security of Bitcoin relies on digital signature technology and network agreement.
+While digital signature ensures ownership, i.e., the funds may only be spend by their owners, and nobody else; the network agreement both prevents double spending and ensures funds.
+In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution.
+When all these requirements are met, the funds are transfered.
+
+According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending.
+The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the e-cash literature.
+
+Since 2009, Bitcoin has been growing and becoming widely used all around the world.
+It started as an experiment based in a seminal work by \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a highly volatile \$192 billion market capitalization, as of this writing \citep{coinmarketcapbtc}.
+There are hundreds of companies investing in different uses of the technology, from exchanges to debit cards, and billions of dollars being invested in the new markets based on Bitcoin's technology.
+
 Despite Bitcoin's huge success, there are still many challenges to be overcome.
-The network must scale to support hundreds of transactions per second, but today its capacity is close to 8 transactions per second.
+The network must scale to support hundreds of transactions per second, but, as we will see, today its capacity is close to 8 transactions per second.
 Another important challenge is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
+
+% TODO Include incentives.
 
@@
 Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behaviour, and it is seen as a problem to be solved.
-
-According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending.
-The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the e-cash literature.
 
@@
 The proof-of-work is a mathematical problem with the following characteristics: (i) it is hard to find a solution; (ii) this difficulty level may be adjusted; and (iii) it is fast to check whether the proposed solution is correct.
+
+% Describe the basics of SHA-256 algorithm.
 
@@
 
+The digital signature scheme used by Bitcoin is based on a pair of private and public keys.
+The private key is used to sign the transaction, while the public key is used to check whether the signature is valid.
+Thus, the owner of some Bitcoin funds is, in fact, the owner of a pair of private and public keys.
+The private key must never be publicly published, as whoever has access to the private key is able to spend its funds.
+In other words, in order to protect their funds, the owners must protect the private key.
+If one looses their private key, unfortunately, access to their funds will be lost forever.
+The public key may be used to a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.
+
 Bitcoins (BTCs) owned by someone are, in fact, unspent outputs in one or more transactions.
@@
 This enables $O(1)$ unspent transaction lookup.
+% Improve explanation of O(1).
 

### 2018-01-19 Chapters/Chapter01.tex
Update on Overleaf.

@@
 The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the e-cash literature.
+
+Bitcoin provides interesting incentives to all players.
+On one hand, users have incentive to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered -- but only in the size in bytes of the transaction --; (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks -- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
+On the other hand, miners have incentive to mine Bitcoin because
 

### 2018-01-20 Chapters/Chapter01.tex
Update on Overleaf.

--- 2018-01-19 Chapters/Chapter01.tex
@@
 The main problem when one is trying to create a digital money is how to prevent double spending.
-As the money is digital, what would prevent counterfeiting?
+As the money is digital, and copies can be made \textit{ad nauseam}, what can prevent counterfeiting?
 That is, what would prevent users from sending copies of the same money to two (or more) people?
@@
 The security of Bitcoin relies on digital signature technology and network agreement.
-While digital signature ensures ownership, i.e., the funds may only be spend by their owners, and nobody else; the network agreement both prevents double spending and ensures funds.
+While digital signature ensures ownership, i.e., the funds may only be spend by their owners, and nobody else; the network agreement both prevents double spending and ensures that all processed transactions have sufficient funds.
 In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution.
@@
 According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending.
-The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the e-cash literature.
+The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the previous e-cash philosophies.
+% literature
 
-Since 2009, Bitcoin has been growing and becoming widely used all around the world.
+Bitcoin provides interesting incentives to all players, namely the users and the miners.
+On one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered -- but only in the size (in bytes) of the transaction --; (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks -- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
+On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes and they may also receive the fees of unconfirmed transactions.
+These incentives have kept the Bitcoin network up and running since 2009 with barely no interruptions (99.99\% uptime).
+
+Since 2009, Bitcoin has been growing and becoming more and more used all around the world.
 It started as an experiment based in a seminal work by \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a highly volatile \$192 billion market capitalization, as of this writing \citep{coinmarketcapbtc}.
@@
 Despite Bitcoin's huge success, there are still many challenges to be overcome.
-The network must scale to support hundreds of transactions per second, but, as we will see, today its capacity is close to 8 transactions per second.
-Another important challenge is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
+We will focus on a specific subset of those challenges, namely the scaling, decentralization, and spam challenges.
+One important challenge that we will skip is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
 
-% TODO Include incentives.
+The network must scale to support hundreds of transactions per second, while its capacity is around only eight transactions per second.
+Thus, the more Bitcoin becomes popular, the more saturated the network is.
+Network saturation has many side effects and may affect the players incentive to keep the network running.
+The transaction fees have to be increased to compete for faster confirmation.
+The pool of unconfirmed transactions grows indefinitely, which may cause some transactions to be discarded due to low memory space available, as the previously predictable confirmation time of transactions becomes unpredictable.
 
-Bitcoin has the most decentralized network between the cryptocurrencies, even so there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
-Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behaviour, and it is seen as a problem to be solved.
+Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
+Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behaviour, and it is also seen as a problem to be solved.
+The more decentralized, the more trustworthy Bitcoin is.
 
-The number of ideas and publications focusing in improving Bitcoin's design is increasing every day.
+Generating new transactions in Bitcoin has a tiny cost, because one only has to generate the transaction itself, digitally sign it, and propagate in the Bitcoin network.
+On one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks.
+One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network.
+
+The number of ideas and publications focusing in improving Bitcoin's design and overcoming those challenges is increasing every day.
 Many of these proposals are organized into BIPs (Bitcoin Improvement Proposals) which are discussed and implemented by the community; while others come in the form of whitepapers and alternative software forks (which would include the need of a protocol upgrade).
-Other scaling proposals are published in blogs and forums, describing new cryptocurrencies.
-Bitcoin's community hardly ever publishes their ideas in academic journals.
+Other proposals are published in blogs and forums, describing new cryptocurrencies.
+Bitcoin's community hardly ever publishes their ideas in academic journals, preferring instead, of BIPs, white papers, and web discussions.
 
-% Spam problem.
+After the launch of Bitcoin, more than 1,000 other cryptocurrencies has been created (REF).
+In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain.
+Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}.
+We are specially interested in one of them: Iota.
 
-%The DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015} has a different design which has neither mining nor confirmation blocks and transaction fees.
-Each transaction has its own proof-of-work, and is used to confirm other transactions in a directed acyclic graph.
-Although this design has been used by a few cryptocurrencies, none of them has enough users and would strengthen that the problems have been solved.
+Iota uses a DAG model, called tangle, which has a different design than Bitcoin's blockchain.
+It has neither mining nor confirmation blocks and transaction fees.
+Each transaction has its own proof-of-work, and is used to confirm other transactions forming a directed acyclic graph.
+In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions.
+Hence, theoretically, it scales to any large number of transactions per second.
+The scaling problem of tangle is exactly the opposite of Bitcoin's.
+It must have at least a given number of transaction per seconds, otherwise the transactions are not confirmed and the cryptocurrency does not work.
 
-The DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015} has different design which has neither mining nor confirmation blocks and transaction fees.
-Each transaction has its own proof-of-work, and is used to confirm other transactions in a directed acyclic graph.
-The present work intends to analyze whether the DAG model may be a viable solution to both Bitcoin's scaling problem and Bitcoin's oversized influence of a few miners and mining pools.
+The present work intends to analyze a new architecture which lies between Bitcoin and Iota and may be a viable solution to both Bitcoin's scaling, centralization, and spam problems.
+We also present a mathematical analysis of Bitcoin's mining, forking, and safety.
 
@@
 
-% Describe the basics of SHA-256 algorithm.
-
-Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given number.
+Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given number $A$.
 As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}.
-Thus, if the given number is $2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero).
-But if the given number is $2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero).
-Hence, finding a solution is a hard problem which difficulty depends on the given number.
-The lesser the given number, the higher the difficulty.
+Thus, if the given number is $A = 2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero).
+But if the given number is $A = 2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero).
+Hence, finding a solution is a hard problem which difficulty depends on the given number $A$.
+The lesser the given number $A$, the higher the difficulty.
 On the other side, checking whether a solution is correct is fast, one just has to apply the SHA-256 twice and compare.
@@
 This enables $O(1)$ unspent transaction lookup.
-% Improve explanation of O(1).
 
@@
 
+Next we will do a mathematical analysis of Bitcoin in order to deeply understand its minings properties, how a fork would affect the network, and its security against attackers.
+
 \chapter{Analysis of Bitcoin}
@@
 
-A hash function $h: \{0, 1\}^\infty \rightarrow \{0, 1\}^{n}$ has the following properties:
+Hash functions has been widely studied in computer science. In short, a hash function $h: \{0, 1\}^\infty \rightarrow \{0, 1\}^{n}$ has the following properties:
 
@@
 
+Bitcoin uses two hash functions: HASH-160 and HASH-256.
+The first has $n=160$ and consists of the composition of \textit{SHA-256} and \textit{RIPEMD-160}.
+The latter, has $n=256$ and applies \textit{SHA-256} twice.
+
+For further information, see \citet{gilbert2003security, dobbertin1996ripemd}.
+
 \section{Bitcoin mining}
 
-Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin hash function.
+Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin \textit{HASH-256} function.
 The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given number.
@@
 The average number of attempts is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = \frac{1-p}{p^2}$.
-
-%As one may fail many times until they find a solution, the probability of finding a solution in the $K$ trial follows a geometric distribution.
-Let $X$ be the number of trials until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \mathbf{P}(a < A)$, i.e., there would be $k-1$ failures until a success in the $k-th$ trial.
-The average number of trials is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = (1-p)/p^2$.
 
@@
 
-% A partir de qual valor de H é razoável dizer que T segue uma distribuição exponencial?
+Now, we would like to understand from which value of $H$ it is reasonable to assume that $T$ follows an exponential distribution.
+
 \begin{theorem}
@@
 
+The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, where $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$.
+These conditions give the following equations: $1 - e^{-t_0/600} = \alpha/2$, and $e^{-t_1/600} = \alpha/2$.
+Solving these equations, we have $t_0 = -600 \ln{(1 - \alpha/2)}$, and $t_1 = -600 \ln{(\alpha/2)}$.
+
+For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ -- or $[0.51, 30.76]$ in minutes. Thus, 90\% of the time between blocks blocks are between 30 seconds and 31 minutes.
+
+Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on.
+Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$.
+As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{600}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{600})$.
+Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.
+
+Many exchanges require at least six confirmations in order to accept a deposit in Bitcoin.
+So, for $n=6$, $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$.
+The symmetrical confidence interval with $\alpha=10\%$ is $[27, 105]$ in minutes.
+Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted -- assuming that your transaction will be confirmed in the very next block.
+
 Let's analyze the probability of finding a new block for a miner who has $\alpha$ percent of the network's total hash rate.
@@
 \begin{theorem}
-	When one miner with $\alpha$ percent of the network's total hash rate multiplies their hash rate by $m$, the probability of finding the next block is multiplied by $\frac{m}{m \alpha + 1 - \alpha}$.
+	When one miner with $\alpha$ percent of the network's total hash rate multiplies their hash rate by $m$, the probability of this miner find the next block is multiplied by $\frac{m}{m \alpha + 1 - \alpha}$.
 	\label{thm-miner-multiply}
@@
 
-That way, it is not exacly correct to say that when one doubles their hash rate, their probability will double as well. It is only true for small miners.
+That way, it is not exactly correct to say that when one doubles their hash rate, their probability will double as well. It is only true for small miners.
 
 %Then, $\mathbf{P}(t_0 < T < t_1) = \mathbf{P}(t_0 < X/H < t_1) = \mathbf{P}(t_0 H < X < t_1 H) = [1 - (1-p)^{t_1 H}] - [1 - (1-p)^{t_0 H}] = (1-p)^{t_0 H} - (1-p)^{t_1 H}$.
-
-The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, i.e., $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, where $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$.
-These conditions give the following equations: $1 - e^{-t_0/600} = \alpha/2$, and $e^{-t_1/600} = \alpha/2$.
-Solving these equations, we have $t_0 = -600 \ln{(1 - \alpha/2)}$, and $t_1 = -600 \ln{(\alpha/2)}$.
-
-For $\alpha = 10\%$, $t_0 = 30.77$ and $t_1 = 1797.44$, or $[0.51, 30.76]$ in minutes.
-
-Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on.
-Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i \sim \text{Erlang}(n, \frac{1}{600})$.
-Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.
-
-Many exchanges use the rule of 6 confirmations.
-So, for $n=6$, we have: $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$.
-The symmetrical confidence interval with level $\alpha=10\%$ is $[27, 105]$ in minutes.
 
@@
 Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{600}}$.
-For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, than $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.
+For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.
 
@@
 
-The given number $A$ is adjusted every 2016 blocks.
-So, let's analyze what happens when the network's hash rate changes before the adjustment.
-First, $\mathbf{P}(13 \text{ days} < Y_{2016} < 15 \text{ days}) = \mathbf{P}(13 \cdot 24 \cdot 3600 < Y_{2016} < 14 \cdot 24 \cdot 3600) = 0.9986$, i.e., it will mostly happen between 13 and 15 days, assuming that the network's hash rate remains constant.
-But, as the network's total hash rate remained constant, the adjusted $A$ will not be changed.
+The difficulty, given by the number $A$, is adjusted every 2016 blocks.
+As, $\mathbf{P}(13 \text{ days} < Y_{2016} < 15 \text{ days}) = \mathbf{P}(13 \cdot 24 \cdot 3600 < Y_{2016} < 14 \cdot 24 \cdot 3600) = 0.9986$, it is expected that the total time to find 2016 blocks will be between 13 and 15 days, assuming that the network's hash rate remains constant.
+If it takes less than the expected time, it means that the network's total hash rate has increased.
+While if it takes more than the expected time, it means that the network's total hash rate has decreased.
+So, let's analyze what happens when the network's hash rate changes significantly.
 
@@
 Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
-So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ follows an exponential distribution with $\lambda = \frac{\alpha}{600}$.
-Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha$ follows an Erlang distribution.
-The time required to find 2016 blocks will be divided by $\alpha$, and the variance will be divided by $\alpha^2$.
-When the network's hash rate increases ($\alpha > 1$), the variance decreases.
-When the network's hash rate decreases ($\alpha < 1$), the variance increases.
+So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{600}$.
+Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{600})$.
+Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$.
+Hence, on one hand, when the network's hash rate increases ($\alpha > 1$), the 2016 blocks will be found earlier.
+On the other hand, when the network's hash rate decreases ($\alpha < 1$), the 2016 blocks will be found later.
 
-For example, if $\alpha = 2$, $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved.
-If $\alpha = 0.5$, $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled.
-So, even if half of the network stops mining, it will take at most 29 days to get back to normal.
+For example, if the network's total hash rate suddenly doubles ($\alpha = 2$), then $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved.
+On the other side, if the network's total hash rate suddenly halves ($\alpha = 0.5$), then $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled.
+It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, and then the adjustment will occur and everything will be back to the normal 10 minutes between blocks.
 
@@
 
-For instance, if we would like to know the impact of smoothly double the hash rate in the next week, $\lambda = 1/600$, $a=2$ and $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$.
-$b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01$.
-So, $b > 0.000163690$ and $n<10.1818$.
-So, for $n=10$, $b=0.000166666$, and $\epsilon = 0.009823 < 0.01$, as expected.
+For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$.
+So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected.
 Finally, $\mathbf{E}[T] = 557.65$.
+In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes.
 If the hash rate had suddenly doubled, the average time between blocks would be 300.

### 2018-01-20 Chapters/Chapter01.tex

@@
 %************************************************
+
+% TODO Include mining-pool proof of hashpower statistics.
+% TODO A transaction must confirm the outputs which it is spending (?)
 
@@
 Bitcoin provides interesting incentives to all players, namely the users and the miners.
-On one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered -- but only in the size (in bytes) of the transaction --; (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks -- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
+On one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction --- (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
 On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes and they may also receive the fees of unconfirmed transactions.
@@
 
-The proof-of-work is a mathematical problem with the following characteristics: (i) it is hard to find a solution; (ii) this difficulty level may be adjusted; and (iii) it is fast to check whether the proposed solution is correct.
+The proof-of-work is a mathematical problem with the following characteristics: (i) it is hard to find a solution; (ii) this hardness level may be adjusted; and (iii) it is fast to check whether the proposed solution is correct.
 
-Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given number $A$.
+Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given threshold $A$.
 As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}.
@@
 Hence, finding a solution is a hard problem which difficulty depends on the given number $A$.
-The lesser the given number $A$, the higher the difficulty.
+The lesser the given threshold $A$, the higher the difficulty.
 On the other side, checking whether a solution is correct is fast, one just has to apply the SHA-256 twice and compare.
@@
 In other words, in order to protect their funds, the owners must protect the private key.
-If one looses their private key, unfortunately, access to their funds will be lost forever.
+If one loses their private key, unfortunately, access to their funds will be lost forever.
 The public key may be used to a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.
@@
 For instance, when one has 3 BTCs in one transaction output and would like to send 1 BTC to a friend, they have to create a transaction with one input spending the 3 BTCs and two outputs, one for the friend with 1 BTC and one's change with 2 BTCs.
-This enables $O(1)$ unspent transaction lookup.
 
@@
 In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average.
-Suppose that the Bitcoin's network is able to calculate $H$ hashes per second -- $H$ is the total hash rate of the network.
+Suppose that the Bitcoin's network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network.
 The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block.
-So, the rule of finding a new block every 10 minutes (600 seconds) -- on average -- leads to the following equation: $\mathbf{E}(T) = 600$.
-So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = 600 \Rightarrow p = \frac{1}{600H}$.
-Finally, $\mathbf{E}(X) = 600H$, $\mathbf{E}(T) = 600$, $\mathbf{V}(X) = (600H)^2 - 600H$, and $\mathbf{V}(T) = 600^2 - 600/H$.
+So, the rule of finding a new block every 10 minutes ($\tau = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \tau = 600$.
+So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \tau = 600 \Rightarrow p = \frac{1}{\tau H}$.
+Finally, $\mathbf{E}(X) = \tau H$, $\mathbf{E}(T) = \tau$, $\mathbf{V}(X) = (\tau H)^2 - \tau H$, and $\mathbf{V}(T) = \tau^2 - \tau/H$.
 
-The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{600H} \right)^{tH}$.
-But, as the Bitcoin's network hash rate is really big, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{600}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{600}$.
+The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\tau H} \right)^{tH}$.
+But, as the Bitcoin's network hash rate is really big, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\tau}$.
 
 \begin{theorem}
-	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{600}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{600}}$.
+	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{\tau}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$.
 \end{theorem}
@@
 	\mathbf{P}(T \leq t) &= 1 - (1-p)^{tH} \\
-		     &= 1 - \left( 1 - \frac{1}{600H} \right)^{tH} \\
-	\text{Replacing $u = 600H$,} \\
-	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{600}} \\
-				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{600}} \\
-				&= 1 - \left( 1/e \right)^{\frac{t}{600}} \\
-				&= 1 - e^{-\frac{t}{600}}
+		     &= 1 - \left( 1 - \frac{1}{\tau H} \right)^{tH} \\
+	\text{Replacing $u = \tau H$,} \\
+	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{\tau}} \\
+				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{\tau}} \\
+				&= 1 - \left( 1/e \right)^{\frac{t}{\tau}} \\
+				&= 1 - e^{-\frac{t}{\tau}}
 \end{align*}
@@
 The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, where $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$.
-These conditions give the following equations: $1 - e^{-t_0/600} = \alpha/2$, and $e^{-t_1/600} = \alpha/2$.
-Solving these equations, we have $t_0 = -600 \ln{(1 - \alpha/2)}$, and $t_1 = -600 \ln{(\alpha/2)}$.
+These conditions give the following equations: $1 - e^{-t_0/\tau} = \alpha/2$, and $e^{-t_1/\tau} = \alpha/2$.
+Solving these equations, we have $t_0 = -\tau \ln{(1 - \alpha/2)}$, and $t_1 = -\tau \ln{(\alpha/2)}$.
 
-For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ -- or $[0.51, 30.76]$ in minutes. Thus, 90\% of the time between blocks blocks are between 30 seconds and 31 minutes.
+For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ --- or $[0.51, 30.76]$ in minutes. Thus, 90\% of the time between blocks blocks are between 30 seconds and 31 minutes.
 
@@
 Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$.
-As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{600}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{600})$.
+As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{\tau}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{\tau})$.
 Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.
@@
 The symmetrical confidence interval with $\alpha=10\%$ is $[27, 105]$ in minutes.
-Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted -- assuming that your transaction will be confirmed in the very next block.
+Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted --- assuming that your transaction will be confirmed in the very next block.
 
@@
 Let $T_\alpha = \frac{X}{\alpha H}$ be the time required for the miner to find a new block.
-As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{600}$.
+As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{\tau}$.
 Hence, we confirm the intuition that the miner with $\alpha$ percent of the network's total hash power will find $\alpha$ percent of the blocks.
@@
 		&= \frac{\lambda_\alpha}{\lambda_\alpha + \lambda_{1-\alpha}} \\
-		&= \frac{\alpha/600}{\alpha/600 + (1-\alpha)/600} \\
+		&= \frac{\alpha/\tau}{\alpha/\tau + (1-\alpha)/\tau} \\
 		&= \frac{\alpha}{\alpha + 1 - \alpha} \\
@@
 		&= \frac{\lambda_{m \alpha}}{\lambda_{m \alpha} + \lambda_{1-\alpha}} \\
-		&= \frac{m \alpha/600}{m \alpha/600 + (1-\alpha)/600} \\
+		&= \frac{m \alpha/\tau}{m \alpha/\tau + (1-\alpha)/\tau} \\
 		&= \frac{m \alpha}{m \alpha + 1-\alpha} \\
@@
 Let $\alpha$ be the percentage of the total hash rate of the node which is outdated, and $\Delta t$ the propagation time in seconds.
-Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{600}}$.
+Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{\tau}}$.
 For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.
@@
 
-%Let's assume that the network's total hash rate will change only once, from $H$ to $\alpha H$ -- so, if $\alpha = 1$, there would be no change, if $\alpha > 1$, there would be an increase, and if $\alpha < 1$, there would be a decrease.
+%Let's assume that the network's total hash rate will change only once, from $H$ to $\alpha H$ --- so, if $\alpha = 1$, there would be no change, if $\alpha > 1$, there would be an increase, and if $\alpha < 1$, there would be a decrease.
 As the time required to find a new block would be $T_\alpha = \frac{X}{\alpha H}$, we got the same random variable as we have gotten when analyzing a miner with $\alpha$ percent of the network's total hash rate.
@@
 Hence, $\mathbf{P}(T \leq t) = \mathbf{P}(X \leq H \int_0^t u(t)dt)$.
-When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{600} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{600} \cdot e^{-\frac{1}{600} \int_0^t u(t)dt}$.
+When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{\tau} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{\tau} \cdot e^{-\frac{1}{\tau} \int_0^t u(t)dt}$.
 
 Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
-So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{600}$.
-Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{600})$.
+So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\tau}$.
+Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\tau})$.
 Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$.
@@
 Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(t) = 1 + a^2t$.
-Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{t + a^2t^2/2}{600}}$.
+Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{t + a^2t^2/2}{\tau}}$.
 
@@
 	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
-		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{600}} \right) dx \\
-		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{600}} dx \\
-		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{600}} dx \\
+		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\tau}} \right) dx \\
+		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\tau}} dx \\
+		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx \\
 \end{align*}
 
-Let $b=\frac{a}{20\sqrt{3}}$ and $c = \frac{10\sqrt{3}}{a} \left( s + \frac{1}{600} \right)$, then $sx + \frac{x+a^2x^2/2}{600} = (bx + c)^2 - c^2$.
+Let $b=\frac{a}{20\sqrt{3}}$ and $c = \frac{10\sqrt{3}}{a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
 Let $y = bx+c \Rightarrow dy = bdx$.
@@
 \begin{align*}
-	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{600}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
+	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
 		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
@@
 \frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{20\sqrt{3\pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{10\sqrt{3\pi}s}{2ab} \\
-\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{1200a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right) \\
-\mathbf{E}[T] &= \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{1200a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right)
+\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right) \\
+\mathbf{E}[T] &= \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right)
 \end{align*}
 
-It is easy to notice that $lim_{a \rightarrow 0} \mathbf{E}[T] = 600$, and $lim_{a \rightarrow \infty} \mathbf{E}[T] = 0$.
+It is easy to notice that $lim_{a \rightarrow 0} \mathbf{E}[T] = \tau = 600$, and $lim_{a \rightarrow \infty} \mathbf{E}[T] = 0$.
 
@@
 Let $\beta H$ be the hash rate of the attackers, and $\gamma H$ be the network's hash rate without the attackers.
-Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{600}$ and $\lambda_{\text{network}} = \frac{\gamma}{600}$, respectively.
+Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{\tau}$ and $\lambda_{\text{network}} = \frac{\gamma}{\tau}$, respectively.
 
@@
 Each step right is a new block found by the attacker with probability $1-p$.
-It ends when the network finds $k$ blocks -- in this example, $k=6$.
+It ends when the network finds $k$ blocks --- in this example, $k=6$.
 The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$.
@@
 We may interpret it as a stable system.
-In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/600$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
+In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\tau$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
 Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
@@
 When $m \ll s$, $z_s \rightarrow 1 + 1/m$.
-So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = 600 \text{ seconds}$.
+So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \tau = 600 \text{ seconds}$.
 In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.
@@
 
-Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/600$ transactions per second, where $s$ is the average number of transactions per block.
+Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\tau = s/600$ transactions per second, where $s$ is the average number of transactions per block.
 
@@
 So, two new transactions may choose the same tips to confirm.
-If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions -- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
+If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
 

### 2018-01-21 Chapters/Chapter01.tex
Update on Overleaf.

@@
 Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted --- assuming that your transaction will be confirmed in the very next block.
+The pdf of $Y_6$ is shown in Figure \ref{fig-bitcoin-time-6-blocks}.
+The 10\% symmetrical confidence interval is shown in the white area.
+
+\begin{figure}[ht]
+\centering\includegraphics[height=5cm]{./images01/time-6-blocks.pdf}
+\caption{Probability density function of $Y_6$, i.e., probability of finding 6 blocks after time $t$. The shaded areas shows the lower 5\% and upper 5\% of the pdf.\label{fig-bitcoin-time-6-blocks}}
+\end{figure}
 

### 2018-01-22 Chapters/Chapter01.tex
Update on Overleaf.

@@
 The first has $n=160$ and consists of the composition of \textit{SHA-256} and \textit{RIPEMD-160}.
-The latter, has $n=256$ and applies \textit{SHA-256} twice.
+The latter has $n=256$ and applies \textit{SHA-256} twice.
 
@@
 
-\section{Bitcoin mining}
+\section{Mining one block}
 
 Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin \textit{HASH-256} function.
-The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given number.
-The lesser the $A$, the harder to find a new block.
+The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given threshold.
+The smaller the $A$, the harder to find a new block.
 In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.
@@
 The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\tau H} \right)^{tH}$.
-But, as the Bitcoin's network hash rate is really big, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\tau}$.
+But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\tau}$.
 
@@
 
-	As $\log(1 + \frac{1}{M}) > 0$, $1 < 1 + \log(1 + \frac{1}{M})$.
+	As $\log(1 + \frac{1}{M}) > 0$ and $1 < 1 + \log(1 + \frac{1}{M})$.
 
@@
 
-For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ --- or $[0.51, 30.76]$ in minutes. Thus, 90\% of the time between blocks blocks are between 30 seconds and 31 minutes.
+For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ (or $[0.51, 30.76]$ in minutes). Thus, 90\% of the time the intervals between blocks are between 30 seconds and 30 minutes.
+
+The fact that the time between blocks follows an exponential distribution with $\lambda = 1/\tau = pH$ may be used to estimate the total network's hash rate (or a miner's hash rate).
+For further information, see \cite{ozisikestimation}.
+
+\section{Mining several blocks}
 
@@
 \begin{figure}[ht]
-\centering\includegraphics[height=5cm]{./images01/time-6-blocks.pdf}
+\centering\includegraphics[width=\textwidth]{./images01/time-6-blocks.pdf}
 \caption{Probability density function of $Y_6$, i.e., probability of finding 6 blocks after time $t$. The shaded areas shows the lower 5\% and upper 5\% of the pdf.\label{fig-bitcoin-time-6-blocks}}
 \end{figure}
+
+\section{Mining for a miner}
 
@@
 \begin{cor}
-	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximatelly multiplies their probability of finding the next block by $m$.
+	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximately multiplies their probability of finding the next block by $m$.
 \end{cor}
@@
 
-%Then, $\mathbf{P}(t_0 < T < t_1) = \mathbf{P}(t_0 < X/H < t_1) = \mathbf{P}(t_0 H < X < t_1 H) = [1 - (1-p)^{t_1 H}] - [1 - (1-p)^{t_0 H}] = (1-p)^{t_0 H} - (1-p)^{t_1 H}$.
+\section{Orphan blocks}
 
@@
 On the other side, if the network's total hash rate suddenly halves ($\alpha = 0.5$), then $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled.
-It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, and then the adjustment will occur and everything will be back to the normal 10 minutes between blocks.
+It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, at which point the adjustment will occur and everything will be back to the normal 10 minutes between blocks.
 
@@
 \begin{theorem}
-	$$a > 1, x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
+	$$a > 1 \text{ and } x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
 \end{theorem}
@@
 In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes.
-If the hash rate had suddenly doubled, the average time between blocks would be 300.
+If the hash rate had suddenly doubled, the average time between blocks would be 5 minutes.
 
@@
 
-\subsection{Hash rate changing linearly}
+\subsection{Piecewise linear model of hash rate change}
 
@@
 Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{t + a^2t^2/2}{\tau}}$.
+
+It is well known that $\mathbf{E}(T) = \int_{0}^{\infty} 1 - \mathbf{P}(T \leq t) dt$.
+Thus, replacing $y = \frac{a^2 t + 1}{a \sqrt{2 \tau}}$, and using the fact that $\int_0^\infty e^{-x^2} dx = \sqrt{pi}/2 \erf(x)$, we have:
+
+\begin{align}
+\mathbf{E}(T)|_{t_1}^{t_2} &= \int_{t_1}^{t_2} \exp \left( - \frac{t + a^2 t^2/2}{\tau} \right) dt \nonumber \\
+	&= \frac{\sqrt{2 \tau}}{a} \exp \left( \frac{1}{2a^2 \tau} \right) \int_{y_1}^{y_2}  \exp(-y^2) dy \nonumber \\
+	&= \frac{\sqrt{2 \tau}}{a} \exp \left( \frac{1}{2a^2 \tau} \right) \frac{\sqrt{\pi}}{2} [\erf(y_1) - \erf(y_2)] \nonumber \\
+	&= \frac{\sqrt{2 \pi \tau}}{2a} \exp \left( \frac{1}{2a^2 \tau} \right) \left[\erf(y_2) - \erf(y_1) \right]
+\end{align}
+
+Where $y_1 = \frac{a^2 t_1 + 1}{a \sqrt{2 \tau}}$ and $y_2 = \frac{a^2 t_2 + 1}{a \sqrt{2 \tau}}$.
+
+Thus, $\mathbf{E}(T) = \mathbf{E}(T)|_0^\infty$.
+When $t_1 = 0 \Rightarrow y_1 = \frac{1}{2 \sqrt{2 \tau}}$ and $t_2 \rightarrow \infty \Rightarrow y_2 \rightarrow \infty \Rightarrow \erf(y_2) = 1$, then:
+
+$$
+\mathbf{E}(T) = \frac{\sqrt{2 \pi \tau}}{2a} \exp \left( \frac{1}{2a^2 \tau} \right) \left[1 - \erf\left(\frac{1}{a \sqrt{2 \tau}}\right)\right]
+$$
+
+The function $\mathbf{E}(T)|_{t_1}^{t_2}$ may be used to a piecewise linear analysis of any hash rate change.
+Let's analyze the hash doubling in one week.
+Then, $u(1 \text{week}) = u(604,800) = 2$, thus $a^2 = \frac{1}{604,800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$.
+Let's partition the interval $[0, 1 \text{week}]$ in hourly intervals, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.
+For each interval $[t_i, t_{i+1}]$, we evaluate $\mathbf{E}(T)|_{t_i}^{t_{i+1}}$.
+The result is presented in Figure .
+
+---
 
@@
 
-Let $b=\frac{a}{20\sqrt{3}}$ and $c = \frac{10\sqrt{3}}{a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
+Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
 Let $y = bx+c \Rightarrow dy = bdx$.
@@
 
-As $\frac{d}{ds}c = \frac{10\sqrt{3}}{a}$,
+As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,
 
 \begin{align*}
-\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{20\sqrt{3\pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{10\sqrt{3\pi}s}{2ab} \\
-\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right) \\
-\mathbf{E}[T] &= \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right)
+\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
+\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
+\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
 \end{align*}
-
-It is easy to notice that $lim_{a \rightarrow 0} \mathbf{E}[T] = \tau = 600$, and $lim_{a \rightarrow \infty} \mathbf{E}[T] = 0$.
 
@@
 
-Attackers are not part of the mining network, because they are trying to generate a different longest chain.
+There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the submarine attack.
+
+In the submarine attack, the attacker's goal is to spend some Bitcoins and, after a while, get their Bitcoins back.
+Thus, the attacker sends some funds to the victim, they wait 6 confirmations, and, finally, the victim confirms the funds.
+But, then, the attacker get their funds back propagating new blocks which accept a double spending transaction.
+In order to work, the attacker must propagate more blocks than the network has mined in the same period, becoming the longest chain.
+Hence, we would like to understand what the odds are that the attacker will be successful.
+
+In order to maximize their odds, the attacker must start to mine the new blocks as soon as they send the funds to the victim.
+In this moment, it starts to mine in the head of the blockchain, just like the rest of the network.
+So, the attacker and the network are in exactly the same point.
 
@@
 
-As \cite{nakamoto2008bitcoin} have done, we will also model the attack using the Gambler's Ruin.
+As \cite{nakamoto2008bitcoin} has done, we will also model the attack using the Gambler's Ruin.
 In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$.
@@
 
-The gambler wins \$1 is the same as the network find a new block, the gambler losses \$1 is the same as the attacker find a new block.
+The gambler winning \$1 is the same as the network finding a new block, the gambler losing \$1 is the same as the attacker finding a new block.
 The initial \$$k$ is the same as the number of blocks the attacker is behind the network.
-Thus, the gambler losses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.
+Thus, the gambler loses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.
 
@@
 
-But this scenario seems not to be the optimal attack, because the attacker has waited the $k$ confirmations before starting the attack.
+But this scenario seems not to be the optimal attack, because the attacker has waited $k$ confirmations before starting the attack.
 A better approach would be to start attacking just after propagating the transaction.
@@
 Then, if the attacker has more blocks than the network, he has successfully attacked.
-Otherwise, we got back to the previous model, in which the attacker must still find more blocks.
+Otherwise, we return to the previous model, in which the attacker must still find more blocks.
 
 \begin{theorem}
-	Let $a_{ij}$ be the probability of moving from state $(i, j)$ to state $(s, k)$ in the Markov chain.
-	$$a_{ij} = \binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}.$$
+	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
 \end{theorem}
 \begin{proof}
+The attacker must find exactly $s$ blocks while the network must find exactly $k$ blocks.
+It is as they would be walking the grid from the point $(0, 0)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}.
+When the attacker finds a block, it would be a movement to the right.
+When the network finds a block, it would be an upward movement.
+No matter the order which the blocks are found, all the paths occur with probability $(1-p)^s p^k$.
 
-In order to go from $(i, j)$ to $(s, k)$, the attacker must find exacly $i-s$ blocks, and the network must find exactly $k-j$ blocks.
-It is as they would be walking the grid from the point $(i, j)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}.
-There are many possible paths and all of them occurs with probability $(1-p)^{s-i} p^{k-j}$.
+The walking ends when $(\cdot, k$) is reached, i.e., when the network finds $k$ blocks, regardless of how many blocks the attacker has found -- i.e., it is not allowed to walk above the line $(\cdot, k)$.
+Thus, the number of paths between $(0, 0)$ and $(s, k)$ moving only upward or to the right, without going into the line $(\cdot, k)$ is exactly the number of paths between $(0, 0)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s$ movements to the right ($r$) and $k-1$ upward movements ($u$).
+This number of permutations is $\frac{(k-1+s)!}{s!(k-1)!} = \binom{(k-1)+(s)}{s}$ because there are $s$ repetitions of the element $r$ and $k-1$ repetitions of the element $u$.
 
-The walking ends when $(\cdot, k$) is reach, i.e., when the network finds $k$ blocks, no matter how many blocks the attacker has found.
-So, it is not allowed to walk through the line $(\cdot, k)$.
-Thus, the number of paths between $(i, j)$ and $(s, k)$ going only up or right, without going through the line $(\cdot, k)$ is exactly the number of paths between $(i, j)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s-i$ right movements ($r$) and $k-j-1$ up movements ($u$).
-The number of permutations is $\frac{(k-j-1+s-i)!}{(s-i)!(k-j-1)!} = \binom{(k-j-1)+(s-i)}{s-i}$ because there are $s-i$ repetitions of the element $r$ and $k-j-1$ repetitions of the element $u$.
-
-Finally, the probability is $\binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}$.
+Finally, the probability is $\binom{k+s-1}{s} (1-p)^{s} p^{k}$.
 
@@
 Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network.
-In the red path, the attacker still have to catch up 3 blocks to have a successfull attack, which happens with probability $\rho^3$, if $p < 0.5$.}
+In the red path, the attacker still have to catch up 3 blocks to have a successful attack, which happens with probability $\rho^3$, if $p < 0.5$.}
   \label{figure:bitcoin-attack-paths}
@@
 
-\begin{cor}
-	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
-\end{cor}
-\begin{proof}
-	This probability $\mathbf{P}(S = s) = a_{00}$.
-\end{proof}
-
-Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.
+Assuming that the attacker starts mining just after publishing the victim's transaction, the probability of the attacker will have found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.
 
@@
 
-Let's assume that all transactions have exactly the same size $S$ and pays exactly the same fee.
-If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transaction in each block.
-Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single zero of $z^s(1+m(1-z))-1$ with $|z_s|>1$.
+Let's assume that all transactions have exactly the same size $S$ and pay exactly the same fee.
+If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block.
+Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$.
 The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$.
@@
 The probabilities form a simple geometric series, and so they are exponentially decreasing.
-We may interpret it as a stable system.
+We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite.
 In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\tau$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.

### 2018-01-22 Chapters/Chapter01.tex
Update on Overleaf.

@@
 The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block.
-So, the rule of finding a new block every 10 minutes ($\tau = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \tau = 600$.
-So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \tau = 600 \Rightarrow p = \frac{1}{\tau H}$.
-Finally, $\mathbf{E}(X) = \tau H$, $\mathbf{E}(T) = \tau$, $\mathbf{V}(X) = (\tau H)^2 - \tau H$, and $\mathbf{V}(T) = \tau^2 - \tau/H$.
-
-The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\tau H} \right)^{tH}$.
-But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\tau}$.
+So, the rule of finding a new block every 10 minutes ($\eta = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \eta = 600$.
+So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \eta = 600 \Rightarrow p = \frac{1}{\eta H}$.
+Finally, $\mathbf{E}(X) = \eta H$, $\mathbf{E}(T) = \eta$, $\mathbf{V}(X) = (\eta H)^2 - \eta H$, and $\mathbf{V}(T) = \eta^2 - \eta/H$.
+
+The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\eta H} \right)^{tH}$.
+But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.
 
 \begin{theorem}
-	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{\tau}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$.
+	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{\eta}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$.
 \end{theorem}
@@
 	\mathbf{P}(T \leq t) &= 1 - (1-p)^{tH} \\
-		     &= 1 - \left( 1 - \frac{1}{\tau H} \right)^{tH} \\
-	\text{Replacing $u = \tau H$,} \\
-	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{\tau}} \\
-				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{\tau}} \\
-				&= 1 - \left( 1/e \right)^{\frac{t}{\tau}} \\
-				&= 1 - e^{-\frac{t}{\tau}}
+		     &= 1 - \left( 1 - \frac{1}{\eta H} \right)^{tH} \\
+	\text{Replacing $u = \eta H$,} \\
+	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{\eta}} \\
+				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{\eta}} \\
+				&= 1 - \left( 1/e \right)^{\frac{t}{\eta}} \\
+				&= 1 - e^{-\frac{t}{\eta}}
 \end{align*}
@@
 The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, where $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$.
-These conditions give the following equations: $1 - e^{-t_0/\tau} = \alpha/2$, and $e^{-t_1/\tau} = \alpha/2$.
-Solving these equations, we have $t_0 = -\tau \ln{(1 - \alpha/2)}$, and $t_1 = -\tau \ln{(\alpha/2)}$.
+These conditions give the following equations: $1 - e^{-t_0/\eta} = \alpha/2$, and $e^{-t_1/\eta} = \alpha/2$.
+Solving these equations, we have $t_0 = -\eta \ln{(1 - \alpha/2)}$, and $t_1 = -\eta \ln{(\alpha/2)}$.
 
@@
 
-The fact that the time between blocks follows an exponential distribution with $\lambda = 1/\tau = pH$ may be used to estimate the total network's hash rate (or a miner's hash rate).
+The fact that the time between blocks follows an exponential distribution with $\lambda = 1/\eta = pH$ may be used to estimate the total network's hash rate (or a miner's hash rate).
 For further information, see \cite{ozisikestimation}.
@@
 Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$.
-As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{\tau}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{\tau})$.
+As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{\eta}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{\eta})$.
 Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.
@@
 Let $T_\alpha = \frac{X}{\alpha H}$ be the time required for the miner to find a new block.
-As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{\tau}$.
+As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{\eta}$.
 Hence, we confirm the intuition that the miner with $\alpha$ percent of the network's total hash power will find $\alpha$ percent of the blocks.
@@
 		&= \frac{\lambda_\alpha}{\lambda_\alpha + \lambda_{1-\alpha}} \\
-		&= \frac{\alpha/\tau}{\alpha/\tau + (1-\alpha)/\tau} \\
+		&= \frac{\alpha/\eta}{\alpha/\eta + (1-\alpha)/\eta} \\
 		&= \frac{\alpha}{\alpha + 1 - \alpha} \\
@@
 		&= \frac{\lambda_{m \alpha}}{\lambda_{m \alpha} + \lambda_{1-\alpha}} \\
-		&= \frac{m \alpha/\tau}{m \alpha/\tau + (1-\alpha)/\tau} \\
+		&= \frac{m \alpha/\eta}{m \alpha/\eta + (1-\alpha)/\eta} \\
 		&= \frac{m \alpha}{m \alpha + 1-\alpha} \\
@@
 Let $\alpha$ be the percentage of the total hash rate of the node which is outdated, and $\Delta t$ the propagation time in seconds.
-Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{\tau}}$.
+Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{\eta}}$.
 For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.
@@
 Hence, $\mathbf{P}(T \leq t) = \mathbf{P}(X \leq H \int_0^t u(t)dt)$.
-When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{\tau} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{\tau} \cdot e^{-\frac{1}{\tau} \int_0^t u(t)dt}$.
+When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{\eta} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{\eta} \cdot e^{-\frac{1}{\eta} \int_0^t u(t)dt}$.
 
 Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
-So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\tau}$.
-Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\tau})$.
+So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\eta}$.
+Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\eta})$.
 Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$.
@@
 
-Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(t) = 1 + a^2t$.
-Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{t + a^2t^2/2}{\tau}}$.
+Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(a, b, t) = a^2t + b$.
+Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{bt + a^2t^2/2}{\eta}}$.
 
 It is well known that $\mathbf{E}(T) = \int_{0}^{\infty} 1 - \mathbf{P}(T \leq t) dt$.
-Thus, replacing $y = \frac{a^2 t + 1}{a \sqrt{2 \tau}}$, and using the fact that $\int_0^\infty e^{-x^2} dx = \sqrt{pi}/2 \erf(x)$, we have:
+Thus, replacing $y = \frac{a^2 t + b}{a \sqrt{2 \eta}}$, and using the fact that $\int_0^\infty e^{-x^2} dx = \frac{\sqrt{\pi}}{2} \erf(x)$, we have:
 
 \begin{align}
-\mathbf{E}(T)|_{t_1}^{t_2} &= \int_{t_1}^{t_2} \exp \left( - \frac{t + a^2 t^2/2}{\tau} \right) dt \nonumber \\
-	&= \frac{\sqrt{2 \tau}}{a} \exp \left( \frac{1}{2a^2 \tau} \right) \int_{y_1}^{y_2}  \exp(-y^2) dy \nonumber \\
-	&= \frac{\sqrt{2 \tau}}{a} \exp \left( \frac{1}{2a^2 \tau} \right) \frac{\sqrt{\pi}}{2} [\erf(y_1) - \erf(y_2)] \nonumber \\
-	&= \frac{\sqrt{2 \pi \tau}}{2a} \exp \left( \frac{1}{2a^2 \tau} \right) \left[\erf(y_2) - \erf(y_1) \right]
+\mathbf{E}(T)|_{t_1}^{t_2} &= \int_{t_1}^{t_2} \exp \left( - \frac{bt + a^2 t^2/2}{\eta} \right) dt \nonumber \\
+	&= \frac{\sqrt{2 \eta}}{a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \int_{y_1}^{y_2}  \exp(-y^2) dy \nonumber \\
+	&= \frac{\sqrt{2 \eta}}{a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \frac{\sqrt{\pi}}{2} [\erf(y_1) - \erf(y_2)] \nonumber \\
+	&= \frac{\sqrt{2 \pi \eta}}{2a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \left[\erf(y_2) - \erf(y_1) \right]
 \end{align}
 
-Where $y_1 = \frac{a^2 t_1 + 1}{a \sqrt{2 \tau}}$ and $y_2 = \frac{a^2 t_2 + 1}{a \sqrt{2 \tau}}$.
+Where $y_1 = \frac{a^2 t_1 + b}{a \sqrt{2 \eta}}$ and $y_2 = \frac{a^2 t_2 + b}{a \sqrt{2 \eta}}$.
 
 Thus, $\mathbf{E}(T) = \mathbf{E}(T)|_0^\infty$.
-When $t_1 = 0 \Rightarrow y_1 = \frac{1}{2 \sqrt{2 \tau}}$ and $t_2 \rightarrow \infty \Rightarrow y_2 \rightarrow \infty \Rightarrow \erf(y_2) = 1$, then:
+When $t_1 = 0 \Rightarrow y_1 = \frac{b^2}{2 \sqrt{2 \eta}}$ and $t_2 \rightarrow \infty \Rightarrow y_2 \rightarrow \infty \Rightarrow \erf(y_2) = 1$, then:
 
 $$
-\mathbf{E}(T) = \frac{\sqrt{2 \pi \tau}}{2a} \exp \left( \frac{1}{2a^2 \tau} \right) \left[1 - \erf\left(\frac{1}{a \sqrt{2 \tau}}\right)\right]
+\mathbf{E}(T) = \frac{\sqrt{2 \pi \eta}}{2a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \left[1 - \erf\left(\frac{1}{a \sqrt{2 \eta}}\right)\right]
 $$
@@
 Let's analyze the hash doubling in one week.
-Then, $u(1 \text{week}) = u(604,800) = 2$, thus $a^2 = \frac{1}{604,800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$.
-Let's partition the interval $[0, 1 \text{week}]$ in hourly intervals, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.
-For each interval $[t_i, t_{i+1}]$, we evaluate $\mathbf{E}(T)|_{t_i}^{t_{i+1}}$.
-The result is presented in Figure .
+Then, $u(1 \text{week}) = u(604800) = 2$, thus $a^2 = \frac{1}{604800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$.
+Let's sample the interval $[0, 1 \text{week}]$ every hour, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.
+
+For each point $t_i$, let $g_i(t) = [H(t-t_i) - H(t-604800)] u(t) + 2 H(t-604800)$, then $\mathbf{E}(T) = \mathbf{E}(T)|_{t_i}^{t_{i+1}} + 1 - e^{\frac{t}{\eta}}$. The result is presented in Figure .
 
@@
 	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
-		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\tau}} \right) dx \\
-		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\tau}} dx \\
-		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx \\
+		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\eta}} \right) dx \\
+		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\eta}} dx \\
+		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\eta}} dx \\
 \end{align*}
 
-Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
+Let $b=\frac{a}{\sqrt(2 \eta)}$ and $c = \frac{\sqrt(2 \eta)}{2a} \left( s + \frac{1}{\eta} \right)$, then $sx + \frac{x+a^2x^2/2}{\eta} = (bx + c)^2 - c^2$.
 Let $y = bx+c \Rightarrow dy = bdx$.
@@
 \begin{align*}
-	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
+	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\eta}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
 		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
@@
 
-As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,
+As $\frac{d}{ds}c = \frac{\sqrt{2 \eta}}{2a}$,
 
 \begin{align*}
-\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
-\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
-\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
+\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \eta \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \eta \pi}s}{4ab} \\
+\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \eta \pi}}{2a} e^{\frac{1}{2 \eta a^2}} \erfc\left(\frac{1}{a \sqrt{2 \eta}}\right) \\
+\mathbf{E}[T] &= \frac{\sqrt{2 \eta \pi}}{2a} e^{\frac{1}{2 \eta a^2}} \erfc\left(\frac{1}{a \sqrt{2 \eta}}\right) \\
 \end{align*}
@@
 Let $\beta H$ be the hash rate of the attackers, and $\gamma H$ be the network's hash rate without the attackers.
-Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{\tau}$ and $\lambda_{\text{network}} = \frac{\gamma}{\tau}$, respectively.
+Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{\eta}$ and $\lambda_{\text{network}} = \frac{\gamma}{\eta}$, respectively.
 
@@
 We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite.
-In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\tau$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
+In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\eta$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
 Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
@@
 When $m \ll s$, $z_s \rightarrow 1 + 1/m$.
-So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \tau = 600 \text{ seconds}$.
+So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$.
 In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.
@@
 
-Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\tau = s/600$ transactions per second, where $s$ is the average number of transactions per block.
+Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
 

### 2018-01-22 Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7.3.tex
Add Kanerva-Table-7.3

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{Kanerva-Table-7.3}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}2}]:} \PY{k+kn}{import} \PY{n+nn}{sdm} \PY{k+kn}{as} \PY{n+nn}{sdmlib}
        \PY{k+kn}{import} \PY{n+nn}{matplotlib.pyplot} \PY{k+kn}{as} \PY{n+nn}{plt}
        \PY{o}{\PYZpc{}}\PY{k}{matplotlib} inline
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}3}]:} \PY{k}{def} \PY{n+nf}{gen\PYZus{}sdm}\PY{p}{(}\PY{n}{scanner\PYZus{}type}\PY{p}{,} \PY{n}{bits}\PY{o}{=}\PY{l+m+mi}{1000}\PY{p}{,} \PY{n}{sample}\PY{o}{=}\PY{l+m+mi}{1000000}\PY{p}{)}\PY{p}{:}
            \PY{n}{address\PYZus{}space} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{AddressSpace}\PY{o}{.}\PY{n}{init\PYZus{}from\PYZus{}b64\PYZus{}file}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{sdm\PYZhy{}10000w.as}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
            \PY{n}{counter} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{Counter}\PY{o}{.}\PY{n}{load\PYZus{}file}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{sdm\PYZhy{}10000w}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
            \PY{n}{sdm} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{SDM}\PY{p}{(}\PY{n}{address\PYZus{}space}\PY{p}{,} \PY{n}{counter}\PY{p}{,} \PY{l+m+mi}{451}\PY{p}{,} \PY{n}{scanner\PYZus{}type}\PY{p}{)}
            \PY{k}{return} \PY{n}{sdm}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}4}]:} \PY{n}{sdm} \PY{o}{=} \PY{n}{gen\PYZus{}sdm}\PY{p}{(}\PY{n}{sdmlib}\PY{o}{.}\PY{n}{SDM\PYZus{}SCANNER\PYZus{}OPENCL}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}5}]:} \PY{n}{b} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{Bitstring}\PY{o}{.}\PY{n}{init\PYZus{}random}\PY{p}{(}\PY{l+m+mi}{1000}\PY{p}{)}
        \PY{n}{sdm}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{b}\PY{p}{,} \PY{n}{b}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}6}]:} \PY{k+kn}{from} \PY{n+nn}{IPython.display} \PY{k+kn}{import} \PY{n}{clear\PYZus{}output}

        \PY{n}{distances} \PY{o}{=} \PY{p}{[}\PY{p}{]}
        \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1001}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{)}
        \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{dist} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{:}
            \PY{n}{clear\PYZus{}output}\PY{p}{(}\PY{n}{wait}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
            \PY{k}{print} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Distance: \PYZob{}:4d\PYZcb{} (\PYZob{}:.2f\PYZcb{}}\PY{l+s+s1}{\PYZpc{}}\PY{l+s+s1}{)}\PY{l+s+s1}{\PYZsq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{dist}\PY{p}{,} \PY{l+m+mf}{100.}\PY{o}{*}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{o}{/}\PY{n+nb}{len}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{)}
            \PY{n}{c} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{Bitstring}\PY{o}{.}\PY{n}{init\PYZus{}from\PYZus{}bitstring}\PY{p}{(}\PY{n}{b}\PY{p}{)}
            \PY{n}{c}\PY{o}{.}\PY{n}{flip\PYZus{}random\PYZus{}bits}\PY{p}{(}\PY{n}{dist}\PY{p}{)}
            \PY{k}{assert} \PY{n}{c}\PY{o}{.}\PY{n}{distance\PYZus{}to}\PY{p}{(}\PY{n}{b}\PY{p}{)} \PY{o}{==} \PY{n}{dist}
            \PY{n}{d} \PY{o}{=} \PY{n}{sdm}\PY{o}{.}\PY{n}{read}\PY{p}{(}\PY{n}{c}\PY{p}{)}
            \PY{n}{distances}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{d}\PY{o}{.}\PY{n}{distance\PYZus{}to}\PY{p}{(}\PY{n}{b}\PY{p}{)}\PY{p}{)}
        \PY{k}{print} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Done!}\PY{l+s+s1}{\PYZsq{}}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Distance: 1000 (100.00\%)
Done!

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}29}]:} \PY{n}{plt}\PY{o}{.}\PY{n}{figure}\PY{p}{(}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{distances}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{x}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{p}{[}\PY{l+m+mi}{500}\PY{p}{]}\PY{o}{*}\PY{n+nb}{len}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Kanerva}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s+s1}{s Figure 7.3}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{New distance (after one read)}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Old distance}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{axis}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1000}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1000}\PY{p}{]}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_5_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    % Add a bibliography block to the postdoc

    \end{document}

### 2018-01-25 Chapters/Introduction.tex
Update on Overleaf.

@@
 Today the field has advanced to a great number of nations, and the amount of applications has grown explosively.
-Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power at the hands of individuals.
+Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power to the hands of individuals.
 Whereas, during the war, computations were mostly done by hand, it is not outlandish to say that this original field can be referred to, today, as computational management science.
@@
 This thesis will present a number of different topics for exploration.
-Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three self-contained scientific papers.
-The first of which will study the possibility of distributed financial ledgers.
+Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three specific, self-contained, scientific papers.
+The first of which studies the possibility of distributed financial ledgers.
 
@@
 The World Bank estimates that there are two billion people without access to financial services.
-As banks are unable to sustain operations in numerous poverty-stricken areas, services such as money transfers, access to credit, digital/distant payments, inflation protection, etc., remain beyond reach for large numbers of people.
-This seems to be one of the factors that perpetuate poverty, and the Bill and Melinda Gates Foundation chose as focus of its ``Level-One project'': to provide basic financial services through cell phones.
+As banks are unable to sustain operations in numerous poverty-stricken areas, services such as money transfers, access to credit, digital/distant payments, inflation protection, etc., remain beyond reach for `the unbanked'.
+This seems to be one of the factors that perpetuate poverty.
+The Bill and Melinda Gates Foundation chose as focus of its ``Level-One project'': to provide basic financial services through cell phones.
 Another initiative, the United Nations World Food Programme has began, in 2017, an experiment in Jordan, in which the organization provides funds for thousands of people towards its goal of food relief.
@@
 
-Bitcoin is the first and most famous cryptocurrency, used worldwide and has a highly volatile market cap, as of this writing, of \$ 192 bi.
+Bitcoin is the first and most famous cryptocurrency, used worldwide, with a highly volatile market cap, as of this writing, of \$ 192 bi.
 Even so, it faces serious scalability challenges; such as serious quality of service and network congestion when the number of transactions per second is high, and an increase in the transaction fees and uncertain delays in transactions' confirmations.
@@
 
-\emph{Iota} is a second cryptocurrency, which, instead of using a blockchain, proposes the use of a ``tangle' architecture': a different way to register the currency ledger across thousands of computers.
+\emph{Iota} is a second cryptocurrency that, instead of using a blockchain, proposes the use of a ``tangle' architecture': a different way to register the currency ledger across thousands of computers.
 Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
@@
 Trains allowed distant places to exchange high volume of goods.
-Airplanes and boats allowed countries to do business overseas.
-And, finally, internet has changed everything -- the way we communicate, behave, do business, do shopping, share ideas, and so forth.
+Airplanes and boats opened countries to overseas business.
+And, finally, the internet has had a profound impact in nearly everyone's life, as it changed everything -- from the way we communicate, behave, do business, do shopping, share ideas, and so forth.
 
-One are of technology that has been redefining business is computer science.
+One area of technology that has been redefining business is computer science.
 Together with the internet, computer science has been one of the most important tools to scale a business model -- and create many others which were impossible before.
-More and more expensive human labor has been replace by algorithms.
+More and more expensive human labor has been replaced by algorithms.
 Managers are able to make better decisions because they receive real time information.
@@
 Netflix increases the quality of their services offering movies specific to the taste of each customer.
-Spotify learns which kind of music you like most and suggests playlists and radios.
+Spotify learns which kind of music users like the most and suggests playlists and radios.
 Banks prevent fraud classifying which patterns are strange to their customers.
-It is not possible to imagine a world with no artificial intelligence.
-Behind many artificial intelligence systems, there is the pattern recognition.
-The capacity to match information from new data with information which has already been seen and is stored in memory.
-It may be used to classification, face recognition, character recognition, and so forth.
+
+It is gradually becoming impossible to imagine a world without artificial intelligence.
+
+Behind many artificial intelligence systems, there is pattern recognition: The capacity to match information from new data with information which has already been seen and is stored in memory.
+It may be used in classification, face recognition, character recognition, and so forth.
 
@@
 Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics of a human memory.
-SDM has already been used to different patter recognition problems, like noise reduction, handwriting recognition, robot automation, and so forth.
+SDM has already been used to different pattern recognition problems, like noise reduction, handwriting recognition, robot automation, and so forth.
 
-We implement a RB-Complete\footnote{Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
+We implement a RB-Complete\footnote{`Ridiculously Buzzword Complete': the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
 The computer code has been used in a previous Ph.D.

### 2018-01-25 Chapters/Introduction.tex
Update on Overleaf.

@@
 
-We implement a RB-Complete\footnote{`Ridiculously Buzzword Complete': the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
+We implement a RB-Complete\footnote{`Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
 The computer code has been used in a previous Ph.D.
@@
 Facebook replied applying exactly the same methodology as Princeton's researchers.
-In their own words: ``Using the same robust methodology featured in the paper, we attempted to find out more about this `Princeton University' -- and you won't believe what we found!''.
+In their own words: ``Using the same robust methodology featured in [Princeton's] paper, we attempted to find out more about this `Princeton University' -- and you won't believe what we found!''.
 Then, they conclude: ``This trend suggests that Princeton will have only half its current enrollment by 2018, and by 2021 it will have no students at all, agreeing with the previous graph of scholarly scholarliness.
@@
 A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---this body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
-Please bear with me as we venture into the world of computational management science.
+
+With these qualifications, we are ready to venture into the world of computational management science.

### 2018-02-08 Chapters/Conclusion.tex
Update on Overleaf.

--- 2018-01-18 Chapters/Introduction.tex
@@
-\chapter{Introduction}
+\chapter{Conclusion}
 \bigskip
@@
 \noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
-
-Though most fields existed prior to the war, after the war they were funded by the public as strategic pieces of the major nations arsenal against future conflagrations.
-One of the fields in question was that of Management Science (also called Operations Research in military circles, as researchers filled the ranks of planners of war operations).
-Management Science had started as an industrial field, in movements stemming from Taylor and the origin of the production line by Henry Ford.
-That was the first moment in industry in which operations were systematically subject to some of the tools of science: measurement, experimentation, hypothesis-testing, statistics, mathematical optimization, etc.
-
-This is from almost 100 years ago.
-Today the field has advanced to a great number of nations, and the amount of applications has grown explosively.
-Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power at the hands of individuals.
-Whereas, during the war, computations were mostly done by hand, it is not outlandish to say that this original field can be referred to, today, as computational management science.
-
-Applied mathematics and computer science serve simultaneously as a theoretical foundation and the major tool available to the field.
-Though this is a doctoral thesis concerning business, in this document one should expect to find the language and nomenclature of mathematical modeling and computer science as our primary and most natural language.
-
-This thesis will present a number of different topics for exploration.
-Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three self-contained scientific papers.
-The first of which will study the possibility of distributed financial ledgers.
-
-\section{Distributed financial ledgers}
-
-United Nations World Food Programme, `Blockchain Against Hunger: Harnessing Technology In Support Of Syrian Refugees', 30 May 2017, https://www.wfp.org/news/news-release/blockchain-against-hunger-harnessing-technology-support-syrian-refugees, accessed in January 17, 2018.
-
-United Nations World Food Programme, 2016, The Year in Review, Report.
-
-Woyke, E., 2017, `How blockchain can bring financial services to the poor', MIT Technology Review, April 18.
-
-The World Bank estimates that there are two billion people without access to financial services.
-As banks are unable to sustain operations in numerous poverty-stricken areas, services such as money transfers, access to credit, digital/distant payments, inflation protection, etc., remain beyond reach for large numbers of people.
-This seems to be one of the factors that perpetuate poverty, and the Bill and Melinda Gates Foundation chose as focus of its ``Level-One project'': to provide basic financial services through cell phones.
-Another initiative, the United Nations World Food Programme has began, in 2017, an experiment in Jordan, in which the organization provides funds for thousands of people towards its goal of food relief.
-An interesting aspect of this program has been the format of the funds distributed:  they have been all on the ethereum blockchain.
-
-The possibility of having a completely digital financial system without the overheads of traditional banking systems has appeared with the release of Bitcoin and similar blockchain technologies.
-This field questions numerous traditional assumptions in computer science, record-keeping, banking \& finance, and economic inclusion.
-The seminal work of \citet{nakamoto2008bitcoin} described the architecture of Bitcoin, a peer-to-peer electronic cash system, also known as cryptocurrency.
-Bitcoin's currency ledger is public and stored in a blockchain across thousands of computers.
-Even so, no one is able to spend either somebody else's funds nor to double spend their own funds.
-In order to be confirmed, each transaction must be both digitally signed by the owner of the money and the funds verified in the blockchain by Bitcoin's miners.
-
-The question of whether Bitcoin (or related works) can scale to billions of people is, however, far from settled.
-
-One of the interesting parts of Bitcoin are the incentives.
-On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick, and the currency issuance rate is well known.
-On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.
-
-The impact of Bitcoin in society -- and hence in the companies and the government -- has been increasing every day.
-People are increasingly using Bitcoin to exchange money and transfer money overseas.
-Companies are looking into Bitcoin as an alternative to reduce banking fees.
-The poor may be included in the finance system through Bitcoin.
-People may hedge their assets against their governments' money issuance and inflation---as in the case of Venezuela.
-
-Bitcoin is the first and most famous cryptocurrency, used worldwide and has a highly volatile market cap, as of this writing, of \$ 192 bi.
-Even so, it faces serious scalability challenges; such as serious quality of service and network congestion when the number of transactions per second is high, and an increase in the transaction fees and uncertain delays in transactions' confirmations.
-
-Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'' developers, which believe that is it risky to increase the the blocksize (in which all transactions are stored).
-It is not known whether a blocksize, say, of 1Gb, would be feasible to sustain the decentralization of the network.
-
-\emph{Iota} is a second cryptocurrency, which, instead of using a blockchain, proposes the use of a ``tangle' architecture': a different way to register the currency ledger across thousands of computers.
-Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
-As we will see, the problem here is exactly the opposite of Bitcoin's.
-Iota needs a minimum of transactions per seconds in order to work properly.
-
-Our analysis suggests an architecture for a distributed currency which is inspired in both Bitcoin's blockchain and Iota's tangle in order to solve the scalability problems.
-While Bitcoin's network saturates when it hits a certain number of transactions per second, Iota's does not work properly with less than a certain number of transactions per second.
-Our proposed architecture seems to work in both scenarios: low and high number of transactions per second.
-
-In this first study we will investigate some issues regarding this possibility, namely: (i) cryptographic security and game-theoretical attacks; (ii) scalability; (iii) self-governance of the system; (iv) appropriate incentive system to all participants.
-
-A second topic that may have an outsized influence on business and that we will be taking a closer look is a model of artificial intelligence.
-
-\section{Artificial intelligence}
-% Technology > Computer Science > Artificial Intelligence > Pattern Recognition > SDM.
-
-Technology has been one of the underlying engines behind economic growth.
-It has been changing the whole society -- people, companies, and governments.
-Cities and houses had to be rethought when cars became popular.
-Trains allowed distant places to exchange high volume of goods.
-Airplanes and boats allowed countries to do business overseas.
-And, finally, internet has changed everything -- the way we communicate, behave, do business, do shopping, share ideas, and so forth.
-
-One are of technology that has been redefining business is computer science.
-Together with the internet, computer science has been one of the most important tools to scale a business model -- and create many others which were impossible before.
-More and more expensive human labor has been replace by algorithms.
-Managers are able to make better decisions because they receive real time information.
-The supply chain has incredibly evolved thanks to advances in logistics supported by routing algorithms, storage algorithms, and many others optimization algorithms.
-
-Artificial intelligence has been disrupting many businesses.
-Uber is able to handle hundred of thousands of requests.
-Amazon optimizes the location of each product based on demand.
-Netflix increases the quality of their services offering movies specific to the taste of each customer.
-Spotify learns which kind of music you like most and suggests playlists and radios.
-Banks prevent fraud classifying which patterns are strange to their customers.
-It is not possible to imagine a world with no artificial intelligence.
-Behind many artificial intelligence systems, there is the pattern recognition.
-The capacity to match information from new data with information which has already been seen and is stored in memory.
-It may be used to classification, face recognition, character recognition, and so forth.
-
-The second paper lies at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence.
-Sparse Distributed Memory, or SDM for short, is a theoretical mathematical construct that seems to reflect a number of neuroscientific and psychologically plausible characteristics of a human memory.
-SDM has already been used to different patter recognition problems, like noise reduction, handwriting recognition, robot automation, and so forth.
-
-We implement a RB-Complete\footnote{Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
-The computer code has been used in a previous Ph.D.
-Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).
-What remains to be done is a full description of the systems' architecture and a performance comparison between different setups.
-
-\section{Diffusion of innovation}
-
-In 2014, a group of Princeton's researchers predicted that Facebook's users would abandon the platform by 2017 \citep{cannarella2014epidemiological}.
-The forecast was done applying a disease spreading model which has correctly predicted the abandonment of ``MySpace''.
-Facebook replied applying exactly the same methodology as Princeton's researchers.
-In their own words: ``Using the same robust methodology featured in the paper, we attempted to find out more about this `Princeton University' -- and you won't believe what we found!''.
-Then, they conclude: ``This trend suggests that Princeton will have only half its current enrollment by 2018, and by 2021 it will have no students at all, agreeing with the previous graph of scholarly scholarliness.
-Based on our robust scientific analysis, future generations will only be able to imagine this now-rubble institution that once walked this earth''.
-
-Whilst this brouhaha reminds one of the dangers of extrapolation, our third paper will revisit the prospects of our esteemed colleagues in Facebook.
-Lying at the intersection of Marketing, Diffusion of Technological Innovation, and modeling, the Bass model of diffusion of innovation will be extended, in order to account for users who, after adopting the innovation for a while, decide to reject it later on (possibly bringing down the number of active users---which is impossible in Bass' original model).
-Four alternative mathematical models are presented and discussed with the Facebook's users dataset.
-
-\section{The fine print...}
-
-Before embarking on the technical topics, small qualifications must be asked from my readers.
-First, as stated above, though these problems have immense and urgent importance to the fields of study in business, the language in which we will approach them and discuss them most naturally will be that of mathematics and computer science.
-There will not be surveys, interviews, questionnaires, or such methods typically used in the social sciences: This is basically a work of modeling.
-
-A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---this body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
-Please bear with me as we venture into the world of computational management science.

### 2018-02-08 Chapters/Conclusion.tex
Update on Overleaf.

@@
+\begin{flushright}{\slshape
+	Software is eating the World.} \\ \medskip
+    %--- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
+    --- \citep{andreesen:2011}
+\end{flushright}
+
+\bigskip
+
+\begingroup
+\let\clearpage\relax
+\let\cleardoublepage\relax
+\let\cleardoublepage\relax
+
 \chapter{Conclusion}
@@
 
-\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
+\noindent Modern management and high technology interact in multiple, profound, ways.
+Software, in particular, seems to have an immense power of entering arenas which seemed, at some point, to require specific hardware or the skill of humans.
+One of the members of this thesis committee, Dr.
+Nichols, will participate through teleconferencing over the open web, with no use of hardware specific for the task.
+The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology...
+as an executive, not as a research scientist or specialized engineer (https://www-03.ibm.com/press/br/pt/biography/53561.wss).
+Professor Andrew Ng tells students at Stanford’s Graduate School of Business that “AI is the new electricity”, as his way to emphasize the potential transformational power of the technology.
+It is not impossible that a purely digital form of money may exist.
+It is not impossible that machines may become intelligent.
+Moreover, it is not impossible that these two processes may be already beginning.
+
+@video{AndrewNgAI, 
+  title={Artificial Intelligence is the New Electricity},
+  author={Andrew Ng},
+  journal={Stanford Graduate School of Business },
+  year={2017},
+  url={https://www.youtube.com/watch?v=21EiKfQYZXc}
+}
+
+Clemenceau once said that “war is too important to be left to the generals”.
+I believe it is not far-fetched to state that technology has become too important to be isolated to the realm of computer science, or engineering, or applied mathematics, or any single discipline.
+The emergence of scientific journals with names such as Computational Management Science; INFORMS Journal on Computing; Ledger; Computational Statistics; ACM Transactions on Economics and Computation, and so forth, show that there are growing communities deeply interested in the intersection of business and the computer sciences.
+To whom, for instance, does the OpenAI project belong?
+To computing or to business?
+It seems quite arbitrary to name a single group.
+What about corporations like Uber?
+AirBnB?
+eBay?
+Amazon?
+That is why this thesis is computational and why it is business.
+This work explores topics that seem, on the surface, to belong to computer science, but their applicability and impact to businesses seem too large to be delegated away.
+
+We have started this study with two possible forms of organization of a purely digital money system; a blockchain and a directed acyclic graph.
+Are these cryptocurrencies just a new financial asset, or are they scalable to the unbanked?
+
+Bitcoin
+
+P.5. CryptoPonzi \#2. Dag/Tangle Results. Some speculative possibilities can be mentioned here:
+
+P.6. We have explored Kanerva’s Sparse Distributed Memory.
+
+P. 7. Pattern classification through SDM; Supervised learning
+
+P.8. Representation and Deep learning with SDM.
+
+P. 9. Finally, we have studied how variations of the Bass Model may reflect systems or technologies that may wither in time.
+
+Another possibility would be to compare the proposed model with a computation of the momentum of Metcalfe’s law in between competitors.
+As the reader may remember, Metcalfe’s law states that the value of a network grows $O(n^2)$ with $n$ being the number of network nodes (Note that some have argued that the value growth rate is actually $O(n log n)$ CITATION CITATION CITATION).
+
+If the proposed model and network effects reflect reality, then there could be an integrated mathematical model that explains and represents both Metcalfe’s law and the variation of the Bass model presented herein.
+
+P.10.  Message / future is open / thesis is closed.

### 2018-02-08 FrontBackmatter/Abstract.tex
Update on Overleaf.

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}
Short summary of the contents in English\dots

\vfill

\pdfbookmark[1]{Resumo}{Resumo}
\chapter*{Resumo}
Aqui entre o resumo\dots

\endgroup

\vfill

### 2018-02-08 FrontBackmatter/Acknowledgments.tex
Update on Overleaf.

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
	Showing gratitude is one of the simplest\\
    yet most powerful things humans can do\\
    for each other.} \\ \medskip
    --- Randy Pausch, The Last Lecture
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

I must, first and foremost, give thanks for the complete and unwavering support and love from all my family. My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart. Thank you for your patience and for understanding that I was physically far away but very close in heart. A special thanks to my dad, Reynaldo, who always taught me about life and who first awakened my passion for technology and engineering.

Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again. Especially, for understanding my absence.

Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.

Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.

Thanks to Celene.

Thanks to Claudio Abreu.

Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.

Finally, thanks to everyone who helped me in any way, for each word, for caring, for your support in difficult moments and for all jokes and talks that made my days worthwhile. Surely this work was done thanks to all of you. After all, nobody does it alone!

\bigskip

\endgroup

### 2018-02-20 Chapters/Chapter01.tex
Several changes

@@
 
-Bitcoin (BTC) is the first digital currency, also known as digital money, internet money and cryptocurrency.
+Bitcoin (BTC) is the first digital currency, also known as digital money, internet money, and cryptocurrency.
 It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank.
@@
 
----
-
-In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.
-
-Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.
-
-\begin{align*}
-	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
-		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\eta}} \right) dx \\
-		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\eta}} dx \\
-		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\eta}} dx \\
-\end{align*}
-
-Let $b=\frac{a}{\sqrt(2 \eta)}$ and $c = \frac{\sqrt(2 \eta)}{2a} \left( s + \frac{1}{\eta} \right)$, then $sx + \frac{x+a^2x^2/2}{\eta} = (bx + c)^2 - c^2$.
-Let $y = bx+c \Rightarrow dy = bdx$.
-
-\begin{align*}
-	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\eta}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
-		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
-		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
-		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
-\end{align*}
-
-$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
-
-Hence,
-$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
-
-As $\frac{d}{ds}c = \frac{\sqrt{2 \eta}}{2a}$,
-
-\begin{align*}
-\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \eta \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \eta \pi}s}{4ab} \\
-\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \eta \pi}}{2a} e^{\frac{1}{2 \eta a^2}} \erfc\left(\frac{1}{a \sqrt{2 \eta}}\right) \\
-\mathbf{E}[T] &= \frac{\sqrt{2 \eta \pi}}{2a} e^{\frac{1}{2 \eta a^2}} \erfc\left(\frac{1}{a \sqrt{2 \eta}}\right) \\
-\end{align*}
-
-$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.
-
-\textcolor{red}{I do not know how to interpret this.}
+%In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.
+%
+%Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.
+%
+%\begin{align*}
+%	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
+%		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\tau}} \right) dx \\
+%		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\tau}} dx \\
+%		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx \\
+%\end{align*}
+%
+%Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
+%Let $y = bx+c \Rightarrow dy = bdx$.
+%
+%\begin{align*}
+%	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
+%		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
+%		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
+%		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
+%\end{align*}
+%
+%$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
+%
+%Hence,
+%$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
+%
+%As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,
+%
+%\begin{align*}
+%\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
+%\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
+%\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
+%\end{align*}
+%
+%$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.
+%
+%\textcolor{red}{I do not know how to interpret this.}
 
@@
 
-There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the submarine attack.
-
-In the submarine attack, the attacker's goal is to spend some Bitcoins and, after a while, get their Bitcoins back.
-Thus, the attacker sends some funds to the victim, they wait 6 confirmations, and, finally, the victim confirms the funds.
-But, then, the attacker get their funds back propagating new blocks which accept a double spending transaction.
-In order to work, the attacker must propagate more blocks than the network has mined in the same period, becoming the longest chain.
+There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.
+
+In the double spending attack, the attacker's send some funds to the victim, let's say a merchant.
+They wait for $k$ confirmations of the transaction, and the victim delivers the good or the service to the attacker.
+Then, the attacker mine enough blocks with a conflicting transaction, double spending the funds which was sent to the victim.
+If the attacker is successful, the original transaction will be \textit{erased} and the victim will be left with no funds at all.
+In order to be successful, the attacker must propagate more blocks than the network in the same period, propagating a chain longer than the main chain.
 Hence, we would like to understand what the odds are that the attacker will be successful.
+This attack was originally discussed by \citet{nakamoto2008bitcoin}.
 
@@
 In this moment, it starts to mine in the head of the blockchain, just like the rest of the network.
-So, the attacker and the network are in exactly the same point.
+So, in the beginning, the attacker and the network are in exactly the same point.
 
@@
 
+When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork.
+It is said to be a soft-fork when the rules are backward compatible, and a hard-fork when the rules are not backward compatible.
+In general, a hard-fork relaxes the constraints, while the latter hardens them.
+
+Suppose that the miners' are split in two groups, $G_1$ and $G_2$, with different rules.
+Let's say $H_1$ and $H_2$ are their hash rate, respectively.
+We have two different scenarios to analyze: (i) when neither of them accepts other's blocks; and (ii) when $G_2$ accepts $G_1$'s blocks, but not the other way around.
+
+Scenario (i) is easy to analyze, because the network would just split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.
+
+Scenario (ii) is more trick.
+As $G_2$ accepts $G_1$ blocks, their hashrate will matter.
+If $H_1 > H_2$, then $G_2$ will frequently skip their blockchain, because $G_1$'s blockchain will be longer most of the time.
+If $H_1 < H_2$, then $G_2$ may have its own blockchain after a while --- a true fork.
+But what would happen if $H_1$ keeps increasing and eventually gets larger than $H_2$?
+It if happens for sufficient time, the whole $G_2$'s blockchain may be discarded in order to move to $G_1$'s blockchain when $G_1$'s gets longer.
+
+\begin{theorem}
+	When $H_1 > H_2$, the 
+\end{theorem}
+
+%What would happen if, after a while, the groups agree again. How would the blockchain merge?
+
 \chapter{DAG model}

### 2018-02-20 Chapters/Chapter02.tex
Fix sdm2.tex path
Update on Overleaf.
Several changes

@@
 
-\input{sdm2.tex}
+\input{./Chapters/sdm2.tex}
 
@@
 
+\section{Read and write operations}
+
 The reading and writing operations are executed in two steps: first, the address space is swept looking for the activated addresses; then, the operation is performed in the counter.
@@
 
-The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner.
-The linear scanner runs in a single core and is the slowest one.
-It was developed only for testing purposes.
-The thread scanner runs at the CPU in multiple threads (and our recommendation is to use the number of threads equals to twice the number of cores).
-The OpenCL scanner runs at the GPUs and support multiple devices.
-The speed of a scan depends on the processors and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark test.
+Counters may have overflow protection depending on compiling options.
+By default, there is no overflow check in the counters, since each counter is a signed 32-bit integer --- which can store any number integer -2,147,483,648 and 2,147,483,647 --- and they will not overflow with less than $2^31-1$ divided by the number of activated hard-locations.
+When $N=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes to overflow a counter.
+
+\section{Implementation details}
+
+\subsection{The distace between two bitstrings}
 
@@
 In order to count the number of ones of the XOR, the framework uses a pre-calculated table of distances --- by default, it is used a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM.
-Finally, to calculate the distance between to bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$.
+Finally, to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$.
 As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$).
 
-Internally, the address space is stored in an array of addresses, which, in turn, is stored in an array of 64-bit integers (Fig.
+The address space is stored in an array of addresses, which, in turn, is stored in arrays of 64-bit integers (Fig.
 \ref{tab:hl-addresses-detail}).
-As the addresses and their 64-bit integers are contiguous in RAM or file, there is a performance gain since they are loaded in chunks.
+As the addresses and their 64-bit integers are contiguous in RAM or file, there is a performance gain because processors load them in chunks.
 
@@
 
-\section{Proposal:  Framework Validation}
+\subsection{Scanning for activated hard-locations}
 
-In order to check whether the framework works as expected and according to the model proposed by \citep{Kanerva1988}, we run simulations which were compared to the theoretical analysis conducted some decades ago.
+Scanning for the activated hard-locations is a problem similar to well-known problems in computation geometry called ``range reporting in higher dimensions''.
+In this case, none of the known algorithms is able to solve our problem faster than $O(n)$.
+The algorithm which seems to best fit in our problem consumes $O(n)$ space and runs in $O(\log^d(n))$ \citep{chazelle1988functional}, which is really slower than $O(n)$ when, for instance, $n=1,000,000$ and $d=1,000$.
+For a review of the algorithms, see \citet{chan2011orthogonal}.
+
+In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}.
+
+It is intriguing that none of those algorithms is able to solve our scanning problem.
+The idea behind those algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space.
+But this approach does not work in our case because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the ${0, 1}^d$ space.
+So, each addresses' bit itself splits the hardlocations in half, but it does not split the search space in half since both halves still must be convered.
+For instance, let's say we have 1,000 dimensions with 1,000,000 hardlocations, and we are scanning within a circle with radius 451, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
+The search space has not been split in half because both halves have been covered (and one of them should have been skipped).
+
+Hence, as our best approach approach seems to be scanning through all hardlocations, we may distribute the scan into many tasks which will be executed independently.
+The tasks may be executed in different processes, threads, or even computers.
+They may also run in the CPU or in the GPU.
+In this case, we may take into account both the time required to distribute the tasks and the time to receive their results.
+
+The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner.
+The linear scanner runs in a single core, is the slowest one, and was developed only for testing purposes; the thread scanner runs at the CPU in multiple threads sharing memory (and our recommendation is to use the number of threads equals to twice the number of CPU cores); and the OpenCL scanner runs in multiple GPU cores and support multiple devices.
+The speed of a scan depends on the CPU and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark.
+
+The OpenCL must be initialized, which just copies the hardlocations' addresses to the GPU's memory. Then, many scans may be executed with no necessity to upload the addresses again.
+
+\section{Framework Validation}
+
+The framework has been validated comparing its results with the expected results from \citet{Kanerva1988}.
+Thus, we run simulations which were then compared to the theoretical analysis conducted some decades ago.
+
+SHOW THE RESULTS --- Kanerva's tables/figures vs Simulation.
 
@@
 
-\section{Proposal: Performance Results}
+\section{Applications}
+
+\subsection{Noise filter}
+
+We developed a noise filter using a SDM.
+
+\subsection{Classification filter}
+
+\section{Performance Results}
 
@@
 \chapter{Conclusion}
+
 Sparse Distributed Memory is a viable model of human memory, yet it does require researchers to (re-)implement a number of parallel algorithms in different architectures.
@@
 % TODO Include in the to do list simulations based on the suggestion of Murilinho.
+
+\chapter{Appendix}
+\section{Generating Kanerva's table 7.3}
+
+\includepdf[
+  pages=-,
+  pagecommand={\pagestyle{headings}},
+  %addtotoc={1,section,1,Quilling Shapes,sec:shapes}
+]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}

### 2018-02-20 Chapters/Conclusion.tex
Fix compiling error
Update on Overleaf.
Several changes

--- 2018-02-08 Chapters/Conclusion.tex
@@
 
-\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
+\begin{flushright}{\slshape
+	Software is eating the World.} \\ \medskip
+    --- Mark Andreesen \citep{andreesen:2011}
+\end{flushright}
+\bigskip
+\bigskip
+
+\noindent Modern management and high technology interact in multiple, profound, ways.
+Software, in particular, seems to have an immense power of entering arenas which seemed, at some point, to require either specific hardware or the skill of humans.
+One of the members of this thesis committee, Dr.
+Nichols, will participate through teleconferencing over the open web, with no use of hardware specific for the task.
+The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer (https://www-03.ibm.com/press/br/pt/biography/53561.wss).
+Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citet{AndrewNgAI}, as his way to emphasize the potential transformational power of the technology.
+It is not impossible that a purely digital form of money may exist.
+It is not impossible that machines may become intelligent.
+Moreover, it is not impossible that these two processes may have already begun.
+
+It is worthwhile, in this concluding section, to reflect on some ideas on what this thesis is and what it is that we, as computational management scientists, can obtain from this sort of study.
+Clemenceau once said that ``war is too important to be left to the generals''.
+I believe it is not far-fetched to state that technology has become too important to be isolated to the realm of computer science, or engineering, or applied mathematics, or any single discipline.
+The emergence of scientific journals with names such as Computational Management Science; INFORMS Journal on Computing; Ledger; Computational Statistics; ACM Transactions on Economics and Computation, and so forth, show that there are growing communities deeply interested in the intersection of business and the computer sciences.
+To whom, for instance, does the OpenAI project belong?
+To computing or to business?
+Recall that the project was created as a risk-management strategy against the far-fetched, but not impossible, possibility of having machines yielding too much power.
+What about corporations like Uber?
+AirBnB?
+Imagine a new method that increases profits by 50\% at a tech company.
+Should this method, if implementable as an algorithm (like PageRank \citet{brin1998anatomy}) or a data structure (like a blockchain) be discussed in conferences of `computer science' or `business'?
+It seems quite arbitrary to name a single group, as a whole new ecosystem seems to be emerging within those two.
+That is why this thesis is computational and why it is business.
+This work explores topics that seem, on the surface, to belong to computer science, but their applicability and impact to businesses seem too large to be delegated away, something ``for the nerds in the fifth floor''.
+As technical decisions become central to the organization of man's life, the technician becomes the visionary, the innovator, the decision-making arbiter, sometimes the billionaire.
+
+We have started this study with two possible forms of organization of a purely digital money system; a blockchain and a directed acyclic graph; we then moved to an analysis of the adoption of new technologies.
+
+The possibility that there will be some form of purely digital money has become very real.
+Consider, just as a matter of comparison, Brazil's most important company: Petrobrás.
+As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobras by ten billion dollars, while Bitcoin's is valued at more than double of Petrobras (195B usd vs 83B usd).
+These technologies should, at a minimum, be taken seriously.
+
+%P.5. CryptoPonzi #2. Dag/Tangle Results. Some speculative possibilities can be mentioned here:
+
+We have explored Kanerva’s Sparse Distributed Memory.
+In AI, SDM seems to be a particularly interesting area for study.
+The model plausibly reflects a number of well known aspects of psychology and neuroscience.
+For example, neurons can easily compute the address decoding scheme of the system.
+Neurons are fragile and may be lost, whereas the information remains preserved, due to the distributed character of the model.
+The ``tip-of-the-tongue'' behavior emerges naturally, and so does Miller’s magic number.
+
+There are three contributions made on SDM:  First, I have illuminated a discrepancy between Kanerva’s theoretical model and the real system dynamics; Also, we have seen that pattern classification through supervised learning is possible without presuming any new SDM mechanism.
+This is in contrast with the literature, that presumes additional mechanisms, like genetic algorithms, to account for supervised learning.
+Finally, we now have a tested open-source framework that offers parallelism and can become a de-facto standard in SDM research.
+The framework (i) carefully reproduces crucial figures from Kanerva’s theoretical book; (ii) shows how noise filtering and (iii) supervised learning can be done, and, through the use of (iv) Jupyter Notebooks, enables the reader to easily reproduce all the results on their own machines.
+This respects all constraints posed by Robert M.
+French in his article on Computational Modeling in Cognitive Science: A Manifesto for Change \citep{TOPS:TOPS1206}.
+
+The ability to rapidly reproduce results, and to build on prior work, is, I believe, fundamental to modern science.
+Consider, for instance, the groundbreaking successes in the arena of deep learning.
+Having standard computer libraries to work with has brought together a community, which reinforces the system, as users also gradually improve these libraries.
+It may be possible to achieve new results with multiple layers of a SDM, yet, having to start development from scratch takes a large opportunity cost from most scientists --- especially those who are less concentrated on the computer science aspects, but still would be able to contribute meaningfully.
+
+Finally, we have studied how variations of the Bass Model may reflect systems or technologies that may wither in time.
+Though some innovations, such as the radio, have gained widespread use in a sustainable form...
+One may want to review the Bass model when one is concerned with rapidly-evolving technological ecosystems.
+Hardly anyone remembers the names AskJeeves, World Wide Web Worm, Lycos, WebCrawler, or AltaVista, early web search engines; later replaced, in the market and by the market, by the almost unnoticed url http://google.stanford.edu \citep{brin1998anatomy}.
+
+Another possibility would be to compare the proposed model with a computation of the momentum of Metcalfe’s law in between competitors.
+As the reader may remember, Metcalfe’s law states that the value of a network grows $O(n^2)$ with $n$ being the number of network nodes.
+If the proposed model and Metcalfe’s network effects reflect reality, then there could be an integrated mathematical model that explains and represents both Metcalfe’s law and the variation of the Bass model presented herein.
+
+With this, I submit this thesis in the hope that all readers, present and future, may find the aforementioned studies as useful, genuine, and legitimate contributions to the thriving field of Computational Management Science.

### 2018-02-20 Chapters/Introduction.tex
Several changes

@@
 In order to be confirmed, each transaction must be both digitally signed by the owner of the money and the funds verified in the blockchain by Bitcoin's miners.
-
 The question of whether Bitcoin (or related works) can scale to billions of people is, however, far from settled.
@@
 One of the interesting parts of Bitcoin are the incentives.
-On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick, and the currency issuance rate is well known.
+On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick and global, and the currency issuance rate is well known.
 On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.
+These incentives put the community together and has maintened Bitcoin alive.
 
-The impact of Bitcoin in society -- and hence in the companies and the government -- has been increasing every day.
+The impact of Bitcoin in society -- and hence in the companies and the government -- has been growing every day.
 People are increasingly using Bitcoin to exchange money and transfer money overseas.
@@
 The poor may be included in the finance system through Bitcoin.
-People may hedge their assets against their governments' money issuance and inflation---as in the case of Venezuela.
+People may hedge their assets against their governments' money issuance and inflation --- as in the case of Venezuela.
 
@@
 
-Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'' developers, which believe that is it risky to increase the the blocksize (in which all transactions are stored).
-It is not known whether a blocksize, say, of 1Gb, would be feasible to sustain the decentralization of the network.
+Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'', which believe that is it risky to increase the the blocksize (in which all transactions are stored).
+It is not known whether a blocksize, say, of 1GB, would be feasible to sustain the decentralization of the network.
 
@@
 Netflix increases the quality of their services offering movies specific to the taste of each customer.
-Spotify learns which kind of music users like the most and suggests playlists and radios.
+Spotify learns which kind of music users like the most and suggests playlists.
 Banks prevent fraud classifying which patterns are strange to their customers.

### 2018-02-20 Chapters/sdm2.tex
Add missing files
Fix typo
Several changes

@@
 \begin{figure}[h]
-\centering\includegraphics[scale=0.75]{images/p_circle_r.pdf}
+\centering\includegraphics[scale=0.75]{./images02/p_circle_r.pdf}
 
@@
 \begin{figure}[h]
-\centering\includegraphics[scale=0.75]{images/p1_inter_p2.pdf}
+\centering\includegraphics[scale=0.75]{./images02/p1_inter_p2.pdf}
 
@@
 In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$.
+We think that a 256-dimensional memory may be important because it presents conformity to Miller's magic number \citep{Linhares2011}.
 
 \begin{figure}[h]
-\centering\includegraphics[scale=0.75]{images/hardlocations.pdf}
+\centering\includegraphics[scale=0.75]{./images02/hardlocations.pdf}
 
@@
 \begin{figure}[h]
-\centering\includegraphics[scale=0.75]{images/p1_p2_iter_read.pdf}
+\centering\includegraphics[scale=0.75]{./images02/p1_p2_iter_read.pdf}
 
@@
 \begin{figure}[h]
-\centering\includegraphics[scale=0.75]{images/p1_after_write.pdf}
+\centering\includegraphics[scale=0.75]{./images02/p1_after_write.pdf}
 
@@
 
-SDM memory or an extension of it has been used in a lot of applications, helping to solve specific problems.
+SDM memory or an extension of it has been used in many applications, helping to solve specific problems.
 \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
@@
 In the latter, all hard-locations have the same weight, because they can contribute to the sum with only one bitstring.
+
+A member of my Master's committee, professor Paulo Murilo, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
+He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative).
+Thus, Kanerva's read would be the same as $z=1$, while Chada's would be the same as $z=0$.
+Hence, we might explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.
 

### 2018-02-20 FrontBackmatter/Acknowledgments.tex
Several changes

@@
 
-I must, first and foremost, give thanks for the complete and unwavering support and love from all my family.
+%I must, first and foremost, give thanks for the complete and unwavering support and love from all my family.
 My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart.
@@
 
-Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again.
+%Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again.
 Especially, for understanding my absence.
 
-Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.
+%Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.
 
-Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
+%Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
 
-Thanks to Celene.
+%Thanks to Celene.
 
-Thanks to Claudio Abreu.
+%Thanks to Claudio Abreu.
 
-Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.
+%Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.
 

### 2018-02-20 Chapters/Conclusion.tex
Minor changes

@@
 One may want to review the Bass model when one is concerned with rapidly-evolving technological ecosystems.
-Hardly anyone remembers the names AskJeeves, World Wide Web Worm, Lycos, WebCrawler, or AltaVista, early web search engines; later replaced, in the market and by the market, by the almost unnoticed url http://google.stanford.edu \citep{brin1998anatomy}.
+Hardly anyone remembers the names AskJeeves, World Wide Web Worm, Lycos, WebCrawler, or AltaVista, early web search engines; later replaced, in the market and by the market, by the almost unnoticed url \url{http://google.stanford.edu} \citep{brin1998anatomy}.
 

### 2018-02-23 linhares-figs-captions.tex
organização original figuras

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Figures to insert in text}
\author{Author}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\section{First section}

Your text goes here.

% FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES
\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_0_15_3.jpg}
  \caption{z  0  15  3.jpg}
  \label{fig:boat1}
\end{figure}

Figure \ref{fig:boat1} shows a boat.

/images02/new-images/

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_all.jpg}
  \caption{z  all.jpg}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_0_05_1.png}
  \caption{iter z  0  0.5 1.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_01-05.png}
  \caption{iter z 01-05.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_05_1_15_3.png}
  \caption{iter z  0.5 1 1.5 3.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_05_1_15.png}
  \caption{iter $z\in \{0.5, 1, 1.5\}$ }
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_all_2.png}
  \caption{iter z all 2.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_all.png}
  \caption{iter z all.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_0_1.png}
  \caption{z 0 1.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_1_15_3_45_6.png}
  \caption{z 1 1.5 3 4.5 6.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_01-05.png}
  \caption{z 01 - 05.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_15_3_45_6.png}
  \caption{z 1.5 3 4.5 6.png}
  \label{fig:boat1}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/z_15_3_45.png}
  \caption{z 1.5 3 4.5.png}
  \label{fig:boat1}
\end{figure}

faltando as imagens do google doc...
\section{Critical Distance}

\begin{figure}[h!]
\centering

\subfloat[]{\includegraphics[width=3.1in]{./images02/new-images/kanerva-read.png}}

\subfloat[]{\includegraphics[width=3.1in]{./images02/new-images/chada-read_z_0.png}}

\caption{The critical distance behavior, as the memory is written with items.  The colors of this heatmap display the hamming distance to the item being read as a function of the distance from that item and the numbers of items registered in memory.  In the $x$-axis, the distance from the item being read is displayed. In the $y$-axis, we display the read operation's behavior as the number of items registered in the memory grows.  In (a), we have Kanerva's original model; whereas (b) displays Chada's read.}
\label{fig:crit-dist-10k-writes}

\end{figure}

\section{Speculations on the use of Information Theory}

\begin{figure}
  \includegraphics[width=\linewidth]{./images02/new-images/information_decay.png}
  \caption{information decay 4.png}
  \label{fig:boat1}
\end{figure}

\subsection{A subsection}

More text.

\end{document}

### 2018-02-23 Chapters/Chapter02.tex
Many changes in SDM text

@@
+% !TEX root = ../partial-sdm.tex
+
 \chapter{Introduction}
@@
 
+SDM memory or an extension of it has been used in many applications, helping to solve specific problems.
+\citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
+\citet{Rajesh1998} used SDM in an autonomous robot.
+\citet{Meng2009} modified SDM to clean patterns from noisy inputs.
+
+\citet{Linhares2011} showed that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
+This is the only work, to the author's knowledge, that explores some of intrinsic properties of SDM.
+
 \chapter{Sparse Distributed Memory}
@@
 It is developed in C language --- which may be loaded in many platforms and programming languages --- with a wrapper in Python.
-The Python module makes it easy to create and execute operations in a Sparse Distributed Memory.
+The Python module makes it easy to create and execute simulations in a Sparse Distributed Memory and works properly in Jupyter Notebook \citep{kluyver2016jupyter}.
+It works in both Python 2 and Python 3.
 
-We split the SDM memory in two parts: the hard-location address space and the hard-location counters.
+We split the SDM memory in two parts: the hard-location addresses and the hard-location counters.
 Thus, the addresses (bitstrings) of the hard-locations are stored in one array, while their counters in another.
-This may be useful to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be same in all of them.
+This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be same in all of them.
 As the slowest part of reading and writing operations is scanning the address space, the performance benefits are high.
@@
 Each part may be stored either in the RAM memory or in a file.
-The RAM memory is interesting for quick experiments, automated tests, and others scenarios in which the SDM may be lost.
-The file is interesting for a long-term SDM --- one may create an SDM file with 10,000 random writes, which will be copied over and over to run multiple experiments.
+The RAM memory is interesting for quick experiments, automated tests, and others scenarios in which the SDM may be lost, while the file is interesting for a long-term SDM, like creating an SDM file with 10,000 random writes, which will be copied over and over to run multiple experiments.
 The file may also be sent to another researcher or may be published with the paper to let others run their own checks and verify the results.
@@
 
-There are some other advantages splitting the address space and the counters.
-When the scanner is run in the CPU, the accessed addresses in the RAM are continuous which improves considerably performance.
-When the scanner is run in the GPU, only the address space must be uploaded to the GPU, reducing the required memory.
-
 Let a SDM memory with $N$ dimensions and $H$ hard-locations.
 Then, in a 64-bit computer, the array of hard-location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard-location counters.
-For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard-locations, using 16-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 1.9 GB of memory.
+For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard-locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.
 
 Basic operations were grouped in four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs).
-Operations include creating new bitstrings, flipping bits, generating bitstring with a specific distance from a given bitstring, scan the address space using different algorithms, write a bitstring to a counter, write in an SDM, read from an SDM, and iteratively read from an SDM until convergence.
+Operations include creating new bitstrings, flipping bits, generating bitstring with a specific distance from a given bitstring, scanning the address space using different algorithms, writing a bitstring to a counter, writing in an SDM, reading from an SDM, and iteratively reading from an SDM until convergence.
 
-\section{Read and write operations}
+\section{Bitstring}
 
-The reading and writing operations are executed in two steps: first, the address space is swept looking for the activated addresses; then, the operation is performed in the counter.
-Reading operation assemblies the bitstring according to the counters of the activated addresses, while the writing operation changes the counters.
+Bitstrings are the main structure of SDM.
+The addresses are represented in bitstrings, as well as the data.
+A bitstring is stored as an array of integers.
+Each integer may be 16-bit, 32-bit, or 64-bit long, depending on the configuration.
+By default, each integer is 64-bit long.
 
-Counters may have overflow protection depending on compiling options.
-By default, there is no overflow check in the counters, since each counter is a signed 32-bit integer --- which can store any number integer -2,147,483,648 and 2,147,483,647 --- and they will not overflow with less than $2^31-1$ divided by the number of activated hard-locations.
-When $N=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes to overflow a counter.
-
-\section{Implementation details}
+For instance, a 1,000-bit bitstring will have $\lceil 1000/64 \rceil = 16$ integers.
+These integers will have a total of $16 \cdot 64 = 1,024$ bits.
+The remaining 24 bits are always zero, so they do not affect the result of any operation.
+Bitstrings store neither how many bits they have nor the array length.
+These pieces of information are only stored in the address space.
 
@@
 
-The hamming distance between two bitstrings is calculated counting the number of ones in the XOR of the bitstrings, i.e., $d(x, y) = \text{number of ones in} x \oplus y$.
-In order to count the number of ones of the XOR, the framework uses a pre-calculated table of distances --- by default, it is used a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM.
+The distance between two bitstrings is calculated by the hamming distance, which is the number of different bits between them.
+
+It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in} x \oplus y$.
+There are several algorithms to calculate the number of ones \citep{warren2013hacker}, but the performance depends on the processor.
+So, we have implemented three different algorithms and one may be selected through compiling flags.
+The default algorithm is to use a builtin instruction from the compiler.
+
+An alternative algorithm available is the lookup.
+It pre-calculates a table of distances --- by default, it is a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM ---, which is accessed a few times.
 Finally, to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$.
@@
 
-The address space is stored in an array of addresses, which, in turn, is stored in arrays of 64-bit integers (Fig.
-\ref{tab:hl-addresses-detail}).
-As the addresses and their 64-bit integers are contiguous in RAM or file, there is a performance gain because processors load them in chunks.
+\section{Address space}
+
+An address space is a fixed collection of bitstrings, and each bitstring represents a hard-location address.
+They store the number of bitstrings, as well as the number of bits, number of integers per bitstring, and the number of remaining bits.
+
+Bitstrings are stored in a contiguous array of 64-bit integers, as shown in Figure \ref{tab:hl-addresses-detail}.
+Hence, there is some performance improvement when they are accessed \citep{pai2004linux}.
 
@@
 
-\caption{Hard-location addresses are stored in a contiguous array.
-In a 64-bit computer, each hard-location address is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil N/64 \rceil$.\label{tab:hl-addresses-detail}}
+\caption{Address space's bitstrings are stored in a contiguous array.
+In a 64-bit computer, each bitstring is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil N/64 \rceil$.\label{tab:hl-addresses-detail}}
 \end{figure}
+
+The scan for activated hard-locations is performed in an address space. It returns the indexes of the bitstrings which were inside the circle. Then, each operation uses these index in a different way.
 
@@
 In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}.
+It provides a fast search when $r/b < 0.11$ or $r/b < 0.06$, where $r$ is the radius and $b$ is the number of bits.
+But, in our case, for a 1,000 bits SDM, $r/b = 0.451$, which changes the runtime to $O(n^{0.993})$.
+This is really close to $O(n)$, but with a larger constant.
+Hence, $O(n)$ is still faster.
 
 It is intriguing that none of those algorithms is able to solve our scanning problem.
-The idea behind those algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space.
-But this approach does not work in our case because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the ${0, 1}^d$ space.
-So, each addresses' bit itself splits the hardlocations in half, but it does not split the search space in half since both halves still must be convered.
+The idea behind those computation geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space.
+But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the ${0, 1}^d$ space.
+So, each addresses' bit itself splits the hardlocations in half, but it does not split the search space in half since both halves still must be convered by the algorithm.
 For instance, let's say we have 1,000 dimensions with 1,000,000 hardlocations, and we are scanning within a circle with radius 451, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
@@
 
-Hence, as our best approach approach seems to be scanning through all hardlocations, we may distribute the scan into many tasks which will be executed independently.
+Hence, as our best approach is to scan through all hardlocations, we may distribute the scan into many tasks which will be executed independently.
 The tasks may be executed in different processes, threads, or even computers.
@@
 
-The OpenCL must be initialized, which just copies the hardlocations' addresses to the GPU's memory. Then, many scans may be executed with no necessity to upload the addresses again.
+The OpenCL must be initialized, which just copies the address space's bitstrings to the GPU's memory. Then, many scans may be executed with no necessity to upload the bitstrings again.
+
+\section{Counters}
+
+Each hard-location has one integer of data per bit. For instance, each hard-location of a 1,000 bits SDM has 1,000 bits. Those integers are stores in a counter.
+
+A counter is an array of integers which stores the data of all hard-locations. So, the counter's array has $N \cdot H$ integers.
+
+When two counters are added in a third counter, there may occur an overflow.
+Counters may have overflow protection depending on compiling options.
+By default, there is no overflow check for performance reasons.
+It is not supposed to be a problem because each counter is a signed 32-bit integer which can store any number between -2,147,483,648 and 2,147,483,647 (and they will not overflow with less than $2^{31}-1$ divided by the number of activated hard-locations).
+For instance, when $N=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes to be possible to overflow a counter.
+
+\section{Read and write operations}
+
+The reading and writing operations are executed in two steps: first, the address space is swept looking for the activated addresses; then, the operation is performed in the counters.
+Reading operation assemblies the bitstring according to the counters of the activated addresses, while the writing operation changes the counters.
+
+The iterated reading keeps reading until it gets exactly the same bitstring (or the number of maximum interations has been reached), i.e., it performs $\eta_{i+1} = \text{read}(\eta_i)$ and stops when $\eta_{k+1} = \eta_{k}$.
+If the initial bitstring is inside the critical distance of $\eta$, it will converge to $\eta$, but, if it is not, it will diverge and reach the maximum number of iterations.
+
+It is available both Kanerva's read and Chada's read.
+The latter was implemented according to a generalization proposed by professor Paulo Murilo (personal communication).
+The generalization has a $z$ parameters, which is the expoent of the counter.
+In this case, the results are floating point instead of integer, which reduces considerably the performance.
+When $z=1$, it is exactly as the Kanerva's read.
+When $z=0$, it is the Chada's read.
+We also explored how SDM would behave for different values of $z$.
+
+There is also another special read operation: the weighted reading.
+The weight is retrieved from a lookup table of integers based on the distance between the hard-location's address and the reading address.
+The weight is multiplied by the counter's value and summed.
+The rest of the read operation is exactly the same.
 
@@
 \chapter{Appendix}
-\section{Generating Kanerva's table 7.3}
+%\section{Generating Kanerva's table 7.3}
 
-\includepdf[
-  pages=-,
-  pagecommand={\pagestyle{headings}},
-  %addtotoc={1,section,1,Quilling Shapes,sec:shapes}
-]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}
+%\includepdf[
+%  pages=-,
+%  pagecommand={\pagestyle{headings}},
+%  %addtotoc={1,section,1,Quilling Shapes,sec:shapes}
+%]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}

### 2018-02-23 Chapters/sdm2.tex
Many changes in SDM text

@@
+% !TEX root = ../partial-sdm.tex
+
 Sparse Distributed Memory (SDM) is a mathematical model for cognitive memory published by \citet{Kanerva1988}.
@@
 
-\section{Applications}
-
-SDM memory or an extension of it has been used in many applications, helping to solve specific problems.
-\citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
-\citet{Rajesh1998} used SDM in an autonomous robot.
-\citet{Meng2009} modified SDM to clean patterns from noisy inputs.
-
-\citet{Linhares2011} showed that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
-This is the only work, to the author's knowledge, that explores some of intrinsic properties of SDM.
-
 \section{Concepts}
@@
 
+My advisor, professor Alexandre Linhares, has proposed another read operation: the weighted reading.
+In it, the sum of the counter's value is weighted based on the distance between the hard-location's address and the reading address.
+The logic behind it is to change the importance of each hard-location inside the circle.
+We have tested how SDM behaves when the weight is calculated using the information given by each bitstring, i.e., $w(d) = -\log_2 \left( 2^{-N} \binom{N}{d} \right) = N - \log_2 \binom{N}{d}$, as in information theory \citep{cover2012elements}.
+
 \section{Critical Distance}
@@
 
-Starting from the premise of SDM as a faithful model of human short-term memory, a better understanding of the critical distance may shed light on our understanding of the thresholds that bind our own memory.
+For further information about critical distance behavior, see \citet{Brogliato2011} and \citet{brogliato2014sparse}.

### 2018-02-23 Chapters/Chapter02.tex
Minor change SDM conclusion

@@
 
+We have generated a Docker image, which makes it even easier to explore the framework.
+After running the container, a Jupyter Notebook is avaiable with sdm-framework and other tools already installed.
+The simulations run in this thesis are promptly available to be re-executed and explored.
+We invite the reader to take a look and explore a little bit.
+
 % TODO Include in the to do list simulations based on the suggestion of Murilinho.

### 2018-02-25 Chapters/Chapter02.tex
Linhares figs

--- 2018-02-20 Chapters/Chapter02.tex
@@
 
+\begin{figure}[h!]
+  \centering
+  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_01-05.png}}
+  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_15_3_45_6.png}}
+
+  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_0_1.png}}
+  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_all_2.png}}
+
+  \caption{(a) and (b) show the behavior of a single read; (c) and (d) present the effects of 6 iterative reads.
+As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
+Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
+It seems that Kanerva did not consider that a single read would only `clean' a small number of dimensions \emph{after the critical distance}.
+What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
+  \label{fig:murillo-generalization-experiments}
+\end{figure}
+
+Physicist Paulo Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
+The results, however, have not yielded performance improvements.
+Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads.
+This is shown in Figure \ref{fig:murillo-generalization-experiments}.
+
 \section{Applications}

### 2018-02-25 Chapters/sdm2.tex
Linhares figs

--- 2018-02-20 Chapters/sdm2.tex
@@
 Starting from the premise of SDM as a faithful model of human short-term memory, a better understanding of the critical distance may shed light on our understanding of the thresholds that bind our own memory.
+
+\begin{figure}[h!]
+\centering
+
+\subfloat[Kanerva's original model]{\includegraphics[width=3.1in]{./images02/new-images/kanerva-read.png}}
+
+\subfloat[Chada's read]{\includegraphics[width=3.1in]{./images02/new-images/chada-read_z_0.png}}
+\caption{How far, in hamming distance, is a read item from the original stored item?
+Kanerva demonstrated that, after a small number of iterative readings (6 here), a critical distance behavior emerges.
+Items read at close distance converge rapidly; whereas farther items do not converge.
+Most striking is the point in which the system displays the tip-of-tongue behavior.
+Described by psychological moments when some features of the item are prominent in one's thoughts, yet the item still cannot be recalled (but an additional cue makes convergence `immediate').
+Mathematically, this is the precise distance in which, despite having a relatively high number of cues (correct bits) about the desired item, the time to convergence is infinite.
+Heatmap colors display the hamming distance the associative memory is able to cleanly converge to---or not.
+In the $x$-axis, the distance from the desired item is displayed.
+In the $y$-axis, we display the read operation's behavior as the number of items registered in the memory grows.
+These graphs are computing intensive, yet they can be easily tested by readers in our provided jupyter notebooks.
+Note the different scales.}
+\label{fig:crit-dist-10k-writes}
+
+\end{figure}
+
+Figure \ref{fig:crit-dist-10k-writes} compares the critical distance behavior under different scenarios.
+This replicates our previous results and is a first part of the process of framework validation, to which we throw our attention next.

### 2018-02-25 linhares-figs-captions.tex
Linhares figs

@@
 % FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_0_15_3.jpg}
-  \caption{z  0  15  3.jpg}
-  \label{fig:boat1}
+\begin{figure}[h!]
+  \centering
+  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_01-05.png}}
+  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_15_3_45_6.png}}
+
+  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_0_1.png}}
+  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_all_2.png}}
+
+  \caption{Physicist Paulo Murillo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
+The results, however, have not yielded performance improvements.
+Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads.
+(a) and (b) show the behavior of a single read; (c) and (d) present the effects of 6 iterative reads.
+As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
+Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
+It seems that Kanerva did not consider that a single read would only `clean' a small number of dimensions \emph{after the critical distance}.
+What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
+  \label{fig:murillo-generalization-experiments}
 \end{figure}
 
-Figure \ref{fig:boat1} shows a boat.
+% FIGURES ITER FIGURES ITER FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES FIGURES
 
-/images02/new-images/
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_all.jpg}
-  \caption{z  all.jpg}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_0_05_1.png}
-  \caption{iter z  0  0.5 1.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_01-05.png}
-  \caption{iter z 01-05.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_05_1_15_3.png}
-  \caption{iter z  0.5 1 1.5 3.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_05_1_15.png}
-  \caption{iter $z\in \{0.5, 1, 1.5\}$ }
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_all_2.png}
-  \caption{iter z all 2.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/iter_z_all.png}
-  \caption{iter z all.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_0_1.png}
-  \caption{z 0 1.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_1_15_3_45_6.png}
-  \caption{z 1 1.5 3 4.5 6.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_01-05.png}
-  \caption{z 01 - 05.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_15_3_45_6.png}
-  \caption{z 1.5 3 4.5 6.png}
-  \label{fig:boat1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{./images02/new-images/z_15_3_45.png}
-  \caption{z 1.5 3 4.5.png}
-  \label{fig:boat1}
-\end{figure}
-
-faltando as imagens do google doc...
-\section{Critical Distance}
+\section{Critical Distance Analysis: a brief review}
 
@@
 
-\subfloat[]{\includegraphics[width=3.1in]{./images02/new-images/kanerva-read.png}}
+\subfloat[Kanerva's original model]{\includegraphics[width=3.1in]{./images02/new-images/kanerva-read.png}}
 
-\subfloat[]{\includegraphics[width=3.1in]{./images02/new-images/chada-read_z_0.png}}
-
-\caption{The critical distance behavior, as the memory is written with items.
-The colors of this heatmap display the hamming distance to the item being read as a function of the distance from that item and the numbers of items registered in memory.
-In the $x$-axis, the distance from the item being read is displayed.
+\subfloat[Chada's read]{\includegraphics[width=3.1in]{./images02/new-images/chada-read_z_0.png}}
+\caption{How far, in hamming distance, is a read item from the original stored item?
+Kanerva demonstrated that, after a small number of iterative readings (6 here), a critical distance behavior emerges.
+Items read at close distance converge rapidly; whereas farther items do not converge.
+Most striking is the point in which the system displays the tip-of-tongue behavior.
+Described by psychological moments when some features of the item are prominent in one's thoughts, yet the item still cannot be recalled (but an additional cue makes convergence `immediate').
+Mathemathically, this is the precise distance in which, despite having a relatively high number of cues (correct bits) about the desired item, the time to convergence is infinite.
+Heatmap colors display the hamming distance the associative memory is able to cleanly converge to---or not.
+In the $x$-axis, the distance from the desired item is displayed.
 In the $y$-axis, we display the read operation's behavior as the number of items registered in the memory grows.
-In (a), we have Kanerva's original model; whereas (b) displays Chada's read.}
+These graphs are computing intensive, yet they can be easily tested by readers in our provided jupyter notebooks.
+Note the different scales.}
 \label{fig:crit-dist-10k-writes}
@@
 \end{figure}
+
+Figure \ref{fig:Crit-Dist-Comparison} compares the critical distance behavior under different scenarios.
 
@@
   \caption{information decay 4.png}
-  \label{fig:boat1}
+  \label{fig:boat10}
 \end{figure}

### 2018-02-25 delete.py

@@ -1,4 +1,9 @@
 import numpy as np
+#import matplotlib
+# matplotlib.use('TkAgg')
+
+
+
 import matplotlib.pyplot as plt
 
 origin = 'lower'

### 2018-02-25 Chapters/sdm2.tex

@@
 Figure \ref{fig:crit-dist-10k-writes} compares the critical distance behavior under different scenarios.
-This replicates our previous results and is a first part of the process of framework validation, to which we throw our attention next.
+This replicates our previous results \citet{Brogliato2011} and \citet{brogliato2014sparse} and is a first part of the process of framework validation, to which we throw our attention next.

### 2018-02-26 Chapters/Chapter02.tex
SDM, review of Intro and Conclusion.

@@
 
-\section{Framework Validation}
+\section{Results (i): Framework Validation}
 
@@
 Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
-It seems that Kanerva did not consider that a single read would only `clean' a small number of dimensions \emph{after the critical distance}.
+It seems that Kanerva may not have considered that a single read would only `clean' a small number of dimensions \emph{after the critical distance}.
 What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
@@
 
-\section{Applications}
-
-\subsection{Noise filter}
-
-We developed a noise filter using a SDM.
-
-\subsection{Classification filter}
-
-\section{Performance Results}
+\section{Results (ii): Performance}
 
@@
 Beyond that, we are running as state-of-the-art devices: (i) an Amazon EC2 p3.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.8xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and 8x NVIDIA K80 GPU.
+
+\section{Results (iii): Noise filtering application}
+
+We developed a noise filter using a SDM.
+
+\section{Results (iv): Supervised classification application}
+
+\section{Results (v): The possibility of unsupervised reinforcement learning}
 

### 2018-02-26 Chapters/Introduction.tex
SDM, review of Intro and Conclusion.

@@
 
-This is from almost 100 years ago.
+This humble beginnings date from almost 100 years ago.
 Today the field has advanced to a great number of nations, and the amount of applications has grown explosively.
 Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power to the hands of individuals.
-Whereas, during the war, computations were mostly done by hand, it is not outlandish to say that this original field can be referred to, today, as computational management science.
+Whilst, during the war, computations were mostly done by hand, the electronic computer took over afterwards; up to an extent that it is not outlandish to say that this original field can be referred to, today, as computational management science.
 
@@
 On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.
-These incentives put the community together and has maintened Bitcoin alive.
+These incentives keep the community together and have maintened Bitcoin alive.
 
@@
 
-We implement a RB-Complete\footnote{`Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
+We implement a RB-Complete\footnote{`Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from previous theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
 The computer code has been used in a previous Ph.D.
 Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).
-What remains to be done is a full description of the systems' architecture and a performance comparison between different setups.
+We also reproduce previous experiments and present new possibilities for SDM.
 
@@
 Whilst this brouhaha reminds one of the dangers of extrapolation, our third paper will revisit the prospects of our esteemed colleagues in Facebook.
-Lying at the intersection of Marketing, Diffusion of Technological Innovation, and modeling, the Bass model of diffusion of innovation will be extended, in order to account for users who, after adopting the innovation for a while, decide to reject it later on (possibly bringing down the number of active users---which is impossible in Bass' original model).
+Lying at the intersection of Marketing, Diffusion of Technological Innovation, and modeling, the Bass model of diffusion of innovation will be extended, in order to account for users who, after adopting the innovation for a while, decide to reject it later on (possibly bringing down the number of active users---something impossible in Bass' original model).
 Four alternative mathematical models are presented and discussed with the Facebook's users dataset.
@@
 
-A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies and will agree that---if the topics discussed herein are handled with the proper effort and care---this body of work should be deemed worthy of the originality and the scientific significance implicit to the work of a Doctoral Thesis.
-
+A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic.
 With these qualifications, we are ready to venture into the world of computational management science.

### 2018-02-26 Chapters/Chapter02.tex
SDM: canetation + info theory.

@@
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
-Thus, even though great results have been achieved using SDM, it requires great effort of researchers to improve someone's work.
+Thus, even though intriguing results have been achieved using SDM, it requires great effort of researchers to improve someone's work.
 
-It is our belief that this tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python.
+It is our belief that such a tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
@@
 The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has automatic tests, (iv) is cross-platform, and (v) is easily extended to fulfill other research models.
-Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier to readers to replicate any previous results and easily understand the source-code of the model.
+Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier for readers to replicate any previous results and easily understand the source-code of the model.
+Moreover, new machanisms and applications are also introduced.
 
-SDM memory or an extension of it has been used in many applications, helping to solve specific problems.
-\citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
+SDM, extensions of it, has been used in many applications.
+For example, \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
 \citet{Rajesh1998} used SDM in an autonomous robot.
 \citet{Meng2009} modified SDM to clean patterns from noisy inputs.
-
 \citet{Linhares2011} showed that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
-This is the only work, to the author's knowledge, that explores some of intrinsic properties of SDM.
 
@@
 The framework implements the basic operations in a Sparse Distributed Memory which may be used to create more complex operations.
-It is developed in C language --- which may be loaded in many platforms and programming languages --- with a wrapper in Python.
+It is developed in C language and the OpenCL parallel framework --- which may be loaded in many platforms and programming languages --- with a wrapper in Python.
 The Python module makes it easy to create and execute simulations in a Sparse Distributed Memory and works properly in Jupyter Notebook \citep{kluyver2016jupyter}.
@@
 Basic operations were grouped in four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs).
-Operations include creating new bitstrings, flipping bits, generating bitstring with a specific distance from a given bitstring, scanning the address space using different algorithms, writing a bitstring to a counter, writing in an SDM, reading from an SDM, and iteratively reading from an SDM until convergence.
+Operations include creating new bitstrings, flipping bits, generating a bitstring with a specific distance from a given bitstring, scanning the address space using different algorithms, writing a bitstring to a counter, writing in an SDM, reading from an SDM, and iteratively reading from an SDM until convergence.
 
@@
 
-\subsection{The distace between two bitstrings}
+\subsection{The distance between two bitstrings}
 
@@
 
-It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in} x \oplus y$.
+It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in } x \oplus y$.
 There are several algorithms to calculate the number of ones \citep{warren2013hacker}, but the performance depends on the processor.
@@
 Bitstrings are stored in a contiguous array of 64-bit integers, as shown in Figure \ref{tab:hl-addresses-detail}.
-Hence, there is some performance improvement when they are accessed \citep{pai2004linux}.
+Hence, basic pointer arithmetic provides us with performance improvements in their access, as processors realize fechts of contiguous chunks of memory  \citep{pai2004linux}.
 
@@
 
-The scan for activated hard-locations is performed in an address space. It returns the indexes of the bitstrings which were inside the circle. Then, each operation uses these index in a different way.
+The scan for activated hard-locations is performed in an address space. It returns the indexes of the bitstrings which were inside the circle. Then, each operation uses this index in a different way.
 
@@
 
-Scanning for the activated hard-locations is a problem similar to well-known problems in computation geometry called ``range reporting in higher dimensions''.
+Scanning for the activated hard-locations is a problem similar to well-known problems in computational geometry called ``range reporting in higher dimensions''.
 In this case, none of the known algorithms is able to solve our problem faster than $O(n)$.
@@
 It is intriguing that none of those algorithms is able to solve our scanning problem.
-The idea behind those computation geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space.
-But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the ${0, 1}^d$ space.
+The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space.
+But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the $\{0, 1\}^d$ space.
 So, each addresses' bit itself splits the hardlocations in half, but it does not split the search space in half since both halves still must be convered by the algorithm.
@@
 For instance, when $N=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes to be possible to overflow a counter.
+Note also that it would be improbable for a hard location dimension to be this biased before the complete saturation of the memory.
 
@@
 
-It is available both Kanerva's read and Chada's read.
-The latter was implemented according to a generalization proposed by professor Paulo Murilo (personal communication).
-The generalization has a $z$ parameters, which is the expoent of the counter.
-In this case, the results are floating point instead of integer, which reduces considerably the performance.
+The framework is available both Kanerva's read and Chada's read.
+The latter was implemented according to a generalization proposed by physics professor Paulo Murilo of UFF during the discussion of my master's defense.
+The generalization brings a parameter $z$, which is the exponent of the counter.
+In this case, the results are floating point instead of integer, which considerably reduces performance.
 When $z=1$, it is exactly as the Kanerva's read.
@@
 
-SHOW THE RESULTS --- Kanerva's tables/figures vs Simulation.
-
 The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
-This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve.
+This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
 
@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images02/kanerva-table-7-2-original.png}
+\centering\includegraphics[width=0.8\textwidth]{images02/kanerva-table-7-2-original.png}
 \caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images02/sdm-10000w-table-7-2.png}
+\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-table-7-2.png}
 \caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
@@
 \ref{sdm-100w-table-7-2}).
-This result strengthen our theory that the disparities in the computational results are due to the interaction effect of different attractors.
+This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
 
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images02/sdm-100w-table-7-2.png}
+\centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
 \caption{Results generated by the framework similar from Kanerva's original Table 7.2. It was a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
@@
 \begin{figure}[h]
-\centering\includegraphics[width=0.5\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
+\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
 \caption{This graph shows the interaction effects more clearly.
@@
 
+\section{Results (vi): Information theoretical write operation}
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Kanerva's model]{\includegraphics[width=\linewidth]{./images02/new-images/kanerva-15iter-reads.jpeg}}
+
+  \subfloat[Write process weighted by the amount of information contained in the distance between the written bitstring and each hard location]{\includegraphics[width=\linewidth]{./images02/new-images/info-theory-15iter-reads.jpeg}}
+
+  \caption{(a) and (b) show the behavior of the critical distance under Kanerva's model and the information-theoretic one, respectively.}
+  \label{fig:info-theory-experiments}
+\end{figure}
+
+My advisor, Alexandre Linhares, has proposed another read operation: the weighted reading.
+In it, the sum of the counter's value is weighted based on the distance between the hard-location's address and the reading address.
+The logic behind it is to vary the importance of each hard-location inside the circle.
+It is only natural that one encodes an item in near hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
+Closer hard locations have lower probabilies and therefore should encode more information.
+We know that the probability of a distance $d$ is $\binom{N}{d}/2^N$, and, acordingly, we have tested how SDM behaves when write weights are calculated using the amount of information given by each bitstring, i.e., $w(d) = -\log_2 \left( 2^{-N} \binom{N}{d} \right) = N - \log_2 \binom{N}{d}$, as in information theory \citep{cover2012elements}.
+
+The results can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
+It seems that the critical distance increases by a number of bits.
+Note that 10 additional bits imply an attractor $2^{10}$ of the size of the original.
+Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have large asymmetries.
+
 \chapter{Conclusion}
@@
 
-We propose to provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypothesis and test them computationally through minimal effort.
-Though the framework is not well-documented for public release at this time, it has already served as the backbone of Chada's Ph.D.
+We propose to provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypotheses and test them computationally through minimal effort.
+The framework is well-documented for public release at this time (http://sdm-framework.readthedocs.io), it has already served as the backbone of Chada's Ph.D.
 thesis.
-The single-line command ``pip install sdm'' will install the framework on posix-like systems, and single-line commands will let users test the framework, generate some of the figures from Kanerva's theoretical predictions in their own machines, and --- if interested enough ---, test their own theories and improve the framework in an open-source fashion.
+The single-line command ``pip install sdm'' will install the framework on posix-like systems, and single-line commands will let users test the framework, generate some of the figures from Kanerva's theoretical predictions in their own machines, and --- if interested enough ---, test their own theories and improve the framework, and the benchmarks used to evaluate the framework, in open-source fashion.
 It is our belief that such work is a necessary component towards accelerating research in this promising field.
@@
 We have generated a Docker image, which makes it even easier to explore the framework.
-After running the container, a Jupyter Notebook is avaiable with sdm-framework and other tools already installed.
+After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.
 The simulations run in this thesis are promptly available to be re-executed and explored.

### 2018-02-26 Chapters/sdm2.tex
SDM: canetation + info theory.

@@
 For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
-One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties that Euclidean distance does.
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance.
 
-Every time you are writing or reading in SDM memory, a number of addresses are activated according to their distance to the center address and your data is written in these activated addresses or read from them.
-These issues will be addressed in due detail further on, but a major difference from a traditional computer memory is that your data are always stored and retrieved in a multitude of addresses.
-This way SDM memory has robustness against loss of addresses.
+Every time write or read in SDM memory activates a number of addresses with close distance.
+The data is written in these activated addresses or read from them.
+These issues will be addressed in due detail further on, but a major difference from a traditional computer memory is that the data are always stored and retrieved in a multitude of addresses.
+This way SDM memory has robustness against loss of addresses (e.g., death of a neuron).
 
-In traditional memory each datum is stored in an address and every time you are looking for a specific datum you have to search through the memory to find out whether you have it.
-In spite of computer scientists having developed many algorithms to perform fast searches, almost all of them do a precise search.
+In traditional memory, each datum is stored in an address and every look up of a specific datum requires a search through the memory.
+In spite of computer scientists having developed beautiful algorithms to perform fast searches, almost all of them do a precise search.
 That is, if you have an imprecise clue of what you need, these algorithms will simply fail.
@@
 If one chooses a radius of 1, the SDM will activate all addresses one bit away or less from the center address.
-So, the datum $00101_{b}$ will be written to the addresses $00101_{b}$, $10101_{b}$, $01101_{b}$, $00001_{b}$, $00111_{b}$ and $00100_{b}$.
+So, the datum $00101_{b}$ will be written to the addresses $00101_{b}$, $10101_{b}$, $01101_{b}$, $00001_{b}$, $00111_{b}$, and $00100_{b}$.
+
 In this case, when one needs to retrieve the data, one could have an imprecise cue at most one bit away from $\eta$, since all addresses one bit away have $\eta$ stored in themselves.
-Extending this train of thought for larger dimensions and radius, much more addresses are activated and one can see why SDM is a distributed memory.
+Extending this train of thought for larger dimensions and radius, exponential numbers of addresses are activated and one can see why SDM is a distributed memory.
 
@@
 $N=2^{n}$.
-For $n=100$ we have $N\approx10^{30}$, which is incredibly big when related to a computer memory.
+For $n=100$ we have $N\approx10^{30}$, which is incredibly large when related to a computer memory.
 Furthermore, \citet{Kanerva1988} suggests $n$ between 100 and 10,000.
@@
 In other words, the read operation depends on many hard-locations to be successful.
-This effect is represented in Figure \ref{fig-p1-pointers}: where all hard-locations inside the circle are activated and they, individually, don't point to $\eta$.
+This effect is represented in Figure \ref{fig-p1-pointers}: where all hard-locations inside the circle are activated and they, individually, do not point to $\eta$.
 But, like vectors, adding them up points to $\eta$.
-If another datum $\nu$ is written into the memory near $\eta$, the shared hard-locations will have information from both of them and wouldn't point to either.
+If another datum $\nu$ is written into the memory near $\eta$, the shared hard-locations will have information from both of them and would not point to either.
 All hard-locations outside of the circle are also pointing somewhere (possibly other data points).
@@
 Instead of summing all hard-location counters, each hard-location evaluates its resulting bitstring individually.
-Then, all resulting bitstring are summed again, and the same rule as Kanerva applies.
+Then, all resulting bitstrings are summed again, and the same rule as Kanerva applies.
 Table \ref{tab:chada-read} shows an example of Chada's read result bitstring.
@@
 Thus, Kanerva's read would be the same as $z=1$, while Chada's would be the same as $z=0$.
-Hence, we might explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.
+Hence, we will here explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.
 
@@
 
-My advisor, professor Alexandre Linhares, has proposed another read operation: the weighted reading.
-In it, the sum of the counter's value is weighted based on the distance between the hard-location's address and the reading address.
-The logic behind it is to change the importance of each hard-location inside the circle.
-We have tested how SDM behaves when the weight is calculated using the information given by each bitstring, i.e., $w(d) = -\log_2 \left( 2^{-N} \binom{N}{d} \right) = N - \log_2 \binom{N}{d}$, as in information theory \citep{cover2012elements}.
-
 \section{Critical Distance}

### 2018-02-26 Chapters/sdm-applications.tex
SDM Applications Classification

% !TEX root = ../partial-sdm.tex

\section{Results (iv): Supervised classification application}

The supervised classification problem consists of categorize data into groups after seeing some samples from each group. First, it is presented pieces of data with their categories. The algorithm learns from these data, which is known as learning phase. Then, new pieces of data are presented and the algorithm must classify them into the already known groups. It is named supervised because  the algorithm will not create the groups itself. It will learn the groups from during the learning phase, in which the groups have already been defined and the pieces of data have already been classified into them.

Although this problem has already been studied (REF), our intention here is to show that a pure SDM may also be used to classify data. \citet{fan1997genetic} has used SDM to solve a classification problem, recognizing handwriting letters from images, but he used a mix of genetic algorithm with SDM, which is very different from the original SDM described by \cite{Kanerva1988}. Even though his algorithm has classified properly, we were intrigued whether a pure SDM would also classify successfully.

Hence, we have developed a supervised classification algorithm based on a pure SDM as our main memory. Our goal was to classify noisy images into their respective letters (case sensitive) and numbers. For some examples, see Figure \ref{fig-classification-examples}.

\begin{figure}[h]
\centering\includegraphics[width=\textwidth]{./images02/classification/example.png}
\caption{Examples of noisy images with uppercase letters, lowercase letters and numbers.
\label{fig-classification-examples}}
\end{figure}

Each image was mapped into a bitstring in which the bits were set according to the color of each pixel of the image. So, white pixels were equal to bit 0, and black pixels to bit 1. This was a bijective mapping (or one-to-one mapping), i.e., there was only one bitstring for each image, and there was only one image for each bitstring.

We had a total of 62 classification group. For each of them, it was generated a random bitstring. Thus, the groups' bitstrings were orthogonal between any two of them.

During the learning phase, we have generated 100 noisy images for each character. For an example, see the generated images for letter A in Figure \ref{fig-classification-training-A}. Then, we have wrote the classification group bitstring into the bitstring associated to each noisy image, i.e., write(bs\_image, bs\_label).

\begin{figure}[h]
\centering\includegraphics[width=\textwidth]{./images02/classification/trainingA.png}
\caption{100 noisy images generated to train label A.
\label{fig-classification-training-A}}
\end{figure}

\section{Results (iii): Noise filtering application}

\section{Results (v): The possibility of unsupervised reinforcement learning}

### 2018-02-26 Chapters/sdm-applications.tex
SDM Classified application

@@
 
-The supervised classification problem consists of categorize data into groups after seeing some samples from each group.
+Supervised classification problem consists of categorize data into groups after seeing some samples from each group.
 First, it is presented pieces of data with their categories.
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[width=\textwidth]{./images02/classification/example.png}
@@
 
-Each image was mapped into a bitstring in which the bits were set according to the color of each pixel of the image.
+The images had 31 pixels of width and 32 pixels of height, totaling 992 pixels per image.
+Each image was mapped into a 1,000 bit bitstring in which the bits were set according to the color of each pixel of the image.
 So, white pixels were equal to bit 0, and black pixels to bit 1.
+The 8 remaining bits were all set to zero.
 This was a bijective mapping (or one-to-one mapping), i.e., there was only one bitstring for each image, and there was only one image for each bitstring.
 
-We had a total of 62 classification group. For each of them, it was generated a random bitstring. Thus, the groups' bitstrings were orthogonal between any two of them.
+% TODO Add image showing the association between pixels and bits.
+
+A total of 62 classification groups have been trained in the SDM.
+For each of them, it was generated a random bitstring.
+Thus, the groups' bitstrings were orthogonal between any two of them.
+There is one image for each of the 62 groups in Figure \ref{fig-classification-groups}.
+Notice that the SDM has never seen a single image with no noise.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images02/classification/groups.png}
+\caption{One noisy image for each of the 62 classification groups.
+\label{fig-classification-groups}}
+\end{figure}
+
+The association of images to groups was stored as sequences in SDM, as detailed by \citet{Kanerva1988} in Chapter 8.
+During the learning phase, the image bitstrings were stored pointing to their groups bitstrings, i.e., write(addr=bs\_image, datum=bs\_label).
+Thus, in order to classify an unknown image, we only had to read from its address and check which group has been found.
+
+% TODO Add image showing the pointers.
 
 During the learning phase, we have generated 100 noisy images for each character.
-For an example, see the generated images for letter A in Figure \ref{fig-classification-training-A}.
+The images had 5\% of noise, i.e., 5\% of their pixels have been randomly flipped.
+For example, see the generated images for letter A in Figure \ref{fig-classification-training-A}.
 Then, we have wrote the classification group bitstring into the bitstring associated to each noisy image, i.e., write(bs\_image, bs\_label).
+For a complete image training set, see Appendix XYZ.
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[width=\textwidth]{./images02/classification/trainingA.png}
@@
 
+Finally, we have assess the performance of our classifier.
+We had done it in three different scenarios: high noise (20\%), low noise (5\%) and no noise.
+See Figures \ref{fig-classification-noise-high} and \ref{fig-classification-no-noise} for images with 20\% noise and no noise.
+The low noise scenario had the same noise as the training set.
+For each scenario, we had classified 620 unknown images with 10 images per group.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=0.75\textwidth]{./images02/classification/noise-high.png}
+\caption{Images generated using a 20\% noise for the high noise scenario.
+\label{fig-classification-noise-high}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=0.75\textwidth]{./images02/classification/no-noise.png}
+\caption{Images generated for the no noise scenario.
+\label{fig-classification-no-noise}}
+\end{figure}
+
+The performance was calculated as the percentage of hits for each group.
+We did not expected the same performance for all groups because some groups become very similar to other depending on the noise level, and this similarity may even confuse a person (see Figure \ref{fig-classification-similarity}).
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[``i'', ``l'', and ``r'' with 20\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/ilr-high-noise.png}}
+
+  \subfloat[``i'', ``l'', and ``r'' with 5\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/ilr-low-noise.png}}
+
+  \subfloat[``c'', ``d'', and ``o'' with 20\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/cdo-high-noise.png}}
+
+  \subfloat[``c'', ``d'', and ``o'' with 5\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/cdo-low-noise.png}}
+
+  \subfloat[``G'', ``O'', and ``Q'' with 20\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/GOQ-high-noise.png}}
+
+  \subfloat[``G'', ``O'', and ``Q'' with 5\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/GOQ-low-noise.png}}
+
+  \caption{Images of different characters which may be confusing depending on the noise level.}
+  \label{fig-classification-similarity}
+\end{figure}
+
+In the no noise scenario, the classifier has hit all characters, except letter ``l'' which was wrongly associated to the group of ``i''.
+We believe that it happened because the classifier had never seen an image with no noise and the difference between the images of ``l'' and ``i'' is smaller than the critical distance.
+So, both groups have been merged and it would converge to only one of them.
+In our simulation, it happened to be the group of ``i''.
+
+In the low noise scenario, it has made few mistakes.
+It correctly classified all images but some from characters ``b'', ``e'', ``f'', ``l'', ``t'', and ``9''.
+It completely classified ``l'' images to the ``i'' group.
+In the other cases, it made just a few mistakes.
+See Figure \ref{fig-classification-low-noise-results} to check the images and their classification.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[Images from character ``b which were classified as {[b, b, b, h, b, o, b, h, b, b]}, respectively.
+It has made 3 misses.]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-b.png}}
+
+  \subfloat[Images from character ``e which were classified as {[e, e, e, e, e, e, e, e, o, e]}, respectively.
+It has made 1 miss.
+]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-e.png}}
+
+  \subfloat[Images from character ``f which were classified as {[i, f, f, I, I, I, f, f, f, f]}, respectively.
+It has made 4 misses.
+]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-f.png}}
+
+  \subfloat[Images from character ``l'' which were classified as {[i, i, i, i, i, i, i, i, i, i]}, respectively.
+It has missed them all, as if both groups have been merged.
+]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-l.png}}
+
+  \subfloat[Images from character ``t'' which were classified as {[t, t, t, t, t, t, t, i, t, t]}, respectively.
+It has made 1 miss.
+]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-t.png}}
+
+  \subfloat[Images from character ``9'' which were classified as {[9, 9, 0, 9, 9, 9, 0, 0, 9, 9]}, respectively.
+It has made 3 misses.
+]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-9.png}}
+
+  \caption{Characters in the low noise scenario in which the classifier has made at least one mistake.
+In all the other cases, it correctly classified the images.
+We may notice that the groups of ``i'' and ``l'' have been completely merged by the classifier, because it cannot distinguish them, not even with no noise.}
+  \label{fig-classification-low-noise-results}
+\end{figure}
+
+The high noise scenario is the most interesting, because, even in a high noise level, the classifier has hit most of the characters.
+It has hit all images for 44 out of 62 groups, and made at least one miss for the other 18 groups.
+The misses may be seen in details in Figure \ref{fig-classificiation-high-noise-misses}.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[Images from character ``B'' which were classified as {[S, B, B, B, B, B, B, B, B, B]}.
+It has made 1 mistake.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-B.png}}
+
+  \subfloat[Images from character ``O'' which were classified as {[G, G, O, O, O, O, O, O, O, O]}.
+It has made 2 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-O.png}}
+
+  \subfloat[Images from character ``T'' which were classified as {[T, T, T, T, T, I, T, T, T, T]}.
+It has made 1 mistake.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-T.png}}
+
+  \subfloat[Images from character ``Y'' which were classified as {[Y, I, Y, Y, Y, Y, Y, Y, Y, Y]}.
+It has made 1 mistake.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-Y.png}}
+
+  \subfloat[Images from character ``b'' which were classified as {[o, o, o, b, o, h, h, b, b, o]}.
+It has made 7 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-b2.png}}
+
+  \subfloat[Images from character ``c'' which were classified as {[c, c, c, c, c, o, c, c, c, o]}.
+It has made 2 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-c.png}}
+
+  \subfloat[Images from character ``e'' which were classified as {[e, o, e, o, o, o, e, o, o, e]}.
+It has made 6 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-e.png}}
+
+  \subfloat[Images from character ``f'' which were classified as {[I, I, I, I, i, I, I, I, I, I]}.
+It has missed them all.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-f.png}}
+
+\end{figure}
+\begin{figure}[!htb]\ContinuedFloat
+
+  \subfloat[Images from character ``i'' which were classified as {[i, i, i, I, i, i, i, i, I, i]}.
+It has made 2 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-i.png}}
+
+  \subfloat[Images from character ``j'' which were classified as {[j, j, j, I, I, j, j, j, j, I]}.
+It has made 3 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-j.png}}
+
+  \subfloat[Images from character ``l'' which were classified as {[I, i, I, I, I, I, i, I, I, i]}.
+It has missed them all.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-l.png}}
+
+  \subfloat[Images from character ``n'' which were classified as {[u, n, n, n, n, n, u, u, u, h]}.
+It has made 5 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-n.png}}
+
+  \subfloat[Images from character ``q'' which were classified as {[q, q, q, q, q, q, q, q, q, g]}.
+It has made 1 mistake.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-q.png}}
+
+  \subfloat[Images from character ``t'' which were classified as {[I, r, I, i, I, i, i, i, I, i]}.
+It has missed them all.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-t2.png}}
+
+  \subfloat[Images from character ``1'' which were classified as {[1, I, 1, I, 1, 1, I, I, 1, I]}.
+It has made 5 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-1.png}}
+
+  \subfloat[Images from character ``7'' which were classified as {[7, 7, 7, I, 7, I, I, 7, 7, 7]}.
+It has made 3 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-7.png}}
+
+\end{figure}
+\begin{figure}[!htb]\ContinuedFloat
+
+  \subfloat[Images from character ``8'' which were classified as {[8, 6, 6, 6, 8, d, 8, 8, d, 6]}.
+It has made 6 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-8.png}}
+
+  \subfloat[Images from character ``9'' which were classified as {[9, 0, 6, 0, 9, 0, 0, 9, 0, 0]}.
+It has made 7 mistakes.
+]{\includegraphics[width=\textwidth]{./images02/classification/high-noise-9.png}}
+
+  \caption{Characters in the high noise scenario in which the classifier has made at least one mistake. In all the other cases, it correctly classified the images.}
+  \label{fig-classification-high-noise-misses}
+\end{figure}
+
+The critical distance plays an important role in the classification error.
+As we have 62 groups and each have been trained with 100 images, there were 6,200 writes to the memory.
+When an image is being classified, it will have to converge to a group, and the convergence depends on the distance between this image and the images from the training set, i.e, in the noise level.
+
+These results show that the SDM may be used as a supervised classification algorithm.
+Although we do not believe that the mapping between images and bitstrings are even close to the way human cognition deals with images, we believe the results are interesting and useful to many possible real world problems.
+
 \section{Results (iii): Noise filtering application}

### 2018-02-27 Chapters/sdm-applications.tex
Chapter 02 SDM

@@
 
-\section{Results (iii): Noise filtering application}
+\section{Results (iii): Supervised image noise filtering application}
+
+Image noise filtering consists in removing the noise from an input, in out case an image.
+Our images are black \& white images and the noise is generated randomly flipping some of their pixels from black to while and vice versa.
+In Figure \ref{fig-filter-progressive-noise}, we may see an image with different levels of noise, from 0\% to 45\% in steps of 5\%.
+It makes no sense to apply 50\% of noise because it would absolutely randomize the image.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images02/filter/progressive-noise.png}
+\caption{Progressive noise into letter ``A'', from 0\% to 45\% in steps of 5\%.
+\label{fig-filter-progressive-noise}}
+\end{figure}
+
+The images have 30 x 30 pixels, totaling 900 pixels per image.
+Each image is mapped into a 1,000 bit bitstring in which the bits are set according to the color of each pixel of the image.
+White pixels are equal to bit 0, and black pixels to bit 1.
+The 100 remaining bits are all set to zero.
+This is a bijective mapping (or one-to-one) from images and bitstrings, i.e., there is one, and only one, bitstring for each image, and vice versa.
+
+In the learning phase, noisy images are generated and they are written into SDM chunked with their labels.
+The chunk was calculated using the exclusive or (XOR) operator.
+So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- \pyth{write(addr=bs_image ^ bs_label, datum=bs_image)}.
+
+Finally, in order to remove the noise of a new image, first we have to classify it (possibly using the already presented classification algorithm), and then we just have to read from the chunked address until it converges.
 
 \section{Results (v): The possibility of unsupervised reinforcement learning}
+
+Reinforcement learning has increasing prominence in the media after AlphaZero has won all games from both the best chess grandmasters in the world and the best chess engines.
+What is incredible about these victories is that AlphaZero has almost no knowledge about chess game and has learned all its movement playing against itself for 4 hours.
+Basically, it knowns only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm.
+
+Reinforcement learning is a machine learning algorithm which learns from the rewards of its actions.
+So, it receives the game state as input, then it decides which action will be taken, and finally it learns from the rewards of all the actions it has chosen.
+In theory, it learns after each reward feedback it receives, improving its decision over time and presenting intelligent behavior.
+A positive reward would indicate that the chosen action should be encouraged.
+While a negative reward would indicate the opposite.
+In some algorithms, there may be a neutral reward which would indicate that the chosen action was neither positive nor negative.
+How each type of reward should be handled depends on each algorithm.
+
+We have done some experiments with an SDM as a memory for a TicTacToe player.
+Basically, it receives the current board state and returns which action should be played.
+In the end of the game, it receives the sequences of boards and the winner, and is supposed to learn from them.
+
+Our algorithm to decide what should be player is very simples: it read the current board from SDM.
+If the reading converges to another board, it chooses the movement which would bring the current board to the one read from SDM.
+If the reading does not converge, it just play randomly.
+
+After a game has finished, it is time to learn from its decisions.
+Thus, if SDM wins the game, it will write the whole sequence of boards to SDM.
+Let $b_0, b_1, b_2, \dots, b_n$ be the board sequence of the game.
+Then it will write $b_0 \rightarrow b_1 \rightarrow b_2 \rightarrow \cdots \rightarrow b_n$, with possibly different weights for each transition.
+If it loses, it will reverse the board (replace X by O and vice versa), and will act as if it had won.
+Hence, it will learn which sequences lead to victory.
+When a new board appears, it may have already seen that situation and will decide according to the sequences which goes towards victory.
+This is our positive reward learning.
+
+It is also important to learn when a draw happens --- after all, it is better to tie than to lose, right?
+In this case, the sequence of boards is also written to SDM, but with no weight at all.
+So, if the board has appeared both in a tie sequence and in a winning sequence, it would be more likely to choose the winning one because it was written with greater weight.
+This is our neutral reward learning.
+
+Finally, we also want to prevent losing games.
+So, when it loses a game, it will stimulate movements different from the chosen ones.
+Thus, for each transition $b_k \rightarrow b_{k+1}$ made by its action, it will write all possible transitions from $b_k$ but $b_{k+1}$.
+
+Internally, every board is mapped into a random bitstring and passed to SDM.
+Thus, SDM knows nothing about the boards themselves.
+It knowns only about their transition and which ones would lead to either a victory or a draw.
+As every two boards are orthogonal, SDM does not known whether two boards are consecutives or not.
+The only link between two boards is the transition written in SDM.
+
+After all, SDM knowns nothing about the boards themselves and yet it may learn how to play TicTacToe.
+
+In order to properly run the discussed algorithms, it is necessary to have two SDMs: a 0-fold and a 1-fold SDM.
+In the 0-fold SDM, every bitstring is written to its own address.
+In the 1-fold SDM, every bitstring points somewhere else.
+So, the transitions are written in the 1-fold SDM, while the boards themselves are written to the 0-fold SDM.
+The boards are written only once in the 0-fold, no matter how many times they appear.
+The transitions may be written more than once in the 1-fold SDM, because it would reinforce that transition.
+
+In more details, the next movement decision consists in one read from the 1-fold SDM, resulting in a bitstring.
+Then this bitstring is used in an interative reading from the 0-fold SDM, which will converge to the bitstring associated with the next board.
+If it does not converge to any board, than SDM will choose a random movement and learn from it.
+
+The weight used when writing a winning sequence is calculated using ...
+
+--- talk about player generations ---
+
+\subsection{Training}
+
+It is an unsupervised algorithm because SDM learns playing against an opponent, who may be another SDM player, a human, or a player whose movements are always aleatory.
+
+Thus, in order to train a SDM player, we just have to keep it playing over and over.
+
+\subsection{Results}

### 2018-02-27 Chapters/sdm-applications.tex
Some test to TTT boards

@@
 
+\begin{forest}
+  TTT/.style args={#1:#2}{
+    make tab/.expanded=\forestove{content},
+    label={\pgfkeysvalueof{/forest/label position #1}:$#2$}
+  },
+  TTT*/.style={
+    make tab=::/::/::,
+    content/.expand once=%
+    \expandafter\vphantom\expandafter{\romannumeral-`0\forestov{content}},
+    draw=none,
+    append after command={(\tikzlastnode.north) edge (\tikzlastnode.south)},
+    for descendants={before computing xy={l*=1.2}},
+  },
+  th/.style=thick,
+  for tree={node options=draw, inner sep=+0pt, parent anchor=south, child anchor=north}
+%
+[::/::/::, TTT=r:1
+  [x::/::/::, TTT=r:-1
+    [x::/::/::, TTT=r:-1
+      [x::/::/::, TTT=r:-1
+      ]
+    ]
+  ]
+]
+\end{forest}
+
 \subsection{Results}

### 2018-02-28 Chapters/sdm-applications.tex
SDM TicTacToe board sequence
Improve TicTacToe board sequence

@@
 
-\section{Results (iv): Supervised classification application}
+\chapter{Results (iv): Supervised classification application}
 
@@
 
-\section{Results (iii): Supervised image noise filtering application}
+\chapter{Results (iii): Supervised image noise filtering application}
 
@@
 
-\section{Results (v): The possibility of unsupervised reinforcement learning}
+\chapter{Results (v): The possibility of unsupervised reinforcement learning}
 
@@
 Thus, if SDM wins the game, it will write the whole sequence of boards to SDM.
-Let $b_0, b_1, b_2, \dots, b_n$ be the board sequence of the game.
+Let $b_0, b_1, b_2, \dots, b_n$ be the board sequence of the game (see Figure \ref{fig-ttt-example}).
 Then it will write $b_0 \rightarrow b_1 \rightarrow b_2 \rightarrow \cdots \rightarrow b_n$, with possibly different weights for each transition.
@@
 This is our positive reward learning.
+
+\begin{figure}
+    \captionsetup[subfigure]{labelformat=empty}
+    %\centering
+
+    \subfloat[$b0$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+     & & \\\hline  & & \\\hline  & &
+    \end{tabular}}
+    {$\xrightarrow{w_0}$}
+    \subfloat[$b1$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+     & & \\\hline  &x& \\\hline  & &
+    \end{tabular}}
+    {$\xrightarrow{w_1}$}
+    \subfloat[$b2$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+     & & \\\hline o&x& \\\hline  & &
+    \end{tabular}}
+    {$\xrightarrow{w_2}$}
+    \subfloat[$b3$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+    x& & \\\hline o&x& \\\hline  & &
+    \end{tabular}}
+    {$\xrightarrow{w_3}$}
+    \subfloat[$b4$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+    x& & \\\hline o&x& \\\hline  & &o
+    \end{tabular}}
+    {$\xrightarrow{w_4}$}
+    \subfloat[$b5$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+    x&x& \\\hline o&x& \\\hline  & &o
+    \end{tabular}}
+    {$\xrightarrow{w_5}$}
+    \subfloat[$b6$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+    x&x&o\\\hline o&x& \\\hline  & &o
+    \end{tabular}}
+    {$\xrightarrow{w_6}$}
+    \subfloat[$b7$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
+    x&x&o\\\hline o&x& \\\hline  &x&o
+    \end{tabular}}
+
+    \caption{Example of a game with 7 movements in which X wins.\label{fig-ttt-example}}
+\end{figure}
 

### 2018-02-28 Chapters/Chapter02.tex
Merge.
SDM Applications Classification
SDM Classified application

@@
 
-\section{Results (i): Framework Validation}
+\chapter{Results (i): Framework Validation}
 
@@
 
-\section{Results (ii): Performance}
+\chapter{Results (ii): Performance}
 
@@
 
-\section{Results (iii): Noise filtering application}
+\input{./Chapters/sdm-applications.tex}
 
-We developed a noise filter using a SDM.
-
-\section{Results (iv): Supervised classification application}
-
-\section{Results (v): The possibility of unsupervised reinforcement learning}
-
-\section{Results (vi): Information theoretical write operation}
+\chapter{Results (vi): Information theoretical write operation}
 
@@
 
+Future work: classification with context using sequences --- for words instead of only letters.
+
 We have generated a Docker image, which makes it even easier to explore the framework.

### 2018-03-01 Chapters/sdm2.tex

@@
 Most notable among these are the tip-of-the-tongue phenomenon, conformity to Miller's magic number \citep{Linhares2011} and robustness against loss of neurons.
+
+The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings.
+The distance between two bitstrings is calculated using the Hamming distance.
+It is defined for two bitstrings of equal length as the number of positions at which the bits are different.
+For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance.
+
+\begin{figure}[ht!]
+  \centering
+  \subfloat[]{\includegraphics[width=2.4in]{./images02/new-images/qn3.png}}
+  \subfloat[]{\includegraphics[width=2.4in]{./images02/new-images/qn7.png}}
+
+  \subfloat[]{\includegraphics[width=4.2in]{./images02/new-images/qn10.png}}
+
+  \caption{Graph $Q_n$, for $n \in$ \{3, 7, 10\}.
+Each node represents a possible bitstring in $\{0,1\}^N$, and two nodes are linked if the bitstrings differ by a single dimension.
+A number of observations can be made here.
+First, the number of nodes grows as $2^N$ as N grows; which makes use of the entire space infeasible as $N>>20$.
+Another interesting observation, better seen in the figures below, is that most of the space lies at the center, at a distance of around 500 from any given vantage point.
+Though Kanerva has derived many combinatorial properties of the space, I believe that this is a aesthetically appealing object on its own, and beautiful results can be found in the graph-theoretical literature.\label{The hypercube is also studied in graph theory.}}
+  \end{figure}
+
+%FOOTNOTE:
+
+Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
+Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^N$ to the set of hard locations it belongs to?
+In other words… though this would break Kanerva's assumption of a uniformly distributed set of hard locations for a perfectly symmetrical set of hard locations, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
+
+%  (See \cite{@article{harary1988survey,
+%    title={A survey of the theory of hypercube graphs},
+%    author={Harary, Frank and Hayes, John P and Wu, Horng-Jyh},
+%    journal={Computers \& Mathematics with Applications},
+%    volume={15},
+%    number={4},
+%    pages={277--289},
+%    year={1988},
+%    publisher={Elsevier}
+%  }}
 
@@
 \end{figure}
-
-The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings.
-The distance between two bitstrings is calculated using the Hamming distance.
-It is defined for two bitstrings of equal length as the number of positions at which the bits are different.
-For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
-One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance.
 

### 2018-03-01 Chapters/sdm2.tex
Fix Qn images problem

--- 2018-02-26 Chapters/sdm2.tex
@@
 Most notable among these are the tip-of-the-tongue phenomenon, conformity to Miller's magic number \citep{Linhares2011} and robustness against loss of neurons.
+
+The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings.
+The distance between two bitstrings is calculated using the Hamming distance.
+It is defined for two bitstrings of equal length as the number of positions at which the bits are different.
+For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance.
+
+\begin{figure}[p!]
+  \centering
+  \subfloat[]{\includegraphics[width=2.2in]{./images02/new-images/qn3.png}}
+  \subfloat[]{\includegraphics[width=2.2in]{./images02/new-images/qn7.png}}
+
+  \subfloat[]{\includegraphics[width=4.0in]{./images02/new-images/qn10.png}}
+  \caption{Graph $Q_n$, for $n \in$ \{3, 7, 10\}.
+Each node represents a possible bitstring in $\{0,1\}^n$, and two nodes are linked if the bitstrings differ by a single dimension.
+A number of observations can be made here.
+First, the number of nodes grows as $2^n$ as $n$ grows; which makes use of the entire space infeasible as $n>>20$.
+Another interesting observation, better seen in the figures below, is that most of the space lies at the center, at a distance of around 500 from any given vantage point.
+Though Kanerva has derived many combinatorial properties of the space, I believe that this is a aesthetically appealing object on its own, and beautiful results can be found in the graph-theoretical literature.\label{hypercubes}}
+\end{figure}
+
+%FOOTNOTE:
+
+Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions, as in Figure \ref{hypercubes}.
+Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it belongs to?
+In other words… though this would break Kanerva's assumption of a uniformly distributed set of hard locations for a perfectly symmetrical set of hard locations, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
+
+%  (See \cite{@article{harary1988survey,
+%    title={A survey of the theory of hypercube graphs},
+%    author={Harary, Frank and Hayes, John P and Wu, Horng-Jyh},
+%    journal={Computers \& Mathematics with Applications},
+%    volume={15},
+%    number={4},
+%    pages={277--289},
+%    year={1988},
+%    publisher={Elsevier}
+%  }}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[scale=0.75]{./images02/p_circle_r.pdf}
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[scale=0.75]{./images02/p1_inter_p2.pdf}
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[scale=0.75]{./images02/hardlocations.pdf}
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[scale=0.75]{./images02/p1_p2_iter_read.pdf}
@@
 
-\begin{figure}[h]
+\begin{figure}[!htb]
 \centering\includegraphics[scale=0.75]{./images02/p1_after_write.pdf}
@@
 
-\begin{figure}[h!]
+\begin{figure}[!htb]
 \centering

### 2018-03-02 Chapters/sdm-applications.tex
Minor changes
sdm + bibtex

@@
 The chunk was calculated using the exclusive or (XOR) operator.
-So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- \pyth{write(addr=bs_image ^ bs_label, datum=bs_image)}.
+So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- write($addr=bs_{image} XOR bs_{label}, datum=bs_{image}$).
 

### 2018-03-02 Chapters/sdm2.tex
sdm + bibtex

@@
   \centering
-  \subfloat[]{\includegraphics[width=2.2in]{./images02/new-images/qn3.png}}
-  \subfloat[]{\includegraphics[width=2.2in]{./images02/new-images/qn7.png}}
+  \subfloat[$Q_3$]{\includegraphics[width=2.2in]{./images02/new-images/qn3.png}}
+  \subfloat[$Q_7$]{\includegraphics[width=2.2in]{./images02/new-images/qn7.png}}
 
-  \subfloat[]{\includegraphics[width=4.0in]{./images02/new-images/qn10.png}}
-  \caption{Graph $Q_n$, for $n \in$ \{3, 7, 10\}.
+  \subfloat[$Q_{10}$]{\includegraphics[width=4.0in]{./images02/new-images/qn10.png}}
+  \caption{Here we have $Q_n$, for $n \in$ \{3, 7, 10\}.
 Each node represents a possible bitstring in $\{0,1\}^n$, and two nodes are linked if the bitstrings differ by a single dimension.
@@
 %FOOTNOTE:
+The graph composed of $\{0,1\}^n$ nodes and links between nodes $iff$ their Hamming distance is one is called the \emph{hypercube graph}, or $Q_n$.
+It is a beautiful combinatorial object on its own for which one can find a survey in \citet{harary1988survey}.
 
@@
 In other words… though this would break Kanerva's assumption of a uniformly distributed set of hard locations for a perfectly symmetrical set of hard locations, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
-
-%  (See \cite{@article{harary1988survey,
-%    title={A survey of the theory of hypercube graphs},
-%    author={Harary, Frank and Hayes, John P and Wu, Horng-Jyh},
-%    journal={Computers \& Mathematics with Applications},
-%    volume={15},
-%    number={4},
-%    pages={277--289},
-%    year={1988},
-%    publisher={Elsevier}
-%  }}
 

### 2018-03-02 Chapters/Chapter02.tex
sdm

@@
 
-Future work: classification with context using sequences --- for words instead of only letters.
+\section{Future work}
+
+Here are some ideas that have been considered during this work, but have had to be left for future research.
+
+\subsection{Multiple levels}
+
+\subsection{Classification with context using sequences --- for words instead of only letters}
+
+\subsection{Symmetrical Hard Locations}
+
+Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
+Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
+In other words… though this would break Kanerva's assumption of a uniformly distributed set of hard locations for a perfectly symmetrical set of hard locations, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
 
@@
 The simulations run in this thesis are promptly available to be re-executed and explored.
-We invite the reader to take a look and explore a little bit.
+We invite readers to take a look and explore a little bit.
 

### 2018-03-02 Chapters/sdm2.tex
sdm

@@
 First, the number of nodes grows as $2^n$ as $n$ grows; which makes use of the entire space infeasible as $n>>20$.
-Another interesting observation, better seen in the figures below, is that most of the space lies at the center, at a distance of around 500 from any given vantage point.
-Though Kanerva has derived many combinatorial properties of the space, I believe that this is a aesthetically appealing object on its own, and beautiful results can be found in the graph-theoretical literature.\label{hypercubes}}
+Another interesting observation, better seen in the figures below, is that most of the space lies at the center, at a distance of around 500 from any given vantage point.\label{hypercubes}}
 \end{figure}
 
-%FOOTNOTE:
-The graph composed of $\{0,1\}^n$ nodes and links between nodes $iff$ their Hamming distance is one is called the \emph{hypercube graph}, or $Q_n$.
-It is a beautiful combinatorial object on its own for which one can find a survey in \citet{harary1988survey}.
+The graph composed of $\{0,1\}^n$ nodes and links between nodes $iff$ their Hamming distance is one is called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}.
+Though Kanerva has derived many combinatorial properties of the space, I believe that this is a aesthetically appealing object on its own, and beautiful results can be found in the graph-theoretical literature.
+A good survey is found in \citet{harary1988survey}.
 
-Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions, as in Figure \ref{hypercubes}.
+Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
 Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it belongs to?

### 2018-03-04 Chapters/sdm2.tex
Future research (triangulation of Qn)

@@
 A number of observations can be made here.
-First, the number of nodes grows as $2^n$ as $n$ grows; which makes use of the entire space infeasible as $n>>20$.
-Another interesting observation, better seen in the figures below, is that most of the space lies at the center, at a distance of around 500 from any given vantage point.\label{hypercubes}}
+First, the number of nodes grows as $2^n$ as $n$ grows; which makes the  space intractable as $n>>20$.
+Another interesting observation, better seen in the figures below, is that most of the space lies `at the center', at a distance of around 500 from any given vantage point.\label{hypercubes}}
 \end{figure}
@@
 A good survey is found in \citet{harary1988survey}.
-
-Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
-Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it belongs to?
-In other words… though this would break Kanerva's assumption of a uniformly distributed set of hard locations for a perfectly symmetrical set of hard locations, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
 

### 2018-03-05 Chapters/sdm2.tex
Parts under work

@@
   \caption{Here we have $Q_n$, for $n \in$ \{3, 7, 10\}.
-Each node represents a possible bitstring in $\{0,1\}^n$, and two nodes are linked if the bitstrings differ by a single dimension.
+Each node corresponds to a bitstring in $\{0,1\}^n$, and two nodes are linked iff the bitstrings differ by a single dimension.
 A number of observations can be made here.
@@
 
-The graph composed of $\{0,1\}^n$ nodes and links between nodes $iff$ their Hamming distance is one is called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}.
-Though Kanerva has derived many combinatorial properties of the space, I believe that this is a aesthetically appealing object on its own, and beautiful results can be found in the graph-theoretical literature.
-A good survey is found in \citet{harary1988survey}.
+The space studied by Kanerva is also called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}.
+For a fixed $n \in \mathbb{Z}$, the graph $G = (V, E)$, in which $v \in V$ iff there is a bijective function $b: V\to \{0,1\}^n$ and $(v_i, v_j) \in E$ iff $H(b(v_i), b(v_j))=1$, where $H$ is the Hamming distance.
+That is, $n$-sized bitstrings correspond to nodes and edges exist between nodes $iff$ they flip a single bit.
+Though Kanerva has derived many combinatorial properties of the space, additional results have been found in the graph-theoretical literature.
+A good survey is provided by \citet{harary1988survey}.
 

### 2018-03-05 Chapters/Chapter02.tex
Parts under work
SDM introduction

@@
 Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+Such model may be applied in all sort of applications because it would replicate human capacity to remember past experiences from clues of the present.
+For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where ones fear come from.
+They just feel it.
+We may interpret this situation as clues of the present --- a dark alley --- recalling past experiences from memory and thus generating the scared feeling.
+Our memory is able to make a parallel between previous experiences and the clues.
+Although one has never been in the exactly same situation, ones brain makes an analogy and recognizes the danger.
+This flexibility into mapping one situation in another is an important human feature which is hard to replicate into computers.
+
 It has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
+\cite{Linhares2011} has showed that SDM respects the limits of short-term memory discussed by \citet{Miller1995} and \citet{Cowan2011}.
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
@@
 
-It is our belief that such a tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python.
+It is our belief that such a tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python in an intuitive platform such as juypyter notebook.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
-This new tool provides a ready to use system in which experiments can be executed almost as soon as they are designed.
+This new tool provides a ready to use system in which experiments can be executed almost as soon as they are designed and it may accelerate researches \citep{shen2014interactive}.
 
@@
 
-The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has automatic tests, (iv) is cross-platform, and (v) is easily extended to fulfill other research models.
+Extensions of SDM has been used in many applications.
+For example, \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
+\citet{Rajesh1998} used a modified SDM in an autonomous robot.
+\citet{Meng2009} modified SDM to clean patterns from noisy inputs.
+\citet{fan1997genetic} extended SDM with genetic algorithms.
+\citet{chada2016you} extended SDM creating the Rotational Sparse Distributed Memory (RSDM), which is used to modeling network motifs, dynamic flexibility, and hierarchical organization, all results from neuroscience literature.
+
+The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extended to fulfill other research models.
 Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier for readers to replicate any previous results and easily understand the source-code of the model.
-Moreover, new machanisms and applications are also introduced.
+Moreover, it is compatible with jupyter notebook and researchers may share their notebooks possibly accelerating the advances in their fields \citep{shen2014interactive}.
 
-SDM, extensions of it, has been used in many applications.
-For example, \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
-\citet{Rajesh1998} used SDM in an autonomous robot.
-\citet{Meng2009} modified SDM to clean patterns from noisy inputs.
-\citet{Linhares2011} showed that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
+Other contributions has also been introduced, which include (i) a noise filtering approach, (ii) a supervised classification algorithm, (iii) and a reiforcement learning algorithm, all of them using only the original SDM proposed by Kanerva, i.e., with no addons.
+Although our applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them has done it adapting SDM to better fit their problems.
+We have presented different approaches with no adaptations at all.
+
+Finally, we have found an anomaly in one of Kanerva's prediction, which we believe is related to SDM capacity.
+We have also tested a generic reading operation proposed by professor Paulo Murilo (personal communication).
 

### 2018-03-05 Chapters/sdm-applications.tex
SDM introduction

--- 2018-02-28 Chapters/sdm-applications.tex
@@
 The chunk was calculated using the exclusive or (XOR) operator.
-So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- \pyth{write(addr=bs_image ^ bs_label, datum=bs_image)}.
+So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- write(addr=bs\_image XOR bs\_label, datum=bs\_image).
 

### 2018-03-05 Chapters/sdm2.tex
Review of the SDM description

@@
 For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
-One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance.
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance in 3 dimensions, i.e., both follow triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which Euclidean distance in 3 dimensions may be interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{and} d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
+Hence, it makes no sense to say that $A$ is also close to $C$.
+This difference in intuition may trick even experienced researchers when analyzing some situations.
 
@@
 
-The main change between Kanerva's read and Chada's read is that, in the former, a hard-location that has more bitstrings written has a greater weight in the decision of each bit.
+The main difference between Kanerva's read and Chada's is that, in the former, a hard-location that has more bitstrings written has a greater weight in the decision of each bit.
 In the latter, all hard-locations have the same weight, because they can contribute to the sum with only one bitstring.
+
+It is important to say that Chada's read came from \citet{anwar2003sparse} which gave a misguided description of the read operation. The original description is the following:
+
+\begin{samepage}
+\begin{quote}
+Similar to writing, retrieving from SDM involves the same concept of access sphere --- you read all the hardlocations within the access sphere of location Y, pool the bit vectors read from all these hard locations and let each of the kth bits of those locations participate in a majority vote for the kth bit of Y.
+\hfill --- \citet{anwar2003sparse}
+\end{quote}
+\end{samepage}
+
+This fact just highlihts how important it is to have a reference implementation that one may read the code to clarify one's understanding about the details of each operation.
+
+\subsection{Generalized read operation}
 
@@
 Hence, we will here explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.
+Mathematically, let $A$ be the set of the counters of the activated hardlocation, and $c_i$ be the counter of the $i$-th bit.
+Then,
+
+$$
+s_i = \sum_{c \in A} \frac{c_i}{|c_i|} |c_i|^z
+$$
+
+Finally, the $i$-th bit of the resulting bitstring is 1 if $s_i > 0$, or 0 if $s_i < 0$, or random if $s_i = 0$.
+Notice that when $z=1$, then $s_i = \sum_{c \in A} c_i$, which is the Kanerva's read; and when $z=0$, then $s_i = \frac{c_i}{|c_i|} = \text{sign}(c_i)$, which is the Chada's read.
 
@@
 
-In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 10 million hard-locations, an access-radius of 451 (within 1000 hard-locations in each circle) and 10 thousand writes of random bitstrings in the memory.
+In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 1 million hard-locations $N'=1,000,000$, an access-radius of 451 (within 1,000 hard-locations in each circle) and 10 thousand writes of random bitstrings in the memory.
 As computer resources were very poor those days, Kanerva couldn't make a more generic analysis.
@@
 Figure \ref{fig:crit-dist-10k-writes} compares the critical distance behavior under different scenarios.
-This replicates our previous results \citet{Brogliato2011} and \citet{brogliato2014sparse} and is a first part of the process of framework validation, to which we throw our attention next.
+This replicates our previous results \citep{Brogliato2011, brogliato2014sparse} and is a first part of the process of framework validation, to which we throw our attention next.

### 2018-03-05 Chapters/Chapter02.tex
info-theory

@@
 
-My advisor, Alexandre Linhares, has proposed another read operation: the weighted reading.
-In it, the sum of the counter's value is weighted based on the distance between the hard-location's address and the reading address.
+My advisor, Alexandre Linhares, has proposed another read operation: an information-theoretical weighted reading.
+In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address.
 The logic behind it is to vary the importance of each hard-location inside the circle.
@@
 Closer hard locations have lower probabilies and therefore should encode more information.
-We know that the probability of a distance $d$ is $\binom{N}{d}/2^N$, and, acordingly, we have tested how SDM behaves when write weights are calculated using the amount of information given by each bitstring, i.e., $w(d) = -\log_2 \left( 2^{-N} \binom{N}{d} \right) = N - \log_2 \binom{N}{d}$, as in information theory \citep{cover2012elements}.
+
+Consider the following. Given any two $n$-sized bitstrings, the probability of a Hamming distance $d$ is given by,
+
+%\begin{align*}
+
+$p(H=d)=\frac{ \binom{n}{d} }{ 2^n },$ \\
+
+$p(H\leq d)={\frac{1}{2^n}}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}  }$, \\
+
+%\end{align}
+
+and, consequently,
+
+$p(H\geq n-d)={\frac{1}{2^n}}{\displaystyle\sum_{i=n-d}^{n}{\binom{n}{i}}  }$, \\
+$p(d+1 \leq H \leq n-d-1)=2^n - {\frac{1}{2^{n-1}}}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}  .}$ \\
+
+accordingly, we have tested how SDM behaves when write weights are calculated using the amount of information given by each bitstring, i.e., $w(d) = -\log_2 \left( 2^{-n} \binom{n}{d} \right) = n - \log_2 \binom{n}{d}$, as in information theory \citep{cover2012elements}.
 
@@
 Note that 10 additional bits imply an attractor $2^{10}$ of the size of the original.
-Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have large asymmetries.
+Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
 

### 2018-03-05 Chapters/Chapter02.tex
SDM Notation and minor changes.

@@
 
+\chapter{Notation}
+
+\begin{tabular}{cp{\textwidth}}
+  $n$ & Number of dimensions, i.e., $n=1,000$ \\
+  $N$ & Length of the binary space --- $|\{0, 1\}^n| = 2^n$ \\
+  $N'$ & Number of hard-locations samples from $\{0, 1\}^n$ \\
+  $H$ & The same as $N'$ \\
+  $\eta$ & A bitstring \\
+  $\eta_x$ & A clue $x$ bits away from $\eta$, i.e., $\text{dist}(\eta, \eta_x) = x$ \\
+  $\xi$ & A bitstring \\
+  $\text{dist}(x, y)$ & Hamming distance between $x$ and $y$
+\end{tabular}\\
+
 \chapter{Sparse Distributed Memory}
@@
 Thus, the addresses (bitstrings) of the hard-locations are stored in one array, while their counters in another.
-This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be same in all of them.
-As the slowest part of reading and writing operations is scanning the address space, the performance benefits are high.
+This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be the same in all of them.
+As the slowest part of reading and writing operations is scanning the address space, the performance benefits are significant.
 
@@
 The RAM memory is interesting for quick experiments, automated tests, and others scenarios in which the SDM may be lost, while the file is interesting for a long-term SDM, like creating an SDM file with 10,000 random writes, which will be copied over and over to run multiple experiments.
-The file may also be sent to another researcher or may be published with the paper to let others run their own checks and verify the results.
+The file may also be sent to another researcher or may be published within the paper to let others run their own checks and verify the results.
 In summary, the framework fits many different uses and necessities.

### 2018-03-05 Chapters/Chapter02.tex
SDM

@@
   $N$ & Length of the binary space --- $|\{0, 1\}^n| = 2^n$ \\
-  $N'$ & Number of hard-locations samples from $\{0, 1\}^n$ \\
+  $N'$ & Number of hard-locations samples from $\{0, 1\}^n$. It typical value is 1,000,000. \\
   $H$ & The same as $N'$ \\
+  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, it typical value is $451$. This value is calculated to activate, on average, one thousand of $N'$. \\
   $\eta$ & A bitstring \\
@@
 The remaining 24 bits are always zero, so they do not affect the result of any operation.
+The memory usage efficiency is $1 - 24/1024 = 97.65\%$.
 Bitstrings store neither how many bits they have nor the array length.
@@
 The distance between two bitstrings is calculated by the hamming distance, which is the number of different bits between them.
+It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in } x \oplus y$.
 
-It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in } x \oplus y$.
 There are several algorithms to calculate the number of ones \citep{warren2013hacker}, but the performance depends on the processor.
@@
 
-An alternative algorithm available is the lookup.
-It pre-calculates a table of distances --- by default, it is a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM ---, which is accessed a few times.
-Finally, to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$.
+There is also the naive algorithm, which really counts the number of ones checking bit by bit. It is available only to testing purposes and should never be used.
+
+The other algorithm available is the lookup.
+It pre-calculates a table with the number of ones of all possible 16-bit integers.
+This table is accessed a few times to calculate the number of ones of a 64-bit integer, i.e., to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:63], y[0:63]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + d(x[32:47], y[32:47]) + d(x[48:63], y[48:63])$ where $x[i:i+15]$ and $y[i, i+15]$ are the 16-bit integers formed by the bits between $i$ and $i+15$ of $x$ and $y$, respectivelly.
+Each 16-bit distance is calculated through a single table access.
 As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$).
+This table uses 65MB of RAM.
+One may change the table from 16-bit integers to 32-bit integers, which would halve the number of accesses but would use 4GB of RAM (instead of 65MB).
 
@@
 Bitstrings are stored in a contiguous array of 64-bit integers, as shown in Figure \ref{tab:hl-addresses-detail}.
-Hence, basic pointer arithmetic provides us with performance improvements in their access, as processors realize fechts of contiguous chunks of memory  \citep{pai2004linux}.
+Hence, basic pointer arithmetic provides us with performance improvements in their access, as processors realize fetches of contiguous chunks of memory  \citep{pai2004linux}.
 
@@
 
-The scan for activated hard-locations is performed in an address space. It returns the indexes of the bitstrings which were inside the circle. Then, each operation uses this index in a different way.
+The scan for activated hard-locations is performed in an address space.
+It returns the indexes of the bitstrings which were inside the circle (and their distances).
+Then, each operation uses these pieces of information in a different way.
 
@@
 Scanning for the activated hard-locations is a problem similar to well-known problems in computational geometry called ``range reporting in higher dimensions''.
-In this case, none of the known algorithms is able to solve our problem faster than $O(n)$.
-The algorithm which seems to best fit in our problem consumes $O(n)$ space and runs in $O(\log^d(n))$ \citep{chazelle1988functional}, which is really slower than $O(n)$ when, for instance, $n=1,000,000$ and $d=1,000$.
-For a review of the algorithms, see \citet{chan2011orthogonal}.
+In this case, none of the known algorithms is able to solve our problem faster than $O(H)$.
+The algorithm which seems to best fit in our problem consumes $O(H)$ space and runs in $O(\log^n(H))$ \citep{chazelle1988functional}, which is really slower than $O(H)$ when, for instance, $H=1,000,000$ and $n=1,000$.
+For a review of the range reporting algorithms, see \citet{chan2011orthogonal}.
 
 In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}.
-It provides a fast search when $r/b < 0.11$ or $r/b < 0.06$, where $r$ is the radius and $b$ is the number of bits.
-But, in our case, for a 1,000 bits SDM, $r/b = 0.451$, which changes the runtime to $O(n^{0.993})$.
-This is really close to $O(n)$, but with a larger constant.
-Hence, $O(n)$ is still faster.
+It provides a fast search when $r/n < 0.11$ or $r/n < 0.06$, where $r$ is the radius and $n$ is the number of bits.
+But, in our case, for a 1,000 bits SDM, $r/n = 0.451$, which changes the runtime to $O(H^{0.993})$.
+This is really close to $O(H)$, but with a larger constant.
+Unfortunately, $O(H)$ is still faster.
 
 It is intriguing that none of those algorithms is able to solve our scanning problem.
-The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space.
-But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the $\{0, 1\}^d$ space.
-So, each addresses' bit itself splits the hardlocations in half, but it does not split the search space in half since both halves still must be convered by the algorithm.
-For instance, let's say we have 1,000 dimensions with 1,000,000 hardlocations, and we are scanning within a circle with radius 451, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
-The search space has not been split in half because both halves have been covered (and one of them should have been skipped).
+The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(H))$ to go through the whole space.
+But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hard-locations' addresses are randomly sampled from the $\{0, 1\}^n$ space.
+Although each addresses' bit itself splits the hardlocations in half, it does not split the search space in half since both halves still must be convered by the algorithm.
+For instance, let's say we have $n=1,000$ dimensions with $H=1,000,000$ hard-locations, and we are scanning within a circle with radius $r=451$, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
+Hence, the search space has not been split in half because both halves have been covered (and one of them should have been skipped).
 
-Hence, as our best approach is to scan through all hardlocations, we may distribute the scan into many tasks which will be executed independently.
+Finally, as our best approach is to scan through all hard-locations, we may distribute the scan into many tasks which will be executed independently.
 The tasks may be executed in different processes, threads, or even computers.
@@
 
-The OpenCL must be initialized, which just copies the address space's bitstrings to the GPU's memory. Then, many scans may be executed with no necessity to upload the bitstrings again.
+The OpenCL must be initialized, which just copies the address space's bitstrings to the GPU's memory.
+Then, many scans may be executed with no necessity to upload the bitstrings again.
+The OpenCL scanner supports running into multiple devices.
 
@@
 
-Each hard-location has one integer of data per bit. For instance, each hard-location of a 1,000 bits SDM has 1,000 bits. Those integers are stores in a counter.
+Each hard-location has one integer of data per bit. For instance, each hard-location of a 1,000 bits SDM has 1,000 bits. Those integers are stored in a counter.
 
-A counter is an array of integers which stores the data of all hard-locations. So, the counter's array has $N \cdot H$ integers.
+A counter is an array of integers which stores the data of all hard-locations. So, the counter's array has $n \cdot H$ integers.
 
 When two counters are added in a third counter, there may occur an overflow.
-Counters may have overflow protection depending on compiling options.
-By default, there is no overflow check for performance reasons.
-It is not supposed to be a problem because each counter is a signed 32-bit integer which can store any number between -2,147,483,648 and 2,147,483,647 (and they will not overflow with less than $2^{31}-1$ divided by the number of activated hard-locations).
-For instance, when $N=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes to be possible to overflow a counter.
-Note also that it would be improbable for a hard location dimension to be this biased before the complete saturation of the memory.
+It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with less writes than $2^{31}-1$ divided by the average number of activated hard-locations.
+For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.
+Note also that it would be more likely to saturate the memory before any overflow.
+
+Anyway, counters may have overflow protection depending on compiling options. By default, there is no overflow check for performance reasons (and because it does not seem necessary).
 
@@
 
-The framework is available both Kanerva's read and Chada's read.
-The latter was implemented according to a generalization proposed by physics professor Paulo Murilo of UFF during the discussion of my master's defense.
-The generalization brings a parameter $z$, which is the exponent of the counter.
+The framework has both Kanerva's read and the generalized read.
+The latter was implemented according to the generalization proposed by professor Paulo Murilo.
+The generalization brings a parameter $z$, which is the exponent.
 In this case, the results are floating point instead of integer, which considerably reduces performance.
@@
 
-There is also another special read operation: the weighted reading.
-The weight is retrieved from a lookup table of integers based on the distance between the hard-location's address and the reading address.
-The weight is multiplied by the counter's value and summed.
+There is another special read operation: the weighted reading.
+In the weighted reading, the value of the counters are multiplied by a weight which depends only on the distance between the reading address and the hard-location address.
+The weight is retrieved from a lookup table of integers indexed by the distance.
 The rest of the read operation is exactly the same.
+
+There is also a weighted writing operation.
+In this case, the weight is applied when the counters are updated, i.e., if the weight is 2, the counters are increased twice when bits are 1, and decreased twice when bits are 0.
+Just as in the weighted reading, the weights depend only on the distance between the writing address and the hard-location address.
+The weights are retrieved from a lookup table of integers indexed by the distance.
 

### 2018-03-05 Chapters/sdm2.tex
SDM

@@
 \begin{quote}
-Similar to writing, retrieving from SDM involves the same concept of access sphere --- you read all the hardlocations within the access sphere of location Y, pool the bit vectors read from all these hard locations and let each of the kth bits of those locations participate in a majority vote for the kth bit of Y.
+%Similar to writing, retrieving from SDM involves the same concept of access sphere --- you read all the hardlocations within the access sphere of location Y, pool the bit vectors read from all these hard locations and let each of the kth bits of those locations participate in a majority vote for the kth bit of Y.
 \hfill --- \citet{anwar2003sparse}
+With our datum distributively stored, the next question is how to retrieve it.
+With this in mind, let us ask first how one reads from a single hard location, $x$.
+Compute $\zeta$, the bit vector read at $x$, by assigning its $i$th bit the value 1 or 0 according as the $i$th counter at $x$ is positive or negative.
+Thus, each bit of $\zeta$ results from a majority rule decision of all the data that have been written at $x$.
+[...] Knowing how to read from a hard location allows us to read from any of the $2^{1000}$ arbitrary locations.
+Suppose $\zeta$ is any location.
+The bit vector, $\xi$, to be read at $\zeta$, [...] Put another way, pool the bit vectors read from hard locations accessible from $\zeta$, and let each of their $i$th bits vote on the $i$th bit of $\xi$.
+\\
+\hfill --- \citet[p.342]{anwar2003sparse}
 \end{quote}
@@
 
-A member of my Master's committee, professor Paulo Murilo, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
+A member of my Master's committee, professor Paulo Murilo\footnote{Universidade Federal Fluminense's physicist doctor Paulo Murilo}, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
 He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative).

### 2018-03-05 Chapters/Chapter02.tex
Info-Theory stuff

--- 2018-03-05 Chapters/Chapter02.tex
@@
 
-\chapter{Results (vi): Information theoretical write operation}
+\chapter{Results (vi): Information-theoretical write operation}
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}
+
+  \subfloat[$w(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}
+
+  \subfloat[stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}
+
+  \caption{Computing the amount of information of a signal to each hard location in its access radius.
+(a) entirety of the space; (b) region of interest; (c) Fast computation is possible through a stepwise function.
+}
+  \label{fig:info-theory-hypothesis}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}
+
+  \subfloat[$w(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}
+
+  \subfloat[stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}
+
+  \caption{Computing the sum of low-likelihood signals. (a) entirety of the space; (b) region of interest; (c) Fast computation through a stepwise function. }
+  \label{fig:info-theory-sum-hypothesis}
+\end{figure}
 
@@
 
-Consider the following. Given any two $n$-sized bitstrings, the probability of a Hamming distance $d$ is given by,
+Consider the following.
+Information Theory \citep{cover2012elements} let us compute the precise amount of information in an event, when given its probability $p$, through the measure of \emph{self-information}:\\
 
-%\begin{align*}
+$I(p)= -log_2(p).$ \\
 
-$p(H=d)=\frac{ \binom{n}{d} }{ 2^n },$ \\
+Now, given any two $n$-sized bitstrings, the probability of their Hamming distance being $d$ is given by,
 
-$p(H\leq d)={\frac{1}{2^n}}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}  }$, \\
+$p(H=d)= {2^{-n} \binom{n}{d} } ,$ \\
 
-%\end{align}
+$p(H\leq d)= 2^{-n} {\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}}  $, \\
 
@@
 
-$p(H\geq n-d)={\frac{1}{2^n}}{\displaystyle\sum_{i=n-d}^{n}{\binom{n}{i}}  }$, \\
-$p(d+1 \leq H \leq n-d-1)=2^n - {\frac{1}{2^{n-1}}}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}  .}$ \\
+$p(H\geq n-d)=2^{-n}{\displaystyle\sum_{i=n-d}^{n}{\binom{n}{i}}  }$, \\
 
-accordingly, we have tested how SDM behaves when write weights are calculated using the amount of information given by each bitstring, i.e., $w(d) = -\log_2 \left( 2^{-n} \binom{n}{d} \right) = n - \log_2 \binom{n}{d}$, as in information theory \citep{cover2012elements}.
+$p(d+1 \leq H \leq n-d-1)=2^n - 2^{1-n}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}, \forall d<n/2}$. \\
+
+Hence the weighted write would, on each hard location, sum (or subtract) the following:  \\
+
+$w(d) = -\log_2 \left( 2^{-n} \binom{n}{d} \right) = n - \log_2 \binom{n}{d}$, as seen in Figure \ref{fig:info-theory-hypothesis}.
+
+It is easy to interpret this data though a binary tree approach.  How many binary questions would be needed to precisely define a bitstring?
+
+  INTERPRET PLUS FIG
+
+Another possibility would be to use the sum of all distances closer (and less likely) locations within $w(d)$,
+
+$w(d) = -\log_2 \left( 2^{-n} \displaystyle\sum_{i=0}^{d}{\binom{n}{i}} \right) = n - \log_2 \displaystyle\sum_{i=0}^{d}{\binom{n}{i}}$, \\
+
+This can be seen in \ref{fig:info-theory-sum-hypothesis}.
 

### 2018-03-05 Chapters/sdm2.tex

@@
 
-A member of my Master's committee, professor Paulo Murilo\footnote{Universidade Federal Fluminense's physicist doctor Paulo Murilo}, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
+A member of my Master's committee, professor Paulo Murilo, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
 He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative).

### 2018-03-05 Chapters/Chapter02.tex
SDM update notation

@@
 \begin{tabular}{cp{\textwidth}}
-  $n$ & Number of dimensions, i.e., $n=1,000$ \\
-  $N$ & Length of the binary space --- $|\{0, 1\}^n| = 2^n$ \\
-  $N'$ & Number of hard-locations samples from $\{0, 1\}^n$. It typical value is 1,000,000. \\
-  $H$ & The same as $N'$ \\
-  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, it typical value is $451$. This value is calculated to activate, on average, one thousand of $N'$. \\
-  $\eta$ & A bitstring \\
-  $\eta_x$ & A clue $x$ bits away from $\eta$, i.e., $\text{dist}(\eta, \eta_x) = x$ \\
-  $\xi$ & A bitstring \\
-  $\text{dist}(x, y)$ & Hamming distance between $x$ and $y$
+  $n$ & Number of dimensions, i.e., $n=1,000$. \\
+  $N$ & Length of the binary space --- $|\{0, 1\}^n| = 2^n$. \\
+  $N'$ & Number of hard-locations samples from $\{0, 1\}^n$. Its typical value is 1,000,000, as suggested by \citet{Kanerva1988}. \\
+  $H$ & Same as $N'$. \\
+  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, its typical value is $451$. This value is calculated to activate, on average, one thousand of $N'$. \\
+  $\eta$ & A bitstring, usually a datum. \\
+  $\eta_x$ & A clue $x$ bits away from $\eta$, i.e., $\text{dist}(\eta, \eta_x) = x$. \\
+  $\xi$ & A bitstring, usually an address. \\
+  $\text{dist}(x, y)$ & Hamming distance between $x$ and $y$ \\
+  $\text{d}(x, y)$ & Same as $\text{dist}(x, y)$
 \end{tabular}\\

### 2018-03-05 Chapters/Chapter02.tex
sdm + bibtex

--- 2018-03-05 Chapters/Chapter02.tex
@@
 
-$p(H=d)= {2^{-n} \binom{n}{d} } ,$ \\
+$p(H=d)= {2^{-n} \binom{n}{d} }$ \\
+
+And the probability of it being at most $d$ is \\
 
@@
 
-  INTERPRET PLUS FIG
+Another possibility would be to use the sum of all distances closer (and less likely) locations within the weighting function $w(d)$,
 
-Another possibility would be to use the sum of all distances closer (and less likely) locations within $w(d)$,
-
-$w(d) = -\log_2 \left( 2^{-n} \displaystyle\sum_{i=0}^{d}{\binom{n}{i}} \right) = n - \log_2 \displaystyle\sum_{i=0}^{d}{\binom{n}{i}}$, \\
+$w(d) = -\log_2 \left( 2^{-n} \displaystyle\sum_{i=0}^{d}{\binom{n}{i}} \right) = n - \log_2 \displaystyle\sum_{i=0}^{d}{\binom{n}{i}}$. \\
 
@@
 Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
+
+Note, finally, that this is not the first time in which a weighted function has been applied to writing in SDM --- \citet{hely1997new} suggest a rather complex spreading model based on floating point signals in the interval [0.05, 1.0] --- they were, however, only able to test their model with 1,000 hard locations.
 
@@
 
-Here are some ideas that have been considered during this work, but have had to be left for future research.
+Here is an interesting questions that have been considered during this work, but have had to be left for future research.
 
@@
 
+\subsection{i versus l}
+
+\subsection{Magic numbers}
+
+Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.  More recently, he suggested the use of 10,000 dimensions.
+
+Each parameter set choice like this will lead to particular numbers --- many of them emergent---, such as the access radius size, critical distance, and so forth.
+
+One intriguing question here is:  is there a `better' number of dimensions and of hard locations?  If so, can such numbers better studied analitically, or numerically?
+
+How should these parameters be compared?  What are the tradeoffs that should be considered?  What are the `best' benchmarks possible?
+
 \subsection{Classification with context using sequences --- for words instead of only letters}
 
-\subsection{Symmetrical Hard Locations}
+\subsection{Symmetrical, rapidly accessible, Hard Locations}
 
-Here is an interesting question that I leave for further research: A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
+A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
 Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
-In other words… though this would break Kanerva's assumption of a uniformly distributed set of hard locations for a perfectly symmetrical set of hard locations, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
+Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
+
+$\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
+
+\subsection{Docker image and jupyter notebooks}
 
@@
 We invite readers to take a look and explore a little bit.
+
+\subsection{From theory to a platform.}
+
+Platforms in the history of computing.
+
+Opens 10 doors.
 

### 2018-03-05 Chapters/Chapter02.tex
SDM distance histogram

--- 2018-03-05 Chapters/Chapter02.tex
@@
 
-The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
-This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
+\section{Distance between random bitstrings}
+
+As predicted by Kanerva, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
+For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
+
+In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with 1,000 bits.
+In total, 20,000 bitstrings have been generated.
+The code is available in the ``Distance between bitstrings'' notebook.
+
+In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
+  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}
+
+  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
+  \label{fig:validation-distance}
+\end{figure}
 
@@
 
+The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
+This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
+
 % TODO Include in the to do list simulations based on the suggestion of Murilinho.

### 2018-03-05 Chapters/Chapter02.tex

--- 2018-03-05 Chapters/Chapter02.tex
@@
 
+The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
+This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
+
 \subsection{From theory to a platform.}

### 2018-03-05 Chapters/Chapter02.tex
SDM

@@
 
-As predicted by Kanerva, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
+As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
 For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
 
-In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with 1,000 bits.
-In total, 20,000 bitstrings have been generated.
-The code is available in the ``Distance between bitstrings'' notebook.
+In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits.
+In total, 20,000 random bitstrings have been generated during the simulation.
+The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.
 
@@
 
-\begin{figure}[h!]
+This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.
+
+\begin{figure}[!htb]
   \centering

### 2018-03-06 Chapters/Chapter02.tex
sdm

--- 2018-03-05 Chapters/Chapter02.tex
@@
 
-Other contributions has also been introduced, which include (i) a noise filtering approach, (ii) a supervised classification algorithm, (iii) and a reiforcement learning algorithm, all of them using only the original SDM proposed by Kanerva, i.e., with no addons.
-Although our applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them has done it adapting SDM to better fit their problems.
+Other contributions have also been introduced, which include (i) a noise filtering approach, (ii) a supervised classification algorithm, (iii) and a reinforcement learning algorithm, all of them using only the original SDM proposed by Kanerva, i.e., with no additional mechanisms, algorithms, data structures, etc.
+Although some of our applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them have done some adapting of SDM to their problems, and none of them have used just the ideas introduced by Kanerva.
 We have presented different approaches with no adaptations at all.
@@
   $n$ & Number of dimensions, i.e., $n=1,000$. \\
-  $N$ & Length of the binary space --- $|\{0, 1\}^n| = 2^n$. \\
+  $N$ & Size of the binary space --- $|\{0, 1\}^n| = 2^n$. \\
   $N'$ & Number of hard-locations samples from $\{0, 1\}^n$. Its typical value is 1,000,000, as suggested by \citet{Kanerva1988}. \\
@@
 So, we have implemented three different algorithms and one may be selected through compiling flags.
-The default algorithm is to use a builtin instruction from the compiler.
+The default algorithm is to use a built-in \_\_popcnt() instruction from the compiler.
 
@@
 This table uses 65MB of RAM.
-One may change the table from 16-bit integers to 32-bit integers, which would halve the number of accesses but would use 4GB of RAM (instead of 65MB).
+One may change the table from 16-bit integers to 32-bit integers, which would halve the number of accesses at the expense of 4GB of RAM (instead of 65MB).
 
@@
 
-\section{Distance between random bitstrings}
-
-As predicted by Kanerva, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
-For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
-
-In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with 1,000 bits.
-In total, 20,000 bitstrings have been generated.
-The code is available in the ``Distance between bitstrings'' notebook.
-
-In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
-  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}
-
-  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
-  \label{fig:validation-distance}
-\end{figure}
+The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
+This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
 
@@
 
-Here is an interesting questions that have been considered during this work, but have had to be left for future research.
+Here are interesting questions that have been considered during this work, but have had to be left for future research.
 
@@
 
-The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
-This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
-
 \subsection{From theory to a platform.}

### 2018-03-06 Chapters/sdm2.tex
sdm

--- 2018-03-05 Chapters/sdm2.tex
@@
 For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
-One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance in 3 dimensions, i.e., both follow triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which Euclidean distance in 3 dimensions may be interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{and} d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance in 3 dimensions, i.e., both follow triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which Euclidean distance in 3 dimensions may be interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{ and } d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
 Hence, it makes no sense to say that $A$ is also close to $C$.

### 2018-03-06 Chapters/sdm-applications.tex

@@
 In the 0-fold SDM, every bitstring is written to its own address.
-In the 1-fold SDM, every bitstring points somewhere else.
+In the 1-fold SDM, every bitstring points to another one.
 So, the transitions are written in the 1-fold SDM, while the boards themselves are written to the 0-fold SDM.
@@
 If it does not converge to any board, than SDM will choose a random movement and learn from it.
+Eventually it may converge to a board which is not a sequence of the current board.
+In this case, SDM will also choose a random movement.
 
@@
 
-It is an unsupervised algorithm because SDM learns playing against an opponent, who may be another SDM player, a human, or a player whose movements are always aleatory.
+It is an unsupervised algorithm because SDM learns playing against an opponent, who may be another SDM player, a human, or a player whose movements are always random.
 
 Thus, in order to train a SDM player, we just have to keep it playing over and over.
+
+\subsection{Results}
 

### 2018-03-06 Chapters/Chapter02.tex
SDM Validation

@@
 
-The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
-This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
+\section{Distance between random bitstrings}
+
+As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
+For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
+
+In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits.
+In total, 20,000 random bitstrings have been generated during the simulation.
+The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.
+
+In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
+
+This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
+  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}
+
+  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
+  \label{fig:validation-distance}
+\end{figure}
+
+\section{Number of activated hard-locations}
+
+In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard-locations in a 1,000 bits SDM.
+He also proposed to activate only 1,000 of them, on average.
+He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard-locations.
+
+We extended his results, calculating the distribution of the number of activated hard-locations.
+As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = p(p-1)H$.
+In this case, $\mu = 1071.85$ and $\sigma = 32.72$.
+
+In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard-locations.
+The code is available in the ``Number of activated hard-locations'' notebook \citep{sdmframework}.
+
+In figure \ref{fig:validation-activated-hls}, we can notice that the theoretical model and the simulation matches.
+Hence, it seems that both the address space generation algotihm and the scan algorithm work properly.
+Notice that the curve is almost the same for $n=1,000$ and $n=256$.
+It happens because the access radius is adjust to have $p$ as close as possible to $0.001$.
+They are not exactly the same because their $p$ differs a little.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[$n=1,000$, $H=1,000,000$,\protect\\ $r=451$, and $p=0.00107185$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-1000.png}}
+  \subfloat[$n=256$, $H=1,000,000$,\protect\\ $r=103$, and $p=0.00106684$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-256.png}}
+
+  \caption{Histogram of the number of activated hard-locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
+  \label{fig:validation-activated-hls}
+\end{figure}
+
+\section{Intersection of two circles}
+
+Kanerva has calculated the intersection of two circles according to the distance between their centers.
+The intersection is important to understand how SDM works, because it affects directly the critical distance.
+When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
+In fact, it converges because they share a sufficient amount of hard-locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge.
+For further information about the relation between the critical distance and the intersection, see \citet{brogliato2014sparse}.
+
+We have calculated the intersection between a random bitstring (bs1) and another bitstrings (bs2) exactly $d$ bits away.
+The former (bs1) is just a random bitstring.
+The latter (bs2) was generated randomly flipping $d$ bits of bs1.
+The code is available in the ``Kanerva's Figure 1.2'' notebook \citep{sdmframework}.
+
+In Figure \ref{fig:validation-intersection}, we can notice that we have obtained the same results as Kanerva. It seems that the random flipping bits algorithm and the scan algorithm work properly.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[{\citet[Figure 1.2, p.25]{Kanerva1988}} ]{\includegraphics[width=0.45\textwidth]{./images02/new-images/kanerva-table-12.jpg}}
+  \subfloat[Generated by SDM-Frameworkm with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}
+
+  \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
+  \label{fig:validation-activated-hls}
+\end{figure}
+
+---
 

### 2018-03-06 Chapters/Chapter02.tex
SDM

@@
 We extended his results, calculating the distribution of the number of activated hard-locations.
-As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = p(p-1)H$.
+As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$.
 In this case, $\mu = 1071.85$ and $\sigma = 32.72$.
@@
   \label{fig:validation-activated-hls}
+\end{figure}
+
+Besides the number of activated hard-locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard-locations.
+Let $A$ be the set of activated hard-locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:
+
+\begin{align}
+P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(d \in A)} \\
+    &= \frac{\binom{n}{x}}{\sum_{k=0}^r \binom{n}{k}} \label{eq:prob-d-inside-circle}
+\end{align}
+
+In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard-locations to the center of 1,000 random circles.
+The code is available in the ``Distances of activated hard-locations'' notebook \citep{sdmframework}.
+
+In figure \ref{fig:validation-distance-activated-hls}, we can notice that the theoretical model and the simulation matches.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images02/new-images/distance-activated-hls.png}
+
+\caption{Histogram of the distances of activated hard-locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
+\label{fig:validation-distance-activated-hls}}
 \end{figure}
@@
   \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
-  \label{fig:validation-activated-hls}
+  \label{fig:validation-intersection}
 \end{figure}

### 2018-03-06 Chapters/Chapter02.tex
sdm

--- 2018-03-06 Chapters/Chapter02.tex
@@
 
-The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation.
-This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.
+\section{Distance between random bitstrings}
+
+As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
+For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
+
+In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits.
+In total, 20,000 random bitstrings have been generated during the simulation.
+The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.
+
+In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
+
+This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
+  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}
+
+  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
+  \label{fig:validation-distance}
+\end{figure}
 
@@
 
-\subsection{Some initial anomalous results}
+\section{Some initial anomalous results}
 
@@
 \centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
-\caption{Results generated by the framework similar from Kanerva's original Table 7.2. It was a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
+\caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
 \label{sdm-100w-table-7-2}}
@@
   \centering
-  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_01-05.png}}
-  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_15_3_45_6.png}}
+  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_01-05.png}}
 
-  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_0_1.png}}
-  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_all_2.png}}
+  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_15_3_45_6.png}}
 
-  \caption{(a) and (b) show the behavior of a single read; (c) and (d) present the effects of 6 iterative reads.
-As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
+  \caption{(a) and (b) show the behavior of a single read; As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
 Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
@@
 
+\begin{figure}[h!]
+  \centering
+  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_0_1.png}}
+
+  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_all_2.png}}
+
+  \caption{(a) and (b) show the behavior of the previous Figure, now executed with 6 iterative reads.
+What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
+  \label{fig:murillo-generalization-experiments-6reads}
+\end{figure}
+
 Physicist Paulo Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
@@
 Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads.
-This is shown in Figure \ref{fig:murillo-generalization-experiments}.
+This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
 
@@
   \centering
-  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}
+  \subfloat[$w_1(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}
 
-  \subfloat[$w(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}
+  \subfloat[$w_1(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}
 
-  \subfloat[stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}
+  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}
 
-  \caption{Computing the amount of information of a signal to each hard location in its access radius.
-(a) entirety of the space; (b) region of interest; (c) Fast computation is possible through a stepwise function.
-}
+  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
+(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
   \label{fig:info-theory-hypothesis}
@@
   \centering
-  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}
+  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}
 
-  \subfloat[$w(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}
+  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}
 
-  \subfloat[stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}
+  \subfloat[stepwise $\left \lfloor{w_2(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}
 
-  \caption{Computing the sum of low-likelihood signals. (a) entirety of the space; (b) region of interest; (c) Fast computation through a stepwise function. }
+  \caption{SOON TO BE DEPRECATED.
+Shannon write operation:  Computing the sum of low-likelihood signals.
+(a) entirety of the space; (b) region of interest; (c) Fast integer computation through a stepwise function.
+}
   \label{fig:info-theory-sum-hypothesis}
@@
 
+\begin{figure}[h!]
+  \centering
+  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon.jpeg}}
+
+  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon-region.jpeg}}
+
+  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon-stepwise.jpeg}}
+
+  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
+(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
+  \label{fig:info-theory-hypothesis}
+\end{figure}
+
 My advisor, Alexandre Linhares, has proposed another read operation: an information-theoretical weighted reading.
@@
 The logic behind it is to vary the importance of each hard-location inside the circle.
-It is only natural that one encodes an item in near hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
+It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
 Closer hard locations have lower probabilies and therefore should encode more information.
@@
 
-The results can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
+The initial results of this \emph{Shannon write} operation can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
 It seems that the critical distance increases by a number of bits.
@@
 Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
+
+But, qualitatively, they seem disappointing.  HEATMAPFIGs
 
@@
 
-Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.  More recently, he suggested the use of 10,000 dimensions.
+Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.
+More recently, he suggested the use of 10,000 dimensions, and on personal discussions suggested that this should be a minimum; as he has been concerned in latent semantic analysis and seems to be the proper scale in that application.
 
@@
 $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
+A reduction from measuring the distance to $10^6$ hard locations to a computation of $~ 10^3$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.
 
-\subsection{Docker image and jupyter notebooks}
+At large scales of computing, the very ability to perform some experiments is a function of sheer performance.
+The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
+More on this in my closing words.
+
+\section{Seeing farther}
+
+Let us revisit, in this concluding thoughts, the emphasis employed over speed of computation.
+At first sight, that might seem like a typical objective of efficiency in computer science.
+But we are not only interested in the computer sciece effects here --- the ambition is different.
+More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.
 
@@
 After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.
-The simulations run in this thesis are promptly available to be re-executed and explored.
-We invite readers to take a look and explore a little bit.
+
+All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested. We invite readers to take a look and explore a little bit.
 

### 2018-03-06 Chapters/Chapter02.tex
sdm conclusion

@@
 
+\section{Number of activated hard-locations}
+
+In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard-locations in a 1,000 bits SDM.
+He also proposed to activate only 1,000 of them, on average.
+He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard-locations.
+
+We extended his results, calculating the distribution of the number of activated hard-locations.
+As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$.
+In this case, $\mu = 1071.85$ and $\sigma = 32.72$.
+
+In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard-locations.
+The code is available in the ``Number of activated hard-locations'' notebook \citep{sdmframework}.
+
+In figure \ref{fig:validation-activated-hls}, we can notice that the theoretical model and the simulation matches.
+Hence, it seems that both the address space generation algotihm and the scan algorithm work properly.
+Notice that the curve is almost the same for $n=1,000$ and $n=256$.
+It happens because the access radius is adjust to have $p$ as close as possible to $0.001$.
+They are not exactly the same because their $p$ differs a little.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[$n=1,000$, $H=1,000,000$,\protect\\ $r=451$, and $p=0.00107185$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-1000.png}}
+  \subfloat[$n=256$, $H=1,000,000$,\protect\\ $r=103$, and $p=0.00106684$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-256.png}}
+
+  \caption{Histogram of the number of activated hard-locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
+  \label{fig:validation-activated-hls}
+\end{figure}
+
+Besides the number of activated hard-locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard-locations.
+Let $A$ be the set of activated hard-locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:
+
+\begin{align}
+P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(d \in A)} \\
+    &= \frac{\binom{n}{x}}{\sum_{k=0}^r \binom{n}{k}} \label{eq:prob-d-inside-circle}
+\end{align}
+
+In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard-locations to the center of 1,000 random circles.
+The code is available in the ``Distances of activated hard-locations'' notebook \citep{sdmframework}.
+
+In figure \ref{fig:validation-distance-activated-hls}, we can notice that the theoretical model and the simulation matches.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images02/new-images/distance-activated-hls.png}
+
+\caption{Histogram of the distances of activated hard-locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
+\label{fig:validation-distance-activated-hls}}
+\end{figure}
+
+\section{Intersection of two circles}
+
+Kanerva has calculated the intersection of two circles according to the distance between their centers.
+The intersection is important to understand how SDM works, because it affects directly the critical distance.
+When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
+In fact, it converges because they share a sufficient amount of hard-locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge.
+For further information about the relation between the critical distance and the intersection, see \citet{brogliato2014sparse}.
+
+We have calculated the intersection between a random bitstring (bs1) and another bitstrings (bs2) exactly $d$ bits away.
+The former (bs1) is just a random bitstring.
+The latter (bs2) was generated randomly flipping $d$ bits of bs1.
+The code is available in the ``Kanerva's Figure 1.2'' notebook \citep{sdmframework}.
+
+In Figure \ref{fig:validation-intersection}, we can notice that we have obtained the same results as Kanerva. It seems that the random flipping bits algorithm and the scan algorithm work properly.
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[{\citet[Figure 1.2, p.25]{Kanerva1988}} ]{\includegraphics[width=0.45\textwidth]{./images02/new-images/kanerva-table-12.jpg}}
+  \subfloat[Generated by SDM-Frameworkm with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}
+
+  \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
+  \label{fig:validation-intersection}
+\end{figure}
+
+---
+
 One particular analysis of interest is that of the distance read at a point $\alpha$.
@@
 
-Let us revisit, in this concluding thoughts, the emphasis employed over speed of computation.
+Let us revisit, in these concluding thoughts, the emphasis employed over speed of computation.
 At first sight, that might seem like a typical objective of efficiency in computer science.
-But we are not only interested in the computer sciece effects here --- the ambition is different.
+But we are not only interested in the computer science effects here --- the ambition is different.
 More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.
+
+If
 

### 2018-03-06 Chapters/Chapter02.tex
sdm conclusion

@@
 \begin{align}
-P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(d \in A)} \\
+P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(a \in A)} \\
     &= \frac{\binom{n}{x}}{\sum_{k=0}^r \binom{n}{k}} \label{eq:prob-d-inside-circle}
@@
 
----
+\section{Storage and retrieval of sequences}
+
+\citet[Ch.8]{Kanerva1988} presented an approach to store and retrieve sequences using $k$ different SDMs, namely $\text{sdm}_1$, $\text{sdm}_2$, $dots$, $\text{sdm}_k$.
+
+Let $a_0, a_1, a_2, \dots, a_n$ be a sequence to be stored in a $k$-fold memory.
+So, all pointers of the form $a_i \rightarrow a_{i+k}$ will be written to $\text{sdm}_k$ memory, i.e., in $\text{sdm}_1$, the following pointers will be written: $a_0 \rightarrow a_1$, $a_1 \rightarrow a_2$, $\dots$, $a_{n-1} \rightarrow a_n$; while in $\text{sdm}_2$, the following pointers will be written: $a_0 \rightarrow a_2$, $a_2 \rightarrow a_3$, $\dots$, $a_{n-2} \rightarrow a_n$; and so forth.
+
+We have tested exactly the same example presented in \citet{Kanerva1988}, p.85.
+We wrote two sequences to a $3$-fold memory: $<A, B, C, D>$ and $<E, B, C, F>$.
+Then, after reading the sequences $<A, B, C>$ and $<E, B, C>$, we have obtained $D$ and $F$, respectively.
+
+Each reading operation was performed summing the counters of all activated hard-locations from all three memories.
+For instance, to read the sequence $<A, B, C>$, we have activated the hard-locations around $C$ in $\text{sdm}_1$, we have also activated the hard-locations around $D$ in $\text{sdm}_2$, and, finally, we have also activated the hard-locations around $A$ in $\text{sdm}_3$.
+After summing the counters of all those hard-locations, we evaluate the resulting bitstring just as in the original read operation.
+
+The code is available in the ``Sequences (Kanerva Ch 8)'' notebook \citep{sdmframework}.
+
+The logic behind how it works is that, when reading the sequence $<A, B, C>$, we have $A$ pointing to $D$, while both $B$ and $C$ point to $D$ and $F$.
+Thus, $D$ appears more often than $F$ and ended up being the result.
+
+Hence, as we have replicated the theoretical results from Kanerva, we have one more evidence that our framework works properly.
+
+\subsection{$k$-fold memory using only one SDM}
+
+We have extended Kanerva's ideas to be able to store and retrieve sequences in $k$-fold memories using only one SDM (instead of $k$ SDMs).
+
+Our idea was to create $k$ random bitstrings, one for each fold.
+We have performed writing and reading exactly as Kanerva's original idea, but, instead of writing to $\text{sdm}_k$, we have written $a_{i+k}$ into the address $a_i \oplus \text{tag}_k$, and, instead of reading from $\text{sdm}_k$, we have read from address $a_i \oplus \text{tag}_k$, where $\oplus$ is the exclusive or (XOR) operator.
+
+It worked as if we had splitted SDM into $k$ regions with low intersection between two of them.
+So, as the interference is minimal, they work like independent SDMs.
+The major disadvantage of this approach is that memory capacity may be reached faster.
+
+Splitting the memory into regions may be an interesting strategy to other sorts of problems, mostly the ones which would need many SDMs and, consequently, would use a lot of RAM.
+
+\section{Convergence }
 
@@
 
-\section{Some initial anomalous results}
+\subsection{Some initial anomalous results}
 

### 2018-03-06 Chapters/Chapter02.tex
SDM

@@
 
-\section{Convergence }
+\section{Convergence of $\eta_x$ to $\eta$}
 
-One particular analysis of interest is that of the distance read at a point $\alpha$.
-Suppose an SDM is trying to read an item written at $\alpha$, but the cues received so far lead to a point of distance $d$ from $\alpha$.
-As one reads at $\alpha+d$, a new bitstring $\beta$ is obtained, leading to our question: what is the new distance from $\alpha$ to $\beta$?
-Is it smaller or larger than $d$?
-That, of course, depends on the ratio between $d$ and the number of dimensions of the memory.
-As we have found out, there are some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
+One particular analysis of Kanerva's interest is that of the distance read at a point $\eta_x$.
+Suppose an SDM is trying to read an item written at $\eta$, but the cues received so far lead to a point of distance $x$ from $\eta$.
+As one reads at $\eta_x$, a new bitstring $\beta$ is obtained, leading to Kanerva's question: what is the new distance from $\eta$ to $\beta$?
+Is it smaller or larger than $x$?
+That, of course, depends on the ratio between $x$ and the number of dimensions of the memory.
 
-\subsection{Some initial anomalous results}
-
-As we ran the simulations reflecting some of Kanerva's graphs, one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model, and we propose that this is due to interaction effects between different attractors.
-
-\citet{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Fig.
-\ref{kanerva-table-7-2}).
-The original prediction considered that the read distance would decline when inside the critical distance in increase afterwards, converging to a \textasciitilde 500-bit distance.
-At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance point.
+\citet[p.70]{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Figure \ref{fig:kanerva-figure-7.3}).
+The original prediction considered that the read distance would decline when inside the critical distance and increase afterwards, converging to a \textasciitilde 500-bit distance.
+At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance bitstring.
+He analyzed specifically an SDM with 1,000 bits and 10,000 random bitstrings written into it.
 
@@
 \caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
-\label{kanerva-table-7-2}}
+\label{fig:kanerva-figure-7.3}}
 \end{figure}
 
-Our preliminary results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Fig.
-\ref{sdm-10000w-table-7-2}).
-Our results were obtained using a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, the same used by Kanerva.
+As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
+%and we propose that this is due to interaction effects between different attractors.
+
+We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
+
+In details, we have created a SDM with $n=1,000$, $H=1,000,000$, and $r=451$.
+Then, we have generated 10,000 random bitstrings and written them into the memory.
+Then, we have generated a reference bitstring (bs\_ref) and written it into the memory.
+Then, we have executed the following steps with $x$ from 0 to 1,000: (i) copy bs\_ref into a new bitstring; (ii) randomly flipped $x$ bits of the copy; (iii) read from the memory in the copy address; and (iv) stored the distance between the returned bitstring and bs\_ref.
+Finally, we have plotted Figure \ref{fig:sdm-10000w-table-7-2}.
 
 \begin{figure}[h]
-\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-table-7-2.png}
+\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2.png}
 \caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
 Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
-\label{sdm-10000w-table-7-2}}
+\label{fig:sdm-10000w-table-7-2}}
 \end{figure}
 
+Figure \ref{fig:sdm-10000w-table-7-2} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance.
+So, we have changed the steps to run $k$ reads and store the average new distance.
+We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6-samples}, which has a way lower noise and still holds the divergence.
+
+\begin{figure}[h]
+\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-6-samples.png}
+\caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
+Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
+\label{fig:sdm-10000w-table-7-2-6-samples}}
+\end{figure}
+
+Our results show that the theoretical prediction is not accurate.
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).
+
+Obviously, these small deviations from Kanerva's original theoretical predictions deserve a qualification.
+Kanerva was working in the 1980s and the 1990s, and had no access to the immense computational power that we do today.
+It is no surprise that some small interaction effects should exist as machines allow us to explore the ideas of his monumental work.
+
 But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Fig.
-\ref{sdm-100w-table-7-2}).
+\ref{fig:sdm-100w-table-7-2}).
 This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
@@
 \caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
-\label{sdm-100w-table-7-2}}
+\label{fig:sdm-100w-table-7-2}}
 \end{figure}
 
-To obtain the results from Fig.
-\ref{sdm-10000w-table-7-2} and \ref{sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
+To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
 Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
@@
 
-Obviously, these small deviations from Kanerva's original theoretical predictions deserve a qualification.
-Kanerva was working in the 1980s and the 1990s, and had no access to the immense computational power that we do today.
-It is no surprise that some small interaction effects should exist as machines allow us to explore the ideas of his monumental work.
+\chapter{Results (ii): Loss of neurons}
+
+\chapter{Results (iii): Generalized read operation}
 
@@
 
-Let us revisit, in these concluding thoughts, the emphasis employed over speed of computation.
+Let us revisit, in this concluding thoughts, the emphasis employed over speed of computation.
 At first sight, that might seem like a typical objective of efficiency in computer science.
-But we are not only interested in the computer science effects here --- the ambition is different.
+But we are not only interested in the computer sciece effects here --- the ambition is different.
 More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.
-
-If
 

### 2018-03-06 Chapters/Chapter02.tex
SDM

@@
 \begin{figure}[h]
-\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2.png}
+\centering
+\subfloat[1 sample for each distance $x$ \label{fig:sdm-10000w-table-7-2-1sample} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2.png}}
+\subfloat[6 samples for each distance $x$ \label{fig:sdm-10000w-table-7-2-6samples} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2-6-samples.png}}
+
 \caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
@@
 
-Figure \ref{fig:sdm-10000w-table-7-2} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance.
+Figure \ref{fig:sdm-10000w-table-7-2-1sample} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance.
 So, we have changed the steps to run $k$ reads and store the average new distance.
-We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6-samples}, which has a way lower noise and still holds the divergence.
-
-\begin{figure}[h]
-\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-6-samples.png}
-\caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
-Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
-\label{fig:sdm-10000w-table-7-2-6-samples}}
-\end{figure}
+We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6samples}, which has a way lower noise and still holds the divergence.
 
@@
 
-But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Fig.
-\ref{fig:sdm-100w-table-7-2}).
+But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Figure \ref{fig:sdm-100w-table-7-2-10samples}).
 This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
+In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractots.
 
 \begin{figure}[h]
-\centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
-\caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
+\centering
+\subfloat[100 writes \label{fig:sdm-100w-table-7-2-10samples} ]{\includegraphics[width=0.47\textwidth]{./images02/sdm-100w-table-7-2.png}}
+\subfloat[Steps of 1,000 writes \label{fig:sdm-table-7-2-steps} ]{\includegraphics[width=0.53\textwidth]{./images02/sdm-table-7-2-steps.png}}
+%\centering\includegraphics[width=\textwidth]{images02/sdm-100w-table-7-2.png}
+
+\caption{Results generated by the framework similar from Kanerva's original Table 7.2.
+Here we have a 1,000 bit, 1,000,000 hard-location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
 \label{fig:sdm-100w-table-7-2}}
@@
 To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
+Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
+Thereby, in order to show the interaction effects more clearly, we changed the single read for an 6-iterative read.
+As we can see in Figure \ref{sdm-10000w-table-7-2-6iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
+
+Hence, our understanding is that the attractors are just preventing the bitstrings to converge directly to 500-bit distance bitstrings, requiring more reading steps to do so.
+They are in other orthogonal bitstrings' critical distance, but sufficiently far not to converge in a single read.
+
+\begin{figure}[h]
+\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-6iter.png}
+\caption{This graph shows the interaction effects more clearly.
+As we change the single read to a 6-iterative read, the effect has vanished and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings.
+Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, except for the iterative read.
+\label{sdm-10000w-table-7-2-6iter}}
+\end{figure}
+
+%To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
 Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
@@
 
-\begin{figure}[h]
-\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
-\caption{This graph shows the interaction effects more clearly.
+%\begin{figure}[h]
+%\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
+%\caption{This graph shows the interaction effects more clearly.
 As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and the opposite.
 Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
-\label{sdm-10000w-notX-table-7-2}}
-\end{figure}
+%\label{sdm-10000w-notX-table-7-2}}
+%\end{figure}
 

### 2018-03-06 Chapters/Conclusion.tex
SDM, review of Intro and Conclusion.
intro + concl tese

@@
 Nichols, will participate through teleconferencing over the open web, with no use of hardware specific for the task.
-The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer (https://www-03.ibm.com/press/br/pt/biography/53561.wss).
+The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer \citep{TonnyMartins}.
 Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citet{AndrewNgAI}, as his way to emphasize the potential transformational power of the technology.
@@
 
-We have started this study with two possible forms of organization of a purely digital money system; a blockchain and a directed acyclic graph; we then moved to an analysis of the adoption of new technologies.
-
-The possibility that there will be some form of purely digital money has become very real.
+The possibility that there will be some form of purely digital money has become very real; and we have started this study with two possible forms of organization of a purely digital money system; a blockchain and a directed acyclic graph.
 Consider, just as a matter of comparison, Brazil's most important company: Petrobrás.
@@
 This respects all constraints posed by Robert M.
-French in his article on Computational Modeling in Cognitive Science: A Manifesto for Change \citep{TOPS:TOPS1206}.
+French in his article on `Computational Modeling in Cognitive Science: A Manifesto for Change' \citep{TOPS:TOPS1206}.
 

### 2018-03-06 Chapters/Introduction.tex
intro + concl tese

@@
 
-United Nations World Food Programme, `Blockchain Against Hunger: Harnessing Technology In Support Of Syrian Refugees', 30 May 2017, https://www.wfp.org/news/news-release/blockchain-against-hunger-harnessing-technology-support-syrian-refugees, accessed in January 17, 2018.
-
-United Nations World Food Programme, 2016, The Year in Review, Report.
-
-Woyke, E., 2017, `How blockchain can bring financial services to the poor', MIT Technology Review, April 18.
-
 The World Bank estimates that there are two billion people without access to financial services.
@@
 Another initiative, the United Nations World Food Programme has began, in 2017, an experiment in Jordan, in which the organization provides funds for thousands of people towards its goal of food relief.
-An interesting aspect of this program has been the format of the funds distributed:  they have been all on the ethereum blockchain.
+An interesting aspect of this program has been the format of the funds distributed:  they have been all on the ethereum blockchain \citep{woyke2017blockchain, UN-you-pay-we-take-the-photo, UNFOODPROG}.
 
@@
 Artificial intelligence has been disrupting many businesses.
-Uber is able to handle hundred of thousands of requests.
+Uber is able to handle hundreds of thousands of requests.
 Amazon optimizes the location of each product based on demand.
@@
 Spotify learns which kind of music users like the most and suggests playlists.
-Banks prevent fraud classifying which patterns are strange to their customers.
+Banks prevent fraud classifying which patterns seem to be erratic towards their customers' previous behavior.
 
@@
 Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).
-We also reproduce previous experiments and present new possibilities for SDM.
+The framework enables us to have a visual exploration previous experiments and new possibilities for SDM.
 

### 2018-03-06 Chapters/Chapter02.tex
SDM

@@
 Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
-Thereby, in order to show the interaction effects more clearly, we changed the single read for an 6-iterative read.
-As we can see in Figure \ref{sdm-10000w-table-7-2-6iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
+Thereby, in order to show the interaction effects more clearly, we changed the single read for an 15-iterative read.
+As we can see in Figure \ref{sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
 
@@
 \begin{figure}[h]
-\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-6iter.png}
+\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-15iter.png}
 \caption{This graph shows the interaction effects more clearly.
@@
 Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, except for the iterative read.
-\label{sdm-10000w-table-7-2-6iter}}
+\label{sdm-10000w-table-7-2-15iter}}
 \end{figure}
@@
 \chapter{Results (ii): Loss of neurons}
+
+In SDM, the data is written distributed among millions of hard-locations, which gives SDM robustness against loss of neurons.
+In other words, SDM should keep convering correctly even when some neurons are dead.
+The question is: how robust it really is?
+How many neurons may die before it starts to forget things.
+
+Looking for answers to those questions, we run a simulation in which we kept killing some neurons and checking whether SDM remained converging to a given bitstring.
+In this simulation, 10,000 random bitstrings were written to a 1,000-bit SDM.
+
+When we say that we killed a neuron, it means that its hard-location's counters have been zeroed and a new random bitstring address has been given to its hard-location.
+As no other bitstring is written into the memory, it is exactly like ignoring this hard-location in the next reading operations, i.e., the hard-location does not affect the next reading.
+
+In Figure \ref{fig:sdm-neuron-death-50k}, we can notice that SDM is absolutely robust up to 50,000 neuron deaths which is 5\% of the all hard-locations.
+
+\begin{figure}[h]
+\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-50k.png}
+\caption{This graph shows the SDM's robustness against loss of neurons. It shows from 0 to 50,000 neuron deaths.
+\label{fig:sdm-neuron-death-50k}}
+\end{figure}
+
+In fact, SDM begins to be affected by loss of neurons after 600,000 deaths (Figure \ref{fig:sdm-neuron-death-1m}), and obviously forgets everything when all neurons are dead.
+
+It is interesting that 500,000 deaths have almost no effect in SDM's recall.
+It is analogous to do an hemispherectomy in a person and, after the procedure, the person is able to recall exactly like before.
+
+\begin{figure}[h]
+\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-1m.png}
+\caption{This graph shows the SDM's robustness against loss of neurons.
+\label{fig:sdm-neuron-death-1m}}
+\end{figure}
+
+Although there is some decrease in SDM after 600,000, it is curious that there is a sudden change between 900,000 (90\%) and 1,000,000 (100\%).
+In Figure \ref{fig:sdm-neuron-death-details} we can see the details of this non-linear change.
+
+\begin{figure}[h]
+\centering\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death.png}
+\caption{This graph shows the SDM's robustness against loss of neurons. In each color, a given amount of neurons have been killed.
+\label{fig:sdm-neuron-death-details}}
+\end{figure}
 

### 2018-03-07 Chapters/Chapter02.tex
SDM
sdm

--- 2018-03-06 Chapters/Chapter02.tex
@@
 
+It is no coincidence that a scientific journal such as Neurocomputing states that “software is scientific method by machine”.
+Indeed, recently, Nature analysed the top-100 cited papers in history, to find
+
+\begin{quote}
+...
+some surprises, not least that it takes a staggering 12,119 citations to rank in the top 100 — and that many of the world’s most famous papers do not make the cut.
+A few that do, such as the first observation of carbon nanotubes (number 36) are indeed classic discoveries.
+But the vast majority describe experimental methods or software that have become essential in their fields.
+[...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.
+\\
+\hfill --- \citet{van2014top}
+\end{quote}
+
+The overarching intention here is to not only provide a starting point, but to provide the Framework in which SDM research can be conducted.
+Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous best, under a particular benchmark set.
+For example, some of the benchmarks that we plan to develop for future research is: how fast is convergence through iterative reading?
+How large is the attractor of the critical distance?
+How well does the system filter noise?
+How well does the system work under the aforementioned supervised learning task?
+And other authors may be able to improve this benchmark set themselves, as is usual in open source development.
+It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.
+
+Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
+We have referred to this modified read operation as Chada read.
+Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
+
+Of course, there is also new work here.
+The mathematics of the model has been to be correct (with a single exception); we have shown how to execute unsupervised learning with nothing besides operations originally proposed by Kanerva; and finally, we have reproduced
+
 \subsection{From theory to a platform.}

### 2018-03-07 Chapters/Chapter02.tex

--- 2018-03-06 Chapters/Chapter02.tex
@@
 
-\chapter{Results (ii): Performance}
+\chapter{Results (iv): Performance}
 

### 2018-03-07 Chapters/Chapter02.tex
SDM

--- 2018-03-06 Chapters/Chapter02.tex
@@
 
+In this chapter we will explore how the generalized read operation works and whether it may improve SDM behavior.
+In this case, we understand that SDM behavior will have been improved when the critical distance increases, which means it is able to recognize things with even less clues than before.
+
+In Figure
+
 \begin{figure}[h!]
   \centering
-  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_01-05.png}}
+  \subfloat[SDM behavior when $z \in \{0.1, 0.2, 0.3, 0.4, 0.5, 1\}$ ]{\includegraphics[width=\textwidth]{./images02/new-images/iter_z_01-05.png}}
 
-  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_15_3_45_6.png}}
+  \subfloat[SDM behavior when $z \in \{1.5, 3, 4.5, 6\}$ ]{\includegraphics[width=\textwidth]{./images02/new-images/z_15_3_45_6.png}}
 

### 2018-03-07 Chapters/Chapter02.tex
sdm review-text-after-read

--- 2018-03-07 Chapters/Chapter02.tex
@@
 Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
-Such model may be applied in all sort of applications because it would replicate human capacity to remember past experiences from clues of the present.
-For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where ones fear come from.
-They just feel it.
-We may interpret this situation as clues of the present --- a dark alley --- recalling past experiences from memory and thus generating the scared feeling.
+This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present.
+For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where one's fear come from.
+One just feels it.
+We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; etc.
+--- recalling past experiences from memory and thus generating the feeling.
 Our memory is able to make a parallel between previous experiences and the clues.
-Although one has never been in the exactly same situation, ones brain makes an analogy and recognizes the danger.
-This flexibility into mapping one situation in another is an important human feature which is hard to replicate into computers.
+Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger.
+This flexibility into mapping one situation in another is an important human feature which is hard to replicate in computers.
 
-It has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\cite{Linhares2011} has showed that SDM respects the limits of short-term memory discussed by \citet{Miller1995} and \citet{Cowan2011}.
+SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
+\citet{Linhares2011} has showed that SDM respects the limits of short-term memory discussed by \citet{Miller1995} and \citet{Cowan2011}.
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
-Thus, even though intriguing results have been achieved using SDM, it requires great effort of researchers to improve someone's work.
+Thus, even though intriguing results have been achieved using SDM, it requires great effort from researchers to build on top of previous work.
 
-It is our belief that such a tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python in an intuitive platform such as juypyter notebook.
+It is our belief that such a tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python, in an intuitive platform such as juypyter notebooks.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
-This new tool provides a ready to use system in which experiments can be executed almost as soon as they are designed and it may accelerate researches \citep{shen2014interactive}.
+This new tool provides a ready to use system in which experiments can be executed almost as soon as they are designed and it may accelerate research \citep{shen2014interactive}.
 
-Our motivation was our own effort in order to run our models.
+Our motivation was our own effort to run our models.
 As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug free.
-Thus, we had to deviate from our main goal --- which was to test our models --- and to focus in the implementation itself.
+Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `idea space' --- and to focus in the implementation itself.
 Furthermore, new members in our research group had to go through different source codes developed by former members.
 
-Extensions of SDM has been used in many applications.
+Extensions of SDM have been used in many applications.
 For example, \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
@@
 
-The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extended to fulfill other research models.
+The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extended to test new research ideas.
 Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier for readers to replicate any previous results and easily understand the source-code of the model.
@@
 Other contributions have also been introduced, which include (i) a noise filtering approach, (ii) a supervised classification algorithm, (iii) and a reinforcement learning algorithm, all of them using only the original SDM proposed by Kanerva, i.e., with no additional mechanisms, algorithms, data structures, etc.
-Although some of our applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them have done some adapting of SDM to their problems, and none of them have used just the ideas introduced by Kanerva.
-We have presented different approaches with no adaptations at all.
+Although some of these applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them have done some adapting of SDM to their problems, and none of them have used just the ideas introduced by Kanerva.
+We have presented different approaches with no adaptations whatsoever.
 
-Finally, we have found an anomaly in one of Kanerva's prediction, which we believe is related to SDM capacity.
-We have also tested a generic reading operation proposed by professor Paulo Murilo (personal communication).
+Finally, I have striven to provide a visual tour of the theory and application of SDM: whenever possible, detailed figures should tell the story --- or at least do the heavy lifting.
+In this study, we will see an anomaly in one of Kanerva's predictions, which I believe is related to SDM capacity.
+We will see tests of a generalized reading operation proposed by Physics Professor Paulo Murilo (personal communication).
+We will see what happens when neurons --- and
+all their information --- is simply and suddenly lost.
+We will see whether information-theory can improve some of Kanerva's ideas.
+From (basic) noise filtering to learning to play tic-tac-toe, we will review the entirety of Dr.
+Pentti Kanerva's proposal.
+
+This time, however, it will be running on a computer.
 
@@
   $n$ & Number of dimensions, i.e., $n=1,000$. \\
-  $N$ & Size of the binary space --- $|\{0, 1\}^n| = 2^n$. \\
+  $N$ & Size of the binary space, $|\{0, 1\}^n| = 2^n$. \\
   $N'$ & Number of hard-locations samples from $\{0, 1\}^n$. Its typical value is 1,000,000, as suggested by \citet{Kanerva1988}. \\
   $H$ & Same as $N'$. \\
-  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, its typical value is $451$. This value is calculated to activate, on average, one thousand of $N'$. \\
+  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, its typical value is $451$. This value is calculated to activate, on average, one thousandth of $N'$. \\
   $\eta$ & A bitstring, usually a datum. \\
@@
   $\xi$ & A bitstring, usually an address. \\
-  $\text{dist}(x, y)$ & Hamming distance between $x$ and $y$ \\
-  $\text{d}(x, y)$ & Same as $\text{dist}(x, y)$
+  $\text{dist}(x, y)$ & Hamming distance between $x$ and $y$. \\
+  $\text{d}(x, y)$ & Same as $\text{dist}(x, y)$.
 \end{tabular}\\
@@
 
-$\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
-A reduction from measuring the distance to $10^6$ hard locations to a computation of $~ 10^3$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.
+Consider the hypercube with $n$ dimensions.
+We want to select a subset of its vertices with cardinality $2^{20}$ that is symmetrically distributed over the space.
+Afterwards, $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
 
+A reduction from measuring the distance to $2^{20}$ hard locations to a computation of $2^{10}$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.
 At large scales of computing, the very ability to perform some experiments is a function of sheer performance.

### 2018-03-07 Chapters/sdm2.tex
sdm review-text-after-read

--- 2018-03-05 Chapters/sdm2.tex
@@
 
-Sparse Distributed Memory (SDM) is a mathematical model for cognitive memory published by \citet{Kanerva1988}.
-It introduces many interesting mathematical properties of $n$-dimensional binary space that, in a memory model, are psychologically plausible.
+Sparse Distributed Memory (SDM) is a mathematical model developed and suggested as a theory of human memory by Finish Scientist Pentti \citet{Kanerva1988}.
+It introduces many interesting mathematical properties of $n$-dimensional binary space that, in a memory model, seem to be remarkably psychologically plausible.
 Most notable among these are the tip-of-the-tongue phenomenon, conformity to Miller's magic number \citep{Linhares2011} and robustness against loss of neurons.
 
-The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings.
-The distance between two bitstrings is calculated using the Hamming distance.
-It is defined for two bitstrings of equal length as the number of positions at which the bits are different.
+The data --- and address space on which it is stored --- are represented by large sequences of bits, called \emph{bitstrings}.
+The \emph{Hamming distance} provides comparisons between bitstrings and is used as a metric for the system.
+The Hamming distance is defined for two bitstrings of equal length as the number of positions in which bits differ.
 For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
-One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance in 3 dimensions, i.e., both follow triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which Euclidean distance in 3 dimensions may be interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{and} d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
-Hence, it makes no sense to say that $A$ is also close to $C$.
-This difference in intuition may trick even experienced researchers when analyzing some situations.
 
@@
 A number of observations can be made here.
-First, the number of nodes grows as $2^n$ as $n$ grows; which makes the  space intractable as $n>>20$.
-Another interesting observation, better seen in the figures below, is that most of the space lies `at the center', at a distance of around 500 from any given vantage point.\label{hypercubes}}
+First, the number of nodes grows as $O(2^n)$; which makes the space rapidly intractable.
+Another interesting observation, better seen in the figures below, is that most of the space lies `at the center', at a distance of around $n/2$ from any given vantage point.\label{hypercubes}}
 \end{figure}
@@
 For a fixed $n \in \mathbb{Z}$, the graph $G = (V, E)$, in which $v \in V$ iff there is a bijective function $b: V\to \{0,1\}^n$ and $(v_i, v_j) \in E$ iff $H(b(v_i), b(v_j))=1$, where $H$ is the Hamming distance.
-That is, $n$-sized bitstrings correspond to nodes and edges exist between nodes $iff$ they flip a single bit.
-Though Kanerva has derived many combinatorial properties of the space, additional results have been found in the graph-theoretical literature.
+That is, $n$-sized bitstrings correspond to nodes, and edges exist between nodes $iff$ they flip a single bit.
+Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community.
 A good survey is provided by \citet{harary1988survey}.
+
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, our 3-dimensional space.
+
+Though both follow the triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which in 3-d Euclidean distance may be loosely interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{ and } d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
+Hence, it makes no sense to say that $A$ is also close to $C$.
+
+There are numerous, beautiful, counter-intuitive notions involved in this space. This difference in intuition may trick even experienced researchers when analyzing some situations.
 

### 2018-03-07 Chapters/Chapter02.tex
SDM OpenCL Kernel

--- 2018-03-07 Chapters/Chapter02.tex
@@
 The OpenCL scanner supports running into multiple devices.
+
+\subsection{OpenCL kernel}
+
+The OpenCL kernel explores the GPU architecture to improve performance.
+Each work group calculates the distance of several bitstrings.
+During the distance calculation, each worker calculates the exclusive OR (XOR) between two 64-bit integers and use the built-in popcount function to count the number of ones.
+Then, they update an array of intermediate distances with their partial distances.
+This array is stored in the local memory and is shared with all workers of the same group.
+This whole step happens simultaneously in the GPU.
+Then, we used a reduction algorithm to sum the intermediate distances array in order to calculate the correct distance.
+This reduction algorithm is also distributed between the workers and runs in $O(\log_2(\text{bs\_step}))$.
+Finally, one of the workers checks whether the distance is less than or equal to the radius to include the bitstring index into the resulting array.
+
+The number of workers in each group is the closest power of two above bs\_len (which is the number of 64-integers that forms a bitstring).
 

### 2018-03-07 Chapters/Chapter02.tex
intro + sdm

--- 2018-03-07 Chapters/Chapter02.tex
@@
 
-The framework has both Kanerva's read and the generalized read.
-The latter was implemented according to the generalization proposed by professor Paulo Murilo.
+The framework has both Kanerva's read and Murilo's generalized read.
 The generalization brings a parameter $z$, which is the exponent.
@@
 
-In this chapter we will explore how the generalized read operation works and whether it may improve SDM behavior.
+In this chapter we will explore how Murilo's generalized read operation works and whether it may improve SDM behavior.
 In this case, we understand that SDM behavior will have been improved when the critical distance increases, which means it is able to recognize things with even less clues than before.
@@
 
-Physicist Paulo Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
+Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
 The results, however, have not yielded performance improvements.

### 2018-03-07 Chapters/Chapter02.tex
SDM Loss of neurons

--- 2018-03-07 Chapters/Chapter02.tex
@@
 This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
-In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractots.
+In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractors.
 
@@
 
-In SDM, the data is written distributed among millions of hard-locations, which gives SDM robustness against loss of neurons.
+In SDM, the data is written distributed among millions of hard-locations, which theoretically gives SDM robustness against loss of neurons.
 In other words, SDM should keep convering correctly even when some neurons are dead.
@@
 How many neurons may die before it starts to forget things.
+These questions have never been addresses before.
 
-Looking for answers to those questions, we run a simulation in which we kept killing some neurons and checking whether SDM remained converging to a given bitstring.
-In this simulation, 10,000 random bitstrings were written to a 1,000-bit SDM.
+Looking for answers to these questions, we run simulations in which we kept killing some neurons and checking whether SDM remained converging to a given bitstring or not.
+In these simulations, 10,000 random bitstrings were written to a 1,000-bit SDM with 1,000,000 hard-locations, and we choose one of them as our target.
+As the bitstrings were all written exactly once, we may generalize the results.
+The code is available in the ``Reseting hard-locations'' notebook \citep{sdmframework}.
 
-When we say that we killed a neuron, it means that its hard-location's counters have been zeroed and a new random bitstring address has been given to its hard-location.
-As no other bitstring is written into the memory, it is exactly like ignoring this hard-location in the next reading operations, i.e., the hard-location does not affect the next reading.
+As neurons are hard-locations in SDM, when we say that a neuron has been killed, we mean that its counters have been zeroed and a new random bitstring address has been assigned.
+During our simulations, no other bitstring has been written after the 10,000.
+Consequently, as their counters willl remain zeroed, it is exactly like ignoring the dead hard-locations in the subsequent reading operations.
 
-In Figure \ref{fig:sdm-neuron-death-50k}, we can notice that SDM is absolutely robust up to 50,000 neuron deaths which is 5\% of the all hard-locations.
-
-\begin{figure}[h]
-\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-50k.png}
-\caption{This graph shows the SDM's robustness against loss of neurons. It shows from 0 to 50,000 neuron deaths.
-\label{fig:sdm-neuron-death-50k}}
+\begin{figure}[!p]
+\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-200k.png}
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
+It shows that a loss of 200,000 neurons, 20\% of the total, does not affect SDM whatsoever.
+\label{fig:sdm-neuron-death-200k}}
 \end{figure}
 
-In fact, SDM begins to be affected by loss of neurons after 600,000 deaths (Figure \ref{fig:sdm-neuron-death-1m}), and obviously forgets everything when all neurons are dead.
+In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is absolutely robust up to 200,000 neuron deaths which is 20\% of all hard-locations.
+This result is pretty impressive and really surprised us.
 
-It is interesting that 500,000 deaths have almost no effect in SDM's recall.
-It is analogous to do an hemispherectomy in a person and, after the procedure, the person is able to recall exactly like before.
+In fact, SDM begins to be significantly affected by loss of neurons after 600,000 neuron deaths (Figure \ref{fig:sdm-neuron-death-1m}), and obviously forgets everything when all neurons are dead.
 
-\begin{figure}[h]
+\begin{figure}[!p]
 \centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-1m.png}
-\caption{This graph shows the SDM's robustness against loss of neurons.
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
+The more neurons are death, the smaller the critical distance, i.e., the worse the SDM recall.
 \label{fig:sdm-neuron-death-1m}}
@@
 
-Although there is some decrease in SDM after 600,000, it is curious that there is a sudden change between 900,000 (90\%) and 1,000,000 (100\%).
+It is interesting that 500,000 neuron deaths has a minor effect in SDM's recall capability (see Figure \ref{fig:sdm-neuron-death-500k}).
+It is analogous to do an hemispherectomy in a person and, after the procedure, the person is able to recall and learn almost just like before.
+In fact, there are clinical reports of children submitted to hemispherectomy who live an almost normal life with minor motor problems.
+
+\begin{figure}[!p]
+\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-500k.png}
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
+Even when 50\% of neurons are dead, SDM recall is barely affected, which is an impressive result and matches with some clinical results of children submitted to hemispherectomy.
+\label{fig:sdm-neuron-death-500k}}
+\end{figure}
+
+An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance is really small, i.e., SDM recall capacity is very diminished.
+After 900,000 neuron deaths the critical distance is zero.
+In this case, everything has been forgot.
+
+Although there is some decrease in SDM recall after 600,000 neuron deaths, it is curious that there is a sudden change between 900,000 (90\%) and 1,000,000 (100\%).
 In Figure \ref{fig:sdm-neuron-death-details} we can see the details of this non-linear change.
+Notice that after 950,000 even the exact clue $\eta_0$ does not converge to $\eta$.
 
-\begin{figure}[h]
+\begin{figure}[!p]
 \centering\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death.png}
-\caption{This graph shows the SDM's robustness against loss of neurons. In each color, a given amount of neurons have been killed.
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
 \label{fig:sdm-neuron-death-details}}
+\end{figure}
+
+We run exactly the same simulation for a 256-bit SDM with 1,000,000 hard-locations.
+The results were even more surprising, because the 256-bit SDM was more robust to loss of neuron than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
+Notice that the loss of 50\% of neurons barely affected the 256-bit SDM which remained functional even facing an enormous loss of 90\% of neurons.
+
+\begin{figure}[!p]
+\centering
+\subfloat[Up to 500,000 neuron deaths ]{\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death-256-500k.png}}
+
+\subfloat[From 600,000 to 1,000,000 neuron deaths ]{\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death-256-1m.png}}
+
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=256$ and $H=1,000,000$.
+\label{fig:sdm-neuron-death-256bits}}
 \end{figure}

### 2018-03-07 Chapters/Introduction.tex
intro + sdm
intro

@@
 
-The impact of Bitcoin in society -- and hence in the companies and the government -- has been growing every day.
+The impact of Bitcoin in society --- and hence in the companies and the government --- has been growing every day.
 People are increasingly using Bitcoin to exchange money and transfer money overseas.
@@
 It may be used in classification, face recognition, character recognition, and so forth.
+Even if AI cannot yet solve numerous hard problems --- such as the Turing Test, Bongard Problems or simply understanding when `Lawyers are sharks' \citep{french1990subcognition, french2000turing, linhares2000glimpse, french2001co} ---, it is clear that the technology must be taken seriously.
 
@@
 The forecast was done applying a disease spreading model which has correctly predicted the abandonment of ``MySpace''.
-Facebook replied applying exactly the same methodology as Princeton's researchers.
-In their own words: ``Using the same robust methodology featured in [Princeton's] paper, we attempted to find out more about this `Princeton University' -- and you won't believe what we found!''.
+Facebook replied after applying Princeton's methodology:
+
+\begin{quote}
+    ``Using the same robust methodology featured in [Princeton's] paper, we attempted to find out more about this `Princeton University' --- and you won't believe what we found!''.
 Then, they conclude: ``This trend suggests that Princeton will have only half its current enrollment by 2018, and by 2021 it will have no students at all, agreeing with the previous graph of scholarly scholarliness.
 Based on our robust scientific analysis, future generations will only be able to imagine this now-rubble institution that once walked this earth''.
+\end{quote}
 

### 2018-03-07 Chapters/Chapter02.tex
SDM

@@
 
-In this chapter we will explore how the generalized read operation works and whether it may improve SDM behavior.
-In this case, we understand that SDM behavior will have been improved when the critical distance increases, which means it is able to recognize things with even less clues than before.
+In this chapter we will explore how the generalized read operation works and whether it may improve SDM behavior or not.
+In this case, we are specifically interested into the impact in the critical distance.
+An increase in , which means it is able to recognize things with even less clues than before.
 

### 2018-03-07 Chapters/sdm2.tex

@@
 
-A member of my Master's committee, professor Paulo Murilo\footnote{Universidade Federal Fluminense's physicist doctor Paulo Murilo}, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
+A member of my Master's committee, Prof.
+Paulo Murilo\footnote{Universidade Federal Fluminense's Physics Professor Paulo Murilo}, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
 He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative).

### 2018-03-07 Chapters/Chapter02.tex
remove git tag

--- 2018-03-07 Chapters/Chapter02.tex
@@
 
-In this chapter we will explore how the generalized read operation works and whether it may improve SDM behavior.
+In this chapter we will explore how Murilo's generalized read operation works and whether it may improve SDM behavior.
 In this case, we understand that SDM behavior will have been improved when the critical distance increases, which means it is able to recognize things with even less clues than before.
 
-In Figure
+% ?????????  In Figure
 
@@
 
-But, qualitatively, they seem disappointing.  HEATMAPFIGs
-
 Note, finally, that this is not the first time in which a weighted function has been applied to writing in SDM --- \citet{hely1997new} suggest a rather complex spreading model based on floating point signals in the interval [0.05, 1.0] --- they were, however, only able to test their model with 1,000 hard locations.

### 2018-03-08 Chapters/Chapter02.tex
sdm

@@
 Of course, there is also new work here.
-The mathematics of the model has been to be correct (with a single exception); we have shown how to execute unsupervised learning with nothing besides operations originally proposed by Kanerva; and finally, we have reproduced
+The mathematics of the model has been shown to be correct (with a single exception); we have shown how to execute unsupervised learning with nothing besides operations originally proposed by Kanerva; we have studied the generalized Murilo read; we have seen and finally, we have reproduced
 

### 2018-03-08 Chapters/Chapter02.tex
SDM

--- 2018-03-07 Chapters/Chapter02.tex
@@
 
-Our intention is to provide comparative performance metrics under different computation engines (CPU, GPU, etc) and different operating systems (Linux, MacOs, Windows, etc).
-Performance can be measured as the average number of scans of all hard locations per second, reads per second, writes per second, etc.
+Performance is an important part of our framework. Running an SDM memory consumes a lot of resource from both CPU and RAM.
+
+For instance, each scan in 1,000,000 hard-locations with 1,000 bits executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
+So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
+The heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} executed $3.05 \cdot 10^{15}$ bit compares.
+For comparison, the number of seconds since Jesus's birth is 63,639,648,000 = $6.36 \cdot 10^{10}$.
+The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$.
+There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.
 
 Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-We also intend to test on machines such as the iMac with dedicated GPU, MacPro with dedicated GPU, and personal computers under Linux with dedicated GPUs.
+Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
 
-Beyond that, we are running as state-of-the-art devices: (i) an Amazon EC2 p3.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.8xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and 8x NVIDIA K80 GPU.
+Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+
+Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds.
+It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.
+
+We have executed the same performance test in each device.
+The test has 3 parts.
+The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans.
+The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners.
+The code is available in the ``Performance test'' notebook \citep{sdmframework}.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
+\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
+\label{fig:performance-imac-scan-1000}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
+\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
+\label{fig:performance-imac-scan-256}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
+\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
+\label{fig:performance-imac-scan-10k}}
+\end{figure}
+
+\begin{center}
+    iMac $n=1,000$ bits
+    \begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 81.62 & 12.25 \\
+    Thread scan & 5000 & 143.68 & 6.95 \\
+    OpenCL scan & 5000 & 238.00 & 4.20 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 1000 & 74.92 & 13.34 \\
+    Thread single read & 1000 & 96.22 & 10.39 \\
+    OpenCL write & 1000 & 126.50 & 7.90 \\
+    OpenCL single read & 1000 & 190.20 & 5.25 \\
+    \hline
+    \end{tabular}
+\end{center}
+
+\begin{center}
+    iMac $n=256$ bits
+    \begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 175.48 & 5.69 \\
+    Thread scan & 5000 & 352.63 & 2.83 \\
+    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 2000 & 304.46 & 3.28 \\
+    Thread single read & 2000 & 391.21 & 2.55 \\
+    OpenCL write & 2000 & 378.44 & 2.64 \\
+    OpenCL single read & 2000 & 466.16 & 2.14 \\
+    \hline
+    \end{tabular}
+\end{center}
+
+\begin{center}
+    iMac $n=10,000$ bits
+    \begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+        Linear scan & 100 & 8.59 & 116.38 \\
+        Thread scan & 500 & 18.66 & 53.56 \\
+        OpenCL scan & 1000 & 77.20 & 12.95 \\
+    \hline
+    \end{tabular}
+\end{center}
 

### 2018-03-08 Chapters/Chapter02.tex

@@
 
-% ?????????  In Figure
-
 \begin{figure}[h!]
@@
 %]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}
+
+\chapter*{List of jupyter notebooks}

### 2018-03-08 Chapters/Chapter02.tex
SDM Performance

@@
 
-Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
-
-Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
-
 Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds.
@@
 
+Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+
+Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
+The complete results are presented in Figures \ref{fig:perf-imac-read-256}, \ref{fig:perf-imac-write-256}, \ref{fig:perf-imac-read-1000}, \ref{fig:perf-imac-write-1000}, \ref{fig:perf-imac-scan-256}, \ref{fig:perf-imac-scan-1000}, and \ref{fig:perf-imac-scan-10k} and Tables \ref{tab:perf-imac-256}, \ref{tab:perf-imac-1000}, and \ref{tab:perf-imac-10k}.
+
+Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+
+It is interesting that which scanner is faster depends also on the SDM settings.
+In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL, but for a 256-bit SDM was the threads.
+What happened here is that the OpenCL kernel chosen was a generic one which performs the scan for any SDM.
+It is always possible to optimize the OpenCL kernel to a specific setting, and it would be faster than the threads.
+By default, the framework chooses a generic kernel which we believe would be reasonable for the most common setups.
+% TODO Run the optimized kernel and show the results.
+
+Most of the time, the bottleneck of both the read and the write operations is the scanner.
+But we have optimized the OpenCL kernel so much for the iMac 2017 that scanner not the bottleneck anymore.
+In the writing operation, it took the same time to scan and to update the counters.
+It is impressive because it update, on average, 1,000 counters of 32-bit integers each, and it was as fast as (i) sending the command to the GPU, (ii) performing 1 billion bit compares, and (iii) downloading the response from the GPU.
+
+Our conclusion is that, if one is really concerned about performance, one should fine tune the OpenCL kernel for one's GPU. It would always be faster than running in the CPU.
+
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
-\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:performance-imac-scan-1000}}
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-256.png}
+\caption{Time to run a single read in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
+\label{fig:perf-imac-read-256}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-256.png}
+\caption{Time to run one write in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
+\label{fig:perf-imac-write-256}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-1000.png}
+\caption{Time to run one write in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
+\label{fig:perf-imac-write-1000}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-1000.png}
+\caption{Time to run a single read in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
+\label{fig:perf-imac-read-1000}}
 \end{figure}
@@
 \caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:performance-imac-scan-256}}
+\label{fig:perf-imac-scan-256}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
+\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
+\label{fig:perf-imac-scan-1000}}
 \end{figure}
@@
 \caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
-\label{fig:performance-imac-scan-10k}}
+\label{fig:perf-imac-scan-10k}}
 \end{figure}
 
-\begin{center}
-    iMac $n=1,000$ bits
+\begin{table}[!htb]
+\centering
 \begin{tabular}{| l | r | r | r |}
@@
 \end{tabular}
-\end{center}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
+\label{tab:perf-imac-1000}}
+\end{table}
 
-\begin{center}
-    iMac $n=256$ bits
+\begin{table}[!htb]
+\centering
 \begin{tabular}{| l | r | r | r |}
@@
 \end{tabular}
-\end{center}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+\label{tab:perf-imac-256}}
+\end{table}
 
-\begin{center}
-    iMac $n=10,000$ bits
+\begin{table}[!htb]
+\centering
 \begin{tabular}{| l | r | r | r |}
@@
 \end{tabular}
-\end{center}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.
+There is no benchmark for read and write operations because RAM is not enough to allocate the counters --- it would consume 37.25 GB of RAM.
+\label{tab:perf-imac-10k}}
+\end{table}
 

### 2018-03-08 Chapters/Chapter02.tex
SDM

@@
 
-In this chapter we will explore how Murilo's generalized read operation works and whether it may improve SDM behavior.
-In this case, we understand that SDM behavior will have been improved when the critical distance increases, which means it is able to recognize things with even less clues than before.
+Murilo observed that the models of Kanerva's read ($z=1$) and Chada's read ($z=0$) were simple variations of a generalized read with an exponent $z$, which suggests experimenting with different values.
+Mathematically, let $A$ be the set of the counters of the activated hardlocation, and $c_i$ be the counter of the $i$-th bit.
+Then,
+
+$$
+s_i = \sum_{c \in A} \frac{c_i}{|c_i|} |c_i|^z
+$$
+
+The sum of $|c_i|^z$ turns the intermediate values from integers to float point numbers. Thus, we have developed a specific read operation which stored the intermediate values in double variables.
+
+The results, however, have not yielded performance improvements.
+Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller critical distance and faster divergence at large-distance reads.
+This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
+
+We understand that the critical distant is an important parameter of SDM.
+The bigger the critical distance, the best, because SDM is able to converge even with farther clues.
+For $z>1$, the bigger the $z$, the smaller the critical distance.
+For $z = 6$, the critical distance almost reaches zero.
+
+It is interesting that Kanerva has proposed $z=1$ without realizing the generalized reading. Even so, he proposed the optimal $z$ --- the one with the highest critical distance.
 
@@
 
-  \caption{(a) and (b) show the behavior of a single read; As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
+  \caption{(a) and (b) show the behavior of a single read.
+As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
 Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
@@
   \centering
-  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_0_1.png}}
+  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=\textwidth]{./images02/new-images/z_0_1.png}}
 
-  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_all_2.png}}
+  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=\textwidth]{./images02/new-images/iter_z_all_2.png}}
 
-  \caption{(a) and (b) show the behavior of the previous Figure, now executed with 6 iterative reads.
+  \caption{(a) and (b) show the behavior of Figure \ref{fig:murillo-generalization-experiments}, now executed with 6-iterative reads.
 What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
@@
 \end{figure}
-
-Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
-The results, however, have not yielded performance improvements.
-Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads.
-This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
 

### 2018-03-08 Chapters/Chapter02.tex
sdm conclusion

--- 2018-03-08 Chapters/Chapter02.tex
@@
 
+% ?????????  In Figure
+
 \begin{figure}[h!]
@@
 
-Performance is an important part of our framework. Running an SDM memory consumes a lot of resource from both CPU and RAM.
-
-For instance, each scan in 1,000,000 hard-locations with 1,000 bits executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
-So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
-The heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} executed $3.05 \cdot 10^{15}$ bit compares.
-For comparison, the number of seconds since Jesus's birth is 63,639,648,000 = $6.36 \cdot 10^{10}$.
-The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$.
-There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.
+Our intention is to provide comparative performance metrics under different computation engines (CPU, GPU, etc) and different operating systems (Linux, MacOs, Windows, etc).
+Performance can be measured as the average number of scans of all hard locations per second, reads per second, writes per second, etc.
 
 Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
+We also intend to test on machines such as the iMac with dedicated GPU, MacPro with dedicated GPU, and personal computers under Linux with dedicated GPUs.
 
-Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
-
-Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds.
-It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.
-
-We have executed the same performance test in each device.
-The test has 3 parts.
-The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans.
-The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners.
-The code is available in the ``Performance test'' notebook \citep{sdmframework}.
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
-\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:performance-imac-scan-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
-\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:performance-imac-scan-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
-\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
-\label{fig:performance-imac-scan-10k}}
-\end{figure}
-
-\begin{center}
-    iMac $n=1,000$ bits
-    \begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 81.62 & 12.25 \\
-    Thread scan & 5000 & 143.68 & 6.95 \\
-    OpenCL scan & 5000 & 238.00 & 4.20 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 1000 & 74.92 & 13.34 \\
-    Thread single read & 1000 & 96.22 & 10.39 \\
-    OpenCL write & 1000 & 126.50 & 7.90 \\
-    OpenCL single read & 1000 & 190.20 & 5.25 \\
-    \hline
-    \end{tabular}
-\end{center}
-
-\begin{center}
-    iMac $n=256$ bits
-    \begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 175.48 & 5.69 \\
-    Thread scan & 5000 & 352.63 & 2.83 \\
-    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 2000 & 304.46 & 3.28 \\
-    Thread single read & 2000 & 391.21 & 2.55 \\
-    OpenCL write & 2000 & 378.44 & 2.64 \\
-    OpenCL single read & 2000 & 466.16 & 2.14 \\
-    \hline
-    \end{tabular}
-\end{center}
-
-\begin{center}
-    iMac $n=10,000$ bits
-    \begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-        Linear scan & 100 & 8.59 & 116.38 \\
-        Thread scan & 500 & 18.66 & 53.56 \\
-        OpenCL scan & 1000 & 77.20 & 12.95 \\
-    \hline
-    \end{tabular}
-\end{center}
+Beyond that, we are running as state-of-the-art devices: (i) an Amazon EC2 p3.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.8xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and 8x NVIDIA K80 GPU.
 
@@
 
+Deep neural networks, alphaGo, alphazero
+
 \subsection{i versus l}
@@
 The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
-More on this in my closing words.
+A little more on this in my closing words.
 
@@
 
-It is no coincidence that a scientific journal such as Neurocomputing states that “software is scientific method by machine”.
-Indeed, recently, Nature analysed the top-100 cited papers in history, to find
+The overarching intention here is to not only provide a starting point, but to provide a Framework in which SDM research can be conducted.
+Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).
+For example, some of the benchmarks that we plan to develop in future research is: how fast is convergence through iterative reading?
+How large is the attractor of the critical distance?
+How well does the system filter noise?
+How well does the system work under the supervised learning task?
+And other authors may be able to improve this benchmark set themselves, as is usual in open source development.
+It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.
+
+Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
+We have referred to this modified read operation as Chada read (the tale is that my friend \text{\&} colleague, Dr.
+Daniel de Magalhães Chada, along with Linhares, did not consult Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterwards).
+Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
+Indeed, a high-quality codebase seems to have become a scientific community's form of unequivocally standing behind a consensus.
+For example, the journal Nature analysed the top-100 cited papers in history, to find:
 
@@
 A few that do, such as the first observation of carbon nanotubes (number 36) are indeed classic discoveries.
-But the vast majority describe experimental methods or software that have become essential in their fields.
-[...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.
-\\
-\hfill --- \citet{van2014top}
+But the vast majority describe experimental methods or \emph{software that have become essential in their fields.
+[...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.} \\
+\hfill --- \citet{van2014top}, emphasis mine.
 \end{quote}
 
-The overarching intention here is to not only provide a starting point, but to provide the Framework in which SDM research can be conducted.
-Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous best, under a particular benchmark set.
-For example, some of the benchmarks that we plan to develop for future research is: how fast is convergence through iterative reading?
-How large is the attractor of the critical distance?
-How well does the system filter noise?
-How well does the system work under the aforementioned supervised learning task?
-And other authors may be able to improve this benchmark set themselves, as is usual in open source development.
-It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.
+It is no coincidence that scientific journals such as BMC Neuroscience, or the Journal of Machine Learning Research have specific sections on open-source software.
+Neurocomputing states that “software is scientific method by machine”.
 
-Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
-We have referred to this modified read operation as Chada read.
-Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
+Of course, for the skeptical reader who may consider software a less worthy pursuit, there is also new work here.
+The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.
+The emphasis might have been on \emph{breadth of topics}, in detriment of depth here or there.
+But this is due to our research group's enthusiam for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.
+If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.
+The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
+Some essays completely shut the door close at the end; this one intends to leave it wide open.
+As the reader may have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
 
-Of course, there is also new work here.
-The mathematics of the model has been shown to be correct (with a single exception); we have shown how to execute unsupervised learning with nothing besides operations originally proposed by Kanerva; we have studied the generalized Murilo read; we have seen and finally, we have reproduced
+\begin{figure}[!p]
+\centering
 
-\subsection{From theory to a platform.}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}
 
-Platforms in the history of computing.
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}
 
-Opens 10 doors.
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}
+
+\caption{Some slides from an old Linhares class \citet{only-one-life}.
+In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
+It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
+Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
+\label{fig:one-life}}
+\end{figure}
 
@@
 %]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}
-
-\chapter*{List of jupyter notebooks}

### 2018-03-08 Chapters/Chapter02.tex
sdm conclusion

@@
 The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
+Ralph Waldo Emerson once said \emph{Do not go where the path may lead, go instead where there is no path and leave a trail.}  Dr.
+Kanerva has left the trail.
+It is my job to pave it and try to deliver an easier road for the next generation.
 Some essays completely shut the door close at the end; this one intends to leave it wide open.
-As the reader may have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
+As the reader might have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
 

### 2018-03-08 Chapters/Chapter02.tex
Intro & sdm conclusion

@@
 
+\begin{figure}[!p]
+\centering
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}
+
+\caption{Some slides from an old Linhares class; a viewpoint that has influenced the choice of topics found in this thesis.
+In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
+It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
+It is, in a sense, the Big Drama of our moment in history.
+Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
+Note that the slides were made in 2007, and expected an AI-based `semantic revolution` by 2015 \citep{only-one-life}.
+In between, both the iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created (Uber, Instagram, etc, on the case of smartphones; and exchanges, miners, investors, payment processors, etc., in the case of blockchains).
+Linhares's thesis is that students should (i) expect new technological platforms, (ii) rapidly identify them, and (iii) skydive, truly throw their lifes, at them.
+It is there that we might find leverage to make a small impact upon the world.
+\label{fig:one-life}}
+\end{figure}
+
 Let us revisit, in these concluding thoughts, the emphasis employed over speed of computation.
@@
 It is no coincidence that scientific journals such as BMC Neuroscience, or the Journal of Machine Learning Research have specific sections on open-source software.
-Neurocomputing states that “software is scientific method by machine”.
+The journal Neurocomputing even goes as far as to state that “software is scientific method by machine”.
 
@@
 The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
-Ralph Waldo Emerson once said \emph{Do not go where the path may lead, go instead where there is no path and leave a trail.}  Dr.
+Ralph Waldo Emerson once said \emph{do not go where the path may lead, go instead where there is no path, and leave a trail.}  Dr.
 Kanerva has left the trail.
-It is my job to pave it and try to deliver an easier road for the next generation.
+It is my job to pave it and throw light at it and try to deliver an easier pathway for the next generation.
 Some essays completely shut the door close at the end; this one intends to leave it wide open.
 As the reader might have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
-
-\begin{figure}[!p]
-\centering
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}
-
-\caption{Some slides from an old Linhares class \citet{only-one-life}.
-In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
-It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
-Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
-\label{fig:one-life}}
-\end{figure}
 

### 2018-03-08 Chapters/Chapter02.tex
SDM Performance data for Amazon EC2 p2.xlarge and p3.2xlarge.

--- 2018-03-08 Chapters/Chapter02.tex
@@
 
-Murilo observed that the models of Kanerva's read ($z=1$) and Chada's read ($z=0$) were simple variations of a generalized read with an exponent $z$, which suggests experimenting with different values.
-Mathematically, let $A$ be the set of the counters of the activated hardlocation, and $c_i$ be the counter of the $i$-th bit.
-Then,
-
-$$
-s_i = \sum_{c \in A} \frac{c_i}{|c_i|} |c_i|^z
-$$
-
-The sum of $|c_i|^z$ turns the intermediate values from integers to float point numbers. Thus, we have developed a specific read operation which stored the intermediate values in double variables.
-
-The results, however, have not yielded performance improvements.
-Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller critical distance and faster divergence at large-distance reads.
-This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
-
-We understand that the critical distant is an important parameter of SDM.
-The bigger the critical distance, the best, because SDM is able to converge even with farther clues.
-For $z>1$, the bigger the $z$, the smaller the critical distance.
-For $z = 6$, the critical distance almost reaches zero.
-
-It is interesting that Kanerva has proposed $z=1$ without realizing the generalized reading. Even so, he proposed the optimal $z$ --- the one with the highest critical distance.
+In this chapter we will explore how Murilo's generalized read operation works and whether it may improve SDM behavior.
+In this case, we understand that SDM behavior will have been improved when the critical distance increases, which means it is able to recognize things with even less clues than before.
 
@@
 
-  \caption{(a) and (b) show the behavior of a single read.
-As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
+  \caption{(a) and (b) show the behavior of a single read; As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
 Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
@@
   \centering
-  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=\textwidth]{./images02/new-images/z_0_1.png}}
-
-  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=\textwidth]{./images02/new-images/iter_z_all_2.png}}
-
-  \caption{(a) and (b) show the behavior of Figure \ref{fig:murillo-generalization-experiments}, now executed with 6-iterative reads.
+  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_0_1.png}}
+
+  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_all_2.png}}
+
+  \caption{(a) and (b) show the behavior of the previous Figure, now executed with 6 iterative reads.
 What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
@@
 \end{figure}
+
+Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values.
+The results, however, have not yielded performance improvements.
+Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads.
+This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
 
@@
 Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.
-There is no benchmark for read and write operations because RAM is not enough to allocate the counters --- it would consume 37.25 GB of RAM.
+There is no benchmark for read and write operations because RAM was not enough to allocate the counters --- it would consume 37.25 GB of RAM.
 \label{tab:perf-imac-10k}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 15.44 & 64.73 \\
+    Thread scan & 5000 & 30.47 & 32.81 \\
+    OpenCL scan & 5000 & 93.63 & 10.67 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 1000 & 23.44 & 42.64 \\
+    Thread single read & 1000 & 28.26 & 35.37 \\
+    OpenCL write & 1000 & 51.78 & 19.31 \\
+    OpenCL single read & 1000 & 85.31 & 11.72 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 52.35 & 19.09 \\
+    Thread scan & 5000 & 100.48 & 9.95 \\
+    OpenCL scan & 5000 & 145.18 & 6.88 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 2000 & 84.73 & 11.80 \\
+    Thread single read & 2000 & 95.26 & 10.49 \\
+    OpenCL write & 2000 & 113.01 & 8.84 \\
+    OpenCL single read & 2000 & 133.19 & 7.50 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 100 & 1.66 & 600.75 \\
+    Thread scan & 500 & 3.37 & 296.56 \\
+    OpenCL scan & 1000 & 21.39 & 46.73 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 500 & 2.60 & 383.50 \\
+    Thread single read & 500 & 3.24 & 307.97 \\
+    OpenCL write & 1000 & 8.18 & 122.17 \\
+    OpenCL single read & 55.47 & 18.02 & 55.47 \\
+    \hline
+\end{tabular}
+\caption{EC p2.xlarge 10,000-bits}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 17.13 & 58.34 \\
+    Thread scan & 5000 & 60.98 & 16.39 \\
+    OpenCL scan & 5000 & 1175.88 & 0.85 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 1000 & 35.12 & 28.47 \\
+    Thread single read & 1000 & 57.31 & 17.44 \\
+    OpenCL write & 1000 & 113.98 & 8.77 \\
+    OpenCL single read & 1000 & 549.12 & 1.82 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 56.80 & 17.60 \\
+    Thread scan & 5000 & 192.54 & 5.19 \\
+    OpenCL scan & 5000 & 1565.28 & 0.63 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 2000 & 131.64 & 7.59 \\
+    Thread single read & 2000 & 161.97 & 6.17 \\
+    OpenCL write & 2000 & 427.38 & 2.33 \\
+    OpenCL single read & 2000 & 987.95 & 1.01 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 100 & 1.84 & 540.97 \\
+    Thread scan & 500 & 5.03 & 198.74 \\
+    OpenCL scan & 1000 & 174.05 & 5.74 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 500 & 4.50 & 222.08 \\
+    Thread single read & 500 & 6.89 & 145.01 \\
+    OpenCL write & 1000 & 12.42 & 80.48 \\
+    OpenCL single read & 1000 & 71.46 & 13.99 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.}
 \end{table}
@@
 
-Deep neural networks, alphaGo, alphazero
-
 \subsection{i versus l}
@@
 The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
-A little more on this in my closing words.
+More on this in my closing words.
 
@@
 
-The overarching intention here is to not only provide a starting point, but to provide a Framework in which SDM research can be conducted.
-Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).
-For example, some of the benchmarks that we plan to develop in future research is: how fast is convergence through iterative reading?
-How large is the attractor of the critical distance?
-How well does the system filter noise?
-How well does the system work under the supervised learning task?
-And other authors may be able to improve this benchmark set themselves, as is usual in open source development.
-It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.
-
-Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
-We have referred to this modified read operation as Chada read (the tale is that my friend \text{\&} colleague, Dr.
-Daniel de Magalhães Chada, along with Linhares, did not consult Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterwards).
-Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
-Indeed, a high-quality codebase seems to have become a scientific community's form of unequivocally standing behind a consensus.
-For example, the journal Nature analysed the top-100 cited papers in history, to find:
+It is no coincidence that a scientific journal such as Neurocomputing states that “software is scientific method by machine”.
+Indeed, recently, Nature analysed the top-100 cited papers in history, to find
 
@@
 A few that do, such as the first observation of carbon nanotubes (number 36) are indeed classic discoveries.
-But the vast majority describe experimental methods or \emph{software that have become essential in their fields.
-[...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.} \\
-\hfill --- \citet{van2014top}, emphasis mine.
+But the vast majority describe experimental methods or software that have become essential in their fields.
+[...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.
+\\
+\hfill --- \citet{van2014top}
 \end{quote}
 
-It is no coincidence that scientific journals such as BMC Neuroscience, or the Journal of Machine Learning Research have specific sections on open-source software.
-Neurocomputing states that “software is scientific method by machine”.
-
-Of course, for the skeptical reader who may consider software a less worthy pursuit, there is also new work here.
-The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.
-The emphasis might have been on \emph{breadth of topics}, in detriment of depth here or there.
-But this is due to our research group's enthusiam for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.
-If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.
-The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
-Some essays completely shut the door close at the end; this one intends to leave it wide open.
-As the reader may have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
-
-\begin{figure}[!p]
-\centering
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}
-
-\caption{Some slides from an old Linhares class \citet{only-one-life}.
-In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
-It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
-Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
-\label{fig:one-life}}
-\end{figure}
+The overarching intention here is to not only provide a starting point, but to provide the Framework in which SDM research can be conducted.
+Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous best, under a particular benchmark set.
+For example, some of the benchmarks that we plan to develop for future research is: how fast is convergence through iterative reading?
+How large is the attractor of the critical distance?
+How well does the system filter noise?
+How well does the system work under the aforementioned supervised learning task?
+And other authors may be able to improve this benchmark set themselves, as is usual in open source development.
+It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.
+
+Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
+We have referred to this modified read operation as Chada read.
+Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
+
+Of course, there is also new work here.
+The mathematics of the model has been shown to be correct (with a single exception); we have shown how to execute unsupervised learning with nothing besides operations originally proposed by Kanerva; we have studied the generalized Murilo read; we have seen and finally, we have reproduced
+
+\subsection{From theory to a platform.}
+
+Platforms in the history of computing.
+
+Opens 10 doors.
 

### 2018-03-09 Chapters/Introduction.tex

@@
 \noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.
+
+\begin{figure}[!p]
+\centering
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}
+
+\caption{Some slides from an old Linhares class; a viewpoint that has influenced the choice of topics found in this thesis.
+In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
+It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
+It is, in a sense, the Big Drama of our moment in history.
+Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
+Note that the slides were made in 2007, and expected an AI-based `semantic revolution` by 2015 \citep{only-one-life}.
+In between, both the iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created (Uber, Instagram, etc, on the case of smartphones; and exchanges, miners, investors, payment processors, etc., in the case of blockchains).
+Linhares's thesis is that students should (i) expect new, unexpected, technological platforms, (ii) rapidly identify them, and (iii) throw their lifes at them, for they offer leverage to make an asymmetric impact.
+\label{fig:onelife}}
+\end{figure}
 
@@
 Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power to the hands of individuals.
-Whilst, during the war, computations were mostly done by hand, the electronic computer took over afterwards; up to an extent that it is not outlandish to say that this original field can be referred to, today, as computational management science.
+Whilst, during the war, computations were mostly done by hand, the electronic computer took over afterwards; up to an extent that it is not outlandish to say that this original field can be referred to, today, as \emph{computational management science}.
 
@@
 
-This thesis will present a number of different topics for exploration.
+This thesis will explore three different topics (inspired by Figure \ref{onelife}).
 Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three specific, self-contained, scientific papers.

### 2018-03-09 Chapters/sdm-applications.tex
SDM

@@
 
-\chapter{Results (iv): Supervised classification application}
+\chapter{Results (v): Supervised classification application}
 
@@
 
+In our simplified scenario, there is neither translation nor rotation.
+Future work may explore how sensible this classification algorithm is to these operations.
+We expect that with proper training, the algorithm will remain classifying the images with a good hit rate.
+
 These results show that the SDM may be used as a supervised classification algorithm.
@@
 
-\chapter{Results (iii): Supervised image noise filtering application}
+\chapter{Results (vi): Supervised image noise filtering application}
 
@@
 The chunk was calculated using the exclusive or (XOR) operator.
-So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- write(addr=bs\_image XOR bs\_label, datum=bs\_image).
+So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- write(addr=bs\_image $\oplus$ bs\_label, datum=bs\_image).
 
@@
 
-\chapter{Results (v): The possibility of unsupervised reinforcement learning}
+\chapter{Results (vii): The possibility of unsupervised reinforcement learning}
 

### 2018-03-09 Chapters/Chapter02.tex
notebook trying to replicate Kanerva's calculation of critical distance

--- 2018-03-08 Chapters/Chapter02.tex
@@
 
-\begin{figure}[!p]
-\centering
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life077.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life078.jpeg}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life079.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life080.jpeg}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life081.jpeg}}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{images02/new-images/Only-one-life082.jpeg}}
-
-\caption{Some slides from an old Linhares class; a viewpoint that has influenced the choice of topics found in this thesis.
-In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
-It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
-It is, in a sense, the Big Drama of our moment in history.
-Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
-Note that the slides were made in 2007, and expected an AI-based `semantic revolution` by 2015 \citep{only-one-life}.
-In between, both the iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created (Uber, Instagram, etc, on the case of smartphones; and exchanges, miners, investors, payment processors, etc., in the case of blockchains).
-Linhares's thesis is that students should (i) expect new technological platforms, (ii) rapidly identify them, and (iii) skydive, truly throw their lifes, at them.
-It is there that we might find leverage to make a small impact upon the world.
-\label{fig:one-life}}
-\end{figure}
-
 Let us revisit, in these concluding thoughts, the emphasis employed over speed of computation.
@@
 The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
-Ralph Waldo Emerson once said \emph{do not go where the path may lead, go instead where there is no path, and leave a trail.}  Dr.
+Ralph Waldo Emerson once said \emph{Do not go where the path may lead.
+Go, instead, where there is no path, and leave a trail.}  Dr.
 Kanerva has left the trail.
-It is my job to pave it and throw light at it and try to deliver an easier pathway for the next generation.
+It is my job to pave it and to throw light at it and to try to deliver an easier pathway for the next generation.
 Some essays completely shut the door close at the end; this one intends to leave it wide open.
-As the reader might have noticed, this final section does not read as an analysis of work previously done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
+As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
 

### 2018-03-09 Chapters/Chapter02.tex

@@
 
+Kernel comparisons
+
+iMac 1000 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	3000	15.028241872787476	199.62414934459346	5.009413957595825
+single_scan1	3000	11.874222040176392	252.6481305343213	3.958074013392131
+single_scan2	3000	13.736932039260864	218.38937482007222	4.578977346420288
+single_scan3	3000	11.063963174819946	271.1505771121497	3.687987724939982
+single_scan4	3000	12.131728887557983	247.2854469305467	4.043909629185994
+single_scan5	3000	11.181737899780273	268.29460920014515	3.7272459665934243
+single_scan5_unroll	3000	10.841995000839233	276.7018431356759	3.6139983336130777
+single_scan6	3000	12.084025859832764	248.26163356468672	4.028008619944255
+
+iMac 256 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	3000	9.098679065704346	329.7181907764943	3.0328930219014483
+single_scan1	3000	8.634926080703735	347.4262514770137	2.8783086935679116
+single_scan2	3000	11.461902141571045	261.7366614149787	3.8206340471903486
+single_scan3	3000	11.187175273895264	268.1642082608964	3.729058424631755
+single_scan4	3000	13.442016124725342	223.18080652215394	4.480672041575114
+single_scan5	3000	11.295457601547241	265.59348950936374	3.76515253384908
+single_scan5_unroll	3000	11.396361827850342	263.24190520773243	3.798787275950114
+single_scan6	3000	13.104900121688843	228.92200414675017	4.368300040562947
+
+iMac 10k bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	500	30.534967184066772	16.374669636484867	61.069934368133545
+single_scan1	500	22.480546236038208	22.24145244293299	44.961092472076416
+single_scan2	500	22.491918802261353	22.23020651976254	44.983837604522705
+single_scan3	500	6.335015773773193	78.92640174156905	12.670031547546387
+single_scan4	500	5.727261781692505	87.30175414685539	11.45452356338501
+single_scan5	500	6.294461011886597	79.43491890025042	12.588922023773193
+single_scan5_unroll	500	5.688975811004639	87.88928211521136	11.377951622009277
+single_scan6	500	5.48170280456543	91.21253337258189	10.96340560913086
+
 \begin{figure}[!htb]

### 2018-03-09 Chapters/Conclusion.tex
small details

@@
 The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer \citep{TonnyMartins}.
-Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citet{AndrewNgAI}, as his way to emphasize the potential transformational power of the technology.
+Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citep{AndrewNgAI}, as his way to emphasize the potential transformational power of the technology.
+
 It is not impossible that a purely digital form of money may exist.
@@
 Consider, just as a matter of comparison, Brazil's most important company: Petrobrás.
-As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobras by ten billion dollars, while Bitcoin's is valued at more than double of Petrobras (195B usd vs 83B usd).
+As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobrás by ten billion dollars, while Bitcoin's is valued at more than double of Petrobrás (195B usd vs 83B usd).
 These technologies should, at a minimum, be taken seriously.

### 2018-03-09 Chapters/Introduction.tex

@@
 
-\caption{Some slides from an old Linhares class; a viewpoint that has influenced the choice of topics found in this thesis.
+\caption{Slides from an old Linhares class; a viewpoint that has influenced the choice of topics found in this thesis.
 In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
 It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
-It is, in a sense, the Big Drama of our moment in history.
+It is, in a sense, a major part of the Big Drama of our moment in history.
 Each time a new computing platform appears, it seems as the opportunities are ripe for the taking; as if a multitude of doors have opened simultaneously.
@@
 In between, both the iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created (Uber, Instagram, etc, on the case of smartphones; and exchanges, miners, investors, payment processors, etc., in the case of blockchains).
-Linhares's thesis is that students should (i) expect new, unexpected, technological platforms, (ii) rapidly identify them, and (iii) throw their lifes at them, for they offer leverage to make an asymmetric impact.
+Linhares's thesis is that students should (i) expect new, unexpected, technological platforms, (ii) rapidly identify them, and (iii) throw their energy at them, as they offer leverage to make an asymmetric impact.
 \label{fig:onelife}}
@@
 
-This thesis will explore three different topics (inspired by Figure \ref{onelife}).
+This thesis will explore three different topics related to \emph{computing business platforms} (Figure \ref{onelife}).
 Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three specific, self-contained, scientific papers.

### 2018-03-09 Chapters/Chapter02.tex
SDM minor changes

@@
 
-Performance is an important part of our framework. Running an SDM memory consumes a lot of resource from both CPU and RAM.
-
-For instance, each scan in 1,000,000 hard-locations with 1,000 bits executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
+Performance matters --- and has always mattered.
+If an experiment takes a few seconds, there is no point arguing whether we should try it.
+If it takes a few hours, maybe we should think first.
+If it takes a few days, it is more important to make a good plan.
+As SDM consumes a lot of both processor and memory resources, some experiments may take a long time.
+
+Each scan a 1,000-bits SDM with 1,000,000 hard-locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
 So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
@@
 There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.
+A modern laptop can increment a counter $5 \cdot 10^{8}$ times per second.
+Hence, a naive implementation of SDM may take several hours to simply write 10,000 random bitstrings.
 
@@
 It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.
+It is a great improvement when compared to the first versions of the code which took 16 hours to generate the same heatmap (and its memory use was already optimized and the computations were distributed in threads).
 
@@
 single_scan6	500	5.48170280456543	91.21253337258189	10.96340560913086
+
+MacBook 1000 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	1000	23.6049609184	42.363976092	23.6049609184
+single_scan1	1000	13.2222340107	75.6301846716	13.2222340107
+single_scan2	1000	47.2828419209	21.1493209667	47.2828419209
+single_scan3	1000	33.0640568733	30.2443225231	33.0640568733
+single_scan4	1000	40.3230810165	24.7996922554	40.3230810165
+single_scan5	1000	31.5100109577	31.7359458028	31.5100109577
+single_scan5_unroll	1000	27.5569870472	36.288437422	27.5569870472
+single_scan6	1000	39.4823219776	25.3277910192	39.4823219776
+
+MacBook 256 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	1000	8.36867690086	119.493202073	8.36867690086
+single_scan1	1000	10.4354770184	95.8269562801	10.4354770184
+single_scan2	1000	23.4823930264	42.5850976465	23.4823930264
+single_scan3	1000	25.5149869919	39.1926517665	25.5149869919
+single_scan4	1000	42.3955621719	23.5873744508	42.3955621719
+single_scan5	1000	24.4205920696	40.9490481291	24.4205920696
+single_scan5_unroll	1000	22.7717738152	43.9140142581	22.7717738152
+single_scan6	1000	42.1847259998	23.7052624214	42.1847259998
+
+EC2 p3.2xlarge 1000 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	3000	2.09311103821	1433.27322117	0.697703679403
+single_scan1	3000	1.64607286453	1822.51956438	0.548690954844
+single_scan2	3000	2.57495617867	1165.06837082	0.858318726222
+single_scan3	3000	3.06142997742	979.934220978	1.02047665914
+single_scan4	3000	3.30994796753	906.358658635	1.10331598918
+single_scan5	3000	2.8592748642	1049.21707163	0.953091621399
+single_scan6	3000	3.12985897064	958.509641533	1.04328632355
+
+EC2 p3.2xlarge 256 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	3000	1.09906792641	2729.58561334	0.366355975469
+single_scan1	3000	1.10967302322	2703.49908236	0.369891007741
+single_scan2	3000	2.21544480324	1354.12987749	0.738481601079
+single_scan3	3000	1.9066619873	1573.43043496	0.635553995768
+single_scan4	3000	3.15382003784	951.227389009	1.05127334595
+single_scan5	3000	1.87652993202	1598.69552241	0.625509977341
+single_scan6	3000	3.04240322113	986.062589983	1.01413440704
+
+EC2 p3.2xlarge 10k bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	500	10.3001351357	48.5430524372	20.6002702713
+single_scan1	500	2.47107696533	202.340925441	4.94215393066
+single_scan2	500	2.5089058876	199.290058057	5.01781177521
+single_scan3	500	3.0364408493	164.666471311	6.07288169861
+single_scan4	500	2.99925804138	166.707896787	5.99851608276
+single_scan5	500	2.68434309959	186.265310152	5.36868619919
+single_scan6	500	2.98077511787	167.741604189	5.96155023575
+
+EC2 p2.xlarge 1000 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	3000	11.3801429272	263.617075743	3.79338097572
+single_scan1	3000	11.8365280628	253.452700325	3.94550935427
+single_scan2	3000	25.6270592213	117.063763505	8.54235307376
+single_scan3	3000	29.2102689743	102.70360751	9.73675632477
+single_scan4	3000	34.1599340439	87.8221836186	11.3866446813
+single_scan5	3000	29.8609650135	100.465607814	9.9536550045
+single_scan6	3000	34.0075969696	88.2155832028	11.3358656565
+
+EC2 p2.xlarge 256 bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	3000	2.28510403633	1312.85051022	0.761701345444
+single_scan1	3000	2.42481589317	1237.20733126	0.808271964391
+single_scan2	3000	16.7731001377	178.857812531	5.59103337924
+single_scan3	3000	18.9349951744	158.436797705	6.31166505814
+single_scan4	3000	30.8822939396	97.1430427373	10.2940979799
+single_scan5	3000	20.0731289387	149.453531094	6.69104297956
+single_scan6	3000	30.8959820271	97.1000046988	10.2986606757
+
+EC2 p2.xlarge 10k bits
+Loops	Total time	Scans per second	Time per scan (ms)
+single_scan0	500	17.7273669243	28.204978333	35.4547338486
+single_scan1	500	28.9013559818	17.3002263394	57.8027119637
+single_scan2	500	31.8554549217	15.6958989042	63.7109098434
+single_scan3	500	19.9614479542	25.0482831279	39.9228959084
+single_scan4	500	22.749904871	21.9781138794	45.499809742
+single_scan5	500	21.7587711811	22.9792388476	43.5175423622
+single_scan6	500	20.7129361629	24.1395037414	41.4258723259
 
@@
 It is no coincidence that scientific journals such as BMC Neuroscience, or the Journal of Machine Learning Research have specific sections on open-source software.
-The journal Neurocomputing even goes as far as to state that “software is scientific method by machine”.
+The journal Neurocomputing states, bluntly: “software is scientific method by machine”.
 

### 2018-03-09 Chapters/sdm-applications.tex
SDM minor changes

@@
 
-\chapter{Results (vi): Supervised image noise filtering application}
+\chapter{Results (vi): Image noise filtering application}
 
@@
 This is a bijective mapping (or one-to-one) from images and bitstrings, i.e., there is one, and only one, bitstring for each image, and vice versa.
+
+In the learning phase, 200 noisy images were generated and written into SDM.
+Half of the letter ``I'' and half of the letter ``T'' (see Figure \ref{fig:filter-training}).
+They were written in their own addresses, i.e., write(address=bs\_image, datum=bs\_image).
+
+\begin{figure}[!htb]
+  \subfloat[Letter ``I'' with 15\% of noise. ]{\includegraphics[width=\textwidth]{./images02/filter/training-i.png}}
+
+  \subfloat[Letter ``T'' with 15\% of noise. ]{\includegraphics[width=\textwidth]{./images02/filter/training-t.png}}
+
+  \caption{Training images written into the SDM. They were written in their own addresses --- write(address=bs\_image, datum=bs\_image).}
+  \label{fig:filter-training}
+\end{figure}
 

### 2018-03-10 Chapters/Chapter02.tex
SDM Splitting into multiple files.

--- 2018-03-09 Chapters/Chapter02.tex
@@
 \chapter{Sparse Distributed Memory}
-
 \input{./Chapters/sdm2.tex}
@@
 \chapter{Framework Architecture}
-
-The framework implements the basic operations in a Sparse Distributed Memory which may be used to create more complex operations.
-It is developed in C language and the OpenCL parallel framework --- which may be loaded in many platforms and programming languages --- with a wrapper in Python.
-The Python module makes it easy to create and execute simulations in a Sparse Distributed Memory and works properly in Jupyter Notebook \citep{kluyver2016jupyter}.
-It works in both Python 2 and Python 3.
-
-We split the SDM memory in two parts: the hard-location addresses and the hard-location counters.
-Thus, the addresses (bitstrings) of the hard-locations are stored in one array, while their counters in another.
-This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be the same in all of them.
-As the slowest part of reading and writing operations is scanning the address space, the performance benefits are significant.
-
-Each part may be stored either in the RAM memory or in a file.
-The RAM memory is interesting for quick experiments, automated tests, and others scenarios in which the SDM may be lost, while the file is interesting for a long-term SDM, like creating an SDM file with 10,000 random writes, which will be copied over and over to run multiple experiments.
-The file may also be sent to another researcher or may be published within the paper to let others run their own checks and verify the results.
-In summary, the framework fits many different uses and necessities.
-
-Let a SDM memory with $N$ dimensions and $H$ hard-locations.
-Then, in a 64-bit computer, the array of hard-location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard-location counters.
-For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard-locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.
-
-Basic operations were grouped in four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs).
-Operations include creating new bitstrings, flipping bits, generating a bitstring with a specific distance from a given bitstring, scanning the address space using different algorithms, writing a bitstring to a counter, writing in an SDM, reading from an SDM, and iteratively reading from an SDM until convergence.
-
-\section{Bitstring}
-
-Bitstrings are the main structure of SDM.
-The addresses are represented in bitstrings, as well as the data.
-A bitstring is stored as an array of integers.
-Each integer may be 16-bit, 32-bit, or 64-bit long, depending on the configuration.
-By default, each integer is 64-bit long.
-
-For instance, a 1,000-bit bitstring will have $\lceil 1000/64 \rceil = 16$ integers.
-These integers will have a total of $16 \cdot 64 = 1,024$ bits.
-The remaining 24 bits are always zero, so they do not affect the result of any operation.
-The memory usage efficiency is $1 - 24/1024 = 97.65\%$.
-Bitstrings store neither how many bits they have nor the array length.
-These pieces of information are only stored in the address space.
-
-\subsection{The distance between two bitstrings}
-
-The distance between two bitstrings is calculated by the hamming distance, which is the number of different bits between them.
-It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in } x \oplus y$.
-
-There are several algorithms to calculate the number of ones \citep{warren2013hacker}, but the performance depends on the processor.
-So, we have implemented three different algorithms and one may be selected through compiling flags.
-The default algorithm is to use a built-in \_\_popcnt() instruction from the compiler.
-
-There is also the naive algorithm, which really counts the number of ones checking bit by bit. It is available only to testing purposes and should never be used.
-
-The other algorithm available is the lookup.
-It pre-calculates a table with the number of ones of all possible 16-bit integers.
-This table is accessed a few times to calculate the number of ones of a 64-bit integer, i.e., to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:63], y[0:63]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + d(x[32:47], y[32:47]) + d(x[48:63], y[48:63])$ where $x[i:i+15]$ and $y[i, i+15]$ are the 16-bit integers formed by the bits between $i$ and $i+15$ of $x$ and $y$, respectivelly.
-Each 16-bit distance is calculated through a single table access.
-As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$).
-This table uses 65MB of RAM.
-One may change the table from 16-bit integers to 32-bit integers, which would halve the number of accesses at the expense of 4GB of RAM (instead of 65MB).
-
-\section{Address space}
-
-An address space is a fixed collection of bitstrings, and each bitstring represents a hard-location address.
-They store the number of bitstrings, as well as the number of bits, number of integers per bitstring, and the number of remaining bits.
-
-Bitstrings are stored in a contiguous array of 64-bit integers, as shown in Figure \ref{tab:hl-addresses-detail}.
-Hence, basic pointer arithmetic provides us with performance improvements in their access, as processors realize fetches of contiguous chunks of memory  \citep{pai2004linux}.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}[
-mycell/.style={draw, minimum size=7mm},
-matrixA/.style={matrix of nodes,
-    nodes={mycell, anchor=center},
-    column sep=-\pgflinewidth,
-    row sep=-\pgflinewidth,
-    },
-matrixB/.style={matrix of nodes,
-    nodes={mycell, anchor=center},
-    column sep=-\pgflinewidth,
-    row sep=-\pgflinewidth,
-}]
-
-\matrix[matrixA] (A) { addr$_1$ & addr$_2$ & addr$_3$ & $\cdots$ & addr$_H$ \\ };
-
-\matrix[matrixB, below=of A] (B) {
-addr$_{k, 1}$ & addr$_{k, 2}$ & addr$_{k, 3}$ & $\cdots$ & addr$_{k, 8 \cdot \lceil N/64 \rceil}$ \\
-};
-
-\draw[dashed] (A-1-1.south west)--(B-1-1.north west);
-\draw[dashed] (A-1-1.south east)--(B-1-5.north east);
-\draw [
-	thick,
-    decoration={
-        brace,
-        mirror,
-		amplitude=0.2cm,
-        raise=0.2cm
-    },
-    decorate
-] (B-1-1.south west) -- (B-1-5.south east)
-node [pos=0.5,anchor=north,yshift=-0.5cm] {N bits};
-
-\end{tikzpicture}
-
-\caption{Address space's bitstrings are stored in a contiguous array.
-In a 64-bit computer, each bitstring is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil N/64 \rceil$.\label{tab:hl-addresses-detail}}
-\end{figure}
-
-The scan for activated hard-locations is performed in an address space.
-It returns the indexes of the bitstrings which were inside the circle (and their distances).
-Then, each operation uses these pieces of information in a different way.
-
-\subsection{Scanning for activated hard-locations}
-
-Scanning for the activated hard-locations is a problem similar to well-known problems in computational geometry called ``range reporting in higher dimensions''.
-In this case, none of the known algorithms is able to solve our problem faster than $O(H)$.
-The algorithm which seems to best fit in our problem consumes $O(H)$ space and runs in $O(\log^n(H))$ \citep{chazelle1988functional}, which is really slower than $O(H)$ when, for instance, $H=1,000,000$ and $n=1,000$.
-For a review of the range reporting algorithms, see \citet{chan2011orthogonal}.
-
-In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}.
-It provides a fast search when $r/n < 0.11$ or $r/n < 0.06$, where $r$ is the radius and $n$ is the number of bits.
-But, in our case, for a 1,000 bits SDM, $r/n = 0.451$, which changes the runtime to $O(H^{0.993})$.
-This is really close to $O(H)$, but with a larger constant.
-Unfortunately, $O(H)$ is still faster.
-
-It is intriguing that none of those algorithms is able to solve our scanning problem.
-The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(H))$ to go through the whole space.
-But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hard-locations' addresses are randomly sampled from the $\{0, 1\}^n$ space.
-Although each addresses' bit itself splits the hardlocations in half, it does not split the search space in half since both halves still must be convered by the algorithm.
-For instance, let's say we have $n=1,000$ dimensions with $H=1,000,000$ hard-locations, and we are scanning within a circle with radius $r=451$, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
-Hence, the search space has not been split in half because both halves have been covered (and one of them should have been skipped).
-
-Finally, as our best approach is to scan through all hard-locations, we may distribute the scan into many tasks which will be executed independently.
-The tasks may be executed in different processes, threads, or even computers.
-They may also run in the CPU or in the GPU.
-In this case, we may take into account both the time required to distribute the tasks and the time to receive their results.
-
-The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner.
-The linear scanner runs in a single core, is the slowest one, and was developed only for testing purposes; the thread scanner runs at the CPU in multiple threads sharing memory (and our recommendation is to use the number of threads equals to twice the number of CPU cores); and the OpenCL scanner runs in multiple GPU cores and support multiple devices.
-The speed of a scan depends on the CPU and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark.
-
-The OpenCL must be initialized, which just copies the address space's bitstrings to the GPU's memory.
-Then, many scans may be executed with no necessity to upload the bitstrings again.
-The OpenCL scanner supports running into multiple devices.
-
-\subsection{OpenCL kernel}
-
-The OpenCL kernel explores the GPU architecture to improve performance.
-Each work group calculates the distance of several bitstrings.
-During the distance calculation, each worker calculates the exclusive OR (XOR) between two 64-bit integers and use the built-in popcount function to count the number of ones.
-Then, they update an array of intermediate distances with their partial distances.
-This array is stored in the local memory and is shared with all workers of the same group.
-This whole step happens simultaneously in the GPU.
-Then, we used a reduction algorithm to sum the intermediate distances array in order to calculate the correct distance.
-This reduction algorithm is also distributed between the workers and runs in $O(\log_2(\text{bs\_step}))$.
-Finally, one of the workers checks whether the distance is less than or equal to the radius to include the bitstring index into the resulting array.
-
-The number of workers in each group is the closest power of two above bs\_len (which is the number of 64-integers that forms a bitstring).
-
-\section{Counters}
-
-Each hard-location has one integer of data per bit. For instance, each hard-location of a 1,000 bits SDM has 1,000 bits. Those integers are stored in a counter.
-
-A counter is an array of integers which stores the data of all hard-locations. So, the counter's array has $n \cdot H$ integers.
-
-When two counters are added in a third counter, there may occur an overflow.
-It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with less writes than $2^{31}-1$ divided by the average number of activated hard-locations.
-For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.
-Note also that it would be more likely to saturate the memory before any overflow.
-
-Anyway, counters may have overflow protection depending on compiling options. By default, there is no overflow check for performance reasons (and because it does not seem necessary).
-
-\section{Read and write operations}
-
-The reading and writing operations are executed in two steps: first, the address space is swept looking for the activated addresses; then, the operation is performed in the counters.
-Reading operation assemblies the bitstring according to the counters of the activated addresses, while the writing operation changes the counters.
-
-The iterated reading keeps reading until it gets exactly the same bitstring (or the number of maximum interations has been reached), i.e., it performs $\eta_{i+1} = \text{read}(\eta_i)$ and stops when $\eta_{k+1} = \eta_{k}$.
-If the initial bitstring is inside the critical distance of $\eta$, it will converge to $\eta$, but, if it is not, it will diverge and reach the maximum number of iterations.
-
-The framework has both Kanerva's read and Murilo's generalized read.
-The generalization brings a parameter $z$, which is the exponent.
-In this case, the results are floating point instead of integer, which considerably reduces performance.
-When $z=1$, it is exactly as the Kanerva's read.
-When $z=0$, it is the Chada's read.
-We also explored how SDM would behave for different values of $z$.
-
-There is another special read operation: the weighted reading.
-In the weighted reading, the value of the counters are multiplied by a weight which depends only on the distance between the reading address and the hard-location address.
-The weight is retrieved from a lookup table of integers indexed by the distance.
-The rest of the read operation is exactly the same.
-
-There is also a weighted writing operation.
-In this case, the weight is applied when the counters are updated, i.e., if the weight is 2, the counters are increased twice when bits are 1, and decreased twice when bits are 0.
-Just as in the weighted reading, the weights depend only on the distance between the writing address and the hard-location address.
-The weights are retrieved from a lookup table of integers indexed by the distance.
+\input{./Chapters/sdm-architecture.tex}
 
 \chapter{Results (i): Framework Validation}
-
-The framework has been validated comparing its results with the expected results from \citet{Kanerva1988}.
-Thus, we run simulations which were then compared to the theoretical analysis conducted some decades ago.
-
-\section{Distance between random bitstrings}
-
-As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
-For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
-
-In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits.
-In total, 20,000 random bitstrings have been generated during the simulation.
-The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.
-
-In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
-
-This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.
-
-\begin{figure}[!htb]
-  \centering
-  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
-  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}
-
-  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
-  \label{fig:validation-distance}
-\end{figure}
-
-\section{Number of activated hard-locations}
-
-In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard-locations in a 1,000 bits SDM.
-He also proposed to activate only 1,000 of them, on average.
-He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard-locations.
-
-We extended his results, calculating the distribution of the number of activated hard-locations.
-As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$.
-In this case, $\mu = 1071.85$ and $\sigma = 32.72$.
-
-In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard-locations.
-The code is available in the ``Number of activated hard-locations'' notebook \citep{sdmframework}.
-
-In figure \ref{fig:validation-activated-hls}, we can notice that the theoretical model and the simulation matches.
-Hence, it seems that both the address space generation algotihm and the scan algorithm work properly.
-Notice that the curve is almost the same for $n=1,000$ and $n=256$.
-It happens because the access radius is adjust to have $p$ as close as possible to $0.001$.
-They are not exactly the same because their $p$ differs a little.
-
-\begin{figure}[!htb]
-  \centering
-  \subfloat[$n=1,000$, $H=1,000,000$,\protect\\ $r=451$, and $p=0.00107185$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-1000.png}}
-  \subfloat[$n=256$, $H=1,000,000$,\protect\\ $r=103$, and $p=0.00106684$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-256.png}}
-
-  \caption{Histogram of the number of activated hard-locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
-  \label{fig:validation-activated-hls}
-\end{figure}
-
-Besides the number of activated hard-locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard-locations.
-Let $A$ be the set of activated hard-locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:
-
-\begin{align}
-P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(a \in A)} \\
-    &= \frac{\binom{n}{x}}{\sum_{k=0}^r \binom{n}{k}} \label{eq:prob-d-inside-circle}
-\end{align}
-
-In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard-locations to the center of 1,000 random circles.
-The code is available in the ``Distances of activated hard-locations'' notebook \citep{sdmframework}.
-
-In figure \ref{fig:validation-distance-activated-hls}, we can notice that the theoretical model and the simulation matches.
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{./images02/new-images/distance-activated-hls.png}
-
-\caption{Histogram of the distances of activated hard-locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
-\label{fig:validation-distance-activated-hls}}
-\end{figure}
-
-\section{Intersection of two circles}
-
-Kanerva has calculated the intersection of two circles according to the distance between their centers.
-The intersection is important to understand how SDM works, because it affects directly the critical distance.
-When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
-In fact, it converges because they share a sufficient amount of hard-locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge.
-For further information about the relation between the critical distance and the intersection, see \citet{brogliato2014sparse}.
-
-We have calculated the intersection between a random bitstring (bs1) and another bitstrings (bs2) exactly $d$ bits away.
-The former (bs1) is just a random bitstring.
-The latter (bs2) was generated randomly flipping $d$ bits of bs1.
-The code is available in the ``Kanerva's Figure 1.2'' notebook \citep{sdmframework}.
-
-In Figure \ref{fig:validation-intersection}, we can notice that we have obtained the same results as Kanerva. It seems that the random flipping bits algorithm and the scan algorithm work properly.
-
-\begin{figure}[!htb]
-  \centering
-  \subfloat[{\citet[Figure 1.2, p.25]{Kanerva1988}} ]{\includegraphics[width=0.45\textwidth]{./images02/new-images/kanerva-table-12.jpg}}
-  \subfloat[Generated by SDM-Frameworkm with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}
-
-  \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
-  \label{fig:validation-intersection}
-\end{figure}
-
-\section{Storage and retrieval of sequences}
-
-\citet[Ch.8]{Kanerva1988} presented an approach to store and retrieve sequences using $k$ different SDMs, namely $\text{sdm}_1$, $\text{sdm}_2$, $dots$, $\text{sdm}_k$.
-
-Let $a_0, a_1, a_2, \dots, a_n$ be a sequence to be stored in a $k$-fold memory.
-So, all pointers of the form $a_i \rightarrow a_{i+k}$ will be written to $\text{sdm}_k$ memory, i.e., in $\text{sdm}_1$, the following pointers will be written: $a_0 \rightarrow a_1$, $a_1 \rightarrow a_2$, $\dots$, $a_{n-1} \rightarrow a_n$; while in $\text{sdm}_2$, the following pointers will be written: $a_0 \rightarrow a_2$, $a_2 \rightarrow a_3$, $\dots$, $a_{n-2} \rightarrow a_n$; and so forth.
-
-We have tested exactly the same example presented in \citet{Kanerva1988}, p.85.
-We wrote two sequences to a $3$-fold memory: $<A, B, C, D>$ and $<E, B, C, F>$.
-Then, after reading the sequences $<A, B, C>$ and $<E, B, C>$, we have obtained $D$ and $F$, respectively.
-
-Each reading operation was performed summing the counters of all activated hard-locations from all three memories.
-For instance, to read the sequence $<A, B, C>$, we have activated the hard-locations around $C$ in $\text{sdm}_1$, we have also activated the hard-locations around $D$ in $\text{sdm}_2$, and, finally, we have also activated the hard-locations around $A$ in $\text{sdm}_3$.
-After summing the counters of all those hard-locations, we evaluate the resulting bitstring just as in the original read operation.
-
-The code is available in the ``Sequences (Kanerva Ch 8)'' notebook \citep{sdmframework}.
-
-The logic behind how it works is that, when reading the sequence $<A, B, C>$, we have $A$ pointing to $D$, while both $B$ and $C$ point to $D$ and $F$.
-Thus, $D$ appears more often than $F$ and ended up being the result.
-
-Hence, as we have replicated the theoretical results from Kanerva, we have one more evidence that our framework works properly.
-
-\subsection{$k$-fold memory using only one SDM}
-
-We have extended Kanerva's ideas to be able to store and retrieve sequences in $k$-fold memories using only one SDM (instead of $k$ SDMs).
-
-Our idea was to create $k$ random bitstrings, one for each fold.
-We have performed writing and reading exactly as Kanerva's original idea, but, instead of writing to $\text{sdm}_k$, we have written $a_{i+k}$ into the address $a_i \oplus \text{tag}_k$, and, instead of reading from $\text{sdm}_k$, we have read from address $a_i \oplus \text{tag}_k$, where $\oplus$ is the exclusive or (XOR) operator.
-
-It worked as if we had splitted SDM into $k$ regions with low intersection between two of them.
-So, as the interference is minimal, they work like independent SDMs.
-The major disadvantage of this approach is that memory capacity may be reached faster.
-
-Splitting the memory into regions may be an interesting strategy to other sorts of problems, mostly the ones which would need many SDMs and, consequently, would use a lot of RAM.
-
-\section{Convergence of $\eta_x$ to $\eta$}
-
-One particular analysis of Kanerva's interest is that of the distance read at a point $\eta_x$.
-Suppose an SDM is trying to read an item written at $\eta$, but the cues received so far lead to a point of distance $x$ from $\eta$.
-As one reads at $\eta_x$, a new bitstring $\beta$ is obtained, leading to Kanerva's question: what is the new distance from $\eta$ to $\beta$?
-Is it smaller or larger than $x$?
-That, of course, depends on the ratio between $x$ and the number of dimensions of the memory.
-
-\citet[p.70]{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Figure \ref{fig:kanerva-figure-7.3}).
-The original prediction considered that the read distance would decline when inside the critical distance and increase afterwards, converging to a \textasciitilde 500-bit distance.
-At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance bitstring.
-He analyzed specifically an SDM with 1,000 bits and 10,000 random bitstrings written into it.
-
-\begin{figure}[h]
-\centering\includegraphics[width=0.8\textwidth]{images02/kanerva-table-7-2-original.png}
-\caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
-\label{fig:kanerva-figure-7.3}}
-\end{figure}
-
-As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
-%and we propose that this is due to interaction effects between different attractors.
-
-We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
-
-In details, we have created a SDM with $n=1,000$, $H=1,000,000$, and $r=451$.
-Then, we have generated 10,000 random bitstrings and written them into the memory.
-Then, we have generated a reference bitstring (bs\_ref) and written it into the memory.
-Then, we have executed the following steps with $x$ from 0 to 1,000: (i) copy bs\_ref into a new bitstring; (ii) randomly flipped $x$ bits of the copy; (iii) read from the memory in the copy address; and (iv) stored the distance between the returned bitstring and bs\_ref.
-Finally, we have plotted Figure \ref{fig:sdm-10000w-table-7-2}.
-
-\begin{figure}[h]
-\centering
-\subfloat[1 sample for each distance $x$ \label{fig:sdm-10000w-table-7-2-1sample} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2.png}}
-\subfloat[6 samples for each distance $x$ \label{fig:sdm-10000w-table-7-2-6samples} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2-6-samples.png}}
-
-\caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
-Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
-\label{fig:sdm-10000w-table-7-2}}
-\end{figure}
-
-Figure \ref{fig:sdm-10000w-table-7-2-1sample} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance.
-So, we have changed the steps to run $k$ reads and store the average new distance.
-We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6samples}, which has a way lower noise and still holds the divergence.
-
-Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).
-
-Obviously, these small deviations from Kanerva's original theoretical predictions deserve a qualification.
-Kanerva was working in the 1980s and the 1990s, and had no access to the immense computational power that we do today.
-It is no surprise that some small interaction effects should exist as machines allow us to explore the ideas of his monumental work.
-
-But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Figure \ref{fig:sdm-100w-table-7-2-10samples}).
-This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
-In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractors.
-
-\begin{figure}[h]
-\centering
-\subfloat[100 writes \label{fig:sdm-100w-table-7-2-10samples} ]{\includegraphics[width=0.47\textwidth]{./images02/sdm-100w-table-7-2.png}}
-\subfloat[Steps of 1,000 writes \label{fig:sdm-table-7-2-steps} ]{\includegraphics[width=0.53\textwidth]{./images02/sdm-table-7-2-steps.png}}
-%\centering\includegraphics[width=\textwidth]{images02/sdm-100w-table-7-2.png}
-
-\caption{Results generated by the framework similar from Kanerva's original Table 7.2.
-Here we have a 1,000 bit, 1,000,000 hard-location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
-\label{fig:sdm-100w-table-7-2}}
-\end{figure}
-
-To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
-Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
-Thereby, in order to show the interaction effects more clearly, we changed the single read for an 15-iterative read.
-As we can see in Figure \ref{sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
-
-Hence, our understanding is that the attractors are just preventing the bitstrings to converge directly to 500-bit distance bitstrings, requiring more reading steps to do so.
-They are in other orthogonal bitstrings' critical distance, but sufficiently far not to converge in a single read.
-
-\begin{figure}[h]
-\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-15iter.png}
-\caption{This graph shows the interaction effects more clearly.
-As we change the single read to a 6-iterative read, the effect has vanished and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings.
-Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, except for the iterative read.
-\label{sdm-10000w-table-7-2-15iter}}
-\end{figure}
-
-%To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
-Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
-Thereby, in order to show the interaction effects more clearly, we wrote a handmade bitstring to the SDM which had all bits inverted in relation to the origin bitstring --- their hamming distance was equal to 1,000.
-Our handmade bitstring was acting as an opposite attractor, and one can see the accelerating effects towards convergence to both attractors: the origin and the handmade bitstrings (Fig.
-\ref{sdm-10000w-notX-table-7-2}).
-Here we had the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
-
-%\begin{figure}[h]
-%\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
-%\caption{This graph shows the interaction effects more clearly.
-As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and the opposite.
-Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
-%\label{sdm-10000w-notX-table-7-2}}
-%\end{figure}
+\input{./Chapters/sdm-validation.tex}
 
 \chapter{Results (ii): Loss of neurons}
-
 In SDM, the data is written distributed among millions of hard-locations, which theoretically gives SDM robustness against loss of neurons.
@@
 \chapter{Results (iv): Performance}
+\input{./Chapters/sdm-performance.tex}
 
-Performance is an important part of our framework. Running an SDM memory consumes a lot of resource from both CPU and RAM.
-
-For instance, each scan in 1,000,000 hard-locations with 1,000 bits executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
-So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
-The heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} executed $3.05 \cdot 10^{15}$ bit compares.
-For comparison, the number of seconds since Jesus's birth is 63,639,648,000 = $6.36 \cdot 10^{10}$.
-The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$.
-There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.
-
-Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds.
-It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.
-
-We have executed the same performance test in each device.
-The test has 3 parts.
-The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans.
-The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners.
-The code is available in the ``Performance test'' notebook \citep{sdmframework}.
-
-Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-
-Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
-The complete results are presented in Figures \ref{fig:perf-imac-read-256}, \ref{fig:perf-imac-write-256}, \ref{fig:perf-imac-read-1000}, \ref{fig:perf-imac-write-1000}, \ref{fig:perf-imac-scan-256}, \ref{fig:perf-imac-scan-1000}, and \ref{fig:perf-imac-scan-10k} and Tables \ref{tab:perf-imac-256}, \ref{tab:perf-imac-1000}, and \ref{tab:perf-imac-10k}.
-
-Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
-
-It is interesting that which scanner is faster depends also on the SDM settings.
-In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL, but for a 256-bit SDM was the threads.
-What happened here is that the OpenCL kernel chosen was a generic one which performs the scan for any SDM.
-It is always possible to optimize the OpenCL kernel to a specific setting, and it would be faster than the threads.
-By default, the framework chooses a generic kernel which we believe would be reasonable for the most common setups.
-% TODO Run the optimized kernel and show the results.
-
-Most of the time, the bottleneck of both the read and the write operations is the scanner.
-But we have optimized the OpenCL kernel so much for the iMac 2017 that scanner not the bottleneck anymore.
-In the writing operation, it took the same time to scan and to update the counters.
-It is impressive because it update, on average, 1,000 counters of 32-bit integers each, and it was as fast as (i) sending the command to the GPU, (ii) performing 1 billion bit compares, and (iii) downloading the response from the GPU.
-
-Our conclusion is that, if one is really concerned about performance, one should fine tune the OpenCL kernel for one's GPU. It would always be faster than running in the CPU.
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-256.png}
-\caption{Time to run a single read in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-read-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-256.png}
-\caption{Time to run one write in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-write-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-1000.png}
-\caption{Time to run one write in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-write-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-1000.png}
-\caption{Time to run a single read in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-read-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
-\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-scan-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
-\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-scan-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
-\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
-\label{fig:perf-imac-scan-10k}}
-\end{figure}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 81.62 & 12.25 \\
-    Thread scan & 5000 & 143.68 & 6.95 \\
-    OpenCL scan & 5000 & 238.00 & 4.20 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 1000 & 74.92 & 13.34 \\
-    Thread single read & 1000 & 96.22 & 10.39 \\
-    OpenCL write & 1000 & 126.50 & 7.90 \\
-    OpenCL single read & 1000 & 190.20 & 5.25 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-1000}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 175.48 & 5.69 \\
-    Thread scan & 5000 & 352.63 & 2.83 \\
-    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 2000 & 304.46 & 3.28 \\
-    Thread single read & 2000 & 391.21 & 2.55 \\
-    OpenCL write & 2000 & 378.44 & 2.64 \\
-    OpenCL single read & 2000 & 466.16 & 2.14 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 100 & 8.59 & 116.38 \\
-    Thread scan & 500 & 18.66 & 53.56 \\
-    OpenCL scan & 1000 & 77.20 & 12.95 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
-Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.
-There is no benchmark for read and write operations because RAM is not enough to allocate the counters --- it would consume 37.25 GB of RAM.
-\label{tab:perf-imac-10k}}
-\end{table}
-
+% Application
 \input{./Chapters/sdm-applications.tex}

### 2018-03-10 Chapters/sdm-architecture.tex
SDM Splitting into multiple files.

% !TEX root = ../partial-sdm.tex

The framework implements the basic operations in a Sparse Distributed Memory which may be used to create more complex operations. It is developed in C language and the OpenCL parallel framework --- which may be loaded in many platforms and programming languages --- with a wrapper in Python. The Python module makes it easy to create and execute simulations in a Sparse Distributed Memory and works properly in Jupyter Notebook \citep{kluyver2016jupyter}. It works in both Python 2 and Python 3.

We split the SDM memory in two parts: the hard-location addresses and the hard-location counters. Thus, the addresses (bitstrings) of the hard-locations are stored in one array, while their counters in another. This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be the same in all of them. As the slowest part of reading and writing operations is scanning the address space, the performance benefits are significant.

Each part may be stored either in the RAM memory or in a file. The RAM memory is interesting for quick experiments, automated tests, and others scenarios in which the SDM may be lost, while the file is interesting for a long-term SDM, like creating an SDM file with 10,000 random writes, which will be copied over and over to run multiple experiments. The file may also be sent to another researcher or may be published within the paper to let others run their own checks and verify the results. In summary, the framework fits many different uses and necessities.

Let a SDM memory with $N$ dimensions and $H$ hard-locations. Then, in a 64-bit computer, the array of hard-location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard-location counters. For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard-locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.

Basic operations were grouped in four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs). Operations include creating new bitstrings, flipping bits, generating a bitstring with a specific distance from a given bitstring, scanning the address space using different algorithms, writing a bitstring to a counter, writing in an SDM, reading from an SDM, and iteratively reading from an SDM until convergence.

\section{Bitstring}

Bitstrings are the main structure of SDM. The addresses are represented in bitstrings, as well as the data. A bitstring is stored as an array of integers. Each integer may be 16-bit, 32-bit, or 64-bit long, depending on the configuration. By default, each integer is 64-bit long.

For instance, a 1,000-bit bitstring will have $\lceil 1000/64 \rceil = 16$ integers. These integers will have a total of $16 \cdot 64 = 1,024$ bits. The remaining 24 bits are always zero, so they do not affect the result of any operation. The memory usage efficiency is $1 - 24/1024 = 97.65\%$. Bitstrings store neither how many bits they have nor the array length. These pieces of information are only stored in the address space.

\subsection{The distance between two bitstrings}

The distance between two bitstrings is calculated by the hamming distance, which is the number of different bits between them. It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in } x \oplus y$.

There are several algorithms to calculate the number of ones \citep{warren2013hacker}, but the performance depends on the processor. So, we have implemented three different algorithms and one may be selected through compiling flags. The default algorithm is to use a built-in \_\_popcnt() instruction from the compiler.

There is also the naive algorithm, which really counts the number of ones checking bit by bit. It is available only to testing purposes and should never be used.

The other algorithm available is the lookup. It pre-calculates a table with the number of ones of all possible 16-bit integers. This table is accessed a few times to calculate the number of ones of a 64-bit integer, i.e., to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:63], y[0:63]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + d(x[32:47], y[32:47]) + d(x[48:63], y[48:63])$ where $x[i:i+15]$ and $y[i, i+15]$ are the 16-bit integers formed by the bits between $i$ and $i+15$ of $x$ and $y$, respectivelly. Each 16-bit distance is calculated through a single table access. As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$). This table uses 65MB of RAM. One may change the table from 16-bit integers to 32-bit integers, which would halve the number of accesses at the expense of 4GB of RAM (instead of 65MB).

\section{Address space}

An address space is a fixed collection of bitstrings, and each bitstring represents a hard-location address. They store the number of bitstrings, as well as the number of bits, number of integers per bitstring, and the number of remaining bits.

Bitstrings are stored in a contiguous array of 64-bit integers, as shown in Figure \ref{tab:hl-addresses-detail}. Hence, basic pointer arithmetic provides us with performance improvements in their access, as processors realize fetches of contiguous chunks of memory  \citep{pai2004linux}.

\begin{figure}
\centering
\begin{tikzpicture}[
mycell/.style={draw, minimum size=7mm},
matrixA/.style={matrix of nodes,
    nodes={mycell, anchor=center},
    column sep=-\pgflinewidth,
    row sep=-\pgflinewidth,
    },
matrixB/.style={matrix of nodes,
    nodes={mycell, anchor=center},
    column sep=-\pgflinewidth,
    row sep=-\pgflinewidth,
}]

\matrix[matrixA] (A) { addr$_1$ & addr$_2$ & addr$_3$ & $\cdots$ & addr$_H$ \\ };

\matrix[matrixB, below=of A] (B) {
addr$_{k, 1}$ & addr$_{k, 2}$ & addr$_{k, 3}$ & $\cdots$ & addr$_{k, 8 \cdot \lceil N/64 \rceil}$ \\
};

\draw[dashed] (A-1-1.south west)--(B-1-1.north west);
\draw[dashed] (A-1-1.south east)--(B-1-5.north east);
\draw [
	thick,
    decoration={
        brace,
        mirror,
		amplitude=0.2cm,
        raise=0.2cm
    },
    decorate
] (B-1-1.south west) -- (B-1-5.south east)
node [pos=0.5,anchor=north,yshift=-0.5cm] {N bits};

\end{tikzpicture}

\caption{Address space's bitstrings are stored in a contiguous array. In a 64-bit computer, each bitstring is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil N/64 \rceil$.\label{tab:hl-addresses-detail}}
\end{figure}

The scan for activated hard-locations is performed in an address space. It returns the indexes of the bitstrings which were inside the circle (and their distances). Then, each operation uses these pieces of information in a different way.

\subsection{Scanning for activated hard-locations}

Scanning for the activated hard-locations is a problem similar to well-known problems in computational geometry called ``range reporting in higher dimensions''. In this case, none of the known algorithms is able to solve our problem faster than $O(H)$. The algorithm which seems to best fit in our problem consumes $O(H)$ space and runs in $O(\log^n(H))$ \citep{chazelle1988functional}, which is really slower than $O(H)$ when, for instance, $H=1,000,000$ and $n=1,000$. For a review of the range reporting algorithms, see \citet{chan2011orthogonal}.

In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}. It provides a fast search when $r/n < 0.11$ or $r/n < 0.06$, where $r$ is the radius and $n$ is the number of bits. But, in our case, for a 1,000 bits SDM, $r/n = 0.451$, which changes the runtime to $O(H^{0.993})$. This is really close to $O(H)$, but with a larger constant. Unfortunately, $O(H)$ is still faster.

It is intriguing that none of those algorithms is able to solve our scanning problem. The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(H))$ to go through the whole space. But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hard-locations' addresses are randomly sampled from the $\{0, 1\}^n$ space. Although each addresses' bit itself splits the hardlocations in half, it does not split the search space in half since both halves still must be convered by the algorithm. For instance, let's say we have $n=1,000$ dimensions with $H=1,000,000$ hard-locations, and we are scanning within a circle with radius $r=451$, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450. Hence, the search space has not been split in half because both halves have been covered (and one of them should have been skipped).

Finally, as our best approach is to scan through all hard-locations, we may distribute the scan into many tasks which will be executed independently. The tasks may be executed in different processes, threads, or even computers. They may also run in the CPU or in the GPU. In this case, we may take into account both the time required to distribute the tasks and the time to receive their results.

The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner. The linear scanner runs in a single core, is the slowest one, and was developed only for testing purposes; the thread scanner runs at the CPU in multiple threads sharing memory (and our recommendation is to use the number of threads equals to twice the number of CPU cores); and the OpenCL scanner runs in multiple GPU cores and support multiple devices. The speed of a scan depends on the CPU and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark.

The OpenCL must be initialized, which just copies the address space's bitstrings to the GPU's memory. Then, many scans may be executed with no necessity to upload the bitstrings again. The OpenCL scanner supports running into multiple devices.

\subsection{OpenCL kernel}

The OpenCL kernel explores the GPU architecture to improve performance. Each work group calculates the distance of several bitstrings. During the distance calculation, each worker calculates the exclusive OR (XOR) between two 64-bit integers and use the built-in popcount function to count the number of ones. Then, they update an array of intermediate distances with their partial distances. This array is stored in the local memory and is shared with all workers of the same group. This whole step happens simultaneously in the GPU. Then, we used a reduction algorithm to sum the intermediate distances array in order to calculate the correct distance. This reduction algorithm is also distributed between the workers and runs in $O(\log_2(\text{bs\_step}))$. Finally, one of the workers checks whether the distance is less than or equal to the radius to include the bitstring index into the resulting array.

The number of workers in each group is the closest power of two above bs\_len (which is the number of 64-integers that forms a bitstring).

\section{Counters}

Each hard-location has one integer of data per bit. For instance, each hard-location of a 1,000 bits SDM has 1,000 bits. Those integers are stored in a counter.

A counter is an array of integers which stores the data of all hard-locations. So, the counter's array has $n \cdot H$ integers.

When two counters are added in a third counter, there may occur an overflow. It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with less writes than $2^{31}-1$ divided by the average number of activated hard-locations. For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.  Note also that it would be more likely to saturate the memory before any overflow.

Anyway, counters may have overflow protection depending on compiling options. By default, there is no overflow check for performance reasons (and because it does not seem necessary).

\section{Read and write operations}

The reading and writing operations are executed in two steps: first, the address space is swept looking for the activated addresses; then, the operation is performed in the counters. Reading operation assemblies the bitstring according to the counters of the activated addresses, while the writing operation changes the counters.

The iterated reading keeps reading until it gets exactly the same bitstring (or the number of maximum interations has been reached), i.e., it performs $\eta_{i+1} = \text{read}(\eta_i)$ and stops when $\eta_{k+1} = \eta_{k}$. If the initial bitstring is inside the critical distance of $\eta$, it will converge to $\eta$, but, if it is not, it will diverge and reach the maximum number of iterations.

The framework has both Kanerva's read and Murilo's generalized read. The generalization brings a parameter $z$, which is the exponent. In this case, the results are floating point instead of integer, which considerably reduces performance. When $z=1$, it is exactly as the Kanerva's read. When $z=0$, it is the Chada's read. We also explored how SDM would behave for different values of $z$.

There is another special read operation: the weighted reading. In the weighted reading, the value of the counters are multiplied by a weight which depends only on the distance between the reading address and the hard-location address. The weight is retrieved from a lookup table of integers indexed by the distance. The rest of the read operation is exactly the same.

There is also a weighted writing operation. In this case, the weight is applied when the counters are updated, i.e., if the weight is 2, the counters are increased twice when bits are 1, and decreased twice when bits are 0. Just as in the weighted reading, the weights depend only on the distance between the writing address and the hard-location address. The weights are retrieved from a lookup table of integers indexed by the distance.

### 2018-03-10 Chapters/sdm-performance.tex
SDM Splitting into multiple files.

% !TEX root = ../partial-sdm.tex

Performance matters --- and has always mattered. If an experiment takes a few seconds, there is no point arguing whether we should try it. If it takes a few hours, maybe we should think first. If it takes a few days, it is more important to make a good plan. As SDM consumes a lot of both processor and memory resources, some experiments may take a long time.

Each scan a 1,000-bits SDM with 1,000,000 hard-locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount. So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares. The heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} executed $3.05 \cdot 10^{15}$ bit compares. For comparison, the number of seconds since Jesus's birth is 63,639,648,000 = $6.36 \cdot 10^{10}$. The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$. There are approximtely $1.8 \cdot 10^{9}$ websites in the internet. A modern laptop can increment a counter $5 \cdot 10^{8}$ times per second. Hence, a naive implementation of SDM may take several hours to simply write 10,000 random bitstrings.

Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds. It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second. It is a great improvement when compared to the first versions of the code which took 16 hours to generate the same heatmap (and its memory use was already optimized and the computations were distributed in threads).

We have executed the same performance test in each device. The test has 3 parts. The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans. The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners. The code is available in the ``Performance test'' notebook \citep{sdmframework}.

Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.

Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU. The complete results are presented in Figures \ref{fig:perf-imac-read-256}, \ref{fig:perf-imac-write-256}, \ref{fig:perf-imac-read-1000}, \ref{fig:perf-imac-write-1000}, \ref{fig:perf-imac-scan-256}, \ref{fig:perf-imac-scan-1000}, and \ref{fig:perf-imac-scan-10k} and Tables \ref{tab:perf-imac-256}, \ref{tab:perf-imac-1000}, and \ref{tab:perf-imac-10k}.

Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.

It is interesting that which scanner is faster depends also on the SDM settings. In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL, but for a 256-bit SDM was the threads. What happened here is that the OpenCL kernel chosen was a generic one which performs the scan for any SDM. It is always possible to optimize the OpenCL kernel to a specific setting, and it would be faster than the threads. By default, the framework chooses a generic kernel which we believe would be reasonable for the most common setups.
% TODO Run the optimized kernel and show the results.

Most of the time, the bottleneck of both the read and the write operations is the scanner. But we have optimized the OpenCL kernel so much for the iMac 2017 that scanner not the bottleneck anymore. In the writing operation, it took the same time to scan and to update the counters. It is impressive because it update, on average, 1,000 counters of 32-bit integers each, and it was as fast as (i) sending the command to the GPU, (ii) performing 1 billion bit compares, and (iii) downloading the response from the GPU.

Our conclusion is that, if one is really concerned about performance, one should fine tune the OpenCL kernel for one's GPU. It would always be faster than running in the CPU.

Kernel comparisons

iMac 1000 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	3000	15.028241872787476	199.62414934459346	5.009413957595825
single\_scan1	3000	11.874222040176392	252.6481305343213	3.958074013392131
single\_scan2	3000	13.736932039260864	218.38937482007222	4.578977346420288
single\_scan3	3000	11.063963174819946	271.1505771121497	3.687987724939982
single\_scan4	3000	12.131728887557983	247.2854469305467	4.043909629185994
single\_scan5	3000	11.181737899780273	268.29460920014515	3.7272459665934243
single\_scan5\_unroll	3000	10.841995000839233	276.7018431356759	3.6139983336130777
single\_scan6	3000	12.084025859832764	248.26163356468672	4.028008619944255

iMac 256 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	3000	9.098679065704346	329.7181907764943	3.0328930219014483
single\_scan1	3000	8.634926080703735	347.4262514770137	2.8783086935679116
single\_scan2	3000	11.461902141571045	261.7366614149787	3.8206340471903486
single\_scan3	3000	11.187175273895264	268.1642082608964	3.729058424631755
single\_scan4	3000	13.442016124725342	223.18080652215394	4.480672041575114
single\_scan5	3000	11.295457601547241	265.59348950936374	3.76515253384908
single\_scan5\_unroll	3000	11.396361827850342	263.24190520773243	3.798787275950114
single\_scan6	3000	13.104900121688843	228.92200414675017	4.368300040562947

iMac 10k bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	500	30.534967184066772	16.374669636484867	61.069934368133545
single\_scan1	500	22.480546236038208	22.24145244293299	44.961092472076416
single\_scan2	500	22.491918802261353	22.23020651976254	44.983837604522705
single\_scan3	500	6.335015773773193	78.92640174156905	12.670031547546387
single\_scan4	500	5.727261781692505	87.30175414685539	11.45452356338501
single\_scan5	500	6.294461011886597	79.43491890025042	12.588922023773193
single\_scan5\_unroll	500	5.688975811004639	87.88928211521136	11.377951622009277
single\_scan6	500	5.48170280456543	91.21253337258189	10.96340560913086

MacBook 1000 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	1000	23.6049609184	42.363976092	23.6049609184
single\_scan1	1000	13.2222340107	75.6301846716	13.2222340107
single\_scan2	1000	47.2828419209	21.1493209667	47.2828419209
single\_scan3	1000	33.0640568733	30.2443225231	33.0640568733
single\_scan4	1000	40.3230810165	24.7996922554	40.3230810165
single\_scan5	1000	31.5100109577	31.7359458028	31.5100109577
single\_scan5\_unroll	1000	27.5569870472	36.288437422	27.5569870472
single\_scan6	1000	39.4823219776	25.3277910192	39.4823219776

MacBook 256 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	1000	8.36867690086	119.493202073	8.36867690086
single\_scan1	1000	10.4354770184	95.8269562801	10.4354770184
single\_scan2	1000	23.4823930264	42.5850976465	23.4823930264
single\_scan3	1000	25.5149869919	39.1926517665	25.5149869919
single\_scan4	1000	42.3955621719	23.5873744508	42.3955621719
single\_scan5	1000	24.4205920696	40.9490481291	24.4205920696
single\_scan5\_unroll	1000	22.7717738152	43.9140142581	22.7717738152
single\_scan6	1000	42.1847259998	23.7052624214	42.1847259998

EC2 p3.2xlarge 1000 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	3000	2.09311103821	1433.27322117	0.697703679403
single\_scan1	3000	1.64607286453	1822.51956438	0.548690954844
single\_scan2	3000	2.57495617867	1165.06837082	0.858318726222
single\_scan3	3000	3.06142997742	979.934220978	1.02047665914
single\_scan4	3000	3.30994796753	906.358658635	1.10331598918
single\_scan5	3000	2.8592748642	1049.21707163	0.953091621399
single\_scan6	3000	3.12985897064	958.509641533	1.04328632355

EC2 p3.2xlarge 256 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	3000	1.09906792641	2729.58561334	0.366355975469
single\_scan1	3000	1.10967302322	2703.49908236	0.369891007741
single\_scan2	3000	2.21544480324	1354.12987749	0.738481601079
single\_scan3	3000	1.9066619873	1573.43043496	0.635553995768
single\_scan4	3000	3.15382003784	951.227389009	1.05127334595
single\_scan5	3000	1.87652993202	1598.69552241	0.625509977341
single\_scan6	3000	3.04240322113	986.062589983	1.01413440704

EC2 p3.2xlarge 10k bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	500	10.3001351357	48.5430524372	20.6002702713
single\_scan1	500	2.47107696533	202.340925441	4.94215393066
single\_scan2	500	2.5089058876	199.290058057	5.01781177521
single\_scan3	500	3.0364408493	164.666471311	6.07288169861
single\_scan4	500	2.99925804138	166.707896787	5.99851608276
single\_scan5	500	2.68434309959	186.265310152	5.36868619919
single\_scan6	500	2.98077511787	167.741604189	5.96155023575

EC2 p2.xlarge 1000 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	3000	11.3801429272	263.617075743	3.79338097572
single\_scan1	3000	11.8365280628	253.452700325	3.94550935427
single\_scan2	3000	25.6270592213	117.063763505	8.54235307376
single\_scan3	3000	29.2102689743	102.70360751	9.73675632477
single\_scan4	3000	34.1599340439	87.8221836186	11.3866446813
single\_scan5	3000	29.8609650135	100.465607814	9.9536550045
single\_scan6	3000	34.0075969696	88.2155832028	11.3358656565

EC2 p2.xlarge 256 bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	3000	2.28510403633	1312.85051022	0.761701345444
single\_scan1	3000	2.42481589317	1237.20733126	0.808271964391
single\_scan2	3000	16.7731001377	178.857812531	5.59103337924
single\_scan3	3000	18.9349951744	158.436797705	6.31166505814
single\_scan4	3000	30.8822939396	97.1430427373	10.2940979799
single\_scan5	3000	20.0731289387	149.453531094	6.69104297956
single\_scan6	3000	30.8959820271	97.1000046988	10.2986606757

EC2 p2.xlarge 10k bits
Loops	Total time	Scans per second	Time per scan (ms)
single\_scan0	500	17.7273669243	28.204978333	35.4547338486
single\_scan1	500	28.9013559818	17.3002263394	57.8027119637
single\_scan2	500	31.8554549217	15.6958989042	63.7109098434
single\_scan3	500	19.9614479542	25.0482831279	39.9228959084
single\_scan4	500	22.749904871	21.9781138794	45.499809742
single\_scan5	500	21.7587711811	22.9792388476	43.5175423622
single\_scan6	500	20.7129361629	24.1395037414	41.4258723259

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-256.png}
\caption{Time to run a single read in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
\label{fig:perf-imac-read-256}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-256.png}
\caption{Time to run one write in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
\label{fig:perf-imac-write-256}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-1000.png}
\caption{Time to run one write in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
\label{fig:perf-imac-write-1000}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-1000.png}
\caption{Time to run a single read in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
\label{fig:perf-imac-read-1000}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
\label{fig:perf-imac-scan-256}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
\label{fig:perf-imac-scan-1000}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
\label{fig:perf-imac-scan-10k}}
\end{figure}

\begin{table}[!htb]
\centering
\begin{tabular}{| l | r | r | r |}
    \hline
    & Loops & Scans / second & Scan time (ms) \\ \hline
    Linear scan & 1000 & 81.62 & 12.25 \\
    Thread scan & 5000 & 143.68 & 6.95 \\
    OpenCL scan & 5000 & 238.00 & 4.20 \\ \hline
    \hline
    & Loops & Ops / second & Op. time (ms) \\ \hline
    Thread write & 1000 & 74.92 & 13.34 \\
    Thread single read & 1000 & 96.22 & 10.39 \\
    OpenCL write & 1000 & 126.50 & 7.90 \\
    OpenCL single read & 1000 & 190.20 & 5.25 \\
    \hline
\end{tabular}
\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
\label{tab:perf-imac-1000}}
\end{table}

\begin{table}[!htb]
\centering
\begin{tabular}{| l | r | r | r |}
    \hline
    & Loops & Scans / second & Scan time (ms) \\ \hline
    Linear scan & 1000 & 175.48 & 5.69 \\
    Thread scan & 5000 & 352.63 & 2.83 \\
    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
    \hline
    & Loops & Ops / second & Op. time (ms) \\ \hline
    Thread write & 2000 & 304.46 & 3.28 \\
    Thread single read & 2000 & 391.21 & 2.55 \\
    OpenCL write & 2000 & 378.44 & 2.64 \\
    OpenCL single read & 2000 & 466.16 & 2.14 \\
    \hline
\end{tabular}
\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
\label{tab:perf-imac-256}}
\end{table}

\begin{table}[!htb]
\centering
\begin{tabular}{| l | r | r | r |}
    \hline
    & Loops & Scans / second & Scan time (ms) \\ \hline
    Linear scan & 100 & 8.59 & 116.38 \\
    Thread scan & 500 & 18.66 & 53.56 \\
    OpenCL scan & 1000 & 77.20 & 12.95 \\
    \hline
\end{tabular}
\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.  There is no benchmark for read and write operations because RAM is not enough to allocate the counters --- it would consume 37.25 GB of RAM.
\label{tab:perf-imac-10k}}
\end{table}

### 2018-03-10 Chapters/sdm-validation.tex
SDM Splitting into multiple files.

% !TEX root = ../partial-sdm.tex

The framework has been validated comparing its results with the expected results from \citet{Kanerva1988}. Thus, we run simulations which were then compared to the theoretical analysis conducted some decades ago.

\section{Distance between random bitstrings}

As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$. For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.

In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits. In total, 20,000 random bitstrings have been generated during the simulation. The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.

In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.

This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.

\begin{figure}[!htb]
  \centering
  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}

  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
  \label{fig:validation-distance}
\end{figure}

\section{Number of activated hard-locations}

In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard-locations in a 1,000 bits SDM. He also proposed to activate only 1,000 of them, on average. He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard-locations.

We extended his results, calculating the distribution of the number of activated hard-locations. As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$. In this case, $\mu = 1071.85$ and $\sigma = 32.72$.

In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard-locations. The code is available in the ``Number of activated hard-locations'' notebook \citep{sdmframework}.

In figure \ref{fig:validation-activated-hls}, we can notice that the theoretical model and the simulation matches. Hence, it seems that both the address space generation algotihm and the scan algorithm work properly. Notice that the curve is almost the same for $n=1,000$ and $n=256$. It happens because the access radius is adjust to have $p$ as close as possible to $0.001$. They are not exactly the same because their $p$ differs a little.

\begin{figure}[!htb]
  \centering
  \subfloat[$n=1,000$, $H=1,000,000$,\protect\\ $r=451$, and $p=0.00107185$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-1000.png}}
  \subfloat[$n=256$, $H=1,000,000$,\protect\\ $r=103$, and $p=0.00106684$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-256.png}}

  \caption{Histogram of the number of activated hard-locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
  \label{fig:validation-activated-hls}
\end{figure}

Besides the number of activated hard-locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard-locations. Let $A$ be the set of activated hard-locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:

\begin{align}
P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(a \in A)} \\
    &= \frac{\binom{n}{x}}{\sum_{k=0}^r \binom{n}{k}} \label{eq:prob-d-inside-circle}
\end{align}

In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard-locations to the center of 1,000 random circles. The code is available in the ``Distances of activated hard-locations'' notebook \citep{sdmframework}.

In figure \ref{fig:validation-distance-activated-hls}, we can notice that the theoretical model and the simulation matches.

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{./images02/new-images/distance-activated-hls.png}

\caption{Histogram of the distances of activated hard-locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
\label{fig:validation-distance-activated-hls}}
\end{figure}

\section{Intersection of two circles}

Kanerva has calculated the intersection of two circles according to the distance between their centers. The intersection is important to understand how SDM works, because it affects directly the critical distance. When $\eta_d$ is inside the critical distance, then it will converge to $\eta$. In fact, it converges because they share a sufficient amount of hard-locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge. For further information about the relation between the critical distance and the intersection, see \citet{brogliato2014sparse}.

We have calculated the intersection between a random bitstring (bs1) and another bitstrings (bs2) exactly $d$ bits away. The former (bs1) is just a random bitstring. The latter (bs2) was generated randomly flipping $d$ bits of bs1. The code is available in the ``Kanerva's Figure 1.2'' notebook \citep{sdmframework}.

In Figure \ref{fig:validation-intersection}, we can notice that we have obtained the same results as Kanerva. It seems that the random flipping bits algorithm and the scan algorithm work properly.

\begin{figure}[!htb]
  \centering
  \subfloat[{\citet[Figure 1.2, p.25]{Kanerva1988}} ]{\includegraphics[width=0.45\textwidth]{./images02/new-images/kanerva-table-12.jpg}}
  \subfloat[Generated by SDM-Frameworkm with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}

  \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
  \label{fig:validation-intersection}
\end{figure}

\section{Storage and retrieval of sequences}

\citet[Ch.8]{Kanerva1988} presented an approach to store and retrieve sequences using $k$ different SDMs, namely $\text{sdm}_1$, $\text{sdm}_2$, $dots$, $\text{sdm}_k$.

Let $a_0, a_1, a_2, \dots, a_n$ be a sequence to be stored in a $k$-fold memory. So, all pointers of the form $a_i \rightarrow a_{i+k}$ will be written to $\text{sdm}_k$ memory, i.e., in $\text{sdm}_1$, the following pointers will be written: $a_0 \rightarrow a_1$, $a_1 \rightarrow a_2$, $\dots$, $a_{n-1} \rightarrow a_n$; while in $\text{sdm}_2$, the following pointers will be written: $a_0 \rightarrow a_2$, $a_2 \rightarrow a_3$, $\dots$, $a_{n-2} \rightarrow a_n$; and so forth.

We have tested exactly the same example presented in \citet{Kanerva1988}, p.85. We wrote two sequences to a $3$-fold memory: $<A, B, C, D>$ and $<E, B, C, F>$. Then, after reading the sequences $<A, B, C>$ and $<E, B, C>$, we have obtained $D$ and $F$, respectively.

Each reading operation was performed summing the counters of all activated hard-locations from all three memories. For instance, to read the sequence $<A, B, C>$, we have activated the hard-locations around $C$ in $\text{sdm}_1$, we have also activated the hard-locations around $D$ in $\text{sdm}_2$, and, finally, we have also activated the hard-locations around $A$ in $\text{sdm}_3$. After summing the counters of all those hard-locations, we evaluate the resulting bitstring just as in the original read operation.

The code is available in the ``Sequences (Kanerva Ch 8)'' notebook \citep{sdmframework}.

The logic behind how it works is that, when reading the sequence $<A, B, C>$, we have $A$ pointing to $D$, while both $B$ and $C$ point to $D$ and $F$. Thus, $D$ appears more often than $F$ and ended up being the result.

Hence, as we have replicated the theoretical results from Kanerva, we have one more evidence that our framework works properly.

\subsection{$k$-fold memory using only one SDM}

We have extended Kanerva's ideas to be able to store and retrieve sequences in $k$-fold memories using only one SDM (instead of $k$ SDMs).

Our idea was to create $k$ random bitstrings, one for each fold. We have performed writing and reading exactly as Kanerva's original idea, but, instead of writing to $\text{sdm}_k$, we have written $a_{i+k}$ into the address $a_i \oplus \text{tag}_k$, and, instead of reading from $\text{sdm}_k$, we have read from address $a_i \oplus \text{tag}_k$, where $\oplus$ is the exclusive or (XOR) operator.

It worked as if we had splitted SDM into $k$ regions with low intersection between two of them. So, as the interference is minimal, they work like independent SDMs. The major disadvantage of this approach is that memory capacity may be reached faster.

Splitting the memory into regions may be an interesting strategy to other sorts of problems, mostly the ones which would need many SDMs and, consequently, would use a lot of RAM.

\section{Convergence of $\eta_x$ to $\eta$}

One particular analysis of Kanerva's interest is that of the distance read at a point $\eta_x$. Suppose an SDM is trying to read an item written at $\eta$, but the cues received so far lead to a point of distance $x$ from $\eta$.  As one reads at $\eta_x$, a new bitstring $\beta$ is obtained, leading to Kanerva's question: what is the new distance from $\eta$ to $\beta$? Is it smaller or larger than $x$? That, of course, depends on the ratio between $x$ and the number of dimensions of the memory.

\citet[p.70]{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Figure \ref{fig:kanerva-figure-7.3}). The original prediction considered that the read distance would decline when inside the critical distance and increase afterwards, converging to a \textasciitilde 500-bit distance.  At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance bitstring. He analyzed specifically an SDM with 1,000 bits and 10,000 random bitstrings written into it.

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/kanerva-table-7-2-original.png}
\caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
\label{fig:kanerva-figure-7.3}}
\end{figure}

As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
%and we propose that this is due to interaction effects between different attractors.

We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.

In details, we have created a SDM with $n=1,000$, $H=1,000,000$, and $r=451$. Then, we have generated 10,000 random bitstrings and written them into the memory. Then, we have generated a reference bitstring (bs\_ref) and written it into the memory. Then, we have executed the following steps with $x$ from 0 to 1,000: (i) copy bs\_ref into a new bitstring; (ii) randomly flipped $x$ bits of the copy; (iii) read from the memory in the copy address; and (iv) stored the distance between the returned bitstring and bs\_ref. Finally, we have plotted Figure \ref{fig:sdm-10000w-table-7-2}.

\begin{figure}[h]
\centering
\subfloat[1 sample for each distance $x$ \label{fig:sdm-10000w-table-7-2-1sample} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2.png}}
\subfloat[6 samples for each distance $x$ \label{fig:sdm-10000w-table-7-2-6samples} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2-6-samples.png}}

\caption{Results generated by the framework diverging from Kanerva's original Table 7.2. Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
\label{fig:sdm-10000w-table-7-2}}
\end{figure}

Figure \ref{fig:sdm-10000w-table-7-2-1sample} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance. So, we have changed the steps to run $k$ reads and store the average new distance. We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6samples}, which has a way lower noise and still holds the divergence.

Our results show that the theoretical prediction is not accurate.  There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).

Obviously, these small deviations from Kanerva's original theoretical predictions deserve a qualification.  Kanerva was working in the 1980s and the 1990s, and had no access to the immense computational power that we do today. It is no surprise that some small interaction effects should exist as machines allow us to explore the ideas of his monumental work.

But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Figure \ref{fig:sdm-100w-table-7-2-10samples}). This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors. In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractors.

\begin{figure}[h]
\centering
\subfloat[100 writes \label{fig:sdm-100w-table-7-2-10samples} ]{\includegraphics[width=0.47\textwidth]{./images02/sdm-100w-table-7-2.png}}
\subfloat[Steps of 1,000 writes \label{fig:sdm-table-7-2-steps} ]{\includegraphics[width=0.53\textwidth]{./images02/sdm-table-7-2-steps.png}}
%\centering\includegraphics[width=\textwidth]{images02/sdm-100w-table-7-2.png}

\caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
\label{fig:sdm-100w-table-7-2}}
\end{figure}

To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin. Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM. Thereby, in order to show the interaction effects more clearly, we changed the single read for an 15-iterative read. As we can see in Figure \ref{sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.

Hence, our understanding is that the attractors are just preventing the bitstrings to converge directly to 500-bit distance bitstrings, requiring more reading steps to do so. They are in other orthogonal bitstrings' critical distance, but sufficiently far not to converge in a single read.

\begin{figure}[h]
\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-15iter.png}
\caption{This graph shows the interaction effects more clearly.  As we change the single read to a 6-iterative read, the effect has vanished and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings. Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, except for the iterative read.
\label{sdm-10000w-table-7-2-15iter}}
\end{figure}

%To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin. Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM. Thereby, in order to show the interaction effects more clearly, we wrote a handmade bitstring to the SDM which had all bits inverted in relation to the origin bitstring --- their hamming distance was equal to 1,000. Our handmade bitstring was acting as an opposite attractor, and one can see the accelerating effects towards convergence to both attractors: the origin and the handmade bitstrings (Fig. \ref{sdm-10000w-notX-table-7-2}). Here we had the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.

%\begin{figure}[h]
%\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
%\caption{This graph shows the interaction effects more clearly.  As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and the opposite. Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
%\label{sdm-10000w-notX-table-7-2}}
%\end{figure}

### 2018-03-10 Chapters/sdm-performance.tex
SDM Add kernel comparison to performance chapter

@@
 
-iMac 1000 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	3000	15.028241872787476	199.62414934459346	5.009413957595825
-single\_scan1	3000	11.874222040176392	252.6481305343213	3.958074013392131
-single\_scan2	3000	13.736932039260864	218.38937482007222	4.578977346420288
-single\_scan3	3000	11.063963174819946	271.1505771121497	3.687987724939982
-single\_scan4	3000	12.131728887557983	247.2854469305467	4.043909629185994
-single\_scan5	3000	11.181737899780273	268.29460920014515	3.7272459665934243
-single\_scan5\_unroll	3000	10.841995000839233	276.7018431356759	3.6139983336130777
-single\_scan6	3000	12.084025859832764	248.26163356468672	4.028008619944255
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 3000 & 15.02 & 199.62 & 5.00 \\
+    single\_scan1 & 3000 & 11.87 & 252.64 & 3.95 \\
+    single\_scan2 & 3000 & 13.73 & 218.38 & 4.57 \\
+    single\_scan3 & 3000 & 11.06 & 271.15 & 3.68 \\
+    single\_scan4 & 3000 & 12.13 & 247.28 & 4.04 \\
+    single\_scan5 & 3000 & 11.18 & 268.29 & 3.72 \\
+    single\_scan5\_unroll & 3000 & 10.84 & 276.70 & 3.6139983336130777 \\
+    single\_scan6 & 3000 & 12.08 & 248.26 & 4.02 \\
+    \hline
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
+\label{tab:perf-imac-256}}
+\end{table}
 
-iMac 256 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	3000	9.098679065704346	329.7181907764943	3.0328930219014483
-single\_scan1	3000	8.634926080703735	347.4262514770137	2.8783086935679116
-single\_scan2	3000	11.461902141571045	261.7366614149787	3.8206340471903486
-single\_scan3	3000	11.187175273895264	268.1642082608964	3.729058424631755
-single\_scan4	3000	13.442016124725342	223.18080652215394	4.480672041575114
-single\_scan5	3000	11.295457601547241	265.59348950936374	3.76515253384908
-single\_scan5\_unroll	3000	11.396361827850342	263.24190520773243	3.798787275950114
-single\_scan6	3000	13.104900121688843	228.92200414675017	4.368300040562947
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 3000 & 9.09 & 329.71 & 3.03 \\
+    single\_scan1 & 3000 & 8.63 & 347.42 & 2.87 \\
+    single\_scan2 & 3000 & 11.46 & 261.73 & 3.82 \\
+    single\_scan3 & 3000 & 11.18 & 268.16 & 3.72 \\
+    single\_scan4 & 3000 & 13.44 & 223.18 & 4.48 \\
+    single\_scan5 & 3000 & 11.29 & 265.59 & 3.76 \\
+    single\_scan5\_unroll & 3000 & 11.39 & 263.24 & 3.79 \\
+    single\_scan6 & 3000 & 13.10 & 228.92 & 4.36 \\
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+\label{tab:perf-imac-256}}
+\end{table}
 
-iMac 10k bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	500	30.534967184066772	16.374669636484867	61.069934368133545
-single\_scan1	500	22.480546236038208	22.24145244293299	44.961092472076416
-single\_scan2	500	22.491918802261353	22.23020651976254	44.983837604522705
-single\_scan3	500	6.335015773773193	78.92640174156905	12.670031547546387
-single\_scan4	500	5.727261781692505	87.30175414685539	11.45452356338501
-single\_scan5	500	6.294461011886597	79.43491890025042	12.588922023773193
-single\_scan5\_unroll	500	5.688975811004639	87.88928211521136	11.377951622009277
-single\_scan6	500	5.48170280456543	91.21253337258189	10.96340560913086
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 500 & 30.53 & 16.37 & 61.06 \\
+    single\_scan1 & 500 & 22.48 & 22.24 & 44.96 \\
+    single\_scan2 & 500 & 22.49 & 22.23 & 44.98 \\
+    single\_scan3 & 500 & 6.33 & 78.92 & 12.67 \\
+    single\_scan4 & 500 & 5.72 & 87.30 & 11.45 \\
+    single\_scan5 & 500 & 6.29 & 79.43 & 12.58 \\
+    single\_scan5\_unroll & 500 & 5.68 & 87.88 & 11.37 \\
+    single\_scan6 & 500 & 5.48 & 91.21 & 10.96 \\
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4850$.
+\label{tab:perf-imac-256}}
+\end{table}
 
-MacBook 1000 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	1000	23.6049609184	42.363976092	23.6049609184
-single\_scan1	1000	13.2222340107	75.6301846716	13.2222340107
-single\_scan2	1000	47.2828419209	21.1493209667	47.2828419209
-single\_scan3	1000	33.0640568733	30.2443225231	33.0640568733
-single\_scan4	1000	40.3230810165	24.7996922554	40.3230810165
-single\_scan5	1000	31.5100109577	31.7359458028	31.5100109577
-single\_scan5\_unroll	1000	27.5569870472	36.288437422	27.5569870472
-single\_scan6	1000	39.4823219776	25.3277910192	39.4823219776
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 1000 & 23.60 & 42.36 & 23.60 \\
+    single\_scan1 & 1000 & 13.22 & 75.63 & 13.22 \\
+    single\_scan2 & 1000 & 47.28 & 21.14 & 47.28 \\
+    single\_scan3 & 1000 & 33.06 & 30.24 & 33.06 \\
+    single\_scan4 & 1000 & 40.32 & 24.79 & 40.32 \\
+    single\_scan5 & 1000 & 31.51 & 31.73 & 31.51 \\
+    single\_scan5\_unroll & 1000 & 27.55 & 36.28 & 27.55 \\
+    single\_scan6 & 1000 & 39.48 & 25.32 & 39.48 \\
+\end{tabular}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
+\label{tab:perf-imac-256}}
+\end{table}
 
-MacBook 256 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	1000	8.36867690086	119.493202073	8.36867690086
-single\_scan1	1000	10.4354770184	95.8269562801	10.4354770184
-single\_scan2	1000	23.4823930264	42.5850976465	23.4823930264
-single\_scan3	1000	25.5149869919	39.1926517665	25.5149869919
-single\_scan4	1000	42.3955621719	23.5873744508	42.3955621719
-single\_scan5	1000	24.4205920696	40.9490481291	24.4205920696
-single\_scan5\_unroll	1000	22.7717738152	43.9140142581	22.7717738152
-single\_scan6	1000	42.1847259998	23.7052624214	42.1847259998
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 1000 & 8.36 & 119.49 & 8.36 \\
+    single\_scan1 & 1000 & 10.43 & 95.82 & 10.43 \\
+    single\_scan2 & 1000 & 23.48 & 42.58 & 23.48 \\
+    single\_scan3 & 1000 & 25.51 & 39.19 & 25.51 \\
+    single\_scan4 & 1000 & 42.39 & 23.58 & 42.39 \\
+    single\_scan5 & 1000 & 24.42 & 40.94 & 24.42 \\
+    single\_scan5\_unroll & 1000 & 22.77 & 43.91 & 22.77 \\
+    single\_scan6 & 1000 & 42.18 & 23.70 & 42.18 \\
+\end{tabular}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+\label{tab:perf-imac-256}}
+\end{table}
 
-EC2 p3.2xlarge 1000 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	3000	2.09311103821	1433.27322117	0.697703679403
-single\_scan1	3000	1.64607286453	1822.51956438	0.548690954844
-single\_scan2	3000	2.57495617867	1165.06837082	0.858318726222
-single\_scan3	3000	3.06142997742	979.934220978	1.02047665914
-single\_scan4	3000	3.30994796753	906.358658635	1.10331598918
-single\_scan5	3000	2.8592748642	1049.21707163	0.953091621399
-single\_scan6	3000	3.12985897064	958.509641533	1.04328632355
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Loops & Total time & Scans per second & Time per scan (ms) \\
+    single\_scan0 & 3000 & 2.09 & 1433.27 & 0.69 \\
+    single\_scan1 & 3000 & 1.64 & 1822.51 & 0.54 \\
+    single\_scan2 & 3000 & 2.57 & 1165.06 & 0.85 \\
+    single\_scan3 & 3000 & 3.06 & 979.93 & 1.02 \\
+    single\_scan4 & 3000 & 3.30 & 906.35 & 1.10 \\
+    single\_scan5 & 3000 & 2.85 & 1049.21 & 0.95 \\
+    single\_scan6 & 3000 & 3.12 & 958.50 & 1.04 \\
+\end{tabular}
+\caption{EC2 p3.2xlarge 1000 bits
+\label{tab:perf-imac-256}}
+\end{table}
 
-EC2 p3.2xlarge 256 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	3000	1.09906792641	2729.58561334	0.366355975469
-single\_scan1	3000	1.10967302322	2703.49908236	0.369891007741
-single\_scan2	3000	2.21544480324	1354.12987749	0.738481601079
-single\_scan3	3000	1.9066619873	1573.43043496	0.635553995768
-single\_scan4	3000	3.15382003784	951.227389009	1.05127334595
-single\_scan5	3000	1.87652993202	1598.69552241	0.625509977341
-single\_scan6	3000	3.04240322113	986.062589983	1.01413440704
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Loops & Total time & Scans per second & Time per scan (ms) \\
+    single\_scan0 & 3000 & 1.09 & 2729.58 & 0.36 \\
+    single\_scan1 & 3000 & 1.10 & 2703.49 & 0.36 \\
+    single\_scan2 & 3000 & 2.21 & 1354.12 & 0.73 \\
+    single\_scan3 & 3000 & 1.90 & 1573.43 & 0.63 \\
+    single\_scan4 & 3000 & 3.15 & 951.22 & 1.05 \\
+    single\_scan5 & 3000 & 1.87 & 1598.69 & 0.62 \\
+    single\_scan6 & 3000 & 3.04 & 986.06 & 1.01 \\
+\end{tabular}
+\caption{EC2 p3.2xlarge 256 bits
+\label{tab:perf-imac-256}}
+\end{table}
 
-EC2 p3.2xlarge 10k bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	500	10.3001351357	48.5430524372	20.6002702713
-single\_scan1	500	2.47107696533	202.340925441	4.94215393066
-single\_scan2	500	2.5089058876	199.290058057	5.01781177521
-single\_scan3	500	3.0364408493	164.666471311	6.07288169861
-single\_scan4	500	2.99925804138	166.707896787	5.99851608276
-single\_scan5	500	2.68434309959	186.265310152	5.36868619919
-single\_scan6	500	2.98077511787	167.741604189	5.96155023575
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Loops & Total time & Scans per second & Time per scan (ms) \\
+    single\_scan0 & 500 & 10.30 & 48.54 & 20.60 \\
+    single\_scan1 & 500 & 2.47 & 202.34 & 4.94 \\
+    single\_scan2 & 500 & 2.50 & 199.29 & 5.01 \\
+    single\_scan3 & 500 & 3.03 & 164.66 & 6.07 \\
+    single\_scan4 & 500 & 2.99 & 166.70 & 5.99 \\
+    single\_scan5 & 500 & 2.68 & 186.26 & 5.36 \\
+    single\_scan6 & 500 & 2.98 & 167.74 & 5.96 \\
+\end{tabular}
+\caption{EC2 p3.2xlarge 10k bits
+\label{tab:perf-imac-256}}
+\end{table}
 
-EC2 p2.xlarge 1000 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	3000	11.3801429272	263.617075743	3.79338097572
-single\_scan1	3000	11.8365280628	253.452700325	3.94550935427
-single\_scan2	3000	25.6270592213	117.063763505	8.54235307376
-single\_scan3	3000	29.2102689743	102.70360751	9.73675632477
-single\_scan4	3000	34.1599340439	87.8221836186	11.3866446813
-single\_scan5	3000	29.8609650135	100.465607814	9.9536550045
-single\_scan6	3000	34.0075969696	88.2155832028	11.3358656565
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Loops & Total time & Scans per second & Time per scan (ms) \\
+    single\_scan0 & 3000 & 11.38 & 263.61 & 3.79 \\
+    single\_scan1 & 3000 & 11.83 & 253.45 & 3.94 \\
+    single\_scan2 & 3000 & 25.62 & 117.06 & 8.54 \\
+    single\_scan3 & 3000 & 29.21 & 102.70 & 9.73 \\
+    single\_scan4 & 3000 & 34.15 & 87.82 & 11.38 \\
+    single\_scan5 & 3000 & 29.86 & 100.46 & 9.95 \\
+    single\_scan6 & 3000 & 34.00 & 88.21 & 11.33 \\
+\end{tabular}
+\caption{EC2 p2.xlarge 1000 bits
+\label{tab:perf-imac-256}}
+\end{table}
 
-EC2 p2.xlarge 256 bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	3000	2.28510403633	1312.85051022	0.761701345444
-single\_scan1	3000	2.42481589317	1237.20733126	0.808271964391
-single\_scan2	3000	16.7731001377	178.857812531	5.59103337924
-single\_scan3	3000	18.9349951744	158.436797705	6.31166505814
-single\_scan4	3000	30.8822939396	97.1430427373	10.2940979799
-single\_scan5	3000	20.0731289387	149.453531094	6.69104297956
-single\_scan6	3000	30.8959820271	97.1000046988	10.2986606757
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Loops & Total time & Scans per second & Time per scan (ms) \\
+    single\_scan0 & 3000 & 2.28 & 1312.85 & 0.76 \\
+    single\_scan1 & 3000 & 2.42 & 1237.20 & 0.80 \\
+    single\_scan2 & 3000 & 16.77 & 178.85 & 5.59 \\
+    single\_scan3 & 3000 & 18.93 & 158.43 & 6.31 \\
+    single\_scan4 & 3000 & 30.88 & 97.14 & 10.29 \\
+    single\_scan5 & 3000 & 20.07 & 149.45 & 6.69 \\
+    single\_scan6 & 3000 & 30.89 & 97.10 & 10.29 \\
+\end{tabular}
+\caption{EC2 p2.xlarge 256 bits
+\label{tab:perf-imac-256}}
+\end{table}
 
-EC2 p2.xlarge 10k bits
-Loops	Total time	Scans per second	Time per scan (ms)
-single\_scan0	500	17.7273669243	28.204978333	35.4547338486
-single\_scan1	500	28.9013559818	17.3002263394	57.8027119637
-single\_scan2	500	31.8554549217	15.6958989042	63.7109098434
-single\_scan3	500	19.9614479542	25.0482831279	39.9228959084
-single\_scan4	500	22.749904871	21.9781138794	45.499809742
-single\_scan5	500	21.7587711811	22.9792388476	43.5175423622
-single\_scan6	500	20.7129361629	24.1395037414	41.4258723259
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Loops & Total time & Scans per second & Time per scan (ms) \\
+    single\_scan0 & 500 & 17.72 & 28.20 & 35.45 \\
+    single\_scan1 & 500 & 28.90 & 17.30 & 57.80 \\
+    single\_scan2 & 500 & 31.85 & 15.69 & 63.71 \\
+    single\_scan3 & 500 & 19.96 & 25.04 & 39.92 \\
+    single\_scan4 & 500 & 22.74 & 21.97 & 45.49 \\
+    single\_scan5 & 500 & 21.75 & 22.97 & 43.51 \\
+    single\_scan6 & 500 & 20.71 & 24.13 & 41.42 \\
+\end{tabular}
+\caption{EC2 p2.xlarge 10k bits
+\label{tab:perf-imac-256}}
+\end{table}
 

### 2018-03-10 Chapters/sdm-performance.tex
SDM performance tables fix

@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 3000 & 15.02 & 199.62 & 5.00 \\
-    single\_scan1 & 3000 & 11.87 & 252.64 & 3.95 \\
-    single\_scan2 & 3000 & 13.73 & 218.38 & 4.57 \\
-    single\_scan3 & 3000 & 11.06 & 271.15 & 3.68 \\
-    single\_scan4 & 3000 & 12.13 & 247.28 & 4.04 \\
-    single\_scan5 & 3000 & 11.18 & 268.29 & 3.72 \\
-    single\_scan5\_unroll & 3000 & 10.84 & 276.70 & 3.6139983336130777 \\
-    single\_scan6 & 3000 & 12.08 & 248.26 & 4.02 \\
+    single\_scan0 & 3000 & 199.62 & 5.00 \\
+    single\_scan1 & 3000 & 252.64 & 3.95 \\
+    single\_scan2 & 3000 & 218.38 & 4.57 \\
+    single\_scan3 & 3000 & 271.15 & 3.68 \\
+    single\_scan4 & 3000 & 247.28 & 4.04 \\
+    single\_scan5 & 3000 & 268.29 & 3.72 \\
+    single\_scan5\_unroll & 3000 & 276.70 & 3.6139983336130777 \\
+    single\_scan6 & 3000 & 248.26 & 4.02 \\
     \hline
@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 3000 & 9.09 & 329.71 & 3.03 \\
-    single\_scan1 & 3000 & 8.63 & 347.42 & 2.87 \\
-    single\_scan2 & 3000 & 11.46 & 261.73 & 3.82 \\
-    single\_scan3 & 3000 & 11.18 & 268.16 & 3.72 \\
-    single\_scan4 & 3000 & 13.44 & 223.18 & 4.48 \\
-    single\_scan5 & 3000 & 11.29 & 265.59 & 3.76 \\
-    single\_scan5\_unroll & 3000 & 11.39 & 263.24 & 3.79 \\
-    single\_scan6 & 3000 & 13.10 & 228.92 & 4.36 \\
+    single\_scan0 & 3000 & 329.71 & 3.03 \\
+    single\_scan1 & 3000 & 347.42 & 2.87 \\
+    single\_scan2 & 3000 & 261.73 & 3.82 \\
+    single\_scan3 & 3000 & 268.16 & 3.72 \\
+    single\_scan4 & 3000 & 223.18 & 4.48 \\
+    single\_scan5 & 3000 & 265.59 & 3.76 \\
+    single\_scan5\_unroll & 3000 & 263.24 & 3.79 \\
+    single\_scan6 & 3000 & 228.92 & 4.36 \\
 \end{tabular}
@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 500 & 30.53 & 16.37 & 61.06 \\
-    single\_scan1 & 500 & 22.48 & 22.24 & 44.96 \\
-    single\_scan2 & 500 & 22.49 & 22.23 & 44.98 \\
-    single\_scan3 & 500 & 6.33 & 78.92 & 12.67 \\
-    single\_scan4 & 500 & 5.72 & 87.30 & 11.45 \\
-    single\_scan5 & 500 & 6.29 & 79.43 & 12.58 \\
-    single\_scan5\_unroll & 500 & 5.68 & 87.88 & 11.37 \\
-    single\_scan6 & 500 & 5.48 & 91.21 & 10.96 \\
+    single\_scan0 & 500 & 16.37 & 61.06 \\
+    single\_scan1 & 500 & 22.24 & 44.96 \\
+    single\_scan2 & 500 & 22.23 & 44.98 \\
+    single\_scan3 & 500 & 78.92 & 12.67 \\
+    single\_scan4 & 500 & 87.30 & 11.45 \\
+    single\_scan5 & 500 & 79.43 & 12.58 \\
+    single\_scan5\_unroll & 500 & 87.88 & 11.37 \\
+    single\_scan6 & 500 & 91.21 & 10.96 \\
 \end{tabular}
@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 23.60 & 42.36 & 23.60 \\
-    single\_scan1 & 1000 & 13.22 & 75.63 & 13.22 \\
-    single\_scan2 & 1000 & 47.28 & 21.14 & 47.28 \\
-    single\_scan3 & 1000 & 33.06 & 30.24 & 33.06 \\
-    single\_scan4 & 1000 & 40.32 & 24.79 & 40.32 \\
-    single\_scan5 & 1000 & 31.51 & 31.73 & 31.51 \\
-    single\_scan5\_unroll & 1000 & 27.55 & 36.28 & 27.55 \\
-    single\_scan6 & 1000 & 39.48 & 25.32 & 39.48 \\
+    single\_scan0 & 1000 & 42.36 & 23.60 \\
+    single\_scan1 & 1000 & 75.63 & 13.22 \\
+    single\_scan2 & 1000 & 21.14 & 47.28 \\
+    single\_scan3 & 1000 & 30.24 & 33.06 \\
+    single\_scan4 & 1000 & 24.79 & 40.32 \\
+    single\_scan5 & 1000 & 31.73 & 31.51 \\
+    single\_scan5\_unroll & 1000 & 36.28 & 27.55 \\
+    single\_scan6 & 1000 & 25.32 & 39.48 \\
 \end{tabular}
@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 8.36 & 119.49 & 8.36 \\
-    single\_scan1 & 1000 & 10.43 & 95.82 & 10.43 \\
-    single\_scan2 & 1000 & 23.48 & 42.58 & 23.48 \\
-    single\_scan3 & 1000 & 25.51 & 39.19 & 25.51 \\
-    single\_scan4 & 1000 & 42.39 & 23.58 & 42.39 \\
-    single\_scan5 & 1000 & 24.42 & 40.94 & 24.42 \\
-    single\_scan5\_unroll & 1000 & 22.77 & 43.91 & 22.77 \\
-    single\_scan6 & 1000 & 42.18 & 23.70 & 42.18 \\
+    single\_scan0 & 1000 & 119.49 & 8.36 \\
+    single\_scan1 & 1000 & 95.82 & 10.43 \\
+    single\_scan2 & 1000 & 42.58 & 23.48 \\
+    single\_scan3 & 1000 & 39.19 & 25.51 \\
+    single\_scan4 & 1000 & 23.58 & 42.39 \\
+    single\_scan5 & 1000 & 40.94 & 24.42 \\
+    single\_scan5\_unroll & 1000 & 43.91 & 22.77 \\
+    single\_scan6 & 1000 & 23.70 & 42.18 \\
 \end{tabular}
@@
     Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 2.09 & 1433.27 & 0.69 \\
-    single\_scan1 & 3000 & 1.64 & 1822.51 & 0.54 \\
-    single\_scan2 & 3000 & 2.57 & 1165.06 & 0.85 \\
-    single\_scan3 & 3000 & 3.06 & 979.93 & 1.02 \\
-    single\_scan4 & 3000 & 3.30 & 906.35 & 1.10 \\
-    single\_scan5 & 3000 & 2.85 & 1049.21 & 0.95 \\
-    single\_scan6 & 3000 & 3.12 & 958.50 & 1.04 \\
+    single\_scan0 & 3000 & 1433.27 & 0.69 \\
+    single\_scan1 & 3000 & 1822.51 & 0.54 \\
+    single\_scan2 & 3000 & 1165.06 & 0.85 \\
+    single\_scan3 & 3000 & 979.93 & 1.02 \\
+    single\_scan4 & 3000 & 906.35 & 1.10 \\
+    single\_scan5 & 3000 & 1049.21 & 0.95 \\
+    single\_scan6 & 3000 & 958.50 & 1.04 \\
 \end{tabular}
@@
     Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 1.09 & 2729.58 & 0.36 \\
-    single\_scan1 & 3000 & 1.10 & 2703.49 & 0.36 \\
-    single\_scan2 & 3000 & 2.21 & 1354.12 & 0.73 \\
-    single\_scan3 & 3000 & 1.90 & 1573.43 & 0.63 \\
-    single\_scan4 & 3000 & 3.15 & 951.22 & 1.05 \\
-    single\_scan5 & 3000 & 1.87 & 1598.69 & 0.62 \\
-    single\_scan6 & 3000 & 3.04 & 986.06 & 1.01 \\
+    single\_scan0 & 3000 & 2729.58 & 0.36 \\
+    single\_scan1 & 3000 & 2703.49 & 0.36 \\
+    single\_scan2 & 3000 & 1354.12 & 0.73 \\
+    single\_scan3 & 3000 & 1573.43 & 0.63 \\
+    single\_scan4 & 3000 & 951.22 & 1.05 \\
+    single\_scan5 & 3000 & 1598.69 & 0.62 \\
+    single\_scan6 & 3000 & 986.06 & 1.01 \\
 \end{tabular}
@@
     Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 500 & 10.30 & 48.54 & 20.60 \\
-    single\_scan1 & 500 & 2.47 & 202.34 & 4.94 \\
-    single\_scan2 & 500 & 2.50 & 199.29 & 5.01 \\
-    single\_scan3 & 500 & 3.03 & 164.66 & 6.07 \\
-    single\_scan4 & 500 & 2.99 & 166.70 & 5.99 \\
-    single\_scan5 & 500 & 2.68 & 186.26 & 5.36 \\
-    single\_scan6 & 500 & 2.98 & 167.74 & 5.96 \\
+    single\_scan0 & 500 & 48.54 & 20.60 \\
+    single\_scan1 & 500 & 202.34 & 4.94 \\
+    single\_scan2 & 500 & 199.29 & 5.01 \\
+    single\_scan3 & 500 & 164.66 & 6.07 \\
+    single\_scan4 & 500 & 166.70 & 5.99 \\
+    single\_scan5 & 500 & 186.26 & 5.36 \\
+    single\_scan6 & 500 & 167.74 & 5.96 \\
 \end{tabular}
@@
     Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 11.38 & 263.61 & 3.79 \\
-    single\_scan1 & 3000 & 11.83 & 253.45 & 3.94 \\
-    single\_scan2 & 3000 & 25.62 & 117.06 & 8.54 \\
-    single\_scan3 & 3000 & 29.21 & 102.70 & 9.73 \\
-    single\_scan4 & 3000 & 34.15 & 87.82 & 11.38 \\
-    single\_scan5 & 3000 & 29.86 & 100.46 & 9.95 \\
-    single\_scan6 & 3000 & 34.00 & 88.21 & 11.33 \\
+    single\_scan0 & 3000 & 263.61 & 3.79 \\
+    single\_scan1 & 3000 & 253.45 & 3.94 \\
+    single\_scan2 & 3000 & 117.06 & 8.54 \\
+    single\_scan3 & 3000 & 102.70 & 9.73 \\
+    single\_scan4 & 3000 & 87.82 & 11.38 \\
+    single\_scan5 & 3000 & 100.46 & 9.95 \\
+    single\_scan6 & 3000 & 88.21 & 11.33 \\
 \end{tabular}
@@
     Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 2.28 & 1312.85 & 0.76 \\
-    single\_scan1 & 3000 & 2.42 & 1237.20 & 0.80 \\
-    single\_scan2 & 3000 & 16.77 & 178.85 & 5.59 \\
-    single\_scan3 & 3000 & 18.93 & 158.43 & 6.31 \\
-    single\_scan4 & 3000 & 30.88 & 97.14 & 10.29 \\
-    single\_scan5 & 3000 & 20.07 & 149.45 & 6.69 \\
-    single\_scan6 & 3000 & 30.89 & 97.10 & 10.29 \\
+    single\_scan0 & 3000 & 1312.85 & 0.76 \\
+    single\_scan1 & 3000 & 1237.20 & 0.80 \\
+    single\_scan2 & 3000 & 178.85 & 5.59 \\
+    single\_scan3 & 3000 & 158.43 & 6.31 \\
+    single\_scan4 & 3000 & 97.14 & 10.29 \\
+    single\_scan5 & 3000 & 149.45 & 6.69 \\
+    single\_scan6 & 3000 & 97.10 & 10.29 \\
 \end{tabular}
@@
     Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 500 & 17.72 & 28.20 & 35.45 \\
-    single\_scan1 & 500 & 28.90 & 17.30 & 57.80 \\
-    single\_scan2 & 500 & 31.85 & 15.69 & 63.71 \\
-    single\_scan3 & 500 & 19.96 & 25.04 & 39.92 \\
-    single\_scan4 & 500 & 22.74 & 21.97 & 45.49 \\
-    single\_scan5 & 500 & 21.75 & 22.97 & 43.51 \\
-    single\_scan6 & 500 & 20.71 & 24.13 & 41.42 \\
+    single\_scan0 & 500 & 28.20 & 35.45 \\
+    single\_scan1 & 500 & 17.30 & 57.80 \\
+    single\_scan2 & 500 & 15.69 & 63.71 \\
+    single\_scan3 & 500 & 25.04 & 39.92 \\
+    single\_scan4 & 500 & 21.97 & 45.49 \\
+    single\_scan5 & 500 & 22.97 & 43.51 \\
+    single\_scan6 & 500 & 24.13 & 41.42 \\
 \end{tabular}

### 2018-03-10 Chapters/sdm-performance.tex
SDM Performance scan, read, write for Amazon EC2 instances

@@
 \end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 15.44 & 64.73 \\
+    Thread scan & 5000 & 30.47 & 32.81 \\
+    OpenCL scan & 5000 & 93.63 & 10.67 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 1000 & 23.44 & 42.64 \\
+    Thread single read & 1000 & 28.26 & 35.37 \\
+    OpenCL write & 1000 & 51.78 & 19.31 \\
+    OpenCL single read & 1000 & 85.31 & 11.72 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 52.35 & 19.09 \\
+    Thread scan & 5000 & 100.48 & 9.95 \\
+    OpenCL scan & 5000 & 145.18 & 6.88 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 2000 & 84.73 & 11.80 \\
+    Thread single read & 2000 & 95.26 & 10.49 \\
+    OpenCL write & 2000 & 113.01 & 8.84 \\
+    OpenCL single read & 2000 & 133.19 & 7.50 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 100 & 1.66 & 600.75 \\
+    Thread scan & 500 & 3.37 & 296.56 \\
+    OpenCL scan & 1000 & 21.39 & 46.73 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 500 & 2.60 & 383.50 \\
+    Thread single read & 500 & 3.24 & 307.97 \\
+    OpenCL write & 1000 & 8.18 & 122.17 \\
+    OpenCL single read & 55.47 & 18.02 & 55.47 \\
+    \hline
+\end{tabular}
+\caption{EC p2.xlarge 10,000-bits}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 17.13 & 58.34 \\
+    Thread scan & 5000 & 60.98 & 16.39 \\
+    OpenCL scan & 5000 & 1175.88 & 0.85 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 1000 & 35.12 & 28.47 \\
+    Thread single read & 1000 & 57.31 & 17.44 \\
+    OpenCL write & 1000 & 113.98 & 8.77 \\
+    OpenCL single read & 1000 & 549.12 & 1.82 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 56.80 & 17.60 \\
+    Thread scan & 5000 & 192.54 & 5.19 \\
+    OpenCL scan & 5000 & 1565.28 & 0.63 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 2000 & 131.64 & 7.59 \\
+    Thread single read & 2000 & 161.97 & 6.17 \\
+    OpenCL write & 2000 & 427.38 & 2.33 \\
+    OpenCL single read & 2000 & 987.95 & 1.01 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 100 & 1.84 & 540.97 \\
+    Thread scan & 500 & 5.03 & 198.74 \\
+    OpenCL scan & 1000 & 174.05 & 5.74 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 500 & 4.50 & 222.08 \\
+    Thread single read & 500 & 6.89 & 145.01 \\
+    OpenCL write & 1000 & 12.42 & 80.48 \\
+    OpenCL single read & 1000 & 71.46 & 13.99 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.}
+\end{table}

### 2018-03-11 Chapters/sdm-performance.tex
SDM Performance

@@
 
-We have executed the same performance test in each device.
-The test has 3 parts.
-The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans.
-The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners.
-The code is available in the ``Performance test'' notebook \citep{sdmframework}.
+We have created a benchmark to be able to compare the performance in different devices.
+So, the same performance test was executable in our devices and the results were reported in tables and figures.
+The benchmark has 3 parts: (i) the first part consists of comparing the available OpenCL kernels to find which works best for that device; than (ii) the second part consists of comparing the linear scanner, the thread scanner, and the OpenCL scanners with the best kernel found in part one; finally, (iii) the third part consists of comparing read and write operations using the thread scanner and the OpenCL scanner with the best kernel.
+Each part was run for three SDM setups: (i) $n=1,000$, $r=451$, and $H=1,000,000$; (ii) $n=256$, $r=103$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+The whole source code is available in the ``Performance test'' notebook \citep{sdmframework}.
+We would like to invite you to run this benchmark in your device and send us the results.
 
-Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+Our first device was a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+It was not possible to run the 10,000-bits in this device because there was no memory available --- it would use 37.25 GB of memory.
 
-Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
+Our second device was an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
 The complete results are presented in Figures \ref{fig:perf-imac-read-256}, \ref{fig:perf-imac-write-256}, \ref{fig:perf-imac-read-1000}, \ref{fig:perf-imac-write-1000}, \ref{fig:perf-imac-scan-256}, \ref{fig:perf-imac-scan-1000}, and \ref{fig:perf-imac-scan-10k} and Tables \ref{tab:perf-imac-256}, \ref{tab:perf-imac-1000}, and \ref{tab:perf-imac-10k}.
 
-Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+Beyond that, we were also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+
+\section{Kernels comparison}
+
+OpenCL is a framework for writting software that executes in \emph{heterogeneous} devices \citep{munshi2009opencl}, like CPUs, GPUs, and other co-processor for hardware acceleration.
+Because they are heterogogeneous, they may differ a lot in architecture and performance, which means there is no one-size-fits-all kernel.
+The slower kernel for one device may be the fastest for another device, and it really happened in our case.
+
+Our framework has 7 kernels available: single\_scan0, single\_scan1, single\_scan2, single\_scan3, single\_scan4, single\_scan5, single\_scan5\_unroll, and single\_scan6.
+Each scan uses a different algorithm to do exactly the same thing: calculate which hard-locations are inside the circle of the given bitstring.
+They differ in how they split the work between work-groups and how they explore the GPU architecture to obtain performance gains.
+
+The difference in which kernel is the best depends also in the SDM setup.
+The best kernel for the 1,000-bits SDM in the MacBook Pro was single\_scan1 with average scan time of 13.22ms (Table \ref{tab:perf-macbook-1000}).
+But, for the 256-bits SDM in the same device, it was single\_scan0 with average scan time of 8.36ms (Table \ref{tab:perf-macbook-256}).
+The comparison of all kernels in MacBook Pro is available in Figures \ref{fig:perf-macbook-kernels-1000} and \ref{fig:perf-macbook-kernels-256}.
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 1000 & 42.36 & 23.60 \\
+    single\_scan1 & 1000 & 75.63 & 13.22 \\
+    single\_scan2 & 1000 & 21.14 & 47.28 \\
+    single\_scan3 & 1000 & 30.24 & 33.06 \\
+    single\_scan4 & 1000 & 24.79 & 40.32 \\
+    single\_scan5 & 1000 & 31.73 & 31.51 \\
+    single\_scan5\_unroll & 1000 & 36.28 & 27.55 \\
+    single\_scan6 & 1000 & 25.32 & 39.48 \\
+    \hline
+\end{tabular}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
+\label{tab:perf-macbook-kernels-1000}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 1000 & 119.49 & 8.36 \\
+    single\_scan1 & 1000 & 95.82 & 10.43 \\
+    single\_scan2 & 1000 & 42.58 & 23.48 \\
+    single\_scan3 & 1000 & 39.19 & 25.51 \\
+    single\_scan4 & 1000 & 23.58 & 42.39 \\
+    single\_scan5 & 1000 & 40.94 & 24.42 \\
+    single\_scan5\_unroll & 1000 & 43.91 & 22.77 \\
+    single\_scan6 & 1000 & 23.70 & 42.18 \\
+    \hline
+\end{tabular}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+\label{tab:perf-macbook-kernels-256}}
+\end{table}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Kernel comparisons in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
+\label{fig:perf-macbook-kernels-1000}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Kernel comparisons in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
+\label{fig:perf-macbook-kernels-256}}
+\end{figure}
+
+\section{Scanners comparison}
 
@@
 Our conclusion is that, if one is really concerned about performance, one should fine tune the OpenCL kernel for one's GPU. It would always be faster than running in the CPU.
-
-Kernel comparisons
 
@@
 \caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-256}}
+\label{tab:perf-imac-kernels-1000}
 \end{table}
@@
     single\_scan6 & 3000 & 228.92 & 4.36 \\
+    \hline
 \end{tabular}
 \caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
+\label{tab:perf-imac-kernels-256}}
 \end{table}
@@
     single\_scan6 & 500 & 91.21 & 10.96 \\
+    \hline
 \end{tabular}
 \caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4850$.
-\label{tab:perf-imac-256}}
+\label{tab:perf-imac-kernels-10k}}
 \end{table}
@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 42.36 & 23.60 \\
-    single\_scan1 & 1000 & 75.63 & 13.22 \\
-    single\_scan2 & 1000 & 21.14 & 47.28 \\
-    single\_scan3 & 1000 & 30.24 & 33.06 \\
-    single\_scan4 & 1000 & 24.79 & 40.32 \\
-    single\_scan5 & 1000 & 31.73 & 31.51 \\
-    single\_scan5\_unroll & 1000 & 36.28 & 27.55 \\
-    single\_scan6 & 1000 & 25.32 & 39.48 \\
-\end{tabular}
-\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 119.49 & 8.36 \\
-    single\_scan1 & 1000 & 95.82 & 10.43 \\
-    single\_scan2 & 1000 & 42.58 & 23.48 \\
-    single\_scan3 & 1000 & 39.19 & 25.51 \\
-    single\_scan4 & 1000 & 23.58 & 42.39 \\
-    single\_scan5 & 1000 & 40.94 & 24.42 \\
-    single\_scan5\_unroll & 1000 & 43.91 & 22.77 \\
-    single\_scan6 & 1000 & 23.70 & 42.18 \\
-\end{tabular}
-\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
     single\_scan0 & 3000 & 1433.27 & 0.69 \\
@@
     single\_scan6 & 3000 & 958.50 & 1.04 \\
+    \hline
 \end{tabular}
@@
     single\_scan6 & 3000 & 986.06 & 1.01 \\
+    \hline
 \end{tabular}
@@
     single\_scan6 & 500 & 167.74 & 5.96 \\
+    \hline
 \end{tabular}
@@
     single\_scan6 & 3000 & 88.21 & 11.33 \\
+    \hline
 \end{tabular}
@@
     single\_scan6 & 3000 & 97.10 & 10.29 \\
+    \hline
 \end{tabular}
@@
     single\_scan6 & 500 & 24.13 & 41.42 \\
+    \hline
 \end{tabular}

### 2018-03-11 Chapters/sdm-performance-tables.tex
SDM Performance almost done

--- 2018-03-10 Chapters/sdm-performance.tex
@@
-% !TEX root = ../partial-sdm.tex
-
-Performance matters --- and has always mattered.
-If an experiment takes a few seconds, there is no point arguing whether we should try it.
-If it takes a few hours, maybe we should think first.
-If it takes a few days, it is more important to make a good plan.
-As SDM consumes a lot of both processor and memory resources, some experiments may take a long time.
-
-Each scan a 1,000-bits SDM with 1,000,000 hard-locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
-So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
-The heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} executed $3.05 \cdot 10^{15}$ bit compares.
-For comparison, the number of seconds since Jesus's birth is 63,639,648,000 = $6.36 \cdot 10^{10}$.
-The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$.
-There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.
-A modern laptop can increment a counter $5 \cdot 10^{8}$ times per second.
-Hence, a naive implementation of SDM may take several hours to simply write 10,000 random bitstrings.
-
-Amazon EC2 p3.2xlarge has generated the heatmap of Figure \ref{fig:cir-dist-10k-writes-kanerva} in 15 minutes and 3 seconds.
-It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.
-It is a great improvement when compared to the first versions of the code which took 16 hours to generate the same heatmap (and its memory use was already optimized and the computations were distributed in threads).
-
-We have executed the same performance test in each device.
-The test has 3 parts.
-The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans.
-The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners.
-The code is available in the ``Performance test'' notebook \citep{sdmframework}.
-
-Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-
-Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
-The complete results are presented in Figures \ref{fig:perf-imac-read-256}, \ref{fig:perf-imac-write-256}, \ref{fig:perf-imac-read-1000}, \ref{fig:perf-imac-write-1000}, \ref{fig:perf-imac-scan-256}, \ref{fig:perf-imac-scan-1000}, and \ref{fig:perf-imac-scan-10k} and Tables \ref{tab:perf-imac-256}, \ref{tab:perf-imac-1000}, and \ref{tab:perf-imac-10k}.
-
-Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
-
-It is interesting that which scanner is faster depends also on the SDM settings.
-In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL, but for a 256-bit SDM was the threads.
-What happened here is that the OpenCL kernel chosen was a generic one which performs the scan for any SDM.
-It is always possible to optimize the OpenCL kernel to a specific setting, and it would be faster than the threads.
-By default, the framework chooses a generic kernel which we believe would be reasonable for the most common setups.
-% TODO Run the optimized kernel and show the results.
-
-Most of the time, the bottleneck of both the read and the write operations is the scanner.
-But we have optimized the OpenCL kernel so much for the iMac 2017 that scanner not the bottleneck anymore.
-In the writing operation, it took the same time to scan and to update the counters.
-It is impressive because it update, on average, 1,000 counters of 32-bit integers each, and it was as fast as (i) sending the command to the GPU, (ii) performing 1 billion bit compares, and (iii) downloading the response from the GPU.
-
-Our conclusion is that, if one is really concerned about performance, one should fine tune the OpenCL kernel for one's GPU. It would always be faster than running in the CPU.
-
-Kernel comparisons
-
 \begin{table}[!htb]
@@
     single\_scan5 & 3000 & 268.29 & 3.72 \\
-    single\_scan5\_unroll & 3000 & 276.70 & 3.6139983336130777 \\
-    single\_scan6 & 3000 & 248.26 & 4.02 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 3000 & 329.71 & 3.03 \\
-    single\_scan1 & 3000 & 347.42 & 2.87 \\
-    single\_scan2 & 3000 & 261.73 & 3.82 \\
-    single\_scan3 & 3000 & 268.16 & 3.72 \\
-    single\_scan4 & 3000 & 223.18 & 4.48 \\
-    single\_scan5 & 3000 & 265.59 & 3.76 \\
-    single\_scan5\_unroll & 3000 & 263.24 & 3.79 \\
-    single\_scan6 & 3000 & 228.92 & 4.36 \\
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 500 & 16.37 & 61.06 \\
-    single\_scan1 & 500 & 22.24 & 44.96 \\
-    single\_scan2 & 500 & 22.23 & 44.98 \\
-    single\_scan3 & 500 & 78.92 & 12.67 \\
-    single\_scan4 & 500 & 87.30 & 11.45 \\
-    single\_scan5 & 500 & 79.43 & 12.58 \\
-    single\_scan5\_unroll & 500 & 87.88 & 11.37 \\
-    single\_scan6 & 500 & 91.21 & 10.96 \\
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4850$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 42.36 & 23.60 \\
-    single\_scan1 & 1000 & 75.63 & 13.22 \\
-    single\_scan2 & 1000 & 21.14 & 47.28 \\
-    single\_scan3 & 1000 & 30.24 & 33.06 \\
-    single\_scan4 & 1000 & 24.79 & 40.32 \\
-    single\_scan5 & 1000 & 31.73 & 31.51 \\
-    single\_scan5\_unroll & 1000 & 36.28 & 27.55 \\
-    single\_scan6 & 1000 & 25.32 & 39.48 \\
-\end{tabular}
-\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 119.49 & 8.36 \\
-    single\_scan1 & 1000 & 95.82 & 10.43 \\
-    single\_scan2 & 1000 & 42.58 & 23.48 \\
-    single\_scan3 & 1000 & 39.19 & 25.51 \\
-    single\_scan4 & 1000 & 23.58 & 42.39 \\
-    single\_scan5 & 1000 & 40.94 & 24.42 \\
-    single\_scan5\_unroll & 1000 & 43.91 & 22.77 \\
-    single\_scan6 & 1000 & 23.70 & 42.18 \\
-\end{tabular}
-\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 1433.27 & 0.69 \\
-    single\_scan1 & 3000 & 1822.51 & 0.54 \\
-    single\_scan2 & 3000 & 1165.06 & 0.85 \\
-    single\_scan3 & 3000 & 979.93 & 1.02 \\
-    single\_scan4 & 3000 & 906.35 & 1.10 \\
-    single\_scan5 & 3000 & 1049.21 & 0.95 \\
-    single\_scan6 & 3000 & 958.50 & 1.04 \\
-\end{tabular}
-\caption{EC2 p3.2xlarge 1000 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 2729.58 & 0.36 \\
-    single\_scan1 & 3000 & 2703.49 & 0.36 \\
-    single\_scan2 & 3000 & 1354.12 & 0.73 \\
-    single\_scan3 & 3000 & 1573.43 & 0.63 \\
-    single\_scan4 & 3000 & 951.22 & 1.05 \\
-    single\_scan5 & 3000 & 1598.69 & 0.62 \\
-    single\_scan6 & 3000 & 986.06 & 1.01 \\
-\end{tabular}
-\caption{EC2 p3.2xlarge 256 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 500 & 48.54 & 20.60 \\
-    single\_scan1 & 500 & 202.34 & 4.94 \\
-    single\_scan2 & 500 & 199.29 & 5.01 \\
-    single\_scan3 & 500 & 164.66 & 6.07 \\
-    single\_scan4 & 500 & 166.70 & 5.99 \\
-    single\_scan5 & 500 & 186.26 & 5.36 \\
-    single\_scan6 & 500 & 167.74 & 5.96 \\
-\end{tabular}
-\caption{EC2 p3.2xlarge 10k bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 263.61 & 3.79 \\
-    single\_scan1 & 3000 & 253.45 & 3.94 \\
-    single\_scan2 & 3000 & 117.06 & 8.54 \\
-    single\_scan3 & 3000 & 102.70 & 9.73 \\
-    single\_scan4 & 3000 & 87.82 & 11.38 \\
-    single\_scan5 & 3000 & 100.46 & 9.95 \\
-    single\_scan6 & 3000 & 88.21 & 11.33 \\
-\end{tabular}
-\caption{EC2 p2.xlarge 1000 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 1312.85 & 0.76 \\
-    single\_scan1 & 3000 & 1237.20 & 0.80 \\
-    single\_scan2 & 3000 & 178.85 & 5.59 \\
-    single\_scan3 & 3000 & 158.43 & 6.31 \\
-    single\_scan4 & 3000 & 97.14 & 10.29 \\
-    single\_scan5 & 3000 & 149.45 & 6.69 \\
-    single\_scan6 & 3000 & 97.10 & 10.29 \\
-\end{tabular}
-\caption{EC2 p2.xlarge 256 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 500 & 28.20 & 35.45 \\
-    single\_scan1 & 500 & 17.30 & 57.80 \\
-    single\_scan2 & 500 & 15.69 & 63.71 \\
-    single\_scan3 & 500 & 25.04 & 39.92 \\
-    single\_scan4 & 500 & 21.97 & 45.49 \\
-    single\_scan5 & 500 & 22.97 & 43.51 \\
-    single\_scan6 & 500 & 24.13 & 41.42 \\
-\end{tabular}
-\caption{EC2 p2.xlarge 10k bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-256.png}
-\caption{Time to run a single read in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-read-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-256.png}
-\caption{Time to run one write in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-write-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-1000.png}
-\caption{Time to run one write in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-write-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-1000.png}
-\caption{Time to run a single read in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-read-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
-\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-scan-256}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
-\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-scan-1000}}
-\end{figure}
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
-\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
-\label{fig:perf-imac-scan-10k}}
-\end{figure}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
+    single\_scan5\_unroll & 3000 & 276.70 & 3.61 \\
+    single\_scan6 & 3000 & 248.26 & 4.02 \\ \hline
     \hline
@@
     & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 175.48 & 5.69 \\
-    Thread scan & 5000 & 352.63 & 2.83 \\
-    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
+    single\_scan0 & 3000 & 199.62 & 5.00 \\
+    single\_scan1 & 3000 & 252.64 & 3.95 \\
+    single\_scan2 & 3000 & 218.38 & 4.57 \\
+    single\_scan3 & 3000 & 271.15 & 3.68 \\
+    single\_scan4 & 3000 & 247.28 & 4.04 \\
+    single\_scan5 & 3000 & 268.29 & 3.72 \\
+    single\_scan5\_unroll & 3000 & 276.70 & 3.61 \\
+    single\_scan6 & 3000 & 248.26 & 4.02 \\
+    \hline
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
+\label{tab:perf-imac-kernels-1000}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 3000 & 329.71 & 3.03 \\
+    single\_scan1 & 3000 & 347.42 & 2.87 \\
+    single\_scan2 & 3000 & 261.73 & 3.82 \\
+    single\_scan3 & 3000 & 268.16 & 3.72 \\
+    single\_scan4 & 3000 & 223.18 & 4.48 \\
+    single\_scan5 & 3000 & 265.59 & 3.76 \\
+    single\_scan5\_unroll & 3000 & 263.24 & 3.79 \\
+    single\_scan6 & 3000 & 228.92 & 4.36 \\
+    \hline
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+\label{tab:perf-imac-kernels-256}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 500 & 16.37 & 61.06 \\
+    single\_scan1 & 500 & 22.24 & 44.96 \\
+    single\_scan2 & 500 & 22.23 & 44.98 \\
+    single\_scan3 & 500 & 78.92 & 12.67 \\
+    single\_scan4 & 500 & 87.30 & 11.45 \\
+    single\_scan5 & 500 & 79.43 & 12.58 \\
+    single\_scan5\_unroll & 500 & 87.88 & 11.37 \\
+    single\_scan6 & 500 & 91.21 & 10.96 \\
+    \hline
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4850$.
+\label{tab:perf-imac-kernels-10k}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 198.03 & 5.04 \\
+    Thread scan & 5000 & 341.77 & 2.92 \\
+    OpenCL scan & 5000 & 355.46 & 2.81 \\ \hline
     \hline
@@
 \label{tab:perf-imac-10k}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 15.44 & 64.73 \\
-    Thread scan & 5000 & 30.47 & 32.81 \\
-    OpenCL scan & 5000 & 93.63 & 10.67 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 1000 & 23.44 & 42.64 \\
-    Thread single read & 1000 & 28.26 & 35.37 \\
-    OpenCL write & 1000 & 51.78 & 19.31 \\
-    OpenCL single read & 1000 & 85.31 & 11.72 \\
-    \hline
-\end{tabular}
-\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 52.35 & 19.09 \\
-    Thread scan & 5000 & 100.48 & 9.95 \\
-    OpenCL scan & 5000 & 145.18 & 6.88 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 2000 & 84.73 & 11.80 \\
-    Thread single read & 2000 & 95.26 & 10.49 \\
-    OpenCL write & 2000 & 113.01 & 8.84 \\
-    OpenCL single read & 2000 & 133.19 & 7.50 \\
-    \hline
-\end{tabular}
-\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 100 & 1.66 & 600.75 \\
-    Thread scan & 500 & 3.37 & 296.56 \\
-    OpenCL scan & 1000 & 21.39 & 46.73 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 500 & 2.60 & 383.50 \\
-    Thread single read & 500 & 3.24 & 307.97 \\
-    OpenCL write & 1000 & 8.18 & 122.17 \\
-    OpenCL single read & 55.47 & 18.02 & 55.47 \\
-    \hline
-\end{tabular}
-\caption{EC p2.xlarge 10,000-bits}
-\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.}
 \end{table}
@@
 \end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 56.80 & 17.60 \\
+    Thread scan & 5000 & 192.54 & 5.19 \\
+    OpenCL scan & 5000 & 1565.28 & 0.63 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 2000 & 131.64 & 7.59 \\
+    Thread single read & 2000 & 161.97 & 6.17 \\
+    OpenCL write & 2000 & 427.38 & 2.33 \\
+    OpenCL single read & 2000 & 987.95 & 1.01 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 100 & 1.84 & 540.97 \\
+    Thread scan & 500 & 5.03 & 198.74 \\
+    OpenCL scan & 1000 & 174.05 & 5.74 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 500 & 4.50 & 222.08 \\
+    Thread single read & 500 & 6.89 & 145.01 \\
+    OpenCL write & 1000 & 12.42 & 80.48 \\
+    OpenCL single read & 1000 & 71.46 & 13.99 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.}
+\end{table}
+
+macbook scan 1000
+	Loops	Total time	Scans per second	Time per scan (ms)
+Linear scan	1000	17.9807908535	55.6149063825	17.9807908535
+Thread scan	3000	30.5303659439	98.2628248057	10.176788648
+OpenCL scan	3000	37.0593481064	80.9512350673	12.3531160355
+
+macbook scan 256
+	Loops	Total time	Scans per second	Time per scan (ms)
+Linear scan	1000	9.07448506355	110.199090416	9.07448506355
+Thread scan	3000	15.4379580021	194.326218506	5.1459860007
+OpenCL scan	3000	24.1792058945	124.07355366	8.05973529816
+
+macbook read/write 1000
+	Loops	Total time	Operation per second	Time per operation (ms)
+Thread write	1000	14.1335010529	70.7538773486	14.1335010529
+Thread single read	1000	11.1241891384	89.8941925166	11.1241891384
+OpenCL write	1000	18.5585401058	53.8835487219	18.5585401058
+OpenCL single read	1000	13.0633058548	76.5503013644	13.0633058548
+
+macbook read/write 256
+	Loops	Total time	Operation per second	Time per operation (ms)
+Thread write	2000	13.4491240978	148.708569082	6.72456204891
+Thread single read	2000	11.7688980103	169.939445329	5.88444900513
+OpenCL write	2000	12.7978179455	156.276640949	6.39890897274
+OpenCL single read	2000	10.5222630501	190.073180121	5.26113152504
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 1000 & 42.36 & 23.60 \\
+    single\_scan1 & 1000 & 75.63 & 13.22 \\
+    single\_scan2 & 1000 & 21.14 & 47.28 \\
+    single\_scan3 & 1000 & 30.24 & 33.06 \\
+    single\_scan4 & 1000 & 24.79 & 40.32 \\
+    single\_scan5 & 1000 & 31.73 & 31.51 \\
+    single\_scan5\_unroll & 1000 & 36.28 & 27.55 \\
+    single\_scan6 & 1000 & 25.32 & 39.48 \\
+    \hline
+\end{tabular}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
+\label{tab:perf-macbook-kernels-1000}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    single\_scan0 & 1000 & 119.49 & 8.36 \\
+    single\_scan1 & 1000 & 95.82 & 10.43 \\
+    single\_scan2 & 1000 & 42.58 & 23.48 \\
+    single\_scan3 & 1000 & 39.19 & 25.51 \\
+    single\_scan4 & 1000 & 23.58 & 42.39 \\
+    single\_scan5 & 1000 & 40.94 & 24.42 \\
+    single\_scan5\_unroll & 1000 & 43.91 & 22.77 \\
+    single\_scan6 & 1000 & 23.70 & 42.18 \\
+    \hline
+\end{tabular}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+\label{tab:perf-macbook-kernels-256}}
+\end{table}
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{| l | r | r | r |}
+    \hline
+    & Loops & Scans / second & Scan time (ms) \\ \hline
+    Linear scan & 1000 & 17.13 & 58.34 \\
+    Thread scan & 5000 & 60.98 & 16.39 \\
+    OpenCL scan & 5000 & 1175.88 & 0.85 \\ \hline
+    \hline
+    & Loops & Ops / second & Op. time (ms) \\ \hline
+    Thread write & 1000 & 35.12 & 28.47 \\
+    Thread single read & 1000 & 57.31 & 17.44 \\
+    OpenCL write & 1000 & 113.98 & 8.77 \\
+    OpenCL single read & 1000 & 549.12 & 1.82 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.}
+\end{table}

### 2018-03-11 Chapters/sdm-performance.tex
SDM Performance almost done

--- 2018-03-10 Chapters/sdm-performance.tex
@@
 
-We have executed the same performance test in each device.
-The test has 3 parts.
-The first part consists of: 1,000 linear scans, 5,000 thread scans, and 5,000 OpenCL scans.
-The second part consists of 5,000 writes, 5,000 single reads, and 1,000 6-iterative readings, with both thread and opencl scanners.
-The code is available in the ``Performance test'' notebook \citep{sdmframework}.
+We have created a benchmark to be able to compare the performance in different devices.
+So, the same performance test was executable in our devices and the results were reported in tables and figures.
+The benchmark has 3 parts: (i) the first part consists of comparing the available OpenCL kernels to find which works best for that device; than (ii) the second part consists of comparing the linear scanner, the thread scanner, and the OpenCL scanners with the best kernel found in part one; finally, (iii) the third part consists of comparing read and write operations using the thread scanner and the OpenCL scanner with the best kernel.
+Each part was run for three SDM setups: (i) $n=1,000$, $r=451$, and $H=1,000,000$; (ii) $n=256$, $r=103$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+The whole source code is available in the ``Performance test'' notebook \citep{sdmframework}.
+We would like to invite you to run this benchmark in your device and send us the results.
 
-Our first device is a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+Our first device was a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+It was not possible to run the 10,000-bits in this device because there was no memory available --- it would use 37.25 GB of memory.
+For its results, see Table \ref{tab:perf-macbook}
 
-Our second device is an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU.
-The complete results are presented in Figures \ref{fig:perf-imac-read-256}, \ref{fig:perf-imac-write-256}, \ref{fig:perf-imac-read-1000}, \ref{fig:perf-imac-write-1000}, \ref{fig:perf-imac-scan-256}, \ref{fig:perf-imac-scan-1000}, and \ref{fig:perf-imac-scan-10k} and Tables \ref{tab:perf-imac-256}, \ref{tab:perf-imac-1000}, and \ref{tab:perf-imac-10k}.
+Our second device was an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU. For its results, see Table \ref{tab:perf-ec2-imac}.
 
-Beyond that, we are also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU, and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+Beyond that, we were also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU (see Table \ref{tab:perf-p2}), and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU (see Table \ref{tab:perf-ec2-p3}).
 
-It is interesting that which scanner is faster depends also on the SDM settings.
-In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL, but for a 256-bit SDM was the threads.
+\section{Kernels comparison}
+
+OpenCL is a framework for writting software that executes in \emph{heterogeneous} devices \citep{munshi2009opencl}, like CPUs, GPUs, and other co-processor for hardware acceleration.
+Because they are heterogogeneous, they may differ a lot in architecture and performance, which means there is no one-size-fits-all kernel.
+The slower kernel for one device may be the fastest for another device, and it really happened in our case.
+
+Our framework has 8 kernels available: single\_scan0, single\_scan1, single\_scan2, single\_scan3, single\_scan4, single\_scan5, single\_scan5\_unroll, and single\_scan6.
+Each scan uses a different algorithm to do exactly the same thing: calculate which hard-locations are inside the circle of the given bitstring.
+They differ in how they split the work between work-groups and how they explore the GPU architecture to obtain performance gains.
+
+The difference in which kernel is the best depends also in the SDM setup.
+The best kernel for the 1,000-bits SDM in the iMac 2017 was single\_scan5\_unroll with average scan time of 3.61ms; but, for the 256-bits SDM in the same device, it was single\_scan0 with average scan time of 3.03ms; while, for the 10,000-bits SDM in the same device, it was single\_scan6 with 10.96ms (see Table \ref{tab:perf-imac}).
+
+We recommend users to run a specific kernel comparison test for their own GPU and SDM settings.
+
+\section{Scanners comparison}
+
+In this section, we are comparing the OpenCL scanner (with the best kernel) with the linear scanner and the threads scanner.
+Again, which one is faster depends on the SDM settings.
+In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL scanner with , but for a 256-bit SDM was the threads scanner.
+
 What happened here is that the OpenCL kernel chosen was a generic one which performs the scan for any SDM.
@@
 
-Most of the time, the bottleneck of both the read and the write operations is the scanner.
-But we have optimized the OpenCL kernel so much for the iMac 2017 that scanner not the bottleneck anymore.
-In the writing operation, it took the same time to scan and to update the counters.
-It is impressive because it update, on average, 1,000 counters of 32-bit integers each, and it was as fast as (i) sending the command to the GPU, (ii) performing 1 billion bit compares, and (iii) downloading the response from the GPU.
+We can notice that Amazon EC2 p3.2xlarge and p2.xlarge's linear and thread scanners, both running on CPU, were much slower than the CPU of both the personal MacBook Pro and the iMac 2017.
+As Amazon EC2 are virtual machines with GPU devices, their CPU is shared with other virtual machines which reduces CPU power significantly.
+Hence, for both virtual machines we have tested, using the GPU really boost the performance, but using the CPU should be avoided.
+See Tables \ref{tab:perf-imac}
 
-Our conclusion is that, if one is really concerned about performance, one should fine tune the OpenCL kernel for one's GPU. It would always be faster than running in the CPU.
+\section{Read and write operations}
 
-Kernel comparisons
+Even though scan is an important part of the operations, we are really interested in the performance of the operations themselves.
+Comparing the times of the thread and OpenCL scanners with the times of their respective operations (either read or write), we can notice that their difference remains almost constant, which means the operation time itself is negligible when compared to the scan time.
+In other words, in order to gain even more performance, we have to pursue ways to improve the scan.
 
@@
 \centering
-\begin{tabular}{| l | r | r | r |}
+\begin{tabular}{lrrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
     \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 3000 & 199.62 & 5.00 \\
-    single\_scan1 & 3000 & 252.64 & 3.95 \\
-    single\_scan2 & 3000 & 218.38 & 4.57 \\
-    single\_scan3 & 3000 & 271.15 & 3.68 \\
-    single\_scan4 & 3000 & 247.28 & 4.04 \\
-    single\_scan5 & 3000 & 268.29 & 3.72 \\
-    single\_scan5\_unroll & 3000 & 276.70 & 3.6139983336130777 \\
-    single\_scan6 & 3000 & 248.26 & 4.02 \\
+	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    single\_scan0 &  8.36 & 23.60 &  \\
+    single\_scan1 & 10.43 & 13.22 &  \\
+    single\_scan2 & 23.48 & 47.28 &  \\
+    single\_scan3 & 25.51 & 33.06 &  \\
+    single\_scan4 & 42.39 & 40.32 &  \\
+    single\_scan5 & 24.42 & 31.51 &  \\
+	single\_scan5\_unroll & 22.77 & 27.55 & & \\
+	single\_scan6 & 42.18 & 39.48 &  \\ \hline
     \hline
+	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 9.07 & 17.98 & \\
+    Thread scan & 5.14 & 10.17 & \\
+    OpenCL scan & 8.05 & 12.35 & \\ \hline
+    \hline
+	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+	Thread write       & 6.72 & 14.13 & \\
+	Thread single read & 5.88 & 11.12 & \\
+	OpenCL write       & 6.39 & 18.55 & \\
+	OpenCL single read & 5.26 & 13.06 & \\
+    \bottomrule
 \end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-256}}
+\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for $n=10,000$ because memory is not enough on either RAM or GPU---it would consume 37.25 GB of RAM and 1.2GB of memory in the GPU.
+For the histogram of durations, see Figures \ref{fig:perf-macbook-kernels}, \ref{fig:perf-macbook-scanners}, \ref{fig:perf-macbook-read}, and \ref{fig:perf-macbook-write}.
+\label{tab:perf-macbook}}
 \end{table}
@@
 \centering
-\begin{tabular}{| l | r | r | r |}
+\begin{tabular}{lrrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
     \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 3000 & 329.71 & 3.03 \\
-    single\_scan1 & 3000 & 347.42 & 2.87 \\
-    single\_scan2 & 3000 & 261.73 & 3.82 \\
-    single\_scan3 & 3000 & 268.16 & 3.72 \\
-    single\_scan4 & 3000 & 223.18 & 4.48 \\
-    single\_scan5 & 3000 & 265.59 & 3.76 \\
-    single\_scan5\_unroll & 3000 & 263.24 & 3.79 \\
-    single\_scan6 & 3000 & 228.92 & 4.36 \\
+	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    single\_scan0 & 3.03 & 5.00 & 61.06 \\
+    single\_scan1 & 2.87 & 3.95 & 44.96 \\
+    single\_scan2 & 3.82 & 4.57 & 44.98 \\
+    single\_scan3 & 3.72 & 3.68 & 12.67 \\
+    single\_scan4 & 4.48 & 4.04 & 11.45 \\
+    single\_scan5 & 3.76 & 3.72 & 12.58 \\
+	single\_scan5\_unroll & 3.79 & 3.61 & 11.37 \\
+	single\_scan6 & 4.36 & 4.02 & 10.96 \\ \hline
+    \hline
+	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 5.04 & 12.25 & 116.38 \\
+    Thread scan & 2.92 &  6.95 &  53.56 \\
+    OpenCL scan & 2.81 &  4.20 &  12.95 \\ \hline
+    \hline
+	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Thread write       & 3.28 & 13.34 \\
+    Thread single read & 2.55 & 10.39 \\
+    OpenCL write       & 2.64 &  7.90 \\
+    OpenCL single read & 2.14 &  5.25 \\
+    \bottomrule
 \end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for read and write operations with $n=10,000$ because RAM is not enough to allocate the counters---it would consume 37.25 GB of RAM.
+For the histogram of durations, see Figures \ref{fig:perf-imac-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.
+\label{tab:perf-imac}}
 \end{table}
@@
 \centering
-\begin{tabular}{| l | r | r | r |}
+\begin{tabular}{lrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
 	\hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 500 & 16.37 & 61.06 \\
-    single\_scan1 & 500 & 22.24 & 44.96 \\
-    single\_scan2 & 500 & 22.23 & 44.98 \\
-    single\_scan3 & 500 & 78.92 & 12.67 \\
-    single\_scan4 & 500 & 87.30 & 11.45 \\
-    single\_scan5 & 500 & 79.43 & 12.58 \\
-    single\_scan5\_unroll & 500 & 87.88 & 11.37 \\
-    single\_scan6 & 500 & 91.21 & 10.96 \\
+	Kernels & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    single\_scan0 &  0.76 &  3.79 & 35.45 \\
+    single\_scan1 &  0.80 &  3.94 & 57.80 \\
+    single\_scan2 &  5.59 &  8.54 & 63.71 \\
+    single\_scan3 &  6.31 &  9.73 & 39.92 \\
+    single\_scan4 & 10.29 & 11.38 & 45.49 \\
+    single\_scan5 &  6.69 &  9.95 & 43.51 \\
+	single\_scan5\_unroll & & & \\
+    single\_scan6 & 10.29 & 11.33 & 41.42 \\
+    \hline
+	Scanners & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 19.09 & 64.73 & 600.75 \\
+    Thread scan &  9.95 & 32.81 & 296.56 \\
+    OpenCL scan &  6.88 & 10.67 &  46.73 \\ \hline
+    \hline
+	Operations & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Thread write       & 11.80 & 42.64 & 383.50 \\
+    Thread single read & 10.49 & 35.37 & 307.97 \\
+    OpenCL write       &  8.84 & 19.31 & 122.17 \\
+    OpenCL single read &  7.50 & 11.72 &  55.47 \\
+    \bottomrule
 \end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4850$.
-\label{tab:perf-imac-256}}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
+Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
+The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
+For the histogram of durations, see Figures \ref{fig:perf-ec2-p2-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.
+\label{tab:perf-ec2-p2}}
 \end{table}
@@
 \centering
-\begin{tabular}{| l | r | r | r |}
+\begin{tabular}{lrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
 	\hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 42.36 & 23.60 \\
-    single\_scan1 & 1000 & 75.63 & 13.22 \\
-    single\_scan2 & 1000 & 21.14 & 47.28 \\
-    single\_scan3 & 1000 & 30.24 & 33.06 \\
-    single\_scan4 & 1000 & 24.79 & 40.32 \\
-    single\_scan5 & 1000 & 31.73 & 31.51 \\
-    single\_scan5\_unroll & 1000 & 36.28 & 27.55 \\
-    single\_scan6 & 1000 & 25.32 & 39.48 \\
+	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    \hline
+    single\_scan0 & 0.36 & 0.69 & 20.60 \\
+    single\_scan1 & 0.36 & 0.54 &  4.94 \\
+    single\_scan2 & 0.73 & 0.85 &  5.01 \\
+    single\_scan3 & 0.63 & 1.02 &  6.07 \\
+    single\_scan4 & 1.05 & 1.10 &  5.99 \\
+    single\_scan5 & 0.62 & 0.95 &  5.36 \\
+	single\_scan5\_unroll & & & \\
+    single\_scan6 & 1.01 & 1.04 &  5.96 \\ \hline
+    \hline
+	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 17.60 & 58.34 & 540.97 \\
+    Thread scan &  5.19 & 16.39 & 198.74 \\
+    OpenCL scan &  0.63 &  0.85 &   5.74 \\ \hline
+    \hline
+	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Thread write       & 7.59 & 28.47 & 222.08 \\
+    Thread single read & 6.17 & 17.44 & 145.01 \\
+    OpenCL write       & 2.33 &  8.77 &  80.48 \\
+    OpenCL single read & 1.01 &  1.82 &  13.99 \\
+    \hline
 \end{tabular}
-\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-256}}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
+The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
+For the histogram of durations, see Figures \ref{fig:perf-ec2-p3-kernels}, \ref{fig:perf-ec2-p3-scanners}, \ref{fig:perf-ec2-p3-read}, and \ref{fig:perf-ec2-p3-write}.
+\label{tab:perf-ec2-p3}}
 \end{table}
 
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    single\_scan0 & 1000 & 119.49 & 8.36 \\
-    single\_scan1 & 1000 & 95.82 & 10.43 \\
-    single\_scan2 & 1000 & 42.58 & 23.48 \\
-    single\_scan3 & 1000 & 39.19 & 25.51 \\
-    single\_scan4 & 1000 & 23.58 & 42.39 \\
-    single\_scan5 & 1000 & 40.94 & 24.42 \\
-    single\_scan5\_unroll & 1000 & 43.91 & 22.77 \\
-    single\_scan6 & 1000 & 23.70 & 42.18 \\
-\end{tabular}
-\caption{MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 1433.27 & 0.69 \\
-    single\_scan1 & 3000 & 1822.51 & 0.54 \\
-    single\_scan2 & 3000 & 1165.06 & 0.85 \\
-    single\_scan3 & 3000 & 979.93 & 1.02 \\
-    single\_scan4 & 3000 & 906.35 & 1.10 \\
-    single\_scan5 & 3000 & 1049.21 & 0.95 \\
-    single\_scan6 & 3000 & 958.50 & 1.04 \\
-\end{tabular}
-\caption{EC2 p3.2xlarge 1000 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 2729.58 & 0.36 \\
-    single\_scan1 & 3000 & 2703.49 & 0.36 \\
-    single\_scan2 & 3000 & 1354.12 & 0.73 \\
-    single\_scan3 & 3000 & 1573.43 & 0.63 \\
-    single\_scan4 & 3000 & 951.22 & 1.05 \\
-    single\_scan5 & 3000 & 1598.69 & 0.62 \\
-    single\_scan6 & 3000 & 986.06 & 1.01 \\
-\end{tabular}
-\caption{EC2 p3.2xlarge 256 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 500 & 48.54 & 20.60 \\
-    single\_scan1 & 500 & 202.34 & 4.94 \\
-    single\_scan2 & 500 & 199.29 & 5.01 \\
-    single\_scan3 & 500 & 164.66 & 6.07 \\
-    single\_scan4 & 500 & 166.70 & 5.99 \\
-    single\_scan5 & 500 & 186.26 & 5.36 \\
-    single\_scan6 & 500 & 167.74 & 5.96 \\
-\end{tabular}
-\caption{EC2 p3.2xlarge 10k bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 263.61 & 3.79 \\
-    single\_scan1 & 3000 & 253.45 & 3.94 \\
-    single\_scan2 & 3000 & 117.06 & 8.54 \\
-    single\_scan3 & 3000 & 102.70 & 9.73 \\
-    single\_scan4 & 3000 & 87.82 & 11.38 \\
-    single\_scan5 & 3000 & 100.46 & 9.95 \\
-    single\_scan6 & 3000 & 88.21 & 11.33 \\
-\end{tabular}
-\caption{EC2 p2.xlarge 1000 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 3000 & 1312.85 & 0.76 \\
-    single\_scan1 & 3000 & 1237.20 & 0.80 \\
-    single\_scan2 & 3000 & 178.85 & 5.59 \\
-    single\_scan3 & 3000 & 158.43 & 6.31 \\
-    single\_scan4 & 3000 & 97.14 & 10.29 \\
-    single\_scan5 & 3000 & 149.45 & 6.69 \\
-    single\_scan6 & 3000 & 97.10 & 10.29 \\
-\end{tabular}
-\caption{EC2 p2.xlarge 256 bits
-\label{tab:perf-imac-256}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Loops & Total time & Scans per second & Time per scan (ms) \\
-    single\_scan0 & 500 & 28.20 & 35.45 \\
-    single\_scan1 & 500 & 17.30 & 57.80 \\
-    single\_scan2 & 500 & 15.69 & 63.71 \\
-    single\_scan3 & 500 & 25.04 & 39.92 \\
-    single\_scan4 & 500 & 21.97 & 45.49 \\
-    single\_scan5 & 500 & 22.97 & 43.51 \\
-    single\_scan6 & 500 & 24.13 & 41.42 \\
-\end{tabular}
-\caption{EC2 p2.xlarge 10k bits
-\label{tab:perf-imac-256}}
-\end{table}
+% MacBook Pro
 
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-256.png}
-\caption{Time to run a single read in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-read-256}}
+\centering
+\subfloat[$n=256$, $r=103$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-kernels-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-kernels-1000.png}}
+
+\caption{Kernel comparisons for MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+\label{fig:perf-macbook-kernels}}
 \end{figure}
@@
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-256.png}
-\caption{Time to run one write in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-write-256}}
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-scan-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-scan-1000.png}}
+
+\caption{Scanner comparisons for MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+\label{fig:perf-macbook-scanners}}
 \end{figure}
@@
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-write-1000.png}
-\caption{Time to run one write in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-write-1000}}
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-read-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-read-1000.png}}
+
+\caption{Read operation comparisons for MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+\label{fig:perf-macbook-read}}
 \end{figure}
@@
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-read-1000.png}
-\caption{Time to run a single read in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-read-1000}}
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-write-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/macbook-write-1000.png}}
+
+\caption{Write operation comparisons for MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
+\label{fig:perf-macbook-write}}
+\end{figure}
+
+% iMac
+
+\begin{figure}[!htb]
+\centering
+\subfloat[$n=256$, $r=103$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/imac-kernels-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/imac-kernels-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/imac-kernels-10k.png}}
+
+\caption{Kernel comparisons for iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+\label{fig:perf-imac-kernels}}
 \end{figure}
@@
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-256.png}
-\caption{Time to run a single scan in a SDM with $n=256$, $H=1,000,000$ and $r=103$.
-\label{fig:perf-imac-scan-256}}
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/imac-scan-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/imac-scan-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/imac-scan-10k.png}}
+
+\caption{Scanner comparisons for iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+\label{fig:perf-imac-scanners}}
 \end{figure}
@@
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-1000.png}
-\caption{Time to run a single scan in a SDM with $n=1,000$, $H=1,000,000$ and $r=451$.
-\label{fig:perf-imac-scan-1000}}
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/imac-read-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/imac-read-1000.png}}
+
+\caption{Read operation comparisons for iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+\label{fig:perf-imac-read}}
 \end{figure}
@@
 \begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{images02/performance/imac-scans-10k.png}
-\caption{Time to run a single scan in a SDM with $n=10,000$, $H=1,000,000$ and $r=4845$.
-\label{fig:perf-imac-scan-10k}}
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/imac-write-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=\textwidth]{images02/performance/imac-write-1000.png}}
+
+\caption{Write operation comparisons for iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+\label{fig:perf-imac-write}}
 \end{figure}
 
-\begin{table}[!htb]
+% EC2 p2.xlarge
+
+\begin{figure}[!htb]
 \centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 81.62 & 12.25 \\
-    Thread scan & 5000 & 143.68 & 6.95 \\
-    OpenCL scan & 5000 & 238.00 & 4.20 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 1000 & 74.92 & 13.34 \\
-    Thread single read & 1000 & 96.22 & 10.39 \\
-    OpenCL write & 1000 & 126.50 & 7.90 \\
-    OpenCL single read & 1000 & 190.20 & 5.25 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=1,000$ bits, $H=1,000,000$, and $r=451$.
-\label{tab:perf-imac-1000}}
-\end{table}
+\subfloat[$n=256$, $r=103$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-kernels-256.png}}
 
-\begin{table}[!htb]
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-kernels-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-kernels-10k.png}}
+
+\caption{Kernel comparisons for Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
+\label{fig:perf-ec2-p2-kernels}}
+\end{figure}
+
+\begin{figure}[!htb]
 \centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 1000 & 175.48 & 5.69 \\
-    Thread scan & 5000 & 352.63 & 2.83 \\
-    OpenCL scan & 5000 & 244.88 & 4.08 \\ \hline
-    \hline
-    & Loops & Ops / second & Op. time (ms) \\ \hline
-    Thread write & 2000 & 304.46 & 3.28 \\
-    Thread single read & 2000 & 391.21 & 2.55 \\
-    OpenCL write & 2000 & 378.44 & 2.64 \\
-    OpenCL single read & 2000 & 466.16 & 2.14 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU. Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-\label{tab:perf-imac-256}}
-\end{table}
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-scan-256.png}}
 
-\begin{table}[!htb]
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-scan-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-scan-10k.png}}
+
+\caption{Scanner comparisons for Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
+\label{fig:perf-ec2-p2-scanners}}
+\end{figure}
+
+\begin{figure}[!htb]
 \centering
-\begin{tabular}{| l | r | r | r |}
-    \hline
-    & Loops & Scans / second & Scan time (ms) \\ \hline
-    Linear scan & 100 & 8.59 & 116.38 \\
-    Thread scan & 500 & 18.66 & 53.56 \\
-    OpenCL scan & 1000 & 77.20 & 12.95 \\
-    \hline
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
-Running an SDM with $n=10,000$ bits, $H=1,000,000$, and $r=4845$.
-There is no benchmark for read and write operations because RAM is not enough to allocate the counters --- it would consume 37.25 GB of RAM.
-\label{tab:perf-imac-10k}}
-\end{table}
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-read-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-read-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-read-10k.png}}
+
+\caption{Read operation comparisons for Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
+\label{fig:perf-ec2-p2-read}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-write-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-write-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p2-write-10k.png}}
+
+\caption{Write operation comparisons for Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
+\label{fig:perf-ec2-p2-write}}
+\end{figure}
+
+% EC2 p3.2xlarge
+
+\begin{figure}[!htb]
+\centering
+\subfloat[$n=256$, $r=103$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-kernels-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-kernels-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-kernels-10k.png}}
+
+\caption{Kernel comparisons for Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+\label{fig:perf-ec2-p3-kernels}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-scan-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-scan-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-scan-10k.png}}
+
+\caption{Scanner comparisons for Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+\label{fig:perf-ec2-p3-scanners}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-read-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-read-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-read-10k.png}}
+
+\caption{Read operation comparisons for Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+\label{fig:perf-ec2-p3-read}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-write-256.png}}
+
+\subfloat[$n=1,000$, $r=451$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-write-1000.png}}
+
+\subfloat[$n=10,000$, $r=4805$, and $H=1,000,000$]{\includegraphics[width=0.7\textwidth]{images02/performance/ec2-p3-write-10k.png}}
+
+\caption{Write operation comparisons for Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+\label{fig:perf-ec2-p3-write}}
+\end{figure}

### 2018-03-11 Chapters/Chapter02.tex
New chapter: Critical distance

@@
 
-\chapter{Results (i): Framework Validation}
+\chapter{Results (i): Framework validation}
 \input{./Chapters/sdm-validation.tex}
 
-\chapter{Results (ii): Loss of neurons}
+\chapter{Results (ii): Critical distance}
+\input{./Chapters/sdm-critical-distance.tex}
+
+\chapter{Results (iii): Loss of neurons}
 In SDM, the data is written distributed among millions of hard-locations, which theoretically gives SDM robustness against loss of neurons.
@@
 
-\chapter{Results (iii): Generalized read operation}
+\chapter{Results (iv): Generalized read operation}
 
@@
 
-\chapter{Results (iv): Performance}
+\chapter{Results (v): Performance}
 \input{./Chapters/sdm-performance.tex}
@@
 
-\chapter{Results (vi): Information-theoretical write operation}
+\chapter{Results (vii): Information-theoretical write operation}
 

### 2018-03-11 Chapters/sdm-critical-distance.tex
New chapter: Critical distance

--- 2018-03-10 Chapters/sdm-validation.tex
@@
 % !TEX root = ../partial-sdm.tex
-
-The framework has been validated comparing its results with the expected results from \citet{Kanerva1988}.
-Thus, we run simulations which were then compared to the theoretical analysis conducted some decades ago.
-
-\section{Distance between random bitstrings}
-
-As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
-For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
-
-In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits.
-In total, 20,000 random bitstrings have been generated during the simulation.
-The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.
-
-In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
-
-This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.
-
-\begin{figure}[!htb]
-  \centering
-  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
-  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}
-
-  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
-  \label{fig:validation-distance}
-\end{figure}
-
-\section{Number of activated hard-locations}
-
-In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard-locations in a 1,000 bits SDM.
-He also proposed to activate only 1,000 of them, on average.
-He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard-locations.
-
-We extended his results, calculating the distribution of the number of activated hard-locations.
-As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$.
-In this case, $\mu = 1071.85$ and $\sigma = 32.72$.
-
-In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard-locations.
-The code is available in the ``Number of activated hard-locations'' notebook \citep{sdmframework}.
-
-In figure \ref{fig:validation-activated-hls}, we can notice that the theoretical model and the simulation matches.
-Hence, it seems that both the address space generation algotihm and the scan algorithm work properly.
-Notice that the curve is almost the same for $n=1,000$ and $n=256$.
-It happens because the access radius is adjust to have $p$ as close as possible to $0.001$.
-They are not exactly the same because their $p$ differs a little.
-
-\begin{figure}[!htb]
-  \centering
-  \subfloat[$n=1,000$, $H=1,000,000$,\protect\\ $r=451$, and $p=0.00107185$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-1000.png}}
-  \subfloat[$n=256$, $H=1,000,000$,\protect\\ $r=103$, and $p=0.00106684$ ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/activated-hls-256.png}}
-
-  \caption{Histogram of the number of activated hard-locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
-  \label{fig:validation-activated-hls}
-\end{figure}
-
-Besides the number of activated hard-locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard-locations.
-Let $A$ be the set of activated hard-locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:
-
-\begin{align}
-P(\text{d}(a, \xi)=x | a \in A) &= \frac{P(\text{d}(a, \xi)=x)}{P(a \in A)} \\
-    &= \frac{\binom{n}{x}}{\sum_{k=0}^r \binom{n}{k}} \label{eq:prob-d-inside-circle}
-\end{align}
-
-In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard-locations to the center of 1,000 random circles.
-The code is available in the ``Distances of activated hard-locations'' notebook \citep{sdmframework}.
-
-In figure \ref{fig:validation-distance-activated-hls}, we can notice that the theoretical model and the simulation matches.
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{./images02/new-images/distance-activated-hls.png}
-
-\caption{Histogram of the distances of activated hard-locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
-\label{fig:validation-distance-activated-hls}}
-\end{figure}
-
-\section{Intersection of two circles}
-
-Kanerva has calculated the intersection of two circles according to the distance between their centers.
-The intersection is important to understand how SDM works, because it affects directly the critical distance.
-When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
-In fact, it converges because they share a sufficient amount of hard-locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge.
-For further information about the relation between the critical distance and the intersection, see \citet{brogliato2014sparse}.
-
-We have calculated the intersection between a random bitstring (bs1) and another bitstrings (bs2) exactly $d$ bits away.
-The former (bs1) is just a random bitstring.
-The latter (bs2) was generated randomly flipping $d$ bits of bs1.
-The code is available in the ``Kanerva's Figure 1.2'' notebook \citep{sdmframework}.
-
-In Figure \ref{fig:validation-intersection}, we can notice that we have obtained the same results as Kanerva. It seems that the random flipping bits algorithm and the scan algorithm work properly.
-
-\begin{figure}[!htb]
-  \centering
-  \subfloat[{\citet[Figure 1.2, p.25]{Kanerva1988}} ]{\includegraphics[width=0.45\textwidth]{./images02/new-images/kanerva-table-12.jpg}}
-  \subfloat[Generated by SDM-Frameworkm with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}
-
-  \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
-  \label{fig:validation-intersection}
-\end{figure}
-
-\section{Storage and retrieval of sequences}
-
-\citet[Ch.8]{Kanerva1988} presented an approach to store and retrieve sequences using $k$ different SDMs, namely $\text{sdm}_1$, $\text{sdm}_2$, $dots$, $\text{sdm}_k$.
-
-Let $a_0, a_1, a_2, \dots, a_n$ be a sequence to be stored in a $k$-fold memory.
-So, all pointers of the form $a_i \rightarrow a_{i+k}$ will be written to $\text{sdm}_k$ memory, i.e., in $\text{sdm}_1$, the following pointers will be written: $a_0 \rightarrow a_1$, $a_1 \rightarrow a_2$, $\dots$, $a_{n-1} \rightarrow a_n$; while in $\text{sdm}_2$, the following pointers will be written: $a_0 \rightarrow a_2$, $a_2 \rightarrow a_3$, $\dots$, $a_{n-2} \rightarrow a_n$; and so forth.
-
-We have tested exactly the same example presented in \citet{Kanerva1988}, p.85.
-We wrote two sequences to a $3$-fold memory: $<A, B, C, D>$ and $<E, B, C, F>$.
-Then, after reading the sequences $<A, B, C>$ and $<E, B, C>$, we have obtained $D$ and $F$, respectively.
-
-Each reading operation was performed summing the counters of all activated hard-locations from all three memories.
-For instance, to read the sequence $<A, B, C>$, we have activated the hard-locations around $C$ in $\text{sdm}_1$, we have also activated the hard-locations around $D$ in $\text{sdm}_2$, and, finally, we have also activated the hard-locations around $A$ in $\text{sdm}_3$.
-After summing the counters of all those hard-locations, we evaluate the resulting bitstring just as in the original read operation.
-
-The code is available in the ``Sequences (Kanerva Ch 8)'' notebook \citep{sdmframework}.
-
-The logic behind how it works is that, when reading the sequence $<A, B, C>$, we have $A$ pointing to $D$, while both $B$ and $C$ point to $D$ and $F$.
-Thus, $D$ appears more often than $F$ and ended up being the result.
-
-Hence, as we have replicated the theoretical results from Kanerva, we have one more evidence that our framework works properly.
-
-\subsection{$k$-fold memory using only one SDM}
-
-We have extended Kanerva's ideas to be able to store and retrieve sequences in $k$-fold memories using only one SDM (instead of $k$ SDMs).
-
-Our idea was to create $k$ random bitstrings, one for each fold.
-We have performed writing and reading exactly as Kanerva's original idea, but, instead of writing to $\text{sdm}_k$, we have written $a_{i+k}$ into the address $a_i \oplus \text{tag}_k$, and, instead of reading from $\text{sdm}_k$, we have read from address $a_i \oplus \text{tag}_k$, where $\oplus$ is the exclusive or (XOR) operator.
-
-It worked as if we had splitted SDM into $k$ regions with low intersection between two of them.
-So, as the interference is minimal, they work like independent SDMs.
-The major disadvantage of this approach is that memory capacity may be reached faster.
-
-Splitting the memory into regions may be an interesting strategy to other sorts of problems, mostly the ones which would need many SDMs and, consequently, would use a lot of RAM.
-
-\section{Convergence of $\eta_x$ to $\eta$}
 
@@
 As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
-%and we propose that this is due to interaction effects between different attractors.
-
 We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
@@
 So, we have changed the steps to run $k$ reads and store the average new distance.
-We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6samples}, which has a way lower noise and still holds the divergence.
+We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6samples}, which has much lower noise and still holds the divergence.
 
@@
 Thereby, in order to show the interaction effects more clearly, we changed the single read for an 15-iterative read.
-As we can see in Figure \ref{sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
+As we can see in Figure \ref{fig:sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
 
@@
 As we change the single read to a 6-iterative read, the effect has vanished and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings.
-Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, except for the iterative read.
-\label{sdm-10000w-table-7-2-15iter}}
+Here we have the exact same configuration of Figure \ref{fig:sdm-10000w-table-7-2}, except for the iterative read.
+\label{fig:sdm-10000w-table-7-2-15iter}}
+\end{figure}
+
+Going further in the analysis, we calculated the probability of missing a bit when reading from SDM.
+After all, that's how Kanerva has originally found the curve.
+To do this, we used the following equations from our previous work \citep{brogliato2014sparse}.
+Let $d$ be the distance to the target, $h$ be the number of hard-locations activated during read and write operations, $s$ be the number of total stored bitstrings, $H$ be the number of total hard-locations, $w$ be the number of times the target was written into SDM, $\theta$ be the total random bitstrings in all $h$ hard-locations activated by read operation, and $\phi(d)$ be the average number of shared hard-locations activated two bitstrings $d$ bits away.
+
+\begin{align}
+\theta &= \frac{sh^2}{H} - w \cdot \phi(d) \\
+P(miss | bit=0) &= 1 - P \left( \sum_{i=1}^\theta X_i < \frac{sh^2}{2H} \right) \\
+P(miss | bit=1) &= P \left( \sum_{i=1}^\theta X_i < \frac{sh^2}{2H} - w \cdot \phi(d) \right) \\
+P(miss) &= \frac{1}{2} \cdot \left[ P(miss | bit=0) + P(miss | bit=1) \right]
+\end{align}
+
+For details and the proof of this equation, see \citet{brogliato2014sparse}.
+Although Kanerva has found a formula for $\phi(d)$ through an unsolved integral, and \citet{de1995geometrical} has proposed another way to calculate $\phi(d)$, we have used our framework to estimate the values of $d$.
+In order to do that, we used a Monte Carlo approach, generating many pair of random bitstrings $d$ bits away from them and calculating the average number of shared hard-locations between them.
+The code is available in the ``Calculate critical distance'' notebook \citep{sdmframework}.
+
+Kanerva's settings according to the parameters of the equation were: $s=10,000$, $H=1,000,000$, and $w=1$.
+We have calculated $\phi(d)$ as explained, and $h = H \cdot 2^{-n} \sum_{i=0}^{r} \binom{n}{i}$, where $n=1,000$ and $r=451$., $\phi(d)$.
+Finally, $h=1,071.85$ and changing $d$ from $0$ to $1000$, we got Figure \ref{fig:kanerva-figure-73-calculated}.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=0.8\textwidth]{./images02/calculated-table-72.png}
+\caption{Kanerva's original Figure 7.3 generated using the equations from \citet{brogliato2014sparse}.
+\label{fig:kanerva-figure-73-calculated}
+}
+\end{figure}
+
+As we can easily notice, we got exactly the same curve as Kanerva. This question has been intriguing us since then and we are still looking for a better explanation then ``attractors''.
+
+\section{Critical distance 209}
+
+The critical distance is defined as $d$ where $P(miss) = d/n$, or, in Figure \ref{fig:kanerva-figure-73-calculated}, the point where the curve meets with the identity function (the black diagonal line).
+Thus, we plot a zoom-in of Figure \ref{fig:kanerva-figure-73-calculated} around $d=209$ in Figure \ref{fig:figure-73-eq-zoom-in} using the same equations \citep{brogliato2014sparse}.
+It was surprising that the meeting does not happen at $d=209$, but around $d=221$.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=0.8\textwidth]{./images02/figure-73-eq-zoom.png}
+\caption{Zoom-in around $d=209$ of Figure \ref{fig:kanerva-figure-73-calculated}.
+\label{fig:figure-73-eq-zoom-in}
+}
+\end{figure}
+
+To confirm that the critical distance is not around 209, but around 221, we also plot a zoom-in of Figure \ref{fig:sdm-10000w-table-7-2} around $d=209$ in Figure \ref{fig:sdm-10000w-zoom}.
+In order to reduce the noise, we increased the samples to $k=180$.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=0.8\textwidth]{./images02/sdm-10000w-zoom-209.png}
+\caption{Zoom-in around $d=209$ of Figure \ref{fig:sdm-10000w-table-7-2}.
+\label{fig:sdm-10000w-zoom}
+}
 \end{figure}

### 2018-03-11 Chapters/sdm-validation.tex
New chapter: Critical distance

@@
 Splitting the memory into regions may be an interesting strategy to other sorts of problems, mostly the ones which would need many SDMs and, consequently, would use a lot of RAM.
-
-\section{Convergence of $\eta_x$ to $\eta$}
-
-One particular analysis of Kanerva's interest is that of the distance read at a point $\eta_x$.
-Suppose an SDM is trying to read an item written at $\eta$, but the cues received so far lead to a point of distance $x$ from $\eta$.
-As one reads at $\eta_x$, a new bitstring $\beta$ is obtained, leading to Kanerva's question: what is the new distance from $\eta$ to $\beta$?
-Is it smaller or larger than $x$?
-That, of course, depends on the ratio between $x$ and the number of dimensions of the memory.
-
-\citet[p.70]{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Figure \ref{fig:kanerva-figure-7.3}).
-The original prediction considered that the read distance would decline when inside the critical distance and increase afterwards, converging to a \textasciitilde 500-bit distance.
-At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance bitstring.
-He analyzed specifically an SDM with 1,000 bits and 10,000 random bitstrings written into it.
-
-\begin{figure}[h]
-\centering\includegraphics[width=0.8\textwidth]{images02/kanerva-table-7-2-original.png}
-\caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
-\label{fig:kanerva-figure-7.3}}
-\end{figure}
-
-As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
-%and we propose that this is due to interaction effects between different attractors.
-
-We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
-
-In details, we have created a SDM with $n=1,000$, $H=1,000,000$, and $r=451$.
-Then, we have generated 10,000 random bitstrings and written them into the memory.
-Then, we have generated a reference bitstring (bs\_ref) and written it into the memory.
-Then, we have executed the following steps with $x$ from 0 to 1,000: (i) copy bs\_ref into a new bitstring; (ii) randomly flipped $x$ bits of the copy; (iii) read from the memory in the copy address; and (iv) stored the distance between the returned bitstring and bs\_ref.
-Finally, we have plotted Figure \ref{fig:sdm-10000w-table-7-2}.
-
-\begin{figure}[h]
-\centering
-\subfloat[1 sample for each distance $x$ \label{fig:sdm-10000w-table-7-2-1sample} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2.png}}
-\subfloat[6 samples for each distance $x$ \label{fig:sdm-10000w-table-7-2-6samples} ]{\includegraphics[width=0.5\textwidth]{./images02/sdm-10000w-table-7-2-6-samples.png}}
-
-\caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
-Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
-\label{fig:sdm-10000w-table-7-2}}
-\end{figure}
-
-Figure \ref{fig:sdm-10000w-table-7-2-1sample} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance.
-So, we have changed the steps to run $k$ reads and store the average new distance.
-We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6samples}, which has a way lower noise and still holds the divergence.
-
-Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).
-
-Obviously, these small deviations from Kanerva's original theoretical predictions deserve a qualification.
-Kanerva was working in the 1980s and the 1990s, and had no access to the immense computational power that we do today.
-It is no surprise that some small interaction effects should exist as machines allow us to explore the ideas of his monumental work.
-
-But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Figure \ref{fig:sdm-100w-table-7-2-10samples}).
-This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
-In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractors.
-
-\begin{figure}[h]
-\centering
-\subfloat[100 writes \label{fig:sdm-100w-table-7-2-10samples} ]{\includegraphics[width=0.47\textwidth]{./images02/sdm-100w-table-7-2.png}}
-\subfloat[Steps of 1,000 writes \label{fig:sdm-table-7-2-steps} ]{\includegraphics[width=0.53\textwidth]{./images02/sdm-table-7-2-steps.png}}
-%\centering\includegraphics[width=\textwidth]{images02/sdm-100w-table-7-2.png}
-
-\caption{Results generated by the framework similar from Kanerva's original Table 7.2.
-Here we have a 1,000 bit, 1,000,000 hard-location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
-\label{fig:sdm-100w-table-7-2}}
-\end{figure}
-
-To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
-Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
-Thereby, in order to show the interaction effects more clearly, we changed the single read for an 15-iterative read.
-As we can see in Figure \ref{sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
-
-Hence, our understanding is that the attractors are just preventing the bitstrings to converge directly to 500-bit distance bitstrings, requiring more reading steps to do so.
-They are in other orthogonal bitstrings' critical distance, but sufficiently far not to converge in a single read.
-
-\begin{figure}[h]
-\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-15iter.png}
-\caption{This graph shows the interaction effects more clearly.
-As we change the single read to a 6-iterative read, the effect has vanished and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings.
-Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, except for the iterative read.
-\label{sdm-10000w-table-7-2-15iter}}
-\end{figure}
-
-%To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin.
-Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
-Thereby, in order to show the interaction effects more clearly, we wrote a handmade bitstring to the SDM which had all bits inverted in relation to the origin bitstring --- their hamming distance was equal to 1,000.
-Our handmade bitstring was acting as an opposite attractor, and one can see the accelerating effects towards convergence to both attractors: the origin and the handmade bitstrings (Fig.
-\ref{sdm-10000w-notX-table-7-2}).
-Here we had the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
-
-%\begin{figure}[h]
-%\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-notX-table-7-2.png}
-%\caption{This graph shows the interaction effects more clearly.
-As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and the opposite.
-Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.
-%\label{sdm-10000w-notX-table-7-2}}
-%\end{figure}

### 2018-03-11 Chapters/Conclusion.tex
petrossaurus
1st after major pull

@@
 \noindent Modern management and high technology interact in multiple, profound, ways.
-Software, in particular, seems to have an immense power of entering arenas which seemed, at some point, to require either specific hardware or the skill of humans.
+Software --- given the widespread availability of general-purpose Turing-complete hardware --- seems to have an immense power of entering arenas which seemed, at some point, to require either specific hardware or the skill of humans.
 One of the members of this thesis committee, Dr.
@@
 Imagine a new method that increases profits by 50\% at a tech company.
-Should this method, if implementable as an algorithm (like PageRank \citet{brin1998anatomy}) or a data structure (like a blockchain) be discussed in conferences of `computer science' or `business'?
+Should this method, if implementable as an algorithm (like PageRank \citep{brin1998anatomy}) or a data structure (like a blockchain) be discussed in conferences of `computer science' or `business'?
 It seems quite arbitrary to name a single group, as a whole new ecosystem seems to be emerging within those two.
 That is why this thesis is computational and why it is business.
-This work explores topics that seem, on the surface, to belong to computer science, but their applicability and impact to businesses seem too large to be delegated away, something ``for the nerds in the fifth floor''.
+This work explores topics that seem, on the surface, to belong to computer science, but their applicability and impact to businesses seem too large, too central, to be delegated away, something ``for those nerds in the fifth floor''.
 As technical decisions become central to the organization of man's life, the technician becomes the visionary, the innovator, the decision-making arbiter, sometimes the billionaire.
@@
 As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobrás by ten billion dollars, while Bitcoin's is valued at more than double of Petrobrás (195B usd vs 83B usd).
-These technologies should, at a minimum, be taken seriously.
+Prices change, of course; but these technologies should, at a minimum, be taken seriously.
 

### 2018-03-11 Chapters/sdm-performance.tex
SDM Performance finished
Performance:  reviewed here.

@@
 If it takes a few days, it is more important to make a good plan.
-As SDM consumes a lot of both processor and memory resources, some experiments may take a long time.
+As SDM consumes large processor and memory resources, some experiments may take a long time.
 
-Each scan a 1,000-bits SDM with 1,000,000 hard-locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
+Each scan on a 1,000-bits SDM with 1,000,000 hard-locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
 So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
@@
 A modern laptop can increment a counter $5 \cdot 10^{8}$ times per second.
-Hence, a naive implementation of SDM may take several hours to simply write 10,000 random bitstrings.
+Hence, a naive implementation of SDM may take several hours --- or days --- to simply write 10,000 random bitstrings.
 
@@
 It has compared $3.37 \cdot 10^{12}$ bits per second through $52.6 \cdot 10^{9} = 52.6 \text{ billion}$ XORs and popcounts per second.
-It is a great improvement when compared to the first versions of the code which took 16 hours to generate the same heatmap (and its memory use was already optimized and the computations were distributed in threads).
+It is a 60-fold improvement over the first versions of the code which took 16 hours to generate the same heatmap (and its memory use was already optimized and the computations were distributed in threads).
 
 We have created a benchmark to be able to compare the performance in different devices.
-So, the same performance test was executable in our devices and the results were reported in tables and figures.
+So, the same performance test was executable in our devices, with results reported in tables and figures.
 The benchmark has 3 parts: (i) the first part consists of comparing the available OpenCL kernels to find which works best for that device; than (ii) the second part consists of comparing the linear scanner, the thread scanner, and the OpenCL scanners with the best kernel found in part one; finally, (iii) the third part consists of comparing read and write operations using the thread scanner and the OpenCL scanner with the best kernel.
@@
 The whole source code is available in the ``Performance test'' notebook \citep{sdmframework}.
-We would like to invite you to run this benchmark in your device and send us the results.
+We would like to invite the reader to run this benchmark and, if possible, share the results.
 
 Our first device was a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-It was not possible to run the 10,000-bits in this device because there was no memory available --- it would use 37.25 GB of memory.
+It was not possible to run the 10,000-bits in this device because there was no memory available --- it needs 37.25 GB of memory.
 For its results, see Table \ref{tab:perf-macbook}
@@
 
-OpenCL is a framework for writting software that executes in \emph{heterogeneous} devices \citep{munshi2009opencl}, like CPUs, GPUs, and other co-processor for hardware acceleration.
-Because they are heterogogeneous, they may differ a lot in architecture and performance, which means there is no one-size-fits-all kernel.
-The slower kernel for one device may be the fastest for another device, and it really happened in our case.
+OpenCL is a framework for writting software that executes in \emph{heterogeneous} devices \citep{munshi2009opencl}, like CPUs, GPUs, FPGA and other co-processor for hardware acceleration.
+Because they are heterogeneous, they may differ a lot in architecture and performance, which means there is no one-size-fits-all kernel.
+A \emph{kernel} is generally a small function on the code that runs in thousands of parallel threads, executing the same steps on different parts of a large vector or matrix.
+The slower kernel for one device may be the fastest for another device, as we will see happening in our case.
 
-Our framework has 8 kernels available: single\_scan0, single\_scan1, single\_scan2, single\_scan3, single\_scan4, single\_scan5, single\_scan5\_unroll, and single\_scan6.
+A total of 8 kernels have been developed for benchmarking in our framework: single\_scan0, single\_scan1, single\_scan2, single\_scan3, single\_scan4, single\_scan5, single\_scan5\_unroll, and single\_scan6.
 Each scan uses a different algorithm to do exactly the same thing: calculate which hard-locations are inside the circle of the given bitstring.
@@
 
-We recommend users to run a specific kernel comparison test for their own GPU and SDM settings.
+We recommend users to run a specific kernel comparison test for their own GPU and SDM settings.  This is available in the jupyter notebooks.
 
@@
 As Amazon EC2 are virtual machines with GPU devices, their CPU is shared with other virtual machines which reduces CPU power significantly.
-Hence, for both virtual machines we have tested, using the GPU really boost the performance, but using the CPU should be avoided.
+Hence, for both virtual machines we have tested, using the GPU seriously boosts performance, but using the CPU should be avoided.
 See Tables \ref{tab:perf-imac}
@@
 
-\begin{table}[!htb]
-\centering
-\begin{tabular}{lrrrr}
-    \toprule
-    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
-    \hline
-	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    single\_scan0 & 3.03 & 5.00 & 61.06 \\
-    single\_scan1 & 2.87 & 3.95 & 44.96 \\
-    single\_scan2 & 3.82 & 4.57 & 44.98 \\
-    single\_scan3 & 3.72 & 3.68 & 12.67 \\
-    single\_scan4 & 4.48 & 4.04 & 11.45 \\
-    single\_scan5 & 3.76 & 3.72 & 12.58 \\
-	single\_scan5\_unroll & 3.79 & 3.61 & 11.37 \\
-	single\_scan6 & 4.36 & 4.02 & 10.96 \\ \hline
-    \hline
-	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    Linear scan & 5.04 & 12.25 & 116.38 \\
-    Thread scan & 2.92 &  6.95 &  53.56 \\
-    OpenCL scan & 2.81 &  4.20 &  12.95 \\ \hline
-    \hline
-	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    Thread write       & 3.28 & 13.34 \\
-    Thread single read & 2.55 & 10.39 \\
-    OpenCL write       & 2.64 &  7.90 \\
-    OpenCL single read & 2.14 &  5.25 \\
-    \bottomrule
-\end{tabular}
-\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
-The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
-There is no benchmark for read and write operations with $n=10,000$ because RAM is not enough to allocate the counters---it would consume 37.25 GB of RAM.
-For the histogram of durations, see Figures \ref{fig:perf-imac-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.
-\label{tab:perf-imac}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{lrrr}
-    \toprule
-    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
-	\hline
-	Kernels & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    single\_scan0 &  0.76 &  3.79 & 35.45 \\
-    single\_scan1 &  0.80 &  3.94 & 57.80 \\
-    single\_scan2 &  5.59 &  8.54 & 63.71 \\
-    single\_scan3 &  6.31 &  9.73 & 39.92 \\
-    single\_scan4 & 10.29 & 11.38 & 45.49 \\
-    single\_scan5 &  6.69 &  9.95 & 43.51 \\
-	single\_scan5\_unroll & & & \\
-    single\_scan6 & 10.29 & 11.33 & 41.42 \\
-    \hline
-	Scanners & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    Linear scan & 19.09 & 64.73 & 600.75 \\
-    Thread scan &  9.95 & 32.81 & 296.56 \\
-    OpenCL scan &  6.88 & 10.67 &  46.73 \\ \hline
-    \hline
-	Operations & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    Thread write       & 11.80 & 42.64 & 383.50 \\
-    Thread single read & 10.49 & 35.37 & 307.97 \\
-    OpenCL write       &  8.84 & 19.31 & 122.17 \\
-    OpenCL single read &  7.50 & 11.72 &  55.47 \\
-    \bottomrule
-\end{tabular}
-\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
-Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
-The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
-There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
-The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
-For the histogram of durations, see Figures \ref{fig:perf-ec2-p2-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.
-\label{tab:perf-ec2-p2}}
-\end{table}
-
-\begin{table}[!htb]
-\centering
-\begin{tabular}{lrrr}
-    \toprule
-    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
-	\hline
-	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    \hline
-    single\_scan0 & 0.36 & 0.69 & 20.60 \\
-    single\_scan1 & 0.36 & 0.54 &  4.94 \\
-    single\_scan2 & 0.73 & 0.85 &  5.01 \\
-    single\_scan3 & 0.63 & 1.02 &  6.07 \\
-    single\_scan4 & 1.05 & 1.10 &  5.99 \\
-    single\_scan5 & 0.62 & 0.95 &  5.36 \\
-	single\_scan5\_unroll & & & \\
-    single\_scan6 & 1.01 & 1.04 &  5.96 \\ \hline
-    \hline
-	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    Linear scan & 17.60 & 58.34 & 540.97 \\
-    Thread scan &  5.19 & 16.39 & 198.74 \\
-    OpenCL scan &  0.63 &  0.85 &   5.74 \\ \hline
-    \hline
-	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
-    Thread write       & 7.59 & 28.47 & 222.08 \\
-    Thread single read & 6.17 & 17.44 & 145.01 \\
-    OpenCL write       & 2.33 &  8.77 &  80.48 \\
-    OpenCL single read & 1.01 &  1.82 &  13.99 \\
-    \hline
-\end{tabular}
-\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
-The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
-There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
-The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
-For the histogram of durations, see Figures \ref{fig:perf-ec2-p3-kernels}, \ref{fig:perf-ec2-p3-scanners}, \ref{fig:perf-ec2-p3-read}, and \ref{fig:perf-ec2-p3-write}.
-\label{tab:perf-ec2-p3}}
-\end{table}
-
-% MacBook Pro
-
 \begin{figure}[!htb]
@@
 
-% iMac
+% -------------------------
+% -------------------------
+% -------------------------
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{lrrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
+    \hline
+	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    single\_scan0 & 3.03 & 5.00 & 61.06 \\
+    single\_scan1 & 2.87 & 3.95 & 44.96 \\
+    single\_scan2 & 3.82 & 4.57 & 44.98 \\
+    single\_scan3 & 3.72 & 3.68 & 12.67 \\
+    single\_scan4 & 4.48 & 4.04 & 11.45 \\
+    single\_scan5 & 3.76 & 3.72 & 12.58 \\
+	single\_scan5\_unroll & 3.79 & 3.61 & 11.37 \\
+	single\_scan6 & 4.36 & 4.02 & 10.96 \\ \hline
+    \hline
+	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 5.04 & 12.25 & 116.38 \\
+    Thread scan & 2.92 &  6.95 &  53.56 \\
+    OpenCL scan & 2.81 &  4.20 &  12.95 \\ \hline
+    \hline
+	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Thread write       & 3.28 & 13.34 \\
+    Thread single read & 2.55 & 10.39 \\
+    OpenCL write       & 2.64 &  7.90 \\
+    OpenCL single read & 2.14 &  5.25 \\
+    \bottomrule
+\end{tabular}
+\caption{iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G GPU.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for read and write operations with $n=10,000$ because RAM is not enough to allocate the counters---it would consume 37.25 GB of RAM.
+For the histogram of durations, see Figures \ref{fig:perf-imac-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.
+\label{tab:perf-imac}}
+\end{table}
 
@@
 
-% EC2 p2.xlarge
+% -------------------------
+% -------------------------
+% -------------------------
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{lrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
+	\hline
+	Kernels & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    single\_scan0 &  0.76 &  3.79 & 35.45 \\
+    single\_scan1 &  0.80 &  3.94 & 57.80 \\
+    single\_scan2 &  5.59 &  8.54 & 63.71 \\
+    single\_scan3 &  6.31 &  9.73 & 39.92 \\
+    single\_scan4 & 10.29 & 11.38 & 45.49 \\
+    single\_scan5 &  6.69 &  9.95 & 43.51 \\
+	single\_scan5\_unroll & & & \\
+    single\_scan6 & 10.29 & 11.33 & 41.42 \\
+    \hline
+	Scanners & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 19.09 & 64.73 & 600.75 \\
+    Thread scan &  9.95 & 32.81 & 296.56 \\
+    OpenCL scan &  6.88 & 10.67 &  46.73 \\ \hline
+    \hline
+	Operations & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Thread write       & 11.80 & 42.64 & 383.50 \\
+    Thread single read & 10.49 & 35.37 & 307.97 \\
+    OpenCL write       &  8.84 & 19.31 & 122.17 \\
+    OpenCL single read &  7.50 & 11.72 &  55.47 \\
+    \bottomrule
+\end{tabular}
+\caption{Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU.
+Running an SDM with $n=256$ bits, $H=1,000,000$, and $r=103$.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
+The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
+For the histogram of durations, see Figures \ref{fig:perf-ec2-p2-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.
+\label{tab:perf-ec2-p2}}
+\end{table}
 
@@
 
-% EC2 p3.2xlarge
+% -------------------------
+% -------------------------
+% -------------------------
+
+\begin{table}[!htb]
+\centering
+\begin{tabular}{lrrr}
+    \toprule
+    & \textbf{256 bits} & \textbf{1,000 bits} & \textbf{10,000 bits} \\ \hline
+	\hline
+	Kernel & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    \hline
+    single\_scan0 & 0.36 & 0.69 & 20.60 \\
+    single\_scan1 & 0.36 & 0.54 &  4.94 \\
+    single\_scan2 & 0.73 & 0.85 &  5.01 \\
+    single\_scan3 & 0.63 & 1.02 &  6.07 \\
+    single\_scan4 & 1.05 & 1.10 &  5.99 \\
+    single\_scan5 & 0.62 & 0.95 &  5.36 \\
+	single\_scan5\_unroll & & & \\
+    single\_scan6 & 1.01 & 1.04 &  5.96 \\ \hline
+    \hline
+	Scanner & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Linear scan & 17.60 & 58.34 & 540.97 \\
+    Thread scan &  5.19 & 16.39 & 198.74 \\
+    OpenCL scan &  0.63 &  0.85 &   5.74 \\ \hline
+    \hline
+	Operation & Duration (ms) & Duration (ms) & Duration (ms) \\ \hline
+    Thread write       & 7.59 & 28.47 & 222.08 \\
+    Thread single read & 6.17 & 17.44 & 145.01 \\
+    OpenCL write       & 2.33 &  8.77 &  80.48 \\
+    OpenCL single read & 1.01 &  1.82 &  13.99 \\
+    \hline
+\end{tabular}
+\caption{Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU.
+The SDM settings were: (i) $n=256$, $r=103$, and $H=1,000,000$; (ii) $n=1,000$, $r=451$, and $H=1,000,000$; and (iii) $n=10,000$, $r=4850$, and $H=1,000,000$.
+There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
+The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
+For the histogram of durations, see Figures \ref{fig:perf-ec2-p3-kernels}, \ref{fig:perf-ec2-p3-scanners}, \ref{fig:perf-ec2-p3-read}, and \ref{fig:perf-ec2-p3-write}.
+\label{tab:perf-ec2-p3}}
+\end{table}
 
@@
 \end{figure}
+
+\subsection{Summary of results}
+
+The results, beyond showing the obvious fact that consumer grade hardware is not comparable to the Amazon instances, indicate a non-trivial issue:  The chosen kernel for scanning the memory is of crucial importance to performance, and this kernel speed is a function of both the hardware used and the particular parameters used in the SDM settings.

### 2018-03-11 Chapters/sdm-critical-distance.tex
Minor fix from previous commits
performance, mostly

@@
 
-One particular analysis of Kanerva's interest is that of the distance read at a point $\eta_x$.
+One particular analysis of Kanerva's interest is given by the limits of recovery.
+That is, given an item read at a distance $x$ from a previously stored $\eta$, does this reading at a $\eta_x$ recover the original?
 Suppose an SDM is trying to read an item written at $\eta$, but the cues received so far lead to a point of distance $x$ from $\eta$.
@@
 Kanerva's settings according to the parameters of the equation were: $s=10,000$, $H=1,000,000$, and $w=1$.
-We have calculated $\phi(d)$ as explained, and $h = H \cdot 2^{-n} \sum_{i=0}^{r} \binom{n}{i}$, where $n=1,000$ and $r=451$., $\phi(d)$.
+We have calculated $\phi(d)$ as explained, and $h = H \cdot 2^{-n} \sum_{i=0}^{r} \binom{n}{i}$, where $n=1,000$ and $r=451$.
 Finally, $h=1,071.85$ and changing $d$ from $0$ to $1000$, we got Figure \ref{fig:kanerva-figure-73-calculated}.

### 2018-03-11 Chapters/sdm-performance.tex
performance, mostly

@@
 If it takes a few hours, maybe we should think first.
-If it takes a few days, it is more important to make a good plan.
+If it takes a few days --- or more ---, it is important to devise a good plan.
 As SDM consumes large processor and memory resources, some experiments may take a long time.
@@
 
-Even though scan is an important part of the operations, we are really interested in the performance of the operations themselves.
+Even though scaning is an important part of the operations, we are really interested in the performance of the enterity of operations themselves.
 Comparing the times of the thread and OpenCL scanners with the times of their respective operations (either read or write), we can notice that their difference remains almost constant, which means the operation time itself is negligible when compared to the scan time.
@@
 The results, beyond showing the obvious fact that consumer grade hardware is not comparable to the Amazon instances, indicate a non-trivial issue:  The chosen kernel for scanning the memory is of crucial importance to performance, and this kernel speed is a function of both the hardware used and the particular parameters used in the SDM settings.
+
+It is reasonable to consider that the performance results obtained are of particular merit, and one particular fact stands out:  The scanning of 1,000,000 hard locations has become, in the desired professional-grade machines, \emph{faster} than the updating of the 1,000 active locations.

### 2018-03-12 Chapters/sdm-critical-distance.tex
New SDM image
crit-distance

@@
 For details and the proof of this equation, see \citet{brogliato2014sparse}.
-Although Kanerva has found a formula for $\phi(d)$ through an unsolved integral, and \citet{de1995geometrical} has proposed another way to calculate $\phi(d)$, we have used our framework to estimate the values of $d$.
+Although Kanerva has found a formula for $\phi(d)$ through an unsolved integral, and \citet{de1995geometrical} have proposed another way to calculate $\phi(d)$, we have used our framework to estimate the values of $d$.
 In order to do that, we used a Monte Carlo approach, generating many pair of random bitstrings $d$ bits away from them and calculating the average number of shared hard-locations between them.
@@
 
-As we can easily notice, we got exactly the same curve as Kanerva. This question has been intriguing us since then and we are still looking for a better explanation then ``attractors''.
+As we can easily notice, we got exactly the same curve as Kanerva.
+This question has been intriguing us since then and we are still looking for a more analytic explanation then merely interference from the other written attractors.
 
@@
 %\caption{This graph shows the interaction effects more clearly.
-As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and the opposite.
+As we include an opposite bitstring, one can see the accelerating effects towards convergence to both attractors: the origin and its polar opposite.
 Here we have the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.

### 2018-03-12 Chapters/sdm-architecture.tex
SDM Kernels source code

@@
 
-The OpenCL kernel explores the GPU architecture to improve performance.
-Each work group calculates the distance of several bitstrings.
+There are 8 OpenCL kernels which explore differently the GPU architecture to improve performance.
+It is necessary because there are several GPU microarchitecture and a single kernel will never be optimal for all of them.
+In simplified form, OpenCL splits the tasks into workgroups which, in turn, split their part of the task into workers.
+The works are like threads in a computer.
+OpenCL specifies four levels of memory hierarchy for the GPUs: global memory, read-only memory, local memory, and private memory.
+The global memory and read-only memory are accessible by all workgroups, while each workgroup has its own local memory, accessible by its workers.
+Finally, each worker has its own private memory.
+The number of workers per workgroup is defined by user and must be multiple of the number of tasks.
+
+All 8 kernels do the same thing: calculate the exclusive OR (XOR) between two 64-bit integers and count the number of bits one in the result.
+They just do it with different approaches.
+For instance, \lstinline{single_scan0} calculates one distance between bitstrings per worker (Listing \ref{lst:single_scan0}); while \lstinline{single_scan2} uses a whole workgroup to calculate each distance, distributing each element of the 64-int integer array per worker (Listing \ref{lst:single_scan2}).
+
+The OpenCL kernels \lstinline{single_scan3} (Listing \ref{lst:single_scan3}), \lstinline{single_scan4} (Listing \ref{lst:single_scan4}), \lstinline{single_scan5} (Listing \ref{lst:single_scan5}), \lstinline{single_scan5_unroll} (Listing \ref{lst:single_scan5_unroll}), \lstinline{single_scan6} (Listing \ref{lst:single_scan6}) explore the GPU architecture to improve the sum of the partial distances.
+Each workgroup calculates the distance of several bitstrings.
 During the distance calculation, each worker calculates the exclusive OR (XOR) between two 64-bit integers and use the built-in popcount function to count the number of ones.
-Then, they update an array of intermediate distances with their partial distances.
-This array is stored in the local memory and is shared with all workers of the same group.
+Then, they update an array of partial distances with their results.
+This array is stored in the local memory and is shared between all workers of the same workgroup.
 This whole step happens simultaneously in the GPU.
-Then, we used a reduction algorithm to sum the intermediate distances array in order to calculate the correct distance.
+Then, a reduction algorithm is used to sum the partial distances array in order to calculate the total distance.
 This reduction algorithm is also distributed between the workers and runs in $O(\log_2(\text{bs\_step}))$.
-Finally, one of the workers checks whether the distance is less than or equal to the radius to include the bitstring index into the resulting array.
+Finally, the first worker of each workgroup checks whether the distance is less than or equal to the radius to include the bitstring index into the resulting array.
 
-The number of workers in each group is the closest power of two above bs\_len (which is the number of 64-integers that forms a bitstring).
+Some of the optimizations may not work in some GPUs, because not all their premises are valid. Before choosing a kernel, one should check whether it works property for one's specific GPU device.
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+label={lst:single_scan0},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan0}}.
+It calculates one distance per worker and let the GPU decide how to distribute this task between workgroups and workers.
+It is the most simple kernel and does not explore any details of the GPU architecture.},
+]
+__kernel
+void single_scan0(
+		__constant const uchar *bitcount_table,
+		__global const ulong *bitstrings,
+		const uint bs_len,
+		const uint sample,
+		__constant const ulong *bs,
+		const uint radius,
+		__global uint *counter,
+		__global uint *selected,
+		__local uint *partial_dist)
+{
+	uint id = get_global_id(0);
+
+	if (id < sample) {
+		ulong a;
+		uint dist;
+
+		const __global ulong *row = bitstrings + id*bs_len;
+
+		dist = 0;
+		for(uint j=0; j<bs_len; j++) {
+			a = row[j] ^ bs[j];
+			dist += popcount(a);
+		}
+		if (dist <= radius) {
+			selected[atomic_inc(counter)] = id;
+		}
+	}
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+label={lst:single_scan1},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan1}}.
+It is just like \lstinline{single_scan0}, but it distribute several distances per workgroup, which, in turn, distribute the distances among their workers.},
+]
+__kernel
+void single_scan1(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint id;
+    ulong a;
+    uint dist;
+    const __global ulong *row;
+
+    for (id=get_global_id(0); id < sample; id += get_global_size(0)) {
+
+        row = bitstrings + id*bs_len;
+
+        dist = 0;
+        for(uint j=0; j<bs_len; j++) {
+            a = row[j] ^ bs[j];
+            dist += popcount(a);
+        }
+        if (dist <= radius) {
+            selected[atomic_inc(counter)] = id;
+        }
+
+    }
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+label={lst:single_scan2},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan2}}.
+It calculates one distance per workgroup, distributing each 64-bit integer operation per worker, and then summing the results obtained by the workers.
+The sum algorith is done by only the first worker of each workgroup.},
+]
+__kernel
+void single_scan2(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint dist;
+    ulong a;
+    uint j;
+
+    for (uint id = get_group_id(0); id < sample; id += get_num_groups(0)) {
+
+        const __global ulong *row = bitstrings + id*bs_len;
+
+        dist = 0;
+        j = get_local_id(0);
+        if (j < bs_len) {
+            a = row[j] ^ bs[j];
+            dist += popcount(a);
+        }
+        partial_dist[get_local_id(0)] = dist;
+
+        barrier(CLK_LOCAL_MEM_FENCE);
+
+        if (get_local_id(0) == 0) {
+            dist = 0;
+            for(uint t = 0; t < bs_len; t++) {
+                dist += partial_dist[t];
+            }
+            if (dist <= radius) {
+                selected[atomic_inc(counter)] = id;
+            }
+        }
+
+        barrier(CLK_LOCAL_MEM_FENCE);
+    }
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+label={lst:single_scan3},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan3}}.
+It calculates one distance per workgroup, distributing each 64-bit integer operation per worker, and then summing the results obtained by the workers.
+The sum algorith is a parallel reduction, in which the workers split the array in two parts and sum the second part in the first part every loop.
+So, the sum is calculated in {$O(\log_2(\text{number of workers per workgroup}))$}.
+This kernel only works when the number of workers per workgroup is a power-of-2.},
+]
+__kernel
+void single_scan3(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint dist;
+    ulong a;
+    uint j;
+
+    for (uint id = get_group_id(0); id < sample; id += get_num_groups(0)) {
+
+        const __global ulong *row = bitstrings + id*bs_len;
+
+        dist = 0;
+        j = get_local_id(0);
+        if (j < bs_len) {
+            a = row[j] ^ bs[j];
+            dist = popcount(a);
+        }
+        partial_dist[get_local_id(0)] = dist;
+
+        // Parallel reduction to sum all partial_dist array.
+        for(uint stride = get_local_size(0)/2; stride > 0; stride /= 2) {
+            barrier(CLK_LOCAL_MEM_FENCE);
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] +=
+                    partial_dist[get_local_id(0) + stride];
+            }
+        }
+
+        if (get_local_id(0) == 0) {
+            if (partial_dist[0] <= radius) {
+                selected[atomic_inc(counter)] = id;
+            }
+        }
+
+        barrier(CLK_LOCAL_MEM_FENCE);
+    }
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+label={lst:single_scan4},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan4}}.
+This kernel is just like \lstinline{single_scan3}, but it works with any number of workers per workgroup.
+The tradeoff is that it includes an aditional step in the parallel reduction algorithm.},
+]
+__kernel
+void single_scan4(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint dist;
+    ulong a;
+    uint j;
+
+    for (uint id = get_group_id(0); id < sample; id += get_num_groups(0)) {
+
+        const __global ulong *row = bitstrings + id*bs_len;
+
+        dist = 0;
+        j = get_local_id(0);
+        if (j < bs_len) {
+            a = row[j] ^ bs[j];
+            dist = popcount(a);
+        }
+        partial_dist[get_local_id(0)] = dist;
+
+        uint old_stride = get_local_size(0);
+        __local uint extra;
+        extra = 0;
+        for(uint stride = get_local_size(0)/2; stride > 0; stride /= 2) {
+            barrier(CLK_LOCAL_MEM_FENCE);
+            if ((old_stride&1) == 1 && get_local_id(0) == old_stride-1) {
+                extra += partial_dist[get_local_id(0)];
+            }
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] +=
+                    partial_dist[get_local_id(0) + stride];
+            }
+            old_stride = stride;
+        }
+
+        if (get_local_id(0) == 0) {
+            if (partial_dist[0] + extra <= radius) {
+                selected[atomic_inc(counter)] = id;
+            }
+        }
+
+        barrier(CLK_LOCAL_MEM_FENCE);
+    }
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+label={lst:single_scan5},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan5}}.
+This kernel is just like \lstinline{single_scan3}, but it explores one more detail of many GPU microarchitecture: the size of the warp.
+As the workers in the same warp are always synchronized, there is no need to synchronize them using a barrier.
+This specific kernel only works when the number of workers per workgroup is a power-of-2.},
+]
+__kernel
+void single_scan5(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint dist;
+    ulong a;
+    uint j;
+
+    for (uint id = get_group_id(0); id < sample; id += get_num_groups(0)) {
+        const __global ulong *row = bitstrings + id*bs_len;
+
+        dist = 0;
+        j = get_local_id(0);
+        if (j < bs_len) {
+            a = row[j] ^ bs[j];
+            dist = popcount(a);
+        }
+        partial_dist[get_local_id(0)] = dist;
+
+        uint stride;
+        for(stride = get_local_size(0)/2; stride > 32; stride /= 2) {
+            barrier(CLK_LOCAL_MEM_FENCE);
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] +=
+                    partial_dist[get_local_id(0) + stride];
+            }
+        }
+        barrier(CLK_LOCAL_MEM_FENCE);
+        for(/**/; stride > 0; stride /= 2) {
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] +=
+                    partial_dist[get_local_id(0) + stride];
+            }
+        }
+
+        if (get_local_id(0) == 0) {
+            if (partial_dist[0] <= radius) {
+                selected[atomic_inc(counter)] = id;
+            }
+        }
+        barrier(CLK_LOCAL_MEM_FENCE);
+    }
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+basicstyle=\scriptsize\ttfamily,
+label={lst:single_scan5_unroll},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan5_unroll}}. This kernel is exactly like \lstinline{single_scan5}, but it unrolls the last for since it has at most 5 loops.},
+]
+__kernel
+void single_scan5_unroll(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint dist;
+    ulong a;
+    uint j;
+
+    for (uint id = get_group_id(0); id < sample; id += get_num_groups(0)) {
+        const __global ulong *row = bitstrings + id*bs_len;
+
+        dist = 0;
+        j = get_local_id(0);
+        if (j < bs_len) {
+            a = row[j] ^ bs[j];
+            dist = popcount(a);
+        }
+        partial_dist[get_local_id(0)] = dist;
+
+        for(uint stride = get_local_size(0)/2; stride > 32; stride /= 2) {
+            barrier(CLK_LOCAL_MEM_FENCE);
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] += partial_dist[get_local_id(0) + stride];
+            }
+        }
+
+        // We do not need to sync because they all run in the same warp.
+        if (get_local_id(0) < 32 && get_local_size(0) >= 64) {
+            partial_dist[get_local_id(0)] += partial_dist[get_local_id(0) + 32];
+        }
+        if (get_local_id(0) < 16 && get_local_size(0) >= 32) {
+            partial_dist[get_local_id(0)] += partial_dist[get_local_id(0) + 16];
+        }
+        if (get_local_id(0) < 8 && get_local_size(0) >= 16) {
+            partial_dist[get_local_id(0)] += partial_dist[get_local_id(0) + 8];
+        }
+        if (get_local_id(0) < 4 && get_local_size(0) >= 8) {
+            partial_dist[get_local_id(0)] += partial_dist[get_local_id(0) + 4];
+        }
+        if (get_local_id(0) < 2 && get_local_size(0) >= 4) {
+            partial_dist[get_local_id(0)] += partial_dist[get_local_id(0) + 2];
+        }
+
+        if (get_local_id(0) == 0) {
+            partial_dist[0] += partial_dist[1];
+            if (partial_dist[0] <= radius) {
+                selected[atomic_inc(counter)] = id;
+            }
+        }
+        barrier(CLK_LOCAL_MEM_FENCE);
+    }
+}
+\end{lstlisting}
+\end{figure}
+
+\begin{figure}[!p]
+\begin{lstlisting}[
+language={[OpenCL]C},
+basicstyle=\scriptsize\ttfamily,
+label={lst:single_scan6},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan6}}.
+This kernel is just like \lstinline{single_scan5}, but it works with any number of workers per work.
+The tradeoff is an aditional step in the parallel reduction algorithm.},
+]
+__kernel
+void single_scan6(
+        __constant const uchar *bitcount_table,
+        __global const ulong *bitstrings,
+        const uint bs_len,
+        const uint sample,
+        __constant const ulong *bs,
+        const uint radius,
+        __global uint *counter,
+        __global uint *selected,
+        __local uint *partial_dist)
+{
+    uint dist;
+    ulong a;
+    uint j;
+
+    for (uint id = get_group_id(0); id < sample; id += get_num_groups(0)) {
+        const __global ulong *row = bitstrings + id*bs_len;
+
+        dist = 0;
+        j = get_local_id(0);
+        if (j < bs_len) {
+            a = row[j] ^ bs[j];
+            dist = popcount(a);
+        }
+        partial_dist[get_local_id(0)] = dist;
+
+        uint old_stride = get_local_size(0);
+        uint stride;
+        __local uint extra;
+        extra = 0;
+        for(stride = get_local_size(0)/2; stride > 32; stride /= 2) {
+            barrier(CLK_LOCAL_MEM_FENCE);
+            if ((old_stride&1) == 1 && get_local_id(0) == old_stride-1) {
+                extra += partial_dist[get_local_id(0)];
+            }
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] +=
+                    partial_dist[get_local_id(0) + stride];
+            }
+            old_stride = stride;
+        }
+        barrier(CLK_LOCAL_MEM_FENCE);
+        for(/**/; stride > 0; stride /= 2) {
+            if ((old_stride&1) == 1 && get_local_id(0) == old_stride-1) {
+                extra += partial_dist[get_local_id(0)];
+            }
+            if (get_local_id(0) < stride) {
+                partial_dist[get_local_id(0)] +=
+                    partial_dist[get_local_id(0) + stride];
+            }
+            old_stride = stride;
+        }
+
+        if (get_local_id(0) == 0) {
+            if (partial_dist[0] + extra <= radius) {
+                selected[atomic_inc(counter)] = id;
+            }
+        }
+        barrier(CLK_LOCAL_MEM_FENCE);
+    }
+}
+\end{lstlisting}
+\end{figure}
 

### 2018-03-12 Chapters/sdm-applications.tex
SDM Splitting into multiple files.
SDM Applications Noise filter

@@
 Future work may explore how sensible this classification algorithm is to these operations.
-We expect that with proper training, the algorithm will remain classifying the images with a good hit rate.
+We expect that, with proper training, the algorithm will remain classifying the images with a good hit rate.
 
@@
 \chapter{Results (vi): Image noise filtering application}
+
+\lstset{
+    language=Python,
+    basicstyle=\small\ttfamily,
+}
 
@@
 The images have 30 x 30 pixels, totaling 900 pixels per image.
-Each image is mapped into a 1,000 bit bitstring in which the bits are set according to the color of each pixel of the image.
-White pixels are equal to bit 0, and black pixels to bit 1.
+Each image is mapped into a bistrings of 1,000 bit in which the bits are set according to the color of each pixel of the image.
+White pixels are mapped to bit 0, and black pixels to bit 1.
 The 100 remaining bits are all set to zero.
@@
 Half of the letter ``I'' and half of the letter ``T'' (see Figure \ref{fig:filter-training}).
-They were written in their own addresses, i.e., write(address=bs\_image, datum=bs\_image).
+They were written into their own addresses, i.e., \lstinline{write(address=bs_image, datum=bs_image)}.
 
@@
   \label{fig:filter-training}
+\end{figure}
+
+Then, in order to test the filtering, we just have read from noisy images, and the results were really impressive.
+We were able to clean images up to 42\% of noise (see Figure \ref{fig:filter-testing}).
+While SDM has never seen a clean version of the letters, it just learned from the learning phase which pixels has appeared more frequently and chose them.
+
+\begin{figure}[!htb]
+  \subfloat[Steps of reading from letter ``T'' with 42\% of noise]{\includegraphics[width=\textwidth]{./images02/filter/T-42noise.png}}
+
+  \subfloat[Steps of reading from letter ``I'' with 42\% of noise]{\includegraphics[width=\textwidth]{./images02/filter/I-42noise.png}}
+
+  \caption{In order to test the SDM as a noise filter, we read from noisy images expecting to get a clean image.
+It is interesting to highlight that SDM has never seen a clean version letters ``T'' and ``I''.
+  \label{fig:filter-testing}
+  }
+\end{figure}
+
+A simplified mathematical analysis would be: During the learning phase, $200$ images with 15\% of noise were written to SDM, so, the average distance between them and the clean image was 150 bits.
+Thus they shared, on average, 175 hard-locations with the clean image.
+In this 175 hard-locations, the counter's value for a black pixel mapped to bit 1 was $(1-0.15) \cdot 200 - 0.15 \cdot 200 = 140$.
+Finally, let's analyse the reading.
+When reading from a noise image with 42\% of noise, the average distance between the noisy image and its clean image is 420, which means they share, on average, 6 hard-locations.
+As the average number of activated hard-locations is 1,072, the sum of their counters will be, on average, $Y = 6 \cdot 140 - \sum_{i=1}^{1072-6} X_i$, where $X_i$ is a Bernoulli trial with probability 0.5.
+Hence, $P(\text{black pixel}) = P(Y > 0) = P(6 \cdot 140 - \sum_{i=1}^{1072-6} X_i > 0) = P(\sum_{i=1}^{1066} X_i < 840) = 99.99\%$.
+But, when reading from a noise image with 45\% of noise, the average number of hard-locations shared with the clean image is only 3.
+Thus the sum of the activated hardlocations' counters will be, on average, $3 \cdot 140 - \sum_{i=1}^{1072-3} X_i$, and $P(\text{black pixel}) = P(\sum_{i=1}^{1069} X_i < 420) = 1.28 \cdot 10^{-12}$.
+The probability abruptly drops from 100\% to 0\% when the noise goes from 42\% to 45\% (see Figure \ref{fig:filter-prob-right-pixel}).
+The analysis for while pixels is exactly the same, but with opposite signs.
+The code to calculate this probability is available in ``Noise filter - Math analysis'' notebook \citep{sdmframework}.
+
+This is a simplified analysis because it does not take into consideration the hardlocations shared by the different letters.
+It works fine for our example because letter ``I'' and ``T'' share, on average, zero hardlocations.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=0.7\textwidth]{./images02/filter/prob-right-pixel.png}
+\caption{Probability of getting the right pixel when reading from an image with noise $p$. It assumes that SDM was trained with 200 images with 15\% noise.
+\label{fig:filter-prob-right-pixel}
+}
 \end{figure}

### 2018-03-12 Chapters/Chapter02.tex
SDM Minor change in generalized reading.
sdm ch2

@@
 
-It is interesting that Kanerva has proposed $z=1$ without realizing the generalized reading. Even so, he proposed the optimal $z$ --- the one with the highest critical distance.
+It is interesting that Kanerva has proposed $z=1$ without realizing the generalized reading. Even so, he proposed the $z$ with the highest critical distance.
 
@@
 
-My advisor, Alexandre Linhares, has proposed another read operation: an information-theoretical weighted reading.
+My advisor, Alexandre Linhares, has proposed another write operation: an information-theoretical weighted write.
 In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address.

### 2018-03-12 Chapters/sdm-applications.tex

@@
 This is a simplified analysis because it does not take into consideration the hardlocations shared by the different letters.
-It works fine for our example because letter ``I'' and ``T'' share, on average, zero hardlocations.
+It works fine for our example because letter ``I'' and ``T'' are almost orthogonal and share, on average, only one hardlocation.
 
@@
 
-In the learning phase, noisy images are generated and they are written into SDM chunked with their labels.
-The chunk was calculated using the exclusive or (XOR) operator.
-So, the image bitstring was written to the address of its bitstrings XOR its label bitstring --- write(addr=bs\_image $\oplus$ bs\_label, datum=bs\_image).
+If the intersection between images becomes too high, the noise filter stops working.
+We have confirmed it writing letters ``B'', ``C'', ``D'' and numeral ``8''.
+They share a great amount of hard-locations and our noise filter stopped working.
+The training sets can be seen in Figure \ref{fig:filter-training2} and the results in Figure \ref{fig:filter-results2}.
 
-Finally, in order to remove the noise of a new image, first we have to classify it (possibly using the already presented classification algorithm), and then we just have to read from the chunked address until it converges.
+A possible solution to this interference problem is to use labels.
+Each label has a random bitstring, which will be chuncked with the images before writing into SDM.
+Hence, before reading, we also have to chunck the image with the label --- which means we need to know the label of each image.
+The chunk was done using the exclusive OR (XOR) operator, i.e., $bs\_chunck = bs\_image \oplus bs\_label$.
+In other words, we run \lstinline{write(address=bs_chunk, datum=bs_label)} during the training, and \lstinline{read(address=bs_chunk)} during the testing.
+We used the same training set as before, and the results can be seen in Figure \ref{fig:filter-results2-chunk}.
+
+The chunck through exclusive OR (XOR) works because of the following theorem:
+
+\begin{theorem}
+If $v_1$ and $v_2$ are random bitstrings, then $\forall a, b, \mathbf{E} \left[ d(a \oplus v_1, b \oplus v_2) \right] = n/2$.
+\end{theorem}
+\begin{proof}
+Let $\mathbb{A} = \{ i | a^i = b^i \}$ be the indexes in which the bits of $a$ are equal to the bits of $b$, and $\mathbb{B} = \{ i | a^i \ne b^i \}$ be indexes in which the bits of $a$ are different from the bits of $b$.
+Thus,
+
+\begin{align*}
+d(a \oplus v_1, b \oplus v_2) &= \sum_{i=1}^n d(a^i \oplus v_1^i, b^i \oplus v_2^i) \\
+    &= \sum_{i=1}^n (a^i \oplus v_1^i) \oplus (b^i \oplus v_2^i) \\
+    &= \sum_{i=1}^n (a^i \oplus b^i) \oplus (v_1^i \oplus v_2^i) \\
+    &= \sum_{i \in \mathbb{A}} (a^i \oplus b^i) \oplus (v_1^i \oplus v_2^i) +
+       \sum_{i \in \mathbb{B}} (a^i \oplus b^i) \oplus (v_1^i \oplus v_2^i)
+\end{align*}
+
+For $i \in \mathbb{A}$, $a^i \oplus b^i = 0$, and follows:
+
+\begin{align*}
+(a^i \oplus b^i) \oplus (v_1^i \oplus v_2^i) &= 0 \oplus (v_1^i \oplus v_2^i) \\
+    &= v_1^i \oplus v_2^i \\
+    &= d(v_1^i, v_2^i)
+\end{align*}
+
+Hence, $\mathbf{E} \left[ \sum_{i \in \mathbb{A}} d(a^i \oplus v_1^i, b^i \oplus v_2^i) \right] = \mathbf{E} \left[ \sum_{i \in \mathbb{A}} d(v_1^i, v_2^i) \right] = |\mathbb{A}|/2$, because $v_1$ and $v_2$ are random bitstrings and their average distance is half the number of bits.
+
+For $i \in \mathbb{B}$, $a^i \oplus b^i = 1$, and follows:
+
+\begin{align*}
+(a^i \oplus b^i) \oplus (v_1^i \oplus v_2^i) &= 1 \oplus (v_1^i \oplus v_2^i) \\
+    &= d(1, v_1^i \oplus v_2^i)
+\end{align*}
+
+Hence, $\mathbf{E} \left[ \sum_{i \in \mathbb{B}} d(a^i \oplus v_1^i, b^i \oplus v_2^i) \right] = \mathbf{E} \left[ \sum_{i \in \mathbb{B}} d(1, v_1^i \oplus v_2^i) \right] = |\mathbb{B}|/2$.
+
+Finally, $\mathbf{E} \left[ d(a \oplus v_1, b \oplus v_2) \right] = |\mathbb{A}|/2 + |\mathbb{B}|/2 = n/2$, since $|\mathbb{A}| + |\mathbb{B}| = n$
+\end{proof}
+
+In our example, we just used the right label to read each image.
+But, we could have used our classification algorithm already presented in this work as a pre-processing step, and then, with the correct label, run the noise filter.
 

### 2018-03-12 Chapters/sdm-applications.tex
SDM Noise filter finished.

@@
 
-\chapter{Results (v): Supervised classification application}
+\chapter{Results (vi): Supervised classification application}
 
@@
 
-\chapter{Results (vi): Image noise filtering application}
+\chapter{Results (vii): Image noise filtering application}
 
@@
 
-If the intersection between images becomes too high, the noise filter stops working.
+If the intersection between images becomes too high, the noise filter stops working properly.
 We have confirmed it writing letters ``B'', ``C'', ``D'' and numeral ``8''.
-They share a great amount of hard-locations and our noise filter stopped working.
+They share a great amount of hard-locations and our noise filter could not filter their noise correctly.
 The training sets can be seen in Figure \ref{fig:filter-training2} and the results in Figure \ref{fig:filter-results2}.
+
+\begin{figure}[!htb]
+  \subfloat[Letter ``B'' with 15\% of noise. ]{\includegraphics[width=0.5\textwidth]{./images02/filter/training2-b.png}}
+  \subfloat[Letter ``C'' with 15\% of noise. ]{\includegraphics[width=0.5\textwidth]{./images02/filter/training2-c.png}}
+
+  \subfloat[Letter ``D'' with 15\% of noise. ]{\includegraphics[width=0.5\textwidth]{./images02/filter/training2-d.png}}
+  \subfloat[Letter ``8'' with 15\% of noise. ]{\includegraphics[width=0.5\textwidth]{./images02/filter/training2-8.png}}
+
+  \caption{Training images in which the intersection between images is too high. They were written in their own addresses --- write(address=bs\_image, datum=bs\_image).}
+  \label{fig:filter-training2}
+\end{figure}
+
+\begin{figure}[!htb]
+  \captionsetup[subfigure]{labelformat=empty}
+  \subfloat[]{\includegraphics[width=0.285\textwidth]{./images02/filter/case2-b-10noise.png}}
+
+  \subfloat[]{\includegraphics[width=0.57\textwidth]{./images02/filter/case2-c-10noise.png}}
+
+  \subfloat[]{\includegraphics[width=0.285\textwidth]{./images02/filter/case2-d-10noise.png}}
+
+  \subfloat[]{\includegraphics[width=0.57\textwidth]{./images02/filter/case2-8-10noise.png}}
+
+  \caption{When the intersection between images becomes too high, there appears some interference in the resulting image.
+All cases have 10\% noise.
+We can notice that the empty space in the right side of the ``C'' letter generates some white pixels in the right side of both ``B'' and ``D'' letters.
+  \label{fig:filter-results2}
+  }
+\end{figure}
 
@@
 
-The chunck through exclusive OR (XOR) works because of the following theorem:
+\begin{figure}[!tb]
+  \captionsetup[subfigure]{labelformat=empty}
+  \subfloat[]{\includegraphics[width=\textwidth]{./images02/filter/labelB-20noise.png}}
+
+  \subfloat[]{\includegraphics[width=\textwidth]{./images02/filter/labelC-20noise.png}}
+
+  \subfloat[]{\includegraphics[width=\textwidth]{./images02/filter/labelD-20noise.png}}
+
+  \subfloat[]{\includegraphics[width=\textwidth]{./images02/filter/label8-20noise.png}}
+
+  \caption{Using labels solves the interference problem when the intersection between images becomes too high. All cases have 20\% noise.
+  \label{fig:filter-results2-chunk}
+  }
+\end{figure}
+
+The chunck through exclusive OR (XOR) works because of Theorem \ref{thm:xor-chuncks}, which says that chuncking the images with labels will generate, on average, orthogonal bitstrings.
+Thus, these orthogonal bitstrings will not interfere with each other because they share, on average, only one hardlocation.
+
+The disadvantage of using labels is that it requires classification of the images.
+In our example, we just used the correct label with each image, but we could have used our classification algorithm as a pre-processing step, and only then run the noise filter.
 
@@
 If $v_1$ and $v_2$ are random bitstrings, then $\forall a, b, \mathbf{E} \left[ d(a \oplus v_1, b \oplus v_2) \right] = n/2$.
+\label{thm:xor-chuncks}
 \end{theorem}
@@
 
-In our example, we just used the right label to read each image.
-But, we could have used our classification algorithm already presented in this work as a pre-processing step, and then, with the correct label, run the noise filter.
-
-\chapter{Results (vii): The possibility of unsupervised reinforcement learning}
+\chapter{Results (viii): The possibility of unsupervised reinforcement learning}
 

### 2018-03-12 Chapters/sdm-applications.tex
SDM minor changes in chapter numbering and in reinforcement learning.

@@
 
-Our algorithm to decide what should be player is very simples: it read the current board from SDM.
+Our algorithm to decide what should be played is very simples: it read the current board from SDM.
 If the reading converges to another board, it chooses the movement which would bring the current board to the one read from SDM.
@@
 Then it will write $b_0 \rightarrow b_1 \rightarrow b_2 \rightarrow \cdots \rightarrow b_n$, with possibly different weights for each transition.
-If it loses, it will reverse the board (replace X by O and vice versa), and will act as if it had won.
+If it loses, it will reverse the board (replacing X by O and vice versa), and will act as if it had won.
 Hence, it will learn which sequences lead to victory.
@@
     \end{tabular}}
-    {$\xrightarrow{w_0}$}
+    {$\rightarrow$}
     \subfloat[$b1$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
@@
     \end{tabular}}
-    {$\xrightarrow{w_1}$}
+    {$\rightarrow$}
     \subfloat[$b2$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
@@
     \end{tabular}}
-    {$\xrightarrow{w_2}$}
+    {$\rightarrow$}
     \subfloat[$b3$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
@@
     \end{tabular}}
-    {$\xrightarrow{w_3}$}
+    {$\rightarrow$}
     \subfloat[$b4$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
@@
     \end{tabular}}
-    {$\xrightarrow{w_4}$}
+    {$\rightarrow$}
     \subfloat[$b5$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
@@
     \end{tabular}}
-    {$\xrightarrow{w_5}$}
+    {$\rightarrow$}
     \subfloat[$b6$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}
@@
     \end{tabular}}
-    {$\xrightarrow{w_6}$}
+    {$\rightarrow$}
     \subfloat[$b7$]{\begin{tabular}{p{\widthof{x}}|p{\widthof{x}}|p{\widthof{x}}}

### 2018-03-14 Chapters/sdm-applications.tex
SDM Reinforcement learning

@@
 Reinforcement learning is a machine learning algorithm which learns from the rewards of its actions.
-So, it receives the game state as input, then it decides which action will be taken, and finally it learns from the rewards of all the actions it has chosen.
+So, it receives the game state as input, it decides which action will be taken, and then it learns from the rewards of all the actions it has chosen.
 In theory, it learns after each reward feedback it receives, improving its decision over time and presenting intelligent behavior.
-A positive reward would indicate that the chosen action should be encouraged.
-While a negative reward would indicate the opposite.
+A positive reward would indicate that the chosen action should be encouraged, while a negative reward would indicate the opposite.
 In some algorithms, there may be a neutral reward which would indicate that the chosen action was neither positive nor negative.
@@
 
-Our algorithm to decide what should be played is very simples: it read the current board from SDM.
-If the reading converges to another board, it chooses the movement which would bring the current board to the one read from SDM.
-If the reading does not converge, it just play randomly.
+In our approach, there were 9 possible actions: one for each cell of the TicTacToe game.
+Action 1 means playing in the first cell of the first row; action 2 means playing in the second cell of the first row; and so one.
+Figure \ref{fig:ttt-action-mapping} shows the TicTacToe board numbering and the link between each cell and its respective action.
+
+Our algorithm to decide what should be played is very simple: it reads the current board from SDM and then it chooses the valid action with highest score.
+To calculate the scores, the bitstring is split into 9 parts, one per action (Figure \ref{fig:ttt-action-mapping}).
+The number of bits 1 in each part indicates the score of its respective action.
+
+\begin{figure}
+\centering
+\begin{tikzpicture}[
+mycell/.style={minimum size=7mm},
+matrixA/.style={matrix of nodes,
+    nodes={mycell, anchor=center},
+    column sep=-\pgflinewidth,
+    row sep=-\pgflinewidth,
+    },
+mycell2/.style={draw, minimum size=7mm},
+matrixB/.style={matrix of nodes,
+    nodes={mycell2, anchor=center},
+    column sep=-\pgflinewidth,
+    row sep=-\pgflinewidth,
+}]
+
+\matrix[matrixA] (A) { 1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9 \\};
+\draw[black] (A-1-1.north east) -- (A-3-1.south east);
+\draw[black] (A-1-2.north east) -- (A-3-2.south east);
+\draw[black] (A-1-1.south west) -- (A-1-3.south east);
+\draw[black] (A-2-1.south west) -- (A-2-3.south east);
+
+\matrix[matrixB, right=of A] (B) {
+Action 1 & Action 2 & Action 3 & $\cdots$ & Action 9 \\
+};
+
+\matrix[matrixB, below=of B] (C) {
+b$_0$ & b$_1$ & b$_2$ & b$_3$ & b$_4$ & b$_5$ & $\cdots$ & b$_k$ \\
+};
+
+\draw [
+	thick,
+    dashed,
+    ->
+] (A-1-1.north) edge[bend left] (B-1-1.north);
+
+\draw [
+	thick,
+    dashed,
+    ->
+] (A-1-2.north) edge[bend left] (B-1-2.north);
+
+\draw [
+	thick,
+    dashed,
+    ->
+] (A-1-3.north) edge[bend left] (B-1-3.north);
+
+\draw [
+	thick,
+    dashed,
+    ->
+] (A-2-3.south) edge[bend left] (B-1-5.north);
+
+\draw[dashed] (B-1-1.south west)--(C-1-1.north west);
+\draw[dashed] (B-1-1.south east)--(C-1-8.north east);
+\draw [
+	thick,
+    decoration={
+        brace,
+        mirror,
+		amplitude=0.2cm,
+        raise=0.2cm
+    },
+    decorate
+] (C-1-1.south west) -- (C-1-8.south east)
+node [pos=0.5,anchor=north,yshift=-0.5cm] {$k$ bits per action};
+
+\end{tikzpicture}
+
+\caption{Each action is a cell in the TicTacToe board and is mapped to slice of the bitstring.
+\label{fig:ttt-action-mapping}}
+\end{figure}
+
+%If the reading converges to another board, it chooses the movement which would bring the current board to the one read from SDM. If the reading does not converge, it just play randomly.
 
 After a game has finished, it is time to learn from its decisions.
-Thus, if SDM wins the game, it will write the whole sequence of boards to SDM.
+Our algorithm has three rewards: positive, negative, and neutral.
+It always learn from both players, no matter who wins or if it was a draw.
+The winner's sequence of actions feed our positive reward learning.
+The loser's sequence of actions feed our negative reward learning.
+If it is a draw, both sequences feed our neutral reward learning.
+
+Each board state has a unique random bitstring.
+For instance, each board in Figure \ref{fig-ttt-example} has its own unique random bitstring.
+Thus, if a specific board state is reached again, SDM will return the scores of each action.
+
 Let $b_0, b_1, b_2, \dots, b_n$ be the board sequence of the game (see Figure \ref{fig-ttt-example}).
-Then it will write $b_0 \rightarrow b_1 \rightarrow b_2 \rightarrow \cdots \rightarrow b_n$, with possibly different weights for each transition.
-If it loses, it will reverse the board (replacing X by O and vice versa), and will act as if it had won.
-Hence, it will learn which sequences lead to victory.
-When a new board appears, it may have already seen that situation and will decide according to the sequences which goes towards victory.
-This is our positive reward learning.
+In order to learn, our algorithm will map each action which goes from $b_k$ to $b_{k+1}$ to a reward bitstring and then will write a pointer from $b_k$ to this reward bitstring, i.e., \lstinline{write(bs_board, bs_reward)}.
 
@@
 
-It is also important to learn when a draw happens --- after all, it is better to tie than to lose, right?
-In this case, the sequence of boards is also written to SDM, but with no weight at all.
-So, if the board has appeared both in a tie sequence and in a winning sequence, it would be more likely to choose the winning one because it was written with greater weight.
-This is our neutral reward learning.
+The reward bitstring may be a positive, negative, or neutral bitstring.
+The positive reward bitstring is randomly generated and then only the bits related to the action will be set to 1 (see Figure \ref{fig:ttt-positive-reward-bitstring}).
+The negative reward bitstring is also randomly generated and then only the bits related to the action will be set to 0.
+Finally, the neutral reward bitstring is only a random bitstring.
 
-Finally, we also want to prevent losing games.
-So, when it loses a game, it will stimulate movements different from the chosen ones.
-Thus, for each transition $b_k \rightarrow b_{k+1}$ made by its action, it will write all possible transitions from $b_k$ but $b_{k+1}$.
+The idea behind these reward bitstrings is to increase the score of positive rewards (all bits set to 1), and to decrease the score of negative rewards (all bits set to 0).
+The neutral reward will have, on average, half of its bits 1 and the other half 0.
+Thus, it is in the middle between a positive reward and a negative reward.
+
+\begin{figure}
+\centering
+
+\begin{tikzpicture}[
+mycell/.style={draw, minimum size=7mm},
+matrixA/.style={matrix of nodes,
+    nodes={mycell, anchor=center},
+    column sep=-\pgflinewidth,
+    row sep=-\pgflinewidth,
+}]
+
+\matrix[matrixA, right=of A] (A) {
+Action 1 & $\cdots$ & Action $k$ & $\cdots$ & Action 9 \\
+};
+
+\matrix[matrixA, below=of B] (B) {
+0 & 1 & 0 & 1 & $\cdots$ & 1 & 1 & 1 & 1 & $\cdots$ & 1 & 0 & 0 & 1 \\
+};
+
+\draw[dashed] (A-1-1.south west)--(B-1-1.north west);
+\draw[dashed] (A-1-1.south east)--(B-1-4.north east);
+\draw[dashed] (A-1-3.south west)--(B-1-6.north west);
+\draw[dashed] (A-1-3.south east)--(B-1-9.north east);
+\draw[dashed] (A-1-5.south west)--(B-1-11.north west);
+\draw[dashed] (A-1-5.south east)--(B-1-14.north east);
+
+\draw [
+	thick,
+    decoration={
+        brace,
+        mirror,
+		amplitude=0.2cm,
+        raise=0.2cm
+    },
+    decorate
+] (B-1-1.south west) -- (B-1-4.south east)
+node [pos=0.5,anchor=north,yshift=-0.5cm] {random bits};
+
+\draw [
+	thick,
+    decoration={
+        brace,
+        mirror,
+		amplitude=0.2cm,
+        raise=0.2cm
+    },
+    decorate
+] (B-1-6.south west) -- (B-1-9.south east)
+node [pos=0.5,anchor=north,yshift=-0.5cm] {all ones};
+
+\draw [
+	thick,
+    decoration={
+        brace,
+        mirror,
+		amplitude=0.2cm,
+        raise=0.2cm
+    },
+    decorate
+] (B-1-11.south west) -- (B-1-14.south east)
+node [pos=0.5,anchor=north,yshift=-0.5cm] {random bits};
+\end{tikzpicture}
+
+\caption{Positive reward bitstrings used in our reinforcement learning algorithm.
+\label{fig:ttt-positive-reward-bitstrings}}
+\end{figure}
+
+Let the $b$ be the sequence of boards and $a$ be the sequence of actions, then $b_0 \xrightarrow{a_0} b_1 \xrightarrow{a_1} b_2 \xrightarrow{a_2} \cdots \xrightarrow{a_{n-1}} b_n$.
+Suppose there is a winner, thus the winner's actions will be the sequence $a_{n-1}, a_{n-3}, a_{n-5}, \dots$, while the loser's actions will be the sequence $a_{n-2}, a_{n-4}, a_{n-6}, \dots$.
+
+The positive reward learning will be writing the following pointers in SDM: $b_{n-1} \rightarrow a_{n-1}$, $b_{n-3} \rightarrow a_{n-3}$, and so on.
+The reward bitstring of $a_{n-1}$ will have all bits set to one in $a_{n-1}$'s slice.
+All other bits will be random.
+
+The negative reward learning will be writing the following pointers in SDM: $b_{n-2} \rightarrow a_{n-2}$, $b_{n-3} \rightarrow a_{n-3}$, and so on.
+The reward bitstring of $a_{n-2}$ will have all bits set to zero in $a_{n-2}$'s slice.
+All other bits will be random.
+
+If it is a draw, all actions will be mapped to the neutral reward bitstring, which is simply a random bitstring.
+
+There is also weights associated to positive, negative, and neutral reward.
+They are used to indicate what goal is more important.
+For instance, if SDM should try to win in the first place, no matter if it may lead to losing, the weight of the positive reward should be higher than the others.
+But, if it is more important to not lose, then the weight of the negative reward should be higher.
+
+In the very beginning, SDM is empty and its counters are zeroed.
+So, any reading will result in a random bitstring, because it will flip a coin for all counters.
+Thus, the chosen action will be random, as their scores will follow a binomial distribution.
+It is exactly the desired behavior --- we play randomly until we learn.
+In fact, it will happen every time an unknown board is seen.
 
 Internally, every board is mapped into a random bitstring and passed to SDM.
-Thus, SDM knows nothing about the boards themselves.
-It knowns only about their transition and which ones would lead to either a victory or a draw.
-As every two boards are orthogonal, SDM does not known whether two boards are consecutives or not.
-The only link between two boards is the transition written in SDM.
+As every two boards are, on average, orthogonal, SDM knows nothing about the boards themselves, neither whether they are consecutives or not.
+It knowns only the score of the actions according to the games it has seen and learned from.
+Hopefully, the actions will lead to a victory or a draw.
 
-After all, SDM knowns nothing about the boards themselves and yet it may learn how to play TicTacToe.
-
-In order to properly run the discussed algorithms, it is necessary to have two SDMs: a 0-fold and a 1-fold SDM.
-In the 0-fold SDM, every bitstring is written to its own address.
-In the 1-fold SDM, every bitstring points to another one.
-So, the transitions are written in the 1-fold SDM, while the boards themselves are written to the 0-fold SDM.
-The boards are written only once in the 0-fold, no matter how many times they appear.
-The transitions may be written more than once in the 1-fold SDM, because it would reinforce that transition.
-
-In more details, the next movement decision consists in one read from the 1-fold SDM, resulting in a bitstring.
-Then this bitstring is used in an interative reading from the 0-fold SDM, which will converge to the bitstring associated with the next board.
-If it does not converge to any board, than SDM will choose a random movement and learn from it.
-Eventually it may converge to a board which is not a sequence of the current board.
-In this case, SDM will also choose a random movement.
-
-The weight used when writing a winning sequence is calculated using ...
-
---- talk about player generations ---
+In more details, the next movement decision consists in one read from SDM, resulting in a bitstring.
+Then the scores of the actions are calculated counting the number of 1 in each part of the bitstring.
+The chosen action is the one which is valid and has the highest score.
 
@@
 
-It is an unsupervised algorithm because SDM learns playing against an opponent, who may be another SDM player, a human, or a player whose movements are always random.
+Our algorithm learned playing games against opponents.
+We had four types of opponents: (i) another SDM player, (ii) a random player whose actions are always random; (iii) a smart player whose actions wins when it can, block the opponent when it can, or are random; and (iv) a human player.
 
-Thus, in order to train a SDM player, we just have to keep it playing over and over.
+The weights of the rewards were chosen to prevent losing. Thus the weight of the negative reward was 5, while the weight of the positive reward was 2 and the neutral was 1.
+
+Every learn cycle had two parts: (i) 100 games learning and (ii) 100 games testing. So, it has never learned during the testing phase and has not affected the measure of the statistics.
 
@@
 
-\begin{forest}
-  TTT/.style args={#1:#2}{
-    make tab/.expanded=\forestove{content},
-    label={\pgfkeysvalueof{/forest/label position #1}:$#2$}
-  },
-  TTT*/.style={
-    make tab=::/::/::,
-    content/.expand once=%
-    \expandafter\vphantom\expandafter{\romannumeral-`0\forestov{content}},
-    draw=none,
-    append after command={(\tikzlastnode.north) edge (\tikzlastnode.south)},
-    for descendants={before computing xy={l*=1.2}},
-  },
-  th/.style=thick,
-  for tree={node options=draw, inner sep=+0pt, parent anchor=south, child anchor=north}
-%
-[::/::/::, TTT=r:1
-  [x::/::/::, TTT=r:-1
-    [x::/::/::, TTT=r:-1
-      [x::/::/::, TTT=r:-1
-      ]
-    ]
-  ]
-]
-\end{forest}
+When playing against the random player, SDM has already started improving after the first 100 training games.
+Its winning rate converges quickly to around 80\%, while the drawing rate starts to grow after 2,000 games and keeps growing until the end.
+The losing rate keeps decreasing until it reaches cycles of 100 games without any loss.
+See Figure \ref{fig:ttt-results-random}.
 
-\subsection{Results}
+When playing against smart player, SDM has started learning how to not lose during the first training cycles.
+The drawing rate grows quickly in the first 500 games and slowly since then.
+The winning rate grew to around 20\% and remained there until 6,000 games, while the drawing rate kept growing.
+Then, after learning how to not lose, SDM started learning how to win, since after 7,000 games, the winning rate started to increase.
+See Figure \ref{fig:ttt-results-smart}.
+
+When playing against another SDM player, both player quickly learns how to not lose.
+During the first 100 testing games, without any learning game, they have behaved like two random players.
+After the 100 training games, they have learned how to not lose and the drawing rate grows quickly, reaching 100\%.
+See Figure \ref{fig:ttt-results-sdm}.
+
+When playing against mixed players, SDM has also adapted.
+In Figure \ref{fig:ttt-results-mixed}, it has played 6,000 games against random player, smart player, and another SDM player.
+In each cycle, one of them was randomly chosen.
+The number of loses over time is decreasing, while the number of wins and draws change a lot.
+It is easy to notice when the other SDM player was chosen, because all games in that cycle have draw.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-random-loses.png}}
+
+\subfloat[Results over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-random.png}}
+
+\caption{Results playing against random player. Each cyle was made of 100 games for training, and then 100 games for measuring statistics.
+\label{fig:ttt-results-random}
+}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-smart-loses.png}}
+
+\subfloat[Results over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-smart.png}}
+
+\caption{Results playing against smart player. Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
+\label{fig:ttt-results-smart}
+}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-sdm-loses.png}}
+
+\subfloat[Results over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-sdm.png}}
+
+\caption{Results playing against another SDM player. Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
+\label{fig:ttt-results-sdm}
+}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-mixed-loses.png}}
+
+\subfloat[Results over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-mixed.png}}
+
+\caption{Results playing against a randomly chosen player between random player, smart player, and another SDM player.
+Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
+\label{fig:ttt-results-mixed}
+}
+\end{figure}

### 2018-03-14 Chapters/Chapter02.tex
SDM minor changes in chapter numbering and in reinforcement learning.
sdm.. details

@@
 One just feels it.
-We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; etc.
+We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; a general lack of activity; that very feeling that something isn't right...
+without knowing what isn't right, or even what `right' means...
+etc.
 --- recalling past experiences from memory and thus generating the feeling.
@@
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
-Thus, even though intriguing results have been achieved using SDM, it requires great effort from researchers to build on top of previous work.
+Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate, effort from researchers to build on top of previous work.
 
-It is our belief that such a tool could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python, in an intuitive platform such as juypyter notebooks.
+It is our belief that a tool such as a framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python, in an intuitive platform such as juypyter notebooks.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
-This new tool provides a ready to use system in which experiments can be executed almost as soon as they are designed and it may accelerate research \citep{shen2014interactive}.
+This new tool would provide a ready to use system in which experiments could be executed almost as soon as designed --- and it might accelerate research \citep{shen2014interactive}.
 
@@
 
-\chapter{Results (vii): Information-theoretical write operation}
+\chapter{Results (ix): Information-theoretical write operation}
 

### 2018-03-14 Chapters/Chapter02.tex
final note...

@@
 \caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
-It shows that a loss of 200,000 neurons, 20\% of the total, does not affect SDM whatsoever.
+It shows that a loss of 200,000 neurons, 20\% of the total, does not seem to affect SDM whatsoever.
 \label{fig:sdm-neuron-death-200k}}
@@
 
-In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is absolutely robust up to 200,000 neuron deaths which is 20\% of all hard-locations.
-This result is pretty impressive and really surprised us.
-
+In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is robust up to 200,000 neuron deaths which is 20\% of all hard-locations.
+Its robustness is astonishing.
 In fact, SDM begins to be significantly affected by loss of neurons after 600,000 neuron deaths (Figure \ref{fig:sdm-neuron-death-1m}), and obviously forgets everything when all neurons are dead.
@@
 \caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
-The more neurons are death, the smaller the critical distance, i.e., the worse the SDM recall.
+The more neurons are lost, the smaller the critical distance, i.e., the worse the SDM recall.
 \label{fig:sdm-neuron-death-1m}}
@@
 It is interesting that 500,000 neuron deaths has a minor effect in SDM's recall capability (see Figure \ref{fig:sdm-neuron-death-500k}).
-It is analogous to do an hemispherectomy in a person and, after the procedure, the person is able to recall and learn almost just like before.
-In fact, there are clinical reports of children submitted to hemispherectomy who live an almost normal life with minor motor problems.
+It is analogous to do an hemispherectomy in a person and, after the procedure, the person being able to recall and learn almost just like before.
+In fact, there are clinical reports of children submitted to hemispherectomy who live an almost normal life with minor function problems.
 
@@
 
-An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance is really small, i.e., SDM recall capacity is very diminished.
+An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance becomes small, i.e., SDM recall capacity is very diminished.
 After 900,000 neuron deaths the critical distance is zero.
-In this case, everything has been forgot.
+In this case, everything has been lost.
 
@@
 We run exactly the same simulation for a 256-bit SDM with 1,000,000 hard-locations.
-The results were even more surprising, because the 256-bit SDM was more robust to loss of neuron than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
+The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
 Notice that the loss of 50\% of neurons barely affected the 256-bit SDM which remained functional even facing an enormous loss of 90\% of neurons.
@@
 
-We propose to provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypotheses and test them computationally through minimal effort.
+We provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypotheses and test them computationally through minimal effort.
 The framework is well-documented for public release at this time (http://sdm-framework.readthedocs.io), it has already served as the backbone of Chada's Ph.D.

### 2018-03-14 Chapters/sdm-applications.tex
final note...

@@
 Thus they shared, on average, 175 hard-locations with the clean image.
-In this 175 hard-locations, the counter's value for a black pixel mapped to bit 1 was $(1-0.15) \cdot 200 - 0.15 \cdot 200 = 140$.
+In these 175 hard-locations, the counter's value for a black pixel mapped to bit 1 was $(1-0.15) \cdot 200 - 0.15 \cdot 200 = 140$.
 Finally, let's analyse the reading.
@@
 This is a simplified analysis because it does not take into consideration the hardlocations shared by the different letters.
-It works fine for our example because letter ``I'' and ``T'' are almost orthogonal and share, on average, only one hardlocation.
+It works fine for our example because letter ``I'' and ``T'' are almost orthogonal and share, on average, only one hard location.
 
@@
 \end{proof}
+
+A final note is in order.
+In no way is it claimed here that these bitstrings should form a plausible representation of letters in the human mind (See, for instance, \citet{hofstadter1995seeing} for a marvelous discussion of the subtleties and fluidity involved in that process).
+These letters should merely be seen as \emph{invariant patterns} that are processed by the memory, which is able to capture their underlying invariant structure even when presented only under heavy noise.
 

### 2018-03-14 Chapters/Chapter02.tex
little garibation on conclusion...

@@
 
-\section{Future work}
-
 Here are interesting questions that have been considered during this work, but have had to be left for future research.
 
-\subsection{Multiple levels}
+% \subsection{Multiple levels}
 
-Deep neural networks, alphaGo, alphazero
+% Deep neural networks, alphaGo, alphazero
 
-\subsection{i versus l}
+% \subsection{i versus l}
 
-\subsection{Magic numbers}
+\section{Magic numbers}
 
@@
 
-\subsection{Classification with context using sequences --- for words instead of only letters}
+% \subsection{Classification with context using sequences --- for words instead of only letters}
 
-\subsection{Symmetrical, rapidly accessible, Hard Locations}
+\section{Symmetrical, rapidly accessible, Hard Locations}
 
@@
 
-\section{Seeing farther}
+\section{Paving and illuminating the pathway}
 
@@
 
-If
-
+All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested.
 We have generated a Docker image, which makes it even easier to explore the framework.
 After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.
-
-All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested. We invite readers to take a look and explore a little bit.
+We invite the reader to take a look and explore a little bit.
 
@@
 Kanerva has left the trail.
-It is my job to pave it and to throw light at it and to try to deliver an easier pathway for the next generation.
+It is my job to pave it and illuminate it and to try to deliver an easier pathway for the next generation.
 Some essays completely shut the door close at the end; this one intends to leave it wide open.

### 2018-03-14 Chapters/sdm-applications.tex
little garibation on conclusion...

@@
 
-\subsection{Training}
+\section{Training}
 
@@
 
-\subsection{Results}
+\section{Results}
 
@@
 In each cycle, one of them was randomly chosen.
-The number of loses over time is decreasing, while the number of wins and draws change a lot.
+The number of losses over time is decreasing, whereas the number of wins and draws change a lot.
 It is easy to notice when the other SDM player was chosen, because all games in that cycle have draw.
@@
 \centering
-\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-random-loses.png}}
+\subfloat[Losses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-random-loses.png}}
 
@@
 \centering
-\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-smart-loses.png}}
+\subfloat[Losses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-smart-loses.png}}
 
@@
 \centering
-\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-sdm-loses.png}}
+\subfloat[Losses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-sdm-loses.png}}
 
@@
 \centering
-\subfloat[Loses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-mixed-loses.png}}
+\subfloat[Losses over time]{\includegraphics[width=\textwidth]{./images02/ttt/results-mixed-loses.png}}
 

### 2018-03-14 Chapters/Chapter02.tex
garibation...

@@
 The framework is well-documented for public release at this time (http://sdm-framework.readthedocs.io), it has already served as the backbone of Chada's Ph.D.
-thesis.
+thesis \citep{chada2016you}.
 The single-line command ``pip install sdm'' will install the framework on posix-like systems, and single-line commands will let users test the framework, generate some of the figures from Kanerva's theoretical predictions in their own machines, and --- if interested enough ---, test their own theories and improve the framework, and the benchmarks used to evaluate the framework, in open-source fashion.
@@
 
+Beyond speed, I have also strived for \emph{ease of use}.
 All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested.
-We have generated a Docker image, which makes it even easier to explore the framework.
+I have generated a Docker image, which makes it even easier to explore the framework.
 After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.
@@
 
-The overarching intention here is to not only provide a starting point, but to provide a Framework in which SDM research can be conducted.
-Consider, for example, having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).
+The overarching intention here is to not only provide a starting point, but to provide a documented Framework in which SDM research can be conducted.
+Consider having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).
 For example, some of the benchmarks that we plan to develop in future research is: how fast is convergence through iterative reading?
@@
 Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
-We have referred to this modified read operation as Chada read (the tale is that my friend \text{\&} colleague, Dr.
-Daniel de Magalhães Chada, along with Linhares, did not consult Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterwards).
+We have referred to this modified read operation as Chada read\footnote{Legend has it that my friend \text{\&} colleague, Dr.
+Daniel de Magalhães Chada, along with Linhares, did not consult and re-check with Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterwards.}.
 Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.

### 2018-03-14 Chapters/Chapter02.tex
quotes

@@
 \chapter{Introduction}
+
+\bigskip
+
+\begin{flushright}{\slshape
+    {Before I came across it, I had read many articles describing intricate properties of neurons, the many varieties of neurotransmitters and second-messenger molecules, synaptic modification processes, complex and hypercomplex cells, columns in the visual cortex, and so on.
+All of this was fascinating but seemed very biological and microscopic:  altogether quite distant from everyday mental experiences.
+I simply had never come across a cogent theory that addressed brain mechanisms on a global level.
+Pentti Kanerva's memory model was a revelation for me: it was the very first piece of research I had ever come across that made me feel I could glimpse the distant goal of understanding how the brain works as a whole.
+It gave me a concrete sense for how familiar mental phenomena could be thought of as distributed patterns of micro-events, thanks to beautiful mathematics.
+	.} \\ \medskip
+    --- Douglas Hofstadter}
+\end{flushright}
+\bigskip
+\bigskip
 

### 2018-03-14 Chapters/Conclusion.tex
quotes opressoras

@@
 To computing or to business?
-Recall that the project was created as a risk-management strategy against the far-fetched, but not impossible, possibility of having machines yielding too much power.
+Recall that the project was created as a risk-management strategy against the far-fetched, science-fiction sounding --- but not impossible ---, possibility of having machines yielding too much power.
 What about corporations like Uber?
@@
 Should this method, if implementable as an algorithm (like PageRank \citep{brin1998anatomy}) or a data structure (like a blockchain) be discussed in conferences of `computer science' or `business'?
-It seems quite arbitrary to name a single group, as a whole new ecosystem seems to be emerging within those two.
+It seems quite arbitrary to name a single group, as a whole new ecosystem seems to have emerged within those two.
 That is why this thesis is computational and why it is business.

### 2018-03-14 Chapters/Chapter02.tex
quotes opressoras

@@
 Pentti Kanerva's memory model was a revelation for me: it was the very first piece of research I had ever come across that made me feel I could glimpse the distant goal of understanding how the brain works as a whole.
-It gave me a concrete sense for how familiar mental phenomena could be thought of as distributed patterns of micro-events, thanks to beautiful mathematics.
-	.} \\ \medskip
+It gave me a concrete sense for how familiar mental phenomena could be thought of as distributed patterns of micro-events, thanks to beautiful mathematics.} \\ \medskip
     --- Douglas Hofstadter}

### 2018-03-14 Chapters/Chapter03.tex
quotes opressoras

@@
 
-The way innovations diffuse in the market is an important and useful topic in marketing, which was made popular through Rogers' work \citep{rogers1962diffusion} and has influenced many marketing researchers.
+\bigskip
+
+\begin{flushright}{\slshape
+    {Why do entrepreneurs appear, not continuously, that is singly in every appropriately chosen interval, but in clusters?
+Exclusively because the appearance of one or a few entrepreneurs facilitates the appearance of others, and these the appearance of more, in ever-increasing numbers.}
+	\\ \medskip
+    --- Joseph Schumpeter}
+\end{flushright}
+\bigskip
+\bigskip
+
+The way innovations diffuse through markets is an important and useful topic in marketing, which was made popular through Rogers' work \citep{rogers1962diffusion} and has influenced many marketing researchers.
 Within innovation diffusion, \citet{bass1969} proposed a model which forecasts how many people will have adopted new products or technologies by a given point in time.

### 2018-03-14 Chapters/Chapter02.tex
SDM Information theory

--- 2018-03-14 Chapters/Chapter02.tex
@@
 
+My advisor, Alexandre Linhares, has proposed another write operation: an information-theoretical weighted write.
+In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address.
+The logic behind it is to vary the importance of each hard-location inside the circle.
+It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
+Closer hard locations have lower probabilies and therefore should encode more information.
+
+Note that this is not the first time in which a weighted function has been applied to writing in SDM --- \citet{hely1997new} suggest a rather complex spreading model based on floating point signals in the interval [0.05, 1.0] --- they were, however, only able to test their model with 1,000 hard locations.
+
+Consider the following.
+Information Theory \citep{cover2012elements} let us compute the precise amount of information in an event, when given its probability $p$, through the measure of \emph{self-information}:
+
+$$
+I(p)= -log_2(p)
+$$
+
+Now, given any two $n$-sized bitstrings, the probability of their Hamming distance being exactly $d$ is given by $P(X=d)= 2^{-n} \binom{n}{d}$, and the probability of it being at most $d$ is:
+
+$$
+P(X\leq d)= 2^{-n} {\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}}
+$$
+
+But we must consider that not all hardlocations are activated in each write operation, which changes our probability function. Thus, let $r$ be the access radius then:
+
+\begin{align*}
+P(X = d | X \leq r) &= \frac{P((X = d) \cap (X \leq r))}{P(X \leq r)} \\
+    &= \frac{P(X = d)}{P(X \leq r)}, \quad \text{as $d \leq r$} \\
+    &= \frac{ 2^{-n} \binom{n}{d} }{ 2^{-n} \sum_{i=0}^{r} \binom{n}{i} } \\
+    &= \frac{\binom{n}{d}}{\sum_{i=0}^{r} \binom{n}{i}}, \quad d \leq r
+\end{align*}
+
+And the probability of it being at most $d$ is:
+
+$$
+P(X \leq d | X \leq r) = \frac{\sum_{i=0}^{d} \binom{n}{i}}{\sum_{i=0}^{r} \binom{n}{i}}, \quad d \leq r
+$$
+
+As expected, $P(X \leq d | X \leq r) = 1$ when $d = r$.
+
+Hence the weighted write would, on each hard location, sum (or subtract) using the following weights, as seen in Figure \ref{fig:info-theory-hypothesis}:
+
+$$
+w(d) = -\log_2 \left( P(X = d | X \leq r) \right) = - \log_2 \binom{n}{d} + \log_2 \sum_{i=0}^{r} \binom{n}{i}, \quad d \leq r
+$$
+
 \begin{figure}[h!]
   \centering
-  \subfloat[$w_1(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}
+  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/info-theory/weights.png}}
 
-  \subfloat[$w_1(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}
+  \subfloat[$w_2(d)$ for the desired range.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region.png}}
 
-  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}
+  \subfloat[Stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region-stepwise.png}}
 
@@
 
+The initial results of this \emph{Shannon write} operation can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
+It seems that, when $n=1,000$, $H=1,000,000$, $r=451$, and 10,000 written random bitstrings, the critical distance increased from around 221 to around 250.
+This increase may be interpreted as an improvement in SDM, because it would converge to the correct bitstring even for farther bitstrings.
+Note that 29 additional bits imply an attractor area $2^{29}$ times larges than the original.
+It may affect memory capacity, possibly increasing it.
+
+Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
+
 \begin{figure}[h!]
   \centering
-  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}
+  \subfloat[Write process weighted by the amount of information contained in the distance between the written bitstring and each hard location \label{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73.png}}
 
-  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}
+  \subfloat[Zoom in Figure \ref{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73-zoom.png}}
 
-  \subfloat[stepwise $\left \lfloor{w_2(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}
+  \subfloat[Behavior of weighted write operation according to the distance from the center and the number of items previously stored in the memory]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-heatmap.png}}
 
-  \caption{SOON TO BE DEPRECATED.
+  \caption{Behavior of the critical distance under the information-theoretic weighted write operation when $n=1,000$, $H=1,000,000$ and $r=451$.}
+  \label{fig:info-theory-experiments}
+\end{figure}
+
+%It is easy to interpret this weight through a binary tree approach.  How many binary questions would be needed to precisely define a bitstring inside the access radius and exactly $d$ bits away?
+
+%Another possibility would be to use the sum of all distances closer (and less likely) locations within the weighting function $w(d)$,
+
+%$w(d) = -\log_2 \left( 2^{-n} \displaystyle\sum_{i=0}^{d}{\binom{n}{i}} \right) = n - \log_2 \displaystyle\sum_{i=0}^{d}{\binom{n}{i}}$. \\
+
+%This can be seen in \ref{fig:info-theory-sum-hypothesis}.
+
+%and, consequently,
+
+%$p(H\geq n-d)=2^{-n}{\displaystyle\sum_{i=n-d}^{n}{\binom{n}{i}}  }$, \\
+
+%$p(d+1 \leq H \leq n-d-1)=2^n - 2^{1-n}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}, \forall d<n/2}$. \\
+
+%\begin{figure}[h!]
+%  \centering
+%  \subfloat[$w_1(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}
+%
+%  \subfloat[$w_1(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}
+%
+%  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}
+
+%  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
+(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
+%  \label{fig:info-theory-hypothesis}
+%\end{figure}
+
+%\begin{figure}[h!]
+%  \centering
+%  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}
+%
+%  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}
+%
+%  \subfloat[stepwise $\left \lfloor{w_2(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}
+%
+%  \caption{SOON TO BE DEPRECATED.
 Shannon write operation:  Computing the sum of low-likelihood signals.
@@
 }
-  \label{fig:info-theory-sum-hypothesis}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Kanerva's model]{\includegraphics[width=\linewidth]{./images02/new-images/kanerva-15iter-reads.jpeg}}
-
-  \subfloat[Write process weighted by the amount of information contained in the distance between the written bitstring and each hard location]{\includegraphics[width=\linewidth]{./images02/new-images/info-theory-15iter-reads.jpeg}}
-
-  \caption{(a) and (b) show the behavior of the critical distance under Kanerva's model and the information-theoretic one, respectively.}
-  \label{fig:info-theory-experiments}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon.jpeg}}
-
-  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon-region.jpeg}}
-
-  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon-stepwise.jpeg}}
-
-  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
-(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
-  \label{fig:info-theory-hypothesis}
-\end{figure}
-
-My advisor, Alexandre Linhares, has proposed another write operation: an information-theoretical weighted write.
-In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address.
-The logic behind it is to vary the importance of each hard-location inside the circle.
-It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
-Closer hard locations have lower probabilies and therefore should encode more information.
-
-Consider the following.
-Information Theory \citep{cover2012elements} let us compute the precise amount of information in an event, when given its probability $p$, through the measure of \emph{self-information}:\\
-
-$I(p)= -log_2(p).$ \\
-
-Now, given any two $n$-sized bitstrings, the probability of their Hamming distance being $d$ is given by,
-
-$p(H=d)= {2^{-n} \binom{n}{d} }$ \\
-
-And the probability of it being at most $d$ is \\
-
-$p(H\leq d)= 2^{-n} {\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}}  $, \\
-
-and, consequently,
-
-$p(H\geq n-d)=2^{-n}{\displaystyle\sum_{i=n-d}^{n}{\binom{n}{i}}  }$, \\
-
-$p(d+1 \leq H \leq n-d-1)=2^n - 2^{1-n}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}, \forall d<n/2}$. \\
-
-Hence the weighted write would, on each hard location, sum (or subtract) the following:  \\
-
-$w(d) = -\log_2 \left( 2^{-n} \binom{n}{d} \right) = n - \log_2 \binom{n}{d}$, as seen in Figure \ref{fig:info-theory-hypothesis}.
-
-It is easy to interpret this data though a binary tree approach.  How many binary questions would be needed to precisely define a bitstring?
-
-Another possibility would be to use the sum of all distances closer (and less likely) locations within the weighting function $w(d)$,
-
-$w(d) = -\log_2 \left( 2^{-n} \displaystyle\sum_{i=0}^{d}{\binom{n}{i}} \right) = n - \log_2 \displaystyle\sum_{i=0}^{d}{\binom{n}{i}}$. \\
-
-This can be seen in \ref{fig:info-theory-sum-hypothesis}.
-
-The initial results of this \emph{Shannon write} operation can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
-It seems that the critical distance increases by a number of bits.
-Note that 10 additional bits imply an attractor $2^{10}$ of the size of the original.
-Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
-
-Note, finally, that this is not the first time in which a weighted function has been applied to writing in SDM --- \citet{hely1997new} suggest a rather complex spreading model based on floating point signals in the interval [0.05, 1.0] --- they were, however, only able to test their model with 1,000 hard locations.
+%  \label{fig:info-theory-sum-hypothesis}
+%\end{figure}
+%
 

### 2018-03-14 Chapters/Chapter02.tex
quotes (S. Ulam.)

--- 2018-03-14 Chapters/Chapter02.tex
@@
 \begin{flushright}{\slshape
-    {Before I came across it, I had read many articles describing intricate properties of neurons, the many varieties of neurotransmitters and second-messenger molecules, synaptic modification processes, complex and hypercomplex cells, columns in the visual cortex, and so on.
-All of this was fascinating but seemed very biological and microscopic:  altogether quite distant from everyday mental experiences.
-I simply had never come across a cogent theory that addressed brain mechanisms on a global level.
-Pentti Kanerva's memory model was a revelation for me: it was the very first piece of research I had ever come across that made me feel I could glimpse the distant goal of understanding how the brain works as a whole.
-It gave me a concrete sense for how familiar mental phenomena could be thought of as distributed patterns of micro-events, thanks to beautiful mathematics.} \\ \medskip
-    --- Douglas Hofstadter}
+    {Pentti Kanerva's memory model was a revelation for me: \\
+    it was the very first piece of research I had ever come across\\
+    that made me feel I could glimpse the distant goal of understanding \\
+    how the brain works as a whole. It gave me a concrete sense for\\
+    how familiar mental phenomena could be thought of as distributed\\
+    patterns of micro-events, thanks to beautiful mathematics.} \\ \medskip
+    --- Douglas Hofstadter
+    \bigskip
+    \bigskip
+
+    {How is memory gradually built up during one's conscious, or\\
+    even unconscious, life and thought?  My guess is that everything \\
+    we experience is classified and registered on very many parallel \\
+    channels in different locations} \\ \medskip
+    --- Stanislaw Ulam}
 \end{flushright}

### 2018-03-14 Chapters/Chapter03.tex
quotes (S. Ulam.)

@@
 
-As this is a working paper, it is also intended to include the analyses of other datasets, like either Twitter's number of \textit{active users}, or WhatsApp's, or Netflix's, or Reddit's, or Dropbox's, or Waze's.
+As this is a working paper, it is also intended to include the analyses of other datasets, like either Twitter's number of \textit{active users}, or WhatsApp's, or Netflix's, or Reddit's, or Dropbox's, or Waze's, etc.
 This would enhance the proposed model power of forecasting.
@@
 Although there is an extensive literature on negative word-of-mouth, this literature does not predict which model would be the best.
-But, if any of these companies discloses the number of \textit{total users} and \textit{active users}, the estimation method could be adapted to estimate the parameters using both pieces of information at the same time, which would make possible to verify which of the models of rejection is the best.
+But, if any of these companies discloses the number of \textit{total users} and \textit{active users}, the estimation method could be adapted to estimate the parameters using both pieces of information at the same time, which would make possible to verify which of the models of rejection provides the best fit.
 

### 2018-03-14 Chapters/Conclusion.tex
quotes (S. Ulam.)

@@
 The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer \citep{TonnyMartins}.
-Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citep{AndrewNgAI}, as his way to emphasize the potential transformational power of the technology.
+Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citep{AndrewNgAI}, as his hyperbolic way to emphasize the potential transformational power of the technology.
 

### 2018-03-14 Chapters/Chapter02.tex
shannon

@@
 We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; a general lack of activity; that very feeling that something isn't right...
-without knowing what isn't right, or even what `right' means...
+without knowing precisely what isn't right, or even what `right' means...
 etc.
@@
   \centering
-  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/info-theory/weights.png}}
+  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/info-theory/weights.png}}
 
-  \subfloat[$w_2(d)$ for the desired range.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region.png}}
+  \subfloat[$w(d)$ for the desired range.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region.png}}
 
-  \subfloat[Stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region-stepwise.png}}
+  \subfloat[Stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region-stepwise.png}}
 
@@
 This increase may be interpreted as an improvement in SDM, because it would converge to the correct bitstring even for farther bitstrings.
-Note that 29 additional bits imply an attractor area $2^{29}$ times larges than the original.
-It may affect memory capacity, possibly increasing it.
-
+Note that 29 additional bits imply an attractor area $2^{29}$ times larger than the original.
+This Shannon write may affect memory capacity --- possibly increasing it.
 Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
+Whereas these are just some initial tests, the idea seems meritorious so far.
+As for future research, we will execute all tests in the thesis and compare this Shannon write with the original Kanerva model.
 

### 2018-03-14 Chapters/Chapter02.tex
SDM Conclusion

--- 2018-03-14 Chapters/Chapter02.tex
@@
 \begin{flushright}{\slshape
-    {Pentti Kanerva's memory model was a revelation for me: \\
-    it was the very first piece of research I had ever come across\\
-    that made me feel I could glimpse the distant goal of understanding \\
-    how the brain works as a whole. It gave me a concrete sense for\\
-    how familiar mental phenomena could be thought of as distributed\\
-    patterns of micro-events, thanks to beautiful mathematics.} \\ \medskip
-    --- Douglas Hofstadter
-    \bigskip
-    \bigskip
-
-    {How is memory gradually built up during one's conscious, or\\
-    even unconscious, life and thought?  My guess is that everything \\
-    we experience is classified and registered on very many parallel \\
-    channels in different locations} \\ \medskip
-    --- Stanislaw Ulam}
+    {Before I came across it, I had read many articles describing intricate properties of neurons, the many varieties of neurotransmitters and second-messenger molecules, synaptic modification processes, complex and hypercomplex cells, columns in the visual cortex, and so on.
+All of this was fascinating but seemed very biological and microscopic:  altogether quite distant from everyday mental experiences.
+I simply had never come across a cogent theory that addressed brain mechanisms on a global level.
+Pentti Kanerva's memory model was a revelation for me: it was the very first piece of research I had ever come across that made me feel I could glimpse the distant goal of understanding how the brain works as a whole.
+It gave me a concrete sense for how familiar mental phenomena could be thought of as distributed patterns of micro-events, thanks to beautiful mathematics.} \\ \medskip
+    --- Douglas Hofstadter}
 \end{flushright}
@@
 
-% \subsection{i versus l}
+\section{``i'' versus ``l'' issue}
+
+The classification algorithm had some problems classifying the letter ``i'' and ``l''.
+It may have happened because their distance is not enough.
+In this case, SDM could not differenciate the details --- it has only considered the big picture.
+Although this behavior is close to how we see things, we also have the ability to zoom and focus in the details, discerning letter ``i'' from letter ``l''.
+
+I have run the classification algorithm under the MNIST database of handwritten digits \citep{deng2012mnist}.
+First, SDM has been trained with the 60,000 training images, and then it classified the 10,000 testing images.
+It has given the correct classification for 77.22\% of the images, which is inferior to the specialized algorithms.
+For instance, in 1998, \citet{lecun1998gradient} has developed algorithms which reaches from 88\%, using linear classifier, to 99.7\%, using convolutional net.
+For a review of the algorithms' performance in the MNIST database, see \cite{mnist}.
+
+Looking into the reason behind images incorrectly classified, I have found that the issue is very related to the ``i'' versus ``l'' issue.
+Some handwritten digits are very close to others, and a ``2'' or a ``7'' may look like a ``1'', for instance.
+So, how can we solve this issue without using anything specific to images?
+The machine learning algorithms use specific tecniques to ace performance.
+I would like to unveil a solution closer to how we behave, even if this leads to smaller performance.
+
+An unexplored idea is to use other SDMs which communicates.
+A first SDM would write the whole picture, just like we have done.
+Another SDM would write specific parts of the image, just like our eye focusing in specific parts.
+When reading, they may compose the counters and give a more precise classification.
 
@@
 
-\section{Symmetrical, rapidly accessible, Hard Locations}
+\section{Symmetrical, rapidly accessible, hard locations}
 

### 2018-03-14 Chapters/Chapter02.tex
i versus l

--- 2018-03-14 Chapters/Chapter02.tex
@@
 
-% \subsection{i versus l}
+\section{``i'' versus ``l''}
+
+The classification algorithm had some problems classifying the letters ``i'' and ``l''.
+It may have happened because of the low distance between them.
+In this case, SDM could not differentiate the details --- it has only considered the big picture.
+Although this behavior is close to how humans see things, we also have the ability to zoom and focus in the details, clearly discerning letter ``i'' from letter ``l''.
+
+I have run the classification algorithm under the MNIST database of handwritten digits \citep{deng2012mnist}.
+First, SDM has been trained with the 60,000 training images, and then it classified the 10,000 testing images.
+In these initial tests, the memory has given the correct classification for 77.22\% of the images, which is inferior to the specialized algorithms.
+For instance, in 1998, \citet{lecun1998gradient} have developed algorithms which achieve from 88\% through a linear classifier, to 99.7\% through a convolutional net.
+For a review of algorithms' performance in the MNIST database, see \cite{mnist}.
+
+Looking into the reason behind images incorrectly classified, I have found that the issue is very related to the ``i'' versus ``l'' issue.
+Some handwritten digits are very close to others, and a ``2'' or a ``7'' may look like a ``1'', for instance.
+So, how can we solve this issue without using anything specific to images?
+Machine learning algorithms use specific techniques to improve performance.
+I would like to unveil a solution psychologically closer to how we behave --- even if that eventually leads to lower performance.
+
+An unexplored idea is to use multiple SDMs which communicate.
+A first SDM would write the whole picture, just like we have done.
+Another SDM would write specific parts of the image, just like our eye focusing in specific parts.
+When reading, they may compose the counters and give a more precise classification.
 

### 2018-03-14 Chapters/Chapter02.tex
details

@@
 
-% \subsection{Multiple levels}
-
-% Deep neural networks, alphaGo, alphazero
-
 \section{``i'' versus ``l''}
@@
 
+% \subsection{Multiple levels}
+
+% Deep neural networks, alphaGo, alphazero
+
 \section{Magic numbers}
@@
 
-Each parameter set choice like this will lead to particular numbers --- many of them emergent---, such as the access radius size, critical distance, and so forth.
+Each parameter set choice like this will lead to particular numbers --- many of them emergent---, such as the access radius size, critical distance, memory capacity, and so forth.
 
-One intriguing question here is:  is there a `better' number of dimensions and of hard locations?  If so, can such numbers better studied analitically, or numerically?
+One intriguing question here is:  is there a `better' number of dimensions and of hard locations?  If so, can such numbers be better studied algebrically or numerically?
 
@@
 
-\section{Paving and illuminating the pathway}
+\section{Illuminating and paving a pathway}
 
@@
 The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
-Ralph Waldo Emerson once said \emph{Do not go where the path may lead.
-Go, instead, where there is no path, and leave a trail.}  Dr.
-Kanerva has left the trail.
-It is my job to pave it and illuminate it and to try to deliver an easier pathway for the next generation.
+Ralph Waldo Emerson once said \emph{do not go where the path may lead.
+Go, instead, where there is no path, and leave a trail.}  Pentti Kanerva has left the trail.
+It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.
 Some essays completely shut the door close at the end; this one intends to leave it wide open.

### 2018-03-14 Chapters/sdm-applications.tex
details

@@
 Finally, let's analyse the reading.
-When reading from a noise image with 42\% of noise, the average distance between the noisy image and its clean image is 420, which means they share, on average, 6 hard-locations.
+When reading from a noisy image with 42\% of noise, the average distance between the noisy image and its clean image is 420, which means they share, on average, 6 hard-locations.
 As the average number of activated hard-locations is 1,072, the sum of their counters will be, on average, $Y = 6 \cdot 140 - \sum_{i=1}^{1072-6} X_i$, where $X_i$ is a Bernoulli trial with probability 0.5.
 Hence, $P(\text{black pixel}) = P(Y > 0) = P(6 \cdot 140 - \sum_{i=1}^{1072-6} X_i > 0) = P(\sum_{i=1}^{1066} X_i < 840) = 99.99\%$.
-But, when reading from a noise image with 45\% of noise, the average number of hard-locations shared with the clean image is only 3.
+But, when reading from a noisy image with 45\% of noise, the average number of hard-locations shared with the clean image is only 3.
 Thus the sum of the activated hardlocations' counters will be, on average, $3 \cdot 140 - \sum_{i=1}^{1072-3} X_i$, and $P(\text{black pixel}) = P(\sum_{i=1}^{1069} X_i < 420) = 1.28 \cdot 10^{-12}$.

### 2018-03-15 Chapters/Chapter02.tex
details
SDM Conclusion i vs l

@@
 First, SDM has been trained with the 60,000 training images, and then it classified the 10,000 testing images.
-In these initial tests, the memory has given the correct classification for 77.22\% of the images, which is inferior to the specialized algorithms.
+In these initial tests, the memory has given the correct classification for 79.22\% of the images, which is inferior to the specialized algorithms.
 For instance, in 1998, \citet{lecun1998gradient} have developed algorithms which achieve from 88\% through a linear classifier, to 99.7\% through a convolutional net.
@@
 Ralph Waldo Emerson once said \emph{do not go where the path may lead.
-Go, instead, where there is no path, and leave a trail.}  Pentti Kanerva has left the trail.
+Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.
 It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.

### 2018-03-15 Chapters/Chapter02.tex
Fix hard location (removing hard-location and hardlocation).

@@
 This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present.
-For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where one's fear come from.
+For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where one's fear comes from.
 One just feels it.
@@
 SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\citet{Linhares2011} has showed that SDM respects the limits of short-term memory discussed by \citet{Miller1995} and \citet{Cowan2011}.
+\citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1995} and \citet{Cowan2011}.
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
@@
 As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug free.
-Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `idea space' --- and to focus in the implementation itself.
-Furthermore, new members in our research group had to go through different source codes developed by former members.
+Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `idea space' --- and to focus on the implementation itself.
+Furthermore, new members of our research group had to go through different source codes developed by former members.
 
@@
   $N$ & Size of the binary space, $|\{0, 1\}^n| = 2^n$. \\
-  $N'$ & Number of hard-locations samples from $\{0, 1\}^n$. Its typical value is 1,000,000, as suggested by \citet{Kanerva1988}. \\
+  $N'$ & Number of hard locations samples from $\{0, 1\}^n$. Its typical value is 1,000,000, as suggested by \citet{Kanerva1988}. \\
   $H$ & Same as $N'$. \\
@@
 \chapter{Results (iii): Loss of neurons}
-In SDM, the data is written distributed among millions of hard-locations, which theoretically gives SDM robustness against loss of neurons.
+In SDM, the data is written distributed among millions of hard locations, which theoretically gives SDM robustness against loss of neurons.
 In other words, SDM should keep convering correctly even when some neurons are dead.
@@
 Looking for answers to these questions, we run simulations in which we kept killing some neurons and checking whether SDM remained converging to a given bitstring or not.
-In these simulations, 10,000 random bitstrings were written to a 1,000-bit SDM with 1,000,000 hard-locations, and we choose one of them as our target.
+In these simulations, 10,000 random bitstrings were written to a 1,000-bit SDM with 1,000,000 hard locations, and we choose one of them as our target.
 As the bitstrings were all written exactly once, we may generalize the results.
-The code is available in the ``Reseting hard-locations'' notebook \citep{sdmframework}.
+The code is available in the ``Reseting hard locations'' notebook \citep{sdmframework}.
 
-As neurons are hard-locations in SDM, when we say that a neuron has been killed, we mean that its counters have been zeroed and a new random bitstring address has been assigned.
+As neurons are hard locations in SDM, when we say that a neuron has been killed, we mean that its counters have been zeroed and a new random bitstring address has been assigned.
 During our simulations, no other bitstring has been written after the 10,000.
-Consequently, as their counters willl remain zeroed, it is exactly like ignoring the dead hard-locations in the subsequent reading operations.
+Consequently, as their counters willl remain zeroed, it is exactly like ignoring the dead hard locations in the subsequent reading operations.
 
@@
 
-In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is robust up to 200,000 neuron deaths which is 20\% of all hard-locations.
+In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is robust up to 200,000 neuron deaths which is 20\% of all hard locations.
 Its robustness is astonishing.
@@
 
-We run exactly the same simulation for a 256-bit SDM with 1,000,000 hard-locations.
+We run exactly the same simulation for a 256-bit SDM with 1,000,000 hard locations.
 The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
@@
 Murilo observed that the models of Kanerva's read ($z=1$) and Chada's read ($z=0$) were simple variations of a generalized read with an exponent $z$, which suggests experimenting with different values.
-Mathematically, let $A$ be the set of the counters of the activated hardlocation, and $c_i$ be the counter of the $i$-th bit.
+Mathematically, let $A$ be the set of the counters of the activated hard location, and $c_i$ be the counter of the $i$-th bit.
 Then,
@@
 My advisor, Alexandre Linhares, has proposed another write operation: an information-theoretical weighted write.
-In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address.
-The logic behind it is to vary the importance of each hard-location inside the circle.
+In it, the sum of the counter's value is weighted based on the distance between each hard location's address and the reading address.
+The logic behind it is to vary the importance of each hard location inside the circle.
 It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
@@
 
-But we must consider that not all hardlocations are activated in each write operation, which changes our probability function. Thus, let $r$ be the access radius then:
+But we must consider that not all hard locations are activated in each write operation, which changes our probability function. Thus, let $r$ be the access radius then:
 
@@
 A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
-Is there an algorithm that separates the area of each hard-location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
+Is there an algorithm that separates the area of each hard location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
 Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.

### 2018-03-15 Chapters/sdm-applications.tex
Fix hard location (removing hard-location and hardlocation).

@@
 A simplified mathematical analysis would be: During the learning phase, $200$ images with 15\% of noise were written to SDM, so, the average distance between them and the clean image was 150 bits.
-Thus they shared, on average, 175 hard-locations with the clean image.
-In these 175 hard-locations, the counter's value for a black pixel mapped to bit 1 was $(1-0.15) \cdot 200 - 0.15 \cdot 200 = 140$.
+Thus they shared, on average, 175 hard locations with the clean image.
+In these 175 hard locations, the counter's value for a black pixel mapped to bit 1 was $(1-0.15) \cdot 200 - 0.15 \cdot 200 = 140$.
 Finally, let's analyse the reading.
-When reading from a noisy image with 42\% of noise, the average distance between the noisy image and its clean image is 420, which means they share, on average, 6 hard-locations.
-As the average number of activated hard-locations is 1,072, the sum of their counters will be, on average, $Y = 6 \cdot 140 - \sum_{i=1}^{1072-6} X_i$, where $X_i$ is a Bernoulli trial with probability 0.5.
+When reading from a noisy image with 42\% of noise, the average distance between the noisy image and its clean image is 420, which means they share, on average, 6 hard locations.
+As the average number of activated hard locations is 1,072, the sum of their counters will be, on average, $Y = 6 \cdot 140 - \sum_{i=1}^{1072-6} X_i$, where $X_i$ is a Bernoulli trial with probability 0.5.
 Hence, $P(\text{black pixel}) = P(Y > 0) = P(6 \cdot 140 - \sum_{i=1}^{1072-6} X_i > 0) = P(\sum_{i=1}^{1066} X_i < 840) = 99.99\%$.
-But, when reading from a noisy image with 45\% of noise, the average number of hard-locations shared with the clean image is only 3.
-Thus the sum of the activated hardlocations' counters will be, on average, $3 \cdot 140 - \sum_{i=1}^{1072-3} X_i$, and $P(\text{black pixel}) = P(\sum_{i=1}^{1069} X_i < 420) = 1.28 \cdot 10^{-12}$.
+But, when reading from a noisy image with 45\% of noise, the average number of hard locations shared with the clean image is only 3.
+Thus the sum of the activated hard locations' counters will be, on average, $3 \cdot 140 - \sum_{i=1}^{1072-3} X_i$, and $P(\text{black pixel}) = P(\sum_{i=1}^{1069} X_i < 420) = 1.28 \cdot 10^{-12}$.
 The probability abruptly drops from 100\% to 0\% when the noise goes from 42\% to 45\% (see Figure \ref{fig:filter-prob-right-pixel}).
@@
 
-This is a simplified analysis because it does not take into consideration the hardlocations shared by the different letters.
+This is a simplified analysis because it does not take into consideration the hard locations shared by the different letters.
 It works fine for our example because letter ``I'' and ``T'' are almost orthogonal and share, on average, only one hard location.
@@
 We have confirmed it writing letters ``B'', ``C'', ``D'' and numeral ``8''.
-They share a great amount of hard-locations and our noise filter could not filter their noise correctly.
+They share a great amount of hard locations and our noise filter could not filter their noise correctly.
 The training sets can be seen in Figure \ref{fig:filter-training2} and the results in Figure \ref{fig:filter-results2}.
@@
 The chunck through exclusive OR (XOR) works because of Theorem \ref{thm:xor-chuncks}, which says that chuncking the images with labels will generate, on average, orthogonal bitstrings.
-Thus, these orthogonal bitstrings will not interfere with each other because they share, on average, only one hardlocation.
+Thus, these orthogonal bitstrings will not interfere with each other because they share, on average, only one hard location.
 

### 2018-03-15 Chapters/sdm-architecture.tex
Fix hard location (removing hard-location and hardlocation).

@@
 We split the SDM memory in two parts: the hard-location addresses and the hard-location counters.
-Thus, the addresses (bitstrings) of the hard-locations are stored in one array, while their counters in another.
-This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard-locations will be the same in all of them.
+Thus, the addresses (bitstrings) of the hard locations are stored in one array, while their counters in another.
+This makes possible to create multiple SDMs using the same address space, which would save computational effort to scan a bitstring in all the SDMs --- since they share the same address space, the activated hard locations will be the same in all of them.
 As the slowest part of reading and writing operations is scanning the address space, the performance benefits are significant.
@@
 
-Let a SDM memory with $N$ dimensions and $H$ hard-locations.
+Let a SDM memory with $N$ dimensions and $H$ hard locations.
 Then, in a 64-bit computer, the array of hard-location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard-location counters.
-For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard-locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.
+For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.
 
@@
 
-There is also the naive algorithm, which really counts the number of ones checking bit by bit. It is available only to testing purposes and should never be used.
+There is also the naive algorithm, which really counts the number of ones checking bit by bit. It is available only for testing purposes and should never be used.
 
@@
 
-The scan for activated hard-locations is performed in an address space.
+The scan for activated hard locations is performed in an address space.
 It returns the indexes of the bitstrings which were inside the circle (and their distances).
@@
 
-\subsection{Scanning for activated hard-locations}
+\subsection{Scanning for activated hard locations}
 
-Scanning for the activated hard-locations is a problem similar to well-known problems in computational geometry called ``range reporting in higher dimensions''.
+Scanning for the activated hard locations is a problem similar to well-known problems in computational geometry called ``range reporting in higher dimensions''.
 In this case, none of the known algorithms is able to solve our problem faster than $O(H)$.
@@
 The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(H))$ to go through the whole space.
-But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hard-locations' addresses are randomly sampled from the $\{0, 1\}^n$ space.
-Although each addresses' bit itself splits the hardlocations in half, it does not split the search space in half since both halves still must be convered by the algorithm.
-For instance, let's say we have $n=1,000$ dimensions with $H=1,000,000$ hard-locations, and we are scanning within a circle with radius $r=451$, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
+But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hard locations' addresses are randomly sampled from the $\{0, 1\}^n$ space.
+Although each addresses' bit itself splits the hard locations in half, it does not split the search space in half since both halves still must be covered by the algorithm.
+For instance, let's say we have $n=1,000$ dimensions with $H=1,000,000$ hard locations, and we are scanning within a circle with radius $r=451$, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450.
 Hence, the search space has not been split in half because both halves have been covered (and one of them should have been skipped).
 
-Finally, as our best approach is to scan through all hard-locations, we may distribute the scan into many tasks which will be executed independently.
+Finally, as our best approach is to scan through all hard locations, we may distribute the scan into many tasks which will be executed independently.
 The tasks may be executed in different processes, threads, or even computers.
@@
 
-A counter is an array of integers which stores the data of all hard-locations. So, the counter's array has $n \cdot H$ integers.
+A counter is an array of integers which stores the data of all hard locations. So, the counter's array has $n \cdot H$ integers.
 
 When two counters are added in a third counter, there may occur an overflow.
-It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with less writes than $2^{31}-1$ divided by the average number of activated hard-locations.
-For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.
+It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with less writes than $2^{31}-1$ divided by the average number of activated hard locations.
+For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.
 Note also that it would be more likely to saturate the memory before any overflow.

### 2018-03-15 Chapters/sdm-critical-distance.tex
Fix hard location (removing hard-location and hardlocation).

@@
 \caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
-Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
+Here we had a 1,000 bit, 1,000,000 hard location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
 \label{fig:sdm-10000w-table-7-2}}
@@
 \caption{Results generated by the framework similar from Kanerva's original Table 7.2.
-Here we have a 1,000 bit, 1,000,000 hard-location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
+Here we have a 1,000 bit, 1,000,000 hard location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
 \label{fig:sdm-100w-table-7-2}}
@@
 To do this, we used the following equations from our previous work \citep{brogliato2014sparse}.
-Let $d$ be the distance to the target, $h$ be the number of hard-locations activated during read and write operations, $s$ be the number of total stored bitstrings, $H$ be the number of total hard-locations, $w$ be the number of times the target was written into SDM, $\theta$ be the total random bitstrings in all $h$ hard-locations activated by read operation, and $\phi(d)$ be the average number of shared hard-locations activated two bitstrings $d$ bits away.
+Let $d$ be the distance to the target, $h$ be the number of hard locations activated during read and write operations, $s$ be the number of total stored bitstrings, $H$ be the number of total hard locations, $w$ be the number of times the target was written into SDM, $\theta$ be the total random bitstrings in all $h$ hard locations activated by read operation, and $\phi(d)$ be the average number of shared hard locations activated two bitstrings $d$ bits away.
 
@@
 Although Kanerva has found a formula for $\phi(d)$ through an unsolved integral, and \citet{de1995geometrical} have proposed another way to calculate $\phi(d)$, we have used our framework to estimate the values of $d$.
-In order to do that, we used a Monte Carlo approach, generating many pair of random bitstrings $d$ bits away from them and calculating the average number of shared hard-locations between them.
+In order to do that, we used a Monte Carlo approach, generating many pair of random bitstrings $d$ bits away from them and calculating the average number of shared hard locations between them.
 The code is available in the ``Calculate critical distance'' notebook \citep{sdmframework}.
@@
 
-\section{Critical distance 209}
+\section{Critical distance of 209}
 

### 2018-03-15 Chapters/sdm-performance.tex
Fix hard location (removing hard-location and hardlocation).

@@
 
-Each scan on a 1,000-bits SDM with 1,000,000 hard-locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
+Each scan on a 1,000-bits SDM with 1,000,000 hard locations executes $10^9$ bit compares through $10^9/64 = 15,625,000$ XORs and calls to the built-in popcount.
 So, 10,000 writes execute $10^{13}$ bit compares, while a 6-iterative reading executes $6 \cdot 10^{12}$ bit compares.
@@
 A total of 8 kernels have been developed for benchmarking in our framework: single\_scan0, single\_scan1, single\_scan2, single\_scan3, single\_scan4, single\_scan5, single\_scan5\_unroll, and single\_scan6.
-Each scan uses a different algorithm to do exactly the same thing: calculate which hard-locations are inside the circle of the given bitstring.
+Each scan uses a different algorithm to do exactly the same thing: calculate which hard locations are inside the circle of the given bitstring.
 They differ in how they split the work between work-groups and how they explore the GPU architecture to obtain performance gains.

### 2018-03-15 Chapters/sdm-validation.tex
Fix hard location (removing hard-location and hardlocation).

@@
 
-\section{Number of activated hard-locations}
+\section{Number of activated hard locations}
 
-In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard-locations in a 1,000 bits SDM.
+In his seminal work, Kanerva proposed to use a sample of 1,000,000 hard locations in a 1,000 bits SDM.
 He also proposed to activate only 1,000 of them, on average.
-He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard-locations.
+He calculated that an access radius of $r=451$ would activate, on average, 0.00107185004892 of the whole space, or, in this case, 1,071.85 hard locations.
 
-We extended his results, calculating the distribution of the number of activated hard-locations.
-As each hard-location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard-locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$.
+We extended his results, calculating the distribution of the number of activated hard locations.
+As each hard location has probability $p=0.00107185004892$ of being activated, the probability of activating exactly $a$ out of $H$ hard locations follows a binomial distribution with mean $\mu = pH$ and standard deviation $\sigma = \sqrt{Hp(p-1)}$.
 In this case, $\mu = 1071.85$ and $\sigma = 32.72$.
 
-In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard-locations.
-The code is available in the ``Number of activated hard-locations'' notebook \citep{sdmframework}.
+In order to validate our scan algorithm, we have run 10,000 scans from a random bitstring and counted the number of activated hard locations.
+The code is available in the ``Number of activated hard locations'' notebook \citep{sdmframework}.
 
@@
 
-  \caption{Histogram of the number of activated hard-locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
+  \caption{Histogram of the number of activated hard locations in 10,000 scans from a random bitstring. The curve in red is the theoretical normal distribution with $\mu = Hp$ and $\sigma = p(p-1)H$.}
   \label{fig:validation-activated-hls}
@@
 
-Besides the number of activated hard-locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard-locations.
-Let $A$ be the set of activated hard-locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:
+Besides the number of activated hard locations, we have also extended Kanerva's results to calculate the distribution of distances between the center of the circle and the activated hard locations.
+Let $A$ be the set of activated hard locations, $\xi$ be the center of the circle, and $r$ be the access radius, then:
 
@@
 
-In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard-locations to the center of 1,000 random circles.
-The code is available in the ``Distances of activated hard-locations'' notebook \citep{sdmframework}.
+In order to check Equation \ref{eq:prob-d-inside-circle}, we have calculated the distances of the activated hard locations to the center of 1,000 random circles.
+The code is available in the ``Distances of activated hard locations'' notebook \citep{sdmframework}.
 
@@
 
-\caption{Histogram of the distances of activated hard-locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
+\caption{Histogram of the distances of activated hard locations to the center of the circles. The curve in red is the theoretical distribution of Equation \ref{eq:prob-d-inside-circle}
 \label{fig:validation-distance-activated-hls}}
@@
 When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
-In fact, it converges because they share a sufficient amount of hard-locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge.
+In fact, it converges because they share a sufficient amount of hard locations, i.e., the intersection of the circle around $\eta_d$ and $\eta$ is enough to converge.
 For further information about the relation between the critical distance and the intersection, see \citet{brogliato2014sparse}.
@@
 
-  \caption{Number of hard-locations in the intersection of circles around two bitstrings $x$ bits away.}
+  \caption{Number of hard locations in the intersection of circles around two bitstrings $x$ bits away.}
   \label{fig:validation-intersection}
@@
 
-Each reading operation was performed summing the counters of all activated hard-locations from all three memories.
-For instance, to read the sequence $<A, B, C>$, we have activated the hard-locations around $C$ in $\text{sdm}_1$, we have also activated the hard-locations around $D$ in $\text{sdm}_2$, and, finally, we have also activated the hard-locations around $A$ in $\text{sdm}_3$.
-After summing the counters of all those hard-locations, we evaluate the resulting bitstring just as in the original read operation.
+Each reading operation was performed summing the counters of all activated hard locations from all three memories.
+For instance, to read the sequence $<A, B, C>$, we have activated the hard locations around $C$ in $\text{sdm}_1$, we have also activated the hard locations around $D$ in $\text{sdm}_2$, and, finally, we have also activated the hard locations around $A$ in $\text{sdm}_3$.
+After summing the counters of all those hard locations, we evaluate the resulting bitstring just as in the original read operation.
 

### 2018-03-15 Chapters/sdm2.tex
Fix hard location (removing hard-location and hardlocation).

@@
 So, $\eta$ will be stored in all these activated addresses, which are around address $\xi$, such as in Figure \ref{fig-addresses-inside-access-radius}.
-An address $\xi'$ is inside the circle if its hamming distance to the center $\xi$ is less than or equal to the radius $r$, i.e.
+An address $\xi'$ is inside the circle if its Hamming distance to the center $\xi$ is less than or equal to the radius $r$, i.e.
 $distance(\xi,\xi')\leq r$.
@@
 
-\subfloat[Kanerva's original model]{\includegraphics[width=3.1in]{./images02/new-images/kanerva-read.png}}
+\subfloat[Kanerva's original model]{\includegraphics[width=3.1in]{./images02/new-images/kanerva-read.png}\label{fig:cir-dist-10k-writes-kanerva}}
 
 \subfloat[Chada's read]{\includegraphics[width=3.1in]{./images02/new-images/chada-read_z_0.png}}
-\caption{How far, in hamming distance, is a read item from the original stored item?
+\caption{How far, in Hamming distance, is a read item from the original stored item?
 Kanerva demonstrated that, after a small number of iterative readings (6 here), a critical distance behavior emerges.
@@
 Mathematically, this is the precise distance in which, despite having a relatively high number of cues (correct bits) about the desired item, the time to convergence is infinite.
-Heatmap colors display the hamming distance the associative memory is able to cleanly converge to---or not.
+Heatmap colors display the Hamming distance the associative memory is able to cleanly converge to---or not.
 In the $x$-axis, the distance from the desired item is displayed.

### 2018-03-15 Chapters/sdm2.tex
SDM replace hard-location by hard location

@@
 To solve the feasibility problem of implementing this memory, Kanerva made a random sample of $\{0,1\}^{n}$, in his work, having $N'$ elements.
-All these addresses in the sample are called hard-locations.
+All these addresses in the sample are called hard locations.
 Other elements of $\{0,1\}^{n}$, not in $N'$, are called virtual neurons.
 This is represented in Figure \ref{fig-hardlocations}.
-All properties of read and write operations presented before remain valid, but limited to hard-locations.
-Kanerva suggests taking a sample of about one million hard-locations.
+All properties of read and write operations presented before remain valid, but limited to hard locations.
+Kanerva suggests taking a sample of about one million hard locations.
 
@@
 In fact, only a fraction of this vectorial space is actually instantiated.
-Following Kanerva's suggestion of one million hard-locations, for $n=100$, only $100\cdot10^{6}/2^{100}=7\cdot10^{-23}$ percent of the whole space exists, and for $n=1,000$ only $100\cdot10^{6}/2^{1000}=7\cdot10^{-294}$ percent.
+Following Kanerva's suggestion of one million hard locations, for $n=100$, only $100\cdot10^{6}/2^{100}=7\cdot10^{-23}$ percent of the whole space exists, and for $n=1,000$ only $100\cdot10^{6}/2^{1000}=7\cdot10^{-294}$ percent.
 
-Kanerva also suggests the selection of a radius that will activate, on average, one one thousandth of the sample, which is 1,000 hard-locations for a sample of one million addresses.
+Kanerva also suggests the selection of a radius that will activate, on average, one one thousandth of the sample, which is 1,000 hard locations for a sample of one million addresses.
 In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$.
@@
 
-Since a cue $\eta_{x}$ near the target bitstring $\eta$ shares many hard-locations with $\eta$, SDM can retrieve data from imprecise cues.
+Since a cue $\eta_{x}$ near the target bitstring $\eta$ shares many hard locations with $\eta$, SDM can retrieve data from imprecise cues.
 Despite this feature, it is very important to know how imprecise this cue could be while still giving accurate results.
@@
 The $\{0,1\}^{n}$ space has $N=2^{n}$ locations from which we instantiate $N'$ samples.
-Each location in our sample is called a hard-location.
-On these hard-locations we do operations of read and write.
+Each location in our sample is called a hard location.
+On these hard locations we do operations of read and write.
 One of the insights of SDM is exactly the way we read and write: using data as addresses in a distributed fashion.
-Each datum $\eta$ is written in every activated hard-location inside the access radius centered on the address, that equals datum, $\xi=\eta$.
+Each datum $\eta$ is written in every activated hard location inside the access radius centered on the address, that equals datum, $\xi=\eta$.
 Kanerva suggested using an access radius $r$ having about one one thousandth of $N'$.
-As an imprecise cue $\eta_{x}$ shares hard-locations with the target bitstring $\eta,$ it is possible to retrieve $\eta$ correctly.
+As an imprecise cue $\eta_{x}$ shares hard locations with the target bitstring $\eta,$ it is possible to retrieve $\eta$ correctly.
 (Actually, probably more than one read is necessary to retrieve exactly $\eta.)$.
@@
 One can readily see that the average distance between two random bitstrings has binomial distribution with mean $n/2$ and standard deviation $\sqrt{n/4}$.
-For a large $n$, most of the space lies close to the mean and has fewer shared hard-locations.
+For a large $n$, most of the space lies close to the mean and has fewer shared hard locations.
 As two bitstrings with distance far from $n/2$ are very improbable, \citet{Kanerva1988} defined that two bitstrings are orthogonal when their distance is $n/2$.
@@
 The write operation needs to store, for each dimension bit which happened more ($0$'s or $1$'s).
-This way, each hard-location has $n$ counters, one for each dimension.
+This way, each hard location has $n$ counters, one for each dimension.
 The counter is incremented for each bit $1$ and decremented for each bit $0$.
@@
 
-The read is performed polling each activated hard-location and statistically choosing the most written bit for each dimension.
-It consists of adding all $n$ counters from the activated hard-locations and, for each bit, choosing bit 1 if the counter is positive, choose bit 0 if the counter if negative, and randomly choose bit 0 or 1 if the counter is zero.
+The read is performed polling each activated hard location and statistically choosing the most written bit for each dimension.
+It consists of adding all $n$ counters from the activated hard locations and, for each bit, choosing bit 1 if the counter is positive, choose bit 0 if the counter if negative, and randomly choose bit 0 or 1 if the counter is zero.
 
@@
 One interesting view is that neurons in SDM work like pointers.
-As we write bitstrings in memory, the hard-locations' counters are updated and some bits are flipped.
-Thus, the activated hard-locations do not necessarily point individually to the bitstring that activated it, but together they point correctly.
-In other words, the read operation depends on many hard-locations to be successful.
-This effect is represented in Figure \ref{fig-p1-pointers}: where all hard-locations inside the circle are activated and they, individually, do not point to $\eta$.
+As we write bitstrings in memory, the hard locations' counters are updated and some bits are flipped.
+Thus, the activated hard locations do not necessarily point individually to the bitstring that activated it, but together they point correctly.
+In other words, the read operation depends on many hard locations to be successful.
+This effect is represented in Figure \ref{fig-p1-pointers}: where all hard locations inside the circle are activated and they, individually, do not point to $\eta$.
 But, like vectors, adding them up points to $\eta$.
-If another datum $\nu$ is written into the memory near $\eta$, the shared hard-locations will have information from both of them and would not point to either.
-All hard-locations outside of the circle are also pointing somewhere (possibly other data points).
+If another datum $\nu$ is written into the memory near $\eta$, the shared hard locations will have information from both of them and would not point to either.
+All hard locations outside of the circle are also pointing somewhere (possibly other data points).
 This is not shown, however, in order to keep the picture clean and easily understandable.
@@
 
-Due to the distribution of hard-locations between two random bitstrings, the vast majority of concepts is orthogonal to all others.
+Due to the distribution of hard locations between two random bitstrings, the vast majority of concepts is orthogonal to all others.
 Consider a non-scientific survey during a cognitive science seminar, where students asked to mention ideas unrelated to the course brought up terms like birthdays, boots, dinosaurs, fever, executive order, x-rays, and so on.
@@
 In his work, Kanerva proposed and analyzed a read algorithm called here Kanerva's read.
-His read takes all activated hard-locations counters and sum them.
+His read takes all activated hard locations counters and sum them.
 The resulting bitstring has bit $1$ where the result is positive, bit $0$ where the result is negative, and a random bit where the result is zero.
-In a word, each bit is chosen according to all written bitstrings in all hard-locations, being equal to the bit more appeared.
+In a word, each bit is chosen according to all written bitstrings in all hard locations, being equal to the bit more appeared.
 Table \ref{tab:kanerva-read} shows an example of Kanerva's read result bitstring.
@@
 Daniel Chada, one member of our research group, proposed another way to read in SDM, in this work called Chada's read.
-Instead of summing all hard-location counters, each hard-location evaluates its resulting bitstring individually.
+Instead of summing all hard location counters, each hard location evaluates its resulting bitstring individually.
 Then, all resulting bitstrings are summed again, and the same rule as Kanerva applies.
@@
 
-The main difference between Kanerva's read and Chada's is that, in the former, a hard-location that has more bitstrings written has a greater weight in the decision of each bit.
-In the latter, all hard-locations have the same weight, because they can contribute to the sum with only one bitstring.
+The main difference between Kanerva's read and Chada's is that, in the former, a hard location that has more bitstrings written has a greater weight in the decision of each bit.
+In the latter, all hard locations have the same weight, because they can contribute to the sum with only one bitstring.
 
@@
 Paulo Murilo\footnote{Universidade Federal Fluminense's Physics Professor Paulo Murilo}, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries.
-He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative).
+He proposed summing all hard location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative).
 Thus, Kanerva's read would be the same as $z=1$, while Chada's would be the same as $z=0$.
@@
 \end{minipage}\caption{Comparison of Kanerva's read and Chada's read. Each $\xi_{i}$ is
-an activated hard-location and the values come from their counters.
+an activated hard location and the values come from their counters.
 Gray cells' value is obtained randomly with probability 50\%.\label{tab:read-operation}}
@@
 
-In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 1 million hard-locations $N'=1,000,000$, an access-radius of 451 (within 1,000 hard-locations in each circle) and 10 thousand writes of random bitstrings in the memory.
+In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 1 million hard locations $N'=1,000,000$, an access-radius of 451 (within 1,000 hard locations in each circle) and 10 thousand writes of random bitstrings in the memory.
 As computer resources were very poor those days, Kanerva couldn't make a more generic analysis.

### 2018-03-15 Chapters/sdm-architecture.tex
SDM replace hard-location by hard location.

@@
 
-We split the SDM memory in two parts: the hard-location addresses and the hard-location counters.
+We split the SDM memory in two parts: the hard location addresses and the hard location counters.
 Thus, the addresses (bitstrings) of the hard locations are stored in one array, while their counters in another.
@@
 Let a SDM memory with $N$ dimensions and $H$ hard locations.
-Then, in a 64-bit computer, the array of hard-location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard-location counters.
+Then, in a 64-bit computer, the array of hard location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard location counters.
 For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.
@@
 
-An address space is a fixed collection of bitstrings, and each bitstring represents a hard-location address.
+An address space is a fixed collection of bitstrings, and each bitstring represents a hard location address.
 They store the number of bitstrings, as well as the number of bits, number of integers per bitstring, and the number of remaining bits.
@@
 
-Each hard-location has one integer of data per bit. For instance, each hard-location of a 1,000 bits SDM has 1,000 bits. Those integers are stored in a counter.
+Each hard location has one integer of data per bit. For instance, each hard location of a 1,000 bits SDM has 1,000 bits. Those integers are stored in a counter.
 
@@
 There is another special read operation: the weighted reading.
-In the weighted reading, the value of the counters are multiplied by a weight which depends only on the distance between the reading address and the hard-location address.
+In the weighted reading, the value of the counters are multiplied by a weight which depends only on the distance between the reading address and the hard location address.
 The weight is retrieved from a lookup table of integers indexed by the distance.
@@
 In this case, the weight is applied when the counters are updated, i.e., if the weight is 2, the counters are increased twice when bits are 1, and decreased twice when bits are 0.
-Just as in the weighted reading, the weights depend only on the distance between the writing address and the hard-location address.
+Just as in the weighted reading, the weights depend only on the distance between the writing address and the hard location address.
 The weights are retrieved from a lookup table of integers indexed by the distance.

### 2018-03-15 Chapters/Chapter02.tex
SDM grammar corrections

@@
 In SDM, the data is written distributed among millions of hard locations, which theoretically gives SDM robustness against loss of neurons.
-In other words, SDM should keep convering correctly even when some neurons are dead.
+In other words, SDM should keep converging correctly even when some neurons are dead.
 The question is: how robust it really is?
-How many neurons may die before it starts to forget things.
-These questions have never been addresses before.
+How many neurons may die before it starts to forget things?
+These questions have never been addressed before.
 
@@
 As the bitstrings were all written exactly once, we may generalize the results.
-The code is available in the ``Reseting hard locations'' notebook \citep{sdmframework}.
+The code is available in the ``Resetting hard locations'' notebook \citep{sdmframework}.
 
@@
 During our simulations, no other bitstring has been written after the 10,000.
-Consequently, as their counters willl remain zeroed, it is exactly like ignoring the dead hard locations in the subsequent reading operations.
+Consequently, as their counters will remain zero, it is exactly like ignoring the dead hard locations in the subsequent reading operations.
 
@@
 
-In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is robust up to 200,000 neuron deaths which is 20\% of all hard locations.
+In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is robust up to 200,000 neuron deaths which are 20\% of all hard locations.
 Its robustness is astonishing.
-In fact, SDM begins to be significantly affected by loss of neurons after 600,000 neuron deaths (Figure \ref{fig:sdm-neuron-death-1m}), and obviously forgets everything when all neurons are dead.
+In fact, SDM begins to be significantly affected by the loss of neurons after 600,000 neuron deaths (Figure \ref{fig:sdm-neuron-death-1m}) and obviously forgets everything when all neurons are dead.
 
@@
 
-It is interesting that 500,000 neuron deaths has a minor effect in SDM's recall capability (see Figure \ref{fig:sdm-neuron-death-500k}).
-It is analogous to do an hemispherectomy in a person and, after the procedure, the person being able to recall and learn almost just like before.
+It is interesting that 500,000 neuron deaths have a minor effect on SDM's recall capability (see Figure \ref{fig:sdm-neuron-death-500k}).
+It is analogous to do a hemispherectomy in a person and, after the procedure, the person being able to recall and learn almost just like before.
 In fact, there are clinical reports of children submitted to hemispherectomy who live an almost normal life with minor function problems.
@@
 An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance becomes small, i.e., SDM recall capacity is very diminished.
-After 900,000 neuron deaths the critical distance is zero.
+After 900,000 neuron deaths, the critical distance is zero.
 In this case, everything has been lost.
@@
 Murilo observed that the models of Kanerva's read ($z=1$) and Chada's read ($z=0$) were simple variations of a generalized read with an exponent $z$, which suggests experimenting with different values.
-Mathematically, let $A$ be the set of the counters of the activated hard location, and $c_i$ be the counter of the $i$-th bit.
+Mathematically, let $A$ be the set of the counters of the activated hard location, and $c_i$ be the counter of the $i$th bit.
 Then,
@@
 
-The sum of $|c_i|^z$ turns the intermediate values from integers to float point numbers. Thus, we have developed a specific read operation which stored the intermediate values in double variables.
+The sum of $|c_i|^z$ turns the intermediate values from integers to floating point numbers. Thus, we have developed a specific read operation which stored the intermediate values in double variables.
 
@@
 
-We understand that the critical distant is an important parameter of SDM.
+We understand that the critical distance is an important parameter of SDM.
 The bigger the critical distance, the best, because SDM is able to converge even with farther clues.
@@
 It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
-Closer hard locations have lower probabilies and therefore should encode more information.
+Closer hard locations have lower probabilities and therefore should encode more information.
 
@@
 Consider the following.
-Information Theory \citep{cover2012elements} let us compute the precise amount of information in an event, when given its probability $p$, through the measure of \emph{self-information}:
+Information Theory \citep{cover2012elements} let us compute the precise amount of information in an event when given its probability $p$, through the measure of \emph{self-information}:
 
@@
 
-We provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypotheses and test them computationally through minimal effort.
+We provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypotheses and test them computationally with minimal effort.
 The framework is well-documented for public release at this time (http://sdm-framework.readthedocs.io), it has already served as the backbone of Chada's Ph.D.
@@
 In this case, SDM could not differentiate the details --- it has only considered the big picture.
-Although this behavior is close to how humans see things, we also have the ability to zoom and focus in the details, clearly discerning letter ``i'' from letter ``l''.
+Although this behavior is close to how humans see things, we also have the ability to zoom and focus on the details, clearly discerning letter ``i'' from letter ``l''.
 
@@
 A first SDM would write the whole picture, just like we have done.
-Another SDM would write specific parts of the image, just like our eye focusing in specific parts.
+Another SDM would write specific parts of the image, just like our eye focusing on specific parts.
 When reading, they may compose the counters and give a more precise classification.
@@
 
-One intriguing question here is:  is there a `better' number of dimensions and of hard locations?  If so, can such numbers be better studied algebrically or numerically?
+One intriguing question here is:  is there a `better' number of dimensions and of hard locations?  If so, can such numbers be better studied algebraically or numerically?
 
@@
 We want to select a subset of its vertices with cardinality $2^{20}$ that is symmetrically distributed over the space.
-Afterwards, $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
+Afterward, $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
 
@@
 Consider having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).
-For example, some of the benchmarks that we plan to develop in future research is: how fast is convergence through iterative reading?
+For example, some of the benchmarks that we plan to develop in future research are: how fast is convergence through iterative reading?
 How large is the attractor of the critical distance?
@@
 Indeed, a high-quality codebase seems to have become a scientific community's form of unequivocally standing behind a consensus.
-For example, the journal Nature analysed the top-100 cited papers in history, to find:
+For example, the journal Nature analyzed the top-100 cited papers in history, to find:
 
@@
 The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.
-The emphasis might have been on \emph{breadth of topics}, in detriment of depth here or there.
-But this is due to our research group's enthusiam for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.
+The emphasis might have been on the \emph{breadth of topics}, in detriment of depth here or there.
+But this is due to our research group's enthusiasm for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.
 If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.
@@
 Some essays completely shut the door close at the end; this one intends to leave it wide open.
-As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desiderata}, a prologue, a yearning for others to join me in imagining the shape of things to come.
+As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.
 

### 2018-03-15 Chapters/sdm-applications.tex
SDM grammar corrections

@@
 
-Supervised classification problem consists of categorize data into groups after seeing some samples from each group.
+Supervised classification problem consists of categorizing data into groups after seeing some samples from each group.
 First, it is presented pieces of data with their categories.
-The algorithm learns from these data, which is known as learning phase.
+The algorithm learns from these data, which is known as the learning phase.
 Then, new pieces of data are presented and the algorithm must classify them into the already known groups.
@@
 The images had 31 pixels of width and 32 pixels of height, totaling 992 pixels per image.
-Each image was mapped into a 1,000 bit bitstring in which the bits were set according to the color of each pixel of the image.
+Each image was mapped into a 1,000-bit bitstring in which the bits were set according to the color of each pixel of the image.
 So, white pixels were equal to bit 0, and black pixels to bit 1.
@@
 The association of images to groups was stored as sequences in SDM, as detailed by \citet{Kanerva1988} in Chapter 8.
-During the learning phase, the image bitstrings were stored pointing to their groups bitstrings, i.e., write(addr=bs\_image, datum=bs\_label).
+During the learning phase, the image bitstrings were stored pointing to their group's bitstrings, i.e., write(addr=bs\_image, datum=bs\_label).
 Thus, in order to classify an unknown image, we only had to read from its address and check which group has been found.
@@
 For example, see the generated images for letter A in Figure \ref{fig-classification-training-A}.
-Then, we have wrote the classification group bitstring into the bitstring associated to each noisy image, i.e., write(bs\_image, bs\_label).
+Then, we have written the classification group bitstring into the bitstring associated to each noisy image, i.e., write(bs\_image, bs\_label).
 For a complete image training set, see Appendix XYZ.
@@
 
-Finally, we have assess the performance of our classifier.
+Finally, we have assessed the performance of our classifier.
 We had done it in three different scenarios: high noise (20\%), low noise (5\%) and no noise.
@@
 The performance was calculated as the percentage of hits for each group.
-We did not expected the same performance for all groups because some groups become very similar to other depending on the noise level, and this similarity may even confuse a person (see Figure \ref{fig-classification-similarity}).
+We did not expect the same performance for all groups because some groups become very similar to other depending on the noise level, and this similarity may even confuse a person (see Figure \ref{fig-classification-similarity}).
 
@@
 
-In the no noise scenario, the classifier has hit all characters, except letter ``l'' which was wrongly associated to the group of ``i''.
+In the no noise scenario, the classifier has hit all characters, except letter ``l'' which was wrongly associated with the group of ``i''.
 We believe that it happened because the classifier had never seen an image with no noise and the difference between the images of ``l'' and ``i'' is smaller than the critical distance.
@@
 The high noise scenario is the most interesting, because, even in a high noise level, the classifier has hit most of the characters.
-It has hit all images for 44 out of 62 groups, and made at least one miss for the other 18 groups.
+It has hit all images for 44 out of 62 groups and made at least one miss for the other 18 groups.
 The misses may be seen in details in Figure \ref{fig-classificiation-high-noise-misses}.
@@
 The critical distance plays an important role in the classification error.
-As we have 62 groups and each have been trained with 100 images, there were 6,200 writes to the memory.
+As we have 62 groups and each has been trained with 100 images, there were 6,200 writes to the memory.
 When an image is being classified, it will have to converge to a group, and the convergence depends on the distance between this image and the images from the training set, i.e, in the noise level.
@@
 These results show that the SDM may be used as a supervised classification algorithm.
-Although we do not believe that the mapping between images and bitstrings are even close to the way human cognition deals with images, we believe the results are interesting and useful to many possible real world problems.
+Although we do not believe that the mapping between images and bitstrings are even close to the way human cognition deals with images, we believe the results are interesting and useful to many possible real-world problems.
 
@@
 
-Image noise filtering consists in removing the noise from an input, in out case an image.
+Image noise filtering consists in removing the noise from an input, in our case an image.
 Our images are black \& white images and the noise is generated randomly flipping some of their pixels from black to while and vice versa.
@@
 The images have 30 x 30 pixels, totaling 900 pixels per image.
-Each image is mapped into a bistrings of 1,000 bit in which the bits are set according to the color of each pixel of the image.
+Each image is mapped into a bistring of 1,000 bit in which the bits are set according to the color of each pixel of the image.
 White pixels are mapped to bit 0, and black pixels to bit 1.
@@
 We were able to clean images up to 42\% of noise (see Figure \ref{fig:filter-testing}).
-While SDM has never seen a clean version of the letters, it just learned from the learning phase which pixels has appeared more frequently and chose them.
+While SDM has never seen a clean version of the letters, it just learned from the learning phase which pixels have appeared more frequently and choose them.
 
@@
 In these 175 hard locations, the counter's value for a black pixel mapped to bit 1 was $(1-0.15) \cdot 200 - 0.15 \cdot 200 = 140$.
-Finally, let's analyse the reading.
+Finally, let's analyze the reading.
 When reading from a noisy image with 42\% of noise, the average distance between the noisy image and its clean image is 420, which means they share, on average, 6 hard locations.
@@
 This is a simplified analysis because it does not take into consideration the hard locations shared by the different letters.
-It works fine for our example because letter ``I'' and ``T'' are almost orthogonal and share, on average, only one hard location.
+It works fine for our example because letters ``I'' and ``T'' are almost orthogonal and share, on average, only one hard location.
 
@@
 If the intersection between images becomes too high, the noise filter stops working properly.
-We have confirmed it writing letters ``B'', ``C'', ``D'' and numeral ``8''.
-They share a great amount of hard locations and our noise filter could not filter their noise correctly.
+We have confirmed it writing the letters ``B'', ``C'', ``D'' and the numeral ``8''.
+They share a great number of hard locations and our noise filter could not filter their noise correctly.
 The training sets can be seen in Figure \ref{fig:filter-training2} and the results in Figure \ref{fig:filter-results2}.
@@
 A possible solution to this interference problem is to use labels.
-Each label has a random bitstring, which will be chuncked with the images before writing into SDM.
-Hence, before reading, we also have to chunck the image with the label --- which means we need to know the label of each image.
+Each label has a random bitstring, which will be chunked with the images before writing into SDM.
+Hence, before reading, we also have to chunk the image with the label --- which means we need to know the label of each image.
 The chunk was done using the exclusive OR (XOR) operator, i.e., $bs\_chunck = bs\_image \oplus bs\_label$.
@@
 
-The chunck through exclusive OR (XOR) works because of Theorem \ref{thm:xor-chuncks}, which says that chuncking the images with labels will generate, on average, orthogonal bitstrings.
+The chunk through exclusive OR (XOR) works because of Theorem \ref{thm:xor-chuncks}, which says that chunking the images with labels will generate, on average, orthogonal bitstrings.
 Thus, these orthogonal bitstrings will not interfere with each other because they share, on average, only one hard location.
@@
 What is incredible about these victories is that AlphaZero has almost no knowledge about chess game and has learned all its movement playing against itself for 4 hours.
-Basically, it knowns only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm.
+Basically, it knows only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm.
 
@@
 Basically, it receives the current board state and returns which action should be played.
-In the end of the game, it receives the sequences of boards and the winner, and is supposed to learn from them.
+At the end of the game, it receives both the sequences of boards and the winner and is supposed to learn from them.
 
@@
 
-Our algorithm to decide what should be played is very simple: it reads the current board from SDM and then it chooses the valid action with highest score.
+Our algorithm to decide what should be played is very simple: it reads the current board from SDM and then it chooses the valid action with the highest score.
 To calculate the scores, the bitstring is split into 9 parts, one per action (Figure \ref{fig:ttt-action-mapping}).
@@
 Our algorithm has three rewards: positive, negative, and neutral.
-It always learn from both players, no matter who wins or if it was a draw.
-The winner's sequence of actions feed our positive reward learning.
-The loser's sequence of actions feed our negative reward learning.
+It always learns from both players, no matter who wins or if it was a draw.
+The winner's sequence of actions feeds our positive reward learning.
+The loser's sequence of actions feeds our negative reward learning.
 If it is a draw, both sequences feed our neutral reward learning.
@@
 For instance, each board in Figure \ref{fig-ttt-example} has its own unique random bitstring.
-Thus, if a specific board state is reached again, SDM will return the scores of each action.
+Thus, if a specific board state is reached again, SDM will return the scores for each action.
 
@@
 The neutral reward will have, on average, half of its bits 1 and the other half 0.
-Thus, it is in the middle between a positive reward and a negative reward.
+Thus, it is in the middle of a positive reward and a negative reward.
 
@@
 
-There is also weights associated to positive, negative, and neutral reward.
+There are also weights associated with positive, negative, and neutral rewards.
 They are used to indicate what goal is more important.
@@
 Internally, every board is mapped into a random bitstring and passed to SDM.
-As every two boards are, on average, orthogonal, SDM knows nothing about the boards themselves, neither whether they are consecutives or not.
-It knowns only the score of the actions according to the games it has seen and learned from.
+As every two boards are, on average, orthogonal, SDM knows nothing about the boards themselves, neither whether they are consecutive or not.
+It knows only the score of the actions according to the games it has seen and learned from.
 Hopefully, the actions will lead to a victory or a draw.
@@
 
-When playing against smart player, SDM has started learning how to not lose during the first training cycles.
+When playing against the smart player, SDM has started learning how to not lose during the first training cycles.
 The drawing rate grows quickly in the first 500 games and slowly since then.
@@
 When playing against mixed players, SDM has also adapted.
-In Figure \ref{fig:ttt-results-mixed}, it has played 6,000 games against random player, smart player, and another SDM player.
+In Figure \ref{fig:ttt-results-mixed}, it has played 6,000 games against the random player, the smart player, and another SDM player.
 In each cycle, one of them was randomly chosen.
 The number of losses over time is decreasing, whereas the number of wins and draws change a lot.
-It is easy to notice when the other SDM player was chosen, because all games in that cycle have draw.
+It is easy to notice when the other SDM player was chosen since all games in that cycle have drawn.
 

### 2018-03-15 Chapters/sdm-architecture.tex
SDM grammar corrections

@@
 
-\subsection{OpenCL kernel}
+\subsection{OpenCL kernels}
 
 There are 8 OpenCL kernels which explore differently the GPU architecture to improve performance.
-It is necessary because there are several GPU microarchitecture and a single kernel will never be optimal for all of them.
+It is necessary because there are several GPU microarchitectures and a single kernel will never be optimal for all of them.
 In simplified form, OpenCL splits the tasks into workgroups which, in turn, split their part of the task into workers.
@@
 Then, a reduction algorithm is used to sum the partial distances array in order to calculate the total distance.
-This reduction algorithm is also distributed between the workers and runs in $O(\log_2(\text{bs\_step}))$.
+This reduction algorithm is also distributed among the workers and runs in $O(\log_2(\text{bs\_step}))$.
 Finally, the first worker of each workgroup checks whether the distance is less than or equal to the radius to include the bitstring index into the resulting array.
 
-Some of the optimizations may not work in some GPUs, because not all their premises are valid. Before choosing a kernel, one should check whether it works property for one's specific GPU device.
+Some of the optimizations may not work in some GPUs because not all their premises are valid. Before choosing a kernel, one should check whether it works property for one's specific GPU device.
 
@@
 When two counters are added in a third counter, there may occur an overflow.
-It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with less writes than $2^{31}-1$ divided by the average number of activated hard locations.
+It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with fewer writes than $2^{31}-1$ divided by the average number of activated hard locations.
 For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.
@@
 
-The iterated reading keeps reading until it gets exactly the same bitstring (or the number of maximum interations has been reached), i.e., it performs $\eta_{i+1} = \text{read}(\eta_i)$ and stops when $\eta_{k+1} = \eta_{k}$.
+The iterated reading keeps reading until it gets exactly the same bitstring (or the number of maximum interactions has been reached), i.e., it performs $\eta_{i+1} = \text{read}(\eta_i)$ and stops when $\eta_{k+1} = \eta_{k}$.
 If the initial bitstring is inside the critical distance of $\eta$, it will converge to $\eta$, but, if it is not, it will diverge and reach the maximum number of iterations.
@@
 The generalization brings a parameter $z$, which is the exponent.
-In this case, the results are floating point instead of integer, which considerably reduces performance.
+In this case, the results are floating points instead of integers, which considerably reduces performance.
 When $z=1$, it is exactly as the Kanerva's read.
@@
 There is another special read operation: the weighted reading.
-In the weighted reading, the value of the counters are multiplied by a weight which depends only on the distance between the reading address and the hard location address.
+In the weighted reading, the value of the counters is multiplied by a weight which depends only on the distance between the reading address and the hard location address.
 The weight is retrieved from a lookup table of integers indexed by the distance.

### 2018-03-15 Chapters/sdm-critical-distance.tex
SDM grammar corrections

@@
 \citet[p.70]{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Figure \ref{fig:kanerva-figure-7.3}).
-The original prediction considered that the read distance would decline when inside the critical distance and increase afterwards, converging to a \textasciitilde 500-bit distance.
+The original prediction considered that the read distance would decline when inside the critical distance and increase afterward, converging to a \textasciitilde 500-bit distance.
 At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance bitstring.
@@
 
-As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
+As we ran the simulations, this one, in particular, struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
 We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.
@@
 This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
-In Figure \ref{fig:sdm-table-7-2-steps} we can notice that, the more random bitstrings are written, the stronger the attractors.
+In Figure \ref{fig:sdm-table-7-2-steps} we can notice that the more random bitstrings are written, the stronger the attractors.
 
@@
 As we can easily notice, we got exactly the same curve as Kanerva.
-This question has been intriguing us since then and we are still looking for a more analytic explanation then merely interference from the other written attractors.
+This question has been intriguing us since then and we are still looking for a more analytic explanation than merely interference from the other written attractors.
 

### 2018-03-15 Chapters/sdm-performance.tex
SDM grammar corrections

@@
 The number of people who have ever lived on earth is estimated to be $1,08 \cdot 10^{11}$.
-There are approximtely $1.8 \cdot 10^{9}$ websites in the internet.
+There are approximately $1.8 \cdot 10^{9}$ websites on the internet.
 A modern laptop can increment a counter $5 \cdot 10^{8}$ times per second.
@@
 
-We have created a benchmark to be able to compare the performance in different devices.
+We have created a benchmark to be able to compare the performance on different devices.
 So, the same performance test was executable in our devices, with results reported in tables and figures.
@@
 Our first device was a personal MacBook Pro Retina 13-inch Late 2013 with a 2.6GHz Intel core i5 processor, 6GB DDR3 RAM, and Intel Iris GPU.
-It was not possible to run the 10,000-bits in this device because there was no memory available --- it needs 37.25 GB of memory.
+It was not possible to run the 10,000-bits on this device because there was no memory available --- it needs 37.25 GB of memory.
 For its results, see Table \ref{tab:perf-macbook}
@@
 
-OpenCL is a framework for writting software that executes in \emph{heterogeneous} devices \citep{munshi2009opencl}, like CPUs, GPUs, FPGA and other co-processor for hardware acceleration.
+OpenCL is a framework for writing software that executes in \emph{heterogeneous} devices \citep{munshi2009opencl}, like CPUs, GPUs, FPGA and other co-processors for hardware acceleration.
 Because they are heterogeneous, they may differ a lot in architecture and performance, which means there is no one-size-fits-all kernel.
@@
 
-The difference in which kernel is the best depends also in the SDM setup.
+The difference in which kernel is the best depends also on the SDM setup.
 The best kernel for the 1,000-bits SDM in the iMac 2017 was single\_scan5\_unroll with average scan time of 3.61ms; but, for the 256-bits SDM in the same device, it was single\_scan0 with average scan time of 3.03ms; while, for the 10,000-bits SDM in the same device, it was single\_scan6 with 10.96ms (see Table \ref{tab:perf-imac}).
@@
 Again, which one is faster depends on the SDM settings.
-In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL scanner with , but for a 256-bit SDM was the threads scanner.
+In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL scanner with \lstinline{single_scan5_unroll}, but for a 256-bit SDM was the threads scanner.
 

### 2018-03-15 Chapters/sdm-validation.tex
SDM grammar corrections

@@
 In figure \ref{fig:validation-activated-hls}, we can notice that the theoretical model and the simulation matches.
-Hence, it seems that both the address space generation algotihm and the scan algorithm work properly.
+Hence, it seems that both the address space generation algorithm and the scan algorithm work properly.
 Notice that the curve is almost the same for $n=1,000$ and $n=256$.
-It happens because the access radius is adjust to have $p$ as close as possible to $0.001$.
+It happens because the access radius is adjusted to get $p$ as close as possible to $0.001$.
 They are not exactly the same because their $p$ differs a little.
@@
 Kanerva has calculated the intersection of two circles according to the distance between their centers.
-The intersection is important to understand how SDM works, because it affects directly the critical distance.
+The intersection is important to understand how SDM works because it affects directly the critical distance.
 When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
@@
 
-We have calculated the intersection between a random bitstring (bs1) and another bitstrings (bs2) exactly $d$ bits away.
+We have calculated the intersection between a random bitstring (bs1) and another bitstring (bs2) exactly $d$ bits away.
 The former (bs1) is just a random bitstring.
@@
 
-It worked as if we had splitted SDM into $k$ regions with low intersection between two of them.
+It worked as if we had split SDM into $k$ regions with low intersection between two of them.
 So, as the interference is minimal, they work like independent SDMs.

### 2018-03-16 Chapters/Chapter02.tex
SDM Minor fixes (and add partial-bass.tex).

@@
 SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1995} and \citet{Cowan2011}.
+\citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2011}.
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
-Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate, effort from researchers to build on top of previous work.
+Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate effort from researchers to build on top of previous work.
 
-It is our belief that a tool such as a framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as python, in an intuitive platform such as juypyter notebooks.
+It is our belief that a tool such as a framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
@@
 As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug free.
-Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `idea space' --- and to focus on the implementation itself.
+Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
 Furthermore, new members of our research group had to go through different source codes developed by former members.

### 2018-03-16 Chapters/Chapter02.tex
SDM fix grammar issues and typos

@@
 Our motivation was our own effort to run our models.
-As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug free.
+As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug-free.
 Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
@@
 Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier for readers to replicate any previous results and easily understand the source-code of the model.
-Moreover, it is compatible with jupyter notebook and researchers may share their notebooks possibly accelerating the advances in their fields \citep{shen2014interactive}.
+Moreover, it is compatible with Jupyter notebook and researchers may share their notebooks possibly accelerating the advances in their fields \citep{shen2014interactive}.
 
@@
   $H$ & Same as $N'$. \\
-  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, its typical value is $451$. This value is calculated to activate, on average, one thousandth of $N'$. \\
+  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, its typical value is $451$. This value is calculated to activate, on average, one-thousandth of $N'$. \\
   $\eta$ & A bitstring, usually a datum. \\
@@
 It seems that Kanerva may not have considered that a single read would only `clean' a small number of dimensions \emph{after the critical distance}.
-What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
+What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthogonal distance 500 after a number of iterative readings.}
   \label{fig:murillo-generalization-experiments}
@@
   \caption{(a) and (b) show the behavior of Figure \ref{fig:murillo-generalization-experiments}, now executed with 6-iterative reads.
-What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
+What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthogonal distance 500 after a number of iterative readings.}
   \label{fig:murillo-generalization-experiments-6reads}
@@
 This Shannon write may affect memory capacity --- possibly increasing it.
-Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.
+Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, the shape of the attractor may have asymmetries.
 Whereas these are just some initial tests, the idea seems meritorious so far.
@@
 
-Consider the misunderstanding concerning the SDM read operation:  Dr Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
+Consider the misunderstanding concerning the SDM read operation:  Dr.
+Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
 We have referred to this modified read operation as Chada read\footnote{Legend has it that my friend \text{\&} colleague, Dr.
-Daniel de Magalhães Chada, along with Linhares, did not consult and re-check with Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterwards.}.
+Daniel de Magalhães Chada, along with Linhares, did not consult and re-check with Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterward.}.
 Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
@@
 
-\chapter*{List of jupyter notebooks}
+\chapter*{List of Jupyter notebooks}

### 2018-03-16 Chapters/sdm-applications.tex
SDM fix grammar issues and typos

@@
 \centering\includegraphics[width=\textwidth]{./images02/classification/example.png}
-\caption{Examples of noisy images with uppercase letters, lowercase letters and numbers.
+\caption{Examples of noisy images with uppercase letters, lowercase letters, and numbers.
 \label{fig-classification-examples}}
@@
   \subfloat[Images from character ``l'' which were classified as {[i, i, i, i, i, i, i, i, i, i]}, respectively.
-It has missed them all, as if both groups have been merged.
+It has missed them all as if both groups have been merged.
 ]{\includegraphics[width=\textwidth]{./images02/classification/low-noise-l.png}}
@@
 The images have 30 x 30 pixels, totaling 900 pixels per image.
-Each image is mapped into a bistring of 1,000 bit in which the bits are set according to the color of each pixel of the image.
+Each image is mapped into a bitstring of 1,000 bit in which the bits are set according to the color of each pixel of the image.
 White pixels are mapped to bit 0, and black pixels to bit 1.
@@
 All cases have 10\% noise.
-We can notice that the empty space in the right side of the ``C'' letter generates some white pixels in the right side of both ``B'' and ``D'' letters.
+We can notice that the empty space on the right side of the ``C'' letter generates some white pixels on the right side of both ``B'' and ``D'' letters.
   \label{fig:filter-results2}
@@
 The winning rate grew to around 20\% and remained there until 6,000 games, while the drawing rate kept growing.
-Then, after learning how to not lose, SDM started learning how to win, since after 7,000 games, the winning rate started to increase.
+Then, after learning how to not lose, SDM started learning how to win, since, after 7,000 games, the winning rate started to increase.
 See Figure \ref{fig:ttt-results-smart}.
@@
 
-\caption{Results playing against random player. Each cyle was made of 100 games for training, and then 100 games for measuring statistics.
+\caption{Results playing against the random player. Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
 \label{fig:ttt-results-random}
@@
 
-\caption{Results playing against smart player. Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
+\caption{Results playing against the smart player. Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
 \label{fig:ttt-results-smart}

### 2018-03-16 Chapters/sdm-architecture.tex
SDM fix grammar issues and typos

@@
 
-We split the SDM memory in two parts: the hard location addresses and the hard location counters.
+We split the SDM memory into two parts: the hard location addresses and the hard location counters.
 Thus, the addresses (bitstrings) of the hard locations are stored in one array, while their counters in another.
@@
 
-Basic operations were grouped in four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs).
+Basic operations were grouped into four sets: (i) for bitstrings, (ii) for addresses, (iii) for counters, and (iv) for memories (SDMs).
 Operations include creating new bitstrings, flipping bits, generating a bitstring with a specific distance from a given bitstring, scanning the address space using different algorithms, writing a bitstring to a counter, writing in an SDM, reading from an SDM, and iteratively reading from an SDM until convergence.
@@
 
-The distance between two bitstrings is calculated by the hamming distance, which is the number of different bits between them.
+The distance between two bitstrings is calculated by the Hamming distance, which is the number of different bits between them.
 It is calculated counting the number of ones in the exclusive or (xor) between the bitstrings, i.e., $d(x, y) = \text{number of ones in } x \oplus y$.
@@
 It pre-calculates a table with the number of ones of all possible 16-bit integers.
-This table is accessed a few times to calculate the number of ones of a 64-bit integer, i.e., to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:63], y[0:63]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + d(x[32:47], y[32:47]) + d(x[48:63], y[48:63])$ where $x[i:i+15]$ and $y[i, i+15]$ are the 16-bit integers formed by the bits between $i$ and $i+15$ of $x$ and $y$, respectivelly.
+This table is accessed a few times to calculate the number of ones of a 64-bit integer, i.e., to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:63], y[0:63]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + d(x[32:47], y[32:47]) + d(x[48:63], y[48:63])$ where $x[i:i+15]$ and $y[i, i+15]$ are the 16-bit integers formed by the bits between $i$ and $i+15$ of $x$ and $y$, respectively.
 Each 16-bit distance is calculated through a single table access.
@@
 Then, many scans may be executed with no necessity to upload the bitstrings again.
-The OpenCL scanner supports running into multiple devices.
+The OpenCL scanner supports running on multiple devices.
 
@@
 It calculates one distance per workgroup, distributing each 64-bit integer operation per worker, and then summing the results obtained by the workers.
-The sum algorith is done by only the first worker of each workgroup.},
+The sum algorithm is done by only the first worker of each workgroup.},
 ]
@@
 It calculates one distance per workgroup, distributing each 64-bit integer operation per worker, and then summing the results obtained by the workers.
-The sum algorith is a parallel reduction, in which the workers split the array in two parts and sum the second part in the first part every loop.
+The sum algorithm is a parallel reduction, in which the workers split the array in two parts and sum the second part in the first part every loop.
 So, the sum is calculated in {$O(\log_2(\text{number of workers per workgroup}))$}.

### 2018-03-16 Chapters/sdm-critical-distance.tex
SDM fix grammar issues and typos

@@
 
-\caption{Results generated by the framework similar from Kanerva's original Table 7.2.
+\caption{Results generated by the framework similar to Kanerva's original Table 7.2.
 Here we have a 1,000 bit, 1,000,000 hard location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
@@
 Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM.
-Thereby, in order to show the interaction effects more clearly, we changed the single read for an 15-iterative read.
+Thereby, in order to show the interaction effects more clearly, we changed the single read for a 15-iterative read.
 As we can see in Figure \ref{fig:sdm-10000w-table-7-2-15iter}, after a distance of 500 bits, all bitstrings converged to 500-bit distance bitstrings, just as described by Kanerva.
@@
 Although Kanerva has found a formula for $\phi(d)$ through an unsolved integral, and \citet{de1995geometrical} have proposed another way to calculate $\phi(d)$, we have used our framework to estimate the values of $d$.
-In order to do that, we used a Monte Carlo approach, generating many pair of random bitstrings $d$ bits away from them and calculating the average number of shared hard locations between them.
+In order to do that, we used a Monte Carlo approach, generating many pairs of random bitstrings $d$ bits away from them and calculating the average number of shared hard locations between them.
 The code is available in the ``Calculate critical distance'' notebook \citep{sdmframework}.

### 2018-03-16 Chapters/sdm-performance.tex
SDM fix grammar issues and typos

@@
 
-We have created a benchmark to be able to compare the performance on different devices.
+We have created a benchmark to be able to compare the performance of different devices.
 So, the same performance test was executable in our devices, with results reported in tables and figures.
@@
 
-We recommend users to run a specific kernel comparison test for their own GPU and SDM settings.  This is available in the jupyter notebooks.
+We recommend users to run a specific kernel comparison test for their own GPU and SDM settings.  This is available in the Jupyter notebooks.
 
@@
 
-Even though scaning is an important part of the operations, we are really interested in the performance of the enterity of operations themselves.
+Even though scanning is an important part of the operations, we are really interested in the performance of the entirety of operations themselves.
 Comparing the times of the thread and OpenCL scanners with the times of their respective operations (either read or write), we can notice that their difference remains almost constant, which means the operation time itself is negligible when compared to the scan time.
@@
 There is no benchmark for kernel single\_scan5\_unroll because it returns the wrong result in this GPU.
-The problem is related with the premises of the optimization used by this kernel, which are not true for this GPU.
+The problem is related to the premises of the optimization used by this kernel, which are not true for this GPU.
 For the histogram of durations, see Figures \ref{fig:perf-ec2-p2-kernels}, \ref{fig:perf-imac-scanners}, \ref{fig:perf-imac-read}, and \ref{fig:perf-imac-write}.

### 2018-03-16 Chapters/sdm-validation.tex
SDM fix grammar issues and typos

@@
 
-As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
+As shown by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$.
 For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.
@@
   \subfloat[{\citet[Figure 1.2, p.25]{Kanerva1988}} ]{\includegraphics[width=0.45\textwidth]{./images02/new-images/kanerva-table-12.jpg}}
-  \subfloat[Generated by SDM-Frameworkm with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}
+  \subfloat[Generated by SDM framework with $n=1,000$ ]{\includegraphics[width=0.55\textwidth]{./images02/new-images/intersection-of-circles.png}}
 

### 2018-03-16 Chapters/sdm2.tex
SDM fix grammar issues and typos

@@
 \caption{Activated addresses inside access \protect \\
-radius $r$ around center address.\label{fig-addresses-inside-access-radius}}
+radius $r$ around the center address.\label{fig-addresses-inside-access-radius}}
 \end{figure}
@@
 
-When reading a cue $\eta_{x}$ that is $x$ bits away of $\eta$, the cue shares many addresses with $\eta$.
+When reading a cue $\eta_{x}$ that is $x$ bits away from $\eta$, the cue shares many addresses with $\eta$.
 The number of shared addresses decreases as the cue's distance to $\eta$ increases, in other words, as $x$ increases.
@@
 
-The addresses of the $\{0,1\}^{n}$ space grows exponentially with the number of dimensions $n$, i.e.
+The addresses of the $\{0,1\}^{n}$ space grow exponentially with the number of dimensions $n$, i.e.
 $N=2^{n}$.
@@
 This is represented in Figure \ref{fig-hardlocations}.
-All properties of read and write operations presented before remain valid, but limited to hard locations.
+All properties of reading and write operations presented before remain valid, but limited to hard locations.
 Kanerva suggests taking a sample of about one million hard locations.
@@
 
-Kanerva also suggests the selection of a radius that will activate, on average, one one thousandth of the sample, which is 1,000 hard locations for a sample of one million addresses.
+Kanerva also suggests the selection of a radius that will activate, on average, one-thousandth of the sample, which is 1,000 hard locations for a sample of one million addresses.
 In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$.
@@
 Each location in our sample is called a hard location.
-On these hard locations we do operations of read and write.
+On these hard locations, we do operations of read and write.
 One of the insights of SDM is exactly the way we read and write: using data as addresses in a distributed fashion.
 Each datum $\eta$ is written in every activated hard location inside the access radius centered on the address, that equals datum, $\xi=\eta$.
-Kanerva suggested using an access radius $r$ having about one one thousandth of $N'$.
+Kanerva suggested using an access radius $r$ having about one-thousandth of $N'$.
 As an imprecise cue $\eta_{x}$ shares hard locations with the target bitstring $\eta,$ it is possible to retrieve $\eta$ correctly.
@@
 The read is performed polling each activated hard location and statistically choosing the most written bit for each dimension.
-It consists of adding all $n$ counters from the activated hard locations and, for each bit, choosing bit 1 if the counter is positive, choose bit 0 if the counter if negative, and randomly choose bit 0 or 1 if the counter is zero.
+It consists of adding all $n$ counters from the activated hard locations and, for each bit, choosing bit 1 if the counter is positive, choosing bit 0 if the counter is negative, and randomly choose bit 0 or 1 if the counter is zero.
 
@@
 How do we know which concepts are linked together?
-(ii) From a concept bitstring how can we list all concepts that are somehow linked to it?
+(ii) From a concept bitstring, how can we list all concepts that are somehow linked to it?
 This second question is called the problem of spreading activation.
@@
 With this in mind, let us ask first how one reads from a single hard location, $x$.
-Compute $\zeta$, the bit vector read at $x$, by assigning its $i$th bit the value 1 or 0 according as the $i$th counter at $x$ is positive or negative.
+Compute $\zeta$, the bit vector read at $x$, by assigning its $i$th bit the value 1 or 0 according to the $i$th counter at $x$ is positive or negative.
 Thus, each bit of $\zeta$ results from a majority rule decision of all the data that have been written at $x$.
@@
 
-This fact just highlihts how important it is to have a reference implementation that one may read the code to clarify one's understanding about the details of each operation.
+This fact just highlights how important it is to have a reference implementation that one may read the code to clarify one's understanding about the details of each operation.
 
@@
 Hence, we will here explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.
-Mathematically, let $A$ be the set of the counters of the activated hardlocation, and $c_i$ be the counter of the $i$-th bit.
+Mathematically, let $A$ be the set of the counters of the activated hard location, and $c_i$ be the counter of the $i$th bit.
 Then,
@@
 
-Finally, the $i$-th bit of the resulting bitstring is 1 if $s_i > 0$, or 0 if $s_i < 0$, or random if $s_i = 0$.
+Finally, the $i$th bit of the resulting bitstring is 1 if $s_i > 0$, or 0 if $s_i < 0$, or random if $s_i = 0$.
 Notice that when $z=1$, then $s_i = \sum_{c \in A} c_i$, which is the Kanerva's read; and when $z=0$, then $s_i = \frac{c_i}{|c_i|} = \text{sign}(c_i)$, which is the Chada's read.
@@
 In the $y$-axis, we display the read operation's behavior as the number of items registered in the memory grows.
-These graphs are computing intensive, yet they can be easily tested by readers in our provided jupyter notebooks.
+These graphs are computing intensive, yet they can be easily tested by readers in our provided Jupyter notebooks.
 Note the different scales.}

### 2018-03-16 Chapters/Introduction.tex
quotes
one life

@@
 \bigskip
+
+\begin{flushright}{\slshape
+	It is still an unending source of surprise for me to see \\
+	how a few scribbles on a blackboard or on a sheet of paper\\
+	could change the course of human affairs.} \\ \medskip
+    --- Stanislaw Ulam
+\end{flushright}
 
@@
 In between, both the iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created (Uber, Instagram, etc, on the case of smartphones; and exchanges, miners, investors, payment processors, etc., in the case of blockchains).
-Linhares's thesis is that students should (i) expect new, unexpected, technological platforms, (ii) rapidly identify them, and (iii) throw their energy at them, as they offer leverage to make an asymmetric impact.
+The key point is that we should (i) expect new, unforeseen, technological platforms, (ii) rapidly identify them, and (iii) throw their energy at them, as they offer leverage to make an asymmetric impact.
 \label{fig:onelife}}

### 2018-03-16 Chapters/Chapter03.tex
Bass model: minor grammar and typo fixes

@@
 An important difference between the number of \textit{active adopters} and the \textit{total adopters} is that first may decrease over time, while the latter cannot.
-First, I discuss the \citet{bass1969} model and its parameters; then I describe the model extension and analyze four models of rejection; next I detail the estimation method; after that I estimate the models by using Facebook's active users dataset; and finally, I discuss the results and conclude.
+First, I discuss the \citet{bass1969} model and its parameters; then I describe the model extension and analyze four models of rejection; next, I detail the estimation method; after that, I estimate the models by using Facebook's active users dataset; and finally, I discuss the results and conclude.
 
@@
 At the time of this writing, Facebook's market cap is \$222.69B, Twitter's is \$23.18B, and Netflix's is \$20.49B\footnote{Data obtained from Yahoo!
-Finance website at December 1\textsuperscript{st}, 2016.}.
+Finance website on December 1\textsuperscript{st}, 2016.}.
 
@@
 The potential market size $m$ is the unknown number of people who will have adopted the innovation after a very long time.
-It is not exactly the target market of the innovation, but a subset of it, as no product difuses over its entire target market.
+It is not exactly the target market of the innovation, but a subset of it, as no product diffuses over its entire target market.
 If a company estimates and updates the model more than once for their product, the change in $m$ is a change in the potential market and could help the company to understand whether their decisions in the meantime have increased or decreased the number of future adopters.
-As, $\lim_{t \rightarrow \infty} Y(t) = m$, the whole potential market will have adopted the innovation at some point.
+As $\lim_{t \rightarrow \infty} Y(t) = m$, the whole potential market will have adopted the innovation at some point.
 
@@
 
-These four models can be grouped in two families, one for the equations \ref{eq:model1} and \ref{eq:model2}, and another for the equations \ref{eq:model3} and \ref{eq:model4}.
+These four models can be grouped into two families, one for the equations \ref{eq:model1} and \ref{eq:model2}, and another for the equations \ref{eq:model3} and \ref{eq:model4}.
 The former relates the rejection to the rate of new people adopting the innovation, as if people decide whether they will use or reject the innovation when they try it.
@@
 
-The right side of the differential equation \ref{eq:model2} is the rate of rejection, i.e., the probability of someone who rejecting the innovation, provided that he/she has not rejected it yet.
+The right side of the differential equation \ref{eq:model2} is the rate of rejection, i.e., the probability of someone rejecting the innovation, provided that he/she has not rejected it yet.
 Thus, in this model, the rate of rejection is proportional to the percentage of people adopting the innovation at time $t$, i.e., the more people adopt the innovation, the more they reject it.
@@
 
-The differential equation \ref{eq:model2} can also be solved for $R(t)$ algebrically.
+The differential equation \ref{eq:model2} can also be solved for $R(t)$ algebraically.
 Using the fact that $-\frac{d}{dt} \log[1 - R(t)] = r(t) / [1 - R(t)] = \nu f(t)$, and integrating both sides of this equation yields:
@@
 Thus, while there are \textit{active users}, a fraction $\nu$ of them will be rejecting the innovation.
-Hence, everyone will have reject the innovation sooner or later.
+Hence, everyone will have rejected the innovation sooner or later.
 
@@
 Thus, while there are \textit{active users}, the rate of rejection will be greater than zero.
-Hence, everyone will have reject the innovation sooner or later.
+Hence, everyone will have rejected the innovation sooner or later.
 
@@
 While it worked well for him, it has not in the present work.
-There are several well known problems in the estimation of parameters, most of them related to approximation of derivatives and instability of the estimators.
+There are several well-known problems in the estimation of parameters, most of them related to approximation of derivatives and instability of the estimators.
 These problems have been noticed by many authors \citep{schmittlein1982maximum, srinivasan1986technical, xie1997kalman}.
@@
 Since there is no explicit solution for the parameters $m$, $p$, $q$, $w$, and $\nu$ which maximize the LogLikelihood function, the parameters were estimated using the Truncated Newton Constrained (TNC) method \citep{dembo1983truncated, nash2000survey, nash1984newton} from the SciPy Python Library \citep{scipy}.
-The constrains were $m > 0$, $p > 0$, $q > 0$, $w \geq 0$, and $\nu \geq 0$ for all models.
+The constraints were $m > 0$, $p > 0$, $q > 0$, $w \geq 0$, and $\nu \geq 0$ for all models.
 The initial guess for the TNC method was the same for all models estimation.
@@
 
-The models parameters were estimated using the number of Facebook's \textit{active users} from December 2004 to March 2013 \citep{facebookdata}.
+The models' parameters were estimated using the number of Facebook's \textit{active users} from December 2004 to March 2013 \citep{facebookdata}.
 In the dataset, $x_i$ was the number of Facebook's \textit{active users} at time $t_i$.
@@
 
-It is also intended to run a back test with the available data.
-First, the parameters of the model is estimated using a subset of the dataset.
-Then, through extrapolation, the number of \textit{active uses} is forecast.
+It is also intended to run a backtest with the available data.
+First, the parameters of the model are estimated using a subset of the dataset.
+Then, through extrapolation, the number of \textit{active users} is forecast.
 Finally, it is compared to this part of the dataset - which must not have been used in the estimation.

### 2018-03-16 Chapters/Chapter02.tex
SDM more grammar and typo fixes

@@
 Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger.
-This flexibility into mapping one situation in another is an important human feature which is hard to replicate in computers.
+This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.
 
@@
 Our motivation was our own effort to run our models.
-As there is no reference implementation, we had to implement our own and run several simulations to ensure that our implementation was correct and bug-free.
-Thus, we had to deviate from our main goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
+As there is no reference implementation, we had to develop our own and run several simulations to ensure that our implementation was correct and bug-free.
+Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
 Furthermore, new members of our research group had to go through different source codes developed by former members.
@@
 Extensions of SDM have been used in many applications.
-For example, \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.
+For example, \citet{Snaider2011} extended SDM to store sequences of vectors and trees efficiently.
 \citet{Rajesh1998} used a modified SDM in an autonomous robot.
@@
 
-An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance becomes small, i.e., SDM recall capacity is very diminished.
+An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance becomes small, i.e., SDM recall capacity is hugely diminished.
 After 900,000 neuron deaths, the critical distance is zero.
@@
 
-We run exactly the same simulation for a 256-bit SDM with 1,000,000 hard locations.
+We run the same simulation for a 256-bit SDM with 1,000,000 hard locations.
 The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
@@
 The results, however, have not yielded performance improvements.
-Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller critical distance and faster divergence at large-distance reads.
+Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows an evident deterioration, with a smaller critical distance and faster divergence at large-distance reads.
 This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
@@
 A first SDM would write the whole picture, just like we have done.
-Another SDM would write specific parts of the image, just like our eye focusing on specific parts.
+Another SDM would write specific regions of the image, just like our eye focusing on specific regions.
 When reading, they may compose the counters and give a more precise classification.
@@
 
-One intriguing question here is:  is there a `better' number of dimensions and of hard locations?  If so, can such numbers be better studied algebraically or numerically?
+One intriguing question here is:  is there a `better' number of dimensions and hard locations?  If so, can such numbers be better studied algebraically or numerically?
 

### 2018-03-16 Chapters/sdm-applications.tex
SDM more grammar and typo fixes

@@
 Then, new pieces of data are presented and the algorithm must classify them into the already known groups.
-It is named supervised because  the algorithm will not create the groups itself.
-It will learn the groups from during the learning phase, in which the groups have already been defined and the pieces of data have already been classified into them.
+It is named ``supervised'' because  the algorithm will not create the groups itself.
+It will learn the groups during the learning phase, in which the groups have already been defined and the pieces of data have already been classified into them.
 
@@
 So, white pixels were equal to bit 0, and black pixels to bit 1.
-The 8 remaining bits were all set to zero.
+The eight remaining bits were all set to zero.
 This was a bijective mapping (or one-to-one mapping), i.e., there was only one bitstring for each image, and there was only one image for each bitstring.
@@
 The low noise scenario had the same noise as the training set.
-For each scenario, we had classified 620 unknown images with 10 images per group.
+For each scenario, we had classified 620 unknown images with ten images per group.
 
@@
 We believe that it happened because the classifier had never seen an image with no noise and the difference between the images of ``l'' and ``i'' is smaller than the critical distance.
-So, both groups have been merged and it would converge to only one of them.
+So, both groups have been merged and it converged to only one of them.
 In our simulation, it happened to be the group of ``i''.
@@
 As we have 62 groups and each has been trained with 100 images, there were 6,200 writes to the memory.
-When an image is being classified, it will have to converge to a group, and the convergence depends on the distance between this image and the images from the training set, i.e, in the noise level.
+When an image is being classified, it will have to converge to a group, and the convergence depends on the distance between this image and the images from the training set, i.e., in the noise level.
 
@@
 These results show that the SDM may be used as a supervised classification algorithm.
-Although we do not believe that the mapping between images and bitstrings are even close to the way human cognition deals with images, we believe the results are interesting and useful to many possible real-world problems.
+Although we do not believe that the mapping between images and bitstrings are even close to the way human cognition deals with images, we believe the results are exciting and useful to many possible real-world problems.
 
@@
 Image noise filtering consists in removing the noise from an input, in our case an image.
-Our images are black \& white images and the noise is generated randomly flipping some of their pixels from black to while and vice versa.
+Our images are black \& white images, and the noise is generated randomly flipping some of their pixels from black to while and vice versa.
 In Figure \ref{fig-filter-progressive-noise}, we may see an image with different levels of noise, from 0\% to 45\% in steps of 5\%.
@@
 Each image is mapped into a bitstring of 1,000 bit in which the bits are set according to the color of each pixel of the image.
-White pixels are mapped to bit 0, and black pixels to bit 1.
+White pixels are assigned to bit 0, and black pixels to bit 1.
 The 100 remaining bits are all set to zero.
@@
 
-Then, in order to test the filtering, we just have read from noisy images, and the results were really impressive.
+Then, in order to test the filtering, we just have read from noisy images, and the results were remarkable.
 We were able to clean images up to 42\% of noise (see Figure \ref{fig:filter-testing}).
@@
 We have confirmed it writing the letters ``B'', ``C'', ``D'' and the numeral ``8''.
-They share a great number of hard locations and our noise filter could not filter their noise correctly.
+They share a high number of hard locations, and our noise filter could not filter their noise correctly.
 The training sets can be seen in Figure \ref{fig:filter-training2} and the results in Figure \ref{fig:filter-results2}.
@@
 
-Our algorithm to decide what should be played is very simple: it reads the current board from SDM and then it chooses the valid action with the highest score.
+Our algorithm to decide what should be played is simple: it reads the current board from SDM and then it chooses the valid action with the highest score.
 To calculate the scores, the bitstring is split into 9 parts, one per action (Figure \ref{fig:ttt-action-mapping}).
@@
 For instance, if SDM should try to win in the first place, no matter if it may lead to losing, the weight of the positive reward should be higher than the others.
-But, if it is more important to not lose, then the weight of the negative reward should be higher.
+But, if it is more important not to lose, then the weight of the negative reward should be higher.
 
-In the very beginning, SDM is empty and its counters are zeroed.
+In the very beginning, SDM is empty, and its counters are zeroed.
 So, any reading will result in a random bitstring, because it will flip a coin for all counters.
 Thus, the chosen action will be random, as their scores will follow a binomial distribution.
-It is exactly the desired behavior --- we play randomly until we learn.
+It is precisely the desired behavior --- we play randomly until we learn.
 In fact, it will happen every time an unknown board is seen.
@@
 When playing against the random player, SDM has already started improving after the first 100 training games.
-Its winning rate converges quickly to around 80\%, while the drawing rate starts to grow after 2,000 games and keeps growing until the end.
+Its winning rate converges quickly to around 80\%, while the drawing rate starts to grow after 2,000 games and keeps rising until the end.
 The losing rate keeps decreasing until it reaches cycles of 100 games without any loss.
@@
 
-When playing against the smart player, SDM has started learning how to not lose during the first training cycles.
+When playing against the smart player, SDM has started learning how not to lose during the first training cycles.
 The drawing rate grows quickly in the first 500 games and slowly since then.
-The winning rate grew to around 20\% and remained there until 6,000 games, while the drawing rate kept growing.
-Then, after learning how to not lose, SDM started learning how to win, since, after 7,000 games, the winning rate started to increase.
+The winning rate grew to around 20\% and remained there until 6,000 games, while the drawing rate kept increasing.
+Then, after learning how not to lose, SDM started learning how to win, since, after 7,000 games, the winning rate started to increase.
 See Figure \ref{fig:ttt-results-smart}.
 
-When playing against another SDM player, both player quickly learns how to not lose.
+When playing against another SDM player, both player quickly learns how not to lose.
 During the first 100 testing games, without any learning game, they have behaved like two random players.
-After the 100 training games, they have learned how to not lose and the drawing rate grows quickly, reaching 100\%.
+After the 100 training games, they have learned how not to lose, and the drawing rate grows quickly, reaching 100\%.
 See Figure \ref{fig:ttt-results-sdm}.

### 2018-03-16 Chapters/sdm-architecture.tex
SDM more grammar and typo fixes

@@
 In this case, none of the known algorithms is able to solve our problem faster than $O(H)$.
-The algorithm which seems to best fit in our problem consumes $O(H)$ space and runs in $O(\log^n(H))$ \citep{chazelle1988functional}, which is really slower than $O(H)$ when, for instance, $H=1,000,000$ and $n=1,000$.
+The algorithm which seems to best fit in our problem consumes $O(H)$ space and runs in $O(\log^n(H))$ \citep{chazelle1988functional}, which is significantly slower than $O(H)$ when, for instance, $H=1,000,000$ and $n=1,000$.
 For a review of the range reporting algorithms, see \citet{chan2011orthogonal}.
@@
 But, in our case, for a 1,000 bits SDM, $r/n = 0.451$, which changes the runtime to $O(H^{0.993})$.
-This is really close to $O(H)$, but with a larger constant.
+This is very close to $O(H)$, but with a larger constant.
 Unfortunately, $O(H)$ is still faster.
@@
 
-There are 8 OpenCL kernels which explore differently the GPU architecture to improve performance.
+There are eight OpenCL kernels which explore differently the GPU architecture to improve performance.
 It is necessary because there are several GPU microarchitectures and a single kernel will never be optimal for all of them.
@@
 
-All 8 kernels do the same thing: calculate the exclusive OR (XOR) between two 64-bit integers and count the number of bits one in the result.
+All eight kernels do the same thing: calculate the exclusive OR (XOR) between two 64-bit integers and count the number of bits one in the result.
 They just do it with different approaches.
@@
 It calculates one distance per worker and let the GPU decide how to distribute this task between workgroups and workers.
-It is the most simple kernel and does not explore any details of the GPU architecture.},
+It is the most straightforward kernel and does not explore any details of the GPU architecture.},
 ]
@@
 caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan1}}.
-It is just like \lstinline{single_scan0}, but it distribute several distances per workgroup, which, in turn, distribute the distances among their workers.},
+It is just like \lstinline{single_scan0}, but it distributes several distances per workgroup, which, in turn, distributes the distances among their workers.},
 ]
@@
 It calculates one distance per workgroup, distributing each 64-bit integer operation per worker, and then summing the results obtained by the workers.
-The sum algorithm is a parallel reduction, in which the workers split the array in two parts and sum the second part in the first part every loop.
+The sum algorithm is a parallel reduction, in which the workers split the array into two parts and sum the second part in the first part every loop.
 So, the sum is calculated in {$O(\log_2(\text{number of workers per workgroup}))$}.
@@
 This kernel is just like \lstinline{single_scan3}, but it explores one more detail of many GPU microarchitecture: the size of the warp.
-As the workers in the same warp are always synchronized, there is no need to synchronize them using a barrier.
+As the workers in the same warp are always synchronized, there is no need to sync them using a barrier.
 This specific kernel only works when the number of workers per workgroup is a power-of-2.},
@@
 This kernel is just like \lstinline{single_scan5}, but it works with any number of workers per work.
-The tradeoff is an aditional step in the parallel reduction algorithm.},
+The tradeoff is an additional step in the parallel reduction algorithm.},
 ]
@@
 In this case, the results are floating points instead of integers, which considerably reduces performance.
-When $z=1$, it is exactly as the Kanerva's read.
+When $z=1$, it is precisely as the Kanerva's read.
 When $z=0$, it is the Chada's read.
@@
 
-There is another special read operation: the weighted reading.
+There is another particular read operation: the weighted reading.
 In the weighted reading, the value of the counters is multiplied by a weight which depends only on the distance between the reading address and the hard location address.
 The weight is retrieved from a lookup table of integers indexed by the distance.
-The rest of the read operation is exactly the same.
+The remaining of the read operation is just the same.
 

### 2018-03-16 Chapters/sdm-critical-distance.tex
SDM more grammar and typo fixes

@@
 \caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
-Here we had a 1,000 bit, 1,000,000 hard location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
+Here we had a 1,000 bit, 1,000,000 hard location SDM with 10,000 random bitstrings written into it, which was also Kanerva's configuration.
 \label{fig:sdm-10000w-table-7-2}}
@@
 
-But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Figure \ref{fig:sdm-100w-table-7-2-10samples}).
-This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.
+However, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Figure \ref{fig:sdm-100w-table-7-2-10samples}).
+This result strengthens our hypothesis that the disparities in the computational outcomes are due to the interaction effect of high numbers of different attractors.
 In Figure \ref{fig:sdm-table-7-2-steps} we can notice that the more random bitstrings are written, the stronger the attractors.
@@
 \caption{Results generated by the framework similar to Kanerva's original Table 7.2.
-Here we have a 1,000 bit, 1,000,000 hard location SDM with (a) exactly 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
+Here we have a 1,000 bit, 1,000,000 hard location SDM with (a) just 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.
 \label{fig:sdm-100w-table-7-2}}
@@
 \caption{This graph shows the interaction effects more clearly.
-As we change the single read to a 6-iterative read, the effect has vanished and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings.
-Here we have the exact same configuration of Figure \ref{fig:sdm-10000w-table-7-2}, except for the iterative read.
+As we change the single read to a 6-iterative read, the effect has vanished, and all bitstrings above $x=500$ have converged to 500-bit distance bitstrings.
+Here we have precisely the same configuration of Figure \ref{fig:sdm-10000w-table-7-2}, except for the iterative read.
 \label{fig:sdm-10000w-table-7-2-15iter}}
@@
 Going further in the analysis, we calculated the probability of missing a bit when reading from SDM.
-After all, that's how Kanerva has originally found the curve.
+After all, that is how Kanerva has originally found the curve.
 To do this, we used the following equations from our previous work \citep{brogliato2014sparse}.
-Let $d$ be the distance to the target, $h$ be the number of hard locations activated during read and write operations, $s$ be the number of total stored bitstrings, $H$ be the number of total hard locations, $w$ be the number of times the target was written into SDM, $\theta$ be the total random bitstrings in all $h$ hard locations activated by read operation, and $\phi(d)$ be the average number of shared hard locations activated two bitstrings $d$ bits away.
+Let $d$ be the distance to the target, $h$ be the number of hard locations activated during reading and write operations, $s$ be the number of total stored bitstrings, $H$ be the number of total hard locations, $w$ be the number of times the target was written into SDM, $\theta$ be the total random bitstrings in all $h$ hard locations activated by read operation, and $\phi(d)$ be the average number of shared hard locations activated two bitstrings $d$ bits away.
 
@@
 As we can easily notice, we got exactly the same curve as Kanerva.
-This question has been intriguing us since then and we are still looking for a more analytic explanation than merely interference from the other written attractors.
+This question has been intriguing us since then, and we are still looking for a more analytic explanation than merely interference from the other written attractors.
 

### 2018-03-16 Chapters/sdm-validation.tex
SDM more grammar and typo fixes

@@
 
-In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.
+In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works appropriately.
 
@@
 Notice that the curve is almost the same for $n=1,000$ and $n=256$.
-It happens because the access radius is adjusted to get $p$ as close as possible to $0.001$.
-They are not exactly the same because their $p$ differs a little.
+This difference happens because the access radius is adjusted to get $p$ as close as possible to $0.001$, which mean they are very close, but not the same.
 
@@
 Kanerva has calculated the intersection of two circles according to the distance between their centers.
-The intersection is important to understand how SDM works because it affects directly the critical distance.
+The intersection is important to understand how SDM works because it directly affects the critical distance.
 When $\eta_d$ is inside the critical distance, then it will converge to $\eta$.
@@
 
-We have tested exactly the same example presented in \citet{Kanerva1988}, p.85.
+We have tested the same example presented in \citet{Kanerva1988}, p.85.
 We wrote two sequences to a $3$-fold memory: $<A, B, C, D>$ and $<E, B, C, F>$.
@@
 The logic behind how it works is that, when reading the sequence $<A, B, C>$, we have $A$ pointing to $D$, while both $B$ and $C$ point to $D$ and $F$.
-Thus, $D$ appears more often than $F$ and ended up being the result.
+Thus, $D$ appeared more often than $F$ and ended up being the result.
 
-Hence, as we have replicated the theoretical results from Kanerva, we have one more evidence that our framework works properly.
+Hence, as we have replicated the theoretical results from Kanerva, we have one more evidence that our framework works appropriately.
 

### 2018-03-16 Chapters/sdm2.tex
SDM more grammar and typo fixes

@@
 
-In traditional memory, each datum is stored in an address and every look up of a specific datum requires a search through the memory.
+In traditional memory, each datum is stored in an address and every lookup of a specific datum requires a search through the memory.
 In spite of computer scientists having developed beautiful algorithms to perform fast searches, almost all of them do a precise search.
@@
 If the cue is sufficiently near the target datum $\eta$, the read output will be closer to $\eta$ than $\eta_{x}$ was.
-Repeating the read operation increasingly gets results closer to $\eta$, until it is exactly the same.
-So, it may be necessary to perform more than one read operation in order to converge to the target data $\eta$.
+Repeating the read operation increasingly gets results closer to $\eta$, until it is precisely the same.
+So, it may be necessary to perform more than one read operation to converge to the target data $\eta$.
 
@@
 
-The addresses of the $\{0,1\}^{n}$ space grow exponentially with the number of dimensions $n$, i.e.
-$N=2^{n}$.
-For $n=100$ we have $N\approx10^{30}$, which is incredibly large when related to a computer memory.
+The addresses of the $\{0,1\}^{n}$ space grow exponentially with the number of dimensions $n$, i.e., $N=2^{n}$.
+For $n=100$ we have $N\approx10^{30}$, which is incredibly large when related to computer memory.
 Furthermore, \citet{Kanerva1988} suggests $n$ between 100 and 10,000.
@@
 This is represented in Figure \ref{fig-hardlocations}.
-All properties of reading and write operations presented before remain valid, but limited to hard locations.
+All properties of reading and write operations presented before remain valid but limited to hard locations.
 Kanerva suggests taking a sample of about one million hard locations.
@@
 In Figure \ref{fig-p1-p2-iterative-read}, the circle has radius equal to the critical distance and every $\eta_{x}$ inside the circle should converge.
-The figure also shows a convergence in four readings.
+The figure also shows convergence in four readings.
 
@@
 A random bitstring is generated with equal probability of $0$'s and $1$'s in each bit.
-One can readily see that the average distance between two random bitstrings has binomial distribution with mean $n/2$ and standard deviation $\sqrt{n/4}$.
+One can readily see that the average distance between two random bitstrings follows the binomial distribution with mean $n/2$ and standard deviation $\sqrt{n/4}$.
 For a large $n$, most of the space lies close to the mean and has fewer shared hard locations.
@@
 
-This mapping from concepts to bitstrings brings us two main questions: (i) Suppose we have a bitstring that is linking two major concepts.
+This mapping from concepts to bitstrings brings us two central questions: (i) Suppose we have a bitstring that is linking two major concepts.
 How do we know which concepts are linked together?

### 2018-03-16 Chapters/Conclusion.tex
Minor grammar fixes in Introduction and Conclusion.

@@
 Nichols, will participate through teleconferencing over the open web, with no use of hardware specific for the task.
-The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer \citep{TonnyMartins}.
+The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI, and cognitive technology… as an executive, not as a research scientist or specialized engineer \citep{TonnyMartins}.
 Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citep{AndrewNgAI}, as his hyperbolic way to emphasize the potential transformational power of the technology.
@@
 In AI, SDM seems to be a particularly interesting area for study.
-The model plausibly reflects a number of well known aspects of psychology and neuroscience.
+The model plausibly reflects a number of well-known aspects of psychology and neuroscience.
 For example, neurons can easily compute the address decoding scheme of the system.

### 2018-03-16 Chapters/Introduction.tex
Minor grammar fixes in Introduction and Conclusion.

@@
 The Bill and Melinda Gates Foundation chose as focus of its ``Level-One project'': to provide basic financial services through cell phones.
-Another initiative, the United Nations World Food Programme has began, in 2017, an experiment in Jordan, in which the organization provides funds for thousands of people towards its goal of food relief.
+Another initiative, the United Nations World Food Programme has begun, in 2017, an experiment in Jordan, in which the organization provides funds for thousands of people towards its goal of food relief.
 An interesting aspect of this program has been the format of the funds distributed:  they have been all on the ethereum blockchain \citep{woyke2017blockchain, UN-you-pay-we-take-the-photo, UNFOODPROG}.
@@
 On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.
-These incentives keep the community together and have maintened Bitcoin alive.
+These incentives keep the community together and have maintained Bitcoin alive.
 
@@
 
-Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'', which believe that is it risky to increase the the blocksize (in which all transactions are stored).
+Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'', which believe that is it risky to increase the blocksize (in which all transactions are stored).
 It is not known whether a blocksize, say, of 1GB, would be feasible to sustain the decentralization of the network.
@@
 More and more expensive human labor has been replaced by algorithms.
-Managers are able to make better decisions because they receive real time information.
+Managers are able to make better decisions because they receive real-time information.
 The supply chain has incredibly evolved thanks to advances in logistics supported by routing algorithms, storage algorithms, and many others optimization algorithms.
@@
 The computer code has been used in a previous Ph.D.
-Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been ran on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).
+Thesis; the code has shown some small discrepancies from theoretical expectations; the code has been run on a number of different architectures and information-processing devices (e.g., CPUs, GPUs).
 The framework enables us to have a visual exploration previous experiments and new possibilities for SDM.

### 2018-03-22 Chapters/Chapter02.tex
mudancas marcelo aqui em casa
SDM Grammar and typo fixes.

@@
 SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2011}.
+\citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
 Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
@@
 Other contributions have also been introduced, which include (i) a noise filtering approach, (ii) a supervised classification algorithm, (iii) and a reinforcement learning algorithm, all of them using only the original SDM proposed by Kanerva, i.e., with no additional mechanisms, algorithms, data structures, etc.
-Although some of these applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them have done some adapting of SDM to their problems, and none of them have used just the ideas introduced by Kanerva.
+Although some of these applications have already been explored in previous work \citep{Meng2009, fan1997genetic, rao1995natural}, all of them have adapted SDM to fit their problems, and none of them have used just the ideas introduced by Kanerva.
 We have presented different approaches with no adaptations whatsoever.
@@
 An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance becomes small, i.e., SDM recall capacity is hugely diminished.
-After 900,000 neuron deaths, the critical distance is zero.
-In this case, everything has been lost.
+After 900,000 neuron deaths, the critical distance is zero, and everything has been lost.
 
@@
 
+Dr.
 Murilo observed that the models of Kanerva's read ($z=1$) and Chada's read ($z=0$) were simple variations of a generalized read with an exponent $z$, which suggests experimenting with different values.
@@
 We understand that the critical distance is an important parameter of SDM.
-The bigger the critical distance, the best, because SDM is able to converge even with farther clues.
+The bigger the critical distance, the better, because SDM is able to converge even with farther clues.
 For $z>1$, the bigger the $z$, the smaller the critical distance.

### 2018-03-22 Chapters/Conclusion.tex
SDM Grammar and typo fixes.

@@
 The possibility that there will be some form of purely digital money has become very real; and we have started this study with two possible forms of organization of a purely digital money system; a blockchain and a directed acyclic graph.
-Consider, just as a matter of comparison, Brazil's most important company: Petrobrás.
-As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobrás by ten billion dollars, while Bitcoin's is valued at more than double of Petrobrás (195B usd vs 83B usd).
+Consider, just as a matter of comparison, Brazil's most important company: Petrobras.
+As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobras by ten billion dollars, while Bitcoin's is valued at more than double of Petrobras (195B usd vs 83B usd).
 Prices change, of course; but these technologies should, at a minimum, be taken seriously.

### 2018-03-22 Chapters/Introduction.tex
SDM Grammar and typo fixes.

@@
 
-\caption{Slides from an old Linhares class; a viewpoint that has influenced the choice of topics found in this thesis.
+\caption{Slides from an old Linhares' class; a viewpoint that has influenced the choice of topics found in this thesis.
 In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
@@
 In between, both the iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created (Uber, Instagram, etc, on the case of smartphones; and exchanges, miners, investors, payment processors, etc., in the case of blockchains).
-The key point is that we should (i) expect new, unforeseen, technological platforms, (ii) rapidly identify them, and (iii) throw their energy at them, as they offer leverage to make an asymmetric impact.
+The key point is that we should (i) expect new, unforeseen, technological platforms, (ii) rapidly identify them, and (iii) throw our energy at them, as they offer leverage to make an asymmetric impact.
 \label{fig:onelife}}
@@
 Of particular interest to us is the advent of the computer, and of engineering efforts that brought exponential growth in computational power to the hands of individuals.
-Whilst, during the war, computations were mostly done by hand, the electronic computer took over afterwards; up to an extent that it is not outlandish to say that this original field can be referred to, today, as \emph{computational management science}.
+Whilst, during the war, computations were mostly done by hand, the electronic computer took over afterward; up to an extent that it is not outlandish to say that this original field can be referred to, today, as \emph{computational management science}.
 
@@
 
-This thesis will explore three different topics related to \emph{computing business platforms} (Figure \ref{onelife}).
+This thesis will explore three different topics related to \emph{computing business platforms} (Figure \ref{fig:onelife}).
 Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three specific, self-contained, scientific papers.
@@
 On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick and global, and the currency issuance rate is well known.
-On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.
+On the other hand, the miners have incentive to be part of Bitcoin's network because, every ten minutes, new coins are found and transactions' fees are collected.
 These incentives keep the community together and have maintained Bitcoin alive.
@@
 
-Behind many artificial intelligence systems, there is pattern recognition: The capacity to match information from new data with information which has already been seen and is stored in memory.
+Behind artificial intelligence systems, there is pattern recognition: The capacity to match information from new data with information which has already been seen and is stored in memory.
 It may be used in classification, face recognition, character recognition, and so forth.

### 2018-03-22 Chapters/sdm-architecture.tex
SDM Grammar and typo fixes.

@@
 
-We split the SDM memory into two parts: the hard location addresses and the hard location counters.
+The SDM memory has been split into two parts: the hard location addresses and the hard location counters.
 Thus, the addresses (bitstrings) of the hard locations are stored in one array, while their counters in another.
@@
 
-Let a SDM memory with $N$ dimensions and $H$ hard locations.
-Then, in a 64-bit computer, the array of hard location addresses will use $H \cdot 8 \cdot \lceil N/64 \rceil$ bytes of memory, and there will be $H \cdot N$ hard location counters.
+Let a SDM memory with $n$ dimensions and $H$ hard locations.
+Then, in a 64-bit computer, the array of hard location addresses will use $H \cdot 8 \cdot \lceil n/64 \rceil$ bytes of memory, and there will be $H \cdot n$ hard location counters.
 For example, in a SDM memory with 1,000 dimensions and 1,000,000 hard locations, using 32-bit integers for the counters, the array of addresses will use 122MB of memory and the counters will use 3.8 GB of memory.
@@
 \matrix[matrixB, below=of A] (B) {
-addr$_{k, 1}$ & addr$_{k, 2}$ & addr$_{k, 3}$ & $\cdots$ & addr$_{k, 8 \cdot \lceil N/64 \rceil}$ \\
+addr$_{k, 1}$ & addr$_{k, 2}$ & addr$_{k, 3}$ & $\cdots$ & addr$_{k, 8 \cdot \lceil n/64 \rceil}$ \\
 };
@@
 ] (B-1-1.south west) -- (B-1-5.south east)
-node [pos=0.5,anchor=north,yshift=-0.5cm] {N bits};
+node [pos=0.5,anchor=north,yshift=-0.5cm] {$n$ bits};
 
@@
 \caption{Address space's bitstrings are stored in a contiguous array.
-In a 64-bit computer, each bitstring is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil N/64 \rceil$.\label{tab:hl-addresses-detail}}
+In a 64-bit computer, each bitstring is stored in a sub-array of 64-bit integers, with length $8 \cdot \lceil n/64 \rceil$.\label{tab:hl-addresses-detail}}
 \end{figure}
@@
 
-In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}.
+In 2014, \citet{norouzi2014fast} published a solution to fast search in hamming space which seems applicable to our problem.
 It provides a fast search when $r/n < 0.11$ or $r/n < 0.06$, where $r$ is the radius and $n$ is the number of bits.
@@
 It is intriguing that none of those algorithms is able to solve our scanning problem.
-The idea behind those computational geometry algorithms is roughly to split the search space in half each step, which would take $O(\log(H))$ to go through the whole space.
+The roughly idea behind those computational geometry algorithms is to split the search space in half each step, which would take $O(\log(H))$ to go through the whole space.
 But this approach does not work because of the high number of dimensions (i.e., 1,000) and because the hard locations' addresses are randomly sampled from the $\{0, 1\}^n$ space.
@@
 
-The OpenCL must be initialized, which just copies the address space's bitstrings to the GPU's memory.
+The OpenCL must be initialized, which basically copies the address space's bitstrings to the GPU's memory.
 Then, many scans may be executed with no necessity to upload the bitstrings again.
@@
 
-There are eight OpenCL kernels which explore differently the GPU architecture to improve performance.
+There are eight OpenCL kernels which differently explore the GPU architecture to improve performance.
 It is necessary because there are several GPU microarchitectures and a single kernel will never be optimal for all of them.
@@
 OpenCL specifies four levels of memory hierarchy for the GPUs: global memory, read-only memory, local memory, and private memory.
-The global memory and read-only memory are accessible by all workgroups, while each workgroup has its own local memory, accessible by its workers.
+The global memory and read-only memory are accessible by all workgroups, while each workgroup has its own local memory, shared by its workers.
 Finally, each worker has its own private memory.
@@
 They just do it with different approaches.
-For instance, \lstinline{single_scan0} calculates one distance between bitstrings per worker (Listing \ref{lst:single_scan0}); while \lstinline{single_scan2} uses a whole workgroup to calculate each distance, distributing each element of the 64-int integer array per worker (Listing \ref{lst:single_scan2}).
+For instance, \lstinline{single_scan0} calculates one distance between bitstrings per worker (Listing \ref{lst:single_scan0}); while \lstinline{single_scan2} uses a whole workgroup to calculate each distance, distributing each element of the 64-bit integer array per worker (Listing \ref{lst:single_scan2}).
 
@@
 caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan0}}.
-It calculates one distance per worker and let the GPU decide how to distribute this task between workgroups and workers.
+It calculates one distance per worker and lets the GPU decide how to distribute this task between workgroups and workers.
 It is the most straightforward kernel and does not explore any details of the GPU architecture.},
@@
 label={lst:single_scan5_unroll},
-caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan5_unroll}}. This kernel is exactly like \lstinline{single_scan5}, but it unrolls the last for since it has at most 5 loops.},
+caption={OpenCL kernel {\lstinline[columns=fixed]{single_scan5_unroll}}. This kernel is exactly like \lstinline{single_scan5}, but it unrolls the last \lstinline{for} since it has at most 5 loops.},
 ]
@@
 
-Each hard location has one integer of data per bit. For instance, each hard location of a 1,000 bits SDM has 1,000 bits. Those integers are stored in a counter.
-
-A counter is an array of integers which stores the data of all hard locations. So, the counter's array has $n \cdot H$ integers.
+A counter is an array of integers which stores the data of all hard locations.
+Each hard location has one integer of data per bit.
+For instance, each hard location of a 1,000 bits SDM has 1,000 integers, totalizing 1,000,000 integers.
+Those integers are stored in a counter.
+So, the counter's array has $n \cdot H$ integers.
 
@@
 It is not supposed to be a problem because, by default, each counter is a signed 32-bit integer that can store any number between -2,147,483,648 and 2,147,483,647, which means they will not overflow with fewer writes than $2^{31}-1$ divided by the average number of activated hard locations.
-For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard locations is 1,000 and it would require at least one million writes before being possible to a counter to overflow.
+For instance, when $n=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard locations is 1,000 and it would require at least one million writes before any overflow is possible.
 Note also that it would be more likely to saturate the memory before any overflow.

### 2018-03-22 Chapters/sdm-critical-distance.tex
SDM Grammar and typo fixes.

@@
 
-\caption{Results generated by the framework diverging from Kanerva's original Table 7.2.
+\caption{Results generated by the framework diverging from Kanerva's original Figure 7.3.
 Here we had a 1,000 bit, 1,000,000 hard location SDM with 10,000 random bitstrings written into it, which was also Kanerva's configuration.
@@
 
-\caption{Results generated by the framework similar to Kanerva's original Table 7.2.
+\caption{Results generated by the framework similar to Kanerva's original Figure 7.3.
 Here we have a 1,000 bit, 1,000,000 hard location SDM with (a) just 100 random bitstrings written into it and (b) steps of 1,000 random bitstrings written into it.

### 2018-03-22 Chapters/sdm-performance.tex
SDM more grammar and typo fixes
SDM Grammar and typo fixes.

@@
 
-Our second device was an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU. For its results, see Table \ref{tab:perf-ec2-imac}.
+Our second device was an iMac Retina 5K 27-inch 2017 with a 3.8GHz Intel core i5 processor, 8GB DDR4 RAM, and a Radeon Pro 580 8G CPU. For its results, see Table \ref{tab:perf-imac}.
 
-Beyond that, we were also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU (see Table \ref{tab:perf-p2}), and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU (see Table \ref{tab:perf-ec2-p3}).
+Beyond that, we were also running in state-of-the-art devices: (i) an Amazon EC2 p2.xlarge with Intel Xeon E5-2686v4 processor, 61GB DDR3 RAM, and NVIDIA K80 GPU (see Table \ref{tab:perf-ec2-p2}), and (ii) an Amazon EC2 p3.2xlarge with Intel Xeon E5-2686v4 processor, 488GB DDR3 RAM, and NVIDIA Tesla V100 GPU (see Table \ref{tab:perf-ec2-p3}).
 
@@
 
-The difference in which kernel is the best depends also on the SDM setup.
+The difference in which kernel is the best also depends on the SDM setup.
 The best kernel for the 1,000-bits SDM in the iMac 2017 was single\_scan5\_unroll with average scan time of 3.61ms; but, for the 256-bits SDM in the same device, it was single\_scan0 with average scan time of 3.03ms; while, for the 10,000-bits SDM in the same device, it was single\_scan6 with 10.96ms (see Table \ref{tab:perf-imac}).
@@
 Again, which one is faster depends on the SDM settings.
-In the iMac 2017, the faster scanner for a 1,000-bits SDM was the OpenCL scanner with \lstinline{single_scan5_unroll}, but for a 256-bit SDM was the threads scanner.
+In the iMac 2017, the fastest scanner for a 1,000-bits SDM was the OpenCL scanner with \lstinline{single_scan5_unroll}, but, for a 256-bit SDM, it was the threads scanner.
 
@@
 We can notice that Amazon EC2 p3.2xlarge and p2.xlarge's linear and thread scanners, both running on CPU, were much slower than the CPU of both the personal MacBook Pro and the iMac 2017.
-As Amazon EC2 are virtual machines with GPU devices, their CPU is shared with other virtual machines which reduces CPU power significantly.
-Hence, for both virtual machines we have tested, using the GPU seriously boosts performance, but using the CPU should be avoided.
-See Tables \ref{tab:perf-imac}
+As Amazon EC2 are virtual machines with GPU devices, their CPU is shared with other virtual machines which significantly reduces CPU power.
+Hence, for both virtual machines we have tested, using the GPU seriously boosts performance, but using the CPU should be avoided (see Table \ref{tab:perf-imac})
 

### 2018-03-22 Chapters/sdm2.tex
SDM Grammar and typo fixes.

@@
 It introduces many interesting mathematical properties of $n$-dimensional binary space that, in a memory model, seem to be remarkably psychologically plausible.
-Most notable among these are the tip-of-the-tongue phenomenon, conformity to Miller's magic number \citep{Linhares2011} and robustness against loss of neurons.
+Most notable among these are the tip-of-the-tongue phenomenon, conformity to Miller's magic number \citep{Linhares2011}, and robustness against loss of neurons.
 
@@
 
-Though both follow the triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which in 3-d Euclidean distance may be loosely interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{ and } d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
+Though both follow the triangle inequality ($d(A, C) \le d(A, B) + d(B, C)$), which in 3-d Euclidean distance may be loosely interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{ and } d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
 Hence, it makes no sense to say that $A$ is also close to $C$.
@@
 
-Every time write or read in SDM memory activates a number of addresses with close distance.
+Every write or read in SDM memory activates a number of addresses with close distance.
 The data is written in these activated addresses or read from them.
@@
 
-A random bitstring is generated with equal probability of $0$'s and $1$'s in each bit.
+A random bitstring is generated with equal probability of $0$'s and $1$'s for each bit.
 One can readily see that the average distance between two random bitstrings follows the binomial distribution with mean $n/2$ and standard deviation $\sqrt{n/4}$.
@@
 
-In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 1 million hard locations $N'=1,000,000$, an access-radius of 451 (within 1,000 hard locations in each circle) and 10 thousand writes of random bitstrings in the memory.
+In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 1 million hard locations ($N'=1,000,000$), an access-radius of 451 (within 1,000 hard locations in each circle), and 10 thousand writes of random bitstrings in the memory.
 As computer resources were very poor those days, Kanerva couldn't make a more generic analysis.

### 2018-03-23 Chapters/Chapter01.tex
TODO
Minor changes
quotes
quotes opressoras
Blockchain Review until bitcoin fork

@@
+% !TEX root = ../partial-blockchain.tex
+
 %************************************************
@@
 
+% TODO Lightning network, é uma solução de scaling? Parece que não.
+
 \chapter{Introduction}
 
-The main problem when one is trying to create a digital money is how to prevent double spending.
+\bigskip
+
+\begin{flushright}{\slshape
+    {We are watching History being made --- or History being repeated.}
+	\\ \medskip
+    --- David Collum, Cornell University, 2013}
+\end{flushright}
+\bigskip
+\bigskip
+
+The main problem when one is trying to create digital money is how to prevent double spending.
 As the money is digital, and copies can be made \textit{ad nauseam}, what can prevent counterfeiting?
 That is, what would prevent users from sending copies of the same money to two (or more) people?
-That is exactly the problem solved by Bitcoin and its underlying blockchain technology.
+That is precisely the problem solved by Bitcoin and its underlying blockchain technology.
 The current solution behind fiat money is having a single issuer, a central bank, and trusting the financial institutions.
@@
 The security of Bitcoin relies on digital signature technology and network agreement.
-While digital signature ensures ownership, i.e., the funds may only be spend by their owners, and nobody else; the network agreement both prevents double spending and ensures that all processed transactions have sufficient funds.
+While digital signature ensures ownership, i.e., the funds may only be spent by their owners, and nobody else; the network agreement both prevents double spending and ensures that all processed transactions have sufficient funds.
 In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution.
-When all these requirements are met, the funds are transfered.
+When all these requirements are met, the funds are transferred.
 
 According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending.
-The no central point of trust and predictable money supply together with a clever solution to the double spending problem is what separates Bitcoin from the previous e-cash philosophies.
+The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the previous e-cash philosophies.
 % literature
@@
 Bitcoin provides interesting incentives to all players, namely the users and the miners.
-On one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction --- (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
-On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes and they may also receive the fees of unconfirmed transactions.
-These incentives have kept the Bitcoin network up and running since 2009 with barely no interruptions (99.99\% uptime).
+On the one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction --- (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
+On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes, and they may also receive the fees of unconfirmed transactions.
+These incentives have kept the Bitcoin network up and running since 2009 with barely any interruptions (99.99\% uptime).
 
 Since 2009, Bitcoin has been growing and becoming more and more used all around the world.
-It started as an experiment based in a seminal work by \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a highly volatile \$192 billion market capitalization, as of this writing \citep{coinmarketcapbtc}.
+It started as an experiment based on a seminal work by \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a highly volatile \$192 billion market capitalization, as of this writing \citep{coinmarketcapbtc}.
 There are hundreds of companies investing in different uses of the technology, from exchanges to debit cards, and billions of dollars being invested in the new markets based on Bitcoin's technology.
@@
 Thus, the more Bitcoin becomes popular, the more saturated the network is.
-Network saturation has many side effects and may affect the players incentive to keep the network running.
+Network saturation has many side effects and may affect the players' incentive to keep the network running.
 The transaction fees have to be increased to compete for faster confirmation.
@@
 
-Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
-Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behaviour, and it is also seen as a problem to be solved.
+Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
+Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior, and it is also seen as a problem to be solved.
 The more decentralized, the more trustworthy Bitcoin is.
 
-Generating new transactions in Bitcoin has a tiny cost, because one only has to generate the transaction itself, digitally sign it, and propagate in the Bitcoin network.
-On one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks.
+Generating new transactions in Bitcoin has a tiny cost because one only has to generate the transaction itself, digitally sign it, and propagate in the Bitcoin network.
+On the one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks.
 One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network.
 
-The number of ideas and publications focusing in improving Bitcoin's design and overcoming those challenges is increasing every day.
+The number of ideas and publications focusing on improving Bitcoin's design and overcoming those challenges is increasing every day.
 Many of these proposals are organized into BIPs (Bitcoin Improvement Proposals) which are discussed and implemented by the community; while others come in the form of whitepapers and alternative software forks (which would include the need of a protocol upgrade).
@@
 
-After the launch of Bitcoin, more than 1,000 other cryptocurrencies has been created (REF).
+After the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created (REF).
 In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain.
 Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}.
-We are specially interested in one of them: Iota.
+We are especially interested in one of them: Iota.
 
@@
 It has neither mining nor confirmation blocks and transaction fees.
-Each transaction has its own proof-of-work, and is used to confirm other transactions forming a directed acyclic graph.
+Each transaction has its own proof-of-work and is used to confirm other transactions, forming a directed acyclic graph.
 In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions.
@@
 The scaling problem of tangle is exactly the opposite of Bitcoin's.
-It must have at least a given number of transaction per seconds, otherwise the transactions are not confirmed and the cryptocurrency does not work.
+It must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work.
 
@@
 In \citeauthor{nakamoto2008bitcoin}'s (2009) seminal paper, there is no distinction between bitcoin and blockchain.
-They are just one thing which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money in the internet, in which all users can agree as to whom is entitled to which funds.
+They are just one thing which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money on the internet, in which all users can agree as to whom is entitled to which funds.
 
@@
 The mechanism that assures the immutability is proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network.
+The network agrees that work should be done in the block at the longest chain in the blockchain.
 
@@
 Thus, the goal is to adjust the difficulty every 14 days.
-If it takes less than 14 days to find 2016 blocks, that means the network's hash power has increased, thus the difficulty is increased.
-If it takes more than 14 days to find 2016 blocks, it means the network's hash power has decreased, thus the difficulty is decreased.
+If it takes less than 14 days to find 2016 blocks, that means the network's hash power has increased; thus the difficulty is increased.
+If it takes more than 14 days to find 2016 blocks, it means the network's hash power has decreased; thus the difficulty is decreased.
 
@@
 A block is found when a solution to the proof-of-work is found.
-Two miners may find blocks in a small interval of time.
-In this case, both will propagate their blocks and the network will randomly choose one of them as the next block.
+When a new block is found, it indirectly confirms all the previous blocks in the chain and their transactions.
+It may happen that two miners find two different blocks in a small interval of time.
+In this case, both miners will propagate their blocks, and the network will randomly choose one of them as the next block.
 This phenomenon is called a fork.
-Thus, when the next block is found, one of those blocks will be confirmed and the other will be ignored and referred to as an orphan block.
-The network agrees that work should be done in the block at the longest chain in the blockchain.
-When a new block is found, it indirectly confirms all the previous blocks in the chain and their transactions.
+Thus, when the next block is found, one of those blocks will be confirmed, and the other will be ignored and referred to as an orphan block.
+As blocks confirm transactions, all the transactions in the orphan block which have not been confirmed by another block will return to the unconfirmed transaction pool.
+Hence, it is not safe to accept a transaction when it is confirmed by only one block.
+It is the idea behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.
 
@@
 
-For instance, no group with more than 50\% of the network's hash power agreed into increasing the maximum block size to increase the number of transactions confirmed by a block, and thus increasing the network's capacity.
-Thus, the capacity remains the same and the community has been discussing the issue in search of a consensus.
+A practical example of a network disagreement is the increase of the block size.
+No group with more than 50\% of the network's hash power agreed into increasing the maximum block size to increase the number of transactions confirmed by a block and thus increasing the network's capacity.
+Hence, the capacity remains the same, and the community has been discussing the issue in search of a consensus.
 
@@
 But, in order to certify ownership of that output, the miner receives instructions in the form of a script with the rules that lead to the desired unspent output address (and this one is displayed as the input address by those websites).
-Because this process requires a digital signature, only the holder of the corresponding private key is able to sign such a transaction.
-
-Next we will do a mathematical analysis of Bitcoin in order to deeply understand its minings properties, how a fork would affect the network, and its security against attackers.
+Because this process requires a digital signature, only the holder of the corresponding private key is able to sign such transaction.
+
+Next, we will do a mathematical analysis of Bitcoin in order to better understand its minings properties, how a fork would affect the network and its security against attackers.
 
 \chapter{Analysis of Bitcoin}
+
+The primary objective of this chapter is to increase our understanding of the Bitcoin through mathematical tools.
 
@@
 When $x \ne y$ and $h(x) = h(y)$, we say that $x$ and $y$ are a collision.
-A hash function is considered to be safe when it is unknown how to find a collision of a given hash.
+A hash function is considered to be safe when it is unknown how to quickly find a collision of a given hash, i.e., one has to check all possible values until the correct one is found (known as the brute-force attack).
 
@@
 The latter has $n=256$ and applies \textit{SHA-256} twice.
-
-For further information, see \citet{gilbert2003security, dobbertin1996ripemd}.
+The first is used in transactions' scripts and the latter in the mining algorithm.
+For both hash functions, it is infeasible to run a brute-force attack because it would demand, on average, either $2^{160}$ or $2^{256}$ trials, and those would take a tremendous amount of time even for the fastest known processors.
+
+For further information about hash functions, see \citet{gilbert2003security, dobbertin1996ripemd}.
 
@@
 
-For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ (or $[0.51, 30.76]$ in minutes). Thus, 90\% of the time the intervals between blocks are between 30 seconds and 30 minutes.
+For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ (or $[0.51, 30.76]$ in minutes).
+Thus, 90\% of the time the intervals between blocks are between 30 seconds and 30 minutes, with average of 10 minutes.
 
@@
 Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on.
-Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$.
+Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$ which is the total time to find the next $n$ blocks.
 As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{\eta}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{\eta})$.
@@
 Many exchanges require at least six confirmations in order to accept a deposit in Bitcoin.
-So, for $n=6$, $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$.
+So, for $n=6$, $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$, i.e., only 55\% of the deposits will be accepted in one hour.
 The symmetrical confidence interval with $\alpha=10\%$ is $[27, 105]$ in minutes.
 Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted --- assuming that your transaction will be confirmed in the very next block.
-The pdf of $Y_6$ is shown in Figure \ref{fig-bitcoin-time-6-blocks}.
-The 10\% symmetrical confidence interval is shown in the white area.
+The pdf of $Y_6$ is shown in Figure \ref{fig-bitcoin-time-6-blocks}, in which the 10\% symmetrical confidence interval is shown in the white area.
+The average total time of six confirmations is $\mathbf{E}(Y_6) = 6 \cdot 600 = 3600 = 60 \text{ minutes}$.
 
@@
 \section{Orphan blocks}
-
 % Probabilidade de orfão
@@
 Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{\eta}}$.
-For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.
+
+Bitcoin peer-to-peer network is a gossip network, where miners are semi-randomly connected to each other, and each miner sends all information it receives to all its peers.
+According to \citet{decker2013information}, the average time for a new block propagate over the network is 12.6 seconds, while the 95\% percentile is 40 seconds, which indicates a long-tail distribution.
+\citet{bitcoinstats} has measured the propagation time between 2013 and 2017.
+During 2017, the worst daily 90\% percentile was 21 seconds.
+Notice that both results may not be contradictory because Bitcoin network is continuously evolving.
+
+For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$, which is almost 0.5\%.
+I would say that a node with 10\% of the total hash rate would be well connected and it would take less time to receive the update, so, the probability would be even smaller than 0.5\%.
+
+Another important factor is that, as Bitcoin is open-source, miners are free to change the gossip algorithm, which leads to the network incentives.
+See \citet{babaioff2012bitcoin} for an analysis of the incentives to miners forward new blocks and transactions in the network.
+
+For further information about gossip algorithms, see \citet{shah2009gossip}.
 
@@
 
+\citet{karame2012two}
+
 There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.
@@
 
+% TODO BIP's and signaling bit --- see BIP 9 and BIP 135.
+
 When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork.

### 2018-03-29 kanerva-answer.tex
Add kanerva-answer.tex.

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfig}

\begin{document}

Let $x$ and $y$ be random bitstrings and $n$ be the SDM's number of bits; let $x_i$ and $y_i$ be the $i$th bit of $x$ and $y$, respectivelly; and $d(x, y)$ be the Hamming distance.

Applying the law of total probability:

\begin{align}
P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) <= r)
\end{align}

We also know that

\begin{align}
P(x_i = y_i | d(x, y) = k) &= \frac{n-k}{n} \\
P(d(x, y) = k | d(x, y) \le r) &= \frac{\binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
\end{align}

Hence,

\begin{align}
P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \frac{n-k}{n} \binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
\end{align}

Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,

\begin{align}
P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}
\end{align}

This equation is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.

\section{Counter bias}

The bias begins in the counters. Let's analyze $i$th counter of a hard location.

Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$), $h$ be the average number of activated hard locations ($h=1,071.85$), $H$ be the number of hard locations ($H=1,000,000$), and $\text{addr}_i$ be the $i$th bit of the hard location's address. Then,

\begin{align}
P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}i = 0) = \sum_{k=1}^{\theta} X_k > \frac{sh}{2H}
\end{align}

Where $\theta = \frac{sh}{H}$ is the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$). Thus, $\sum_{k=1}^{\theta} X_k \sim \text{Binomial} \left( \theta, p \right)$.

Finally, $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = \text{Binomial.cdf} \left( \frac{sh}{2H} \right)$.

For ``random at x'', $p=0.5$, $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters. But, for ``x at x'', $p=0.5529$ and the probabilities depend on $s$. For $s=10,000$, they are equal to 0.6362. For $s=20,000$, they are equal to 0.6888. For $s=30,000$, they are equal to 0.7268. The more random bitstrings are written into the memory, the more a hard location to point to itself. See Figure \ref{fig:sdm-corr-prob} --- and notice that I still have to figure out why the mean is correct, but the standard deviation is not. As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').

\begin{figure}[h!]
  \centering
  \subfloat[``random at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/random_at_x_counters.png}}
  \subfloat[``x at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_counters.png}}

  \caption{Autocorrelation in the counters in autoassociative memories (``x at x''). The histogram was obtained through simulation. The red curve is the theoretical distribution.}
  \label{fig:sdm-corr-prob}
\end{figure}

Not let's analyze the value of the counters for ``x at x''. In write operation, the counters are incremented for every bit 1 and decremented for every bit 0. So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location, $Y_i = \sum_{k=1}^{\theta} X_k$ bits 1, and $\theta - Y_i$ bits 0. Thus, when $\text{addr}_i = 1$, $\text{cnt}_i = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$. And, when $\text{addr}_i = 0$, $\text{cnt}_i = \theta - 2Y_i$.

As $Y_i \sim \text{Binomial}(\theta, p)$, for $\theta$ sufficiently large, we may approximate $Y_i \sim \mathcal{N}(\mu = \theta p, \sigma^2 = \theta p (1-p))$. Hence, as $\text{cnt}_i = 2Y_i - \theta$, $\mathbf{E}[2Y_i - \theta] = 2 \mathbf{E}[Y_i] - \theta$, and $\mathbf{V}[2Y_i - \theta] = 4 \mathbf{V}[Y_i]$, then,

\begin{align}
\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1) \theta, \sigma^2 = 4 \theta p (1-p)) \\
\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1) \theta, \sigma^2 = 4 \theta p (1-p))
\end{align}

In our case, $p=0.5529$, $s=10,000$, $h=1,071.85$, and $H=1,000,000$, so $\theta = 10.7185$ and $\text{cnt}_i \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.5985)$. For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = \theta$. See Figure \ref{fig:sdm-corr-counters}.

\begin{figure}[h!]
  \centering
  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}

  \caption{Autocorrelation in the counters in autoassociative memories (``x at x''). The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
  \label{fig:sdm-corr-counters}
\end{figure}

\section{Read bias}

Now that we know the distribution of $\text{cnt}_i | \text{addr}_i$, we may go to the read operation. During the read operation, on average, $h$ hard locations are activated and their counters are summed up. So, for the $i$th bit,

\begin{align}
\text{acc}_i = \sum_{k=1}^{h} \text{cnt}_k
\end{align}

Let $\eta$ be the reading address and $\eta_i$ the $i$th bit of it. Then,

\begin{align}
\left[ \text{acc}_i|\eta_i=1 \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] \\
\left[ \text{acc}_i|\eta_i=0 \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=0 \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=1 \right]
\end{align}

We may analyze only one case, because the probability of the other is exactly the same.

Each sum is a sum of normally distributed random variables, so

\begin{align}
\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] &\sim \mathcal{N}(\mu = (2p-1) \theta ph, \sigma^2 = 4\theta p(1-p) ph) \label{eqn:sdm-eta1-addr1} \\
\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] &\sim \mathcal{N}(\mu = -(2p-1) \theta (1-p)h, \sigma^2 = 4\theta p(1-p) (1-p)h) \label{eqn:sdm-eta1-addr0}
\end{align}

In our case, $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=672.12, \sigma^2=6281.00)$, and $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=-543.49, \sigma^2=5078.99)$. See Figure \ref{fig:sdm-read-sums} --- we can notice there a small but significant difference between the theoretical and the simulated mean.

\begin{figure}[h!]
  \centering
  \subfloat[Equation \ref{eqn:sdm-eta1-addr1}  ($\text{addr}_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr1.png}}
  \subfloat[Equation \ref{eqn:sdm-eta1-addr0} ($\text{addr}_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr0.png}}

  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
  \label{fig:sdm-read-sums}
\end{figure}

Hence,

\begin{align}
\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \theta h, \sigma^2 = 4 \theta p (1-p) h) \label{eqn:sdm-eta1} \\
\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \theta h, \sigma^2 = 4 \theta p (1-p) h) \label{eqn:sdm-eta0}
\end{align}

In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 11359.99)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 11359.99)$. See Figure \ref{fig:sdm-read} --- we can notice that the small difference in the means from Figure \ref{fig:sdm-read-sums} has propagated to these images.

\begin{figure}[h!]
  \centering
  \subfloat[Equation \ref{eqn:sdm-eta1}  ($\eta_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1.png}}
  \subfloat[Equation \ref{eqn:sdm-eta0} ($\eta_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta0.png}}

  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
  \label{fig:sdm-read}
\end{figure}

Finally,

\begin{align}
P(wrong) &= P(\text{acc}_i < 0 | \eta_i = 1) \cdot P(\eta_i = 1) + P(\text{acc}_i > 0 | \eta_i = 0) \cdot P(\eta_i = 0) \\
    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{1-\mathcal{N}_{\eta_i=0}.\text{cdf}(0)}{2} \\
    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} \\
    &= \mathcal{N}_{\eta_i=1}.\text{cdf}(0)
\end{align}

In our case, $P(wrong) = 0.1137518032308093$.

In order to check this probability, I have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculate the distance from the result of a single read. As the $P(wrong) = 0.11375$, I expected to get an average distance of 113.75 with a standard deviation of 10.04. See Figure \ref{fig:sdm-read-random-bs} --- We can notice a big difference between the theoretical model and the simulation. Using $\mu=102$ and $\sigma=121$, the curves match. I'm still looking for the mistake in the equations. I believe the problem is in Equation \ref{eqn:sdm-eta1-addr1}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.75\textwidth]{./images02/autocorrelation/read-random-bs.png}

  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
  \label{fig:sdm-read-random-bs}
\end{figure}

\end{document}

### 2018-04-03 Chapters/Conclusion.tex
anomaly:  1st inclusion in thesis.

@@
 
-There are three contributions made on SDM:  First, I have illuminated a discrepancy between Kanerva’s theoretical model and the real system dynamics; Also, we have seen that pattern classification through supervised learning is possible without presuming any new SDM mechanism.
+There are at least three contributions\footnote{As many issues have been explored in less detail, it might be advisable to leave it to history to decide whether these explorations were actually contributions.} made on SDM:  First, I have illuminated a discrepancy between Kanerva’s theoretical model and the real system dynamics; Also, we have seen that pattern classification through supervised learning is possible without presuming any new SDM mechanism.
 This is in contrast with the literature, that presumes additional mechanisms, like genetic algorithms, to account for supervised learning.

### 2018-04-03 Chapters/sdm-critical-distance.tex
anomaly:  1st inclusion in thesis.

@@
 
-As we can easily notice, we got exactly the same curve as Kanerva.
-This question has been intriguing us since then, and we are still looking for a more analytic explanation than merely interference from the other written attractors.
+As one can easily notice, we have got exactly the same curve as Kanerva.
+Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the expected $n/2$ equator distance.
+This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors.
+Let us turn back to mathematics to study this anomaly.
+
+\section{A deviation from the equator distance?}
+
+Dr. Kanerva writes:
+
+\begin{quote}
+    You have done an incredibly thorough analysis of SDM.
+I like the puzzle in your message and believe that your simulations are correct and to be learned from.
+So what to make of the difference compared to my Figure 7.3 (and your Figure \ref{fig:kanerva-figure-73-calculated})?
+I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
+You say it in \\
+
+   "Our results show that the theoretical prediction is not accurate.
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
+   (Figure \ref{fig:sdm-10000w-table-7-2})." \\
+
+   I think that is correct.
+It also brings to mind a comment Louis Jaeckel made when we worked at NASA Ames.
+He pointed out that autoassociative storage (each vector is stored with itself as the address) introduces autocorrelation that my formula for Figure 7.2 did not take into account.
+When we read from memory, each stored vector exerts a pull toward itself, which also means that each bit of a retrieved vector is slightly biased toward the same bit of the read address, regardless of the read address.
+We never worked out the math.
+\end{quote}
+
+This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
+Imagine the `counter's eye view':  each individual counter `likes' to write its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
+
+Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
+Whilst the probability of a shared bit-value between two random bitstrings is $1/2$, an address only activates hard-locations close to it.
+
+So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:
+
+\begin{align}
+P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
+\end{align}
+
+We also know that
+
+\begin{align}
+P(x_i = y_i | d(x, y) = k) &= \frac{n-k}{n} \\
+P(d(x, y) = k | d(x, y) \le r) &= \frac{\binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
+\end{align}
+
+Hence,
+
+\begin{align}
+P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \frac{n-k}{n} \binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
+\end{align}
+
+Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,
+
+\begin{align}
+P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}
+\end{align}
+
+This equation is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
 

### 2018-04-03 kanerva-answer.tex
anomaly:  1st inclusion in thesis.
SDM Answer to Kanerva updated.

@@
 \begin{align}
-P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) <= r)
+P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
 \end{align}
@@
 
+Hence, let $Z$ be the number of activated hard location with the same bit as the reading address, then: $Z = \sum_{i=1}^{h} X_i$, where $\mathbf{E}[h] = h$, $\mathbf{V}[h] = H p_1 (1-p_1)$, $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and $X_i \sim \text{Bernoulli}(p)$.
+By the central limit theorem, $Z$ is normally distributed with $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[W_i | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[X_i|h]] + \mathbf{V}[\mathbf{E}[X_i|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+See Figure \ref{fig:sdm-same-bit-histogram} for a comparison between the theoretical model and a simulation.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.75\textwidth]{./images02/autocorrelation/same-bit-histogram.png}
+
+  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
+  \label{fig:sdm-same-bit-histogram}
+\end{figure}
+
 \section{Counter bias}
 
-The bias begins in the counters. Let's analyze $i$th counter of a hard location.
+The bias begins in the counters. Let's analyze the $i$th counter of a hard location.
 
 Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$), $h$ be the average number of activated hard locations ($h=1,071.85$), $H$ be the number of hard locations ($H=1,000,000$), and $\text{addr}_i$ be the $i$th bit of the hard location's address.
-Then,
+
+Let $\theta = \frac{sh}{H}$ be the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$).
+Thus, $Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1, \sigma_1^2)$ with $\mu_1 = p \theta$, and $\sigma_1^2 = p(1-p) \theta + p^2 s^2 p_1 (1 - p_1) / H$.
+
+During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0.
+So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location, $Y_i$ bits 1, and $\theta - Y_i$ bits 0.
+Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$; and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.
+
+Hence, as $\text{cnt}_i = 2Y_i - \theta$, $\mathbf{E}[2Y_i - \theta] = 2 \mathbf{E}[Y_i] - \theta$, and $\mathbf{V}[2Y_i - \theta] = 4 \mathbf{V}[Y_i]$, then,
 
 \begin{align}
-P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}i = 0) = \sum_{k=1}^{\theta} X_k > \frac{sh}{2H}
+\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_2 = (2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2) \\
+\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu_2 = -(2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)
 \end{align}
 
-Where $\theta = \frac{sh}{H}$ is the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$).
-Thus, $\sum_{k=1}^{\theta} X_k \sim \text{Binomial} \left( \theta, p \right)$.
+In our case, $p=0.5529$, $s=10,000$, $h=1,071.85$, and $H=1,000,000$, so $\theta = 10.7185$ and $\text{cnt}_i \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7294)$.
+For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.8255$.
+See Figure \ref{fig:sdm-corr-counters}.
 
-Finally, $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = \text{Binomial.cdf} \left( \frac{sh}{2H} \right)$.
+\begin{figure}[h!]
+  \centering
+  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
+  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
 
-For ``random at x'', $p=0.5$, $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters.
-But, for ``x at x'', $p=0.5529$ and the probabilities depend on $s$.
-For $s=10,000$, they are equal to 0.6362.
-For $s=20,000$, they are equal to 0.6888.
-For $s=30,000$, they are equal to 0.7268.
+  \caption{The value of the counters after $s=10,000$ writes shows the autocorrelation in the counters in autoassociative memories (``x at x'').
+The histogram was obtained through simulation.
+The red curve is the theoretical normal distribution.}
+  \label{fig:sdm-corr-counters}
+\end{figure}
+
+Finally,
+
+\begin{align}
+P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 1 - \mathcal{N}.\text{cdf}(0)
+\end{align}
+
+For ``random at x'', $p=0.5$ implies $\mu_2 = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
+
+However, for ``x at x'', $p=0.5529$ and the probabilities depend on $s$.
+For $s=10,000$, they are equal to 0.6354.
+For $s=20,000$, they are equal to 0.6867.
+For $s=30,000$, they are equal to 0.7232.
 The more random bitstrings are written into the memory, the more a hard location to point to itself.
@@
   \label{fig:sdm-corr-prob}
-\end{figure}
-
-Not let's analyze the value of the counters for ``x at x''.
-In write operation, the counters are incremented for every bit 1 and decremented for every bit 0.
-So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location, $Y_i = \sum_{k=1}^{\theta} X_k$ bits 1, and $\theta - Y_i$ bits 0.
-Thus, when $\text{addr}_i = 1$, $\text{cnt}_i = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$.
-And, when $\text{addr}_i = 0$, $\text{cnt}_i = \theta - 2Y_i$.
-
-As $Y_i \sim \text{Binomial}(\theta, p)$, for $\theta$ sufficiently large, we may approximate $Y_i \sim \mathcal{N}(\mu = \theta p, \sigma^2 = \theta p (1-p))$.
-Hence, as $\text{cnt}_i = 2Y_i - \theta$, $\mathbf{E}[2Y_i - \theta] = 2 \mathbf{E}[Y_i] - \theta$, and $\mathbf{V}[2Y_i - \theta] = 4 \mathbf{V}[Y_i]$, then,
-
-\begin{align}
-\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1) \theta, \sigma^2 = 4 \theta p (1-p)) \\
-\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1) \theta, \sigma^2 = 4 \theta p (1-p))
-\end{align}
-
-In our case, $p=0.5529$, $s=10,000$, $h=1,071.85$, and $H=1,000,000$, so $\theta = 10.7185$ and $\text{cnt}_i \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.5985)$.
-For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = \theta$.
-See Figure \ref{fig:sdm-corr-counters}.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
-  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
-
-  \caption{Autocorrelation in the counters in autoassociative memories (``x at x''). The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
-  \label{fig:sdm-corr-counters}
 \end{figure}
@@
 \begin{align}
-\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] &\sim \mathcal{N}(\mu = (2p-1) \theta ph, \sigma^2 = 4\theta p(1-p) ph) \label{eqn:sdm-eta1-addr1} \\
-\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] &\sim \mathcal{N}(\mu = -(2p-1) \theta (1-p)h, \sigma^2 = 4\theta p(1-p) (1-p)h) \label{eqn:sdm-eta1-addr0}
+\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma^2 = \sigma_1^2 ph) \label{eqn:sdm-eta1-addr1} \\
+\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] &\sim \mathcal{N}(\mu = -\mu_2 (1-p)h, \sigma^2 = \sigma_1^2 (1-p)h) \label{eqn:sdm-eta1-addr0}
 \end{align}

### 2018-04-03 kanerva-answer.tex
trying out

--- 2018-03-29 kanerva-answer.tex
@@
 
-Let $x$ and $y$ be random bitstrings and $n$ be the SDM's number of bits; let $x_i$ and $y_i$ be the $i$th bit of $x$ and $y$, respectivelly; and $d(x, y)$ be the Hamming distance.
+[Thesis text....] ...As one can easily notice, we have got exactly the same curve as Kanerva.
+Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the expected $n/2$ equator distance.
+This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors.
+Let us turn back to mathematics to study this anomaly.
 
-Applying the law of total probability:
+\section{A deviation from the equator distance?}
+
+Dr. Kanerva writes:
+
+\begin{quote}
+    You have done an incredibly thorough analysis of SDM.
+I like the puzzle in your message and believe that your simulations are correct and to be learned from.
+So what to make of the difference compared to my Figure 7.3 (and your Figure \ref{fig:kanerva-figure-73-calculated})?
+I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
+You say it in \\
+
+   "Our results show that the theoretical prediction is not accurate.
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
+   (Figure \ref{fig:sdm-10000w-table-7-2})." \\
+
+   I think that is correct.
+It also brings to mind a comment Louis Jaeckel made when we worked at NASA Ames.
+He pointed out that autoassociative storage (each vector is stored with itself as the address) introduces autocorrelation that my formula for Figure 7.2 did not take into account.
+When we read from memory, each stored vector exerts a pull toward itself, which also means that each bit of a retrieved vector is slightly biased toward the same bit of the read address, regardless of the read address.
+We never worked out the math.
+\end{quote}
+
+This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
+
+Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
+Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
+
+So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:
 
 \begin{align}
-P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) <= r)
+P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
 \end{align}
@@
 This equation is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+
+Linhares here...
+
+Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$
+
+is it?
+
+%\begin{align}
+
+  $\frac{n-k}{n} \binom{n}{k} = \frac{n-k}{n} \frac{n!}{k!(n-k)!}$ \\
+  $ = \frac{(n-k)(n-1)!}{k!(n-k)!} $
+  $ = \frac{(n-1)!}{k!(n-k-1)!}$
+
+%\end{align}
 

### 2018-04-03 kanerva-answer.tex
test

--- 2018-04-03 kanerva-answer.tex
@@
 
-Let $x$ and $y$ be random bitstrings and $n$ be the SDM's number of bits; let $x_i$ and $y_i$ be the $i$th bit of $x$ and $y$, respectivelly; and $d(x, y)$ be the Hamming distance.
+[Thesis text....] ...As one can easily notice, we have got exactly the same curve as Kanerva.
+Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the expected $n/2$ equator distance.
+This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors.
+Let us turn back to mathematics to study this anomaly.
 
-Applying the law of total probability:
+\section{A deviation from the equator distance?}
+
+Dr. Kanerva writes:
+
+\begin{quote}
+    You have done an incredibly thorough analysis of SDM.
+I like the puzzle in your message and believe that your simulations are correct and to be learned from.
+So what to make of the difference compared to my Figure 7.3 (and your Figure \ref{fig:kanerva-figure-73-calculated})?
+I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
+You say it in \\
+
+   "Our results show that the theoretical prediction is not accurate.
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
+   (Figure \ref{fig:sdm-10000w-table-7-2})." \\
+
+   I think that is correct.
+It also brings to mind a comment Louis Jaeckel made when we worked at NASA Ames.
+He pointed out that autoassociative storage (each vector is stored with itself as the address) introduces autocorrelation that my formula for Figure 7.2 did not take into account.
+When we read from memory, each stored vector exerts a pull toward itself, which also means that each bit of a retrieved vector is slightly biased toward the same bit of the read address, regardless of the read address.
+We never worked out the math.
+\end{quote}
+
+This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
+
+Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
+Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
+
+So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:
 
@@
 Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,
+
+THIS STEP IS A LITTLE WTF TO ME...
+
+$\frac{n-k}{n} \binom{n}{k} = \frac{n-k}{n} \binom{n}{n-k} = \frac{n-k}{n} \frac{n!}{k!(n-k)!}$ \\
+
+$ = \frac{(n-k)(n-1)!}{k!(n-k)!} $
+
+$ = \frac{(n-1)!}{k!(n-k-1)!}$
 

### 2018-04-03 kanerva-answer.tex
SDM Answer to Kanerva updated.

@@
 \begin{align}
-\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma^2 = \sigma_1^2 ph) \label{eqn:sdm-eta1-addr1} \\
-\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] &\sim \mathcal{N}(\mu = -\mu_2 (1-p)h, \sigma^2 = \sigma_1^2 (1-p)h) \label{eqn:sdm-eta1-addr0}
+\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma_3^2 = \sigma_2^2 ph + \mu_2^2 h p(1-p)) \label{eqn:sdm-eta1-addr1} \\
+\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] &\sim \mathcal{N}(\mu_3 = -\mu_2 (1-p)h, \sigma_3^2 = \sigma_2^2 (1-p)h + \mu_2^2 h p (1-p)) \label{eqn:sdm-eta1-addr0}
 \end{align}
@@
 \begin{align}
-\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \theta h, \sigma^2 = 4 \theta p (1-p) h) \label{eqn:sdm-eta1} \\
-\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \theta h, \sigma^2 = 4 \theta p (1-p) h) \label{eqn:sdm-eta0}
+\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta1} \\
+\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta0}
 \end{align}
 
-In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 11359.99)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 11359.99)$.
+In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12181.95)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12181.95)$.
 See Figure \ref{fig:sdm-read} --- we can notice that the small difference in the means from Figure \ref{fig:sdm-read-sums} has propagated to these images.
@@
 
-In our case, $P(wrong) = 0.1137518032308093$.
+In our case, $P(wrong) = 0.12193065104931683$.
 
 In order to check this probability, I have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculate the distance from the result of a single read.
-As the $P(wrong) = 0.11375$, I expected to get an average distance of 113.75 with a standard deviation of 10.04.
+As the $P(wrong) = 0.12193$, I expected to get an average distance of 121.93 with a standard deviation of 10.34.
 See Figure \ref{fig:sdm-read-random-bs} --- We can notice a big difference between the theoretical model and the simulation.
-Using $\mu=102$ and $\sigma=121$, the curves match.
+Using $\mu=221$ and $\sigma^2=168$, the curves match.
 I'm still looking for the mistake in the equations.

### 2018-04-03 kanerva-answer.tex
toying around with text

--- 2018-04-03 kanerva-answer.tex
@@
 
-This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
+This is an important observation.
+A hard location is activated because it shares many dimensions with the items read from or written onto it.
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
 
@@
 Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,
-
-THIS STEP IS A LITTLE WTF TO ME...
-
-$\frac{n-k}{n} \binom{n}{k} = \frac{n-k}{n} \binom{n}{n-k} = \frac{n-k}{n} \frac{n!}{k!(n-k)!}$ \\
-
-$ = \frac{(n-k)(n-1)!}{k!(n-k)!} $
-
-$ = \frac{(n-1)!}{k!(n-k-1)!}$
 

### 2018-04-03 kanerva-answer.tex
SDM Answer to Dr K.

@@
 Hence, let $Z$ be the number of activated hard location with the same bit as the reading address, then: $Z = \sum_{i=1}^{h} X_i$, where $\mathbf{E}[h] = h$, $\mathbf{V}[h] = H p_1 (1-p_1)$, $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and $X_i \sim \text{Bernoulli}(p)$.
-By the central limit theorem, $Z$ is normally distributed with $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[W_i | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[X_i|h]] + \mathbf{V}[\mathbf{E}[X_i|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+By the central limit theorem, $Z$ is normally distributed.
+
+Applying the law of total average and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+
 See Figure \ref{fig:sdm-same-bit-histogram} for a comparison between the theoretical model and a simulation.
@@
 
-Let $\theta = \frac{sh}{H}$ be the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$).
-Thus, $Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1, \sigma_1^2)$ with $\mu_1 = p \theta$, and $\sigma_1^2 = p(1-p) \theta + p^2 s^2 p_1 (1 - p_1) / H$.
+Let $\theta = \frac{sh}{H}$ be the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$). Thus,
+
+\begin{align}
+Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1 = p \theta, \sigma_1^2 = p(1-p) \theta + p^2 s^2 p_1 (1 - p_1) / H)
+\end{align}
 
@@
 For $s=30,000$, they are equal to 0.7232.
-The more random bitstrings are written into the memory, the more a hard location to point to itself.
+The more random bitstrings are written into the memory, the more the hard locations point to themselves.
 See Figure \ref{fig:sdm-corr-prob} --- and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
@@
 
-Let $\eta$ be the reading address and $\eta_i$ the $i$th bit of it. Then,
+Let $\eta$ be the reading address and $\eta_i$ the $i$th bit of it.
+Then, let's split the $h$ activated hard locations into two groups: (i) the ones with the same bit as $\eta_i$ with $ph$ hard locations, and (ii) the ones with the opposite bit as $\eta_i$ with $(1-p)h$ hard locations.
 
 \begin{align}
-\left[ \text{acc}_i|\eta_i=1 \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] \\
-\left[ \text{acc}_i|\eta_i=0 \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=0 \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=1 \right]
+\left[ \text{acc}_i|\eta_i \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=\eta_i \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_i \right]
 \end{align}
-
-We may analyze only one case, because the probability of the other is exactly the same.
 
@@
 \begin{align}
-\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma_3^2 = \sigma_2^2 ph + \mu_2^2 h p(1-p)) \label{eqn:sdm-eta1-addr1} \\
-\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k=0 \right] &\sim \mathcal{N}(\mu_3 = -\mu_2 (1-p)h, \sigma_3^2 = \sigma_2^2 (1-p)h + \mu_2^2 h p (1-p)) \label{eqn:sdm-eta1-addr0}
+\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma_3^2 = \sigma_2^2 ph + \mu_2^2 h p(1-p)) \label{eqn:sdm-eta1-addr1} \\
+\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] &\sim \mathcal{N}(\mu_3 = -\mu_2 (1-p)h, \sigma_3^2 = \sigma_2^2 (1-p)h + \mu_2^2 h p (1-p)) \label{eqn:sdm-eta1-addr0}
 \end{align}

### 2018-04-03 Chapters/Chapter02.tex
Swedish guys 1

@@
 
-The classification algorithm had some problems classifying the letters ``i'' and ``l''.
-It may have happened because of the low distance between them.
-In this case, SDM could not differentiate the details --- it has only considered the big picture.
+The classification algorithm had some problems classifying the patterns ``i'' and ``l'', due to the low distance between them.
+In this case, SDM could not discriminate the differences --- it has only considered the big picture.
 Although this behavior is close to how humans see things, we also have the ability to zoom and focus on the details, clearly discerning letter ``i'' from letter ``l''.
@@
 
-% \subsection{Multiple levels}
+% \subsection{Deep learning, multiple SDMs, and the animal behavior of Dr. Linhares}
 
@@
 Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
+Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model seems to require $O(2^n)$ hard-locations, and may introduce non-uniform asymmetries in the space.}.
 

### 2018-04-03 Chapters/Chapter02.tex
theorem form

@@
 
-Each parameter set choice like this will lead to particular numbers --- many of them emergent---, such as the access radius size, critical distance, memory capacity, and so forth.
+Each parameter set choice like this will lead to particular numbers --- many of them emergent ---, such as the access radius size, critical distance, memory capacity, and so forth.
+For example, the number of dimensions $n$ leads to the number of standard deviations available on the space: $2 \sqrt{n}$.
+One then must choose a proper multiple $m\times\sigma$: Kanerva (and most in the literature) have been using $m=3$, as $\mu-3\sigma$ selects approximately $1/1000$ of the space.
+These constants will determine the size of the access radius.
+Then one must choose the number of hard locations, which will determine how many hard-locations are activated on average; and so on and so forth...
 
-One intriguing question here is:  is there a `better' number of dimensions and hard locations?  If so, can such numbers be better studied algebraically or numerically?
-
-How should these parameters be compared?  What are the tradeoffs that should be considered?  What are the `best' benchmarks possible?
+One intriguing question here is:  is there a `better' number of dimensions and hard locations?
+If so, can such numbers be better studied algebraically or numerically?
+How should these parameters be compared?
+What are the tradeoffs that should be considered?
+What are the `best' benchmarks possible?
 
@@
 Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
-Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model seems to require $O(2^n)$ hard-locations, and may introduce non-uniform asymmetries in the space.}.
+Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model may introduce non-uniform asymmetries in the space, and seems to require a staggering $O(2^n)$ hard-locations.}.
 
@@
 
-\section{Illuminating and paving a pathway}
+\section{Illuminating, paving, clearing}
 
@@
 The emphasis might have been on the \emph{breadth of topics}, in detriment of depth here or there.
+The work on, say, reinforcement learning, is most definitely not the last word we will see on the subject, but a challenge left for readers to contemplate.
 But this is due to our research group's enthusiasm for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.
 If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.
-The work on, say, reinforcement learning, is most definitely not the definitive work we will see on the subject, but a challenge left for readers to contemplate.
-Ralph Waldo Emerson once said \emph{do not go where the path may lead.
+
+Ralph Waldo Emerson once said: \emph{do not go where the path may lead.
 Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.
@@
 
-% TODO Include in the to do list simulations based on the suggestion of Murilinho.
+% TODO
 

### 2018-04-03 Chapters/sdm-critical-distance.tex
theorem form

@@
 This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
 

### 2018-04-03 kanerva-answer.tex
theorem form

@@
 \usepackage{subfig}
+\usepackage{amsthm}
+
+\newtheorem{theorem}{Theorem}[section]
+\usepackage{etoolbox}
+\AtBeginEnvironment{theorem}{\large}
 
@@
 [Thesis text....] ...As one can easily notice, we have got exactly the same curve as Kanerva.
-Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the expected $n/2$ equator distance.
+Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the $n/2$ equator distance.
+We have not, however.
 This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors.
@@
 
-   "Our results show that the theoretical prediction is not accurate.
+   ``Our results show that the theoretical prediction is not accurate.
 There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
-   (Figure \ref{fig:sdm-10000w-table-7-2})." \\
+   (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\
 
@@
 
-So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:
+So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?
+\bigskip
+
+\begin{theorem}
+%\emph{(Lagrange's Theorem)}
+\label{T1}
+$P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}$
+\end{theorem}
+
+\begin{proof}
+    Applying the law of total probability to the left-hand expression we obtain
 
     \begin{align}
-P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
+    \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) P(d(x, y) = k | d(x, y) \le r)
     \end{align}
@@
 
-Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,
+    Finally, the combinatorial identity
 
     \begin{align}
-P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}
+    \frac{n-k}{n} \binom{n}{k} = (n-k)\binom{n-1}{k} =(n-k)\binom{n-1}{n-k-1} = \binom{n-1}{k}
     \end{align}
+
+    closes the theorem.
+
+\qed
+\end{proof}
 

### 2018-04-03 kanerva-answer.tex
SDM Answer to Kanerva.

@@
     \begin{align}
-    \frac{n-k}{n} \binom{n}{k} = (n-k)\binom{n-1}{k} =(n-k)\binom{n-1}{n-k-1} = \binom{n-1}{k}
+    \frac{n-k}{n} \binom{n}{k} = \frac{(n-k)}{n} \frac{n!}{(n-k)! k!} = \frac{(n-1)!}{k! (n-1-k)!} = \binom{n-1}{k}
     \end{align}
@@
 
-\qed
 \end{proof}
@@
 In our case, $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=672.12, \sigma^2=6281.00)$, and $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=-543.49, \sigma^2=5078.99)$.
-See Figure \ref{fig:sdm-read-sums} --- we can notice there a small but significant difference between the theoretical and the simulated mean.
+See Figure \ref{fig:sdm-read-sums} --- we can notice that the average is correct but the variance is too small.
 
@@
 In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12181.95)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12181.95)$.
-See Figure \ref{fig:sdm-read} --- we can notice that the small difference in the means from Figure \ref{fig:sdm-read-sums} has propagated to these images.
+See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} has propagated to these images.
 
@@
 
-In our case, $P(wrong) = 0.12193065104931683$.
+Using the empirical variance of $\sigma^2 = 27838.3029124$, we calculate $P(wrong) = 0.22037771219874325$.
 
 In order to check this probability, I have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculate the distance from the result of a single read.
-As the $P(wrong) = 0.12193$, I expected to get an average distance of 121.93 with a standard deviation of 10.34.
-See Figure \ref{fig:sdm-read-random-bs} --- We can notice a big difference between the theoretical model and the simulation.
-Using $\mu=221$ and $\sigma^2=168$, the curves match.
-I'm still looking for the mistake in the equations.
-I believe the problem is in Equation \ref{eqn:sdm-eta1-addr1}.
+As the $P(wrong) = 0.22037$, I expected to get an average distance of 220.37 with a standard deviation of 13.10.
+See Figure \ref{fig:sdm-read-random-bs} for the comparison between the simulated and the theoretical outcomes.
 
-\begin{figure}[h!]
+Figure \ref{fig:sdm-single-read-dist} shows the new distance between $\eta_d$ and $\text{read}(\eta_d)$, where $\eta_d$ is $d$ bits away from $\eta$.
+As for $d \ge 520$ there is no intersection between $\eta$ and $\eta_d$, our models applies and explains the horizontal line around distance 220.
+
+\begin{figure}[!htb]
   \centering
-  \includegraphics[width=0.75\textwidth]{./images02/autocorrelation/read-random-bs.png}
+  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/read-random-bs.png}
 
@@
 
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/single-read-dist.png}
+
+  \caption{New distance after a single read operation in a bitstring $\eta_d$, which is $d$ bits away from $\eta$.
+The new distance was calculated between $\eta_d$ and $\text{read}(\eta_d)$.
+Notice that when $d \ge 520$, the intersection between $\eta$ and $\eta_d$ is zero, which means there is only random bitstrings written into the activated hard locations.
+The distance 220 equals $1000 \cdot 0.220$ which is the probability find in Figure \ref{fig:sdm-read-random-bs}.}
+  \label{fig:sdm-single-read-dist}
+\end{figure}
+
 \end{document}

### 2018-04-03 kanerva-answer.tex
dfrac aumenta a fonte da divisao no teorema

@@
 
-Dr. Kanerva writes:
+Dr. Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.}:
 
@@
 I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
-You say it in \\
+You say in it\\
 
    ``Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
-   (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\
 
@@
 \label{T1}
-$P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}$
+$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}$
 \end{theorem}

### 2018-04-03 kanerva-answer.tex
going very slow... theorem 2...

@@
 A hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
 
@@
 Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
+Let us call these shared bitvalues a \emph{bitmatch in dimension $i$}.
 
-So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?
+So, what is the probability of bitmatches given that we know the access radius $r$ between the address and a hard-location?
 \bigskip
@@
 \begin{theorem}
-%\emph{(Lagrange's Theorem)}
+\emph{Each dimension has a small pull bias, which can be measured by}
 \label{T1}
-$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}$
+$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}.$
 \end{theorem}
@@
 \begin{proof}
+    The left-hand expression $P(x_i = y_i | d(x, y) \le r)$ computes the probability of a bitmatch in $i$, given that we know that $x$ and $y$ are in the access radius defined by $r$, i.e., $d(x, y)\le r$.
+
     Applying the law of total probability to the left-hand expression we obtain
@@
 
-This equation is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+Theorem \ref{T1} is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory).
+When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+Each bit of a hard location does indeed have a small pull bias.
+What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values otherwise.
 
-Hence, let $Z$ be the number of activated hard location with the same bit as the reading address, then: $Z = \sum_{i=1}^{h} X_i$, where $\mathbf{E}[h] = h$, $\mathbf{V}[h] = H p_1 (1-p_1)$, $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and $X_i \sim \text{Bernoulli}(p)$.
+So far we have looked only at a single dimension.  Let us think about how the entire bitstring is affected by this result.
+
+\begin{theorem}
+\emph{Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with }
+\label{T2}
+$\mathbf{E}[Z] = ph$ and $\mathbf{V}[Z] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+
+\end{theorem}
+
+Let $Z$ be the number of activated hard locations with the same bit as the reading address:
+
+$Z = \sum_{i=1}^{h} X_i$,
+
+where $\mathbf{E}[h] = h$, \\
+$\mathbf{V}[h] = H p_1 (1-p_1)$, \\
+$p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and \\
+$X_i \sim \text{Bernoulli}(p)$. \\
+
 By the central limit theorem, $Z$ is normally distributed.
@@
 
-  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
+  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution found in \ref{T2}.}
   \label{fig:sdm-same-bit-histogram}
@@
 \begin{align}
-\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_2 = (2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2) \\
-\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu_2 = -(2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)
+\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_2 = (2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)\label{cntaddr1} \\
+\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu_2 = -(2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)\label{cntaddr0}
 \end{align}
@@
 The histogram was obtained through simulation.
-The red curve is the theoretical normal distribution.}
+The red curve is the theoretical normal distribution found in (\ref{cntaddr1}) and (\ref{cntaddr0}).}
   \label{fig:sdm-corr-counters}

### 2018-04-03 kanerva-answer.tex
teorema 1.2

@@
 \usepackage{etoolbox}
-\AtBeginEnvironment{theorem}{\large}
+%\AtBeginEnvironment{theorem}{\large}
 
@@
 
-So far we have looked only at a single dimension.  Let us think about how the entire bitstring is affected by this result.
+So far we have looked only at a single pair of bitstrings, the probability of a single bitmatch between bitstrings within the access radius distance.
+Now let us consider the number of activated hard locations exhibiting this bitmatch.
 
@@
 
-  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution found in \ref{T2}.}
+  \caption{Given an address $x$ and a dimension $i$, how many hard locations with bitmatches in $i$ are activated by reading at $x$?
+The histogram was obtained through numerical simulation.
+The red curve is the theoretical normal distribution found in Theorem \ref{T2}.}
   \label{fig:sdm-same-bit-histogram}

### 2018-04-04 kanerva-answer.tex
teorema2

@@
 
-Theorem \ref{T1} is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory).
+Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).
 When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
@@
 
-For ``random at x'', $p=0.5$ implies $\mu_2 = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
+For ``random written at x'', $p=0.5$ implies $\mu_2 = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
 
-However, for ``x at x'', $p=0.5529$ and the probabilities depend on $s$.
+However, for ``x written at x'', $p=0.5529$ and the probabilities depend on $s$.
 For $s=10,000$, they are equal to 0.6354.
@@
 
-  \caption{Autocorrelation in the counters in autoassociative memories (``x at x''). The histogram was obtained through simulation. The red curve is the theoretical distribution.}
+  \caption{Autocorrelation in the counters in autoassociative memories (``x written at x''). The histogram was obtained through simulation. The red curve is the theoretical distribution.}
   \label{fig:sdm-corr-prob}

### 2018-04-04 kanerva-answer.tex
teorema 2

@@
 Each bit of a hard location does indeed have a small pull bias.
-What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values otherwise.
+What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0.
 
@@
 
+\begin{proof}
 Let $Z$ be the number of activated hard locations with the same bit as the reading address:
-
 $Z = \sum_{i=1}^{h} X_i$,
 
-where $\mathbf{E}[h] = h$, \\
-$\mathbf{V}[h] = H p_1 (1-p_1)$, \\
-$p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and \\
-$X_i \sim \text{Bernoulli}(p)$. \\
+where $\mathbf{E}[h] = h$, $\mathbf{V}[h] = H p_1 (1-p_1)$, $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and $X_i \sim \text{Bernoulli}(p)$.
 
@@
 Applying the law of total average and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+
+\end{proof}
 

### 2018-04-04 kanerva-answer.tex
SDM Answer to Kanerva

@@
 \begin{align}
-Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1 = p \theta, \sigma_1^2 = p(1-p) \theta + p^2 s^2 p_1 (1 - p_1) / H)
+Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1, \sigma_1^2)
+\end{align}
+
+Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \theta$, since $\mathbf{E}[\theta] = \theta$.
+
+Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \theta + p^2 s^2 p_1 (1-p_1)/H$, since $\mathbf{V}[\theta] = s^2 p_1 (1-p_1)/H$.
+
+Thus,
+
+\begin{align}
+\mu_1 &= p \theta \\
+\sigma_1^2 &= p^2 s^2 p_1 (1 - p_1) / H
 \end{align}

### 2018-04-04 kanerva-answer.tex
details

--- 2018-04-04 kanerva-answer.tex
@@
 
-Dr. Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.}:
+Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.
+Some centuries ago, dear reader, all scientific publishing was the exchange of such letters...}:
 
@@
 
-Applying the law of total average and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+Applying the law of total averages and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
 
@@
 The histogram was obtained through simulation.
-The red curve is the theoretical normal distribution found in (\ref{cntaddr1}) and (\ref{cntaddr0}).}
+The red curve is the theoretical normal distribution found in equations (\ref{cntaddr1}) and (\ref{cntaddr0}).}
   \label{fig:sdm-corr-counters}

### 2018-04-04 kanerva-answer.tex
SDM Answer to Kanerva (missing line in previous commit).

@@
 \begin{align}
-Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1 = p \theta, \sigma_1^2 = p(1-p) \theta + p^2 s^2 p_1 (1 - p_1) / H)
+Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1, \sigma_1^2)
+\end{align}
+
+Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \theta$, since $\mathbf{E}[\theta] = \theta$.
+
+Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \theta + p^2 s^2 p_1 (1-p_1)/H$, since $\mathbf{V}[\theta] = s^2 p_1 (1-p_1)/H$.
+
+Thus,
+
+\begin{align}
+\mu_1 &= p \theta \\
+\sigma_1^2 &= p(1-p) \theta + p^2 s^2 p_1 (1 - p_1) / H
 \end{align}

### 2018-04-04 kanerva-answer.tex
SDM Answer to Kanerva

--- 2018-04-04 kanerva-answer.tex
@@
 
+Let $h$ be the number of activated hard locations.
+As the probability of activating a specific hard location is constant, $h \sim \text{Binomial}(H, p_1)$.
+Thus, $\mathbf{E}[h] = \mu_h = Hp_1$ and $\mathbf{V}[h] = \sigma^2_h = Hp_1(1-p_1)$, where $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$.
+
+Let $Z$ be the number of activated hard locations with the same bit as the reading address.
+Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.
+
 \begin{theorem}
-\emph{Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with }
+Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with $\mathbf{E}[Z] = \mu_Z = p \mu_h$ and $\mathbf{V}[Z] = \sigma_Z^2 = p(1-p) \mu_h + p^2 \sigma^2_h$.
 \label{T2}
-$\mathbf{E}[Z] = ph$ and $\mathbf{V}[Z] = hp(1-p) + p^2 H p_1 (1-p_1)$.
-
 \end{theorem}
@@
 \begin{proof}
-Let $Z$ be the number of activated hard locations with the same bit as the reading address:
-$Z = \sum_{i=1}^{h} X_i$,
+As $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.
 
-where $\mathbf{E}[h] = h$, $\mathbf{V}[h] = H p_1 (1-p_1)$, $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$, and $X_i \sim \text{Bernoulli}(p)$.
+Applying the law of total average, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = p \mu_h$.
 
-By the central limit theorem, $Z$ is normally distributed.
-
-Applying the law of total average and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
-
+Applying the law of total variance, $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = p(1-p)\mu_h + p^2 \sigma^2_h$.
 \end{proof}
@@
 
-Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$), $h$ be the average number of activated hard locations ($h=1,071.85$), $H$ be the number of hard locations ($H=1,000,000$), and $\text{addr}_i$ be the $i$th bit of the hard location's address.
+Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$) and $\text{addr}_i$ be the $i$th bit of the hard location's address.
 
-Let $\theta = \frac{sh}{H}$ be the average number of bitstrings written in each hard location, and $X_k \sim \text{Bernoulli}(p)$ (where $p = P(x_i = y_i | d(x, y) \le r)$). Thus,
+Let $\theta$ be the average number of bitstrings written in each hard location.
+As there are $s$ bitstrings written into the memory, and the probability of activating a specific hard location is constant, $\theta \sim \text{Binomial}(s, p_1)$.
+Thus, $\mathbf{E}[\theta] = \mu_\theta = s p_1$ and $\mathbf{V}[\theta] = \sigma^2_\theta = s p_1 (1 - p_1)$.
+
+Let $Y_i$ be the number of bitmatches in the $i$bit of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
+
+\begin{theorem}
+Giving the number of written bitstrings $s$, $\mathbf{E}[Y_i] = \mu_Y = p \mu_\theta$ and $\mathbf{V}[Y_i] = \sigma^2_Y = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
+\end{theorem}
+\begin{proof}
+Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \mu_\theta$.
+
+Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
+\end{proof}
+
+During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0.
+So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location with $Y_i$ bitmatches and $\theta - Y_i$ non-bitmatches.
+Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$ and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.
+
+\begin{theorem}
+$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mu_{\text{cnt}} = (2p-1) \mu_\theta$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt} = 4p(1-p) \mu_\theta + (2p-1)^2 \sigma^2_\theta$.
+\end{theorem}
+
+\begin{proof}
+
+$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{E}[2Y_i - \theta] = \mathbf{E}[2Y_i] - \mathbf{E}[\theta] = 2 \mathbf{E}[Y_i] - \mu_\theta = 2 p \mu_\theta - \mu_\theta = (2p-1) \mu_\theta$.
+
+Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] &= \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
+
+Let's solve each part independently. Thus,
+
+$\mathbf{V}[2Y_i - \theta | \theta] = \mathbf{V}[2Y_i | \theta] = 4 \mathbf{V}[Y_i | \theta] = 4 \mathbf{V}[\sum_{k=1}^\theta X_k] = 4 \theta p (1-p)$.
+
+$\mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] = \mathbf{E}[4 \theta p (1-p)] = 4p(1-p) \mathbf{E}[\theta] = 4p(1-p) \mu_\theta$.
+
+Finally,
+
+$\mathbf{E}[2Y_i - \theta | \theta] = 2 \mathbf{E}[Y_i | \theta] - \mathbf{E}[\theta | \theta] = 2p \theta - \theta = (2p-1) \theta$.
+
+$\mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]] = \mathbf{V}[(2p-1) \theta] = (2p-1)^2 \mathbf{V}[\theta] = (2p-1)^2 \sigma^2_\theta$.
+\end{proof}
+
+\begin{theorem}
+$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = - \mu_{\text{cnt}}$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt}$.
+\end{theorem}
+\begin{proof}
+Notice that $[\text{cnt}_i | \text{addr}_i = 0] = -[\text{cnt}_i | \text{addr}_i = 1]$.
+Thus, $\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = -\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1]$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 0] = \mathbf{V}[\text{cnt}_i | \text{addr}_i = 1]$.
+\end{proof}
+
+In summary,
 
 \begin{align}
-Y_i = \sum_{k=1}^{\theta} X_k \sim \mathcal{N}(\mu_1, \sigma_1^2)
+\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr1} \\
+\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(-\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr0}
 \end{align}
 
-Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \theta$, since $\mathbf{E}[\theta] = \theta$.
-
-Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \theta + p^2 s^2 p_1 (1-p_1)/H$, since $\mathbf{V}[\theta] = s^2 p_1 (1-p_1)/H$.
-
-Thus,
-
-\begin{align}
-\mu_1 &= p \theta \\
-\sigma_1^2 &= p^2 s^2 p_1 (1 - p_1) / H
-\end{align}
-
-During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0.
-So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location, $Y_i$ bits 1, and $\theta - Y_i$ bits 0.
-Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$; and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.
-
-Hence, as $\text{cnt}_i = 2Y_i - \theta$, $\mathbf{E}[2Y_i - \theta] = 2 \mathbf{E}[Y_i] - \theta$, and $\mathbf{V}[2Y_i - \theta] = 4 \mathbf{V}[Y_i]$, then,
-
-\begin{align}
-\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_2 = (2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)\label{cntaddr1} \\
-\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(\mu_2 = -(2p-1) \theta, \sigma_2^2 = 4 \sigma_1^2)\label{cntaddr0}
-\end{align}
-
-In our case, $p=0.5529$, $s=10,000$, $h=1,071.85$, and $H=1,000,000$, so $\theta = 10.7185$ and $\text{cnt}_i \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7294)$.
-For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.8255$.
+In our case, $p=0.5529$, $s=10,000$, and $H=1,000,000$, so $[\text{cnt}_i | \text{addr}_i=1] \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7184)$.
+For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.7185$.
 See Figure \ref{fig:sdm-corr-counters}.
@@
 
-For ``random written at x'', $p=0.5$ implies $\mu_2 = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
+For ``random written at x'', $p=0.5$ implies $\mu_\text{cnt} = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
 
@@
 The more random bitstrings are written into the memory, the more the hard locations point to themselves.
-See Figure \ref{fig:sdm-corr-prob} --- and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
+
+See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
 As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').

### 2018-04-04 kanerva-answer.tex
details

@@
 
-Dr. Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.}:
+Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.
+To think that some centuries ago, all scientific publishing was the exchange of such letters.}:
 
@@
 
-Applying the law of total average, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = p \mu_h$.
+By the central limit theorem, $Z$ is normally distributed.
+
+Applying the law of total averages and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
 

### 2018-04-04 kanerva-answer.tex
trying to reclean...
SDM Fix error from previous commit.

--- 2018-04-04 kanerva-answer.tex
@@
 
-Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] &= \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
+Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
 

### 2018-04-04 kanerva-answer.tex
details...

--- 2018-04-04 kanerva-answer.tex
@@
 
-Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] &= \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
+Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
 
-Let's solve each part independently. Thus,
+Let us solve each part independently. Thus,
 

### 2018-04-04 kanerva-answer.tex
SDM Answer to Kanerva.

--- 2018-04-04 kanerva-answer.tex
@@
 See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
-As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').
+Let $D$ be the number of counters aligned with $\text{addr}_i$.
+The standard deviation was calculated using the fact that $[D|\theta] \sim \text{Binomial}(1000, P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta))$.
 
@@
 Let $\eta$ be the reading address and $\eta_i$ the $i$th bit of it.
-Then, let's split the $h$ activated hard locations into two groups: (i) the ones with the same bit as $\eta_i$ with $ph$ hard locations, and (ii) the ones with the opposite bit as $\eta_i$ with $(1-p)h$ hard locations.
+Then, let's split the $h$ activated hard locations into two groups: (i) the ones with the same bit as $\eta_i$ with $Z$ hard locations, and (ii) the ones with the opposite bit as $\eta_i$ with $h-Z$ hard locations.
 
 \begin{align}
-\left[ \text{acc}_i|\eta_i \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=\eta_i \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_i \right]
+\left[ \text{acc}_i|\eta_i \right] &= \sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=\eta_i \right] + \sum_{k=1}^{h-Z} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_i \right]
 \end{align}
@@
 \begin{align}
-\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma_3^2 = \sigma_2^2 ph + \mu_2^2 h p(1-p)) \label{eqn:sdm-eta1-addr1} \\
-\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] &\sim \mathcal{N}(\mu_3 = -\mu_2 (1-p)h, \sigma_3^2 = \sigma_2^2 (1-p)h + \mu_2^2 h p (1-p)) \label{eqn:sdm-eta1-addr0}
+\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \mathcal{N}(\mu = \mu_\text{cnt} \mu_Z, \sigma^2 = \sigma_\text{cnt}^2 \mu_Z + \mu_\text{cnt}^2 \sigma^2_Z) \label{eqn:sdm-eta1-addr1} \\
+\sum_{k=1}^{h-Z} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] &\sim \mathcal{N}(\mu = -\mu_\text{cnt} (1-p) \mu_h, \sigma^2 = \sigma^2_\text{cnt} (1-p) \mu_h + \mu_\text{cnt}^2 \sigma^2_{h-Z}) \label{eqn:sdm-eta1-addr0}
 \end{align}
 
-In our case, $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=672.12, \sigma^2=6281.00)$, and $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=-543.49, \sigma^2=5078.99)$.
+In our case, $\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=672.12, \sigma^2=7113.87)$, and $\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=-543.49, \sigma^2=5752.54)$.
 See Figure \ref{fig:sdm-read-sums} --- we can notice that the average is correct but the variance is too small.
@@
 \begin{align}
-\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta1} \\
-\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta0}
+\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \mu_\theta \mu_h, \sigma^2 = \sigma_\text{cnt}^2 \mu_h + 2 \mu_\text{cnt}^2 \sigma^2_h) \label{eqn:sdm-eta1} \\
+\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \mu_\theta \mu_h, \sigma^2 = \sigma_\text{cnt}^2 \mu_h + 2 \mu_\text{cnt}^2 \sigma^2_h) \label{eqn:sdm-eta0}
 \end{align}

### 2018-04-04 kanerva-answer.tex

@@
 
-Let's solve each part independently. Thus,
+Let us solve each part independently. Thus,
 

### 2018-04-04 Chapters/Chapter02.tex
Merge branch 'master' of https://github.com/Alex-Linhares/msbrogli-phd

@@
 \begin{flushright}{\slshape
+    {How is memory gradually built up during one's conscious, or\\
+    even unconscious, life and thought?  My guess is that everything \\
+    we experience is classified and registered on very many parallel \\
+    channels in different locations} \\ \medskip
+    --- Stanislaw Ulam}
+
+    \bigskip
+    \bigskip
     {Pentti Kanerva's memory model was a revelation for me: \\
@@
 --- Douglas Hofstadter
-    \bigskip
-    \bigskip
 
-    {How is memory gradually built up during one's conscious, or\\
-    even unconscious, life and thought?  My guess is that everything \\
-    we experience is classified and registered on very many parallel \\
-    channels in different locations} \\ \medskip
-    --- Stanislaw Ulam}
 \end{flushright}

### 2018-04-04 Chapters/sdm-critical-distance.tex
Merge branch 'master' of https://github.com/Alex-Linhares/msbrogli-phd

--- 2018-04-03 Chapters/sdm-critical-distance.tex
@@
 
-Dr. Kanerva writes:
+Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.
+To think that some centuries ago, all scientific publishing was the exchange of such letters.}:
 
@@
 I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
-You say it in \\
+You say in it\\
 
-   "Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
-   (Figure \ref{fig:sdm-10000w-table-7-2})." \\
+   ``Our results show that the theoretical prediction is not accurate.
+There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\
 
@@
 
-This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
+This is an important observation.
+A hard location is activated because it shares many dimensions with the items read from or written onto it.
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
 
-Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
-Whilst the probability of a shared bit-value between two random bitstrings is $1/2$, an address only activates hard-locations close to it.
+Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
+Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
+Let us call these shared bitvalues a \emph{bitmatch in dimension $i$}.
 
-So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:
+So, what is the probability of bitmatches given that we know the access radius $r$ between the address and a hard-location?
+\bigskip
+
+\begin{theorem}
+\emph{Each dimension has a small pull bias, which can be measured by}
+\label{T1}
+$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}.$
+\end{theorem}
+
+\begin{proof}
+    The left-hand expression $P(x_i = y_i | d(x, y) \le r)$ computes the probability of a bitmatch in $i$, given that we know that $x$ and $y$ are in the access radius defined by $r$, i.e., $d(x, y)\le r$.
+
+    Applying the law of total probability to the left-hand expression we obtain
 
     \begin{align}
-P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
+    \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) P(d(x, y) = k | d(x, y) \le r)
     \end{align}
@@
 
-Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,
+    Finally, the combinatorial identity
 
     \begin{align}
-P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}
+    \frac{n-k}{n} \binom{n}{k} = \frac{(n-k)}{n} \frac{n!}{(n-k)! k!} = \frac{(n-1)!}{k! (n-1-k)!} = \binom{n-1}{k}
     \end{align}
 
-This equation is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+    closes the theorem.
+
+\end{proof}
+
+Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).
+When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+Each bit of a hard location does indeed have a small pull bias.
+What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0.
+
+So far we have looked only at a single pair of bitstrings, the probability of a single bitmatch between bitstrings within the access radius distance.
+Now let us consider the number of activated hard locations exhibiting this bitmatch.
+
+Let $h$ be the number of activated hard locations.
+As the probability of activating a specific hard location is constant, $h \sim \text{Binomial}(H, p_1)$.
+Thus, $\mathbf{E}[h] = \mu_h = Hp_1$ and $\mathbf{V}[h] = \sigma^2_h = Hp_1(1-p_1)$, where $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$.
+
+Let $Z$ be the number of activated hard locations with the same bit as the reading address.
+Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.
+
+\begin{theorem}
+Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with $\mathbf{E}[Z] = \mu_Z = p \mu_h$ and $\mathbf{V}[Z] = \sigma_Z^2 = p(1-p) \mu_h + p^2 \sigma^2_h$.
+\label{T2}
+\end{theorem}
+
+\begin{proof}
+As $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.
+
+By the central limit theorem, $Z$ is normally distributed.
+
+Applying the law of total averages and the law of total variance, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = ph$, and $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = hp(1-p) + p^2 H p_1 (1-p_1)$.
+
+Applying the law of total variance, $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = p(1-p)\mu_h + p^2 \sigma^2_h$.
+\end{proof}
+
+See Figure \ref{fig:sdm-same-bit-histogram} for a comparison between the theoretical model and a simulation.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.75\textwidth]{./images02/autocorrelation/same-bit-histogram.png}
+
+  \caption{Given an address $x$ and a dimension $i$, how many hard locations with bitmatches in $i$ are activated by reading at $x$?
+The histogram was obtained through numerical simulation.
+The red curve is the theoretical normal distribution found in Theorem \ref{T2}.}
+  \label{fig:sdm-same-bit-histogram}
+\end{figure}
+
+\section{Counter bias}
+
+The bias begins in the counters. Let's analyze the $i$th counter of a hard location.
+
+Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$) and $\text{addr}_i$ be the $i$th bit of the hard location's address.
+
+Let $\theta$ be the average number of bitstrings written in each hard location.
+As there are $s$ bitstrings written into the memory, and the probability of activating a specific hard location is constant, $\theta \sim \text{Binomial}(s, p_1)$.
+Thus, $\mathbf{E}[\theta] = \mu_\theta = s p_1$ and $\mathbf{V}[\theta] = \sigma^2_\theta = s p_1 (1 - p_1)$.
+
+Let $Y_i$ be the number of bitmatches in the $i$bit of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
+
+\begin{theorem}
+Giving the number of written bitstrings $s$, $\mathbf{E}[Y_i] = \mu_Y = p \mu_\theta$ and $\mathbf{V}[Y_i] = \sigma^2_Y = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
+\end{theorem}
+\begin{proof}
+Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \mu_\theta$.
+
+Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
+\end{proof}
+
+During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0.
+So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location with $Y_i$ bitmatches and $\theta - Y_i$ non-bitmatches.
+Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$ and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.
+
+\begin{theorem}
+$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mu_{\text{cnt}} = (2p-1) \mu_\theta$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt} = 4p(1-p) \mu_\theta + (2p-1)^2 \sigma^2_\theta$.
+\end{theorem}
+
+\begin{proof}
+
+$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{E}[2Y_i - \theta] = \mathbf{E}[2Y_i] - \mathbf{E}[\theta] = 2 \mathbf{E}[Y_i] - \mu_\theta = 2 p \mu_\theta - \mu_\theta = (2p-1) \mu_\theta$.
+
+Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
+
+Let us solve each part independently. Thus,
+
+$\mathbf{V}[2Y_i - \theta | \theta] = \mathbf{V}[2Y_i | \theta] = 4 \mathbf{V}[Y_i | \theta] = 4 \mathbf{V}[\sum_{k=1}^\theta X_k] = 4 \theta p (1-p)$.
+
+$\mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] = \mathbf{E}[4 \theta p (1-p)] = 4p(1-p) \mathbf{E}[\theta] = 4p(1-p) \mu_\theta$.
+
+Finally,
+
+$\mathbf{E}[2Y_i - \theta | \theta] = 2 \mathbf{E}[Y_i | \theta] - \mathbf{E}[\theta | \theta] = 2p \theta - \theta = (2p-1) \theta$.
+
+$\mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]] = \mathbf{V}[(2p-1) \theta] = (2p-1)^2 \mathbf{V}[\theta] = (2p-1)^2 \sigma^2_\theta$.
+\end{proof}
+
+\begin{theorem}
+$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = - \mu_{\text{cnt}}$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt}$.
+\end{theorem}
+\begin{proof}
+Notice that $[\text{cnt}_i | \text{addr}_i = 0] = -[\text{cnt}_i | \text{addr}_i = 1]$.
+Thus, $\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = -\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1]$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 0] = \mathbf{V}[\text{cnt}_i | \text{addr}_i = 1]$.
+\end{proof}
+
+In summary,
+
+\begin{align}
+\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr1} \\
+\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(-\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr0}
+\end{align}
+
+In our case, $p=0.5529$, $s=10,000$, and $H=1,000,000$, so $[\text{cnt}_i | \text{addr}_i=1] \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7184)$.
+For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.7185$.
+See Figure \ref{fig:sdm-corr-counters}.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
+  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
+
+  \caption{The value of the counters after $s=10,000$ writes shows the autocorrelation in the counters in autoassociative memories (``x at x'').
+The histogram was obtained through simulation.
+The red curve is the theoretical normal distribution found in equations (\ref{cntaddr1}) and (\ref{cntaddr0}).}
+  \label{fig:sdm-corr-counters}
+\end{figure}
+
+Finally,
+
+\begin{align}
+P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 1 - \mathcal{N}.\text{cdf}(0)
+\end{align}
+
+For ``random written at x'', $p=0.5$ implies $\mu_\text{cnt} = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
+
+However, for ``x written at x'', $p=0.5529$ and the probabilities depend on $s$.
+For $s=10,000$, they are equal to 0.6354.
+For $s=20,000$, they are equal to 0.6867.
+For $s=30,000$, they are equal to 0.7232.
+The more random bitstrings are written into the memory, the more the hard locations point to themselves.
+
+See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
+As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[``random at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/random_at_x_counters.png}}
+  \subfloat[``x at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_counters.png}}
+
+  \caption{Autocorrelation in the counters in autoassociative memories (``x written at x''). The histogram was obtained through simulation. The red curve is the theoretical distribution.}
+  \label{fig:sdm-corr-prob}
+\end{figure}
+
+\section{Read bias}
+
+Now that we know the distribution of $\text{cnt}_i | \text{addr}_i$, we may go to the read operation.
+During the read operation, on average, $h$ hard locations are activated and their counters are summed up.
+So, for the $i$th bit,
+
+\begin{align}
+\text{acc}_i = \sum_{k=1}^{h} \text{cnt}_k
+\end{align}
+
+Let $\eta$ be the reading address and $\eta_i$ the $i$th bit of it.
+Then, let's split the $h$ activated hard locations into two groups: (i) the ones with the same bit as $\eta_i$ with $ph$ hard locations, and (ii) the ones with the opposite bit as $\eta_i$ with $(1-p)h$ hard locations.
+
+\begin{align}
+\left[ \text{acc}_i|\eta_i \right] &= \sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=\eta_i \right] + \sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_i \right]
+\end{align}
+
+Each sum is a sum of normally distributed random variables, so
+
+\begin{align}
+\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \mathcal{N}(\mu_3 = \mu_2 ph, \sigma_3^2 = \sigma_2^2 ph + \mu_2^2 h p(1-p)) \label{eqn:sdm-eta1-addr1} \\
+\sum_{k=1}^{(1-p)h} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] &\sim \mathcal{N}(\mu_3 = -\mu_2 (1-p)h, \sigma_3^2 = \sigma_2^2 (1-p)h + \mu_2^2 h p (1-p)) \label{eqn:sdm-eta1-addr0}
+\end{align}
+
+In our case, $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=672.12, \sigma^2=6281.00)$, and $\sum_{k=1}^{ph} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=-543.49, \sigma^2=5078.99)$.
+See Figure \ref{fig:sdm-read-sums} --- we can notice that the average is correct but the variance is too small.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Equation \ref{eqn:sdm-eta1-addr1}  ($\text{addr}_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr1.png}}
+  \subfloat[Equation \ref{eqn:sdm-eta1-addr0} ($\text{addr}_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr0.png}}
+
+  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
+  \label{fig:sdm-read-sums}
+\end{figure}
+
+Hence,
+
+\begin{align}
+\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta1} \\
+\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta0}
+\end{align}
+
+In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12181.95)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12181.95)$.
+See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} has propagated to these images.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Equation \ref{eqn:sdm-eta1}  ($\eta_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1.png}}
+  \subfloat[Equation \ref{eqn:sdm-eta0} ($\eta_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta0.png}}
+
+  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
+  \label{fig:sdm-read}
+\end{figure}
+
+Finally,
+
+\begin{align}
+P(wrong) &= P(\text{acc}_i < 0 | \eta_i = 1) \cdot P(\eta_i = 1) + P(\text{acc}_i > 0 | \eta_i = 0) \cdot P(\eta_i = 0) \\
+    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{1-\mathcal{N}_{\eta_i=0}.\text{cdf}(0)}{2} \\
+    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} \\
+    &= \mathcal{N}_{\eta_i=1}.\text{cdf}(0)
+\end{align}
+
+Using the empirical variance of $\sigma^2 = 27838.3029124$, we calculate $P(wrong) = 0.22037771219874325$.
+
+In order to check this probability, I have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculate the distance from the result of a single read.
+As the $P(wrong) = 0.22037$, I expected to get an average distance of 220.37 with a standard deviation of 13.10.
+See Figure \ref{fig:sdm-read-random-bs} for the comparison between the simulated and the theoretical outcomes.
+
+Figure \ref{fig:sdm-single-read-dist} shows the new distance between $\eta_d$ and $\text{read}(\eta_d)$, where $\eta_d$ is $d$ bits away from $\eta$.
+As for $d \ge 520$ there is no intersection between $\eta$ and $\eta_d$, our models applies and explains the horizontal line around distance 220.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/read-random-bs.png}
+
+  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
+  \label{fig:sdm-read-random-bs}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/single-read-dist.png}
+
+  \caption{New distance after a single read operation in a bitstring $\eta_d$, which is $d$ bits away from $\eta$.
+The new distance was calculated between $\eta_d$ and $\text{read}(\eta_d)$.
+Notice that when $d \ge 520$, the intersection between $\eta$ and $\eta_d$ is zero, which means there is only random bitstrings written into the activated hard locations.
+The distance 220 equals $1000 \cdot 0.220$ which is the probability find in Figure \ref{fig:sdm-read-random-bs}.}
+  \label{fig:sdm-single-read-dist}
+\end{figure}
+
+%\end{document}
 

### 2018-04-05 kanerva-answer.tex
SDM Answer to Kanerva.

@@
 
+Let $D$ be the number of counters aligned with $\text{addr}_i$.
+The standard deviation was calculated using the fact that $[D|\theta] \sim \text{Binomial}(1000, q=P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta))$.
+
+Applying the law of total variance, $\mathbf{V}[D] = \mathbf{E}[\mathbf{V}[D|\theta]] + \mathbf{V}[\mathbf{E}[D|\theta]] = \mathbf{E}[1000 q (1-q)] + \mathbf{V}[1000 q] = 1000 \mathbf{E}[q-q^2] + 1000^2 \mathbf{V}[q] = 1000 \mathbf{E}[q](1-\mathbf{E}[q]) + 1000(1000-1)\mathbf{V}[q]$, where $\mathbf{E}[q] = \sum_\theta P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta) P(\theta)$ and $\mathbf{E}[q^2] = \sum_\theta [P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta)]^2 P(\theta)$.
+
+Doing the math, $\mathbf{E}[q] = 0.402922$ and $\mathbf{E}[q^2] = 0.634433$.
+Thus, $\mathbf{V}[q] = \mathbf{E}[q^2] - (\mathbf{E}[q])^2 = 0.0004166$.
+Hence, $\mathbf{V}[D] = 648.2041$.
 See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
-Let $D$ be the number of counters aligned with $\text{addr}_i$.
-The standard deviation was calculated using the fact that $[D|\theta] \sim \text{Binomial}(1000, P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta))$.
 
@@
 
-In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12181.95)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12181.95)$.
+In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12865.69)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12865.69)$.
 See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} has propagated to these images.

### 2018-04-05 Chapters/sdm-critical-distance.tex
SDM Update section A deviation from the equator distance?

@@
 
+Let $D$ be the number of counters aligned with $\text{addr}_i$.
+The standard deviation was calculated using the fact that $[D|\theta] \sim \text{Binomial}(1000, q=P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta))$.
+
+Applying the law of total variance, $\mathbf{V}[D] = \mathbf{E}[\mathbf{V}[D|\theta]] + \mathbf{V}[\mathbf{E}[D|\theta]] = \mathbf{E}[1000 q (1-q)] + \mathbf{V}[1000 q] = 1000 \mathbf{E}[q-q^2] + 1000^2 \mathbf{V}[q] = 1000 \mathbf{E}[q](1-\mathbf{E}[q]) + 1000(1000-1)\mathbf{V}[q]$, where $\mathbf{E}[q] = \sum_\theta P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta) P(\theta)$ and $\mathbf{E}[q^2] = \sum_\theta [P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta)]^2 P(\theta)$.
+
+Doing the math, $\mathbf{E}[q] = 0.402922$ and $\mathbf{E}[q^2] = 0.634433$.
+Thus, $\mathbf{V}[q] = \mathbf{E}[q^2] - (\mathbf{E}[q])^2 = 0.0004166$.
+Hence, $\mathbf{V}[D] = 648.2041$.
 See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
-As each of the $n$ counters of a hard location may be equal or not with the same probability, I assumed it would follow a Binomial distribution (and it worked for ``random at x'').
 
@@
 \begin{align}
-\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta1} \\
-\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \theta h, \sigma^2 = \sigma_2^2 h + 2 \mu_2^2 hp(1-p)) \label{eqn:sdm-eta0}
+\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \mu_\theta \mu_h, \sigma^2 = \sigma_\text{cnt}^2 \mu_h + 2 \mu_\text{cnt}^2 \sigma^2_h) \label{eqn:sdm-eta1} \\
+\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \mu_\theta \mu_h, \sigma^2 = \sigma_\text{cnt}^2 \mu_h + 2 \mu_\text{cnt}^2 \sigma^2_h) \label{eqn:sdm-eta0}
 \end{align}
 
-In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12181.95)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12181.95)$.
+In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12865.69)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12865.69)$.
 See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} has propagated to these images.
@@
 
-%\end{document}
-
 \section{Critical distance of 209}

### 2018-04-05 Chapters/sdm-autocorrelation.tex
SDM Reorganizing new section in a separate file.

--- 2018-04-05 kanerva-answer.tex
@@
-\documentclass[12pt]{article}
-
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{subfig}
-\usepackage{amsthm}
-
-\newtheorem{theorem}{Theorem}[section]
-\usepackage{etoolbox}
-%\AtBeginEnvironment{theorem}{\large}
-
-\begin{document}
-
-[Thesis text....] ...As one can easily notice, we have got exactly the same curve as Kanerva.
-Both his and our model expect that, after reading, say, from 550 bits of distance from a written bitstring, we should obtain the $n/2$ equator distance.
-We have not, however.
-This question has intrigued us, and here we look for a more analytic explanation than merely interference from the other written attractors.
-Let us turn back to mathematics to study this anomaly.
-
 \section{A deviation from the equator distance?}
@@
 \end{figure}
-
-\end{document}

### 2018-04-05 Chapters/sdm-critical-distance.tex
SDM Reorganizing new section in a separate file.

--- 2018-04-03 Chapters/sdm-critical-distance.tex
@@
 
-\section{A deviation from the equator distance?}
-
-Dr. Kanerva writes:
-
-\begin{quote}
-    You have done an incredibly thorough analysis of SDM.
-I like the puzzle in your message and believe that your simulations are correct and to be learned from.
-So what to make of the difference compared to my Figure 7.3 (and your Figure \ref{fig:kanerva-figure-73-calculated})?
-I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
-You say it in \\
-
-   "Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond ~500 bits
-   (Figure \ref{fig:sdm-10000w-table-7-2})." \\
-
-   I think that is correct.
-It also brings to mind a comment Louis Jaeckel made when we worked at NASA Ames.
-He pointed out that autoassociative storage (each vector is stored with itself as the address) introduces autocorrelation that my formula for Figure 7.2 did not take into account.
-When we read from memory, each stored vector exerts a pull toward itself, which also means that each bit of a retrieved vector is slightly biased toward the same bit of the read address, regardless of the read address.
-We never worked out the math.
-\end{quote}
-
-This is an important insight: a hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address more than it likes the opposite; as each hard-location has a say in its own area (and nowhere else).
-
-Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
-Whilst the probability of a shared bit-value between two random bitstrings is $1/2$, an address only activates hard-locations close to it.
-
-So what is the probability of shared bit-values given that we know the access radius $r$ between the address and a hard-location?  Applying the law of total probability:
-
-\begin{align}
-P(x_i = y_i | d(x, y) \le r) = \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) \cdot P(d(x, y) = k | d(x, y) \le r)
-\end{align}
-
-We also know that
-
-\begin{align}
-P(x_i = y_i | d(x, y) = k) &= \frac{n-k}{n} \\
-P(d(x, y) = k | d(x, y) \le r) &= \frac{\binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
-\end{align}
-
-Hence,
-
-\begin{align}
-P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \frac{n-k}{n} \binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
-\end{align}
-
-Finally, as $\frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}$,
-
-\begin{align}
-P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}
-\end{align}
-
-This equation is valid for both ``x at x'' (autoassociative memory) and ``random at x'' (heteroassociative memory). When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+\input{./Chapters/sdm-autocorrelation.tex}
 

### 2018-04-08 Chapters/Chapter0A.tex
fundamental research theorem

%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
\chapter{Recent Results in Theory of Computing - I}

\section{The Halting Problem is Solvable}

A fundamental question in the graduate computer science curriculum can be posed as follows: Given an average grad student doing a Ph.D, will the student ever complete his dissertation? This problem has been termed the "Halting Problem," and it has been an open problem thus far. In the following, we show that the halting problem is solvable. Furthermore, the problem can be solved within the time stipulated by the Graduate College for Ph.Ds or, in the worst case, with only a constant number of petitions for extensions.

The halting problem was first formulated by Alan Turing, who observed a number of his graduate students being apparently busy all the time but never graduating. Turing tried to solve the problem by first stopping all assistantships after the sixth year and then by purging all games from the research computers. Needless to say, his efforts were fruitless. Later, Church almost succeeded in solving the problem when he placed notices in grad students' mailboxes indicating attractive jobs in industry with several orders of magnitude higher remuneration. The so called Church's thesis was that the halting problem is solvable, given enough financial motivation. Church's idea backfired when grads found out that they have to actually work to earn money in the outside world. Thus, far from solving the halting problem, Church aggravated it (After this, we are not sure whether Church himself graduated). Recently, Cook et al have shown that the halting problem falls under a new complexity class, "NP Hairy." (NP hairy is the class of hopelessly complicated problems with no known solutions. The hardest problem in NP hairy has been shown to be the problem of trying to claim standard deductions in the 1040 form).

In the following, we show that the halting problem is indeed solvable. For this, we assume the existence of a "Super Grad," who is capable of working in any area in CS (except possibly numerical analysis). For notational convenience, we call this super grad, S sub G sup i,j sub * (written using a funky theoretical CS font). The property of Super grad is that, given the description of any grad (mostly in terms of the number of newsfiles he/she reads every day) and a description of his/her thesis topic, Super grad will either halt with a dissertation or keep publishing technical reports indefinitely. Now, we give Super grad a description of himself and his own thesis topic. If Super grad halts, we are done (and so is he) otherwise we get a stream of technical reports. But by the "fundamental research theorem" of CS Departments (refer to the graduate study manual) any five arbitrary technical reports on unrelated topics can be compiled into a Ph.D thesis. Thus, we are done in the second case too.

Finally, how long does it take for a dissertation to be completed? The time is either less than or equal to the duration allowed by the Grad College for the completion of a Ph.D or it is greater. In the latter case, infinite number of petitions can be filed for extensions. Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{Bala Rajagopalan, Internet usenet rec.humor.funny, June 12 1989, `with apologies to non-cs types'.}. (QED)

### 2018-04-08 Chapters/Introduction.tex
fundamental research theorem

@@
 
-A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic.
+A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic\footnote{e.g., the ``fundamental research theorem'' in the Appendix}.
 With these qualifications, we are ready to venture into the world of computational management science.

### 2018-04-08 kanerva-answer.tex
fundamental research theorem

--- 2018-04-04 kanerva-answer.tex
@@
 
-\section{A deviation from the equator distance?}
-
-Dr. Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.}:
-
-\begin{quote}
-    You have done an incredibly thorough analysis of SDM.
-I like the puzzle in your message and believe that your simulations are correct and to be learned from.
-So what to make of the difference compared to my Figure 7.3 (and your Figure \ref{fig:kanerva-figure-73-calculated})?
-I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
-You say in it\\
-
-   ``Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\
-
-   I think that is correct.
-It also brings to mind a comment Louis Jaeckel made when we worked at NASA Ames.
-He pointed out that autoassociative storage (each vector is stored with itself as the address) introduces autocorrelation that my formula for Figure 7.2 did not take into account.
-When we read from memory, each stored vector exerts a pull toward itself, which also means that each bit of a retrieved vector is slightly biased toward the same bit of the read address, regardless of the read address.
-We never worked out the math.
-\end{quote}
-
-This is an important observation.
-A hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
-
-Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
-Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
-Let us call these shared bitvalues a \emph{bitmatch in dimension $i$}.
-
-So, what is the probability of bitmatches given that we know the access radius $r$ between the address and a hard-location?
-\bigskip
-
-\begin{theorem}
-\emph{Each dimension has a small pull bias, which can be measured by}
-\label{T1}
-$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}.$
-\end{theorem}
-
-\begin{proof}
-    The left-hand expression $P(x_i = y_i | d(x, y) \le r)$ computes the probability of a bitmatch in $i$, given that we know that $x$ and $y$ are in the access radius defined by $r$, i.e., $d(x, y)\le r$.
-
-    Applying the law of total probability to the left-hand expression we obtain
-
-    \begin{align}
-    \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) P(d(x, y) = k | d(x, y) \le r)
-    \end{align}
-
-    We also know that
-
-    \begin{align}
-    P(x_i = y_i | d(x, y) = k) &= \frac{n-k}{n} \\
-    P(d(x, y) = k | d(x, y) \le r) &= \frac{\binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
-    \end{align}
-
-    Hence,
-
-    \begin{align}
-    P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \frac{n-k}{n} \binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
-    \end{align}
-
-    Finally, the combinatorial identity
-
-    \begin{align}
-    \frac{n-k}{n} \binom{n}{k} = \frac{(n-k)}{n} \frac{n!}{(n-k)! k!} = \frac{(n-1)!}{k! (n-1-k)!} = \binom{n-1}{k}
-    \end{align}
-
-    closes the theorem.
-
-\end{proof}
-
-Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).
-When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
-Each bit of a hard location does indeed have a small pull bias.
-What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0.
-
-So far we have looked only at a single pair of bitstrings, the probability of a single bitmatch between bitstrings within the access radius distance.
-Now let us consider the number of activated hard locations exhibiting this bitmatch.
-
-Let $h$ be the number of activated hard locations.
-As the probability of activating a specific hard location is constant, $h \sim \text{Binomial}(H, p_1)$.
-Thus, $\mathbf{E}[h] = \mu_h = Hp_1$ and $\mathbf{V}[h] = \sigma^2_h = Hp_1(1-p_1)$, where $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$.
-
-Let $Z$ be the number of activated hard locations with the same bit as the reading address.
-Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.
-
-\begin{theorem}
-Given a reading address $x$ and a dimension $i$, the number of activated hard-locations with bitmatches at $i$ follows a normal distribution with $\mathbf{E}[Z] = \mu_Z = p \mu_h$ and $\mathbf{V}[Z] = \sigma_Z^2 = p(1-p) \mu_h + p^2 \sigma^2_h$.
-\label{T2}
-\end{theorem}
-
-\begin{proof}
-As $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.
-
-Applying the law of total average, $\mathbf{E}[Z] = \mathbf{E}[\mathbf{E}[Z | h]] = \mathbf{E}[ph] = p \mathbf{E}[h] = p \mu_h$.
-
-Applying the law of total variance, $\mathbf{V}[Z] = \mathbf{E}[\mathbf{V}[Z|h]] + \mathbf{V}[\mathbf{E}[Z|h]] = \mathbf{E}[hp(1-p)] + \mathbf{V}[ph] = p(1-p) \mathbf{E}[h] + p^2 \mathbf{V}[h] = p(1-p)\mu_h + p^2 \sigma^2_h$.
-\end{proof}
-
-See Figure \ref{fig:sdm-same-bit-histogram} for a comparison between the theoretical model and a simulation.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.75\textwidth]{./images02/autocorrelation/same-bit-histogram.png}
-
-  \caption{Given an address $x$ and a dimension $i$, how many hard locations with bitmatches in $i$ are activated by reading at $x$?
-The histogram was obtained through numerical simulation.
-The red curve is the theoretical normal distribution found in Theorem \ref{T2}.}
-  \label{fig:sdm-same-bit-histogram}
-\end{figure}
-
-\section{Counter bias}
-
-The bias begins in the counters. Let's analyze the $i$th counter of a hard location.
-
-Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$) and $\text{addr}_i$ be the $i$th bit of the hard location's address.
-
-Let $\theta$ be the average number of bitstrings written in each hard location.
-As there are $s$ bitstrings written into the memory, and the probability of activating a specific hard location is constant, $\theta \sim \text{Binomial}(s, p_1)$.
-Thus, $\mathbf{E}[\theta] = \mu_\theta = s p_1$ and $\mathbf{V}[\theta] = \sigma^2_\theta = s p_1 (1 - p_1)$.
-
-Let $Y_i$ be the number of bitmatches in the $i$bit of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
-
-\begin{theorem}
-Giving the number of written bitstrings $s$, $\mathbf{E}[Y_i] = \mu_Y = p \mu_\theta$ and $\mathbf{V}[Y_i] = \sigma^2_Y = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
-\end{theorem}
-\begin{proof}
-Applying the law of total expectation, $\mathbf{E}[Y] = \mathbf{E}[\mathbf{E}[Y|\theta]] = \mathbf{E}[p \theta] = p \mathbf{E}[\theta] = p \mu_\theta$.
-
-Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
-\end{proof}
-
-During a write operation, the counters are incremented for every bit 1 and decremented for every bit 0.
-So, after $s$ writes, there will be $\theta$ bitstrings written in each hard location with $Y_i$ bitmatches and $\theta - Y_i$ non-bitmatches.
-Thus, $[\text{cnt}_i | \text{addr}_i = 1] = (Y_i) - (\theta - Y_i) = 2Y_i - \theta$ and $[\text{cnt}_i | \text{addr}_i = 0] = \theta - 2Y_i$.
-
-\begin{theorem}
-$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mu_{\text{cnt}} = (2p-1) \mu_\theta$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt} = 4p(1-p) \mu_\theta + (2p-1)^2 \sigma^2_\theta$.
-\end{theorem}
-
-\begin{proof}
-
-$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{E}[2Y_i - \theta] = \mathbf{E}[2Y_i] - \mathbf{E}[\theta] = 2 \mathbf{E}[Y_i] - \mu_\theta = 2 p \mu_\theta - \mu_\theta = (2p-1) \mu_\theta$.
-
-Applying the law of total variance, $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \mathbf{V}[2Y_i - \theta] = \mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] + \mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]]$.
-
-Let's solve each part independently. Thus,
-
-$\mathbf{V}[2Y_i - \theta | \theta] = \mathbf{V}[2Y_i | \theta] = 4 \mathbf{V}[Y_i | \theta] = 4 \mathbf{V}[\sum_{k=1}^\theta X_k] = 4 \theta p (1-p)$.
-
-$\mathbf{E}[\mathbf{V}[2Y_i - \theta | \theta]] = \mathbf{E}[4 \theta p (1-p)] = 4p(1-p) \mathbf{E}[\theta] = 4p(1-p) \mu_\theta$.
-
-Finally,
-
-$\mathbf{E}[2Y_i - \theta | \theta] = 2 \mathbf{E}[Y_i | \theta] - \mathbf{E}[\theta | \theta] = 2p \theta - \theta = (2p-1) \theta$.
-
-$\mathbf{V}[\mathbf{E}[2Y_i - \theta | \theta]] = \mathbf{V}[(2p-1) \theta] = (2p-1)^2 \mathbf{V}[\theta] = (2p-1)^2 \sigma^2_\theta$.
-\end{proof}
-
-\begin{theorem}
-$\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = - \mu_{\text{cnt}}$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 1] = \sigma^2_\text{cnt}$.
-\end{theorem}
-\begin{proof}
-Notice that $[\text{cnt}_i | \text{addr}_i = 0] = -[\text{cnt}_i | \text{addr}_i = 1]$.
-Thus, $\mathbf{E}[\text{cnt}_i | \text{addr}_i = 0] = -\mathbf{E}[\text{cnt}_i | \text{addr}_i = 1]$ and $\mathbf{V}[\text{cnt}_i | \text{addr}_i = 0] = \mathbf{V}[\text{cnt}_i | \text{addr}_i = 1]$.
-\end{proof}
-
-In summary,
-
-\begin{align}
-\left[ \text{cnt}_i | \text{addr}_i=1 \right] &\sim \mathcal{N}(\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr1} \\
-\left[ \text{cnt}_i | \text{addr}_i=0 \right] &\sim \mathcal{N}(-\mu_\text{cnt}, \sigma^2_\text{cnt})\label{cntaddr0}
-\end{align}
-
-In our case, $p=0.5529$, $s=10,000$, and $H=1,000,000$, so $[\text{cnt}_i | \text{addr}_i=1] \sim \mathcal{N}(\mu=1.1341, \sigma^2 = 10.7184)$.
-For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.7185$.
-See Figure \ref{fig:sdm-corr-counters}.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
-  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
-
-  \caption{The value of the counters after $s=10,000$ writes shows the autocorrelation in the counters in autoassociative memories (``x at x'').
-The histogram was obtained through simulation.
-The red curve is the theoretical normal distribution found in (\ref{cntaddr1}) and (\ref{cntaddr0}).}
-  \label{fig:sdm-corr-counters}
-\end{figure}
-
-Finally,
-
-\begin{align}
-P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 1 - \mathcal{N}.\text{cdf}(0)
-\end{align}
-
-For ``random written at x'', $p=0.5$ implies $\mu_\text{cnt} = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
-
-However, for ``x written at x'', $p=0.5529$ and the probabilities depend on $s$.
-For $s=10,000$, they are equal to 0.6354.
-For $s=20,000$, they are equal to 0.6867.
-For $s=30,000$, they are equal to 0.7232.
-The more random bitstrings are written into the memory, the more the hard locations point to themselves.
-
-See Figure \ref{fig:sdm-corr-prob} and notice that I still have to figure out why the mean is correct, but the standard deviation is not.
-Let $D$ be the number of counters aligned with $\text{addr}_i$.
-The standard deviation was calculated using the fact that $[D|\theta] \sim \text{Binomial}(1000, P(\text{cnt}_i > 0 | \text{addr}_i=1, \theta))$.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[``random at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/random_at_x_counters.png}}
-  \subfloat[``x at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_counters.png}}
-
-  \caption{Autocorrelation in the counters in autoassociative memories (``x written at x''). The histogram was obtained through simulation. The red curve is the theoretical distribution.}
-  \label{fig:sdm-corr-prob}
-\end{figure}
-
-\section{Read bias}
-
-Now that we know the distribution of $\text{cnt}_i | \text{addr}_i$, we may go to the read operation.
-During the read operation, on average, $h$ hard locations are activated and their counters are summed up.
-So, for the $i$th bit,
-
-\begin{align}
-\text{acc}_i = \sum_{k=1}^{h} \text{cnt}_k
-\end{align}
-
-Let $\eta$ be the reading address and $\eta_i$ the $i$th bit of it.
-Then, let's split the $h$ activated hard locations into two groups: (i) the ones with the same bit as $\eta_i$ with $Z$ hard locations, and (ii) the ones with the opposite bit as $\eta_i$ with $h-Z$ hard locations.
-
-\begin{align}
-\left[ \text{acc}_i|\eta_i \right] &= \sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=\eta_i \right] + \sum_{k=1}^{h-Z} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_i \right]
-\end{align}
-
-Each sum is a sum of normally distributed random variables, so
-
-\begin{align}
-\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \mathcal{N}(\mu = \mu_\text{cnt} \mu_Z, \sigma^2 = \sigma_\text{cnt}^2 \mu_Z + \mu_\text{cnt}^2 \sigma^2_Z) \label{eqn:sdm-eta1-addr1} \\
-\sum_{k=1}^{h-Z} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] &\sim \mathcal{N}(\mu = -\mu_\text{cnt} (1-p) \mu_h, \sigma^2 = \sigma^2_\text{cnt} (1-p) \mu_h + \mu_\text{cnt}^2 \sigma^2_{h-Z}) \label{eqn:sdm-eta1-addr0}
-\end{align}
-
-In our case, $\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=672.12, \sigma^2=7113.87)$, and $\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=1 \right] \sim \mathcal{N}(\mu=-543.49, \sigma^2=5752.54)$.
-See Figure \ref{fig:sdm-read-sums} --- we can notice that the average is correct but the variance is too small.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Equation \ref{eqn:sdm-eta1-addr1}  ($\text{addr}_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr1.png}}
-  \subfloat[Equation \ref{eqn:sdm-eta1-addr0} ($\text{addr}_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr0.png}}
-
-  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
-  \label{fig:sdm-read-sums}
-\end{figure}
-
-Hence,
-
-\begin{align}
-\left[ \text{acc}_i|\eta_i=1 \right] &\sim \mathcal{N}(\mu = (2p-1)^2 \mu_\theta \mu_h, \sigma^2 = \sigma_\text{cnt}^2 \mu_h + 2 \mu_\text{cnt}^2 \sigma^2_h) \label{eqn:sdm-eta1} \\
-\left[ \text{acc}_i|\eta_i=0 \right] &\sim \mathcal{N}(\mu = -(2p-1)^2 \mu_\theta \mu_h, \sigma^2 = \sigma_\text{cnt}^2 \mu_h + 2 \mu_\text{cnt}^2 \sigma^2_h) \label{eqn:sdm-eta0}
-\end{align}
-
-In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12181.95)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12181.95)$.
-See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} has propagated to these images.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Equation \ref{eqn:sdm-eta1}  ($\eta_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1.png}}
-  \subfloat[Equation \ref{eqn:sdm-eta0} ($\eta_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta0.png}}
-
-  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
-  \label{fig:sdm-read}
-\end{figure}
-
-Finally,
-
-\begin{align}
-P(wrong) &= P(\text{acc}_i < 0 | \eta_i = 1) \cdot P(\eta_i = 1) + P(\text{acc}_i > 0 | \eta_i = 0) \cdot P(\eta_i = 0) \\
-    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{1-\mathcal{N}_{\eta_i=0}.\text{cdf}(0)}{2} \\
-    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} \\
-    &= \mathcal{N}_{\eta_i=1}.\text{cdf}(0)
-\end{align}
-
-Using the empirical variance of $\sigma^2 = 27838.3029124$, we calculate $P(wrong) = 0.22037771219874325$.
-
-In order to check this probability, I have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculate the distance from the result of a single read.
-As the $P(wrong) = 0.22037$, I expected to get an average distance of 220.37 with a standard deviation of 13.10.
-See Figure \ref{fig:sdm-read-random-bs} for the comparison between the simulated and the theoretical outcomes.
-
-Figure \ref{fig:sdm-single-read-dist} shows the new distance between $\eta_d$ and $\text{read}(\eta_d)$, where $\eta_d$ is $d$ bits away from $\eta$.
-As for $d \ge 520$ there is no intersection between $\eta$ and $\eta_d$, our models applies and explains the horizontal line around distance 220.
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/read-random-bs.png}
-
-  \caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
-  \label{fig:sdm-read-random-bs}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/single-read-dist.png}
-
-  \caption{New distance after a single read operation in a bitstring $\eta_d$, which is $d$ bits away from $\eta$.
-The new distance was calculated between $\eta_d$ and $\text{read}(\eta_d)$.
-Notice that when $d \ge 520$, the intersection between $\eta$ and $\eta_d$ is zero, which means there is only random bitstrings written into the activated hard locations.
-The distance 220 equals $1000 \cdot 0.220$ which is the probability find in Figure \ref{fig:sdm-read-random-bs}.}
-  \label{fig:sdm-single-read-dist}
-\end{figure}
+\input{./Chapters/sdm-autocorrelation.tex}
 

### 2018-04-08 Chapters/Chapter0A.tex
det

@@
 In the latter case, infinite number of petitions can be filed for extensions.
-Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{Bala Rajagopalan, Internet usenet rec.humor.funny, June 12 1989, `with apologies to non-cs types'.}.
+Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-cs types': from Bala Rajagopalan, posted on Internet Usenet rec.humor.funny in June 12 1989.}.
 (QED)

### 2018-04-08 Chapters/Introduction.tex
det

--- 2018-03-22 Chapters/Introduction.tex
@@
 
-A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic.
+A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic\footnote{e.g., a reference to the ``fundamental research theorem'', found in the Appendix, seems inescapable.}.
 With these qualifications, we are ready to venture into the world of computational management science.

### 2018-04-08 Chapters/Introduction.tex
reprinted

--- 2018-03-22 Chapters/Introduction.tex
@@
 
-A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic.
+A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic\footnote{e.g., a reference to the ``fundamental research theorem'', reprinted in the Appendix, seems inescapable.}.
 With these qualifications, we are ready to venture into the world of computational management science.

### 2018-04-08 Chapters/Chapter02.tex
comparison w/ other implementations

@@
 
+Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry:
+
+\begin{enumerate}
+    \item APL implementation [59]
+
+    \item LISP implementation for the Connection Machine [60]
+
+    \item FPGA implementation \citep{silva_reconfigurable_2004}
+
+    \item The original hardware implementation developed by NASA [4]
+
+    \item C Binary Vector Symbols (CBVS): includes SDM implementation in C as a part of vector symbolic architecture[54] developed by EISLAB at Luleå University of Technology: http://pendicular.net/cbvs.php[55]
+
+    \item CommonSense ToolKit (CSTK) for realtime sensor data processing developed at the Lancaster University includes implementation of SDM in C++: http://cstk.sourceforge.net/[56]
+
+    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java: http://ccrg.cs.memphis.edu/framework.html
+
+\end{enumerate}
+
+Let us analyse these.  The Connection Machine is obsolete.  APL, while influential, is not a mainstream language in science. The NASA implementation is hardware-based, hence not open-source.
+
+The FPGA implementation by \citet{silva_reconfigurable_2004} has yielded a fast scan of hard locations, provided one has access to the proper hardware.
+Their article claims a four-fold speedup over assembly language; but it does not deal with parallel processing details.
+For example, it is unclear whether there was more than a single thread running on the software implementation.
+Note that the framework presented here is also able to reconfigure field-programmable gate arrays, through the OpenCL heterogeneous computing platform ability to interface with Hardware Description Language and hence, reconfigure FPGAs \citep{waidyasooriya2018design}.
+
 It is our belief that a tool such as a framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks.
@@
 Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
-Furthermore, new members of our research group had to go through different source codes developed by former members.
+Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.
+
+The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extensible to test new research ideas.
 
@@
 
-The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extended to test new research ideas.
 Our reference implementation may, hopefully, accelerate research into the model's dynamics and make it easier for readers to replicate any previous results and easily understand the source-code of the model.

### 2018-04-08 Chapters/Chapter02.tex
comparison

@@
 \begin{enumerate}
-    \item APL implementation [59]
+    \item The original hardware implementation developed by NASA [4]
 
@@
 
+    \item APL implementation [59]
+
     \item FPGA implementation \citep{silva_reconfigurable_2004}
-
-    \item The original hardware implementation developed by NASA [4]
 
@@
 
-Let us analyse these.  The Connection Machine is obsolete.  APL, while influential, is not a mainstream language in science. The NASA implementation is hardware-based, hence not open-source.
+Let us analyse these.  The Connection Machine is obsolete.  The NASA implementation is hardware-based, hence not open-source.  APL, while influential, is not a mainstream language in science.
 
@@
 For example, it is unclear whether there was more than a single thread running on the software implementation.
-Note that the framework presented here is also able to reconfigure field-programmable gate arrays, through the OpenCL heterogeneous computing platform ability to interface with Hardware Description Language and hence, reconfigure FPGAs \citep{waidyasooriya2018design}.
+Note that the framework presented here is also able to reconfigure field-programmable gate arrays, through the OpenCL heterogeneous computing platform ability to interface with Hardware Description Language and hence, reconfigure FPGAs \citep{waidyasooriya2018design, czajkowski_opencl_2012} for our tasks.
 

### 2018-04-08 Chapters/Chapter0A.tex
comparison

--- 2018-04-08 Chapters/Chapter0A.tex
@@
 In the latter case, infinite number of petitions can be filed for extensions.
-Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{Bala Rajagopalan, Internet usenet rec.humor.funny, June 12 1989, `with apologies to non-cs types'.}.
+Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-cs types': this wonderful result is from Bala Rajagopalan, posted on Internet Usenet rec.humor.funny in June 12 1989.}.
 (QED)

### 2018-04-09 Chapters/Chapter02.tex
comparison

@@
 
-Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry:
+\section{The wasted effort of duplicated work}
+
+Our motivation is our own effort to run our models.
+As there is no reference implementation, we had to develop our own and run several simulations to ensure that our implementation was correct and bug-free.
+Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
+Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.
+Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}:
 
 \begin{enumerate}
-    \item The original hardware implementation developed by NASA [4]
+    \item The original hardware implementation developed by NASA [4];
 
-    \item LISP implementation for the Connection Machine [60]
+    \item LISP implementation for the Connection Machine [60];
 
-    \item APL implementation [59]
+    \item APL implementation [59];
 
-    \item FPGA implementation \citep{silva_reconfigurable_2004}
+    \item FPGA implementation \citep{silva_reconfigurable_2004};
 
-    \item C Binary Vector Symbols (CBVS): includes SDM implementation in C as a part of vector symbolic architecture[54] developed by EISLAB at Luleå University of Technology: http://pendicular.net/cbvs.php[55]
+    \item C Binary Vector Symbols (CBVS): includes SDM implementation in C as a part of vector symbolic architecture[54] developed by EISLAB at Luleå University of Technology: http://pendicular.net/cbvs.php[55];
 
-    \item CommonSense ToolKit (CSTK) for realtime sensor data processing developed at the Lancaster University includes implementation of SDM in C++: http://cstk.sourceforge.net/[56]
+    \item CommonSense ToolKit (CSTK) for realtime sensor data processing developed at the Lancaster University includes implementation of SDM in C++: http://cstk.sourceforge.net/[56];
 
-    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java: http://ccrg.cs.memphis.edu/framework.html
+    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java: http://ccrg.cs.memphis.edu/framework.html;
 
@@
 
-It is our belief that a tool such as a framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks.
+Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas.
+It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to Python\footnote{Python is sometimes called a `glue language'.
+That is, in my opinion, not the best metaphor.
+A glue connects two things leaving an inflexible structure.
+Python is perhaps best described as the interstate highway system of Programming; if something is out there, there is a way to reach it with Python.
+In the comparison with Java, for instance, take the $popcnt(xor(b_i,b_j))$ operation, executed billions of times in SDM.
+How easy is it to program that for a particular GPU or FPGA with Java?}.
+It is not available on an open-source repository.
+It has a non-standard license, strange to the open-source community, ``the LIDA Framework Software NonExclusive, Non-Commercial Use License''.
+Potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''...
+\emph{`before Memphis can accept it'}.
+
+The closest implementations to ours, in philosophy at least, is the one in `the common sense toolkit'.
+It is executed in C++, with a normal open-source license, and hosted on an open-source code repository.
+It is, however, strikingly dissimilar to ours on the following aspects:
+
+\begin{enumerate}
+    \item SDM is but a part of the system; the description of the system reads that cstk is `A toolkit for processing and visualising sensor data in real time with support for use with embedded platforms.'
+
+    \item The whole SDM code is composed of 143 \emph{lines} of C++ in the $cstk/cstk-devonly/sdm$ folder.
+0
+    \item There is no work on making the system parallel.
+
+    \item There is strong coupling between location address and location data, which makes experimentation hard.
+
+    \item There are no tests or examples to be found instantly.
+
+    \item Finally, the last commit to this repository seems to have been made in 2005?
+
+    \item There are no publishable / published scientific applications or experiments available to be reproduced.
+
+    \item there is no tutorial, installation instructions, performance benchmarks, framework validation or SDM Documentation.
+
+\end{enumerate}
+
+Note that all these criticisms apply to the implementations in both the `common sense toolkit' and the `C Binary Vector Symbols' REF.
+While these implementations have around 150 lines of C++; at last count, the \emph{documentation of our implementation} had over 100 \emph{pages}.
+We aim a higher here, and, to us, SDM is at center stage.
+
+There is obviously a demand for use of SDM; but each group has been tied to their own \emph{ad-hoc} needs, and there has not been the emergence of a community centered on a tool.
+It is our belief that a tool such as standard open-source framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
-This new tool would provide a ready to use system in which experiments could be executed almost as soon as designed --- and it might accelerate research \citep{shen2014interactive}.
+This new tool would provide a ready to use system in which experiments could be executed almost as soon as designed --- and provide the needed replication of studies \citep{shen2014interactive}.
 
-Our motivation was our own effort to run our models.
-As there is no reference implementation, we had to develop our own and run several simulations to ensure that our implementation was correct and bug-free.
-Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
-Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.
+The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends\footnote{CPUs, GPUs, FPGAs} and operations, (iii) is fully validated against the mathematical model, (iv) is cross-platform\footnote{Unix, Linux, MacOs, Windows, Amazon Web Services, etc.}, and (v) is easily extensible to test new research ideas --- and to let others replicate the results.
 
-The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends and operations, (iii) has been validated against the mathematical model, (iv) is cross-platform, and (v) is easily extensible to test new research ideas.
-
-Extensions of SDM have been used in many applications.
+Another issue is \emph{extensibility}: Extensions of SDM have been used in many applications.
 For example, \citet{Snaider2011} extended SDM to store sequences of vectors and trees efficiently.

### 2018-04-09 Chapters/Chapter01.tex
details here and there

@@
 \begin{flushright}{\slshape
-    {We are watching History being made --- or History being repeated.}
+    {Once I understood the logic behind change addresses...\\
+    $ripemd160( SHA256( pubkey ) )$... the inneficiency... \\
+    the wasted space; the wasted energy... something clicked.\\
+    \medskip
+
+    This is military-grade cryptography and full-blown paranoia.\\
+    Whoever did it hears voices whispering `they know everything.' \\
+    \medskip
+
+    Other than that, it's a get-rich-quick-scheme \\
+    that looks exactly like a get-rich-quick-scheme.}
+	\\ \medskip
+    --- Alex Linhares, EBAPE/FGV
+
+        \bigskip
+        \bigskip
+        \bigskip
+        \bigskip
+
+    {We are watching History being made.\\
+    Or History being repeated.}
 	\\ \medskip
@@
 \bigskip
-
-The main problem when one is trying to create digital money is how to prevent double spending.
+\bigskip
+\bigskip
+
+The main problem with digital money is how to prevent double spending.
 As the money is digital, and copies can be made \textit{ad nauseam}, what can prevent counterfeiting?
@@
 That is precisely the problem solved by Bitcoin and its underlying blockchain technology.
-The current solution behind fiat money is having a single issuer, a central bank, and trusting the financial institutions.
+The current solution behind fiat money is having a single issuer, a central bank --- then trusting the financial institutions.
 

### 2018-04-09 Chapters/Chapter02.tex
details here and there

@@
 
-    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java: http://ccrg.cs.memphis.edu/framework.html;
+    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java.
+\footnote{http://ccrg.cs.memphis.edu/framework.html};
 
@@
 
-Let us analyse these.  The Connection Machine is obsolete.  The NASA implementation is hardware-based, hence not open-source.  APL, while influential, is not a mainstream language in science.
+Let us analyse these.  The Connection Machine is obsolete.  The NASA implementation is hardware-based and obsolete.  APL, while influential, is not a mainstream language in science.
 
@@
 
-Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas.
+Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas \citep{Anwar2003, snaider_integer_2013, snaider_modular_2014, franklin_lida:_2014}.
 It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to Python\footnote{Python is sometimes called a `glue language'.
@@
 It is not available on an open-source repository.
-It has a non-standard license, strange to the open-source community, ``the LIDA Framework Software NonExclusive, Non-Commercial Use License''.
-Potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''...
-\emph{`before Memphis can accept it'}.
+It has a non-standard license, strange to the open-source community, \emph{the LIDA Framework Software NonExclusive, Non-Commercial Use License}.
+Moreover, potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''...
+\emph{`before Memphis can accept it'} \footnote{What they are attempting to do with this bureaucracy remains unknown, the history of computing has not been kind to those who favored centralization \citep{ferguson_computer_2002}.}.
 
@@
     \item The whole SDM code is composed of 143 \emph{lines} of C++ in the $cstk/cstk-devonly/sdm$ folder.
-0
+
     \item There is no work on making the system parallel.
@@
 
-    \item There are no publishable / published scientific applications or experiments available to be reproduced.
+    \item There are no publishable or published scientific applications or experiments available to be reproduced at installation time.
 
@@
 \citet{fan1997genetic} extended SDM with genetic algorithms.
-\citet{chada2016you} extended SDM creating the Rotational Sparse Distributed Memory (RSDM), which is used to modeling network motifs, dynamic flexibility, and hierarchical organization, all results from neuroscience literature.
+\citet{chada2016you} extended SDM creating the Rotational Sparse Distributed Memory (RSDM), which models network motifs, dynamic flexibility, and hierarchical organization --- reflecting results from the neuroscience literature.
 
@@
 
-This time, however, it will be running on a computer.
+This time, however, it will be running on a computer:  implicit assumptions will emerge; anomalies between expected theory and numerical results will appear; and reproducible experiments will be conducted.
 

### 2018-04-09 Chapters/Chapter0A.tex
details here and there

--- 2018-04-08 Chapters/Chapter0A.tex
@@
 In the latter case, infinite number of petitions can be filed for extensions.
-Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{Bala Rajagopalan, Internet usenet rec.humor.funny, June 12 1989, `with apologies to non-cs types'.}.
-(QED)
+Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-CS types': this wonderful result is from Bala Rajagopalan, posted on the late rec.humor.funny Usenet in June 12 1989.}.
+
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+
+\hfill \emph{Quod Erat Demonstrandum}

### 2018-04-09 Chapters/sdm-autocorrelation.tex
details here and there

@@
+% !TEX root = ../partial-sdm.tex
+
 \section{A deviation from the equator distance?}
@@
 
-The bias begins in the counters. Let's analyze the $i$th counter of a hard location.
+The previous theorems show that there is bias in the counters. Let us analyze the $i$th counter of a hard location.
 
@@
 
-Let $Y_i$ be the number of bitmatches in the $i$bit of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
+Let $Y_i$ be the number of bitmatches in the $i$ of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
 
 \begin{theorem}
-Giving the number of written bitstrings $s$, $\mathbf{E}[Y_i] = \mu_Y = p \mu_\theta$ and $\mathbf{V}[Y_i] = \sigma^2_Y = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
+Given the number of written bitstrings $s$, $\mathbf{E}[Y_i] = \mu_Y = p \mu_\theta$ and $\mathbf{V}[Y_i] = \sigma^2_Y = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
 \end{theorem}

### 2018-04-09 Chapters/Chapter02.tex
50 refs on Kanerva chapter

@@
 
-The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends\footnote{CPUs, GPUs, FPGAs} and operations, (iii) is fully validated against the mathematical model, (iv) is cross-platform\footnote{Unix, Linux, MacOs, Windows, Amazon Web Services, etc.}, and (v) is easily extensible to test new research ideas --- and to let others replicate the results.
+The main contribution of this work is a reference implementation which yields (i) orders of magnitude gains in performance, (ii) has several backends\footnote{CPUs, GPUs, FPGAs} and operations, (iii) is fully validated against the mathematical model, (iv) is cross-platform\footnote{Unix, Linux, MacOs, Windows, Amazon Web Services, etc.}, and (v) is easily extensible to test new research ideas --- and to let others replicate the studies.
 
@@
 Note that 29 additional bits imply an attractor area $2^{29}$ times larger than the original.
-This Shannon write may affect memory capacity --- possibly increasing it.
+This Shannon write may affect memory capacity \citep{chou_capacity_1989, chou_capacity_1988, sjodin_improving_1995} --- possibly increasing it.
 Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, the shape of the attractor may have asymmetries.

### 2018-04-09 Chapters/sdm-applications.tex
50 refs on Kanerva chapter

@@
 In Figure \ref{fig-filter-progressive-noise}, we may see an image with different levels of noise, from 0\% to 45\% in steps of 5\%.
-It makes no sense to apply 50\% of noise because it would absolutely randomize the image.
+It makes no sense to apply 50\% of noise as it would absolutely randomize the image.
 
@@
 Half of the letter ``I'' and half of the letter ``T'' (see Figure \ref{fig:filter-training}).
-They were written into their own addresses, i.e., \lstinline{write(address=bs_image, datum=bs_image)}.
+They were written into their own addresses, i.e., \lstinline{write(address=bs_image, datum=bs_image)}, following, for instance, \citet{hong_character_1991, marshall_technical_1989}.
 

### 2018-04-09 Chapters/Chapter02.tex
sociology of science comment

@@
 
-This time, however, it will be running on a computer:  implicit assumptions will emerge; anomalies between expected theory and numerical results will appear; and reproducible experiments will be conducted.
+This time, however, it will be running on a computer:   anomalies between expected theory and numerical results will appear; implicit, hidden, assumptions will emerge; and reproducible experiments will be conducted.
 
@@
 But we are not only interested in the computer science effects here --- the ambition is different.
-More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.
+More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  communities form around open-source code, specialists in a particular arena magnify the capabilities of the code, debates and meetings are regularly held, and deep divisive disagreements lead to community splits in multiple potential directions of exploration.
+In a very real sense, given a shared code to work and experiment on, we can see farther.
 
-Beyond speed, I have also strived for \emph{ease of use}.
+Beyond speed, I have also strived for \emph{ease of use} and \emph{replicable studies}.
 All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested.
-I have generated a Docker image, which makes it even easier to explore the framework.
+Most importantly, they also serve as a tutorial for learning about the framework's design, and perhaps as a teaching tool for SDM itself.
+I have generated a Docker image, which makes it even easier to explore the framework by delegating heavy computational work to cloud services.
 After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.

### 2018-04-09 Chapters/Chapter0A.tex
closing for the day...

@@
 In the latter case, infinite number of petitions can be filed for extensions.
-Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-CS types': this wonderful result is from Bala Rajagopalan, posted on the late rec.humor.funny Usenet in June 12 1989.}.
+Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-CS types': this seminal, wonderful, result is from Bala Rajagopalan, published on the late rec.humor.funny Usenet in June 12, 1989.}.
 

### 2018-04-09 Chapters/Conclusion.tex
closing for the day...

@@
 
-With this, I submit this thesis in the hope that all readers, present and future, may find the aforementioned studies as useful, genuine, and legitimate contributions to the thriving field of Computational Management Science.
+With this, I submit this thesis in the hope that all readers, present and future, may find the aforementioned studies as useful, genuine, and legitimate contributions to the thriving field of Computational Management Science\footnote{Finally, for the skeptical reader that may argue against the decision towards `arbitrary technical reports on unrelated topics', a reference to the ``fundamental research theorem'', reprinted in Appendix \ref{Recent Results in Theory of Computing - I}, seems, of course, virtually inescapable.}.

### 2018-04-09 Chapters/Introduction.tex
closing for the day...

--- 2018-04-08 Chapters/Introduction.tex
@@
 
-A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic\footnote{e.g., the ``fundamental research theorem'' in the Appendix}.
+A second and final qualification: It is my hope that readers of this thesis will accept the format of self-contained studies, as just as valid as a monograph on a particular topic.
 With these qualifications, we are ready to venture into the world of computational management science.

### 2018-04-09 Chapters/Chapter01.tex
Blockchain: add Hathor's introduction

@@
 
-The present work intends to analyze a new architecture which lies between Bitcoin and Iota and may be a viable solution to both Bitcoin's scaling, centralization, and spam problems.
+The present work intends to analyze a new architecture which lies between Bitcoin's and Iota's and may be a viable solution to both Bitcoin's scaling, centralization, and spam problems.
 We also present a mathematical analysis of Bitcoin's mining, forking, and safety.
@@
 
+\section{Hathor's Introduction}
+
+Hathor's architecture lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
+Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG).
+For this, each transaction has its own proof-of-work which is solved by the issuer before propagating the transactions in the network.
+Like Bitcoin, miners find new ``blocks'' every 10 minutes in which they collect the fees and newly generated tokens.
+Each transaction has an ``accumulated weight'' which express the required effort to break the transaction, similar to Bitcoin's number of confirmations.
+
+This way, there will be two difficulty levels in the system: one for new transactions which are just moving tokens around, and another one for ``blocks'' which are generating new tokens and collecting fees.
+The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions.
+The second is adjusted every 2016 blocks to keep the pace of one block every 10 minutes.
+
+There is a trade-off about the required proof-of-work of new transactions.
+The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.
+
+Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
+It is always possible given that there is at least one confirmation path between any transaction and a tip.
+
+The transactions are classified into three groups: (i) confirmed transactions, (ii) confirming transactions, and (iii) unconfirmed transactions (tips).
+The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
+The unconfirmed transactions (tips) are the very new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
+The confirming transactions are in the middle.
+They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.
+
+A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two confirming transactions or tips.
+There may be any number of outputs provided that they comply with the number of newly generated tokens.
+Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.
+
+In the low load scenario, there is a small number of new transactions coming into the network, which means they give a minor contribution to confirmations.
+In this case, the confirmation is held mostly by ``blocks''.
+On the other hand, in the high load scenarios, there is a large number of new transactions giving a major contribution to confirmations.
+In this case, the ``blocks'' strengthen the confirmation, but most of them will have already been confirmed before the new blocks are found.
+The higher the number of new transaction, the faster the transactions are confirmed.
+The blocks assure a ``maximum confirmation time''.
+
+The incentive schema which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
+
+Let $A$ be a transaction, $w_A$ be its weight, and $\text{acc}_A$ be its accumulated weight. Then, $\text{acc}_A = w_A + \sum_{A \leadsto P} w_P$.
+
 \chapter{Conclusion}

### 2018-04-10 Chapters/Chapter01.tex
Blockchain: Iota and Hathor's description.

@@
 
-After the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created (REF).
+After the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created \citep{coinmarketcap}.
 In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain.
@@
 It has neither mining nor confirmation blocks and transaction fees.
-Each transaction has its own proof-of-work and is used to confirm other transactions, forming a directed acyclic graph.
+Each transaction has its own proof-of-work and is used to confirm other transactions, forming a directed acyclic graph of transactions.
+Thus, a transaction is said to be confirmed when there are enough proof-of-work from the transactions confirming it directly or indirectly.
+There is no other way to confirm transactions but generating new transactions.
+
 In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions.
 Hence, theoretically, it scales to any large number of transactions per second.
-The scaling problem of tangle is exactly the opposite of Bitcoin's.
-It must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work.
-
-The present work intends to analyze a new architecture which lies between Bitcoin's and Iota's and may be a viable solution to both Bitcoin's scaling, centralization, and spam problems.
-We also present a mathematical analysis of Bitcoin's mining, forking, and safety.
+The scaling problem of tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work.
+While Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator which works as a trustworthy node \citep{iotacoordinator}.
+This central coordinator just elucidates that the tangle does not seem to work properly under a low volume of transactions.
+
+Iota's tokens are pre-mined, which means all of them have been issued in the genesis transaction, and there is no token issuance at all.
+There is no miners and users can only transfer tokens between themselves.
+This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it.
+As users perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away \citep{debo2010prices}.
+
+The present work intends to analyze a new architecture, named Hathor, which lies between Bitcoin and Iota and may be a viable solution to both scaling, centralization, and spam problems.
+We also present a mathematical analysis of Bitcoin's architecture.
 
@@
 
-\chapter{DAG model}
-
-DAG model proposes a whole different approach to confirmations.
+\chapter{Iota \& Tangle}
+
+Iota's underlying technology is Tangle, which has DAG-based architecture with a whole different approach to confirmations.
 It proposes that there is no need for a block to confirm transactions, as transactions can confirm themselves.
-Here, each transaction has its own proof-of-work, named weight, and they must confirm two other, previous, transactions.
-Hence, each transaction has an accumulated weight, which is the accumulated proof-of-work that has confirmed it so far.
-In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph, as in Fig.
+Here, each transaction has its own proof-of-work, named weight, and they must confirm two other previous transactions.
+In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph (DAG), as in Fig.
 \ref{fig-tangle-example}.
 
-Like the Blockchain, the DAG model is another technology to store immutable data and may be the underlying technology to different applications, such as cryptocurrencies, digital contracts (Ethereum-like), digital notaries, and so forth.
-The main question which this work proposes to answer is: Is it reliable to use DAG model?
-In which conditions?
-
-The transactions may be either confirmed or unconfirmed.
+Like Blockchain, Tangle is another technology to store immutable data and may be the underlying technology to different applications, such as cryptocurrencies, digital contracts (Ethereum-like), digital notaries, and so forth.
+
+Transactions may be either confirmed or unconfirmed.
 The confirmed transactions have been already confirmed by at least one more transaction.
@@
 
-Transactions may have any format, including Bitcoin-like transactions---with inputs, outputs, and scripts.
-But it also may be completely different, just like in the cryptocurrency Iota \citep{tangle2016}.
-This work will not discuss details of the transactions' format, because we understand that it does not affect exactly the network scalability and security.
+Different from Bitcoin, transactions does not have scripts to check whether one may spend the tokens.
+Instead, it uses a fixed Winternitz hash-based digital signature \citep{dods2005hash}, i.e., whoever correctly signs the transaction may tranfer the tokens.
+It also supports multi-signature scheme (see \cite{iotamultisign}).
+Not allowing scripts is a disadvantage when compared to Bitcoin because future applications are limited and changing it would require a major modification in transaction format.
+
+At the beginning, the digital signature algorithm relied on the Curl hash function, which is a ternary hash function designed by Iota's developers.
+This hash function was replaced by the Kerl hash function after \cite{heilman2017iota} has found a critical vulnerability which enabled practical signature forgery attacks.
+The Kerl hash function is a variation of SHA-3 also designed by Iota's developers \citep{iotakerl}.
+Like the Curl hash function, Kerl has not been deeply studied by cryptography researchers and may have critical vulnerabilities.
+
+Iota's tokens are pre-mined, which means they have been issued in the genesis transaction and no more tokens will ever be issued.
+This means there are no miners in Iota's network and only the users keep the network alive through their new transactions.
+If, for any reason, the number of transactions per second plummet, the time to confirmed the most recent transactions will increase significantly.
+This possibly poses a serious risk to the network stability and users' trustworthy.
+
+As there are no miners, there are also no fees, which is a major incentive to new comers.
+Tokens may be freely transfered without any loses, even for tiny amounts, enabling micropayments.
+For example, online workers paid per hour may receive their payments every hour, reducing the default risk.
+This allows untrusted parties to work together reducing the risk for both.
+The contractor always wants to pay in the end, while the workers always wants to get paid in the beginning.
+
+Not having fees is also also an important feature to the Internet-Of-Things (IoT) technologies.
+The IoT is a network of smart devices that are connected to the internet and exchange data.
+It allows, for instance, your refrigerator to do your groceries and automatically pay using your Iota tokens.
+It also allows your electric car to automatically pay for recharge.
+For further information, see \cite{fleisch2010internet}.
+
+Theoretically, Iota's network benefits from high volume of transactions.
+The more transactions are coming, the faster previous transactions are confirmed.
+This is a major feature of Tangle which contrasts with Bitcoin's difficulty to scale.
+On the other hand, low volume of transactions is a primary problem for Iota, since it would take too long to confirm previous transactions.
+As it is a known problem, Iota has created a central coordinator which works as a trustworthy node, clearing transactions.
+In other words, every transaction directly or indirectly confirmed by the central coordinator is assumed to be cleared.
+Iota's developers claims that the coordinator will be turned off when the network outgrows a minimum (unknown) size.
+
+An important factor of Iota is how the new transactions choose which transactions they will confirm.
+There are several possible approaches, such as randomly selecting two of the unconfirmed transactions (tips).
+In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7.
+It may be on purpose, or maybe transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction.
+The selection algorithm seems to be important to protect the network against double spend attacks.
+For further information, see \cite{tangle2016}.
 
@@
 
-The accumulated weight of a transaction A is the sum of all weights of the transactions which confirm A, including A itself, i.e., $w_A + \sum_{A \leadsto P} w_P$.
-For example, in Fig.
-\ref{fig-tangle-example}, the accumulated weight of transaction 3 is the sum of the weights of the transactions 5, 6, 7, and 8.
-The accumulated weight may be interpreted as how hard it is to rollback a transaction.
+Transactions have an accumulated weight which may be interpreted as how hard it is to rollback a transaction.
 It is analogous to the number of confirmations of a block in Bitcoin.
-
+The higher the accumulated weight, the safer the transaction.
+Let $A$ be a transaction, its accumulated weight is the sum of all weights of the transactions which confirm $A$, including A itself, i.e., $w_A + \sum_{A \leadsto P} w_P$.
+For example, in Figure \ref{fig-tangle-example}, the accumulated weight of transaction 3 is the sum of the weights of the transactions 3, 5, 6, 7, and 8.
+
+The score of a transaction is a measure of how much proof-of-work has been done before the transaction has been created.
+As the heighest score of the network increases over time, comparing a transaction's score with the highest score of the network indicates the ``age'' of the transaction.
 The score of a transaction A is the sum of all weights of the transactions which are being confirmed by A, including A itself, i.e., $w_A + \sum_{P \leadsto A} w_P$.
 For example, in Fig.
-\ref{fig-tangle-example}, the score of transaction 3 is the sum of the weights of the transactions 1 and 2.
-It is a measure of how much proof-of-work has been done before confirming this transaction.
-It may indicate whether the confirmed transactions have received enough attention (and hashpower) or whether it may have received tangential attention.
-
+\ref{fig-tangle-example}, the score of transaction 3 is the sum of the weights of the transactions 1, 2, and 3.
+
+Another measure of the ``age'' of a transaction is its height.
 The height of a transaction A is the length of the longest path from transaction A to the genesis transaction.
@@
 \ref{fig-tangle-example}, the height of transaction 5 is four ($5 \rightarrow 3 \rightarrow 2 \rightarrow 1 \rightarrow$ genesis).
-It may be interpreted as the “`age'' of the transaction.
 The lower the height, the older the transaction.
 
-The depth of a transaction A is the length of the longest path in the inverted graph from transaction A to any unconfirmed transaction (tip).
+The depth of a transaction A is a measure of the youth of the transaction.
+It is the length of the longest path in the inverted graph from transaction A to any unconfirmed transaction (tip).
 For example, in Fig.
@@
 It is the opposite of the height.
-It may be interpreted as the youth of the transaction.
 The lower the depth, the younger the transaction.
-When a new transaction is confirming two transactions with high depth, it is referred as lazy transaction.
-
-An important factor of DAG model is how the new transactions choose which transactions they will confirm.
-There are several possible approaches, such as randomly selecting two of the unconfirmed transactions (tips).
-In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7.
-It may be on purpose, or maybe transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction.
-The selection algorithm seems to be important to protect the network against double spend attacks.
+When a new transaction is confirming two transactions with high depth, it is referred to as lazy transaction.
 
@@
 \ref{fig-tangle-swarm}, the reader can notice that the number of new transactions was increased for a while, and then decreased back to the original value.
-The DAG has behaved well when exposed to a high load scenario, since it reduced the number of tips to only three.
+The Iota has behaved well when exposed to a high load scenario, since it reduced the number of tips to only three after the high demand has ceased.
 It is like a moving swarm which gets wider when the number of new transactions increases and gets thinner when the number of new transactions decreases.
@@
 
-Conflicting transactions may happen when two or more transactions try to spend the ``same money'' --- or, in the Bitcoin's transaction format, try to spend the same output.
+Conflicting transactions may happen when two or more transactions try to spend the same tokens --- or, in the Bitcoin's transaction format, try to spend the same output.
 In this case, the network must choose which of the transactions will be accepted and the other one will be invalidated, even when both have already been confirmed.
@@
 \ref{fig-tangle-conflict}).
-But it may be not enough to prevent a nuclear submarine attack.
+But it may be not enough to prevent some attacks like the nuclear submarine attack.
 
@@
 
-\section{Nuclear submarine attack}
-
 The nuclear submarine attack, also known as the parasite chain attack, is when the attacker generates a separate DAG (or a side DAG), with many transactions and a lot of proof-of-work.
@@
 
-If the transactions have no conflict with any transaction of the main DAG, i.e., there is no transaction spending the ``same money'', then it seems easy to handle the transactions.
-But, as it is an attack, probably there will be some conflicts, and it is not easy to choose which transaction should be invalidated.
-As the attacker has been generating a separate DAG, the conflicting transaction may have an accumulated weight similar or greater than the transaction in the main DAG.
-Thus, using only the accumulated weight may not be enough to prevent this attack.
-
-For example, the attacker may generate a transaction which transfers all their funds to another address.
-Then, they start to generate many new transactions which confirms themselves and even confirms some of the transactions in the main DAG.
-But none of these transactions are propagated to the network.
-Then, the attacker buys something in the real world, pays with cryptocurrency, and wait until the payment gets the accumulated weight demanded by the merchant.
+If the transactions have no conflict with any transaction of the main DAG, i.e., there is no transaction spending the same tokens, then it is easy to handle the transactions.
+But, as it is an attack, there will be some conflicts, and it is not easy to choose which transaction should be invalidated.
+As the attacker has been generating a separate DAG, the conflicting transaction may have an accumulated weight similar or greater than the transaction which is already in the main DAG.
+Hence, using only the accumulated weight may not be enough to prevent this attack.
+
+For example, the attackers generate (and do not propagate) a transaction that transfers all their funds to another address.
+Then, they start to generate many new transactions which confirm themselves and even confirm some of the transactions in the main DAG, but none of these transactions are also propagated to the network.
+Afterward, the attacker buys something in the real world, pays with cryptocurrency, and wait until the payment gets the accumulated weight demanded by the merchant.
 Finally, the attacker suddenly propagates all the transactions to the network in a small window of time.
-This may cause the network to accept the attacker's original transaction instead of the one used to pay the merchant.
+If the criteria is to validate the transaction with higher accumulated weight, the network will accept the attackers' original transaction instead of the one used to pay the merchant.
 Hence, the merchant transaction is invalidated, and the double spend attack has succeeded.
 
-\section{Proposal: Questions to be explored}
+By default, Iota uses the Markov Chain Monte Carlo (MCMC) algorithm to select the two tips.
+For further information about attacks and strategies to prevent them, including the MCMC algorithm, see \cite{tangle2016}.
+
+\chapter{Hathor's architecture}
+
+This work introduces the Hathor's architecture which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
+Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG).
+For this, each transaction has its own proof-of-work which is solved by the issuer before propagating the transactions in the network.
+Like Bitcoin, miners find new ``blocks'' every 10 minutes in which they collect the fees and newly generated tokens.
+Each transaction has an ``accumulated weight'' which express the required effort to break the transaction, similar to Bitcoin's number of confirmations.
+
+This way, there will be two difficulty levels in the system: one for new transactions which are just moving tokens around, and another one for ``blocks'' which are generating new tokens and collecting fees.
+The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions.
+The second is adjusted every 2016 blocks to keep the pace of one block every 10 minutes.
+
+There is a trade-off about the required proof-of-work of new transactions.
+The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.
+
+Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
+It is always possible given that there is at least one confirmation path between any transaction and a tip.
+
+The transactions are classified into three groups: (i) confirmed transactions, (ii) confirming transactions, and (iii) unconfirmed transactions (tips).
+The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
+The unconfirmed transactions (tips) are the very new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
+The confirming transactions are in the middle.
+They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.
+
+A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two confirming transactions or tips.
+There may be any number of outputs provided that they comply with the number of newly generated tokens.
+Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.
+
+In the low load scenario, there is a small number of new transactions coming into the network, which means they give a minor contribution to confirmations.
+In this case, the confirmation is held mostly by ``blocks''.
+On the other hand, in the high load scenarios, there is a large number of new transactions giving a major contribution to confirmations.
+In this case, the ``blocks'' strengthen the confirmation, but most of them will have already been confirmed before the new blocks are found.
+The higher the number of new transaction, the faster the transactions are confirmed.
+The blocks assure a ``maximum confirmation time''.
+
+The incentive schema which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
+
+Let $A$ be a transaction, $w_A$ be its weight, and $\text{acc}_A$ be its accumulated weight. Then, $\text{acc}_A = w_A + \sum_{A \leadsto P} w_P$.
+
+\chapter{Proposal: Questions to be explored}
+ The main question which this work proposes to answer is: Is it reliable to use DAG model? In which conditions?
 
@@
 
-\section{Hathor's Introduction}
-
-Hathor's architecture lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
-Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG).
-For this, each transaction has its own proof-of-work which is solved by the issuer before propagating the transactions in the network.
-Like Bitcoin, miners find new ``blocks'' every 10 minutes in which they collect the fees and newly generated tokens.
-Each transaction has an ``accumulated weight'' which express the required effort to break the transaction, similar to Bitcoin's number of confirmations.
-
-This way, there will be two difficulty levels in the system: one for new transactions which are just moving tokens around, and another one for ``blocks'' which are generating new tokens and collecting fees.
-The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions.
-The second is adjusted every 2016 blocks to keep the pace of one block every 10 minutes.
-
-There is a trade-off about the required proof-of-work of new transactions.
-The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.
-
-Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
-It is always possible given that there is at least one confirmation path between any transaction and a tip.
-
-The transactions are classified into three groups: (i) confirmed transactions, (ii) confirming transactions, and (iii) unconfirmed transactions (tips).
-The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
-The unconfirmed transactions (tips) are the very new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
-The confirming transactions are in the middle.
-They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.
-
-A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two confirming transactions or tips.
-There may be any number of outputs provided that they comply with the number of newly generated tokens.
-Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.
-
-In the low load scenario, there is a small number of new transactions coming into the network, which means they give a minor contribution to confirmations.
-In this case, the confirmation is held mostly by ``blocks''.
-On the other hand, in the high load scenarios, there is a large number of new transactions giving a major contribution to confirmations.
-In this case, the ``blocks'' strengthen the confirmation, but most of them will have already been confirmed before the new blocks are found.
-The higher the number of new transaction, the faster the transactions are confirmed.
-The blocks assure a ``maximum confirmation time''.
-
-The incentive schema which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
-
-Let $A$ be a transaction, $w_A$ be its weight, and $\text{acc}_A$ be its accumulated weight. Then, $\text{acc}_A = w_A + \sum_{A \leadsto P} w_P$.
-
 \chapter{Conclusion}

### 2018-04-10 Chapters/Chapter01.tex

@@
 
-The transactions are classified into three groups: (i) confirmed transactions, (ii) confirming transactions, and (iii) unconfirmed transactions (tips).
+The transactions are classified into three groups: (i) confirmed transactions, (ii) in-progress transactions, and (iii) unconfirmed transactions (tips).
 The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
-The unconfirmed transactions (tips) are the very new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
-The confirming transactions are in the middle.
+The unconfirmed transactions (tips) are the brand new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
+The in-progress transactions are in the middle.
 They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.
 
-A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two confirming transactions or tips.
+A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two in-progress transactions or tips.
 There may be any number of outputs provided that they comply with the number of newly generated tokens.
@@
 
-The incentive schema which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
-
-Let $A$ be a transaction, $w_A$ be its weight, and $\text{acc}_A$ be its accumulated weight. Then, $\text{acc}_A = w_A + \sum_{A \leadsto P} w_P$.
+The incentive scheme which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
+
+The concepts of weight and accumulated weight are the same as Iota's, but their equations are different.
+The weight of a transaction is calculated as $\log_2(k)$ where $k$ is the average number of hashes required to solve its proof-of-work again, while the accumulated weighted of a transaction A with weight $w_A$ is calculated as $\log_2(2^{w_A} + \sum_{A \leadsto P} 2^{w_P})$.
+The $\log$ function is used to reduce the magnitude of the numbers.
+Hence, the power-of-2 of the accumulated weight is a measure of the average number of hashes required to solve the proof-of-work of all transactions that are confirming A.
+
+In Bitcoin, it is usually required 6 confirmations before accepting a transaction.
+Let $H$ be the total hash rate of the network.
+Then, as $\mathbf{E}(Y_6) = 60 \text{ minutes}$, it will be necessary to calculate, on average, $\mathbf{E}(Y_6) \cdot H = 60 \cdot 60 \cdot H$ hashes to solve the proof-of-work of 6 blocks.
+Finally, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(60 \cdot 60 \cdot H) = \log_2(3600) + \log_2(H) = 4 + \log_2(225) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
 

### 2018-04-10 Chapters/Chapter02.tex
sdm re-org
recall theory intro

@@
 
-Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
-This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present.
-For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where one's fear comes from.
+Suddenly, you feel the danger.
+You tighten your grip on the child's hand and step up your walking.
+In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everthing you think of is to get out to safety.
+You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exatcly is causing your fear.
 One just feels it.
@@
 Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger.
+The whole process happens so fast it feels like it only took a single unit of psychological time.
+
+Let us compare this to what a child might do.
+
+Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present.
+
 This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.
+
+\section{Desiderata for a theory of memory}
+
+Psychological
+miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue... how does this recall process work?
+
+Neuroscientific (bioplausible)
+neurons can die; what do neurons do?  how are things stored?  how are they recalled?
+
+Computational (fully experimental, with no voodoo-like mysterious properties)
+
+3 sparse distributed memory 11
+
+3.1 Neurons as pointers 17
+
+3.2 Concepts 17
+
+3.3 Read operation 18
+
+3.3.1 Generalized read operation
+
+3.4 Critical Distance 20
+
+4 framework architecture 23
+
+4.1 Bitstring 24
+
+4.2.2 OpenCL kernels
+
+4.3 Counters 27
+
+4.4 Read and write operations
+
+5 results (i): performance
+
+5.1 Kernels comparison 38
+
+5.2 Scanners comparison 39
+
+5.3 Read and write operations
+
+6.1 Distance between random bitstrings
+
+6.2 Number of activated hard locations
+
+6.3 Intersection of two circles 63
+
+6.4 Storage and retrieval of sequences
+
+5.3.1 Summary of results
+
+6 results (ii): framework validation
+
+4.1.1 The distance between two bitstrings 24
+
+4.2 Address space 25
+
+4.2.1 Scanning for activated hard locations 25
+
+6.4.1 k-fold memory using only one SDM
+
+7 results (iii): loss of neurons 67
+
+8 results (iv): critical distance 73
+
+8.1 A deviation from the equator distance? 77
+
+8.2 Counter bias 80
+
+8.3 Read bias 83
+
+8.4 Critical distance of 209 85
+
+9 results (v): generalized read operation 87
+
+10 results (vi): supervised classification
+application
+
+11 results (vii): image noise filtering application 101
+
+12 results (viii): the possibility of unsupervised reinforcement learning 107
+
+13 results (ix): information-theoretical write operation 115
+
+\section{The wasted effort of duplicated work}
 
@@
 Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate effort from researchers to build on top of previous work.
-
-\section{The wasted effort of duplicated work}
 
@@
 
-\chapter{Results (i): Framework validation}
+\chapter{Results (i): Performance}
+\input{./Chapters/sdm-performance.tex}
+
+\chapter{Results (ii): Framework validation}
 \input{./Chapters/sdm-validation.tex}
-
-\chapter{Results (ii): Critical distance}
-\input{./Chapters/sdm-critical-distance.tex}
 
@@
 
-\chapter{Results (iv): Generalized read operation}
+\chapter{Results (iv): Critical distance}
+\input{./Chapters/sdm-critical-distance.tex}
+
+\chapter{Results (v): Generalized read operation}
 
@@
 
-\chapter{Results (v): Performance}
-\input{./Chapters/sdm-performance.tex}
+%\chapter{Results (v): Performance}
+%\input{./Chapters/sdm-performance.tex}
 

### 2018-04-10 Chapters/sdm-applications.tex
recall theory intro

@@
 
-The images had 31 pixels of width and 32 pixels of height, totaling 992 pixels per image.
+The images had 31 pixels of width and 32 pixels of height, totaling 992 pixels per image\footnote{A crucial note is in order.
+While others have proposed similar models \citep{hong_character_1991}, in no way is it claimed here that these bitstrings should form a plausible representation of letters in the human mind (See, for instance, \citet{hofstadter1995seeing} for a marvelous discussion of the subtleties and fluidity involved in that process).
+These letters should merely be seen as \emph{invariant patterns} that are processed by the memory, which is able to capture their underlying invariant structure even when presented only under heavy noise.
+Recall that any rezising, any rotation, or any mere shifting of these letters will take them out of the critical distance for SDM reading.}.
 Each image was mapped into a 1,000-bit bitstring in which the bits were set according to the color of each pixel of the image.
@@
 There is one image for each of the 62 groups in Figure \ref{fig-classification-groups}.
-Notice that the SDM has never seen a single image with no noise.
+Notice that the memory has never seen a single image without noise.
 
@@
 
-A final note is in order.
-In no way is it claimed here that these bitstrings should form a plausible representation of letters in the human mind (See, for instance, \citet{hofstadter1995seeing} for a marvelous discussion of the subtleties and fluidity involved in that process).
-These letters should merely be seen as \emph{invariant patterns} that are processed by the memory, which is able to capture their underlying invariant structure even when presented only under heavy noise.
-
 \chapter{Results (viii): The possibility of unsupervised reinforcement learning}

### 2018-04-10 Chapters/Chapter01.tex
Blockchain: Hathor's architecture.

@@
 
-The present work intends to analyze a new architecture, named Hathor, which lies between Bitcoin and Iota and may be a viable solution to both scaling, centralization, and spam problems.
+The present work intends to propose and analyze a new architecture, named Hathor, which lies between Bitcoin and Iota and may be a viable solution to both scaling, centralization, and spam problems.
 We also present a mathematical analysis of Bitcoin's architecture.
@@
 The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions.
-The second is adjusted every 2016 blocks to keep the pace of one block every 10 minutes.
+The second is adjusted every 2,016 blocks to keep the pace of block on every 2 minutes.
 
@@
 Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
-It is always possible given that there is at least one confirmation path between any transaction and a tip.
+It is always possible since there is at least one confirmation path between any transaction and a tip.
+This is important for security reasons.
 
@@
 Then, as $\mathbf{E}(Y_6) = 60 \text{ minutes}$, it will be necessary to calculate, on average, $\mathbf{E}(Y_6) \cdot H = 60 \cdot 60 \cdot H$ hashes to solve the proof-of-work of 6 blocks.
-Finally, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(60 \cdot 60 \cdot H) = \log_2(3600) + \log_2(H) = 4 + \log_2(225) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
+Therefore, in order to have the same level of security as Bitcoin, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(60 \cdot 60 \cdot H) = 4 + \log_2(225) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
+
+\section{Time between blocks}
+
+The hash function used for the proof-of-work (PoW) is the same as Bitcoin: \emph{SHA-256} applied twice. Thus, most of the math analysis we have done before is just the same.
+
+Let $X$ be the number of trials to solve the PoW and $T$ be the time between blocks.
+We already know that $X$ follows a geometric distribution with $p = \frac{A}{2^{256}}$, where $A$ is inversely proportional to the difficulty, i.e., the smaller the $A$, the higher the difficulty.
+We also know that $T$ follows an exponential distribution with $\lambda = \frac{1}{\eta}$, where $\eta$ is the average time between blocks.
+As proved before, $\eta p H = 1$, thus let's define $w = \log_2(\mathbf{E}(X)) = \log_2(1/p) = \log_2(\eta H) = \log_2(\eta) + \log_2(H)$.
+
+For the blocks, $\eta = 128$, thus, $\mathbf{E}(T) = 128 \text{ seconds}$, and $\mathbf{V}(T) = 16,384$.
+The symmetrical confidence interval with $\alpha = 10\%$ is $[6.56, 383.45]$, i.e., 90\% of the cases the distance between blocks will be between 6 seconds and just under 7 minutes.
+
+Let $H$ be the hash rate of the miners, then, the weight of the blocks is calculated by
+
+$$w_\text{blocks} = \log_2(128) + \log_2(H) = 7 + \log_2(H)$$
+
+This weight will be updated every 2,016 blocks to take into consideration the change of $H$.
+First, $H$ will be estimated using the fact that $\mathbf{E}(Y_{2016}) = \frac{2016}{\lambda} = 2016 \eta = \frac{2016}{pH} = \frac{2016 \cdot 2^w}{H}$.
+Thus, $H = \frac{2016 \cdot 2^w}{\Delta t} \Rightarrow \log_2(H) = w + \log_2(2016) - \log_2(\Delta t)$, where $\Delta t$ is the time between the latest weight update and now.
+Finally,
+
+$$w_\text{new} = 7 + w_\text{old} + \log_2(2016) - \log_2(\Delta t)$$
+
+Notice that, if the hash rate does not change, then $\Delta t = 128 \cdot 2016$ and $w_\text{new} = w_\text{old}$.
+
+\section{Weight of the transactions}
+
+There is a trade-off which must be considered in the weight of the transactions: the higher the weight, the better to prevent spams, but the worse to microtransactions.
+So, trying to fulfill both necessities, the weight will be a function of the transaction's size (in bytes) and total amount:
+
+$$w_\text{tx} = \left\lfloor \log_2(\text{size}) + \log_2(\text{amount}) + 0.5 \right\rfloor$$
+
+Although the transaction size may change a bit depending on the implementation, a typical transaction with 2 inputs and 2 outputs would have 188 bytes.
+So, for instance, transfering 50,000 tokens would require a weight of 23, which means an average of $2^{23}$ trials to solve the PoW.
+
+\section{Issuance rate}
+
+Bitcoin issuance rate started in 50 tokens per block and decreases over time.
+Every 210,000 blocks---4 years, on average---it halves the number of tokens issued per block.
+As Bitcoin smallest fraction is $10^{-8}$, after the 33th reduction it will stop issuing new tokens since $2^{33} > 50 \cdot 10^{-8}$.
+In total, the number of issued tokens will never exceed 21 million.
+For further information, see \cite{bitcoinsupply}.
+
+Using the same strategy, Hathor's issuance rate starts in 50,000 tokens per block and decreases over time.
+Different from Bitcoin, Hathor's tokens are always integers and cannot be split.
+Every 1,000,000 blocks, the number of tokens issued per block is halved.
+Thus, the supply is limited to 100 billion issued tokens during its whole life.
+To calculate the supply limit use the following equation:
+
+\begin{align*}
+    \sum_{k=0}^{\infty} 1,000,000 \cdot \left\lfloor \frac{50000}{2^k} \right\rfloor \approx
+    \sum_{k=0}^{\infty} 1,000,000 \cdot \frac{50000}{2^k} = 100 \text{ billion}
+\end{align*}
+
+As the average time between blocks is 128 seconds, it will take around 3 years, 9 months and 20 days between halves.
+
+\section{Transaction fees}
+
+The value of the fee is calculate as the difference between the transaction's outputs and the inputs.
+For instance, a transaction with inputs summing 10,000 HTR and outputs summing 9,000 HTR are paying 1,000 HTR of fees.
+
+Each transaction may pay a fee to the next block which confirms it.
+But, even if the fee is zero, the miners are forced to confirm at least two pending transactions (in-progress or unconfirmed), and these pending transactions confirms other transactions which may have no fees.
+In summary, transactions with no fees cannot be left behind and will always be confirmed by both blocks and other transactions.
+
+One may arguee that it allows the whole network to never pay fees and they are right.
+In the beginning, miners' incentive is driven by new tokens instead of fees.
+In the long term, it may be necessary to require a minimum fee to keep the miners working.
+
+\section{Transaction validation}
+
+A transaction will be considered valid when it complies with the following rules: (i) it spends only unspent outputs; (ii) the sum of the inputs is greater than or equal to the sum of the outputs; (iii) the number of inputs is at most 256; (iv) the number of outputs is at most 256; (v) it confirms at least two pending transactions; (vi) it solves the PoW with the correct weight.
+
+The transaction has a time field which is used to inform when the transaction was generated.
+The time field must be in UTC time and may be within at most 60 minutes from the current time.
+Otherwise, the transaction will be discarded.
+
+In case of transaction conflict, in which two transactions try to spend the same tokens, the one with higher accumulated weight is chosen and the other is invalidated.
+Although it is not possible in Iota because of the submarine attack, Hathor does not have the same problem.
+In Hathor, like in Bitcoin, the submarine attack is only possible if the attacker has a hash rate higher than the whole network, including the miners.
+In other words, when analysing the double spending attack, Hathor is as safe as Bitcoin.
+
+The digital signature is used to ensure that only the owners may spend their tokens.
+It will be calculated signing the transaction's input and output only.
+This allows the transaction to be signed in one device and to be sent to another device that will choose which transaction will be confirmed and will solve the proof-of-work.
+
+Services of solving proof-of-work may also be offered by companies.
+They give their customers a wallet address and they send the payment inside of the transaction itself.
+This allows IoT devices to save energy or to delegate the task of solving the PoW to whoever is receiving the tokens.
+
+\section{Orphan blocks}
+
+There is no orphan blocks in Hathor.
+Unless a block confirms either directly or indirectly an invalid transaction, every block is valid.
+There is no need to left a block behind since its proof-of-work is also protecting transactions.
+
+\section{Network}
+
+Like in all other cryptocurrencies, nodes will communicate through a peer-to-peer network.
+They will exchange transactions and blocks, verifying whether they are valid or not, while they will also answer queries.
+
+Thus, if a thin client wants to transfer some tokens, it just connects to the network and queries for the tips.
+With the response, it solves the proof-of-work and propagates the new transaction.
+In order to avoid malicious nodes, the thin client may ask to more than one node and randomly pick two transactions.
 
 \chapter{Proposal: Questions to be explored}
- The main question which this work proposes to answer is: Is it reliable to use DAG model? In which conditions?
 
@@
 In this paper, we will analyze the network scaling and security.
-How a DAG network scales, simulating different loads and measuring the bandwidth, computational effort, and storage space necessary to handle all the transactions in time.
+How a Hathor network scales, simulating different loads and measuring the bandwidth, computational effort, and storage space necessary to handle all the transactions in time.
 What is the minimum transaction's aggregated weight which it may be considered unlikely to be reversed?
@@
 
-As the DAG network is distributed, its users may use different parameters --- they cannot violate the network's rules, but they may use different tip selection algorithms, for instance.
-We would like to explore how different users' parameters may affect the network, how much is necessary to cause either a hard-fork or a soft-fork.
-
-\cite{tangle2016} suggests that constraining transactions' weight in a range would both prevent spam, because it would be necessary a minimum work to propagate a new transaction, and attacks, because an attacker would not be allowed to propagate a transaction with very high weight.
+What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
+
+%\cite{tangle2016} suggests that constraining transactions' weight in a range would both prevent spam, because it would be necessary a minimum work to propagate a new transaction, and attacks, because an attacker would not be allowed to propagate a transaction with very high weight.
 We agree with the spam argument, and would like how a minimum weight would affect mobile devices's new transactions.
@@
 
-We have another suggestion to prevent nuclear submarine attacks: set a maximum depth for the transactions confirmed by a new transaction.
+%We have another suggestion to prevent nuclear submarine attacks: set a maximum depth for the transactions confirmed by a new transaction.
 So, new transactions would have to confirming newer transactions instead of old transactions, and a nuclear submarine attack would be controlled because the attacker would be not be able to create a large separate DAG.
@@
 
-If we would like to include fee transactions, how would it be distributed between confirmations?
+%If we would like to include fee transactions, how would it be distributed between confirmations?
 As transactions may have multiple direct confirmations, it is not obvious who would receive the transaction fee.
@@
 
-In order to solve the fee distribution problem, we will simulate how the network would behave if we include blocks.
+%In order to solve the fee distribution problem, we will simulate how the network would behave if we include blocks.
 These blocks would be like Bitcoin's, with an adjustable proof-of-work according to the network's hashpower.
@@
 
-\chapter{Analysis of DAG Model}
+\chapter{Analysis of Hathor}
 

### 2018-04-11 Chapters/Chapter01.tex
Hathor: simulations results (only images)

@@
 
+\section{Transaction confirmation}
+
 The concepts of weight and accumulated weight are the same as Iota's, but their equations are different.
@@
 Then, as $\mathbf{E}(Y_6) = 60 \text{ minutes}$, it will be necessary to calculate, on average, $\mathbf{E}(Y_6) \cdot H = 60 \cdot 60 \cdot H$ hashes to solve the proof-of-work of 6 blocks.
-Therefore, in order to have the same level of security as Bitcoin, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(60 \cdot 60 \cdot H) = 4 + \log_2(225) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
+Therefore, in order to have the same level of security as Bitcoin, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(6 \cdot 128 \cdot H) = 7 + \log_2(6) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
 
@@
 
-The methodology we have been using is computer simulation.
-Through the simulation of many scenarios of DAG model with different parameters, we will understand how the network behaves in complex scenarios, mainly when the load is suddenly increased, and when the network is under attack.
-We will first test our hypothesis through simulation and than prove them mathematically.
+The methodology we have used is computer simulation.
+Through the simulation of many scenarios of Hathor, we will understand how the network behaves in complex scenarios, including when the load suddenly increases, and when the network is under attack.
 
@@
 
-When a new transaction is added to the DAG, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
+When a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work.
+When it finishes solving the proof-of-work, it is propagated and becomes a tip.
+So, two new transactions may choose the same tips to confirm.
+If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
+
+When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
 The topological sort is used to calculate the longest path from all other transactions to the new transaction, and this longest path is used to update the depth of the whole DAG.
@@
 
-The necessary time to generate the proof-of-work is estimated by the function $k \cdot 2^{weight}$, where $k$ is a constant related to the computing power of the agent and weight is randomly chosen.
-
-The time between two new transactions is sampled from the exponential distribution with parameter $\lambda$.
+The amount of time spent to solve the proof-of-work is a sample from an exponential distribution.
+
+The time between two new transactions is sampled from an exponential distribution with parameter $\lambda$.
 The value of $\lambda$ is changed over time, causing changes in the load of the network.
@@
 
-The DAG Model's mining starts just like Bitcoin's, i.e., $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.
-Then, it follows that solving a transaction proof-of-work follows a geometric distribution.
-Let $X$ be the number of attempts until a success, then $\mathbf{P}(X=k) = (1-p)^{k-1}p$, where $p=\frac{A}{2^{256}}$.
-Also, $\mathbf{P}(X \leq k)=1-(1-p)^k$.
-The average number of attempts is $\mathbf{E}(X)=1/p$.
-
-Let $w$ be the transaction's weight, i.e., $\mathbf{E}(X)=w \Rightarrow w = 1/p$.
-Let $T = X/H$ be the time required to solve the transaction's proof-of-work, where $H$ is the hash rate of the transaction's owner.
-Then, $\mathbf{P}(T < t) = \mathbf{P}(X < tH) = 1 - (1 - p)^{tH} = 1 - \left( 1 - \frac{1}{w} \right)^{tH} = 1 - e^{tH \log\left(1 - \frac{1}{w} \right)}$, i.e., $T$ follows an exponential distribution with $\lambda = -H \log\left(1 - \frac{1}{w} \right)$.
-
-% https://math.stackexchange.com/questions/635443/sum-of-exponential-random-variable-with-different-means
-
 The number of transactions solving the proof-of-work may be modeled by a Birth and Death process, since if we have $k$ transactions solving the proof-of-work only two things may happen: either a new transaction will be created or one the transactions will finish solving the proof-of-work.
@@
 Hence, the probability of any transaction finishes its proof-of-work before a new transaction emerges is $p_k = 1 - q_k = \frac{\sum_{i=1}^k \lambda_i}{\mu + \sum_{i=1}^k \lambda_i}$.
-
-Every time a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work.
-When it finishes solving the proof-of-work, it is propagated and becomes a tip.
-So, two new transactions may choose the same tips to confirm.
-If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
-
-The following theorems will be mathematically proved. They have already been used in the implementation of the simulator.
 

### 2018-04-11 Chapters/Chapter01.tex
Hathor: minor text update.

@@
 
-In this paper, we will analyze the network scaling and security.
+What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
+
+%In this paper, we will analyze the network scaling and security.
 How a Hathor network scales, simulating different loads and measuring the bandwidth, computational effort, and storage space necessary to handle all the transactions in time.
@@
 
-We are interested in how the rate of new transactions affects how long it takes to a new transaction to be confirmed for the first time.
+%We are interested in how the rate of new transactions affects how long it takes to a new transaction to be confirmed for the first time.
 We had already run some simulations under normal load (Fig.
@@
 
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist.png}
-\caption{Histogram of how long has been a transaction waiting until its first confirmation.
+%\begin{figure}[ht]
+%\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist.png}
+%\caption{Histogram of how long has been a transaction waiting until its first confirmation.
 It was a simulation of 15 minutes with new transactions rate changing between 1 and 15 tx/s.\label{fig-tangle-hist}}
-\end{figure}
-
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist-2.png}
-\caption{Histogram of how long has been a transaction waiting until its first confirmation.
+%\end{figure}
+
+%\begin{figure}[ht]
+%\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist-2.png}
+%\caption{Histogram of how long has been a transaction waiting until its first confirmation.
 It was a simulation of 5 minutes with new transactions rate of 50 tx/s, i.e., very high load.\label{fig-tangle-hist-2}}
-\end{figure}
-
-Besides the time to the first confirmation, it is also important to measure how long it takes to a new transaction reach a specific accumulated weight.
+%\end{figure}
+
+%Besides the time to the first confirmation, it is also important to measure how long it takes to a new transaction reach a specific accumulated weight.
 It is useful for exchanges and merchants to set their minimum requirements.
@@
 
-What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
-
-%\cite{tangle2016} suggests that constraining transactions' weight in a range would both prevent spam, because it would be necessary a minimum work to propagate a new transaction, and attacks, because an attacker would not be allowed to propagate a transaction with very high weight.
+%OLD \cite{tangle2016} suggests that constraining transactions' weight in a range would both prevent spam, because it would be necessary a minimum work to propagate a new transaction, and attacks, because an attacker would not be allowed to propagate a transaction with very high weight.
 We agree with the spam argument, and would like how a minimum weight would affect mobile devices's new transactions.
@@
 
-%We have another suggestion to prevent nuclear submarine attacks: set a maximum depth for the transactions confirmed by a new transaction.
+%OLD We have another suggestion to prevent nuclear submarine attacks: set a maximum depth for the transactions confirmed by a new transaction.
 So, new transactions would have to confirming newer transactions instead of old transactions, and a nuclear submarine attack would be controlled because the attacker would be not be able to create a large separate DAG.
@@
 
-%If we would like to include fee transactions, how would it be distributed between confirmations?
+%OLD If we would like to include fee transactions, how would it be distributed between confirmations?
 As transactions may have multiple direct confirmations, it is not obvious who would receive the transaction fee.
@@
 
-%In order to solve the fee distribution problem, we will simulate how the network would behave if we include blocks.
+%OLD In order to solve the fee distribution problem, we will simulate how the network would behave if we include blocks.
 These blocks would be like Bitcoin's, with an adjustable proof-of-work according to the network's hashpower.

### 2018-04-12 Chapters/Chapter01.tex
Hathor: some results and cleaning.

@@
 
-In Bitcoin, it is usually required 6 confirmations before accepting a transaction.
+In Bitcoin, it is well-known that one should wait at least ``six confirmations'' before accepting a transaction.
+This Bitcoin's criteria is based on some math presented in Satoshi's seminal work \citep{nakamoto2008bitcoin}.
+Adopting six confirmations is the same as demanding from attackers a minimal effort of six times the network's hash rate to successfully double spend those tokens.
 Let $H$ be the total hash rate of the network.
 Then, as $\mathbf{E}(Y_6) = 60 \text{ minutes}$, it will be necessary to calculate, on average, $\mathbf{E}(Y_6) \cdot H = 60 \cdot 60 \cdot H$ hashes to solve the proof-of-work of 6 blocks.
+
 Therefore, in order to have the same level of security as Bitcoin, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(6 \cdot 128 \cdot H) = 7 + \log_2(6) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
@@
 
-\chapter{Proposal: Questions to be explored}
-
-Given these preliminaries, these are some questions with which we will be concerned.
-
-What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
+\chapter{Methodology}
+
+The methodology we have used is computer simulation.
+Through the simulation of many scenarios of Hathor, we will understand how the network behaves in complex scenarios, including when the load suddenly increases, and when the network is under attack.
+
+The simulator has been developed using an event-based design which is capable of running hours of simulation in just a few minutes.
+It creates agents who decide to make a transaction, then they select which transactions will be confirmed, then they spend some time working in the proof-of-work, and, finally, they propagate the transaction to the network.
+The other agents receive the transaction and may accept or deny it.
+The agents may use different parameters among themselves.
+
+When a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work.
+When it finishes solving the proof-of-work, it is propagated and becomes a tip.
+So, two new transactions may choose the same tips to confirm.
+If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
+
+When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
+The topological sort is used to calculate the longest path from all other transactions to the new transaction, and this longest path is used to update the depth of the whole DAG.
+If the new transaction confirms only transactions which has already been confirmed, the depth update is skipped thanks to Theorem \ref{theorem-new-tx-not-tip}.
+
+The amount of time spent to solve the proof-of-work is a sample from an exponential distribution.
+
+The time between two new transactions is sampled from an exponential distribution with parameter $\lambda$.
+The value of $\lambda$ is changed over time, causing changes in the load of the network.
+We consider 1 tx/s to be a low load, and 15 tx/s to be a high load.
+Even though we have already tested up to 100 tx/s.
+
+The simulator will output different reports: (i) snapshots of the DAG at interesting moments, such as Fig.
+\ref{fig-tangle-example}, \ref{fig-tangle-swarm}, and \ref{fig-tangle-conflict}; (ii) histogram of confirmation time, such as Fig.
+\ref{fig-tangle-hist} and \ref{fig-tangle-hist-2}; and many others.
+
+\chapter{Analysis of Hathor}
+
+Hathor's architecture lies between Iota and Bitcoin's architectures.
+It is similar to Bitcoin's architecture when the number of transactions per second is low, while it is similar to Iota's architecture when the number of transactions per second is high.
+In order to check this statement, I have run some simulations.
+
+First, two simulations with the extreme cases: (i) no miners, (ii) no transactions.
+The first should have exactly the same behavior as Bitcoin's blocks with one block after the other forming a long chain.
+The latter should have a similar behavior as Iota's structure, forming a Direct Acyclic Graph (DAG) with transactions confirming transactions.
+We can notice in Figure \ref{fig:hathor-similarities} that both cases seem to be correct.
+
+Then, I have run a simulation with both miners and transactions.
+As we can see in Figure \ref{fig:hathor-dag}, Hathor's structure is a mix of Iota and Bitcoin's structure.
+The transactions are forming a DAG while, in parallel, the blocks are forming a chain.
+Differently from Bitcoin, there is no orphan block which means some blocks may be found close to the previous one.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[No miners]{\includegraphics[width=0.5\textwidth]{./images01/sim/no_miners.pdf}}
+\subfloat[No transactions]{\includegraphics[width=0.5\textwidth]{./images01/sim/only_miners.pdf}}
+\caption{Visualization of a Hathor's graph in two particular cases: (a) no miners, (b) no transactions.
+It shows that when there are no miners, Hathor is similar to Iota (same structure, but different parameters), and when there is no transactions, it is similar to Bitcoin.\label{fig:hathor-similarities}}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images01/sim/hathor.pdf}
+\caption{Visualization of a Hathor's graph with transactions and blocks. Red boxes are blocks, and white circles are simple transactions. The arrows show the confirmations.\label{fig:hathor-dag}}
+\end{figure}
+
+Another evidence that Hathor lies between Iota and Bitcoin is found when comparing the time to confirm a transaction.
+In this context, a transaction is said to be confirmed when it has reached an accumulated weight similar to six times the hash rate of the whole network (miners and new transactions).
+This criteria is equivalent to the well-known ``6 confirmations'' of Bitcoin, which is adopted by almost the whole network.
+
+Thus, I have run a simulation in which miners are majority and there are few transactions.
+In Figure \ref{fig:hathor-tct-low-mid}, we may see a good fit between the confirmation time of a transaction and the theoretical distribution of the time to find six blocks in Bitcoin (which is $Y_6$ and follows an Erlang distribution).
+When the load is increased, Hathor's confirmation time is reduced and diverges from Bitcoin's distribution.
+The reasoning is that confirmations coming from other transactions start to play an important role and accelerates the speed of confirmations.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Low load]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-low-load.png}}
+\subfloat[Mid load \label{fig:hathor-tct-mid-load}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-mid-load.png}}
+
+\caption{Confirmation time in two scenarios: (a) low load, (b) mid load.
+The red curve is the distribuion of the time to find six blocks in Bitcoin (which follows an Erlang distribution).
+As we can notice, in the low load scenario, Hathor's confirmation time behaves just like Bitcoin's.
+When the load is increased, it starts to diverge from Bitcoin's distribution.
+\label{fig:hathor-tct-low-mid}}
+\end{figure}
+
+We may see Hathor's confirmation time moving from Bitcoin's to Iota's in Figure \ref{fig:hathor-tct-many}.
+Notice that the confirmation timer is getting smaller as the number of transactions per second increases.
+Figure \ref{fig:hathor-tct-many-right} is a zoom-in in the right side, and we can see again the good fit between Hathor Bitcoin's confirmation time under low load.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[The whole picture]{\includegraphics[width=\textwidth]{./images01/sim/tct-many-loads-2.png}}
+
+\subfloat[Zoom in the left part of the above chart]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-many-loads-3.png}}
+\subfloat[Zoom in the right part of the above chart \label{fig:hathor-tct-many-right}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-many-loads-4.png}}
+
+\caption{Confirmation time in many scenarios, moving from a low load ($\lambda_\text{TX} = 0.015625$) to a high load ($\lambda_\text{TX} = 2$). \label{fig:hathor-tct-many}}
+\end{figure}
+
+But, what would happen if, instead of changing the number of transactions per second, we change the relative hash power between miners and transactions?
+In previous simulations, the miners had a hash rate in the same magnitude as the transactions.
+In Figure \ref{fig:hathor-tct-100k-10k}, we can see the same simulation as in Figure \ref{fig:hathor-tct-mid-load}, but with the miners' hash rate ten times the transactions'.
+Besides the difference in the shape of the distribution, we can see that it is moving back towards Bitcoin's confirmation time distribution.
+It also makes sense because increasing the miners' hash rate increases the required minimum accumulated weighted for confirmed transactions.
+Therefore, more transactions are necessary to give more accumulated weight.
+As the number of transactions per second was not changed, most of the work of confirmations were done by blocks (and not by transactions).
+To confirm this idea, I kept the miners' hash rate ten times the transactions' and increased 16 times the number of transactions per second.
+As we can see in Figure \ref{fig:hathor-tct-100k-10k-high}, when the number of transactions per second is increased, its role in the accumulated weight also increases and it goes farther from Bitcoin's distribution.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Same load as in Figure \ref{fig:hathor-tct-mid-load}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-100k-10k.png}}
+\subfloat[High load, 16 times of (a)\label{fig:hathor-tct-100k-10k-high}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-100k-10k-high-load.png}}
+
+\caption{Confirmation time with miners' hash rate ten times the transactions'. \label{fig:hathor-tct-100k-10k}}
+\end{figure}
+
+Finally, even with both blocks and transactions, Hathor's blocks are similar to Bitcoin's blocks, and they share the same math.
+To confirm that, see Figure {fig:hathor-tbb}, where the red curve is Bitcoin's distribution of time between blocks and the blue histogram is Hathor's time between blocks.
+I also made several tests adding and removing miners to check the difficulty adjustment and it worked properly.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[1 miner]{\includegraphics[width=0.5\textwidth]{./images01/sim/tbb-1.png}}
+\subfloat[2 miners after difficulty adjustment]{\includegraphics[width=0.5\textwidth]{./images01/sim/tbb-2-after.png}}
+
+\caption{Histogram of time between blocks. The red curve the Bitcoin's theoretical distribution of time between blocks. As we can notice, the fit is very good. \label{fig:hathor-tbb}}
+\end{figure}
+
+Afterward, to a better visualization of a Hathor's network, I run a simulation classifying the transactions in either confirmed, in-progress, or unconfirmed (tips).
+I also showed the transactions solving the proof-of-work which had not been propagated yet.
+See Figure \ref{fig:hathor-dag-big}} and notice the chain of blocks inside the DAG.
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images01/sim/hathor-2.pdf}
+\caption{Visualization of a Hathor's graph with transactions and blocks.
+Red boxes are blocks; green circles are confirmed transactions; white circles are in-progress transactions; yellow circles are unconfirmed transactions (tips); and grey circles are transactions solving the proof-of-work which have not been propagated yet.
+The arrows show the confirmation chain.
+Block's arrows are in bold.
+\label{fig:hathor-dag-big}}
+\end{figure}
+
+%The number of transactions solving the proof-of-work may be modeled by a Birth and Death process, since if we have $k$ transactions solving the proof-of-work only two things may happen: either a new transaction will be created or one the transactions will finish solving the proof-of-work.
+Let the time between new transactions follows an exponential distribution with $\mu$ parameter, then the probability of a new transaction emerges before any transaction finishes its proof-of-work is $q_k = \mathbf{P}(T_{\text{new tx}} = \min\{T_{\text{new tx}}, T_1, T_2, \dots, T_k\}) = \frac{\mu}{\mu + \sum_{i=1}^k \lambda_i}$.
+Hence, the probability of any transaction finishes its proof-of-work before a new transaction emerges is $p_k = 1 - q_k = \frac{\sum_{i=1}^k \lambda_i}{\mu + \sum_{i=1}^k \lambda_i}$.
+
+\begin{theorem}
+\label{theorem-new-tx-not-tip}
+When a new transaction confirms an already confirmed transaction, the depth of all transactions remains the same, i.e., the depth of the transactions are only changed when a new transaction confirm an unconfirmed transaction.
+\end{theorem}
+
+\begin{theorem}
+The height of a transaction is equal to the maximum height of its parents plus one.
+\end{theorem}
+
+\begin{theorem}
+If new transactions choose randomly the unconfirmed transactions to be confirmed, then no unconfirmed transaction will be left behind, i.e., all transactions will be confirmed in due time.
+\end{theorem}
+
+\chapter{Conclusion}
+
+Bitcoin's underlying technology blockchain has been called by many as a major invention, even comparable to the invention of the internet.
+But it is unlikely that Bitcoin and blockchain have achieved the final or most optimal design for a secure and scalable electronic transaction system.
+In this work, I proposed and analysed a new architecture named Hathor, which seems a scalable alternative to Bitcoin.
+
+Today, Bitcoin network can barely handle 8 transactions per second without increasing the unconfirmed transaction list to hundreds of thousands --- several transactions take days to be confirmed.
+In order to increase Bitcoin's capacity, its community has first proposed and implemented segregated witness, which was not enough.
+Finally, they proposed the lightning network, which is in development and should be available in the next months.
+
+- What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
+
+%\chapter{Proposal: Questions to be explored}
+
+%Given these preliminaries, these are some questions with which we will be concerned.
 
@@
 %\end{enumerate}
-
-\chapter{Methodology}
-
-The methodology we have used is computer simulation.
-Through the simulation of many scenarios of Hathor, we will understand how the network behaves in complex scenarios, including when the load suddenly increases, and when the network is under attack.
-
-The simulator has been developed using an event-based design which is capable of running hours of simulation in just a few minutes.
-It creates agents who decide to make a transaction, then they select which transactions will be confirmed, then they spend some time working in the proof-of-work, and, finally, they propagate the transaction to the network.
-The other agents receive the transaction and may accept or deny it.
-The agents may use different parameters among themselves.
-
-When a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work.
-When it finishes solving the proof-of-work, it is propagated and becomes a tip.
-So, two new transactions may choose the same tips to confirm.
-If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
-
-When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
-The topological sort is used to calculate the longest path from all other transactions to the new transaction, and this longest path is used to update the depth of the whole DAG.
-If the new transaction confirms only transactions which has already been confirmed, the depth update is skipped thanks to Theorem \ref{theorem-new-tx-not-tip}.
-
-The amount of time spent to solve the proof-of-work is a sample from an exponential distribution.
-
-The time between two new transactions is sampled from an exponential distribution with parameter $\lambda$.
-The value of $\lambda$ is changed over time, causing changes in the load of the network.
-We consider 1 tx/s to be a low load, and 15 tx/s to be a high load.
-Even though we have already tested up to 100 tx/s.
-
-The simulator will output different reports: (i) snapshots of the DAG at interesting moments, such as Fig.
-\ref{fig-tangle-example}, \ref{fig-tangle-swarm}, and \ref{fig-tangle-conflict}; (ii) histogram of confirmation time, such as Fig.
-\ref{fig-tangle-hist} and \ref{fig-tangle-hist-2}; and many others.
-
-\chapter{Analysis of Hathor}
-
-In order to a further understanding of the DAG Model, we will analyze the mathematical model behind it.
-
-The number of transactions solving the proof-of-work may be modeled by a Birth and Death process, since if we have $k$ transactions solving the proof-of-work only two things may happen: either a new transaction will be created or one the transactions will finish solving the proof-of-work.
-Let the time between new transactions follows an exponential distribution with $\mu$ parameter, then the probability of a new transaction emerges before any transaction finishes its proof-of-work is $q_k = \mathbf{P}(T_{\text{new tx}} = \min\{T_{\text{new tx}}, T_1, T_2, \dots, T_k\}) = \frac{\mu}{\mu + \sum_{i=1}^k \lambda_i}$.
-Hence, the probability of any transaction finishes its proof-of-work before a new transaction emerges is $p_k = 1 - q_k = \frac{\sum_{i=1}^k \lambda_i}{\mu + \sum_{i=1}^k \lambda_i}$.
-
-\begin{theorem}
-\label{theorem-new-tx-not-tip}
-When a new transaction confirms an already confirmed transaction, the depth of all transactions remains the same, i.e., the depth of the transactions are only changed when a new transaction confirm an unconfirmed transaction.
-\end{theorem}
-
-\begin{theorem}
-The height of a transaction is equal to the maximum height of its parents plus one.
-\end{theorem}
-
-\begin{theorem}
-If new transactions choose randomly the unconfirmed transactions to be confirmed, then no unconfirmed transaction will be left behind, i.e., all transactions will be confirmed in due time.
-\end{theorem}
-
-\chapter{Conclusion}
-
-Bitcoin's underlying technology blockchain has been called by many as a major invention comparable to the invention of the internet.
-But it is unlikely that Bitcoin and blockchain have achieved the final or most optimal design for a secure and scalable electronic transaction system.
-In this work, we will investigate whether DAG model is suitable to be a real alternative to blockchain.
-
-Preliminary results obtained from our simulator have shown that, using a random tip selection strategy, the network seems to support up to 50 transactions per second, with no tip abandoned (Fig.
-\ref{fig-tangle-swarm}) or taking too long to have at least one confirmation (Fig.
-\ref{fig-tangle-hist-2}).
-Today, Bitcoin network can barely handle 8 transactions per second without increasing the unconfirmed transaction list to hundreds of thousands --- several transactions take days to be confirmed.
-
-We have not yet tested the security of the network against double spend attacks.
-It will be one of the next steps and we expect to block a nuclear submarine attack including a rule for invalidate new conflicting transactions which confirm transactions with depth greater than a given threshold.
-We are unsure whether this rule would invalidate legitimate transactions on a high load scenario.
-
-We also plan to improve the analysis of the confirmation time.
-So far, we have the histograms of how long it takes to a transaction to be confirmed for the first time (Fig.
-\ref{fig-tangle-hist} and \ref{fig-tangle-hist-2}).
-We will also generate the histograms of how long it takes to a transaction to have an accumulated weight of at least a given number.
-It is an important histogram because it helps merchants and users to decide how long they should wait to consider the transaction almost irreversible.
-
-We will also work on the mathematical modeling of DAG networks, and the equations will help us to understand the limits of these networks.

### 2018-04-12 Chapters/Chapter01.tex
Hathor: splitting Hathor's architecture and analysis of Bitcoin in separate files.

@@
 \chapter{Analysis of Bitcoin}
-
-The primary objective of this chapter is to increase our understanding of the Bitcoin through mathematical tools.
-
-\section{Hash function}
-
-Hash functions has been widely studied in computer science. In short, a hash function $h: \{0, 1\}^\infty \rightarrow \{0, 1\}^{n}$ has the following properties:
-
-\begin{enumerate}
-	\item $x = y \Rightarrow h(x) = h(y)$
-	\item $h(x) \sim \mathcal{U}(0, 2^{n}-1)$, where $\mathcal{U}$ is the uniform distribution, i.e., $\forall a \in [0, 2^{n}-1], \mathbf{P}(h(x)=a) = \frac{1}{2^{n}}$
-\end{enumerate}
-
-In other words, when two inputs are the same, they have the same output.
-But, when the inputs are different, their outputs are uniformly distributed.
-Clearly, the hash functions are surjective but not injective.
-They are not injective because the image of $h$ has only $2^n$ elements and the domain has infinite elements.
-When $x \ne y$ and $h(x) = h(y)$, we say that $x$ and $y$ are a collision.
-A hash function is considered to be safe when it is unknown how to quickly find a collision of a given hash, i.e., one has to check all possible values until the correct one is found (known as the brute-force attack).
-
-Bitcoin uses two hash functions: HASH-160 and HASH-256.
-The first has $n=160$ and consists of the composition of \textit{SHA-256} and \textit{RIPEMD-160}.
-The latter has $n=256$ and applies \textit{SHA-256} twice.
-The first is used in transactions' scripts and the latter in the mining algorithm.
-For both hash functions, it is infeasible to run a brute-force attack because it would demand, on average, either $2^{160}$ or $2^{256}$ trials, and those would take a tremendous amount of time even for the fastest known processors.
-
-For further information about hash functions, see \citet{gilbert2003security, dobbertin1996ripemd}.
-
-\section{Mining one block}
-
-Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin \textit{HASH-256} function.
-The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given threshold.
-The smaller the $A$, the harder to find a new block.
-In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.
-
-Hence, in order to find a new block, one must try different inputs ($x_1, x_2, \dots, x_k$) until they find a solution, i.e., all attempts will fail ($h(x_i) \geq A$ for $i < k$) but the last ($h(x_k) < A$).
-The probability of finding a solution exactly in the $k^{th}$ attempt follows a geometric distribution.
-Let $X$ be the number of attempts until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \frac{A}{2^{256}}$.
-Also, we have $\mathbf{P}(X \leq k) = 1 - (1-p)^k$.
-The average number of attempts is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = \frac{1-p}{p^2}$.
-
-In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average.
-Suppose that the Bitcoin's network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network.
-The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block.
-So, the rule of finding a new block every 10 minutes ($\eta = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \eta = 600$.
-So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \eta = 600 \Rightarrow p = \frac{1}{\eta H}$.
-Finally, $\mathbf{E}(X) = \eta H$, $\mathbf{E}(T) = \eta$, $\mathbf{V}(X) = (\eta H)^2 - \eta H$, and $\mathbf{V}(T) = \eta^2 - \eta/H$.
-
-The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\eta H} \right)^{tH}$.
-But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.
-
-\begin{theorem}
-	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{\eta}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$.
-\end{theorem}
-\begin{proof}
-\begin{align*}
-	\mathbf{P}(T \leq t) &= 1 - (1-p)^{tH} \\
-		     &= 1 - \left( 1 - \frac{1}{\eta H} \right)^{tH} \\
-	\text{Replacing $u = \eta H$,} \\
-	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{\eta}} \\
-				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{\eta}} \\
-				&= 1 - \left( 1/e \right)^{\frac{t}{\eta}} \\
-				&= 1 - e^{-\frac{t}{\eta}}
-\end{align*}
-\end{proof}
-
-Now, we would like to understand from which value of $H$ it is reasonable to assume that $T$ follows an exponential distribution.
-
-\begin{theorem}
-	$x > M \Rightarrow |(1+1/x)^x - e| < e/M$.
-\end{theorem}
-\begin{proof}
-	Let's use the classical inequality $\frac{x}{1+x} < \log(1+x) < x$ for $x > -1$.
-So, $\frac{1/x}{1+1/x} < \log(1+x) < 1/x$.
-Simplifying, $\frac{1/x}{1+1/x} = 1/(1+x)$.
-Thus, $1/(1+x) < \log(1+1/x) < 1/x \Rightarrow x/(1+x) < x \log(1 + 1/x) < 1$.
-
-	As $\log(1 + \frac{1}{M}) > 0$ and $1 < 1 + \log(1 + \frac{1}{M})$.
-
-	$x > M \Rightarrow 1/x < 1/M \Rightarrow 1+1/x < 1+1/M \Rightarrow 1/(1+1/x) > 1/(1+1/M) \Rightarrow x/(1+x) > M/(1+M)$.
-
-	Again, $\log(1+x) < x \Rightarrow \log(1-1/M) < -1/M \Rightarrow 1 + \log(1-1/M) < (M-1)/M < M/(1+M)$, since $(x-1)/x < x/(x+1)$.
-
-	Hence, $1 + \log(1-1/M) < M/(1+M) < x/(1+x) < x\log(1+1/x)$, and $x\log(1+1/x) < 1 < 1 + \log(1 + \frac{1}{M})$.
-
-	Finally,
-
-	\begin{align*}
-		1 + \log(1-1/M) &< x\log(1+1/x) < 1 + \log(1 + \frac{1}{M}) \\
-		e^{1 + \log(1-1/M)} &< e^{x\log(1+1/x)} < e^{1 + \log(1 + \frac{1}{M})} \\
-		e \cdot e^{\log(1-1/M)} &< e^{\log((1+1/x)^x)} < e \cdot e^{\log(1 + \frac{1}{M})} \\
-		e(1-1/M) &< (1+1/x)^x < e(1 + \frac{1}{M}) \\
-		e-e/M &< (1+1/x)^x < e + e/M \\
-		-e/M &< (1+1/x)^x - e < e/M \\
-	\end{align*}
-	Therefore, $\left| \left( 1+1/x \right)^x - e \right| < e/M$.
-\end{proof}
-
-We may consider $H$ big enough to say that $T$ follows an exponential distribution when $e/H < \epsilon$, where $\epsilon$ is the maximum approximation error.
-When $\epsilon = 10^{-6} \Rightarrow H > e \cdot 10^6$.
-So, when $H > 2.6 \text{Mh/s}$, our approximation is good enough.
-
-The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, where $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$.
-These conditions give the following equations: $1 - e^{-t_0/\eta} = \alpha/2$, and $e^{-t_1/\eta} = \alpha/2$.
-Solving these equations, we have $t_0 = -\eta \ln{(1 - \alpha/2)}$, and $t_1 = -\eta \ln{(\alpha/2)}$.
-
-For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ (or $[0.51, 30.76]$ in minutes).
-Thus, 90\% of the time the intervals between blocks are between 30 seconds and 30 minutes, with average of 10 minutes.
-
-The fact that the time between blocks follows an exponential distribution with $\lambda = 1/\eta = pH$ may be used to estimate the total network's hash rate (or a miner's hash rate).
-For further information, see \cite{ozisikestimation}.
-
-\section{Mining several blocks}
-
-Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on.
-Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$ which is the total time to find the next $n$ blocks.
-As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{\eta}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{\eta})$.
-Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.
-
-Many exchanges require at least six confirmations in order to accept a deposit in Bitcoin.
-So, for $n=6$, $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$, i.e., only 55\% of the deposits will be accepted in one hour.
-The symmetrical confidence interval with $\alpha=10\%$ is $[27, 105]$ in minutes.
-Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted --- assuming that your transaction will be confirmed in the very next block.
-The pdf of $Y_6$ is shown in Figure \ref{fig-bitcoin-time-6-blocks}, in which the 10\% symmetrical confidence interval is shown in the white area.
-The average total time of six confirmations is $\mathbf{E}(Y_6) = 6 \cdot 600 = 3600 = 60 \text{ minutes}$.
-
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images01/time-6-blocks.pdf}
-\caption{Probability density function of $Y_6$, i.e., probability of finding 6 blocks after time $t$. The shaded areas shows the lower 5\% and upper 5\% of the pdf.\label{fig-bitcoin-time-6-blocks}}
-\end{figure}
-
-\section{Mining for a miner}
-
-Let's analyze the probability of finding a new block for a miner who has $\alpha$ percent of the network's total hash rate.
-Let $T_\alpha = \frac{X}{\alpha H}$ be the time required for the miner to find a new block.
-As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{\eta}$.
-Hence, we confirm the intuition that the miner with $\alpha$ percent of the network's total hash power will find $\alpha$ percent of the blocks.
-
-\begin{theorem}
-	When the miner with $\alpha$ percent of the network's total hash rate is part of the mining network, $\mathbf{P}(\text{next block is from $T_\alpha$}) = \alpha.$
-\end{theorem}
-\begin{proof}
-\begin{align*}
-	\mathbf{P}(\text{next block is from $T_\alpha$}) &= P \left( T_\alpha = \min\{T_\alpha, T_{1-\alpha}\} \right) \\
-		&= \frac{\lambda_\alpha}{\lambda_\alpha + \lambda_{1-\alpha}} \\
-		&= \frac{\alpha/\eta}{\alpha/\eta + (1-\alpha)/\eta} \\
-		&= \frac{\alpha}{\alpha + 1 - \alpha} \\
-		&= \alpha.
-\end{align*}
-\end{proof}
-
-\begin{theorem}
-	When one miner with $\alpha$ percent of the network's total hash rate multiplies their hash rate by $m$, the probability of this miner find the next block is multiplied by $\frac{m}{m \alpha + 1 - \alpha}$.
-	\label{thm-miner-multiply}
-\end{theorem}
-\begin{proof}
-	When miners increase their hash rate, they also increase the network's total hash rate.
-Let $H$ be the network's hash rate before the increase.
-Thus, the network's total hash rate after the increase is $H + (m-1) \alpha H = (1 - \alpha + m \alpha) H$.
-So,
-\begin{align*}
-	\mathbf{P}(\text{next block is from $T_{m \alpha}$}) &= P \left( T_{m \alpha} = \min\{T_{m \alpha}, T_{1-\alpha}\} \right) \\
-		&= \frac{\lambda_{m \alpha}}{\lambda_{m \alpha} + \lambda_{1-\alpha}} \\
-		&= \frac{m \alpha/\eta}{m \alpha/\eta + (1-\alpha)/\eta} \\
-		&= \frac{m \alpha}{m \alpha + 1-\alpha} \\
-		&= \alpha \left( \frac{m}{m \alpha + 1 - \alpha} \right).
-\end{align*}
-\end{proof}
-
-\begin{cor}
-	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximately multiplies their probability of finding the next block by $m$.
-\end{cor}
-\begin{proof}
-	$$\lim_{\alpha \rightarrow 0} \mathbf{P}(\text{next block is from $T_{m \alpha}$}) = \lim_{\alpha \rightarrow 0} \frac{m}{m \alpha + 1 - \alpha} = m.$$
-\end{proof}
-
-That way, it is not exactly correct to say that when one doubles their hash rate, their probability will double as well. It is only true for small miners.
-
-\section{Orphan blocks}
-% Probabilidade de orfão
-An orphan block would be created if a new block is found during the propagation time of a new block.
-Let $\alpha$ be the percentage of the total hash rate of the node which is outdated, and $\Delta t$ the propagation time in seconds.
-Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{\eta}}$.
-
-Bitcoin peer-to-peer network is a gossip network, where miners are semi-randomly connected to each other, and each miner sends all information it receives to all its peers.
-According to \citet{decker2013information}, the average time for a new block propagate over the network is 12.6 seconds, while the 95\% percentile is 40 seconds, which indicates a long-tail distribution.
-\citet{bitcoinstats} has measured the propagation time between 2013 and 2017.
-During 2017, the worst daily 90\% percentile was 21 seconds.
-Notice that both results may not be contradictory because Bitcoin network is continuously evolving.
-
-For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$, which is almost 0.5\%.
-I would say that a node with 10\% of the total hash rate would be well connected and it would take less time to receive the update, so, the probability would be even smaller than 0.5\%.
-
-Another important factor is that, as Bitcoin is open-source, miners are free to change the gossip algorithm, which leads to the network incentives.
-See \citet{babaioff2012bitcoin} for an analysis of the incentives to miners forward new blocks and transactions in the network.
-
-For further information about gossip algorithms, see \citet{shah2009gossip}.
-
-\section{Analysis of network's hash rate change}
-% Análise: Mudança no network's hash rate antes de atualizar o $A$.
-
-The difficulty, given by the number $A$, is adjusted every 2016 blocks.
-As, $\mathbf{P}(13 \text{ days} < Y_{2016} < 15 \text{ days}) = \mathbf{P}(13 \cdot 24 \cdot 3600 < Y_{2016} < 14 \cdot 24 \cdot 3600) = 0.9986$, it is expected that the total time to find 2016 blocks will be between 13 and 15 days, assuming that the network's hash rate remains constant.
-If it takes less than the expected time, it means that the network's total hash rate has increased.
-While if it takes more than the expected time, it means that the network's total hash rate has decreased.
-So, let's analyze what happens when the network's hash rate changes significantly.
-
-%Let's assume that the network's total hash rate will change only once, from $H$ to $\alpha H$ --- so, if $\alpha = 1$, there would be no change, if $\alpha > 1$, there would be an increase, and if $\alpha < 1$, there would be a decrease.
-As the time required to find a new block would be $T_\alpha = \frac{X}{\alpha H}$, we got the same random variable as we have gotten when analyzing a miner with $\alpha$ percent of the network's total hash rate.
-But, the analysis will be different because here $\alpha$ may be greater than one.
-
-Let $H \cdot u(t)$ be the network's total hash rate over time.
-So, the number of hashes calculated in $t$ seconds is $H \int_0^t u(t) dt$.
-Hence, $\mathbf{P}(T \leq t) = \mathbf{P}(X \leq H \int_0^t u(t)dt)$.
-When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{\eta} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{\eta} \cdot e^{-\frac{1}{\eta} \int_0^t u(t)dt}$.
-
-Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
-So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\eta}$.
-Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\eta})$.
-Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$.
-Hence, on one hand, when the network's hash rate increases ($\alpha > 1$), the 2016 blocks will be found earlier.
-On the other hand, when the network's hash rate decreases ($\alpha < 1$), the 2016 blocks will be found later.
-
-For example, if the network's total hash rate suddenly doubles ($\alpha = 2$), then $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved.
-On the other side, if the network's total hash rate suddenly halves ($\alpha = 0.5$), then $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled.
-It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, at which point the adjustment will occur and everything will be back to the normal 10 minutes between blocks.
-
-% TODO Step-wise analysis
-
-\subsection{Hash rate smoothly changing}
-
-Let $u(t) = \frac{1+abx}{1+bx}$.
-It is an useful function because $u(0) = 1$ and $\lim_{t \rightarrow \infty} u(t) = a$.
-The bigger the $b$, the faster $u(t) \rightarrow a$.
-For example, if $a=2$, it means $H$ would be smoothly doubling.
-If $a=0.5$, it means $H$ would be smoothly halving.
-
-It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a \Rightarrow \int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
-$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
-$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
-
-Assuming that $n = \frac{\lambda(a-1)}{b}$ is integer, we have:
-$$F_T(t) = 1 - (1+bt)^n e^{-\lambda at}$$
-
-Thus,
-\begin{align*}
-\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
-	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
-	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
-	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
-	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
-\end{align*}
-
-Hence, as $\mathcal{L}\{f_T(t)\} = s \mathcal{L}\{F_T(t)\}$,
-
-$$\mathcal{L}\{f_T(t)\} = 1 - \sum_{k=0}^n \binom{n}{k} \frac{s b^k k!}{(s+\lambda a)^{k+1}}$$
-
-Then,
-
-\begin{align*}
-\frac{d}{ds} \mathcal{L}\{f_T(t)\}
-	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{d}{ds} \frac{s}{(s+\lambda a)^{k+1}} \\
-	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \left[ \frac{1}{(s+a\lambda)^{k+1}} - \frac{s(k+1)}{(s+a \lambda)^{k+1}} \right] \\
-\frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0}
-	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{1}{(\lambda a)^{k+1}} \\
-	&= - \frac{1}{a\lambda} \sum_{k=0}^n \binom{n}{k} k! \left( \frac{b}{\lambda a} \right)^k \\
-	&= - \frac{1}{a\lambda} \sum_{k=0}^n \frac{n!}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \\
-	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \right] \\
-	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{n-k} \right] \tag{$k \rightarrow n-k$} \\
-	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{-k} \right] \\
-	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right]
-\end{align*}
-
-Finally, as $\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0}$,
-$$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n!
-\left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right] \text{, where $n = \frac{\lambda(a-1)}{b}$}$$
-
-Let's check this equation for already known scenarios.
-When $a=1$, then $n=0$ and $\mathbf{E}[T] = 1/\lambda$.
-When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied.
-In fact, $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.
-
-\begin{theorem}
-	$$a > 1 \text{ and } x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
-\end{theorem}
-\begin{proof}
-$x > M \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM}$.
-As $1-a<0$, $\frac{1-a}{1+bx} > \frac{1-a}{1+bM}$.
-Thus, $\frac{1-a}{1+bM} < \frac{1-a}{1+bx} + a - a = \frac{1+abx}{1+bx} - a < 0 < \frac{a-1}{1+bM}$.
-Hence, $-\frac{a-1}{1+bM} < \frac{1+abx}{1+bx} - a < \frac{a-1}{1+bM}$.
-\end{proof}
-
-For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$.
-So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected.
-Finally, $\mathbf{E}[T] = 557.65$.
-In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes.
-If the hash rate had suddenly doubled, the average time between blocks would be 5 minutes.
-
-%$$\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0} = \frac{b^n n! 600^{n+1}}{a^{n+1}} e^{\frac{a}{600b}} \text{, where $n = \frac{a-1}{600b}$}$$
-
-%$a > 1, x > M \Rightarrow 1 + bx > 1 + bM \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM} \Rightarrow \frac{1-a}{1+bx} + a > \frac{1-a}{1+bM} + a \Rightarrow \frac{1+abx}{1+bx} - a > \frac{1-a}{1+bM}$.
-
-\subsection{Piecewise linear model of hash rate change}
-
-Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(a, b, t) = a^2t + b$.
-Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{bt + a^2t^2/2}{\eta}}$.
-
-It is well known that $\mathbf{E}(T) = \int_{0}^{\infty} 1 - \mathbf{P}(T \leq t) dt$.
-Thus, replacing $y = \frac{a^2 t + b}{a \sqrt{2 \eta}}$, and using the fact that $\int_0^\infty e^{-x^2} dx = \frac{\sqrt{\pi}}{2} \erf(x)$, we have:
-
-\begin{align}
-\mathbf{E}(T)|_{t_1}^{t_2} &= \int_{t_1}^{t_2} \exp \left( - \frac{bt + a^2 t^2/2}{\eta} \right) dt \nonumber \\
-	&= \frac{\sqrt{2 \eta}}{a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \int_{y_1}^{y_2}  \exp(-y^2) dy \nonumber \\
-	&= \frac{\sqrt{2 \eta}}{a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \frac{\sqrt{\pi}}{2} [\erf(y_1) - \erf(y_2)] \nonumber \\
-	&= \frac{\sqrt{2 \pi \eta}}{2a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \left[\erf(y_2) - \erf(y_1) \right]
-\end{align}
-
-Where $y_1 = \frac{a^2 t_1 + b}{a \sqrt{2 \eta}}$ and $y_2 = \frac{a^2 t_2 + b}{a \sqrt{2 \eta}}$.
-
-Thus, $\mathbf{E}(T) = \mathbf{E}(T)|_0^\infty$.
-When $t_1 = 0 \Rightarrow y_1 = \frac{b^2}{2 \sqrt{2 \eta}}$ and $t_2 \rightarrow \infty \Rightarrow y_2 \rightarrow \infty \Rightarrow \erf(y_2) = 1$, then:
-
-$$
-\mathbf{E}(T) = \frac{\sqrt{2 \pi \eta}}{2a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \left[1 - \erf\left(\frac{1}{a \sqrt{2 \eta}}\right)\right]
-$$
-
-The function $\mathbf{E}(T)|_{t_1}^{t_2}$ may be used to a piecewise linear analysis of any hash rate change.
-Let's analyze the hash doubling in one week.
-Then, $u(1 \text{week}) = u(604800) = 2$, thus $a^2 = \frac{1}{604800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$.
-Let's sample the interval $[0, 1 \text{week}]$ every hour, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.
-
-For each point $t_i$, let $g_i(t) = [H(t-t_i) - H(t-604800)] u(t) + 2 H(t-604800)$, then $\mathbf{E}(T) = \mathbf{E}(T)|_{t_i}^{t_{i+1}} + 1 - e^{\frac{t}{\eta}}$. The result is presented in Figure .
-
-%In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.
-%
-%Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.
-%
-%\begin{align*}
-%	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
-%		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\tau}} \right) dx \\
-%		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\tau}} dx \\
-%		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx \\
-%\end{align*}
-%
-%Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
-%Let $y = bx+c \Rightarrow dy = bdx$.
-%
-%\begin{align*}
-%	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
-%		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
-%		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
-%		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
-%\end{align*}
-%
-%$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
-%
-%Hence,
-%$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
-%
-%As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,
-%
-%\begin{align*}
-%\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
-%\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
-%\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
-%\end{align*}
-%
-%$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.
-%
-%\textcolor{red}{I do not know how to interpret this.}
-
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
-\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
-\end{figure}
-
-\section{Attack in the Bitcoin network}
-% Ataque submarino
-% Probabilidade de "correr por fora"
-
-\citet{karame2012two}
-
-There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.
-
-In the double spending attack, the attacker's send some funds to the victim, let's say a merchant.
-They wait for $k$ confirmations of the transaction, and the victim delivers the good or the service to the attacker.
-Then, the attacker mine enough blocks with a conflicting transaction, double spending the funds which was sent to the victim.
-If the attacker is successful, the original transaction will be \textit{erased} and the victim will be left with no funds at all.
-In order to be successful, the attacker must propagate more blocks than the network in the same period, propagating a chain longer than the main chain.
-Hence, we would like to understand what the odds are that the attacker will be successful.
-This attack was originally discussed by \citet{nakamoto2008bitcoin}.
-
-In order to maximize their odds, the attacker must start to mine the new blocks as soon as they send the funds to the victim.
-In this moment, it starts to mine in the head of the blockchain, just like the rest of the network.
-So, in the beginning, the attacker and the network are in exactly the same point.
-
-Let $\beta H$ be the hash rate of the attackers, and $\gamma H$ be the network's hash rate without the attackers.
-Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{\eta}$ and $\lambda_{\text{network}} = \frac{\gamma}{\eta}$, respectively.
-
-As \cite{nakamoto2008bitcoin} has done, we will also model the attack using the Gambler's Ruin.
-In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$.
-The rounds are independent.
-The gambler starts with \$$k$ plays continuously until he either accumulates a target amount of \$$m$, or loses all his money.
-Let $\rho = \frac{1-p}{p}$, then the probability of losing his fortune is:
-
-$$
-\mathbf{P}(\text{losing his fortune}) =
-\begin{cases}
-	\frac{\rho^k - \rho^m}{1-\rho^m} \text{, if $\rho \ne 1$,} \\
-	\frac{m-k}{m} \text{, if $\rho = 1$.}
-\end{cases}
-$$
-
-When $m \rightarrow +\infty$,
-
-$$
-\mathbf{P}(\text{losing his fortune}) =
-\begin{cases}
-	\rho^k \text{, if $\rho < 1$,} \\
-	1 \text{, if $\rho \geq 1$.}
-\end{cases}
-$$
-
-% TODO Expected time to absorption.
-
-The gambler winning \$1 is the same as the network finding a new block, the gambler losing \$1 is the same as the attacker finding a new block.
-The initial \$$k$ is the same as the number of blocks the attacker is behind the network.
-Thus, the gambler loses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.
-
-In our case, $p = \frac{\lambda_{\text{network}}}{\lambda_{\text{network}} + \lambda_{\text{attacker}}} = \frac{\gamma}{\beta + \gamma}$, thus $\rho = \frac{\beta}{\gamma}$.
-Hence, $\rho < 1 \Leftrightarrow \beta < \gamma$.
-
-Suppose that the attacker is mining with the network.
-Suddently, he stops mining with the network and starts attacking, i.e., starts to mine in another chain.
-In this scenario, since the attacker's hash rate is not mining with the network anymore, $\gamma = 1 - \beta$.
-Thus, $\beta < \gamma \Rightarrow \beta < 0.5 \Leftrightarrow \rho < 1$.
-Here comes the conclusion that, if the attacker has 50\% or more of the network's hash rate, then his attack will be certainly successful.
-We got exactly the same equations and conclusions as \cite{nakamoto2008bitcoin}.
-
-But this scenario seems not to be the optimal attack, because the attacker has waited $k$ confirmations before starting the attack.
-A better approach would be to start attacking just after propagating the transaction.
-In this case, our previous model is not good, because even if the attacker have found more blocks than the network, he cannot propagate those blocks before the network has found $k$ confirmations.
-So, we have to model the probabilities before the network has found the $k$ block.
-Then, if the attacker has more blocks than the network, he has successfully attacked.
-Otherwise, we return to the previous model, in which the attacker must still find more blocks.
-
-\begin{theorem}
-	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
-\end{theorem}
-\begin{proof}
-The attacker must find exactly $s$ blocks while the network must find exactly $k$ blocks.
-It is as they would be walking the grid from the point $(0, 0)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}.
-When the attacker finds a block, it would be a movement to the right.
-When the network finds a block, it would be an upward movement.
-No matter the order which the blocks are found, all the paths occur with probability $(1-p)^s p^k$.
-
-The walking ends when $(\cdot, k$) is reached, i.e., when the network finds $k$ blocks, regardless of how many blocks the attacker has found -- i.e., it is not allowed to walk above the line $(\cdot, k)$.
-Thus, the number of paths between $(0, 0)$ and $(s, k)$ moving only upward or to the right, without going into the line $(\cdot, k)$ is exactly the number of paths between $(0, 0)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s$ movements to the right ($r$) and $k-1$ upward movements ($u$).
-This number of permutations is $\frac{(k-1+s)!}{s!(k-1)!} = \binom{(k-1)+(s)}{s}$ because there are $s$ repetitions of the element $r$ and $k-1$ repetitions of the element $u$.
-
-Finally, the probability is $\binom{k+s-1}{s} (1-p)^{s} p^{k}$.
-
-\begin{figure}[ht]
-  \centering
-  \begin{tikzpicture}
-    \coordinate (Origin)   at (0,0);
-    \coordinate (XAxisMin) at (-1,0);
-    \coordinate (XAxisMax) at (5,0);
-    \coordinate (YAxisMin) at (0,-1);
-    \coordinate (YAxisMax) at (0,5);
-    %\draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
-    %\draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis
-
-    \clip (-0.5,-0.5) rectangle (10.5cm,7cm); % Clips the picture...
-    %\pgftransformcm{1}{0}{0}{1}{\pgfpoint{0cm}{0cm}}
-          % This is actually the transformation matrix entries that
-          % gives the slanted unit vectors. You might check it on
-           % MATLAB etc. . I got it by guessing.
-    \draw[style=help lines,dashed] (0,0) grid[step=1cm] (14,6);
-          % Draws a grid in the new coordinates.
-          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
-              % Puts the shaded rectangle
-    \foreach \x in {0,1,...,10}{% Two indices running over each
-      \foreach \y in {0,1,...,6}{% node on the grid we have drawn
-        \node[draw,circle,inner sep=1pt,fill] at (\x,\y) {};
-            % Places a dot at those points
-      }
-    }
-
-    \draw [ultra thick,-latex,red] (0,0) -- (0,1) node [midway, left] {};
-    \draw [ultra thick,-latex,red] (0,1) -- (0,2) node [midway, left] {};
-    \draw [ultra thick,-latex,red] (0,2) -- (1,2) node [midway, above] {};
-    \draw [ultra thick,-latex,red] (1,2) -- (1,3) node [midway, right] {};
-    \draw [ultra thick,-latex,red] (1,3) -- (2,3) node [midway, right] {};
-    \draw [ultra thick,-latex,red] (2,3) -- (2,4) node [midway, right] {};
-    \draw [ultra thick,-latex,red] (2,4) -- (2,5) node [midway, right] {};
-    \draw [ultra thick,-latex,red] (2,5) -- (3,5) node [midway, right] {};
-	\draw [ultra thick,-latex,red] (3,5) -- (3,6) node [above] {(3, 6)};
-
-    \draw [ultra thick,-latex,blue] (0,0) -- (1,0) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (1,0) -- (2,0) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (2,0) -- (2,1) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (2,1) -- (2,2) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (2,2) -- (2,3) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (2,3) -- (3,3) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (3,3) -- (4,3) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (4,3) -- (5,3) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (5,3) -- (5,4) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (5,4) -- (6,4) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (6,4) -- (6,5) node [midway, left] {};
-    \draw [ultra thick,-latex,blue] (6,5) -- (7,5) node [midway, left] {};
-	\draw [ultra thick,-latex,blue] (7,5) -- (7,6) node [above] {(7,6)};
-  \end{tikzpicture}
-  \caption{Both the attacker and the network are mining.
-Each step up is a new block found by the network with probability $p$.
-Each step right is a new block found by the attacker with probability $1-p$.
-It ends when the network finds $k$ blocks --- in this example, $k=6$.
-The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$.
-Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network.
-In the red path, the attacker still have to catch up 3 blocks to have a successful attack, which happens with probability $\rho^3$, if $p < 0.5$.}
-  \label{figure:bitcoin-attack-paths}
-\end{figure}
-\end{proof}
-
-Assuming that the attacker starts mining just after publishing the victim's transaction, the probability of the attacker will have found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.
-
-\begin{theorem}
-	$$\mathbf{P}(S \geq k) = 1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} (1-p)^s p^k.$$
-\end{theorem}
-\begin{proof}
-	Let's use the following identity:
-	$$\frac{1}{(1-z)^{a+1}} = \sum_{i=0}^{\infty} \binom{i+a}{i} z^i \text{, for $|z|<1$}$$
-
-	Thus, replacing $z=1-p$, $i=s$, and $a=k-1$, we have:
-
-	$$\frac{1}{p^{k}} = \sum_{s=0}^{\infty} \binom{s+k-1}{s} (1-p)^s$$
-	$$1 = \sum_{s=0}^{\infty} \binom{s+k-1}{s} (1-p)^s p^k.$$
-
-	Now, just split $\sum_{s=0}^{\infty} = \sum_{s=0}^{k-1} + \sum_{s=k}^{\infty}$ and it is done.
-\end{proof}
-
-Using this last theorem, we moved from an infinity sum to a finity sum.
-
-\begin{theorem}
-	Let $p = \frac{\gamma}{\beta + \gamma}$.
-$$
-\mathbf{P}(\text{successful attack}) =
-\begin{cases}
-%1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} (1-p)^s p^k \left(1 - \left( \frac{1-p}{p} \right)^{k-s} \right)
-	1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} \left( (1-p)^s p^k - (1-p)^k p^s \right) \text{, $p \geq 0.5$} \\
-	1 \text{, $p < 0.5$}.\\
-\end{cases}
-$$
-\end{theorem}
-\begin{proof}
-$$
-\mathbf{P}(\text{successful attack}) = \mathbf{P}(S \geq k) + \sum_{i=0}^{k-1} \mathbf{P}(s=i) \rho^{k-i}
-$$
-\end{proof}
-
-For $k=6$, $p=0.9$, $\mathbf{P}(\text{successful attack}) = 0.0005914121600000266$.
-
-For $k=6$, $p=0.7$, $\mathbf{P}(\text{successful attack}) = 0.15644958192000014$.
-
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images01/fig-bitcoin-attack.png}
-\caption{Probability of a successful attack according to the network's hash rate of the attacker ($\beta$).\label{fig-bitcoin-attack}}
-\end{figure}
-
-%The number of blocks found in a given time interval $\Delta t$ is given by the Poisson distribution.
-So, the number of blocks found by the attackers $N_{\text{attacker}}$ follows a Poisson distribution with parameter $\lambda_{\text{attacker}} \Delta t = \frac{\beta \Delta t}{600}$, and the number of blocks found by the network $N_{\text{network}}$ follows a Poission distribution with parameter $\lambda_{\text{network}} \Delta t = \frac{\gamma \Delta t}{600}$.
-We would like to calculate $\mathbf{P}(N_{\text{attackers}} - N_{\text{network}} \geq k)$.
-
-%$\mathbf{P}(Y_{n+k}^{\text{attacker}} < Y_{n}^{\text{network}})$
-
-\section{Confirmation time and network capacity}
-% Tempo na fila para confirmar transação
-
-Let's say that when a new transaction is propagated it is enqueued in the unconfirmed transaction queue.
-Then, when a new block is found, some of these transactions in the queue are confirmed.
-We are interested in some measures of the queue, like the expected time to confirm a transaction and the queue's length.
-
-Let's assume that all transactions have exactly the same size $S$ and pay exactly the same fee.
-If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block.
-Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$.
-The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$.
-Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$.
-The probabilities form a simple geometric series, and so they are exponentially decreasing.
-We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite.
-In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\eta$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
-Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
-When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.
-
-The average waiting time of a transaction to be confirmed, when $m < s$, is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.
-
-When $m \ll s$, $z_s \rightarrow 1 + 1/m$.
-So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$.
-In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.
-
-When $\lambda_{\text{TX}} \rightarrow s$, $z_s \rightarrow 1$. So, $\mathbf{E}(\pi) \rightarrow +\infty$.
-
-Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
-
-\section{Fork analysis}
-
-% TODO BIP's and signaling bit --- see BIP 9 and BIP 135.
-
-When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork.
-It is said to be a soft-fork when the rules are backward compatible, and a hard-fork when the rules are not backward compatible.
-In general, a hard-fork relaxes the constraints, while the latter hardens them.
-
-Suppose that the miners' are split in two groups, $G_1$ and $G_2$, with different rules.
-Let's say $H_1$ and $H_2$ are their hash rate, respectively.
-We have two different scenarios to analyze: (i) when neither of them accepts other's blocks; and (ii) when $G_2$ accepts $G_1$'s blocks, but not the other way around.
-
-Scenario (i) is easy to analyze, because the network would just split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.
-
-Scenario (ii) is more trick.
-As $G_2$ accepts $G_1$ blocks, their hashrate will matter.
-If $H_1 > H_2$, then $G_2$ will frequently skip their blockchain, because $G_1$'s blockchain will be longer most of the time.
-If $H_1 < H_2$, then $G_2$ may have its own blockchain after a while --- a true fork.
-But what would happen if $H_1$ keeps increasing and eventually gets larger than $H_2$?
-It if happens for sufficient time, the whole $G_2$'s blockchain may be discarded in order to move to $G_1$'s blockchain when $G_1$'s gets longer.
-
-\begin{theorem}
-	When $H_1 > H_2$, the
-\end{theorem}
-
-%What would happen if, after a while, the groups agree again. How would the blockchain merge?
+\input{./Chapters/hathor-bitcoin-math.tex}
 
@@
 \chapter{Hathor's architecture}
-
-This work introduces the Hathor's architecture which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
-Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG).
-For this, each transaction has its own proof-of-work which is solved by the issuer before propagating the transactions in the network.
-Like Bitcoin, miners find new ``blocks'' every 10 minutes in which they collect the fees and newly generated tokens.
-Each transaction has an ``accumulated weight'' which express the required effort to break the transaction, similar to Bitcoin's number of confirmations.
-
-This way, there will be two difficulty levels in the system: one for new transactions which are just moving tokens around, and another one for ``blocks'' which are generating new tokens and collecting fees.
-The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions.
-The second is adjusted every 2,016 blocks to keep the pace of block on every 2 minutes.
-
-There is a trade-off about the required proof-of-work of new transactions.
-The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.
-
-Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
-It is always possible since there is at least one confirmation path between any transaction and a tip.
-This is important for security reasons.
-
-The transactions are classified into three groups: (i) confirmed transactions, (ii) in-progress transactions, and (iii) unconfirmed transactions (tips).
-The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
-The unconfirmed transactions (tips) are the brand new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
-The in-progress transactions are in the middle.
-They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.
-
-A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two in-progress transactions or tips.
-There may be any number of outputs provided that they comply with the number of newly generated tokens.
-Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.
-
-In the low load scenario, there is a small number of new transactions coming into the network, which means they give a minor contribution to confirmations.
-In this case, the confirmation is held mostly by ``blocks''.
-On the other hand, in the high load scenarios, there is a large number of new transactions giving a major contribution to confirmations.
-In this case, the ``blocks'' strengthen the confirmation, but most of them will have already been confirmed before the new blocks are found.
-The higher the number of new transaction, the faster the transactions are confirmed.
-The blocks assure a ``maximum confirmation time''.
-
-The incentive scheme which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
-
-\section{Transaction confirmation}
-
-The concepts of weight and accumulated weight are the same as Iota's, but their equations are different.
-The weight of a transaction is calculated as $\log_2(k)$ where $k$ is the average number of hashes required to solve its proof-of-work again, while the accumulated weighted of a transaction A with weight $w_A$ is calculated as $\log_2(2^{w_A} + \sum_{A \leadsto P} 2^{w_P})$.
-The $\log$ function is used to reduce the magnitude of the numbers.
-Hence, the power-of-2 of the accumulated weight is a measure of the average number of hashes required to solve the proof-of-work of all transactions that are confirming A.
-
-In Bitcoin, it is well-known that one should wait at least ``six confirmations'' before accepting a transaction.
-This Bitcoin's criteria is based on some math presented in Satoshi's seminal work \citep{nakamoto2008bitcoin}.
-Adopting six confirmations is the same as demanding from attackers a minimal effort of six times the network's hash rate to successfully double spend those tokens.
-Let $H$ be the total hash rate of the network.
-Then, as $\mathbf{E}(Y_6) = 60 \text{ minutes}$, it will be necessary to calculate, on average, $\mathbf{E}(Y_6) \cdot H = 60 \cdot 60 \cdot H$ hashes to solve the proof-of-work of 6 blocks.
-
-Therefore, in order to have the same level of security as Bitcoin, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(6 \cdot 128 \cdot H) = 7 + \log_2(6) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.
-
-\section{Time between blocks}
-
-The hash function used for the proof-of-work (PoW) is the same as Bitcoin: \emph{SHA-256} applied twice. Thus, most of the math analysis we have done before is just the same.
-
-Let $X$ be the number of trials to solve the PoW and $T$ be the time between blocks.
-We already know that $X$ follows a geometric distribution with $p = \frac{A}{2^{256}}$, where $A$ is inversely proportional to the difficulty, i.e., the smaller the $A$, the higher the difficulty.
-We also know that $T$ follows an exponential distribution with $\lambda = \frac{1}{\eta}$, where $\eta$ is the average time between blocks.
-As proved before, $\eta p H = 1$, thus let's define $w = \log_2(\mathbf{E}(X)) = \log_2(1/p) = \log_2(\eta H) = \log_2(\eta) + \log_2(H)$.
-
-For the blocks, $\eta = 128$, thus, $\mathbf{E}(T) = 128 \text{ seconds}$, and $\mathbf{V}(T) = 16,384$.
-The symmetrical confidence interval with $\alpha = 10\%$ is $[6.56, 383.45]$, i.e., 90\% of the cases the distance between blocks will be between 6 seconds and just under 7 minutes.
-
-Let $H$ be the hash rate of the miners, then, the weight of the blocks is calculated by
-
-$$w_\text{blocks} = \log_2(128) + \log_2(H) = 7 + \log_2(H)$$
-
-This weight will be updated every 2,016 blocks to take into consideration the change of $H$.
-First, $H$ will be estimated using the fact that $\mathbf{E}(Y_{2016}) = \frac{2016}{\lambda} = 2016 \eta = \frac{2016}{pH} = \frac{2016 \cdot 2^w}{H}$.
-Thus, $H = \frac{2016 \cdot 2^w}{\Delta t} \Rightarrow \log_2(H) = w + \log_2(2016) - \log_2(\Delta t)$, where $\Delta t$ is the time between the latest weight update and now.
-Finally,
-
-$$w_\text{new} = 7 + w_\text{old} + \log_2(2016) - \log_2(\Delta t)$$
-
-Notice that, if the hash rate does not change, then $\Delta t = 128 \cdot 2016$ and $w_\text{new} = w_\text{old}$.
-
-\section{Weight of the transactions}
-
-There is a trade-off which must be considered in the weight of the transactions: the higher the weight, the better to prevent spams, but the worse to microtransactions.
-So, trying to fulfill both necessities, the weight will be a function of the transaction's size (in bytes) and total amount:
-
-$$w_\text{tx} = \left\lfloor \log_2(\text{size}) + \log_2(\text{amount}) + 0.5 \right\rfloor$$
-
-Although the transaction size may change a bit depending on the implementation, a typical transaction with 2 inputs and 2 outputs would have 188 bytes.
-So, for instance, transfering 50,000 tokens would require a weight of 23, which means an average of $2^{23}$ trials to solve the PoW.
-
-\section{Issuance rate}
-
-Bitcoin issuance rate started in 50 tokens per block and decreases over time.
-Every 210,000 blocks---4 years, on average---it halves the number of tokens issued per block.
-As Bitcoin smallest fraction is $10^{-8}$, after the 33th reduction it will stop issuing new tokens since $2^{33} > 50 \cdot 10^{-8}$.
-In total, the number of issued tokens will never exceed 21 million.
-For further information, see \cite{bitcoinsupply}.
-
-Using the same strategy, Hathor's issuance rate starts in 50,000 tokens per block and decreases over time.
-Different from Bitcoin, Hathor's tokens are always integers and cannot be split.
-Every 1,000,000 blocks, the number of tokens issued per block is halved.
-Thus, the supply is limited to 100 billion issued tokens during its whole life.
-To calculate the supply limit use the following equation:
-
-\begin{align*}
-    \sum_{k=0}^{\infty} 1,000,000 \cdot \left\lfloor \frac{50000}{2^k} \right\rfloor \approx
-    \sum_{k=0}^{\infty} 1,000,000 \cdot \frac{50000}{2^k} = 100 \text{ billion}
-\end{align*}
-
-As the average time between blocks is 128 seconds, it will take around 3 years, 9 months and 20 days between halves.
-
-\section{Transaction fees}
-
-The value of the fee is calculate as the difference between the transaction's outputs and the inputs.
-For instance, a transaction with inputs summing 10,000 HTR and outputs summing 9,000 HTR are paying 1,000 HTR of fees.
-
-Each transaction may pay a fee to the next block which confirms it.
-But, even if the fee is zero, the miners are forced to confirm at least two pending transactions (in-progress or unconfirmed), and these pending transactions confirms other transactions which may have no fees.
-In summary, transactions with no fees cannot be left behind and will always be confirmed by both blocks and other transactions.
-
-One may arguee that it allows the whole network to never pay fees and they are right.
-In the beginning, miners' incentive is driven by new tokens instead of fees.
-In the long term, it may be necessary to require a minimum fee to keep the miners working.
-
-\section{Transaction validation}
-
-A transaction will be considered valid when it complies with the following rules: (i) it spends only unspent outputs; (ii) the sum of the inputs is greater than or equal to the sum of the outputs; (iii) the number of inputs is at most 256; (iv) the number of outputs is at most 256; (v) it confirms at least two pending transactions; (vi) it solves the PoW with the correct weight.
-
-The transaction has a time field which is used to inform when the transaction was generated.
-The time field must be in UTC time and may be within at most 60 minutes from the current time.
-Otherwise, the transaction will be discarded.
-
-In case of transaction conflict, in which two transactions try to spend the same tokens, the one with higher accumulated weight is chosen and the other is invalidated.
-Although it is not possible in Iota because of the submarine attack, Hathor does not have the same problem.
-In Hathor, like in Bitcoin, the submarine attack is only possible if the attacker has a hash rate higher than the whole network, including the miners.
-In other words, when analysing the double spending attack, Hathor is as safe as Bitcoin.
-
-The digital signature is used to ensure that only the owners may spend their tokens.
-It will be calculated signing the transaction's input and output only.
-This allows the transaction to be signed in one device and to be sent to another device that will choose which transaction will be confirmed and will solve the proof-of-work.
-
-Services of solving proof-of-work may also be offered by companies.
-They give their customers a wallet address and they send the payment inside of the transaction itself.
-This allows IoT devices to save energy or to delegate the task of solving the PoW to whoever is receiving the tokens.
-
-\section{Orphan blocks}
-
-There is no orphan blocks in Hathor.
-Unless a block confirms either directly or indirectly an invalid transaction, every block is valid.
-There is no need to left a block behind since its proof-of-work is also protecting transactions.
-
-\section{Network}
-
-Like in all other cryptocurrencies, nodes will communicate through a peer-to-peer network.
-They will exchange transactions and blocks, verifying whether they are valid or not, while they will also answer queries.
-
-Thus, if a thin client wants to transfer some tokens, it just connects to the network and queries for the tips.
-With the response, it solves the proof-of-work and propagates the new transaction.
-In order to avoid malicious nodes, the thin client may ask to more than one node and randomly pick two transactions.
+\input{./Chapters/hathor-architecture.tex}
 
@@
 I also showed the transactions solving the proof-of-work which had not been propagated yet.
-See Figure \ref{fig:hathor-dag-big}} and notice the chain of blocks inside the DAG.
+See Figure \ref{fig:hathor-dag-big} and notice the chain of blocks inside the DAG.
 

### 2018-04-12 Chapters/hathor-architecture.tex
Hathor: splitting Hathor's architecture and analysis of Bitcoin in separate files.

This work introduces the Hathor's architecture which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues. Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG). For this, each transaction has its own proof-of-work which is solved by the issuer before propagating the transactions in the network. Like Bitcoin, miners find new ``blocks'' every 10 minutes in which they collect the fees and newly generated tokens. Each transaction has an ``accumulated weight'' which express the required effort to break the transaction, similar to Bitcoin's number of confirmations.

This way, there will be two difficulty levels in the system: one for new transactions which are just moving tokens around, and another one for ``blocks'' which are generating new tokens and collecting fees. The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions. The second is adjusted every 2,016 blocks to keep the pace of block on every 2 minutes.

There is a trade-off about the required proof-of-work of new transactions. The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.

Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them. It is always possible since there is at least one confirmation path between any transaction and a tip. This is important for security reasons.

The transactions are classified into three groups: (i) confirmed transactions, (ii) in-progress transactions, and (iii) unconfirmed transactions (tips). The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level. The unconfirmed transactions (tips) are the brand new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero. The in-progress transactions are in the middle. They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.

A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two in-progress transactions or tips. There may be any number of outputs provided that they comply with the number of newly generated tokens. Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.

In the low load scenario, there is a small number of new transactions coming into the network, which means they give a minor contribution to confirmations. In this case, the confirmation is held mostly by ``blocks''. On the other hand, in the high load scenarios, there is a large number of new transactions giving a major contribution to confirmations. In this case, the ``blocks'' strengthen the confirmation, but most of them will have already been confirmed before the new blocks are found. The higher the number of new transaction, the faster the transactions are confirmed. The blocks assure a ``maximum confirmation time''.

The incentive scheme which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.

\section{Transaction confirmation}

The concepts of weight and accumulated weight are the same as Iota's, but their equations are different. The weight of a transaction is calculated as $\log_2(k)$ where $k$ is the average number of hashes required to solve its proof-of-work again, while the accumulated weighted of a transaction A with weight $w_A$ is calculated as $\log_2(2^{w_A} + \sum_{A \leadsto P} 2^{w_P})$. The $\log$ function is used to reduce the magnitude of the numbers. Hence, the power-of-2 of the accumulated weight is a measure of the average number of hashes required to solve the proof-of-work of all transactions that are confirming A.

In Bitcoin, it is well-known that one should wait at least ``six confirmations'' before accepting a transaction. This Bitcoin's criteria is based on some math presented in Satoshi's seminal work \citep{nakamoto2008bitcoin}. Adopting six confirmations is the same as demanding from attackers a minimal effort of six times the network's hash rate to successfully double spend those tokens. Let $H$ be the total hash rate of the network. Then, as $\mathbf{E}(Y_6) = 60 \text{ minutes}$, it will be necessary to calculate, on average, $\mathbf{E}(Y_6) \cdot H = 60 \cdot 60 \cdot H$ hashes to solve the proof-of-work of 6 blocks.

Therefore, in order to have the same level of security as Bitcoin, a transaction is said to be confirmed when its accumulated weight is greater than or equal to $\log_2(\mathbf{E}(Y_6) \cdot H) = \log_2(6 \cdot 128 \cdot H) = 7 + \log_2(6) + \log_2(H)$, where $H$ is calculated as the total hash rate of the miners plus the total hash rate of new transactions.

\section{Time between blocks}

The hash function used for the proof-of-work (PoW) is the same as Bitcoin: \emph{SHA-256} applied twice. Thus, most of the math analysis we have done before is just the same.

Let $X$ be the number of trials to solve the PoW and $T$ be the time between blocks. We already know that $X$ follows a geometric distribution with $p = \frac{A}{2^{256}}$, where $A$ is inversely proportional to the difficulty, i.e., the smaller the $A$, the higher the difficulty. We also know that $T$ follows an exponential distribution with $\lambda = \frac{1}{\eta}$, where $\eta$ is the average time between blocks. As proved before, $\eta p H = 1$, thus let's define $w = \log_2(\mathbf{E}(X)) = \log_2(1/p) = \log_2(\eta H) = \log_2(\eta) + \log_2(H)$.

For the blocks, $\eta = 128$, thus, $\mathbf{E}(T) = 128 \text{ seconds}$, and $\mathbf{V}(T) = 16,384$. The symmetrical confidence interval with $\alpha = 10\%$ is $[6.56, 383.45]$, i.e., 90\% of the cases the distance between blocks will be between 6 seconds and just under 7 minutes.

Let $H$ be the hash rate of the miners, then, the weight of the blocks is calculated by

$$w_\text{blocks} = \log_2(128) + \log_2(H) = 7 + \log_2(H)$$

This weight will be updated every 2,016 blocks to take into consideration the change of $H$. First, $H$ will be estimated using the fact that $\mathbf{E}(Y_{2016}) = \frac{2016}{\lambda} = 2016 \eta = \frac{2016}{pH} = \frac{2016 \cdot 2^w}{H}$. Thus, $H = \frac{2016 \cdot 2^w}{\Delta t} \Rightarrow \log_2(H) = w + \log_2(2016) - \log_2(\Delta t)$, where $\Delta t$ is the time between the latest weight update and now. Finally,

$$w_\text{new} = 7 + w_\text{old} + \log_2(2016) - \log_2(\Delta t)$$

Notice that, if the hash rate does not change, then $\Delta t = 128 \cdot 2016$ and $w_\text{new} = w_\text{old}$.

\section{Weight of the transactions}

There is a trade-off which must be considered in the weight of the transactions: the higher the weight, the better to prevent spams, but the worse to microtransactions. So, trying to fulfill both necessities, the weight will be a function of the transaction's size (in bytes) and total amount:

$$w_\text{tx} = \left\lfloor \log_2(\text{size}) + \log_2(\text{amount}) + 0.5 \right\rfloor$$

Although the transaction size may change a bit depending on the implementation, a typical transaction with 2 inputs and 2 outputs would have 188 bytes. So, for instance, transfering 50,000 tokens would require a weight of 23, which means an average of $2^{23}$ trials to solve the PoW.

\section{Issuance rate}

Bitcoin issuance rate started in 50 tokens per block and decreases over time. Every 210,000 blocks---4 years, on average---it halves the number of tokens issued per block. As Bitcoin smallest fraction is $10^{-8}$, after the 33th reduction it will stop issuing new tokens since $2^{33} > 50 \cdot 10^{-8}$. In total, the number of issued tokens will never exceed 21 million. For further information, see \cite{bitcoinsupply}.

Using the same strategy, Hathor's issuance rate starts in 50,000 tokens per block and decreases over time. Different from Bitcoin, Hathor's tokens are always integers and cannot be split. Every 1,000,000 blocks, the number of tokens issued per block is halved. Thus, the supply is limited to 100 billion issued tokens during its whole life. To calculate the supply limit use the following equation:

\begin{align*}
    \sum_{k=0}^{\infty} 1,000,000 \cdot \left\lfloor \frac{50000}{2^k} \right\rfloor \approx
    \sum_{k=0}^{\infty} 1,000,000 \cdot \frac{50000}{2^k} = 100 \text{ billion}
\end{align*}

As the average time between blocks is 128 seconds, it will take around 3 years, 9 months and 20 days between halves.

\section{Transaction fees}

The value of the fee is calculate as the difference between the transaction's outputs and the inputs. For instance, a transaction with inputs summing 10,000 HTR and outputs summing 9,000 HTR are paying 1,000 HTR of fees.

Each transaction may pay a fee to the next block which confirms it. But, even if the fee is zero, the miners are forced to confirm at least two pending transactions (in-progress or unconfirmed), and these pending transactions confirms other transactions which may have no fees. In summary, transactions with no fees cannot be left behind and will always be confirmed by both blocks and other transactions.

One may arguee that it allows the whole network to never pay fees and they are right. In the beginning, miners' incentive is driven by new tokens instead of fees. In the long term, it may be necessary to require a minimum fee to keep the miners working.

\section{Transaction validation}

A transaction will be considered valid when it complies with the following rules: (i) it spends only unspent outputs; (ii) the sum of the inputs is greater than or equal to the sum of the outputs; (iii) the number of inputs is at most 256; (iv) the number of outputs is at most 256; (v) it confirms at least two pending transactions; (vi) it solves the PoW with the correct weight.

The transaction has a time field which is used to inform when the transaction was generated. The time field must be in UTC time and may be within at most 60 minutes from the current time. Otherwise, the transaction will be discarded.

In case of transaction conflict, in which two transactions try to spend the same tokens, the one with higher accumulated weight is chosen and the other is invalidated. Although it is not possible in Iota because of the submarine attack, Hathor does not have the same problem. In Hathor, like in Bitcoin, the submarine attack is only possible if the attacker has a hash rate higher than the whole network, including the miners. In other words, when analysing the double spending attack, Hathor is as safe as Bitcoin.

The digital signature is used to ensure that only the owners may spend their tokens. It will be calculated signing the transaction's input and output only. This allows the transaction to be signed in one device and to be sent to another device that will choose which transaction will be confirmed and will solve the proof-of-work.

Services of solving proof-of-work may also be offered by companies. They give their customers a wallet address and they send the payment inside of the transaction itself. This allows IoT devices to save energy or to delegate the task of solving the PoW to whoever is receiving the tokens.

\section{Orphan blocks}

There is no orphan blocks in Hathor. Unless a block confirms either directly or indirectly an invalid transaction, every block is valid. There is no need to left a block behind since its proof-of-work is also protecting transactions.

\section{Network}

Like in all other cryptocurrencies, nodes will communicate through a peer-to-peer network. They will exchange transactions and blocks, verifying whether they are valid or not, while they will also answer queries.

Thus, if a thin client wants to transfer some tokens, it just connects to the network and queries for the tips. With the response, it solves the proof-of-work and propagates the new transaction. In order to avoid malicious nodes, the thin client may ask to more than one node and randomly pick two transactions.

### 2018-04-12 Chapters/hathor-bitcoin-math.tex
Hathor: splitting Hathor's architecture and analysis of Bitcoin in separate files.

The primary objective of this chapter is to increase our understanding of the Bitcoin through mathematical tools.

\section{Hash function}

Hash functions has been widely studied in computer science. In short, a hash function $h: \{0, 1\}^\infty \rightarrow \{0, 1\}^{n}$ has the following properties:

\begin{enumerate}
	\item $x = y \Rightarrow h(x) = h(y)$
	\item $h(x) \sim \mathcal{U}(0, 2^{n}-1)$, where $\mathcal{U}$ is the uniform distribution, i.e., $\forall a \in [0, 2^{n}-1], \mathbf{P}(h(x)=a) = \frac{1}{2^{n}}$
\end{enumerate}

In other words, when two inputs are the same, they have the same output. But, when the inputs are different, their outputs are uniformly distributed. Clearly, the hash functions are surjective but not injective. They are not injective because the image of $h$ has only $2^n$ elements and the domain has infinite elements. When $x \ne y$ and $h(x) = h(y)$, we say that $x$ and $y$ are a collision. A hash function is considered to be safe when it is unknown how to quickly find a collision of a given hash, i.e., one has to check all possible values until the correct one is found (known as the brute-force attack).

Bitcoin uses two hash functions: HASH-160 and HASH-256. The first has $n=160$ and consists of the composition of \textit{SHA-256} and \textit{RIPEMD-160}. The latter has $n=256$ and applies \textit{SHA-256} twice. The first is used in transactions' scripts and the latter in the mining algorithm. For both hash functions, it is infeasible to run a brute-force attack because it would demand, on average, either $2^{160}$ or $2^{256}$ trials, and those would take a tremendous amount of time even for the fastest known processors.

For further information about hash functions, see \citet{gilbert2003security, dobbertin1996ripemd}.

\section{Mining one block}

Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin \textit{HASH-256} function. The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given threshold. The smaller the $A$, the harder to find a new block. In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.

Hence, in order to find a new block, one must try different inputs ($x_1, x_2, \dots, x_k$) until they find a solution, i.e., all attempts will fail ($h(x_i) \geq A$ for $i < k$) but the last ($h(x_k) < A$). The probability of finding a solution exactly in the $k^{th}$ attempt follows a geometric distribution. Let $X$ be the number of attempts until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \frac{A}{2^{256}}$. Also, we have $\mathbf{P}(X \leq k) = 1 - (1-p)^k$. The average number of attempts is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = \frac{1-p}{p^2}$.

In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average. Suppose that the Bitcoin's network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network. The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block. So, the rule of finding a new block every 10 minutes ($\eta = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \eta = 600$. So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \eta = 600 \Rightarrow p = \frac{1}{\eta H}$. Finally, $\mathbf{E}(X) = \eta H$, $\mathbf{E}(T) = \eta$, $\mathbf{V}(X) = (\eta H)^2 - \eta H$, and $\mathbf{V}(T) = \eta^2 - \eta/H$.

The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\eta H} \right)^{tH}$. But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.

\begin{theorem}
	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{\eta}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$.
\end{theorem}
\begin{proof}
\begin{align*}
	\mathbf{P}(T \leq t) &= 1 - (1-p)^{tH} \\
		     &= 1 - \left( 1 - \frac{1}{\eta H} \right)^{tH} \\
	\text{Replacing $u = \eta H$,} \\
	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{\eta}} \\
				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{\eta}} \\
				&= 1 - \left( 1/e \right)^{\frac{t}{\eta}} \\
				&= 1 - e^{-\frac{t}{\eta}}
\end{align*}
\end{proof}

Now, we would like to understand from which value of $H$ it is reasonable to assume that $T$ follows an exponential distribution.

\begin{theorem}
	$x > M \Rightarrow |(1+1/x)^x - e| < e/M$.
\end{theorem}
\begin{proof}
	Let's use the classical inequality $\frac{x}{1+x} < \log(1+x) < x$ for $x > -1$. So, $\frac{1/x}{1+1/x} < \log(1+x) < 1/x$. Simplifying, $\frac{1/x}{1+1/x} = 1/(1+x)$.  Thus, $1/(1+x) < \log(1+1/x) < 1/x \Rightarrow x/(1+x) < x \log(1 + 1/x) < 1$.

	As $\log(1 + \frac{1}{M}) > 0$ and $1 < 1 + \log(1 + \frac{1}{M})$.

	$x > M \Rightarrow 1/x < 1/M \Rightarrow 1+1/x < 1+1/M \Rightarrow 1/(1+1/x) > 1/(1+1/M) \Rightarrow x/(1+x) > M/(1+M)$.

	Again, $\log(1+x) < x \Rightarrow \log(1-1/M) < -1/M \Rightarrow 1 + \log(1-1/M) < (M-1)/M < M/(1+M)$, since $(x-1)/x < x/(x+1)$.

	Hence, $1 + \log(1-1/M) < M/(1+M) < x/(1+x) < x\log(1+1/x)$, and $x\log(1+1/x) < 1 < 1 + \log(1 + \frac{1}{M})$.

	Finally,

	\begin{align*}
		1 + \log(1-1/M) &< x\log(1+1/x) < 1 + \log(1 + \frac{1}{M}) \\
		e^{1 + \log(1-1/M)} &< e^{x\log(1+1/x)} < e^{1 + \log(1 + \frac{1}{M})} \\
		e \cdot e^{\log(1-1/M)} &< e^{\log((1+1/x)^x)} < e \cdot e^{\log(1 + \frac{1}{M})} \\
		e(1-1/M) &< (1+1/x)^x < e(1 + \frac{1}{M}) \\
		e-e/M &< (1+1/x)^x < e + e/M \\
		-e/M &< (1+1/x)^x - e < e/M \\
	\end{align*}
	Therefore, $\left| \left( 1+1/x \right)^x - e \right| < e/M$.
\end{proof}

We may consider $H$ big enough to say that $T$ follows an exponential distribution when $e/H < \epsilon$, where $\epsilon$ is the maximum approximation error. When $\epsilon = 10^{-6} \Rightarrow H > e \cdot 10^6$. So, when $H > 2.6 \text{Mh/s}$, our approximation is good enough.

The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, where $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$. These conditions give the following equations: $1 - e^{-t_0/\eta} = \alpha/2$, and $e^{-t_1/\eta} = \alpha/2$. Solving these equations, we have $t_0 = -\eta \ln{(1 - \alpha/2)}$, and $t_1 = -\eta \ln{(\alpha/2)}$.

For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ (or $[0.51, 30.76]$ in minutes). Thus, 90\% of the time the intervals between blocks are between 30 seconds and 30 minutes, with average of 10 minutes.

The fact that the time between blocks follows an exponential distribution with $\lambda = 1/\eta = pH$ may be used to estimate the total network's hash rate (or a miner's hash rate). For further information, see \cite{ozisikestimation}.

\section{Mining several blocks}

Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on. Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i$ which is the total time to find the next $n$ blocks. As $Y_n$ is the sum of random variables which follow an exponential distribution with same $\lambda = \frac{1}{\eta}$, then $Y_n \sim \text{Erlang}(n, \frac{1}{\eta})$. Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.

Many exchanges require at least six confirmations in order to accept a deposit in Bitcoin. So, for $n=6$, $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$, i.e., only 55\% of the deposits will be accepted in one hour. The symmetrical confidence interval with $\alpha=10\%$ is $[27, 105]$ in minutes. Thus, 90\% of the times, it will take between 27 minutes and 1 hour and 45 minutes to have your deposit accepted --- assuming that your transaction will be confirmed in the very next block. The pdf of $Y_6$ is shown in Figure \ref{fig-bitcoin-time-6-blocks}, in which the 10\% symmetrical confidence interval is shown in the white area. The average total time of six confirmations is $\mathbf{E}(Y_6) = 6 \cdot 600 = 3600 = 60 \text{ minutes}$.

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/time-6-blocks.pdf}
\caption{Probability density function of $Y_6$, i.e., probability of finding 6 blocks after time $t$. The shaded areas shows the lower 5\% and upper 5\% of the pdf.\label{fig-bitcoin-time-6-blocks}}
\end{figure}

\section{Mining for a miner}

Let's analyze the probability of finding a new block for a miner who has $\alpha$ percent of the network's total hash rate. Let $T_\alpha = \frac{X}{\alpha H}$ be the time required for the miner to find a new block. As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{\eta}$. Hence, we confirm the intuition that the miner with $\alpha$ percent of the network's total hash power will find $\alpha$ percent of the blocks.

\begin{theorem}
	When the miner with $\alpha$ percent of the network's total hash rate is part of the mining network, $\mathbf{P}(\text{next block is from $T_\alpha$}) = \alpha.$
\end{theorem}
\begin{proof}
\begin{align*}
	\mathbf{P}(\text{next block is from $T_\alpha$}) &= P \left( T_\alpha = \min\{T_\alpha, T_{1-\alpha}\} \right) \\
		&= \frac{\lambda_\alpha}{\lambda_\alpha + \lambda_{1-\alpha}} \\
		&= \frac{\alpha/\eta}{\alpha/\eta + (1-\alpha)/\eta} \\
		&= \frac{\alpha}{\alpha + 1 - \alpha} \\
		&= \alpha.
\end{align*}
\end{proof}

\begin{theorem}
	When one miner with $\alpha$ percent of the network's total hash rate multiplies their hash rate by $m$, the probability of this miner find the next block is multiplied by $\frac{m}{m \alpha + 1 - \alpha}$.
	\label{thm-miner-multiply}
\end{theorem}
\begin{proof}
	When miners increase their hash rate, they also increase the network's total hash rate. Let $H$ be the network's hash rate before the increase. Thus, the network's total hash rate after the increase is $H + (m-1) \alpha H = (1 - \alpha + m \alpha) H$. So,
\begin{align*}
	\mathbf{P}(\text{next block is from $T_{m \alpha}$}) &= P \left( T_{m \alpha} = \min\{T_{m \alpha}, T_{1-\alpha}\} \right) \\
		&= \frac{\lambda_{m \alpha}}{\lambda_{m \alpha} + \lambda_{1-\alpha}} \\
		&= \frac{m \alpha/\eta}{m \alpha/\eta + (1-\alpha)/\eta} \\
		&= \frac{m \alpha}{m \alpha + 1-\alpha} \\
		&= \alpha \left( \frac{m}{m \alpha + 1 - \alpha} \right).
\end{align*}
\end{proof}

\begin{cor}
	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximately multiplies their probability of finding the next block by $m$.
\end{cor}
\begin{proof}
	$$\lim_{\alpha \rightarrow 0} \mathbf{P}(\text{next block is from $T_{m \alpha}$}) = \lim_{\alpha \rightarrow 0} \frac{m}{m \alpha + 1 - \alpha} = m.$$
\end{proof}

That way, it is not exactly correct to say that when one doubles their hash rate, their probability will double as well. It is only true for small miners.

\section{Orphan blocks}
% Probabilidade de orfão
An orphan block would be created if a new block is found during the propagation time of a new block. Let $\alpha$ be the percentage of the total hash rate of the node which is outdated, and $\Delta t$ the propagation time in seconds. Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{\eta}}$.

Bitcoin peer-to-peer network is a gossip network, where miners are semi-randomly connected to each other, and each miner sends all information it receives to all its peers. According to \citet{decker2013information}, the average time for a new block propagate over the network is 12.6 seconds, while the 95\% percentile is 40 seconds, which indicates a long-tail distribution. \citet{bitcoinstats} has measured the propagation time between 2013 and 2017. During 2017, the worst daily 90\% percentile was 21 seconds. Notice that both results may not be contradictory because Bitcoin network is continuously evolving.

For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, then $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$, which is almost 0.5\%. I would say that a node with 10\% of the total hash rate would be well connected and it would take less time to receive the update, so, the probability would be even smaller than 0.5\%.

Another important factor is that, as Bitcoin is open-source, miners are free to change the gossip algorithm, which leads to the network incentives. See \citet{babaioff2012bitcoin} for an analysis of the incentives to miners forward new blocks and transactions in the network.

For further information about gossip algorithms, see \citet{shah2009gossip}.

\section{Analysis of network's hash rate change}
% Análise: Mudança no network's hash rate antes de atualizar o $A$.

The difficulty, given by the number $A$, is adjusted every 2016 blocks. As, $\mathbf{P}(13 \text{ days} < Y_{2016} < 15 \text{ days}) = \mathbf{P}(13 \cdot 24 \cdot 3600 < Y_{2016} < 14 \cdot 24 \cdot 3600) = 0.9986$, it is expected that the total time to find 2016 blocks will be between 13 and 15 days, assuming that the network's hash rate remains constant. If it takes less than the expected time, it means that the network's total hash rate has increased. While if it takes more than the expected time, it means that the network's total hash rate has decreased. So, let's analyze what happens when the network's hash rate changes significantly.

%Let's assume that the network's total hash rate will change only once, from $H$ to $\alpha H$ --- so, if $\alpha = 1$, there would be no change, if $\alpha > 1$, there would be an increase, and if $\alpha < 1$, there would be a decrease. As the time required to find a new block would be $T_\alpha = \frac{X}{\alpha H}$, we got the same random variable as we have gotten when analyzing a miner with $\alpha$ percent of the network's total hash rate. But, the analysis will be different because here $\alpha$ may be greater than one.

Let $H \cdot u(t)$ be the network's total hash rate over time. So, the number of hashes calculated in $t$ seconds is $H \int_0^t u(t) dt$. Hence, $\mathbf{P}(T \leq t) = \mathbf{P}(X \leq H \int_0^t u(t)dt)$. When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{\eta} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{\eta} \cdot e^{-\frac{1}{\eta} \int_0^t u(t)dt}$.

Let's say that the network's total hash rate has suddenly multiplied by $\alpha$. So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\eta}$. Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\eta})$. Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$. Hence, on one hand, when the network's hash rate increases ($\alpha > 1$), the 2016 blocks will be found earlier. On the other hand, when the network's hash rate decreases ($\alpha < 1$), the 2016 blocks will be found later.

For example, if the network's total hash rate suddenly doubles ($\alpha = 2$), then $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved. On the other side, if the network's total hash rate suddenly halves ($\alpha = 0.5$), then $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled. It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, at which point the adjustment will occur and everything will be back to the normal 10 minutes between blocks.

% TODO Step-wise analysis

\subsection{Hash rate smoothly changing}

Let $u(t) = \frac{1+abx}{1+bx}$. It is an useful function because $u(0) = 1$ and $\lim_{t \rightarrow \infty} u(t) = a$. The bigger the $b$, the faster $u(t) \rightarrow a$. For example, if $a=2$, it means $H$ would be smoothly doubling. If $a=0.5$, it means $H$ would be smoothly halving.

It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a \Rightarrow \int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

Assuming that $n = \frac{\lambda(a-1)}{b}$ is integer, we have:
$$F_T(t) = 1 - (1+bt)^n e^{-\lambda at}$$

Thus,
\begin{align*}
\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
\end{align*}

Hence, as $\mathcal{L}\{f_T(t)\} = s \mathcal{L}\{F_T(t)\}$,

$$\mathcal{L}\{f_T(t)\} = 1 - \sum_{k=0}^n \binom{n}{k} \frac{s b^k k!}{(s+\lambda a)^{k+1}}$$

Then,

\begin{align*}
\frac{d}{ds} \mathcal{L}\{f_T(t)\}
	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{d}{ds} \frac{s}{(s+\lambda a)^{k+1}} \\
	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \left[ \frac{1}{(s+a\lambda)^{k+1}} - \frac{s(k+1)}{(s+a \lambda)^{k+1}} \right] \\
\frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0}
	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{1}{(\lambda a)^{k+1}} \\
	&= - \frac{1}{a\lambda} \sum_{k=0}^n \binom{n}{k} k! \left( \frac{b}{\lambda a} \right)^k \\
	&= - \frac{1}{a\lambda} \sum_{k=0}^n \frac{n!}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \\
	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \right] \\
	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{n-k} \right] \tag{$k \rightarrow n-k$} \\
	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{-k} \right] \\
	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right]
\end{align*}

Finally, as $\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0}$,
$$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right] \text{, where $n = \frac{\lambda(a-1)}{b}$}$$

Let's check this equation for already known scenarios. When $a=1$, then $n=0$ and $\mathbf{E}[T] = 1/\lambda$. When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied. In fact, $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.

\begin{theorem}
	$$a > 1 \text{ and } x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
\end{theorem}
\begin{proof}
$x > M \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM}$. As $1-a<0$, $\frac{1-a}{1+bx} > \frac{1-a}{1+bM}$. Thus, $\frac{1-a}{1+bM} < \frac{1-a}{1+bx} + a - a = \frac{1+abx}{1+bx} - a < 0 < \frac{a-1}{1+bM}$. Hence, $-\frac{a-1}{1+bM} < \frac{1+abx}{1+bx} - a < \frac{a-1}{1+bM}$.
\end{proof}

For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$. So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected. Finally, $\mathbf{E}[T] = 557.65$. In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes. If the hash rate had suddenly doubled, the average time between blocks would be 5 minutes.

%$$\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0} = \frac{b^n n! 600^{n+1}}{a^{n+1}} e^{\frac{a}{600b}} \text{, where $n = \frac{a-1}{600b}$}$$

%$a > 1, x > M \Rightarrow 1 + bx > 1 + bM \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM} \Rightarrow \frac{1-a}{1+bx} + a > \frac{1-a}{1+bM} + a \Rightarrow \frac{1+abx}{1+bx} - a > \frac{1-a}{1+bM}$.

\subsection{Piecewise linear model of hash rate change}

Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(a, b, t) = a^2t + b$. Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{bt + a^2t^2/2}{\eta}}$.

It is well known that $\mathbf{E}(T) = \int_{0}^{\infty} 1 - \mathbf{P}(T \leq t) dt$. Thus, replacing $y = \frac{a^2 t + b}{a \sqrt{2 \eta}}$, and using the fact that $\int_0^\infty e^{-x^2} dx = \frac{\sqrt{\pi}}{2} \erf(x)$, we have:

\begin{align}
\mathbf{E}(T)|_{t_1}^{t_2} &= \int_{t_1}^{t_2} \exp \left( - \frac{bt + a^2 t^2/2}{\eta} \right) dt \nonumber \\
	&= \frac{\sqrt{2 \eta}}{a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \int_{y_1}^{y_2}  \exp(-y^2) dy \nonumber \\
	&= \frac{\sqrt{2 \eta}}{a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \frac{\sqrt{\pi}}{2} [\erf(y_1) - \erf(y_2)] \nonumber \\
	&= \frac{\sqrt{2 \pi \eta}}{2a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \left[\erf(y_2) - \erf(y_1) \right]
\end{align}

Where $y_1 = \frac{a^2 t_1 + b}{a \sqrt{2 \eta}}$ and $y_2 = \frac{a^2 t_2 + b}{a \sqrt{2 \eta}}$.

Thus, $\mathbf{E}(T) = \mathbf{E}(T)|_0^\infty$. When $t_1 = 0 \Rightarrow y_1 = \frac{b^2}{2 \sqrt{2 \eta}}$ and $t_2 \rightarrow \infty \Rightarrow y_2 \rightarrow \infty \Rightarrow \erf(y_2) = 1$, then:

$$
\mathbf{E}(T) = \frac{\sqrt{2 \pi \eta}}{2a} \exp \left( \frac{b^2}{2a^2 \eta} \right) \left[1 - \erf\left(\frac{1}{a \sqrt{2 \eta}}\right)\right]
$$

The function $\mathbf{E}(T)|_{t_1}^{t_2}$ may be used to a piecewise linear analysis of any hash rate change. Let's analyze the hash doubling in one week. Then, $u(1 \text{week}) = u(604800) = 2$, thus $a^2 = \frac{1}{604800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$. Let's sample the interval $[0, 1 \text{week}]$ every hour, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.

For each point $t_i$, let $g_i(t) = [H(t-t_i) - H(t-604800)] u(t) + 2 H(t-604800)$, then $\mathbf{E}(T) = \mathbf{E}(T)|_{t_i}^{t_{i+1}} + 1 - e^{\frac{t}{\eta}}$. The result is presented in Figure .

%In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.
%
%Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.
%
%\begin{align*}
%	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
%		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\tau}} \right) dx \\
%		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\tau}} dx \\
%		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx \\
%\end{align*}
%
%Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
%Let $y = bx+c \Rightarrow dy = bdx$.
%
%\begin{align*}
%	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
%		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
%		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
%		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
%\end{align*}
%
%$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
%
%Hence,
%$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
%
%As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,
%
%\begin{align*}
%\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
%\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
%\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
%\end{align*}
%
%$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.
%
%\textcolor{red}{I do not know how to interpret this.}

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
\end{figure}

\section{Attack in the Bitcoin network}
% Ataque submarino
% Probabilidade de "correr por fora"

\citet{karame2012two}

There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.

In the double spending attack, the attacker's send some funds to the victim, let's say a merchant. They wait for $k$ confirmations of the transaction, and the victim delivers the good or the service to the attacker. Then, the attacker mine enough blocks with a conflicting transaction, double spending the funds which was sent to the victim. If the attacker is successful, the original transaction will be \textit{erased} and the victim will be left with no funds at all. In order to be successful, the attacker must propagate more blocks than the network in the same period, propagating a chain longer than the main chain. Hence, we would like to understand what the odds are that the attacker will be successful. This attack was originally discussed by \citet{nakamoto2008bitcoin}.

In order to maximize their odds, the attacker must start to mine the new blocks as soon as they send the funds to the victim. In this moment, it starts to mine in the head of the blockchain, just like the rest of the network. So, in the beginning, the attacker and the network are in exactly the same point.

Let $\beta H$ be the hash rate of the attackers, and $\gamma H$ be the network's hash rate without the attackers. Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{\eta}$ and $\lambda_{\text{network}} = \frac{\gamma}{\eta}$, respectively.

As \cite{nakamoto2008bitcoin} has done, we will also model the attack using the Gambler's Ruin. In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$. The rounds are independent. The gambler starts with \$$k$ plays continuously until he either accumulates a target amount of \$$m$, or loses all his money. Let $\rho = \frac{1-p}{p}$, then the probability of losing his fortune is:

$$
\mathbf{P}(\text{losing his fortune}) =
\begin{cases}
	\frac{\rho^k - \rho^m}{1-\rho^m} \text{, if $\rho \ne 1$,} \\
	\frac{m-k}{m} \text{, if $\rho = 1$.}
\end{cases}
$$

When $m \rightarrow +\infty$,

$$
\mathbf{P}(\text{losing his fortune}) =
\begin{cases}
	\rho^k \text{, if $\rho < 1$,} \\
	1 \text{, if $\rho \geq 1$.}
\end{cases}
$$

% TODO Expected time to absorption.

The gambler winning \$1 is the same as the network finding a new block, the gambler losing \$1 is the same as the attacker finding a new block. The initial \$$k$ is the same as the number of blocks the attacker is behind the network. Thus, the gambler loses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.

In our case, $p = \frac{\lambda_{\text{network}}}{\lambda_{\text{network}} + \lambda_{\text{attacker}}} = \frac{\gamma}{\beta + \gamma}$, thus $\rho = \frac{\beta}{\gamma}$. Hence, $\rho < 1 \Leftrightarrow \beta < \gamma$.

Suppose that the attacker is mining with the network. Suddently, he stops mining with the network and starts attacking, i.e., starts to mine in another chain. In this scenario, since the attacker's hash rate is not mining with the network anymore, $\gamma = 1 - \beta$. Thus, $\beta < \gamma \Rightarrow \beta < 0.5 \Leftrightarrow \rho < 1$. Here comes the conclusion that, if the attacker has 50\% or more of the network's hash rate, then his attack will be certainly successful. We got exactly the same equations and conclusions as \cite{nakamoto2008bitcoin}.

But this scenario seems not to be the optimal attack, because the attacker has waited $k$ confirmations before starting the attack. A better approach would be to start attacking just after propagating the transaction. In this case, our previous model is not good, because even if the attacker have found more blocks than the network, he cannot propagate those blocks before the network has found $k$ confirmations. So, we have to model the probabilities before the network has found the $k$ block. Then, if the attacker has more blocks than the network, he has successfully attacked. Otherwise, we return to the previous model, in which the attacker must still find more blocks.

\begin{theorem}
	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
\end{theorem}
\begin{proof}
The attacker must find exactly $s$ blocks while the network must find exactly $k$ blocks. It is as they would be walking the grid from the point $(0, 0)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}. When the attacker finds a block, it would be a movement to the right. When the network finds a block, it would be an upward movement. No matter the order which the blocks are found, all the paths occur with probability $(1-p)^s p^k$.

The walking ends when $(\cdot, k$) is reached, i.e., when the network finds $k$ blocks, regardless of how many blocks the attacker has found -- i.e., it is not allowed to walk above the line $(\cdot, k)$. Thus, the number of paths between $(0, 0)$ and $(s, k)$ moving only upward or to the right, without going into the line $(\cdot, k)$ is exactly the number of paths between $(0, 0)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s$ movements to the right ($r$) and $k-1$ upward movements ($u$). This number of permutations is $\frac{(k-1+s)!}{s!(k-1)!} = \binom{(k-1)+(s)}{s}$ because there are $s$ repetitions of the element $r$ and $k-1$ repetitions of the element $u$.

Finally, the probability is $\binom{k+s-1}{s} (1-p)^{s} p^{k}$.

\begin{figure}[ht]
  \centering
  \begin{tikzpicture}
    \coordinate (Origin)   at (0,0);
    \coordinate (XAxisMin) at (-1,0);
    \coordinate (XAxisMax) at (5,0);
    \coordinate (YAxisMin) at (0,-1);
    \coordinate (YAxisMax) at (0,5);
    %\draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
    %\draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis

    \clip (-0.5,-0.5) rectangle (10.5cm,7cm); % Clips the picture...
    %\pgftransformcm{1}{0}{0}{1}{\pgfpoint{0cm}{0cm}}
          % This is actually the transformation matrix entries that
          % gives the slanted unit vectors. You might check it on
           % MATLAB etc. . I got it by guessing.
    \draw[style=help lines,dashed] (0,0) grid[step=1cm] (14,6);
          % Draws a grid in the new coordinates.
          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
              % Puts the shaded rectangle
    \foreach \x in {0,1,...,10}{% Two indices running over each
      \foreach \y in {0,1,...,6}{% node on the grid we have drawn
        \node[draw,circle,inner sep=1pt,fill] at (\x,\y) {};
            % Places a dot at those points
      }
    }

    \draw [ultra thick,-latex,red] (0,0) -- (0,1) node [midway, left] {};
    \draw [ultra thick,-latex,red] (0,1) -- (0,2) node [midway, left] {};
    \draw [ultra thick,-latex,red] (0,2) -- (1,2) node [midway, above] {};
    \draw [ultra thick,-latex,red] (1,2) -- (1,3) node [midway, right] {};
    \draw [ultra thick,-latex,red] (1,3) -- (2,3) node [midway, right] {};
    \draw [ultra thick,-latex,red] (2,3) -- (2,4) node [midway, right] {};
    \draw [ultra thick,-latex,red] (2,4) -- (2,5) node [midway, right] {};
    \draw [ultra thick,-latex,red] (2,5) -- (3,5) node [midway, right] {};
	\draw [ultra thick,-latex,red] (3,5) -- (3,6) node [above] {(3, 6)};

    \draw [ultra thick,-latex,blue] (0,0) -- (1,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (1,0) -- (2,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,0) -- (2,1) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,1) -- (2,2) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,2) -- (2,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,3) -- (3,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (3,3) -- (4,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (4,3) -- (5,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,3) -- (5,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,4) -- (6,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,4) -- (6,5) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,5) -- (7,5) node [midway, left] {};
	\draw [ultra thick,-latex,blue] (7,5) -- (7,6) node [above] {(7,6)};
  \end{tikzpicture}
  \caption{Both the attacker and the network are mining. Each step up is a new block found by the network with probability $p$. Each step right is a new block found by the attacker with probability $1-p$. It ends when the network finds $k$ blocks --- in this example, $k=6$. The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$. Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network. In the red path, the attacker still have to catch up 3 blocks to have a successful attack, which happens with probability $\rho^3$, if $p < 0.5$.}
  \label{figure:bitcoin-attack-paths}
\end{figure}
\end{proof}

Assuming that the attacker starts mining just after publishing the victim's transaction, the probability of the attacker will have found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.

\begin{theorem}
	$$\mathbf{P}(S \geq k) = 1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} (1-p)^s p^k.$$
\end{theorem}
\begin{proof}
	Let's use the following identity:
	$$\frac{1}{(1-z)^{a+1}} = \sum_{i=0}^{\infty} \binom{i+a}{i} z^i \text{, for $|z|<1$}$$

	Thus, replacing $z=1-p$, $i=s$, and $a=k-1$, we have:

	$$\frac{1}{p^{k}} = \sum_{s=0}^{\infty} \binom{s+k-1}{s} (1-p)^s$$
	$$1 = \sum_{s=0}^{\infty} \binom{s+k-1}{s} (1-p)^s p^k.$$

	Now, just split $\sum_{s=0}^{\infty} = \sum_{s=0}^{k-1} + \sum_{s=k}^{\infty}$ and it is done.
\end{proof}

Using this last theorem, we moved from an infinity sum to a finity sum.

\begin{theorem}
	Let $p = \frac{\gamma}{\beta + \gamma}$.
$$
\mathbf{P}(\text{successful attack}) =
\begin{cases}
%1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} (1-p)^s p^k \left(1 - \left( \frac{1-p}{p} \right)^{k-s} \right)
	1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} \left( (1-p)^s p^k - (1-p)^k p^s \right) \text{, $p \geq 0.5$} \\
	1 \text{, $p < 0.5$}.\\
\end{cases}
$$
\end{theorem}
\begin{proof}
$$
\mathbf{P}(\text{successful attack}) = \mathbf{P}(S \geq k) + \sum_{i=0}^{k-1} \mathbf{P}(s=i) \rho^{k-i}
$$
\end{proof}

For $k=6$, $p=0.9$, $\mathbf{P}(\text{successful attack}) = 0.0005914121600000266$.

For $k=6$, $p=0.7$, $\mathbf{P}(\text{successful attack}) = 0.15644958192000014$.

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/fig-bitcoin-attack.png}
\caption{Probability of a successful attack according to the network's hash rate of the attacker ($\beta$).\label{fig-bitcoin-attack}}
\end{figure}

%The number of blocks found in a given time interval $\Delta t$ is given by the Poisson distribution. So, the number of blocks found by the attackers $N_{\text{attacker}}$ follows a Poisson distribution with parameter $\lambda_{\text{attacker}} \Delta t = \frac{\beta \Delta t}{600}$, and the number of blocks found by the network $N_{\text{network}}$ follows a Poission distribution with parameter $\lambda_{\text{network}} \Delta t = \frac{\gamma \Delta t}{600}$. We would like to calculate $\mathbf{P}(N_{\text{attackers}} - N_{\text{network}} \geq k)$.

%$\mathbf{P}(Y_{n+k}^{\text{attacker}} < Y_{n}^{\text{network}})$

\section{Confirmation time and network capacity}
% Tempo na fila para confirmar transação

Let's say that when a new transaction is propagated it is enqueued in the unconfirmed transaction queue. Then, when a new block is found, some of these transactions in the queue are confirmed. We are interested in some measures of the queue, like the expected time to confirm a transaction and the queue's length.

Let's assume that all transactions have exactly the same size $S$ and pay exactly the same fee. If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block. Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$. The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$. Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$. The probabilities form a simple geometric series, and so they are exponentially decreasing. We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite. In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\eta$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$. Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle. When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.

The average waiting time of a transaction to be confirmed, when $m < s$, is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.

When $m \ll s$, $z_s \rightarrow 1 + 1/m$. So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$. In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.

When $\lambda_{\text{TX}} \rightarrow s$, $z_s \rightarrow 1$. So, $\mathbf{E}(\pi) \rightarrow +\infty$.

Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.

\section{Fork analysis}

% TODO BIP's and signaling bit --- see BIP 9 and BIP 135.

When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork. It is said to be a soft-fork when the rules are backward compatible, and a hard-fork when the rules are not backward compatible. In general, a hard-fork relaxes the constraints, while the latter hardens them.

Suppose that the miners' are split in two groups, $G_1$ and $G_2$, with different rules. Let's say $H_1$ and $H_2$ are their hash rate, respectively. We have two different scenarios to analyze: (i) when neither of them accepts other's blocks; and (ii) when $G_2$ accepts $G_1$'s blocks, but not the other way around.

Scenario (i) is easy to analyze, because the network would just split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.

Scenario (ii) is more trick. As $G_2$ accepts $G_1$ blocks, their hashrate will matter. If $H_1 > H_2$, then $G_2$ will frequently skip their blockchain, because $G_1$'s blockchain will be longer most of the time. If $H_1 < H_2$, then $G_2$ may have its own blockchain after a while --- a true fork. But what would happen if $H_1$ keeps increasing and eventually gets larger than $H_2$? It if happens for sufficient time, the whole $G_2$'s blockchain may be discarded in order to move to $G_1$'s blockchain when $G_1$'s gets longer.

\begin{theorem}
	When $H_1 > H_2$, the
\end{theorem}

%What would happen if, after a while, the groups agree again. How would the blockchain merge?

### 2018-04-13 Chapters/Chapter02.tex
INTRO: 1.3 seemingly complete.  1.2 and 1.1 remaining

@@
 In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everthing you think of is to get out to safety.
-You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exatcly is causing your fear.
+You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exactly is causing your fear.
 One just feels it.
@@
 
-Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
-This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present.
+Sparse Distributed Memory (SDM) \citep{Kanerva1988} (see also \citep{kanerva_sparse_1993, denning_sparse_1989} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from the subtlest of clues.
 
@@
 
-SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
+As expected for such an important thery, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
 \citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
-Despite all those applications, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
+Despite this, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
 Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate effort from researchers to build on top of previous work.
@@
 \begin{enumerate}
-    \item The original hardware implementation developed by NASA [4];
+    \item The original 1989 hardware implementation developed in NASA by \citet{flynn_sparse_1989};
 
-    \item LISP implementation for the Connection Machine [60];
+    \item a 1995 LISP implementation for the Connection Machine by \citet{turk_kanervas_1995};
 
-    \item APL implementation [59];
+    \item a 1992 APL implementation by \citet{surkan_wsdm:_1992};
 
-    \item FPGA implementation \citep{silva_reconfigurable_2004};
+    \item a 2004 FPGA implementation by \citet{silva_reconfigurable_2004};
 
-    \item C Binary Vector Symbols (CBVS): includes SDM implementation in C as a part of vector symbolic architecture[54] developed by EISLAB at Luleå University of Technology: http://pendicular.net/cbvs.php[55];
+    \item a 2005 C++ implementation by \citet{berchtold_processing_2005} from Lancaster University, in the `CommonSense ToolKit' (CSTK) \citep{noauthor_cstk:_nodate} for realtime sensor data includes SDM as one of its classification algorithms;
 
-    \item CommonSense ToolKit (CSTK) for realtime sensor data processing developed at the Lancaster University includes implementation of SDM in C++: http://cstk.sourceforge.net/[56];
+    \item  a 2015 `C Binary Vector Symbols (CBVS)':  includes SDM implementation as a part of  vector symbolic architecture developed by \citet{emruli_vector_2015} from EISLAB at Luleå University of Technology \footnote{The code is available at http://pendicular.net/cbvs.php};
 
-    \item Learning Intelligent Distribution Agent (LIDA) developed by Stan Franklin's lab at the University of Memphis includes implementation of SDM in Java.
+    \item a 2013 Java implementation `Learning Intelligent Distribution Agent' (LIDA) developed by \citep{franklin_lida:_2014, snaider_integer_2013, snaider_modular_2014} Stan Franklin from the University of Memphis includes implementation.
 \footnote{http://ccrg.cs.memphis.edu/framework.html};
@@
 
-Let us analyse these.  The Connection Machine is obsolete.  The NASA implementation is hardware-based and obsolete.  APL, while influential, is not a mainstream language in science.
+Let us analyze these.  The Connection Machine is obsolete.  The NASA implementation is hardware-based and obsolete.  APL, while reasonably influential, is not a mainstream language in science.
 
-The FPGA implementation by \citet{silva_reconfigurable_2004} has yielded a fast scan of hard locations, provided one has access to the proper hardware.
+The FPGA implementation by \citet{silva_reconfigurable_2004} has yielded a fast scan of hard locations at low energy costs, provided one has access to the proper hardware.
 Their article claims a four-fold speedup over assembly language; but it does not deal with parallel processing details.
@@
 
-Note that all these criticisms apply to the implementations in both the `common sense toolkit' and the `C Binary Vector Symbols' REF.
-While these implementations have around 150 lines of C++; at last count, the \emph{documentation of our implementation} had over 100 \emph{pages}.
-We aim a higher here, and, to us, SDM is at center stage.
+Note that all these criticisms apply to the implementations in both the `common sense toolkit' and the `C Binary Vector Symbols' \citep{berchtold_processing_2005, noauthor_cstk:_nodate, emruli_vector_2015}.
+While these implementations have around 150 lines of C++; at last count, the \emph{documentation of our implementation} had \emph{over 100 pages} \citep{linhares_sparse_2018}: they have aimed at running code, and we aim at improving a community and industry standard.
 

### 2018-04-15 Chapters/Chapter02.tex
The incredible animal behavior of Dr Linhares

@@
 
+% TODO GARY KLEIN !!!
+GARY KLEIN!!!!
+
 Suddenly, you feel the danger.
@@
 
+% TODO Venn Diagram Psychology, Neuroscience, Cognitive modelling
+% TODO Graph between these disciplines and how they might interact.
+
 Psychological
-miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue... how does this recall process work?
+orthogonality of concepts; life flashing before one's eyes; miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue...
+how does this recall process work?
 
@@
 
-Computational (fully experimental, with no voodoo-like mysterious properties)
+Cognitive modelling (fully experimental, with no voodoo-like mysterious properties)
 
@@
 
-\section{The wasted effort of duplicated work}
+\section{The wasted effort of duplicated, AD-HOC, work}
 
 As expected for such an important thery, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\citet{Linhares2011} has shown that SDM respects the limits of short-term memory discussed by \citet{Miller1955} and \citet{Cowan2001}.
+\citet{Linhares2011} has shown that SDM respects the limits of short-term memory \citep{Miller1955, Cowan2001}.
 Despite this, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
@@
 
-    \item a 2013 Java implementation `Learning Intelligent Distribution Agent' (LIDA) developed by \citep{franklin_lida:_2014, snaider_integer_2013, snaider_modular_2014} Stan Franklin from the University of Memphis includes implementation.
-\footnote{http://ccrg.cs.memphis.edu/framework.html};
+    \item a 2013 Java implementation `Learning Intelligent Distribution Agent' (LIDA) developed by \citep{franklin_lida:_2014, snaider_integer_2013, snaider_modular_2014} Stan Franklin's group from the University of Memphis includes implementation.
+\footnote{http://ccrg.cs.memphis.edu/framework.html; see also http://ccrg.cs.memphis.edu/projects.html where they link to a github repository.};
 
@@
 Then there is LIDA --- a whole cognitive architecture based on Hofstadter's Fluid Concepts, Kanerva's SDM, and other ideas \citep{Anwar2003, snaider_integer_2013, snaider_modular_2014, franklin_lida:_2014}.
-It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to Python\footnote{Python is sometimes called a `glue language'.
+It is developed in Java; which makes it difficult to connect to the lowest levels of hardware; to connect to GPUs or FPGAs, and to other languages --- at least in comparison to the combination Python and OpenCL proposed here\footnote{Python is sometimes called a `glue language'.
 That is, in my opinion, not the best metaphor.
@@
 How easy is it to program that for a particular GPU or FPGA with Java?}.
-It is not available on an open-source repository.
 It has a non-standard license, strange to the open-source community, \emph{the LIDA Framework Software NonExclusive, Non-Commercial Use License}.
+We have not found any parallelism in their code \citep{ccrg_ccrg_nodate}, which may make simulations slow or unfeasible.
 Moreover, potential contributors must sign an ``Agreement Regarding Contributory Code for the LIDA Framework Software''...
-\emph{`before Memphis can accept it'} \footnote{What they are attempting to do with this bureaucracy remains unknown, the history of computing has not been kind to those who favored centralization \citep{ferguson_computer_2002}.}.
+\emph{`before Memphis can accept it'} \footnote{What they are attempting to do with this bureaucracy remains unclear, the history of computing has not been kind to those who favored centralization \citep{ferguson_computer_2002}.
+We certainly refrain from contributing given the legal uncertainties of non-standard licenses and dubious processes --- even as we would like to link these libraries}.
 
@@
 The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
-Notice that the loss of 50\% of neurons barely affected the 256-bit SDM which remained functional even facing an enormous loss of 90\% of neurons.
+Notice that the loss of 50\% of neurons barely affected the 256-bit SDM...
+and it even remained functional when facing an enormous loss of 90\% of neurons!
+What these results seem to imply is not only that the memory is extremely robust, but that a very small expected number of neurons activated in each access radius would preserve the most valueable information.
+This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{Magic numbers}.
+Is there a thing such as excessive robustness, and, if so, when does robustness become waste?\footnote{Consider, for instance, error-correcting codes in Information Theory or the analogous computer ECC memory in Electrical Engineering: tradeoffs between robustness and waste must be considered at each design decision; Why should it be different in the modeling of human memory?}
 
@@
 
+\begin{figure}[h!]
+  \centering
+  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/info-theory/weights.png}}
+
+  \subfloat[$w(d)$ for the desired range.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region.png}}
+
+  \subfloat[Stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region-stepwise.png}}
+
+  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
+(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
+  \label{fig:info-theory-hypothesis}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Write process weighted by the amount of information contained in the distance between the written bitstring and each hard location \label{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73.png}}
+
+  \subfloat[Zoom in Figure \ref{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73-zoom.png}}
+
+  \subfloat[Behavior of weighted write operation according to the distance from the center and the number of items previously stored in the memory]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-heatmap.png}}
+
+  \caption{Behavior of the critical distance under the information-theoretic weighted write operation when $n=1,000$, $H=1,000,000$ and $r=451$.}
+  \label{fig:info-theory-experiments}
+\end{figure}
+
 My advisor, Alexandre Linhares, has proposed another write operation: an information-theoretical weighted write.
@@
 
-\begin{figure}[h!]
-  \centering
-  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/info-theory/weights.png}}
-
-  \subfloat[$w(d)$ for the desired range.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region.png}}
-
-  \subfloat[Stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region-stepwise.png}}
-
-  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
-(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
-  \label{fig:info-theory-hypothesis}
-\end{figure}
-
 The initial results of this \emph{Shannon write} operation can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
@@
 As for future research, we will execute all tests in the thesis and compare this Shannon write with the original Kanerva model.
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Write process weighted by the amount of information contained in the distance between the written bitstring and each hard location \label{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73.png}}
-
-  \subfloat[Zoom in Figure \ref{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73-zoom.png}}
-
-  \subfloat[Behavior of weighted write operation according to the distance from the center and the number of items previously stored in the memory]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-heatmap.png}}
-
-  \caption{Behavior of the critical distance under the information-theoretic weighted write operation when $n=1,000$, $H=1,000,000$ and $r=451$.}
-  \label{fig:info-theory-experiments}
-\end{figure}
 
@@
 
+\section{Magic numbers}
+
+Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.
+More recently, he suggested the use of 10,000 dimensions, and on personal discussions suggested that this should be a minimum; as he has been concerned in latent semantic analysis and seems to be the proper scale in that application.
+
+Each parameter set choice like this will lead to particular numbers --- many of them emergent ---, such as the access radius size, critical distance, memory capacity, and so forth.
+For example, the number of dimensions $n$ leads to the number of standard deviations available on the space: $2 \sqrt{n}$.
+One then must choose a proper multiple $m\times\sigma$: Kanerva (and most in the literature) have been using $m=3$, as $\mu-3\sigma$ selects approximately $1/1000$ of the space.
+These constants will determine the size of the access radius.
+Then one must choose the number of hard locations, which will determine how many hard-locations are activated on average; and so on and so forth...
+
+One intriguing question here is:  is there a `better' number of dimensions and hard locations?
+If so, can such numbers be better studied algebraically or numerically?
+How should these parameters be compared?
+What are the tradeoffs that should be considered?
+What are the `best' benchmarks possible?
+
+% Post-defense TODO: Use the robustness of neurons as a guide.  A more quantitative study of this loss of neurons... between d=0 and d=critical_distance
+% Post-defense TODO: magic numbers
+% Post-defense TODO: mu - m sigma, in which the expected number of neurons is, say, 256 (instead of >1000).
+% Post-defense TODO: study this.
+The marginal increase of 10% of Hard locations increases by how many percentage points the critical distance?
+In fact, is there a way to study this algebrically?
+That would be so cool!
+
+% \subsection{Classification with context using sequences --- for words instead of only letters}
+
+\section{Symmetrical, rapidly accessible, hard locations}
+
+A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
+Is there an algorithm that separates the area of each hard location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
+Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
+Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model may introduce non-uniform asymmetries in the space, and seems to require a staggering $O(2^n)$ hard-locations.}.
+
+Consider the hypercube with $n$ dimensions.
+We want to select a subset of its vertices with cardinality $2^{20}$ that is symmetrically distributed over the space.
+Afterward, $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
+
+A reduction from measuring the distance to $2^{20}$ hard locations to a computation of $2^{10}$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.
+At large scales of computing, the very ability to perform some experiments is a function of sheer performance.
+The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
+A little more on this in my closing words.
+
+\section{Flow versus flash:  blind driving; life flashing before one's eyes; and temperature}
+
+NOT FOR THE THESIS; move this.  Lots of blah-blah here about psychologicallymodelling these things etc.
+
+\begin{table}
+\begin{tabular}{|c|c|}
+\hline
+English & My whole life flashed before my eyes\tabularnewline
+\hline
+\hline
+Arabic & Wa mana shereet hayaty endance raayt al mawt beayny\tabularnewline
+\hline
+Dutch & Het leven aan je voorbij zien gaan\tabularnewline
+\hline
+French & J'ai vu toute ma vie défiler\tabularnewline
+\hline
+German & Mein ganzes Leben passierte Revue\tabularnewline
+\hline
+Italian & Mi é passata tutta la vita davanti agli occhi\tabularnewline
+\hline
+Latin Script & Zendegiam dar yek lahzeh az joloyeh Cheshmhayam gozasht\tabularnewline
+\hline
+Norwegian & Livet passerte (i revy) foran øynene mine\tabularnewline
+\hline
+Persian & \tabularnewline
+\hline
+Portuguese & A vida passou diante dos meus olhos\tabularnewline
+\hline
+Romanian & Mie mi-a trecut toata viata prin fata ochilor\tabularnewline
+\hline
+Russian & \tabularnewline
+\hline
+Spanish & Toda mi vida me pasó delante de los ojos\tabularnewline
+\hline
+Swedish & Livet passerade revy framför mina ögon\tabularnewline
+\hline
+\end{tabular}\caption{An expression that seems universal. Could it be that these are merely
+conscious, psychological reports, of what one feels to have experienced?
+Could this be explained by a theory of human memory?  It seems to be related to Hofstadter's temperature, here at truly high levels... Is it possible to simulate this?}
+% TODO include persian and russian back, through inkscape pdf
+% https://tex.stackexchange.com/a/370073/33832
+\end{table}\label{table:lflashing}
+
+% TODO look at previous paper on this and speculate it here in SDM, somewhere
+% TODO include on book from the following:
+%
+
+\begin{quote}
+..
+body leaning in, until he just collapsed dead before completing his little circle in the grass.
+The NVA were right behind Sergeant Jefferson, coming through the hedgerow, and Power's life flashed before his eyes.
+He saw it all in a second, and he realized then and there that everything he had ever done, every decision he had ever made, it had all been to get him in this hole on this day.
+I'm dying' right here.
+That's it.
+It's over.
+Power shot an NVA who burst through the hedgerow with his ...
+\end{quote}
+
+Consider the above and that the expression on table \ref{table:lflashing} seem to be universal...
+
+% https://books.google.com.br/books?id=OURuAAAAMAAJ&q=%22life+flashed+before+his+eyes%22+subject:%22History%22&dq=%22life+flashed+before+his+eyes%22+subject:%22History%22&hl=en&sa=X&ved=0ahUKEwjdr9aiu7faAhWHvJAKHdlGC7IQ6AEIXDAJ
+
+% https://www.amazon.com/gp/offer-listing/044021310X/ref=olp_page_1?ie=UTF8&f_all=true
+%
+
 \section{``i'' versus ``l''}
@@
 
-% \subsection{Deep learning, multiple SDMs, and the animal behavior of Dr. Linhares}
+% Post-Defense TODO:  SDM creating Voronoi Tesselations of concepts, with different sized areas for each.  This should solve i versus l.
 
-% Deep neural networks, alphaGo, alphazero
+\section{Deep learning, multiple SDMs, and the incredible animal behavior of Dr. Linhares}
 
-\section{Magic numbers}
+In the name of all that is sacred, it is probably advisable to leave the following comments as originally made by my advisor, Dr. Linhares. To quote:
 
-Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.
-More recently, he suggested the use of 10,000 dimensions, and on personal discussions suggested that this should be a minimum; as he has been concerned in latent semantic analysis and seems to be the proper scale in that application.
+\begin{figure}[h!]
+  \centering
 
-Each parameter set choice like this will lead to particular numbers --- many of them emergent ---, such as the access radius size, critical distance, memory capacity, and so forth.
-For example, the number of dimensions $n$ leads to the number of standard deviations available on the space: $2 \sqrt{n}$.
-One then must choose a proper multiple $m\times\sigma$: Kanerva (and most in the literature) have been using $m=3$, as $\mu-3\sigma$ selects approximately $1/1000$ of the space.
-These constants will determine the size of the access radius.
-Then one must choose the number of hard locations, which will determine how many hard-locations are activated on average; and so on and so forth...
+  \subfloat[Bongard problem 71]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp071.png}}
 
-One intriguing question here is:  is there a `better' number of dimensions and hard locations?
-If so, can such numbers be better studied algebraically or numerically?
-How should these parameters be compared?
-What are the tradeoffs that should be considered?
-What are the `best' benchmarks possible?
+  \subfloat[Bongard problem 72]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp072.png}}
 
-% \subsection{Classification with context using sequences --- for words instead of only letters}
+  \subfloat[Bongard problem 73]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp073.png}}
 
-\section{Symmetrical, rapidly accessible, hard locations}
+  \caption{Bongard problems 71 - 73, courtesy of noted Bongardologist, Dr. Harry Foundalis.  What distinguishes the boxes on the left hand side to those on the right hand side?}
+  \label{fig:bongard-problems-71}
+\end{figure}
 
-A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
-Is there an algorithm that separates the area of each hard location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
-Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
-Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model may introduce non-uniform asymmetries in the space, and seems to require a staggering $O(2^n)$ hard-locations.}.
+\begin{figure}[h!]
+  \centering
 
-Consider the hypercube with $n$ dimensions.
-We want to select a subset of its vertices with cardinality $2^{20}$ that is symmetrically distributed over the space.
-Afterward, $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
+  \subfloat[Bongard problem 74]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp074.png}}
 
-A reduction from measuring the distance to $2^{20}$ hard locations to a computation of $2^{10}$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.
-At large scales of computing, the very ability to perform some experiments is a function of sheer performance.
-The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
-A little more on this in my closing words.
+  \subfloat[Bongard problem 75]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp075.png}}
 
-\section{Illuminating, paving, clearing}
+  \subfloat[Bongard problem 76]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp076.png}}
+
+  \label{fig:bongard-problems-74}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+
+  \subfloat[Bongard problem 77]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp077.png}}
+
+  \subfloat[Bongard problem 78]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp078.png}}
+
+  \subfloat[Bongard problem 79]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp079.png}}
+
+  \label{fig:bongard-problems-77}
+\end{figure}
+
+\begin{quote}
+
+    Those deep neural networks...
+alphaGo...
+alphazero...
+very impressive, and to me very right and proper in a sense and also very wrong and misguided in another.
+I have not seriously studied the topic.
+Eric Nichols explained it to me in a napkin, and I watched some youtube videos and browsed the books.
+So what follows, like all great science, is completely based on gossip, hearsay, prejudice and anecdote.
+Or at least that's how I believe all great \emph{hunches} over what is promising and what is a dead-end come from; that aesthetic sense that forms over time and seems to guide us towards what feels beautiful and correct and away from what seems foolish and wasteful...
+
+    You know those animals that create those burrows underground, creatures like the prairie dogs?
+There, underground, they build all sorts of passageways from here to where they want to go; and in the process they discover all that \emph{can't be passageway}.
+What can't be passageway --- what can't be corridor --- will never be used as passageway, by definition.
+And most of the world can't be passageway.
+Think about the animal behavior: the beast won't give up!
+To me it's all very similar to what these systems seem to be doing...
+it seems they are molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...
+You have these monstrous networks creating layers upon layers in which they differentiate and distribute the errors throughout all the nooks and cranies of, say, a domain like the game of Go.
+Then, after some biblical number-crunching, it seems that there may be a very subtle way...
+however tortuous, however tight, however inprobable, however unconfortable.
+
+    And these machines are exhibiting intelligence.
+The victories in the game of Go seem to have shown that alphaGo did have vague, non-explicit, non-formal, but very strong intuitions about what the Go universe looks like.
+So here is what I like and I find beautiful:  These gigantic clusters of inter-related micro impressions that the system has acquired and can relate to.
+I imagine multiple SDMs exchanging information; gossiping about all the dead-ends out there and trying to find some tiny tortuous pathway, and it feels to me that that's a good and proper way to move forward in research.
+
+    But it feels like a mistake to say that that is thinking.
+It's not.
+It is to me a great model of someone getting so attuned into an area that they are able to simply flow through tasks; in the positive psychology sense of the word flow.
+It's very powerful and very intelligent:  but it is a form of intelligence without thinking.
+Intelligence that can't make leaps from one arena to another; Intelligence that can't work on Bongard problems.
+So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type...
+in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71, fig:bongard-problems-71, fig:bongard-problems-71}).
+
+    Cognitive Science is, to me, the study of plausible interaction between system 1 and system 2.
+We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa).
+We need both system 1 and system 2 to be plausible, in the psychological sense and in the neuroscientific sense.
+AlphaGo, for instance, has this wonderful model of the game, reflecting many of the properties of system 1...
+but then falls back into silliness as soon as it turns to the other side of the problem:  it uses some montecarlo methods in very efficient fashion to go on to to the next probing of this wonderful system 1...
+
+    Hofstadter has nailed the interactions and the content...
+analogies, fluid concepts, things seen through a different perspective, variations on a theme, etc.
+Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.
+Attaching something as beautiful as SDM to, say, some hyper-efficient alpha-beta pruning algorithm is nothing short of heresy against all the gods of science and must be punished with death without mercy.
+Seems like the most reasonable path going forward...
+
+\end{quote}
+
+\section{Not a real Conclusion}
 
@@
 
-% TODO
+% TODO  Appendix?
 

### 2018-04-15 Chapters/Chapter01.tex

@@
 
+\section{Governance}
+
+- Punishment for miners/transactions which does not comply with the rules.
+
+- Rule change policy.
+
+\section{Expected number of tips}
+
+Explain transient and stationary state. The balance happens in the stationary state.
+
+In order to estimate the balance in the number of tips, let's consider the system in which the rate of inward tips is $\lambda_\text{TX}$, while the rate of outward tips is $\rho$.
+Let $K$ be the number of tips in the system.
+Thus, after $\Delta t$ seconds, $\Delta \text{tips} = \lambda_\text{TX} \Delta t - \rho K \Delta t$.
+In the balance, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
+
+Finally, the rate of outward tips is twice the rate in which devices solve the proof-of-work of transactions. Let $H$ be the devices' hash rate, then, $\rho = 2 \cdot H \cdot 2^{-w_\text{TX}}$.
+
+Twice?! It is a number between 1 and 2, but none of them.
+
+Therefore, in the balance,
+
+$$K = \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H}$$
+
+$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$
+
+\section{Attacks \& Conflicting transactions}
+
 \chapter{Methodology}

### 2018-04-15 Chapters/Chapter02.tex
Bongard problem citations

@@
 
+highway hypnosis; absentminded driving; Going on Autopilot; PNAS study
+
 NOT FOR THE THESIS; move this.  Lots of blah-blah here about psychologicallymodelling these things etc.
@@
 
-\section{Deep learning, multiple SDMs, and the incredible animal behavior of Dr. Linhares}
+\section{Deep learning, multiple SDMs --- and the incredible animal behavior of Dr. Linhares}
 
@@
 Eric Nichols explained it to me in a napkin, and I watched some youtube videos and browsed the books.
-So what follows, like all great science, is completely based on gossip, hearsay, prejudice and anecdote.
+So what follows, like all great science, is completely based on gossip and hearsay, prejudice and anecdote.
 Or at least that's how I believe all great \emph{hunches} over what is promising and what is a dead-end come from; that aesthetic sense that forms over time and seems to guide us towards what feels beautiful and correct and away from what seems foolish and wasteful...
@@
 So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type...
-in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71, fig:bongard-problems-71, fig:bongard-problems-71}).
+in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71, fig:bongard-problems-74, fig:bongard-problems-77}; see \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013}).
 
@@
 Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.
-Attaching something as beautiful as SDM to, say, some hyper-efficient alpha-beta pruning algorithm is nothing short of heresy against all the gods of science and must be punished with death without mercy.
-Seems like the most reasonable path going forward...
+Attaching something as beautiful as SDM to, say, some hyper-efficient centralized alpha-beta pruning algorithm is nothing short of heresy against all the gods of science for which the villain must be put to death without mercy.
+Seems like the only reasonable path going forward...
 

### 2018-04-15 Chapters/sdm-autocorrelation.tex
Bongard problem citations

@@
 A hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
+Imagine the `counter's eye view':  each individual counter `likes'\footnote{If I were a cell I would try to avoid the neighborhood of a neuron.
+Neurons have serious authoritarian tendencies, violently bossing innocent cells around --- sometimes through electric shocks.} to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
 

### 2018-04-15 Chapters/Chapter02.tex
Desiderata Table

@@
 And most of the world can't be passageway.
-Think about the animal behavior: the beast won't give up!
+Think about the animal behavior: endless dead-end nooks and cranies yet the beast won't give up!
 To me it's all very similar to what these systems seem to be doing...
@@
 So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type...
-in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71, fig:bongard-problems-74, fig:bongard-problems-77}; see \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013}).
+in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71} \ref{fig:bongard-problems-74} \ref{fig:bongard-problems-77}; see \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013}).
 
-    Cognitive Science is, to me, the study of plausible interaction between system 1 and system 2.
+    Cognitive Science has to study the plausible interaction between system 1 and system 2.
 We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa).
@@
 Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.
-Attaching something as beautiful as SDM to, say, some hyper-efficient centralized alpha-beta pruning algorithm is nothing short of heresy against all the gods of science for which the villain must be put to death without mercy.
+Attaching something as beautiful as SDM to, say, some hyper-efficient centralized alpha-beta pruning algorithm is nothing short of heresy against all the gods of science for which the villain shall be put to death without clemency or mercy.
 Seems like the only reasonable path going forward...
@@
 \section{Not a real Conclusion}
+
+% Preview source code for paragraph 8
+
+\begin{table}[ht]
+\begin{tabular}{|p{4cm}|p{7cm}|}  %{|c|c|}
+\hline
+Project & Observation\tabularnewline
+\hline
+\hline
+Source-code  & C, OpenCL, Python (NVIDIA CUDA, FPGAs desirable in the future)\tabularnewline
+\hline
+Step-by-step notebook studies & replication and dissemination of studies\tabularnewline
+\hline
+Documentation of framework & 100+ pages currently\tabularnewline
+\hline
+Open-Source Creative Commons Book & Under consideration (to be created from this thesis)\tabularnewline
+\hline
+Collection of SDM links, results, resources & modeled after `Awesome Machine Learning'\citep{misiti_awesome-machine-learning:_2018} and `Awesome Artificial Intelligence'\citep{lewis_awesome-artificial-intelligence:_2018} lists, community
+managed\tabularnewline
+\hline
+Slides from the Jupyter Notebooks & online course, currently in planning phase\tabularnewline
+\hline
+Videos following the Course Slides  & online course, currently in planning phase\tabularnewline
+\hline
+\end{tabular}
+
+\caption{Desiderata of open-source, creative-common deliverables, longer-term.
+Some of these items are mature for wide dissemination, whereas others are in planning phase.
+For an online SDM course, for instance, we might partner with a Ph.D.
+program in the Teaching of Mathematics and Physics, such as \citet{im/ufrj_programa_nodate}.}
+
+\end{table}
 
@@
 Some essays completely shut the door close at the end; this one intends to leave it wide open.
+This is not a real conclusion.
 As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.

### 2018-04-17 Chapters/Chapter02.tex
details, jupyter --> talex, etc

@@
     \bigskip
-    {Pentti Kanerva's memory model was a revelation for me: \\
+    {\slshape Pentti Kanerva's memory model was a revelation for me: \\
 it was the very first piece of research I had ever come across\\
@@
 how familiar mental phenomena could be thought of as distributed\\
-patterns of micro-events, thanks to beautiful mathematics.} \\ \medskip
---- Douglas Hofstadter
+patterns of micro-events, thanks to beautiful mathematics. \\ \medskip
+--- Douglas Hofstadter}
 
@@
 
+\begin{flushright}{\slshape
+
+    \begin{quote}
+    The most exciting phrase to hear in science, the one that heralds new discoveries, is not `Eureka!' (I found it!) but ``That's funny...''
+    — Isaac Asimov
+    \end{quote}
+
+    \bigskip
+\begin{quote}
+Alexander Fleming, while explaining the staphylococcal work to Pryce, to demonstrate a point, picked up one of the top layer of the discarded plates when he noticed the lysis of the staphylococci (his cry of Eureka was ‘That’s funny!’) and then and there made a subculture of the mould.
+\end{quote}
+
+\bigskip
+\bigskip
+
+}
+\end{flushright}
+
 Sparse Distributed Memory is a viable model of human memory, yet it does require researchers to (re-)implement a number of parallel algorithms in different architectures.
@@
 Here are interesting questions that have been considered during this work, but have had to be left for future research.
+
+\section{Another `funny thing'...}
+
+Over the course of this work we looked into a `funny thing' that had appeared in my master's thesis:  the graphs of convergence did not reflect the original expectations put forth by Kanerva, and neither my own previous calculations.
+These numerical results pointed us back to the math and to some interesting random variables not studied before.
+
+Another funny thing that has appeared in my master's thesis was that the critical distance collapsed when the network was exclusively writing `random-at-x'.
+This is something I would like to look into but have not had the time to conduct a proper study yet.
 
@@
 The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
-A little more on this in my closing words.
-
-\section{Flow versus flash:  blind driving; life flashing before one's eyes; and temperature}
-
-highway hypnosis; absentminded driving; Going on Autopilot; PNAS study
-
-NOT FOR THE THESIS; move this.  Lots of blah-blah here about psychologicallymodelling these things etc.
-
-\begin{table}
-\begin{tabular}{|c|c|}
-\hline
-English & My whole life flashed before my eyes\tabularnewline
-\hline
-\hline
-Arabic & Wa mana shereet hayaty endance raayt al mawt beayny\tabularnewline
-\hline
-Dutch & Het leven aan je voorbij zien gaan\tabularnewline
-\hline
-French & J'ai vu toute ma vie défiler\tabularnewline
-\hline
-German & Mein ganzes Leben passierte Revue\tabularnewline
-\hline
-Italian & Mi é passata tutta la vita davanti agli occhi\tabularnewline
-\hline
-Latin Script & Zendegiam dar yek lahzeh az joloyeh Cheshmhayam gozasht\tabularnewline
-\hline
-Norwegian & Livet passerte (i revy) foran øynene mine\tabularnewline
-\hline
-Persian & \tabularnewline
-\hline
-Portuguese & A vida passou diante dos meus olhos\tabularnewline
-\hline
-Romanian & Mie mi-a trecut toata viata prin fata ochilor\tabularnewline
-\hline
-Russian & \tabularnewline
-\hline
-Spanish & Toda mi vida me pasó delante de los ojos\tabularnewline
-\hline
-Swedish & Livet passerade revy framför mina ögon\tabularnewline
-\hline
-\end{tabular}\caption{An expression that seems universal. Could it be that these are merely
-conscious, psychological reports, of what one feels to have experienced?
-Could this be explained by a theory of human memory?  It seems to be related to Hofstadter's temperature, here at truly high levels... Is it possible to simulate this?}
-% TODO include persian and russian back, through inkscape pdf
-% https://tex.stackexchange.com/a/370073/33832
-\end{table}\label{table:lflashing}
-
-% TODO look at previous paper on this and speculate it here in SDM, somewhere
-% TODO include on book from the following:
-%
-
-\begin{quote}
-..
-body leaning in, until he just collapsed dead before completing his little circle in the grass.
-The NVA were right behind Sergeant Jefferson, coming through the hedgerow, and Power's life flashed before his eyes.
-He saw it all in a second, and he realized then and there that everything he had ever done, every decision he had ever made, it had all been to get him in this hole on this day.
-I'm dying' right here.
-That's it.
-It's over.
-Power shot an NVA who burst through the hedgerow with his ...
-\end{quote}
-
-Consider the above and that the expression on table \ref{table:lflashing} seem to be universal...
-
-% https://books.google.com.br/books?id=OURuAAAAMAAJ&q=%22life+flashed+before+his+eyes%22+subject:%22History%22&dq=%22life+flashed+before+his+eyes%22+subject:%22History%22&hl=en&sa=X&ved=0ahUKEwjdr9aiu7faAhWHvJAKHdlGC7IQ6AEIXDAJ
-
-% https://www.amazon.com/gp/offer-listing/044021310X/ref=olp_page_1?ie=UTF8&f_all=true
-%
+A little more on this will follow below.
 
@@
 
-In the name of all that is sacred, it is probably advisable to leave the following comments as originally made by my advisor, Dr. Linhares. To quote:
+There is the intention of studying what, if any, capabilities do multiple SDMs have...
+yet it is probably advisable to leave the following comments as originally made by my advisor, Dr.
+Linhares.
+To quote:
+
+\begin{table}[ht]
+\begin{tabular}{|p{4cm}|p{7cm}|}  %{|c|c|}
+\hline
+System 1	& System 2\tabularnewline
+\hline
+\hline
+Unconscious Reasoning	& Conscious Reasoning\tabularnewline
+\hline
+Implicit	& Explicit\tabularnewline
+\hline
+Automatic	& Controlled\tabularnewline
+\hline
+Low Effort	& High Effort\tabularnewline
+\hline
+Large Capacity	& Small Capacity\tabularnewline
+\hline
+Rapid	& Slow\tabularnewline
+\hline
+Default Process	& Inhibitory\tabularnewline
+\hline
+Associative	& Rule-Based\tabularnewline
+\hline
+Contextualized	& Abstract\tabularnewline
+\hline
+Domain Specific	& Domain General\tabularnewline
+\hline
+Evolutionarily Old	& Evolutionarily Recent\tabularnewline
+\hline
+Nonverbal	& Linked to language\tabularnewline
+\hline
+Includes recognition, perception, orientation	& Includes rule following, comparisons, weighing of options\tabularnewline
+\hline
+Modular Cognition	& Fluid Intelligence\tabularnewline
+\hline
+Independent of working memory	& Limited by working memory capacity\tabularnewline
+\hline
+Non-Logical	& Logical\tabularnewline
+\hline
+Parallel	& Serial\tabularnewline
+\hline
+\end{tabular}
+
+\caption{Dual process theories.}
+
+\end{table}\label{table:twosys}
 
@@
 
+    So here is my hunch:  \emph{SDM is System 1 and System 1 is SDM} (e.g., as in Table \ref{table:twosys}).
+I haven't seen anyone say this out loud, but Kanerva has pretty much nailed system 1 --- I feel he is at least 90\% right.
+
     You know those animals that create those burrows underground, creatures like the prairie dogs?
@@
 To me it's all very similar to what these systems seem to be doing...
-it seems they are molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...
+it seems they are crafting and molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...
 You have these monstrous networks creating layers upon layers in which they differentiate and distribute the errors throughout all the nooks and cranies of, say, a domain like the game of Go.
 Then, after some biblical number-crunching, it seems that there may be a very subtle way...
-however tortuous, however tight, however inprobable, however unconfortable.
+however tortuous, however tight, however improbable, however unconfortable.
 
     And these machines are exhibiting intelligence.
-The victories in the game of Go seem to have shown that alphaGo did have vague, non-explicit, non-formal, but very strong intuitions about what the Go universe looks like.
+The victories in the game of Go seem to have shown that alphaGo did have vague, non-explicit, non-formal, but very strong intuitions about what the nooks and cranies of the Go universe looks like.
 So here is what I like and I find beautiful:  These gigantic clusters of inter-related micro impressions that the system has acquired and can relate to.
-I imagine multiple SDMs exchanging information; gossiping about all the dead-ends out there and trying to find some tiny tortuous pathway, and it feels to me that that's a good and proper way to move forward in research.
+I imagine multiple SDMs exchanging information; gossiping to each other about all the dead-ends out there and trying to find some tiny tortuous pathway, and it feels to me that that's a good and proper way to move forward in research.
 
@@
 Intelligence that can't make leaps from one arena to another; Intelligence that can't work on Bongard problems.
-So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type...
-in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71} \ref{fig:bongard-problems-74} \ref{fig:bongard-problems-77}; see \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013}).
+So what I would fear as a dead-end in research would be to have another AI winter...
+decades of massive, massive number-crunching and machine learning of this type...
+in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figure \ref{fig:bongard-problems-71}).
+Bongard problems demand high-gossip between system 1 and system 2; they demand a high dose of imagination and mental imagery manipulation; they demand discrimination and demand clusterization, among many other things that have been spelled out before by \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013} and others.
 
-    Cognitive Science has to study the plausible interaction between system 1 and system 2.
+    Cognitive Science has to study the \emph{plausible} interaction between system 1 and system 2.
 We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa).
@@
 AlphaGo, for instance, has this wonderful model of the game, reflecting many of the properties of system 1...
-but then falls back into silliness as soon as it turns to the other side of the problem:  it uses some montecarlo methods in very efficient fashion to go on to to the next probing of this wonderful system 1...
+but then falls back into silliness as soon as it turns to the other side of the problem:  it uses montecarlo methods in very efficient fashion to probe some 70,000 thousands of positions per second and go on to to the next probing of this wonderful system 1...
+
+    It works; it is impressive; it is intelligent; but it is not thinking; and it hasn't thought us much about our cognition.
 
     Hofstadter has nailed the interactions and the content...
-analogies, fluid concepts, things seen through a different perspective, variations on a theme, etc.
+analogies, fluid concepts, things seen through a different perspective, abstract roles, variations on a theme, etc.
 Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.
-Attaching something as beautiful as SDM to, say, some hyper-efficient centralized alpha-beta pruning algorithm is nothing short of heresy against all the gods of science for which the villain shall be put to death without clemency or mercy.
+Isn't it time to work on System 2 models that are counterparts to SDM --- in fact, isn't it obvious that that's our glaring omission?
+Attaching something as beautiful as SDM to, say, some centralized hyper-efficient montecarlo algorithm is nothing short of heresy against all the gods of science for which the villain shall be put to death without clemency or mercy.
 Seems like the only reasonable path going forward...
@@
 For an online SDM course, for instance, we might partner with a Ph.D.
-program in the Teaching of Mathematics and Physics, such as \citet{im/ufrj_programa_nodate}.}
+program in the Teaching of Mathematics and Physics, such as \citet{im/ufrj_programa_nodate}, who are intent on developing and measuring the effects of new educational tools.}
 
@@
 Indeed, a high-quality codebase seems to have become a scientific community's form of unequivocally standing behind a consensus.
-For example, the journal Nature analyzed the top-100 cited papers in history, to find:
+For example, the journal \emph{Nature} analyzed the top-100 cited papers in history, to find:
 
@@
 
-It is no coincidence that scientific journals such as BMC Neuroscience, or the Journal of Machine Learning Research have specific sections on open-source software.
-The journal Neurocomputing states, bluntly: “software is scientific method by machine”.
+It is no coincidence that scientific journals such as \emph{BMC Neuroscience}, or the \emph{Journal of Machine Learning Research} have specific sections on open-source software.
+The journal \emph{Neurocomputing} states, bluntly: “software is scientific method by machine”.
 
@@
 Some essays completely shut the door close at the end; this one intends to leave it wide open.
-This is not a real conclusion.
 As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.
-
-% TODO  Appendix?
 

### 2018-04-17 SDM_the_search_for_powers_of_2_everywhere.tex
details, jupyter --> talex, etc

--- 2018-01-22 Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7.3.tex
@@
     % Document parameters
-    \title{Kanerva-Table-7.3}
+    \title{SDM\_the\_search\_for\_powers\_of\_2\_everywhere}
 
@@
 
+\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
+\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
+\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
+\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
+\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
+\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
+\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
+\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
+\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
 \expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
+\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
+\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
+\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
+\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
+\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
 \expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
-\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
-\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
+\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
+\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
+\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
 \expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
@@
 \expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
+\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
+\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
 \expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
+\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
+\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
 \expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
-\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
-\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
-\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
-\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
-\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
-\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
-\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
-\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
-\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
-\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
-\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
-\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
-\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
-\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
-\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
-\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
-\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
 
@@
 
+    \maketitle
+
+    \section{Sparse Distributed Memory: How many
+subspaces?}\label{sparse-distributed-memory-how-many-subspaces}
+
+In his original analysis, Kanerva uses... ...1,000 dimensions
+...1,000,000 circles ...with radius=451 bits so that a random bistring
+will be stored in aproximately 1,000 hard locations.
+
+A first observation here is that powers of 10 are not well suited,
+either for the mathematical analysis of the space, or for computer
+science. Let us play a little with numbers here, starting from Kanerva's
+parameters. Our goal will be to find how many circles we should use and
+the radius of a circle. Can we find some optimum set of parameters?
+
     \begin{Verbatim}[commandchars=\\\{\}]
-{\color{incolor}In [{\color{incolor}2}]:} \PY{k+kn}{import} \PY{n+nn}{sdm} \PY{k+kn}{as} \PY{n+nn}{sdmlib}
-        \PY{k+kn}{import} \PY{n+nn}{matplotlib.pyplot} \PY{k+kn}{as} \PY{n+nn}{plt}
-        \PY{o}{\PYZpc{}}\PY{k}{matplotlib} inline
+{\color{incolor}In [{\color{incolor}3}]:} \PY{n}{n} \PY{o}{=} \PY{l+m+mi}{1000} \PY{c+c1}{\PYZsh{} number of dimensions}
+        \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{number of dimensions=}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}\PY{n}{n}\PY{p}{)}
 \end{Verbatim}
@@
     \begin{Verbatim}[commandchars=\\\{\}]
-{\color{incolor}In [{\color{incolor}3}]:} \PY{k}{def} \PY{n+nf}{gen\PYZus{}sdm}\PY{p}{(}\PY{n}{scanner\PYZus{}type}\PY{p}{,} \PY{n}{bits}\PY{o}{=}\PY{l+m+mi}{1000}\PY{p}{,} \PY{n}{sample}\PY{o}{=}\PY{l+m+mi}{1000000}\PY{p}{)}\PY{p}{:}
-            \PY{n}{address\PYZus{}space} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{AddressSpace}\PY{o}{.}\PY{n}{init\PYZus{}from\PYZus{}b64\PYZus{}file}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{sdm\PYZhy{}10000w.as}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-            \PY{n}{counter} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{Counter}\PY{o}{.}\PY{n}{load\PYZus{}file}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{sdm\PYZhy{}10000w}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-            \PY{n}{sdm} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{SDM}\PY{p}{(}\PY{n}{address\PYZus{}space}\PY{p}{,} \PY{n}{counter}\PY{p}{,} \PY{l+m+mi}{451}\PY{p}{,} \PY{n}{scanner\PYZus{}type}\PY{p}{)}
-            \PY{k}{return} \PY{n}{sdm}
-\end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-{\color{incolor}In [{\color{incolor}4}]:} \PY{n}{sdm} \PY{o}{=} \PY{n}{gen\PYZus{}sdm}\PY{p}{(}\PY{n}{sdmlib}\PY{o}{.}\PY{n}{SDM\PYZus{}SCANNER\PYZus{}OPENCL}\PY{p}{)}
-\end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-{\color{incolor}In [{\color{incolor}5}]:} \PY{n}{b} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{Bitstring}\PY{o}{.}\PY{n}{init\PYZus{}random}\PY{p}{(}\PY{l+m+mi}{1000}\PY{p}{)}
-        \PY{n}{sdm}\PY{o}{.}\PY{n}{write}\PY{p}{(}\PY{n}{b}\PY{p}{,} \PY{n}{b}\PY{p}{)}
-\end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-{\color{incolor}In [{\color{incolor}6}]:} \PY{k+kn}{from} \PY{n+nn}{IPython.display} \PY{k+kn}{import} \PY{n}{clear\PYZus{}output}
-
-        \PY{n}{distances} \PY{o}{=} \PY{p}{[}\PY{p}{]}
-        \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1001}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{)}
-        \PY{k}{for} \PY{n}{i}\PY{p}{,} \PY{n}{dist} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{:}
-            \PY{n}{clear\PYZus{}output}\PY{p}{(}\PY{n}{wait}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
-            \PY{k}{print} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Distance: \PYZob{}:4d\PYZcb{} (\PYZob{}:.2f\PYZcb{}}\PY{l+s+s1}{\PYZpc{}}\PY{l+s+s1}{)}\PY{l+s+s1}{\PYZsq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{dist}\PY{p}{,} \PY{l+m+mf}{100.}\PY{o}{*}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{o}{/}\PY{n+nb}{len}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{)}
-            \PY{n}{c} \PY{o}{=} \PY{n}{sdmlib}\PY{o}{.}\PY{n}{Bitstring}\PY{o}{.}\PY{n}{init\PYZus{}from\PYZus{}bitstring}\PY{p}{(}\PY{n}{b}\PY{p}{)}
-            \PY{n}{c}\PY{o}{.}\PY{n}{flip\PYZus{}random\PYZus{}bits}\PY{p}{(}\PY{n}{dist}\PY{p}{)}
-            \PY{k}{assert} \PY{n}{c}\PY{o}{.}\PY{n}{distance\PYZus{}to}\PY{p}{(}\PY{n}{b}\PY{p}{)} \PY{o}{==} \PY{n}{dist}
-            \PY{n}{d} \PY{o}{=} \PY{n}{sdm}\PY{o}{.}\PY{n}{read}\PY{p}{(}\PY{n}{c}\PY{p}{)}
-            \PY{n}{distances}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{d}\PY{o}{.}\PY{n}{distance\PYZus{}to}\PY{p}{(}\PY{n}{b}\PY{p}{)}\PY{p}{)}
-        \PY{k}{print} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Done!}\PY{l+s+s1}{\PYZsq{}}
-\end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-Distance: 1000 (100.00\%)
-Done!
+number of dimensions= 1000
 
@@
     \begin{Verbatim}[commandchars=\\\{\}]
-{\color{incolor}In [{\color{incolor}29}]:} \PY{n}{plt}\PY{o}{.}\PY{n}{figure}\PY{p}{(}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{distances}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{x}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{p}{[}\PY{l+m+mi}{500}\PY{p}{]}\PY{o}{*}\PY{n+nb}{len}\PY{p}{(}\PY{n}{x}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{k\PYZhy{}\PYZhy{}}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Kanerva}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s+s1}{s Figure 7.3}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{New distance (after one read)}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Old distance}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
-         \PY{n}{plt}\PY{o}{.}\PY{n}{axis}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1000}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1000}\PY{p}{]}\PY{p}{)}\PY{p}{;}
+{\color{incolor}In [{\color{incolor}4}]:} \PY{k}{def} \PY{n+nf}{mu}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{:}
+            \PY{k}{return} \PY{n}{n}\PY{o}{/}\PY{l+m+mi}{2}
+
+        \PY{k+kn}{import} \PY{n+nn}{math}
+
+        \PY{k}{def} \PY{n+nf}{sigma}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{:}
+            \PY{k}{return} \PY{n}{math}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{o}{/}\PY{l+m+mi}{2}
+
+        \PY{k}{def} \PY{n+nf}{percentage\PYZus{}of\PYZus{}n}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{:}
+            \PY{k}{return} \PY{p}{(}\PY{l+m+mi}{100}\PY{o}{*}\PY{p}{(}\PY{l+m+mi}{6}\PY{o}{*}\PY{n}{sigma}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{)}\PY{p}{)} \PY{o}{/} \PY{n}{n}
 \end{Verbatim}
 
-    \begin{center}
-    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_5_0.png}
-    \end{center}
-    { \hspace*{\fill} \\}
+    \begin{Verbatim}[commandchars=\\\{\}]
+{\color{incolor}In [{\color{incolor}5}]:} \PY{k}{def} \PY{n+nf}{analysis}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{:}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{************ ANALYSIS FOR n=}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}\PY{n}{n}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{orthoghonal distance=}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}\PY{n}{mu}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{standard deviation=}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,}\PY{n}{sigma}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{the space offers }\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n+nb}{str}\PY{p}{(}\PY{n}{n}\PY{o}{/}\PY{o}{/}\PY{n}{sigma}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ standard deviations.}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{3 sigma=}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+m+mf}{3.09}\PY{o}{*}\PY{n}{sigma}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{the distance from a pole to the equator=}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{math}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{standard deviations}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{the percentage of n (in relation to the range R Bits) is}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{percentage\PYZus{}of\PYZus{}n}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{)}\PY{c+c1}{\PYZsh{} When using mu \PYZhy{} 3 sigma...}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Because 3 standard deviations contain \PYZti{}1/1000 of a normal distribution, }\PY{l+s+s1}{\PYZsq{}} \PYZbs{}
+                    \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{approximately one in a thousand items will be found in a circle of}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PYZbs{}
+                    \PY{n}{mu}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{o}{\PYZhy{}}\PY{l+m+mf}{3.09}\PY{o}{*}\PY{n}{sigma}\PY{p}{(}\PY{n}{n}\PY{p}{)}\PY{p}{,}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ radius}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
+            \PY{n+nb}{print} \PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{...........................................}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
+\end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+{\color{incolor}In [{\color{incolor}6}]:} \PY{k}{for} \PY{n}{n} \PY{o+ow}{in} \PY{p}{[}\PY{l+m+mi}{100}\PY{p}{,} \PY{l+m+mi}{1000}\PY{p}{,} \PY{l+m+mi}{10000}\PY{p}{,} \PY{l+m+mi}{256}\PY{p}{,} \PY{l+m+mi}{1024}\PY{p}{]}\PY{p}{:}
+            \PY{n}{analysis}\PY{p}{(}\PY{n}{n}\PY{p}{)}
+\end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+************ ANALYSIS FOR n= 100
+orthoghonal distance= 50.0
+standard deviation= 5.0
+the space offers  20.0  standard deviations.
+3 sigma= 15.45
+the distance from a pole to the equator= 10.0 standard deviations
+the percentage of n (in relation to the range R Bits) is 30.0
+Because 3 standard deviations contain \textasciitilde{}1/1000 of a normal distribution, approximately one in a thousand items will be found in a circle of 34.55  radius
+{\ldots}
+
+************ ANALYSIS FOR n= 1000
+orthoghonal distance= 500.0
+standard deviation= 15.811388300841896
+the space offers  63.0  standard deviations.
+3 sigma= 48.857189849601454
+the distance from a pole to the equator= 31.622776601683793 standard deviations
+the percentage of n (in relation to the range R Bits) is 9.486832980505138
+Because 3 standard deviations contain \textasciitilde{}1/1000 of a normal distribution, approximately one in a thousand items will be found in a circle of 451.14281015039853  radius
+{\ldots}
+
+************ ANALYSIS FOR n= 10000
+orthoghonal distance= 5000.0
+standard deviation= 50.0
+the space offers  200.0  standard deviations.
+3 sigma= 154.5
+the distance from a pole to the equator= 100.0 standard deviations
+the percentage of n (in relation to the range R Bits) is 3.0
+Because 3 standard deviations contain \textasciitilde{}1/1000 of a normal distribution, approximately one in a thousand items will be found in a circle of 4845.5  radius
+{\ldots}
+
+************ ANALYSIS FOR n= 256
+orthoghonal distance= 128.0
+standard deviation= 8.0
+the space offers  32.0  standard deviations.
+3 sigma= 24.72
+the distance from a pole to the equator= 16.0 standard deviations
+the percentage of n (in relation to the range R Bits) is 18.75
+Because 3 standard deviations contain \textasciitilde{}1/1000 of a normal distribution, approximately one in a thousand items will be found in a circle of 103.28  radius
+{\ldots}
+
+************ ANALYSIS FOR n= 1024
+orthoghonal distance= 512.0
+standard deviation= 16.0
+the space offers  64.0  standard deviations.
+3 sigma= 49.44
+the distance from a pole to the equator= 32.0 standard deviations
+the percentage of n (in relation to the range R Bits) is 9.375
+Because 3 standard deviations contain \textasciitilde{}1/1000 of a normal distribution, approximately one in a thousand items will be found in a circle of 462.56  radius
+{\ldots}
+
+    \end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+{\color{incolor}In [{\color{incolor}86}]:} \PY{n}{print\PYZus{}analysis}\PY{p}{(}\PY{l+m+mi}{1024}\PY{p}{)}
+\end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+************ ANALYSIS FOR n= 1024
+orthoghonal distance= 512.0
+standard deviation= 16.0
+3 sigma= 48.0
+the distance from a pole to the equator= 32.0 standard deviations
+the percentage of n (in relation to the range R Bits) is 9.375
+Because 3 standard deviations contain \textasciitilde{}1/1000 of a normal distribution, approximately one in a thousand items will be found in a circle of 464.0  radius
+\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#
+
+    \end{Verbatim}
+
+    Let us focus on the word \emph{aproximately}. The estimation is given by
+usual tables with metrics of the normal curve.
+
+The \emph{unit} of analysis provided is that of the standard deviation.
+
+Our intention is to make the bit the unit of analysis, and estimate the
+statistics of the system in a more precise form.
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+{\color{incolor}In [{\color{incolor} }]:}
+\end{Verbatim}
 

### 2018-04-17 Chapters/Introduction.tex
Intro, references after Plous93

@@
 This thesis will explore three different topics related to \emph{computing business platforms} (Figure \ref{fig:onelife}).
-Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three specific, self-contained, scientific papers.
-The first of which studies the possibility of distributed financial ledgers.
+Though the range of the topics is large, as it usually is in management science, it is my hope to convince readers of the value of this doctoral thesis brought by three specific, self-contained, scientific papers --- the first of which studies the possibility of distributed financial ledgers.
 
@@
 
-A second topic that may have an outsized influence on business and that we will be taking a closer look is a model of artificial intelligence.
+A second topic that may have an outsized influence on business and that we will be taking a closer look is a model of artificial intelligence --- and of human decision making.
 
-\section{Artificial intelligence}
+\section{Sparse Distributed Memory}
+
+A significant part of the thesis is devoted to \emph{Sparse Distributed Memory}, or SDM for short.
+There are two reasons for that: first, it is an important topic in artificial intelligence and computational cognitive science --- a crucial piece of technology that may eventually have vast influence throughout society.
+
+But it is perhaps more than that.
+Perhaps it is a crucial item in understanding human decision making.
+SDM may, perhaps, bring us a formal model of \emph{recognition-primed decision}, put forth by Gary Klein as a theory of decision-making.
+
+\subsection{Decisions with serious skin in the game}
+
+The recent field of \emph{naturalistic decision making} stands as a new alternative model of study of decision-making.
+It bears contrast to both the classical model of rational choice and to the program of heuristics and biases.
+
+The classical model of \emph{rational choice} and optimization which has been the basis of studies in economics (for instance, in game theory), management science and operations research (in for example mathematical programming models), and in artificial intelligence (the symbols and search paradigm), proposes a set of standard, quantitative, methods in order to `rationally' select a choice.
+Under this theory, the decision-maker:
+
+\begin{enumerate}
+	\item identifies a set of options,
+	\item identifies ways of evaluating these options,
+	\item weighs each evaluation dimension,
+	\item calculates a rating of each option, and, finally,
+	\item selects the one with the maximum score.
+\end{enumerate}
+
+Note that this model implies that, in order to have a number of choices to choose from, one must first have (i) perceived a problem, and (ii) perceived a set of alternative choices --- where do the choices come from?
+What psychological process brings forth their emergence?
+The rational model will deal only with the phase of (iii) selecting one choice from the set.
+
+Despite its widespread use in a number of distinct areas, the rational choice model has not found to be psychologically plausible, for a number of reasons \citep{plous_psychology_1993}.
+One of the reasons is that the chosen alternative depends on how decision-makers initially frame a problem (Kahneman and Tversky, 1979).
+There has been strong criticism of the rational choice model from the heuristics and biases research program, in which problems are carefully devised to show that one's intuitions generally depart from the expected optima, and are generally inconsistent with what would be expected as rational.
+A large number of biases that depart from rational choice have been found (see, for instance, \citet{plous_psychology_1993}), placing strain on the traditional rational actor doctrine.
+Yet the heuristics and biases studies are concentrated on carefully devised questionnaires applied mostly to undergraduate students --- not on real world settings with serious skin in the game.
+
+Thus a new field of naturalistic decision-making emerged, in which the focus is centered around real life settings and decisions being made under rapidly changing circumstances.
+A number of studies have been conducted, from firefighters to nurses to chess players to military personnel.
+One of the most interesting theories to emerge from naturalistic decision-making, the recognition-primed decision model, was devised by Gary Klein and his colleagues (Klein, 1999).
+
+\subsubsection{Recognition-primed decision}
+
+Consider the following cases:
+
+EXAMPLE \#1.
+The Cuban World Chess Champion José Raoul Capablanca once remarked about his personal, subjective, experience: `I know at sight what a position contains.
+What could happen?
+What is going to happen?
+You figure it out, I know it!' In another occasion, talking about the numerous possibilities that less-skilled players usually consider on each board position, he bluntly remarked: `I see only one move: The best one.'  Perhaps the reader may think that Capablanca was quite simply being arrogant.
+But there is evidence to the contrary, that expert decision-makers actually are biased towards very high quality choices.
+We believe that, in fact, Capablanca was telling us an important fact about expert human psychology and decision-making, which would later be documented in recognition-primed decision studies.
+
+EXAMPLE \#2.
+A baby at an infirmary suddenly turns blue.
+Within seconds, a nurse has a diagnosis and a potential action.
+In this case, the nurse thinks the baby has a pneumopericardium, which means the sac surrounding the baby's heart is inflated with air, and the resulting pressure detracts from the heart's pumping of blood.
+There is a problem with this diagnosis, though.
+The electrocardiogram is showing a healthy 80 beats per minute.
+If nothing is done, the baby will die within a few minutes.
+The doctor walks into the room to find the nurse screaming for silence and listening to the baby's heart with an stethoscope.
+She is now sure of her diagnosis, and she gives the doctor a syringe: ``stick the heart, it's a pneumopericardium, I know it''.
+Given the electrocardiogram, other nurses are skeptical, until the x-ray operator screams out: ``she's right!''  Her intuitive diagnosis ultimately saves the baby's life.
+
+Klein (1999) conducted a series of studies with decision-makers under rapidly changing scenarios.
+During interviews, when questioned how a specific decision (or course of action) was adopted, decision-makers such as the nurse would proclaim, to Klein's frustration, that they `did not make decisions'.
+One experienced firefighter proclaimed `I don't make decisions--I don't remember when I've ever made a decision' (Klein, 1999, p.10).
+Decision-makers did not seem to be comparing alternative courses of actions, as classical models would predict.
+`It is usually obvious what to do in any given situation' (p.11).
+Repeated statements of the sort by different decision-makers led Klein to propose a psychologically plausible model of decision-making which radically departed from the established view of `comparing alternatives and selecting the optimum'.
+
+Klein (1999) proposed a model of recognition-primed decision, in which experienced decision-makers would find themselves immersed in complex situations and rapidly take adequate courses of action.
+Decision-makers would rapidly perceive cues from any situation and retrieve from episodic memory similar situations (Tulving 1983), which would bring assessments and diagnoses and plausible courses of action.
+Because priming mechanisms are automatic and unconscious (Bargh and Chartrand 1999, Bargh et al.
+2001), these decision-makers reported doing `the obvious' action in different situations.
+This `obvious' course of action, Klein proposes, is brought from long-term episodic memory by priming mechanisms.
+Hence, decision-makers would not be selecting among distinct alternatives, but rather simply performing the automatically-provided action.
+
+Even if the `obvious' action seemed plausible for a theory, another problem remained:  if decision-makers did not compare alternatives, then how could they know that a course of action was good?
+In subsequent interviews, evidence emerged that decision-makers would be using the simulation heuristic, proposed by Kahneman and Tversky (1982).
+That is, facing a particular situation, experienced decision-makers would be primed towards a particular course of action, to the detriment of most alternative courses of action.
+This primed alternative would be `simulated', or `run through', one's mind, and, if found acceptable during the simulation processing, would be acted upon without further deliberation.
+If problems emerged during mental simulation, another different course of action would be primed.
+Thus was born a theory of intuitive decision-making, in which experienced people would not be selecting choices from a vast set of alternatives, but instead `testing' their initially primed predispositions with a simulation heuristic.
+
+This model, of course, applied only to expert decision-makers with years of experience.
+It involves access to a large episodic memory in order to rapidly retrieve a suitable course of action.
+This was initially found surprising by Klein:
+
+\begin{quote}
+	Before we did this study, we believed the novices impulsively jumped at the first option they could think of, whereas experts carefully deliberated about the merits of different courses of action.
+Now it seemed that it was the experts who could generate a single course of action, while novices needed to compare different approaches.
+--- Klein (1999, p.21)
+\end{quote}
+
+Because priming mechanisms that brought plausible actions to mind are unconscious, people would report having ``done the obvious thing to do''.
+Decision-makers would be unable to visualize the cognitive processes underlying their decisions, and would in many cases even believe that they had skills of the `fantastic' variety:  One firefighter demands that his whole crew abandon operations inside a house, just to see it collapse seconds afterward.
+A radar operator would `chill' after spotting a new track, and would fire counter missiles against it, based on the `feeling' that it was a hostile missile.
+It took over a year for this radar operator, after being interviewed by Klein, to understand the incredibly subtle cues that he was responding to whenever he perceived the new radar track.
+Unable to reasonably explain their life-saving, rapid, decisions, both the firefighter and the radar operator thought that they had ESP or other fantastic abilities.
+Careful probing would show that they were able to unconsciously perceive subtle cues, which primed them towards adequate responses.
+
+Beyond a formalization of this process of memory recall, Sparse Distributed Memory will offer us a plausible, both psychologically and neuroscientifically, path towards artificial intelligence.
+
+\subsection{Artificial Intelligence}
+
 % Technology > Computer Science > Artificial Intelligence > Pattern Recognition > SDM.
@@
 
-We implement a RB-Complete\footnote{`Ridiculously Buzzword Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from previous theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
+We implement a BW-Complete\footnote{`BuzzWord-Complete: the model is (i) Open-Source, (ii) Cross-Platform; (iii) highly parallel; (iv) able to execute on CPUs and/or GPUs; (v) it can be run on the `cloud'; etc.} SDM framework that not only shows small discrepancies from previous theoretical expectations, but also may be of use to other researchers interested in testing their own hypotheses and theories of SDM.
 The computer code has been used in a previous Ph.D.

### 2018-04-17 Chapters/Introduction.tex
Intro with refs...

@@
 Despite its widespread use in a number of distinct areas, the rational choice model has not found to be psychologically plausible, for a number of reasons \citep{plous_psychology_1993}.
-One of the reasons is that the chosen alternative depends on how decision-makers initially frame a problem (Kahneman and Tversky, 1979).
+One of the reasons is that the chosen alternative depends on how decision-makers initially frame a problem \citep{kahneman1979prospect}.
 There has been strong criticism of the rational choice model from the heuristics and biases research program, in which problems are carefully devised to show that one's intuitions generally depart from the expected optima, and are generally inconsistent with what would be expected as rational.
@@
 A number of studies have been conducted, from firefighters to nurses to chess players to military personnel.
-One of the most interesting theories to emerge from naturalistic decision-making, the recognition-primed decision model, was devised by Gary Klein and his colleagues (Klein, 1999).
+One of the most interesting theories to emerge from naturalistic decision-making, the recognition-primed decision model, was devised by \citet{klein1999sources} and his colleagues.
 
@@
 
-Klein (1999) conducted a series of studies with decision-makers under rapidly changing scenarios.
+\citet{klein1999sources} conducted a series of studies with decision-makers under rapidly changing scenarios.
 During interviews, when questioned how a specific decision (or course of action) was adopted, decision-makers such as the nurse would proclaim, to Klein's frustration, that they `did not make decisions'.
-One experienced firefighter proclaimed `I don't make decisions--I don't remember when I've ever made a decision' (Klein, 1999, p.10).
+One experienced firefighter proclaimed `I don't make decisions--I don't remember when I've ever made a decision' \citep{klein1999sources}, p.10.
 Decision-makers did not seem to be comparing alternative courses of actions, as classical models would predict.
@@
 
-Klein (1999) proposed a model of recognition-primed decision, in which experienced decision-makers would find themselves immersed in complex situations and rapidly take adequate courses of action.
-Decision-makers would rapidly perceive cues from any situation and retrieve from episodic memory similar situations (Tulving 1983), which would bring assessments and diagnoses and plausible courses of action.
-Because priming mechanisms are automatic and unconscious (Bargh and Chartrand 1999, Bargh et al.
-2001), these decision-makers reported doing `the obvious' action in different situations.
+\citet{klein1999sources} proposed a model of \emph{recognition-primed decision}, in which experienced decision-makers would find themselves immersed in complex situations and rapidly take adequate courses of action.
+Decision-makers would rapidly perceive cues from any situation and retrieve from episodic memory similar situations, which would bring assessments and diagnoses and plausible courses of action.
+Because priming mechanisms are automatic and unconscious \citep{bargh1999unbearable, bargh2001automated}, these decision-makers reported doing `the obvious' action in different situations.
 This `obvious' course of action, Klein proposes, is brought from long-term episodic memory by priming mechanisms.
@@
 Even if the `obvious' action seemed plausible for a theory, another problem remained:  if decision-makers did not compare alternatives, then how could they know that a course of action was good?
-In subsequent interviews, evidence emerged that decision-makers would be using the simulation heuristic, proposed by Kahneman and Tversky (1982).
+In subsequent interviews, evidence emerged that decision-makers would be using the simulation heuristic, proposed by \citet{kahneman1982simulation}.
 That is, facing a particular situation, experienced decision-makers would be primed towards a particular course of action, to the detriment of most alternative courses of action.
@@
 Now it seemed that it was the experts who could generate a single course of action, while novices needed to compare different approaches.
---- Klein (1999, p.21)
+(p.21)
 \end{quote}
@@
 Because priming mechanisms that brought plausible actions to mind are unconscious, people would report having ``done the obvious thing to do''.
-Decision-makers would be unable to visualize the cognitive processes underlying their decisions, and would in many cases even believe that they had skills of the `fantastic' variety:  One firefighter demands that his whole crew abandon operations inside a house, just to see it collapse seconds afterward.
+Decision-makers would be unable to visualize the cognitive processes underlying their decisions, and would in many cases even believe that they had skills of the `fantastic' persuasion:  One firefighter demands that his whole crew abandon operations inside a house, just to see it collapse seconds afterward.
 A radar operator would `chill' after spotting a new track, and would fire counter missiles against it, based on the `feeling' that it was a hostile missile.
-It took over a year for this radar operator, after being interviewed by Klein, to understand the incredibly subtle cues that he was responding to whenever he perceived the new radar track.
+It took over a year for this radar operator, after being carefully and repeatedly interviewed, to understand the incredibly subtle cues that he was responding to whenever he perceived the new radar track.
 Unable to reasonably explain their life-saving, rapid, decisions, both the firefighter and the radar operator thought that they had ESP or other fantastic abilities.
@@
 Beyond a formalization of this process of memory recall, Sparse Distributed Memory will offer us a plausible, both psychologically and neuroscientifically, path towards artificial intelligence.
+We'll need to take a turn here.
 

### 2018-04-17 Chapters/Chapter02.tex
papers TODO

@@
 in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figure \ref{fig:bongard-problems-71}).
-Bongard problems demand high-gossip between system 1 and system 2; they demand a high dose of imagination and mental imagery manipulation; they demand discrimination and demand clusterization, among many other things that have been spelled out before by \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013} and others.
+Bongard problems demand high-gossip between system 1 and system 2; they demand a high dose of imagination and mental imagery manipulation; they demand discrimination and demand clusterization, among many other things that have been spelled out before by \citet{bongard_recognition_1968, douglas1980godel, Linhares2000} and the only serious, deep, study being citet{foundalis_phaeaco:_2006}.
 

### 2018-04-17 Chapters/Chapter0A.tex
papers TODO

--- 2018-04-09 Chapters/Chapter0A.tex
@@
 In the latter case, infinite number of petitions can be filed for extensions.
-Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-CS types': this wonderful result is from Bala Rajagopalan, posted on the late rec.humor.funny Usenet in June 12 1989.}.
+Since the Grad College never remembers previous petitions, the total number of petitions received by the Grad College is always one, a small constant\footnote{`With apologies to non-CS types': this seminal, wonderful, result is from Bala Rajagopalan, published on the late rec.humor.funny Usenet in June 12, 1989.
+Thank you for reading my thesis!}.
 

### 2018-04-17 Chapters/Chapter02.tex
only sdm intro todo...

--- 2018-04-17 Chapters/Chapter02.tex
@@
 
-Psychological
-orthogonality of concepts; life flashing before one's eyes; miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue...
-how does this recall process work?
+Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
+\begin{enumerate}
+    \item Why are most concepts orthogonal, unrelated to each other?
+    \item Why is there Miller's magic number, i.e., we can't hold too many things in mind at once?
+    \item Why do we at times instantly recall an experience; other times we can't recall anything at all; and still other times we get into this strange tip-of-the-tongue situation...
+it is clearly `there'...
+but where?
+(and `what' is `there'?)
+    \item How does this recall process work?  What is remembering?
+    \item Why do neurons die and we still remember most everything?
+    \item What do neurons actually do?
+\end{enumerate}
 
-Neuroscientific (bioplausible)
-neurons can die; what do neurons do?  how are things stored?  how are they recalled?
-
-Cognitive modelling (fully experimental, with no voodoo-like mysterious properties)
-
-3 sparse distributed memory 11
-
-3.1 Neurons as pointers 17
-
-3.2 Concepts 17
-
-3.3 Read operation 18
-
-3.3.1 Generalized read operation
-
-3.4 Critical Distance 20
-
-4 framework architecture 23
-
-4.1 Bitstring 24
-
-4.2.2 OpenCL kernels
-
-4.3 Counters 27
-
-4.4 Read and write operations
-
-5 results (i): performance
-
-5.1 Kernels comparison 38
-
-5.2 Scanners comparison 39
-
-5.3 Read and write operations
-
-6.1 Distance between random bitstrings
-
-6.2 Number of activated hard locations
-
-6.3 Intersection of two circles 63
-
-6.4 Storage and retrieval of sequences
-
-5.3.1 Summary of results
-
-6 results (ii): framework validation
-
-4.1.1 The distance between two bitstrings 24
-
-4.2 Address space 25
-
-4.2.1 Scanning for activated hard locations 25
-
-6.4.1 k-fold memory using only one SDM
-
-7 results (iii): loss of neurons 67
-
-8 results (iv): critical distance 73
-
-8.1 A deviation from the equator distance? 77
-
-8.2 Counter bias 80
-
-8.3 Read bias 83
-
-8.4 Critical distance of 209 85
-
-9 results (v): generalized read operation 87
-
-10 results (vi): supervised classification
-application
-
-11 results (vii): image noise filtering application 101
-
-12 results (viii): the possibility of unsupervised reinforcement learning 107
-
-13 results (ix): information-theoretical write operation 115
+These are some of the questions touched by SDM.
+It is a theory of long-term memory that constructs the process of recall from a very microscopic level of neuron-firings.
+While there are numerous groups working on related research, it seems that most are simply bypassing each other; starting from scratch instead of building on top of what already exists.
 
@@
 
-As expected for such an important thery, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\citet{Linhares2011} has shown that SDM respects the limits of short-term memory \citep{Miller1955, Cowan2001}.
+As expected for such an important theory, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
+\citet{Linhares2011} has argued that SDM respects the limits of short-term memory \citep{Miller1955, Cowan2001}.
+Kanerva's book has over 1,000 citations in Google Scholar.
 Despite this, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
@@
 in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figure \ref{fig:bongard-problems-71}).
-Bongard problems demand high-gossip between system 1 and system 2; they demand a high dose of imagination and mental imagery manipulation; they demand discrimination and demand clusterization, among many other things that have been spelled out before by \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013} and others.
+Bongard problems demand high-gossip between system 1 and system 2; they demand a high dose of imagination and mental imagery manipulation; they demand discrimination and demand clusterization, among many other things that have been spelled out before by \citet{bongard_recognition_1968, douglas1980godel, Linhares2000} and the landmark, deep, study being \citet{foundalis_phaeaco:_2006}.
 

### 2018-04-17 Chapters/sdm-autocorrelation.tex
only sdm intro todo...

--- 2018-04-09 Chapters/sdm-autocorrelation.tex
@@
 A hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
+Imagine the `counter's eye view':  each individual counter `likes'\footnote{If you are a cell I then I have some serious piece of advice to you.
+I would strongly suggest avoiding the neighborhood of a neuron.
+Neurons have serious authoritarian tendencies, bossing around --- generally electrifying --- the hopeless innocent cells around them just because.} to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
 

### 2018-04-18 Chapters/Chapter02.tex
atom has lost it

@@
 \chapter*{List of Jupyter notebooks}
+
+\begin{enumerate}
+  \item Calculate critical distance.ipynb
+  \item Calculate radius for any SDM.ipynb
+  \item Classification Test 1.ipynb
+  \item Classification Test 2.ipynb
+  \item Classification Test 3-Copy1.ipynb
+  \item Classification Test 3.ipynb
+  \item Critical Distance - New.ipynb
+  \item Critical Distance.ipynb
+  \item Distance between bitstrings.ipynb
+  \item Distances of activated hard-locations.ipynb
+  \item Generates a new SDM.ipynb
+  \item index.rst
+  \item Intersection of activated hard-locations.ipynb
+  \item Kanerva-Table-7 - Experiment.ipynb
+  \item Kanerva-Table-7.3 - Generic Read.ipynb
+  \item Kanerva-Table-7.3 - Multiple dimensions.ipynb
+  \item Kanerva-Table-7.3 - Weighted table write.ipynb
+  \item Kanerva-Table-7.3.ipynb
+  \item Kanerva's Figure 1.2.ipynb
+  \item Noise filter 2.ipynb
+  \item Noise filter A-Z 0-1 Generator.ipynb
+  \item Noise filter with labels.ipynb
+  \item Noise filter.ipynb
+  \item Number of activated hard-locations.ipynb
+  \item Performance test.ipynb
+  \item Reseting hard-locations.ipynb
+  \item Sequences (Kanerva Ch 8).ipynb
+  \item Sequences in one SDM (Kanerva Ch 8).ipynb
+  \item table74.pickle
+  \item TicTacToe - Old (TO BE DELETED).ipynb
+  \item TicTacToe.ipynb
+  \item tmp-TO-BE-DELETED-2.ipynb
+  \item tmp-TO-BE-DELETED.ipynb
+  \item Untitled.ipynb
+  \item Untitled1.ipynb
+  \item Weighted operations using information.ipynb
+\end{enumerate}

### 2018-04-19 Chapters/sdm-autocorrelation.tex
SDM Autocorrelation vanishes when n -> infty.

@@
 What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0.
+
+\begin{theorem}
+Let $r$ be the access radius given that $f$ percent of the hard locations are activated. Then, $\lim_{n \rightarrow \infty} r/n = 1/2$.
+\label{thm:sdm-access-radius}
+\end{theorem}
+\begin{proof}
+Let $\text{ppf}$ be the inverse of the cdf of the standard normal distribution. Then,
+
+\begin{align}
+r &= \frac{d}{2} + \text{ppf}(f) \frac{\sqrt{d}}{2} \\
+\frac{r}{d} = \frac{1}{2} + \text{pff}(f) \frac{1}{2 \sqrt{d}}
+\end{align}
+
+Thus, $d \rightarrow \infty \Rightarrow r/d \rightarrow 1/2$.
+\end{proof}
+
+\begin{theorem}
+$\lim_{n \rightarrow \infty} P(x_i = y_i | d(x, y) \le r) = 1/2$
+\label{thm:sdm-autocorrelation-convergence}
+\end{theorem}
+\begin{proof}
+
+From \ref{thm:sdm-access-radius}, we know that $\lim_{n \rightarrow \infty} = d/2$.
+
+For $k$ even, then:
+
+$$
+\sum_{k=0}^{r} \binom{n}{k} = \sum_{k=0}^{n/2} \binom{n}{k} = \frac{2^d}{2} = 2^{d-1}
+$$
+
+$$
+\sum_{k=0}^{r} \binom{n-1}{k} = \sum_{k=0}^{n/2} \binom{n}{k} = \frac{2^{d-1}}{2} - \frac{1}{2} \binom{d-1}{d/2}
+$$
+
+Finally,
+
+\begin{align}
+P(x_i = y_i | d(x, y) \le r) &= \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}} \\
+    &= \frac{2^{d-2} - \frac{1}{2} \binom{d-1}{d/2}}{2^{d-1}} \\
+    &= \frac{2^{d-2}}{2^{d-1}} - \frac{1}{2^d} \binom{d-1}{d/2} \\
+    &= \frac{1}{2} - \frac{1}{2^d} \binom{d-1}{d/2}
+\end{align}
+
+As $d \rightarrow \infty \Rightarrow \frac{1}{2^d} \binom{d-1}{d/2} \rightarrow 0$, the proof is done.
+
+For $k$ odd, the proof is equivalent.
+
+\end{proof}
+
+In Figure \ref{fig:sdm-10kbits-figure-7-3}, $n=10,000$ and we can notice that the autocorrelation has vanished as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=0.75\textwidth]{./images02/sdm-10000bits-10000w-table-7-2.png}
+
+  \caption{Given an address $x$ and a dimension $i$, how many hard locations with bitmatches in $i$ are activated by reading at $x$?
+The histogram was obtained through numerical simulation.
+The red curve is the theoretical normal distribution found in Theorem \ref{T2}.}
+  \label{fig:sdm-10kbits-figure-7-3}
+\end{figure}
 

### 2018-04-19 Chapters/sdm-autocorrelation.tex
SDM Minor change from previous commit.
SDM Autocorrelation theorems.

--- 2018-04-17 Chapters/sdm-autocorrelation.tex
@@
 What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0.
+
+\begin{lemma}
+Let $r$ be the access radius given that $f$ percent of the hard locations are activated. Then, $\lim_{n \rightarrow \infty} r/n = 1/2$.
+\label{thm:sdm-access-radius}
+\end{lemma}
+\begin{proof}
+
+As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 0.5 ($B(n, 0.5)$).
+For $n$ sufficiently large, the Binomial distribution can be approximated by a Normal distribution, i.e., $B(n, 0.5) \rightarrow \mathcal{N}(\mu = n/2, \sigma^2 = n/4)$.
+
+Let $\Phi(x)$ be the cdf of the standard normal distribution. Let $z = \frac{r - n/2}{\sqrt{n}/2}$. Thus, $P(d(x, y) \le r) = \Phi(z)$. As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$.
+
+Calculating the inverse, $z = \Phi^{-1}(f)$. Then,
+
+\begin{align}
+z &= \Phi^{-1}(f) \\
+\frac{r - n/2}{\sqrt{n}/2} &= \Phi^{-1}(f) \\
+r &= \frac{n}{2} + \Phi^{-1}(f) \frac{\sqrt{n}}{2} \\
+\frac{r}{n} &= \frac{1}{2} + \Phi^{-1}(f) \frac{1}{2 \sqrt{n}}
+\end{align}
+
+Therefore, $n \rightarrow \infty \Rightarrow r/n \rightarrow 1/2$.
+\end{proof}
+
+\begin{theorem}
+The autocorrelation vanishes when $n \rightarrow \infty$, i.e., $\lim_{n \rightarrow \infty} P(x_i = y_i | d(x, y) \le r) = 1/2$.
+\label{thm:sdm-autocorrelation-convergence}
+\end{theorem}
+\begin{proof}
+
+From Lemma \ref{thm:sdm-access-radius}, we know that $r = n/2$ for $n$ sufficiently large.
+
+Suppose that $n$ is an even integer, then,
+
+$$
+\sum_{k=0}^{r} \binom{n}{k} = \sum_{k=0}^{n/2} \binom{n}{k} = \frac{1}{2} \sum_{k=0}^{n} \binom{n}{k} = \frac{2^n}{2} = 2^{n-1}
+$$
+
+And, also,
+
+\begin{align*}
+\sum_{k=0}^{r} \binom{n-1}{k} &= \sum_{k=0}^{n/2} \binom{n-1}{k} \\
+    &= \frac{1}{2} \left[ \sum_{k=0}^{n-1} \binom{n-1}{k} - \binom{n-1}{n/2} \right] \\
+    &= \frac{1}{2} \left[ 2^{n-1} - \binom{n-1}{n/2} \right] \\
+    &= 2^{n-2} - \frac{1}{2} \binom{n-1}{n/2}
+\end{align*}
+
+Finally,
+
+\begin{align}
+P(x_i = y_i | d(x, y) \le r) &= \frac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}} \\
+    &= \frac{2^{n-2} - \frac{1}{2} \binom{n-1}{n/2}}{2^{n-1}} \\
+    &= \frac{2^{n-2}}{2^{n-1}} - \frac{1}{2^n} \binom{n-1}{n/2} \\
+    &= \frac{1}{2} - \frac{1}{2^n} \binom{n-1}{n/2}
+\end{align}
+
+As $n \rightarrow \infty \Rightarrow \frac{1}{2^n} \binom{n-1}{n/2} \rightarrow 0$, the proof is done for $n$ even.
+
+When $n$ is an odd integer, the steps of the proof are similar. Therefore, the proof is complete.
+
+\end{proof}
+
+In Figure \ref{fig:sdm-10kbits-figure-7-3}, $n=10,000$ and we can notice that the autocorrelation has vanished as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=0.75\textwidth]{./images02/sdm-10000bits-10000w-table-7-2.png}
+
+  \caption{Given an address $x$ and a dimension $i$, how many hard locations with bitmatches in $i$ are activated by reading at $x$?
+The histogram was obtained through numerical simulation.
+The red curve is the theoretical normal distribution found in Theorem \ref{T2}.}
+  \label{fig:sdm-10kbits-figure-7-3}
+\end{figure}
 

### 2018-04-19 Chapters/sdm-autocorrelation.tex
SDM Autocorrelation theorems.

@@
 
+\begin{lemma}
+Let $\Phi(x)$ be the cdf of the standard normal distribution.
+Then, $n \rightarrow \infty \Rightarrow P(x_i = y_i | d(x, y) \le r) = \frac{1}{2} \frac{\Phi(z_1)}{\Phi(z_2)}$, where $z_1 = \frac{2r - n + 1}{\sqrt{n-1}}$ and $z_2 = \frac{2r - n}{\sqrt{n}}$.
+\label{thm:sdm-T1-n-large}
+\end{lemma}
+\begin{proof}
+
+From the approximation of the Binomial distribution $B(a, 0.5)$ by the Normal distribution $\mathcal{N}(\mu = a/2, \sigma^2 = a/4)$, we conclude that, for $a$ sufficiently large, the cdf of the Binomial is approximately equal to the cdf of the Normal distribution.
+Thus,
+
+$$
+\frac{1}{2^a} \sum_{k=0}^{b} \binom{a}{k} = \Phi \left( \frac{b-a/2}{\sqrt{a}/2} \right) = \Phi \left( \frac{2b-a}{\sqrt{a}} \right)$
+$$
+
+Thus,
+
+$$
+\sum_{k=0}^{b} \binom{a}{k} = 2^a \Phi \left( \frac{2b-a}{\sqrt{a}} \right)$
+$$
+
+The result comes directly from applyting the equation above in $P(x_i = y_i | d(x, y) \le r)$.
+
+\end{proof}
+
 \begin{theorem}
@@
 From Lemma \ref{thm:sdm-access-radius}, we know that $r = n/2$ for $n$ sufficiently large.
+Thus, replacing $r = n/2$ in Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = \frac{\Phi(z_1)}{2 \Phi(z_2)}$, where $z_1 = \frac{1}{\sqrt{n-1}}$ and $z_2 = 0$.
+
+As $n \rightarrow \infty$, $z_1 \rightarrow 0$, and $P(x_i = y_i | d(x, y) \le r) = \frac{\Phi(0)}{2 \Phi(0)} = 1/2$. \qed
+
+Another way to proof is to divide into two cases:
 
@@
 
-In Figure \ref{fig:sdm-10kbits-figure-7-3}, $n=10,000$ and we can notice that the autocorrelation has vanished as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
+In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has vanished as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
+In fact, in this case, $P(x_i = y_i | d(x, y) \le r) = 0.5039895557907589$.
 
@@
 
-  \caption{Given an address $x$ and a dimension $i$, how many hard locations with bitmatches in $i$ are activated by reading at $x$?
-The histogram was obtained through numerical simulation.
-The red curve is the theoretical normal distribution found in Theorem \ref{T2}.}
+  \caption{The same setup as in Figure \ref{fig:sdm-10000w-table-7-2}, but for $n=10,000$. It shows that the interference has gone away when $n$ is sufficiently large.}
   \label{fig:sdm-10kbits-figure-7-3}

### 2018-04-19 Chapters/sdm-autocorrelation.tex
SDM Autocorrelation.

@@
 $$
-\frac{1}{2^a} \sum_{k=0}^{b} \binom{a}{k} = \Phi \left( \frac{b-a/2}{\sqrt{a}/2} \right) = \Phi \left( \frac{2b-a}{\sqrt{a}} \right)$
+\frac{1}{2^a} \sum_{k=0}^{b} \binom{a}{k} = \Phi \left( \frac{b-a/2}{\sqrt{a}/2} \right) = \Phi \left( \frac{2b-a}{\sqrt{a}} \right)
 $$
@@
 $$
-\sum_{k=0}^{b} \binom{a}{k} = 2^a \Phi \left( \frac{2b-a}{\sqrt{a}} \right)$
+\sum_{k=0}^{b} \binom{a}{k} = 2^a \Phi \left( \frac{2b-a}{\sqrt{a}} \right)
 $$
@@
 
-In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has vanished as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
-In fact, in this case, $P(x_i = y_i | d(x, y) \le r) = 0.5039895557907589$.
+In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has reduced significantly as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
+In fact, in this case, using Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = 0.5168761116022098$.
 
@@
 
-  \caption{The same setup as in Figure \ref{fig:sdm-10000w-table-7-2}, but for $n=10,000$. It shows that the interference has gone away when $n$ is sufficiently large.}
+  \caption{The same setup as in Figure \ref{fig:sdm-10000w-table-7-2}, but for $n=10,000$. It shows that the interference has almost gone away when $n$ is sufficiently large.}
   \label{fig:sdm-10kbits-figure-7-3}

### 2018-04-20 Chapters/sdm-autocorrelation.tex
SDM Autocorrelation.

@@
 
-As $n \rightarrow \infty \Rightarrow \frac{1}{2^n} \binom{n-1}{n/2} \rightarrow 0$, the proof is done for $n$ even.
+Stirling's approximation yields that, for $n$ sufficiently large, $\binom{n}{n/2} \sim \frac{2^n}{\sqrt{\pi n/2}}$.
+Thus, $\frac{1}{2^n} \binom{n}{n/2} \sim \frac{\sqrt{2}}{\sqrt{\pi n}}$, which yields $\lim_{n \rightarrow \infty} \frac{1}{2^n} \binom{n}{n/2} = 0$.
+Finally, as $\binom{n-1}{n/2} \le \binom{n}{n/2}$, by the squeeze theorem, $\lim_{n \rightarrow \infty} \frac{1}{2^n} \binom{n-1}{n/2} = 0$, which closes the proof for $n$ even.
 

### 2018-04-22 Chapters/sdm-autocorrelation.tex
SDM Autocorrelation typo.
recompiling..

@@
 A hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes'\footnote{If you are a cell I then I have some serious piece of advice to you.
+Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
+
+%\footnote{If you are a cell I then I have some serious piece of advice to you.
 I would strongly suggest avoiding the neighborhood of a neuron.
-Neurons have serious authoritarian tendencies, bossing around --- generally electrifying --- the hopeless innocent cells around them just because.} to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
+Neurons have serious authoritarian tendencies, bossing around --- generally electrifying --- the hopeless innocent cells around them just because.}
 

### 2018-04-23 Chapters/Chapter01.tex
Hathor: Introduction.

@@
 % TODO Lightning network, é uma solução de scaling? Parece que não.
+
+% TODO Governance (!!!)
+% TODO Rule change policy (maybe PoS)
+% TODO Miner/transaction punishment
+% TODO Different hash function for miners and transactions (!!!)
+%      Scrypt and SHA-256?
 
@@
 
-The main problem with digital money is how to prevent double spending.
+The primary problem for creating digital money is how to prevent double spending.
 As the money is digital, and copies can be made \textit{ad nauseam}, what can prevent counterfeiting?
 That is, what would prevent users from sending copies of the same money to two (or more) people?
-That is precisely the problem solved by Bitcoin and its underlying blockchain technology.
-The current solution behind fiat money is having a single issuer, a central bank --- then trusting the financial institutions.
+That is precisely the problem solved by Bitcoin and its underlying Blockchain technology.
+The current solution behind fiat money is having a single issuer, a central bank --- then trusting the financial institutions and regulators.
 
-Bitcoin (BTC) is the first digital currency, also known as digital money, internet money, and cryptocurrency.
-It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank.
-Bitcoin is distributed since its ledger is public and is stored in thousands of computers.
-It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community.
+The concept of transfering money using cryptography as an underlying technology was shortly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}.
+However, it was only in 1988 that \citet{chaum1988untraceable} created the term \emph{electronic cash} and also proposed a basic and practical scheme which yielded untraceability yet allowed to trace double spendings.
+
+According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending \citep{chaum1983blind, okamoto1995efficient, camenisch2005compact, canard2007divisible}.
+The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the previous e-cash philosophies.
+
+Bitcoin (BTC) is a digital currency, also known as digital money, internet money, and cryptocurrency.
+It is the first currency based on cryptography techniques which is distributed, decentralized, and with no central bank.
+
+Bitcoin is a computer network in which nodes act like clerks performing clearing.
+A transaction clearing consists of ensuring that the transaction is settled according to the rules.
+In order to do that, every node stores a copy of Bitcoin's ledger, which records both all transactions and users' balance.
+When new transactions are added to the ledger, the balances are updated.
+It is said that Bitcoin is distributed because its ledger is public and is stored in thousands of computers.
+Even though the ledger is public, balances are anonymous and no one knows who owns which funds.
+
+Bitcoin is considered decentralized because there is no authority (or government) who decides its future.
+Every decision must be accepted by its community.
+When the majority of Bitcoin's network agrees on a decision, they just have to update their clearing process accordingly.
+
 The security of Bitcoin relies on digital signature technology and network agreement.
@@
 In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution.
-When all these requirements are met, the funds are transferred.
+Only when all these requirements are met, the funds are transferred.
 
-According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending.
-The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the previous e-cash philosophies.
-% literature
-
-Bitcoin provides interesting incentives to all players, namely the users and the miners.
-On the one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction --- (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
-On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes, and they may also receive the fees of unconfirmed transactions.
+Bitcoin provides interesting incentives to all players, users and miners.
+On the one hand, users may have incentives to use Bitcoin because (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction ---; (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes Bitcoin's supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
+On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes, and they may also collect the fees of unconfirmed transactions.
 These incentives have kept the Bitcoin network up and running since 2009 with barely any interruptions (99.99\% uptime).
@@
 Despite Bitcoin's huge success, there are still many challenges to be overcome.
-We will focus on a specific subset of those challenges, namely the scaling, decentralization, and spam challenges.
+We will focus on one specific of those challenges: scaling.
 One important challenge that we will skip is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
@@
 
+\section{Scripts \& Contracts}
+
 \chapter{Methodology}

### 2018-04-23 Chapters/Chapter02.tex
recompiling..
SDM Typos.

@@
 You tighten your grip on the child's hand and step up your walking.
-In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everthing you think of is to get out to safety.
+In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everything you think of is to get out to safety.
 You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exactly is causing your fear.

### 2018-04-25 FrontBackmatter/Abstract.tex
Abstract and acnowledgments (incomplete).

@@
 \chapter*{Abstract}
-Short summary of the contents in English\dots
+
+This thesis presents three specific, self-contained, scientific papers in the Computational Management Science area.
+Modern management and high technology interact in multiple, profound, ways.
+Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'', as his hyperbolic way to emphasize the potential transformational power of the technology.
+
+The first paper is inspired by the possibility that there will be some form of purely digital money and studies distributed ledgers, proposing and analyzing Hathor, an alternative architecture towards a scalable cryptocurrency.
+
+The second paper may be a crucial item in understanding human decision making, perhaps, bringing us a formal model of recognition-primed decision.
+Lying at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence, it presents an open-source, cross-platform, and highly parallel framework of the Sparse Distributed Memory and analyzes the dynamics of the memory with some applications.
+
+Last but not least, the third paper lies at the intersection of marketing, diffusion of technological innovation, and modeling, extending the famous Bass model to account for users who, after adopting the innovation for a while, decide to reject it later on.
 

### 2018-04-25 FrontBackmatter/Acknowledgments.tex
Abstract and acnowledgments (incomplete).

--- 2018-02-08 FrontBackmatter/Acknowledgments.tex
@@
 Thank you for your patience and for understanding that I was physically far away but very close in heart.
-A special thanks to my dad, Reynaldo, who always taught me about life and who first awakened my passion for technology and engineering.
+A special thanks to my dad, Reynaldo, who always inspired me and who first awakened my passion for technology and engineering.
 
-Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again.
-Especially, for understanding my absence.
+Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love. Especially, for understanding my absence.
 
-Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.
+I'm very grateful for the friendship, guidance, and invaluable help of my advisor Alexandre Linhares.
+Without him this work would never be possible.
+His passion for cognitive sciences, artificial intelligence, and economics has influenced everyone around him.
+His ideas have profoundly influenced my thoughts and my life since we have met.
+I must not forget his wife Paula, who became a dear friend and has always supported me as well, even when I was working in her living room until the sunrise.
+Thank you for this amazing journey, my friends.
 
-Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
+Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
+I will never forget our all-you-can-eat adventures in Bros restaurant, the snacks in Catarina, and the drinks in Bigode bar.
+You have become close friends who I will carry for life.
 
-Thanks to Celene.
+Thanks to my long-time partner and dear friend Claudio Abreu, who always encouraged me to take risks and to improve myself. He is the most fearless person I have ever met.
 
-Thanks to Claudio Abreu.
-
-Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.
+Thanks to EBAPE / FGV for their great faculty, staff, and infrastructure.
+A special thanks to Celene Silva Melo for her empathy and support besides all the help with paperwork, funding, and many other things.
+Another special thanks to doctor Luiz Antônio Joia, who trusted me in the first place and guided my first steps into the world of academy.
 

### 2018-04-26 Chapters/Chapter01.tex
Minor abstract and SDM changes.

@@
 Despite Bitcoin's huge success, there are still many challenges to be overcome.
-We will focus on one specific of those challenges: scaling.
+We will focus on the following of those challenges: scaling, spamming, and centralization.
 One important challenge that we will skip is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
@@
 
+The scaling problem is not precisely an issue of Bitcoin, but an issue of the Blockchain technology.
+Hence, all other Blockchain-based cryptocurrencies have the same limitations, such as Litecoin, Bitcoin Cash, and Ethereum.
+One may arguee that increasing the block size is a feasible solution to scaling, but I would reply that it is a temporary solution which buys some time until the next network saturation.
+
 Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
@@
 
-Generating new transactions in Bitcoin has a tiny cost because one only has to generate the transaction itself, digitally sign it, and propagate in the Bitcoin network.
+Generating new transactions in Bitcoin has a tiny computational cost because one only has to generate the transaction itself, digitally sign it, and propagate it in the Bitcoin network.
 On the one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks.

### 2018-04-26 FrontBackmatter/Abstract.tex
Minor abstract and SDM changes.

@@
 
-\vfill
+%\vfill
 
-\pdfbookmark[1]{Resumo}{Resumo}
 \chapter*{Resumo}
-Aqui entre o resumo\dots
+A presente tese é formada por três trabalhos científicos na área de Management Science Computacional.
+A gestão moderna e a alta tecnologia interagem em múltiplas e profundas formas.
+O professor Andre Ng diz aos seus estudantes na Escola de Negócios de Stanford que ``Inteligência Artificial é a nova eletricidade'', como sua forma hiperbólica de enfatizar o potencial transformador da tecnologia.
+
+O primeiro trabalho é inspirado na possibilidade de que haverá alguma forma de dinheiro digital e estuda ledger distribuídas, propondo e analisando o Hathor, uma arquitetura alternativa para criptomoedas escaláveis.
+
+O segundo trabalho pode ser um item crucial no entendimento de tomadas de decisão, nos trazendo um modelo formal de \emph{recognition-primed decisions}.
+Situada na intersecção entre psicologia cognitiva, ciência da computação, neuro-ciência e inteligência artifical, ele apresenta um framework open-source, multi-plataforma e altamente paralelo da Sparse Distributed Memory e analisa a dinâmica da memória e algumas aplicações.
+
+O terceiro e último trabalho se situa na intersecção entre marketing, difusão de inovação tecnologica e modelagem, extendendo o famoso modelo de Bass para levar em consideração usuário que, após adotar a tecnologia por um tempo, decidiram rejeitá-la.
 
@@
 
-\vfill
+%\vfill

### 2018-04-26 Chapters/Chapter01.tex
Minor Hathor improvements.

@@
 On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes, and they may also collect the fees of unconfirmed transactions.
+It is important to highlight that anyone can become a miner, and there is no entry cost besides the mining equipment.
 These incentives have kept the Bitcoin network up and running since 2009 with barely any interruptions (99.99\% uptime).
+For further information about incentives, see \citet{ma2018market, catalini2016some}.
 
@@
 
-Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power.
+Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power (see \citet{gencer2018decentralization}).
 Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior, and it is also seen as a problem to be solved.

### 2018-04-27 Chapters/sdm2.tex
The incredible animal behavior of Dr Linhares

@@
 Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community.
-A good survey is provided by \citet{harary1988survey}.
+A good survey is provided by \citet{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.
 
@@
 
-It is important to say that Chada's read came from \citet{anwar2003sparse} which gave a misguided description of the read operation. The original description is the following:
+It is important to say that Chada's read came from \citet{Anwar2003} which gave a misguided description of the read operation. The original description is the following:
 
@@
 %Similar to writing, retrieving from SDM involves the same concept of access sphere --- you read all the hardlocations within the access sphere of location Y, pool the bit vectors read from all these hard locations and let each of the kth bits of those locations participate in a majority vote for the kth bit of Y.
-\hfill --- \citet{anwar2003sparse}
+\hfill --- \citet{Anwar2003}
 With our datum distributively stored, the next question is how to retrieve it.
@@
 \\
-\hfill --- \citet[p.342]{anwar2003sparse}
+\hfill --- \citet[p.342]{Anwar2003}
 \end{quote}

### 2018-04-27 Chapters/Chapter02.tex
sdm ready?

--- 2018-04-18 Chapters/Chapter02.tex
@@
 % TODO GARY KLEIN !!!
-GARY KLEIN!!!!
 
@@
 You tighten your grip on the child's hand and step up your walking.
-In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everthing you think of is to get out to safety.
+In these large cities, walking on a dark alley was never recommended, but somehow that's what you have done --- and now everything you can think of is to get back to safety.
 You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exactly is causing your fear.
 One just feels it.
-We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; a general lack of activity; that very feeling that something isn't right...
-without knowing precisely what isn't right, or even what `right' means...
-etc.
---- recalling past experiences from memory and thus generating the feeling.
+
+We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind; a general lack of activity; that very feeling that something isn't right...
+without knowing precisely what isn't right, or even what `right' means --- recalling past experiences from memory and thus generating the feeling.
 Our memory is able to make a parallel between previous experiences and the clues.
-Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger.
-The whole process happens so fast it feels like it only took a single unit of psychological time.
+Although one has never been in the exactly same situation, one's brain involuntarily, through analogy with previous experiences, recognizes the possibility of danger.
+The whole process happens so fast it feels like it only took a single indivisible unit of psychological time.
+A single cycle, to use the computing word.
 
+As much as we can rationalize our ability to understand the why of a feeling like this, it seems unlikely that we will be able to nail all reasons down into rules.
 Let us compare this to what a child might do.
+While an adult may respond by discreetly stepping up the walking speed, a child might open her eyes and be surprised and ask, indiscretely, questions about the walking speed...
+Because children do not have the accumulated experience of adults in dealing with large metropolian areas, their behavior differs markedly from ours.
 
@@
 but where?
-(and `what' is `there'?)
+(and `what' is `there'?\footnote{Psychologists have documented interesting properties about this state:  it happens, for instance, around every 1-week cycle; and it happens mostly over proper names, subjects have access to the first letter and to the number of syllables better than chance, etc.
+Also interesting is their clever way of triggerring the state: through definitions of rarely-used words.
+For some psychological studies about the tip-of-tongue, see \citet{meyer_tip---tongue_1992, brown_review_1991, brown_tip_1966}.}.)
     \item How does this recall process work?  What is remembering?
     \item Why do neurons die and we still remember most everything?
-    \item What do neurons actually do?
+    \item What do neurons actually do?  What is their primary function?
 \end{enumerate}
@@
 
-\section{The wasted effort of duplicated, AD-HOC, work}
+\section{The wasted effort of duplicated, ad hoc, work}
 
@@
     \begin{quote}
-    The most exciting phrase to hear in science, the one that heralds new discoveries, is not `Eureka!' (I found it!) but ``That's funny...''
+    The most exciting phrase to hear in science, the one that heralds new discoveries, is not `Eureka!' (I found it!) but ``That's funny...'' \\
     — Isaac Asimov
-    \end{quote}
-
-    \bigskip
-\begin{quote}
-Alexander Fleming, while explaining the staphylococcal work to Pryce, to demonstrate a point, picked up one of the top layer of the discarded plates when he noticed the lysis of the staphylococci (his cry of Eureka was ‘That’s funny!’) and then and there made a subculture of the mould.
     \end{quote}
@@
   \item Generates a new SDM.ipynb
-  \item index.rst
   \item Intersection of activated hard-locations.ipynb
@@
   \item Sequences in one SDM (Kanerva Ch 8).ipynb
-  \item table74.pickle
-  \item TicTacToe - Old (TO BE DELETED).ipynb
   \item TicTacToe.ipynb
-  \item tmp-TO-BE-DELETED-2.ipynb
-  \item tmp-TO-BE-DELETED.ipynb
-  \item Untitled.ipynb
-  \item Untitled1.ipynb
   \item Weighted operations using information.ipynb

### 2018-04-27 Chapters/Chapter01.tex
Hathor: reorganizing the text.

@@
 According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending \citep{chaum1983blind, okamoto1995efficient, camenisch2005compact, canard2007divisible}.
+Some papers even propose solutions in a similar trajectory to Bitcoin, such as hash chain \citep{zongkai2004new} and preventing double spending using peer-to-peer networks \citet{osipkov2007combating, hoepman2007distributed}.
 The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the previous e-cash philosophies.
@@
 Even though the ledger is public, balances are anonymous and no one knows who owns which funds.
+If an attacker tries to change anything, the remaining of the network is able to detect it and ignore the change.
 
 Bitcoin is considered decentralized because there is no authority (or government) who decides its future.
-Every decision must be accepted by its community.
-When the majority of Bitcoin's network agrees on a decision, they just have to update their clearing process accordingly.
+Every decision must be accepted by its community and no one can enforce their will.
+Every change proposal must be submitted to the community who will discuss the matter and come to a verdict.
+If the majority of Bitcoin's community agrees on a decision, they just have to update their clearing process accordingly and the changes are applied.
 
@@
 Despite Bitcoin's huge success, there are still many challenges to be overcome.
-We will focus on the following of those challenges: scaling, spamming, and centralization.
+We will focus on the following challenges: scaling, spamming, and centralization.
 One important challenge that we will skip is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.
@@
 Hence, all other Blockchain-based cryptocurrencies have the same limitations, such as Litecoin, Bitcoin Cash, and Ethereum.
-One may arguee that increasing the block size is a feasible solution to scaling, but I would reply that it is a temporary solution which buys some time until the next network saturation.
+One may arguee that increasing the maximum block size is a feasible solution to scaling, but I would say that it is just a temporary solution which buys some time until the next network saturation.
 
-Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power (see \citet{gencer2018decentralization}).
-Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior, and it is also seen as a problem to be solved.
+Bitcoin seems to have the most decentralized network among the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power (for details, see \citet{gencer2018decentralization}).
+Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior.
+They may also cooperate in an attack, avoiding transactions which seemed confirmed.
 The more decentralized, the more trustworthy Bitcoin is.
+This centralization problem is seem as an important challenge.
 
@@
 One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network.
+This spam problem has happened several times and affects Bitcoin's trustworthy.
+\citet{bitcoinspam2017} reports a possible spam attack lasting at least 18 months.
 
@@
 Each transaction has its own proof-of-work and is used to confirm other transactions, forming a directed acyclic graph of transactions.
-Thus, a transaction is said to be confirmed when there are enough proof-of-work from the transactions confirming it directly or indirectly.
+Thus, a transaction is said to be confirmed when there is enough proof-of-work from the transactions confirming it directly or indirectly.
 There is no other way to confirm transactions but generating new transactions.
@@
 In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions.
-Hence, theoretically, it scales to any large number of transactions per second.
+Therefore, theoretically, it scales to any large number of transactions per second.
 The scaling problem of tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work.
 While Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator which works as a trustworthy node \citep{iotacoordinator}.
-This central coordinator just elucidates that the tangle does not seem to work properly under a low volume of transactions.
 
-Iota's tokens are pre-mined, which means all of them have been issued in the genesis transaction, and there is no token issuance at all.
-There is no miners and users can only transfer tokens between themselves.
-This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it.
-As users perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away \citep{debo2010prices}.
+Every transaction confirmed by the central coordinator is assumed to be valid and cannot be reverted.
+The remaining of the network can verify a confirmation through the central coordinator's digital signature.
+The coordinator will not be necessary anymore when the number of transactions per second reaches a minimum value, but Iota's developers cannot say precisely what is this minimum value.
+This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severily affect Iota's trustworthy).
 
@@
 The blockchain technology works through the creation of new blocks.
-Each new block confirms that all the previous blocks are valid and have not been tampered with.
-The mechanism that assures the immutability is proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network.
+Each new block confirms that all previous blocks are valid and have not been tampered with.
+The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network.
 The network agrees that work should be done in the block at the longest chain in the blockchain.
@@
 The lesser the given threshold $A$, the higher the difficulty.
-On the other side, checking whether a solution is correct is fast, one just has to apply the SHA-256 twice and compare.
+On the other hand, checking whether a solution is correct is fast since one just has to apply the SHA-256 twice and compare.
+
+When miners are finding a solution to a new block, they are mining or working in the new block.
+A block is found when a solution to the proof-of-work is found.
+When a new block is found, it indirectly confirms all the previous blocks in the chain and their transactions.
+
+It may happen that two miners find two different blocks in a small interval of time.
+In this case, miners will propagate their blocks and part of the network will choose one of them as the next block, whereas the other part of the network will choose the other one.
+This phenomenon is called a fork.
+Thus, when the next block is found, one of those blocks will be confirmed, and the other will be ignored and referred to as an orphan block.
+As blocks confirm transactions, all the transactions in the orphan block which have not been confirmed by another block will return to the unconfirmed transaction pool.
+Hence, it is not safe to accept a transaction when it is confirmed by only one block.
+It is the idea behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.
 
@@
 If it takes more than 14 days to find 2016 blocks, it means the network's hash power has decreased; thus the difficulty is decreased.
-
-When miners are finding a solution to a new block, they are mining or working in the new block.
-A block is found when a solution to the proof-of-work is found.
-When a new block is found, it indirectly confirms all the previous blocks in the chain and their transactions.
-It may happen that two miners find two different blocks in a small interval of time.
-In this case, both miners will propagate their blocks, and the network will randomly choose one of them as the next block.
-This phenomenon is called a fork.
-Thus, when the next block is found, one of those blocks will be confirmed, and the other will be ignored and referred to as an orphan block.
-As blocks confirm transactions, all the transactions in the orphan block which have not been confirmed by another block will return to the unconfirmed transaction pool.
-Hence, it is not safe to accept a transaction when it is confirmed by only one block.
-It is the idea behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.
 
@@
 Hence, the capacity remains the same, and the community has been discussing the issue in search of a consensus.
+For further details of the discussion, see \citet{bitcoinblocksize}.
 
@@
 
-Next, we will do a mathematical analysis of Bitcoin in order to better understand its minings properties, how a fork would affect the network and its security against attackers.
-
-\chapter{Analysis of Bitcoin}
-\input{./Chapters/hathor-bitcoin-math.tex}
+Next, we will describe the dynamics of Iota and its underlying technology Tangle. Its major difference from Bitcoin is that it has neither miners nor blocks.
 
@@
 Iota's tokens are pre-mined, which means they have been issued in the genesis transaction and no more tokens will ever be issued.
-This means there are no miners in Iota's network and only the users keep the network alive through their new transactions.
-If, for any reason, the number of transactions per second plummet, the time to confirmed the most recent transactions will increase significantly.
+This means that there are no miners in Iota's network and only the users keep the network alive through their new transactions.
+This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it.
+As users perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away, making it even harder to reach the minimum required to work properly \citep{debo2010prices}.
+Even if the minimum has been reached, if the number of transactions per second plummet, the time to confirm the most recent transactions will increase significantly.
 This possibly poses a serious risk to the network stability and users' trustworthy.
 
-As there are no miners, there are also no fees, which is a major incentive to new comers.
+As there are no miners, there are also no fees, which is a major incentive to newcomers.
 Tokens may be freely transfered without any loses, even for tiny amounts, enabling micropayments.
@@
 
-Not having fees is also also an important feature to the Internet-Of-Things (IoT) technologies.
-The IoT is a network of smart devices that are connected to the internet and exchange data.
-It allows, for instance, your refrigerator to do your groceries and automatically pay using your Iota tokens.
-It also allows your electric car to automatically pay for recharge.
+The Internet-Of-Things (IoT) technologies, which is a network of smart devices that are connected to the internet and exchange data, also benefits from not having fees.
+Imagine your refrigerator doing your groceries and automatically paying using your Iota tokens, or your electric car automatically paying for recharge.
+In both situations, transactions with no fees are interesting because these devices can freely make many microtransactions without any loss.
 For further information, see \cite{fleisch2010internet}.
@@
 
+Next, we will do a mathematical analysis of Bitcoin in order to better understand its minings properties, how a fork would affect the network and its security against attackers.
+
+\chapter{Analysis of Bitcoin}
+\input{./Chapters/hathor-bitcoin-math.tex}
+
 \chapter{Hathor's architecture}
 \input{./Chapters/hathor-architecture.tex}
-
-\section{Governance}
-
-- Punishment for miners/transactions which does not comply with the rules.
-
-- Rule change policy.
-
-\section{Expected number of tips}
-
-Explain transient and stationary state. The balance happens in the stationary state.
-
-In order to estimate the balance in the number of tips, let's consider the system in which the rate of inward tips is $\lambda_\text{TX}$, while the rate of outward tips is $\rho$.
-Let $K$ be the number of tips in the system.
-Thus, after $\Delta t$ seconds, $\Delta \text{tips} = \lambda_\text{TX} \Delta t - \rho K \Delta t$.
-In the balance, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
-
-Finally, the rate of outward tips is twice the rate in which devices solve the proof-of-work of transactions. Let $H$ be the devices' hash rate, then, $\rho = 2 \cdot H \cdot 2^{-w_\text{TX}}$.
-
-Twice?! It is a number between 1 and 2, but none of them.
-
-Therefore, in the balance,
-
-$$K = \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H}$$
-
-$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$
-
-\section{Attacks \& Conflicting transactions}
-
-\section{Scripts \& Contracts}
 
@@
 When a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work.
-When it finishes solving the proof-of-work, it is propagated and becomes a tip.
+When it finishes solving the proof-of-work, the transaction is propagated and becomes a tip.
 So, two new transactions may choose the same tips to confirm.
@@
 
-When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
+When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions.
+The depth-first search is interrupted when it reaches a transaction which the accumulated weight is bigger than a threshold.
+This interruption increases the overall performance of the simulation.
+
+%, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}.
 The topological sort is used to calculate the longest path from all other transactions to the new transaction, and this longest path is used to update the depth of the whole DAG.
@@
 
-The amount of time spent to solve the proof-of-work is a sample from an exponential distribution.
+The amount of time spent to solve the proof-of-work is sampled from an exponential distribution.
 
-The time between two new transactions is sampled from an exponential distribution with parameter $\lambda$.
+The time between two new transactions is also sampled from an exponential distribution with parameter $\lambda$.
 The value of $\lambda$ is changed over time, causing changes in the load of the network.
-We consider 1 tx/s to be a low load, and 15 tx/s to be a high load.
-Even though we have already tested up to 100 tx/s.
 
-The simulator will output different reports: (i) snapshots of the DAG at interesting moments, such as Fig.
+%The simulator will output different reports: (i) snapshots of the DAG at interesting moments, such as Fig.
 \ref{fig-tangle-example}, \ref{fig-tangle-swarm}, and \ref{fig-tangle-conflict}; (ii) histogram of confirmation time, such as Fig.
@@
 It is similar to Bitcoin's architecture when the number of transactions per second is low, while it is similar to Iota's architecture when the number of transactions per second is high.
-In order to check this statement, I have run some simulations.
+In order to check this statement, I have performed some simulations.
 
@@
 
-\begin{theorem}
-\label{theorem-new-tx-not-tip}
-When a new transaction confirms an already confirmed transaction, the depth of all transactions remains the same, i.e., the depth of the transactions are only changed when a new transaction confirm an unconfirmed transaction.
-\end{theorem}
+%\begin{theorem}
+%\label{theorem-new-tx-not-tip}
+%When a new transaction confirms an already confirmed transaction, the depth of all transactions remains the same, i.e., the depth of the transactions are only changed when a new transaction confirm an unconfirmed transaction.
+%\end{theorem}
 
-\begin{theorem}
-The height of a transaction is equal to the maximum height of its parents plus one.
-\end{theorem}
+%\begin{theorem}
+%The height of a transaction is equal to the maximum height of its parents plus one.
+%\end{theorem}
 
-\begin{theorem}
-If new transactions choose randomly the unconfirmed transactions to be confirmed, then no unconfirmed transaction will be left behind, i.e., all transactions will be confirmed in due time.
-\end{theorem}
+%\begin{theorem}
+%If new transactions choose randomly the unconfirmed transactions to be confirmed, then no unconfirmed transaction will be left behind, i.e., all transactions will be confirmed in due time.
+%\end{theorem}
 

### 2018-04-27 Chapters/hathor-architecture.tex
The incredible animal behavior of Dr Linhares
Hathor: reorganizing the text.

@@
+% !TEX root = ../partial-blockchain.tex
+
 This work introduces the Hathor's architecture which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
+
 Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG).
@@
 
-This way, there will be two difficulty levels in the system: one for new transactions which are just moving tokens around, and another one for ``blocks'' which are generating new tokens and collecting fees.
-The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions.
-The second is adjusted every 2,016 blocks to keep the pace of block on every 2 minutes.
+In Hathor, there are two difficulty levels: (i) one for new transactions which are just moving tokens around, and (ii) another one for ``blocks'' which are generating new tokens and collecting fees.
+The first may be adjusted to prevent spammers, which would spend too many resources to generate a great number of new transactions, whereas the latter is adjusted every 2,016 blocks to keep the pace of blocks on every 2 minutes.
 
-There is a trade-off about the required proof-of-work of new transactions.
+Both miners and users will be working on proof-of-work, decentralizing even more the network's hash rate.
+Even though the users' difficulty is less than the miners', the hash rate will increase every new user.
+The more transactions arrive, the higher the total hash rate.
+It may have good consequences in governance, which we will further discuss.
+
+There is a trade-off about the difficulty of new transactions.
 The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.
+This difficulty may even be increased when a spam attack is in course and reduced when it is gone.
+If the difficulty is too high, IoT devices may sign their transactions and send them to another devices which have a greater hash rate and will solve their proof-of-work faster.
 
-Each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
+It also seems interesting to have this difficulty depending on new transaction's size (in bytes) and amount being moved.
+The idea here is to require more work when high amounts are at stake.
+It would not affect IoT devices, which are expected to usually move smaller amounts.
+Regarding the transaction's size, requiring more work for bigger transactions may make sense because they may prevent abuses, such as a denial-of-service attack using enormous transactions which would consume a lot of node's bandwidth and disk space.
+
+Another important security matter is that each transaction has to confirm all its inputs, i.e., there must be a confirmation path between all the transactions of the inputs and the transaction which are spending them.
 It is always possible since there is at least one confirmation path between any transaction and a tip.
-This is important for security reasons.
+It ensures that, when a conflict is resolved, only the sub-DAG with root at the invalidated transaction will be affected.
+The remaining of the DAG remains the same.
 
@@
 They have already been confirmed a few times, but not enough to reach the minimum level required to be a confirmed transaction.
+For simplicity, the pending transactions encompass both in-progress and unconfirmed transactions.
 
-A ``block'' is just a regular transaction with no inputs which confirms a previous block and at least two in-progress transactions or tips.
+A \emph{block} is just a regular transaction with no inputs which confirms a previous block and at least two in-progress transactions or tips.
 There may be any number of outputs provided that they comply with the number of newly generated tokens.
 Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.
+The blocks are ordered according to their timestamp.
+If two blocks have the same timestamp, the block hash is used as a tiebreaker.
 
 In the low load scenario, there is a small number of new transactions coming into the network, which means they give a minor contribution to confirmations.
-In this case, the confirmation is held mostly by ``blocks''.
+In this case, the confirmation is held mostly by blocks.
 On the other hand, in the high load scenarios, there is a large number of new transactions giving a major contribution to confirmations.
-In this case, the ``blocks'' strengthen the confirmation, but most of them will have already been confirmed before the new blocks are found.
+In this case, the blocks strengthen the confirmations, but many of them will have already been confirmed before the next blocks are found.
 The higher the number of new transaction, the faster the transactions are confirmed.
@@
 
-The incentive scheme which keeps the network running is the same as Bitcoin's. Miners go towards fees and newly generate coins, while users go towards exchanging their tokens.
+The incentive scheme which keeps the network running is the same as Bitcoin's.
+Miners go towards fees and newly generate coins, whereas users just want to exchange their tokens.
+When there is no new transcation to be confirmed, the miners keep the network up and running while they find new blocks.
 
@@
 
-The concepts of weight and accumulated weight are the same as Iota's, but their equations are different.
-The weight of a transaction is calculated as $\log_2(k)$ where $k$ is the average number of hashes required to solve its proof-of-work again, while the accumulated weighted of a transaction A with weight $w_A$ is calculated as $\log_2(2^{w_A} + \sum_{A \leadsto P} 2^{w_P})$.
-The $\log$ function is used to reduce the magnitude of the numbers.
-Hence, the power-of-2 of the accumulated weight is a measure of the average number of hashes required to solve the proof-of-work of all transactions that are confirming A.
+Hathor uses similar concepts of weight and accumulated weight as Iota's. The weight depends only in the transaction itself, whereas the accumulated weight depends on its confirmations.
+
+The weight of a transaction is calculated as $w = \log_2(k)$ where $k$ is the average number of hashes required to solve its proof-of-work.
+
+The accumulated weight is the average number of hashes required to solve the proof-of-work of the transaction itself plus all the transactions which confirm it.
+Let $A$ be a transaction, its accumulated weight $w_A$ is calculated as $\log_2(2^{w_A} + \sum_{A \leadsto P} 2^{w_P})$.
 
@@
 
-The hash function used for the proof-of-work (PoW) is the same as Bitcoin: \emph{SHA-256} applied twice. Thus, most of the math analysis we have done before is just the same.
+The hash function used for the proof-of-work (PoW) is the same as Bitcoin: \emph{SHA-256} applied twice. Thus, most of the math analysis we have already done before is just the same.
 
@@
 
-Notice that, if the hash rate does not change, then $\Delta t = 128 \cdot 2016$ and $w_\text{new} = w_\text{old}$.
+Notice that, if the hash rate has not changed, then $\Delta t = 128 \cdot 2016$ and $w_\text{new} = w_\text{old}$.
 
@@
 
-$$w_\text{tx} = \left\lfloor \log_2(\text{size}) + \log_2(\text{amount}) + 0.5 \right\rfloor$$
+$$w_\text{tx} = \log_2(\text{size}) + \log_2(\text{amount}) + 0.5$$
 
-Although the transaction size may change a bit depending on the implementation, a typical transaction with 2 inputs and 2 outputs would have 188 bytes.
-So, for instance, transfering 50,000 tokens would require a weight of 23, which means an average of $2^{23}$ trials to solve the PoW.
+Although the transaction size depends on the implementation, a typical transaction with 2 inputs and 2 outputs would have approximately 188 bytes.
+So, for instance, transfering 50,000 tokens would require a weight of 23, which means an average of $2^{23}$ trials to solve the proof-of-work.
 
 \section{Issuance rate}
+
+Hathor issues tokens every block, thus it is similar to Bitcoin.
+Even so, an important decision is whether it will issue a limited number of tokens or not.
+Bitcoin chose to issue a limited number of tokens.
+On the other hand, Ethereum and others chose to issue an ilimited number of tokens.
+
+Both ways, the number of issued tokens is predictable.
+This feature itself brings confidence to the community, who will not face monetary intervention (unless they agree to) --- this comes in contrast with fiat money in which the central authority may change the monetary policy to adjust to political demands.
 
@@
 
-Using the same strategy, Hathor's issuance rate starts in 50,000 tokens per block and decreases over time.
+%Using the same strategy, Hathor's issuance rate starts in 50,000 tokens per block and decreases over time.
 Different from Bitcoin, Hathor's tokens are always integers and cannot be split.
@@
 
-\begin{align*}
-    \sum_{k=0}^{\infty} 1,000,000 \cdot \left\lfloor \frac{50000}{2^k} \right\rfloor \approx
-    \sum_{k=0}^{\infty} 1,000,000 \cdot \frac{50000}{2^k} = 100 \text{ billion}
-\end{align*}
+%\begin{align*}
+%    \sum_{k=0}^{\infty} 1,000,000 \cdot \left\lfloor \frac{50000}{2^k} \right\rfloor \approx
+%    \sum_{k=0}^{\infty} 1,000,000 \cdot \frac{50000}{2^k} = 100 \text{ billion}
+%\end{align*}
 
-As the average time between blocks is 128 seconds, it will take around 3 years, 9 months and 20 days between halves.
+%As the average time between blocks is 128 seconds, it will take around 3 years, 9 months and 20 days between halves.
 
@@
 
-The value of the fee is calculate as the difference between the transaction's outputs and the inputs.
-For instance, a transaction with inputs summing 10,000 HTR and outputs summing 9,000 HTR are paying 1,000 HTR of fees.
+In Bitcoin, the value of the fee is calculate as the difference between the transaction's outputs and the inputs.
+For instance, a transaction with inputs summing 8,000 and outputs summing 7,000 is paying 1,000 of fees.
 
-Each transaction may pay a fee to the next block which confirms it.
-But, even if the fee is zero, the miners are forced to confirm at least two pending transactions (in-progress or unconfirmed), and these pending transactions confirms other transactions which may have no fees.
+In Hathor, each transaction may pay a fee to the next block which confirms it.
+But, even if the fee is zero, the miners are forced to confirm at least two pending transactions.
+These two pending transactions also confirm other transactions which may have no fees.
 In summary, transactions with no fees cannot be left behind and will always be confirmed by both blocks and other transactions.
@@
 In the beginning, miners' incentive is driven by new tokens instead of fees.
-In the long term, it may be necessary to require a minimum fee to keep the miners working.
+In the long term, it may be necessary to require a minimum fee to keep the miners working, depending on the chosen issuance policy.
 
@@
 
-A transaction will be considered valid when it complies with the following rules: (i) it spends only unspent outputs; (ii) the sum of the inputs is greater than or equal to the sum of the outputs; (iii) the number of inputs is at most 256; (iv) the number of outputs is at most 256; (v) it confirms at least two pending transactions; (vi) it solves the PoW with the correct weight.
+A transaction will be considered valid when it complies with the following rules: (i) it spends only unspent outputs; (ii) the sum of the inputs is greater than or equal to the sum of the outputs; (iii) the number of inputs is at most 256; (iv) the number of outputs is at most 256; (v) it confirms at least two pending transactions; (vi) it solves the proof-of-work with the correct weight.
 
-The transaction has a time field which is used to inform when the transaction was generated.
-The time field must be in UTC time and may be within at most 60 minutes from the current time.
-Otherwise, the transaction will be discarded.
-
-In case of transaction conflict, in which two transactions try to spend the same tokens, the one with higher accumulated weight is chosen and the other is invalidated.
-Although it is not possible in Iota because of the submarine attack, Hathor does not have the same problem.
-In Hathor, like in Bitcoin, the submarine attack is only possible if the attacker has a hash rate higher than the whole network, including the miners.
-In other words, when analysing the double spending attack, Hathor is as safe as Bitcoin.
+The transaction has a timestamp field which is used to record when the transaction was generated.
+This timestamp field must be in UTC time to prevent timezone issues.
+It must also be within at most 5 minutes from the current time, otherwise the transaction will be discarded.
 
@@
 They give their customers a wallet address and they send the payment inside of the transaction itself.
-This allows IoT devices to save energy or to delegate the task of solving the PoW to whoever is receiving the tokens.
+This allows IoT devices to save energy, delegating the task of solving the proof-of-work.
+
+In case of transaction conflict, in which two transactions try to spend the same tokens, the one with higher accumulated weight is chosen and the other is invalidated.
+Although it is not a possible policy in Iota because of the submarine attack, Hathor does not have the same problem.
+In Hathor, like in Bitcoin, the submarine attack is only possible if the attacker has a hash rate higher than the whole network, including the miners.
+In other words, when analysing the double spending attack, Hathor is as safe as Bitcoin.
 
@@
 
-There is no orphan blocks in Hathor.
+Different from Bitcoin, there is no orphan blocks in Hathor.
 Unless a block confirms either directly or indirectly an invalid transaction, every block is valid.
-There is no need to left a block behind since its proof-of-work is also protecting transactions.
+There is no need to left a block behind since its proof-of-work is increasing transactions' accumulated weight.
 
-\section{Network}
+If two blocks are trying to collect fees from the same transactions, the one with higher timestamp will be discarded.
 
-Like in all other cryptocurrencies, nodes will communicate through a peer-to-peer network.
-They will exchange transactions and blocks, verifying whether they are valid or not, while they will also answer queries.
+\section{Governance}
 
-Thus, if a thin client wants to transfer some tokens, it just connects to the network and queries for the tips.
-With the response, it solves the proof-of-work and propagates the new transaction.
-In order to avoid malicious nodes, the thin client may ask to more than one node and randomly pick two transactions.
+- Punishment for miners/transactions which does not comply with the rules.
+
+- Rule change policy.
+
+\section{Expected number of tips}
+
+Explain transient and stationary state. The balance happens in the stationary state.
+
+In order to estimate the balance in the number of tips, let's consider the system in which the rate of inward tips is $\lambda_\text{TX}$, while the rate of outward tips is $\rho$.
+Let $K$ be the number of tips in the system.
+Thus, after $\Delta t$ seconds, $\Delta \text{tips} = \lambda_\text{TX} \Delta t - \rho K \Delta t$.
+In the balance, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
+
+Finally, the rate of outward tips is twice the rate in which devices solve the proof-of-work of transactions. Let $H$ be the devices' hash rate, then, $\rho = 2 \cdot H \cdot 2^{-w_\text{TX}}$.
+
+Twice?! It is a number between 1 and 2, but none of them.
+
+Therefore, in the balance,
+
+$$K = \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H}$$
+
+$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$
+
+\section{Attacks \& Conflicting transactions}
+
+\section{Scripts \& Contracts}

### 2018-05-03 Chapters/Chapter01.tex
Hathor governance and other minor changes.

@@
 It is said that Bitcoin is distributed because its ledger is public and is stored in thousands of computers.
-Even though the ledger is public, balances are anonymous and no one knows who owns which funds.
+Even though the ledger is public, balances are anonymous and no one knows who owns which funds\footnote{There are some techniques which may de-anonymize transactions in specific situations, even when users are using Tor network.
+For further information, see \citet{shentu2015research, biryukov2014deanonymisation, jawaheri2018small}.}.
 If an attacker tries to change anything, the remaining of the network is able to detect it and ignore the change.
@@
 
+% TODO After the year of 2140, there will be only fees to be collected.
 Bitcoin provides interesting incentives to all players, users and miners.
@@
 This spam problem has happened several times and affects Bitcoin's trustworthy.
-\citet{bitcoinspam2017} reports a possible spam attack lasting at least 18 months.
+\citet{bitcoinspam2017} reports a possible spam attack lasting at least contiguous 18 months.
 
@@
 It has neither mining nor confirmation blocks and transaction fees.
-Each transaction has its own proof-of-work and is used to confirm other transactions, forming a directed acyclic graph of transactions.
+Each transaction has its own proof-of-work\footnote{The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with transactions.
+It will be explained later in details.} and is used to confirm other transactions, forming a directed acyclic graph of transactions.
 Thus, a transaction is said to be confirmed when there is enough proof-of-work from the transactions confirming it directly or indirectly.
@@
 
-Next, we will do a mathematical analysis of Bitcoin in order to better understand its minings properties, how a fork would affect the network and its security against attackers.
+Next, I will do a mathematical analysis of Bitcoin in order to better understand its minings properties, how a fork would affect the network and its security against attackers.
+It is not necessary to do a mathematical analysis of Iota, because it has already been done in \citet{tangle2016}.
 

### 2018-05-03 Chapters/hathor-architecture.tex
Hathor governance and other minor changes.

@@
 
-- Punishment for miners/transactions which does not comply with the rules.
+In general, cryptocurrencies are decentralized, which means there is no central authority who decides its future, i.e., no one can enforce their will.
+Every decision must be accepted by its community, which means the community must agree.
+But what happens if they do not agree?
+When a consensus is not reached, the rules remain the same and the cryptocurrency may stall.
+The lack of a central authority may generate long debates, split the community, slow down strategic decision-making, and, ultimately, come to a ``civil war''---it is precisely what happened between Bitcoin Core (BTC) and Bitcoin Cash (BCH) in 2017 \citep{bloomberg2017civilwar, forbes2017civilwar}.
 
-- Rule change policy.
+Governance is an important part of a cryptocurrency because it must evolve, which means its community must agree into changing the rules.
+Governance is an agreement of how the community will proceed to change the rules.
+
+Despite the large literature available about governance, what separates Blockchain-based cryptocurrencies from them is the decentralization (against the hierarchical model).
+The number of papers about governance in decentralized cryptocurrencies has been growing, but it still lacks a solution.
+\citet{hacker2017corporate} resorts to the theory of complex systems and proposes a governance framework for decentralized cryptocurrencies, which is, in summary, a centralized coordination entity.
+\citet{hsieh2017internal} has analyzed the effects of governance in returns using panel data on several cryptocurrencies.
+They present a deeper discussion about the parts of a governance mechanism and concludes that
+
+\begin{quote}
+``...
+on the one hand, investors value cryptocurrencies’ core value proposition, rooted in decentralization; but on the other hand, are suspicious of decentralized governance at higher levels in the organization because they could slow down strategic decision-making (e.g., regarding the introduction of new innovations) or create information asymmetries between investors and technologists.''
+\end{quote}
+
+I believe that the solution to a good governance will come from financial incentives to all players to find a common ground.
+In Bitcoin, users have less bargaining power than miners because they do not contribute with work, whereas, in Hathor, both miners and users are working together.
+So, when it comes to changing the rules, the bargaining power is more distributed than in Bitcoin.
+The distribution depends on the ratio of the miners' hashpower and the users' hashpower.
+The higher the ratio, the closer to Bitcoin's governance.
+The lower the ratio, the higher the bargaining power of the users.
+
+%- Punishment for miners/transactions which does not comply with the rules.
+%- Rule change policy.
 

### 2018-05-03 Chapters/Chapter01.tex
Minor typo

--- 2018-04-27 Chapters/Chapter01.tex
@@
 According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending \citep{chaum1983blind, okamoto1995efficient, camenisch2005compact, canard2007divisible}.
-Some papers even propose solutions in a similar trajectory to Bitcoin, such as hash chain \citep{zongkai2004new} and preventing double spending using peer-to-peer networks \citet{osipkov2007combating, hoepman2007distributed}.
+Some papers even propose solutions in a similar trajectory to Bitcoin, such as hash chain \citep{zongkai2004new} and preventing double spending using peer-to-peer networks \citep{osipkov2007combating, hoepman2007distributed}.
 The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the previous e-cash philosophies.

### 2018-05-03 Chapters/hathor-bitcoin-math.tex
The incredible animal behavior of Dr Linhares

@@
-The primary objective of this chapter is to increase our understanding of the Bitcoin through mathematical tools.
+The primary objective of this chapter is to increase the understanding of the Bitcoin through mathematical tools.
 
@@
 
+Analyze when $k=2$ (for instance, Binance requires only two confirmations for deposits with less than $1,000$).
+
 %The number of blocks found in a given time interval $\Delta t$ is given by the Poisson distribution.
@@
 
-Scenario (i) is easy to analyze, because the network would just split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.
+Scenario (i) is easy to analyze, as the network would split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.
 

### 2018-05-04 Chapters/Chapter01.tex
Hathor typos and grammar fixes.

@@
     This is military-grade cryptography and full-blown paranoia.\\
-    Whoever did it hears voices whispering `they know everything.' \\
+    %Whoever did it hears voices whispering `they know everything.' \\
     \medskip
@@
 	\\ \medskip
-    --- Alex Linhares, EBAPE/FGV
+    --- Alexandre Linhares, EBAPE/FGV
 
@@
 As the money is digital, and copies can be made \textit{ad nauseam}, what can prevent counterfeiting?
-That is, what would prevent users from sending copies of the same money to two (or more) people?
+What would prevent users from sending copies of the same money to two (or more) people?
 That is precisely the problem solved by Bitcoin and its underlying Blockchain technology.
@@
 Hence, all other Blockchain-based cryptocurrencies have the same limitations, such as Litecoin, Bitcoin Cash, and Ethereum.
-One may arguee that increasing the maximum block size is a feasible solution to scaling, but I would say that it is just a temporary solution which buys some time until the next network saturation.
+One may argue that increasing the maximum block size is a feasible solution to scaling, but I would say that it is just a temporary solution which buys some time until the next network saturation.
 
@@
 Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior.
-They may also cooperate in an attack, avoiding transactions which seemed confirmed.
+They may also cooperate in an attack, voiding transactions which seemed confirmed.
 The more decentralized, the more trustworthy Bitcoin is.
-This centralization problem is seem as an important challenge.
+This centralization problem is seen as an important challenge.
 
@@
 In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain.
-Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}.
+Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the Directed Acyclic Graph (DAG) model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}.
 We are especially interested in one of them: Iota.
@@
 The coordinator will not be necessary anymore when the number of transactions per second reaches a minimum value, but Iota's developers cannot say precisely what is this minimum value.
-This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severily affect Iota's trustworthy).
+This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severily affect Iota's trustworthiness).
 

### 2018-05-04 Chapters/hathor-architecture.tex
Hathor typos and grammar fixes.

@@
 Both miners and users will be working on proof-of-work, decentralizing even more the network's hash rate.
-Even though the users' difficulty is less than the miners', the hash rate will increase every new user.
+Even though the users' difficulty is less than the miners', the hash rate will increase with every new user.
 The more transactions arrive, the higher the total hash rate.
-It may have good consequences in governance, which we will further discuss.
+This may have good consequences in governance, which we will further discuss.
 
-There is a trade-off about the difficulty of new transactions.
+There is a trade-off about the \textit{difficulty} of new transactions.
 The higher it is, the harder it is to generate new transactions, preventing spammers but also making it harder for IoT devices generate new transactions.
@@
 It would not affect IoT devices, which are expected to usually move smaller amounts.
-Regarding the transaction's size, requiring more work for bigger transactions may make sense because they may prevent abuses, such as a denial-of-service attack using enormous transactions which would consume a lot of node's bandwidth and disk space.
+Regarding the transaction's size, requiring more work for larger transactions may make sense because they may prevent abuses, such as a denial-of-service attack using enormous transactions which would consume a lot of node's bandwidth and disk space.
 
@@
 Miners go towards fees and newly generate coins, whereas users just want to exchange their tokens.
-When there is no new transcation to be confirmed, the miners keep the network up and running while they find new blocks.
+When there is no new transaction to be confirmed, the miners keep the network up and running while they find new blocks.
 
@@
 
-Explain transient and stationary state. The balance happens in the stationary state.
+It is intuitive that the number of tips depends on the number of transactions per second.
+The higher the number of transactions per second, the higher the number of tips.
+We can model the number of tips at time $t$ as a stochastic process, which means it changes over time following some probability distribution.
+Stochastic processes have two important states: transient and steady.
+A process is in transient state when its properties are changing over time, whereas it is in steady state when its properties has already converged.
 
-In order to estimate the balance in the number of tips, let's consider the system in which the rate of inward tips is $\lambda_\text{TX}$, while the rate of outward tips is $\rho$.
-Let $K$ be the number of tips in the system.
+Let's assume that the number of transactions per second is constant and does not change over time.
+Thus, at the beginning, there will be no tips, and the number will increase until it converges to its stable quantity.
+The process is in transient state until it reaches the stability.
+Then, it reaches the steady state.
+Notice that the number of tips also change in steady state, but its average does not change.
+It just floats around the average.
+
+If the number of transactions per second increases, the process returns to the transient state and moves towards the new steady state.
+In practice, the number of transactions per second is always changing, so is the steady state.
+
+In order to estimate the number of tips in the steady state, let's consider the process in which the rate of inward tips is $\lambda_\text{TX}$, while the rate of outward tips is $\rho$.
+Let $K$ be the number of tips at a given time.
 Thus, after $\Delta t$ seconds, $\Delta \text{tips} = \lambda_\text{TX} \Delta t - \rho K \Delta t$.
-In the balance, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
+In the steady state, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
 
 Finally, the rate of outward tips is twice the rate in which devices solve the proof-of-work of transactions. Let $H$ be the devices' hash rate, then, $\rho = 2 \cdot H \cdot 2^{-w_\text{TX}}$.
+%Twice?! It is a number between 1 and 2, but none of them.
 
-Twice?! It is a number between 1 and 2, but none of them.
-
-Therefore, in the balance,
+Therefore, in the steady state,
 
@@
 
-$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$
+%$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$
+For further information about the stochastic process of the total number of tips, see \citet{tangle2016}.
 
-\section{Attacks \& Conflicting transactions}
-
-\section{Scripts \& Contracts}
+%\section{Attacks \& Conflicting transactions}
+%\section{Scripts \& Contracts}

### 2018-05-04 Chapters/Chapter01.tex
Hathor: Alex's comments and minor fixes.

@@
 
-The concept of transfering money using cryptography as an underlying technology was shortly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}.
+The concept of transferring money using cryptography as an underlying technology was shortly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}.
 However, it was only in 1988 that \citet{chaum1988untraceable} created the term \emph{electronic cash} and also proposed a basic and practical scheme which yielded untraceability yet allowed to trace double spendings.
@@
 Bitcoin (BTC) is a digital currency, also known as digital money, internet money, and cryptocurrency.
-It is the first currency based on cryptography techniques which is distributed, decentralized, and with no central bank.
+It is the first currency based on cryptography techniques which are distributed, decentralized, and with no central bank.
 
@@
 It is said that Bitcoin is distributed because its ledger is public and is stored in thousands of computers.
-Even though the ledger is public, balances are anonymous and no one knows who owns which funds\footnote{There are some techniques which may de-anonymize transactions in specific situations, even when users are using Tor network.
+Even though the ledger is public, balances are anonymous, and no one knows who owns which funds\footnote{There are some techniques which may de-anonymize transactions in specific situations, even when users are using Tor network.
 For further information, see \citet{shentu2015research, biryukov2014deanonymisation, jawaheri2018small}.}.
@@
 Bitcoin is considered decentralized because there is no authority (or government) who decides its future.
-Every decision must be accepted by its community and no one can enforce their will.
+Every decision must be accepted by its community, and no one can enforce their will.
 Every change proposal must be submitted to the community who will discuss the matter and come to a verdict.
-If the majority of Bitcoin's community agrees on a decision, they just have to update their clearing process accordingly and the changes are applied.
+If the majority of Bitcoin's community agrees on a decision, they just have to update their clearing process accordingly, and the changes are applied.
 
@@
 % TODO After the year of 2140, there will be only fees to be collected.
-Bitcoin provides interesting incentives to all players, users and miners.
-On the one hand, users may have incentives to use Bitcoin because (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction ---; (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes Bitcoin's supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
+Bitcoin provides interesting incentives to all players (users and miners).
+On the one hand, users may have incentives to use Bitcoin because (i) the fees are small and do not depend on the amount being transferred --- but only in the size (in bytes) of the transaction ---; (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes Bitcoin's supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance.
 On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes, and they may also collect the fees of unconfirmed transactions.
@@
 The coordinator will not be necessary anymore when the number of transactions per second reaches a minimum value, but Iota's developers cannot say precisely what is this minimum value.
-This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severily affect Iota's trustworthiness).
+This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severely affect Iota's trustworthiness).
 
@@
 The simulator has been developed using an event-based design which is capable of running hours of simulation in just a few minutes.
-It creates agents who decide to make a transaction, then they select which transactions will be confirmed, then they spend some time working in the proof-of-work, and, finally, they propagate the transaction to the network.
+It creates agents who decide to make a transaction, then they select which transactions will be confirmed, next they spend some time working in the proof-of-work, and, finally, they propagate the transaction to the network.
 The other agents receive the transaction and may accept or deny it.
@@
 So, two new transactions may choose the same tips to confirm.
-If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.
+If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not been propagated yet.
 
-When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions.
-The depth-first search is interrupted when it reaches a transaction which the accumulated weight is bigger than a threshold.
-This interruption increases the overall performance of the simulation.
+When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} to update the aggregated weight of the directly and indirectly confirmed transactions.
+The depth-first search is interrupted when it reaches a transaction which the accumulated weight is larger than a given threshold.
+This interruption significantly increases the overall performance of the simulation.
+If the accumulated weight is an important metric, the whole network may be updated in specific times to get a measurement (instead of every new transaction).
 
@@
 
-The amount of time spent to solve the proof-of-work is sampled from an exponential distribution.
+Simulator's random variables simulator are all sampled from their distributions.
+The time between two transactions is sampled from an exponential distribution with $\lambda_\text{TX}$.
+The number of attempts to find a solution of the proof-of-work is sampled from a geometric distribution with $p=2^{w}$, where $w$ can be either $w_\text{TX}$ or $w_\text{block}$.
+The amount of time spent to solve the proof-of-work is calculated dividing the number of attempts by the hash rate of the device (which could be either a miner or an user).
+The time between blocks is just the amount of time spent to solve the proof-of-work with $w_\text{block}$.
 
-The time between two new transactions is also sampled from an exponential distribution with parameter $\lambda$.
-The value of $\lambda$ is changed over time, causing changes in the load of the network.
+Transactions (and blocks) do not have inputs, outputs, and scripts.
+They have only pointers to other transactions, which form the DAG.
+They also store their weight, accumulated weight, timestamp, and some statistics used for reports.
+
+New miners or users may be added or removed any time during a simulation.
+This allows the simulation of many different scenarios, such as increasing the number of miners, a sudden increase in the number of transactions per second, a sudden decrease in the number of miners, and so forth.
 
@@
 
+% TODO Include the parameters of the simulations.
+
 Hathor's architecture lies between Iota and Bitcoin's architectures.
 It is similar to Bitcoin's architecture when the number of transactions per second is low, while it is similar to Iota's architecture when the number of transactions per second is high.
+There is no ``switching'' between Bitcoin and Iota.
+It just behaves like one or another according to the network.
 In order to check this statement, I have performed some simulations.
 
-First, two simulations with the extreme cases: (i) no miners, (ii) no transactions.
-The first should have exactly the same behavior as Bitcoin's blocks with one block after the other forming a long chain.
-The latter should have a similar behavior as Iota's structure, forming a Direct Acyclic Graph (DAG) with transactions confirming transactions.
-We can notice in Figure \ref{fig:hathor-similarities} that both cases seem to be correct.
+First, two simulations in the extreme cases: (i) no miners, (ii) no transactions.
+The first should have precisely the same behavior as Bitcoin's blocks with one block after the other forming a long chain.
+The latter should have a similar behavior as Iota's, forming a Direct Acyclic Graph (DAG) with new transactions confirming previous ones.
+We can see in Figure \ref{fig:hathor-similarities} that both cases seem to be correct.
 
-Then, I have run a simulation with both miners and transactions.
+Next, I have run a more realistic simulation, with both miners and transactions.
 As we can see in Figure \ref{fig:hathor-dag}, Hathor's structure is a mix of Iota and Bitcoin's structure.
-The transactions are forming a DAG while, in parallel, the blocks are forming a chain.
-Differently from Bitcoin, there is no orphan block which means some blocks may be found close to the previous one.
+The transactions are forming a DAG while, in parallel, the blocks are forming a chain inside the DAG.
 
@@
 In this context, a transaction is said to be confirmed when it has reached an accumulated weight similar to six times the hash rate of the whole network (miners and new transactions).
-This criteria is equivalent to the well-known ``6 confirmations'' of Bitcoin, which is adopted by almost the whole network.
+This criteria is equivalent to the well-known ``6 confirmations'' of Bitcoin, which is adopted by almost the whole ecosystem.
 
@@
 In Figure \ref{fig:hathor-tct-low-mid}, we may see a good fit between the confirmation time of a transaction and the theoretical distribution of the time to find six blocks in Bitcoin (which is $Y_6$ and follows an Erlang distribution).
-When the load is increased, Hathor's confirmation time is reduced and diverges from Bitcoin's distribution.
-The reasoning is that confirmations coming from other transactions start to play an important role and accelerates the speed of confirmations.
+The blocks create ``maximum confirmation time'', since they are found with a precise pace, i.e., when there is not enough new transactions coming, the confirmation is done by the blocks.
+But, when the load is increased, Hathor's confirmation time is reduced and diverges from Bitcoin's time distribution.
+The reasoning is that confirmations coming from other transactions start to play an important role and accelerates the speed of confirmations, i.e., there is no need to wait for the next blocks, because the transactions are confirming themselves.
+This is what allows Hathor to scale and support higher volumes, indeed.
 
@@
 Finally, they proposed the lightning network, which is in development and should be available in the next months.
+I believe these proposals relieve the network---a temporary solution---, but do not solve the scalability problem.
 
-- What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
+Hathor's architecture allows a great number of transactions per second, since new transactions confirm previous ones (and there is no such thing as ``maximum block size'').
+The more transactions are coming, the faster previous transactions will be confirmed.
+It is the opposite of Bitcoin because the network benefits from high volume scenarios.
+
+As the transactions also have a proof-of-work, it becomes harder to performan a spam attack.
+The attacker would spend a considerable amount of computational resources to solve the proof-of-work of all transactions.
+The amount of work depends on the transaction's weight.
+Future work may explore automatic adjustments in transaction's weight to improve spam prevention.
+For instance, the network can detect a high number of new transactions coming and increase the transaction's weight for a while.
+Or else, the transaction's weight may be a function of the time between an output and its spending transaction, so, transferring the same tokens over and over in a small window of time would require a higher weight.
+
+%- What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?
 

### 2018-05-04 Chapters/hathor-architecture.tex
Hathor: Alex's comments and minor fixes.

@@
 
-This work introduces the Hathor's architecture which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
+This work introduces the Hathor's architecture, which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
 
-Like Iota, new transactions confirm others, forming a Directed Acyclic Graph (DAG).
+Like Iota, new transactions confirm previous ones, forming a Directed Acyclic Graph (DAG).
 For this, each transaction has its own proof-of-work which is solved by the issuer before propagating the transactions in the network.
@@
 It is always possible since there is at least one confirmation path between any transaction and a tip.
-It ensures that, when a conflict is resolved, only the sub-DAG with root at the invalidated transaction will be affected.
-The remaining of the DAG remains the same.
+This ensures that, when a conflict is resolved, only the sub-DAG with root at the invalidated transaction will be affected.
+The remaining parts of the DAG remains the same.
 
 The transactions are classified into three groups: (i) confirmed transactions, (ii) in-progress transactions, and (iii) unconfirmed transactions (tips).
-The confirmed transaction are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
+The confirmed transactions are the ones which have already been settled, i.e., their accumulated weights have reached a minimum level.
 The unconfirmed transactions (tips) are the brand new transactions which have not been confirmed even once yet, i.e., their accumulated weights are zero.
@@
 In this case, the blocks strengthen the confirmations, but many of them will have already been confirmed before the next blocks are found.
-The higher the number of new transaction, the faster the transactions are confirmed.
+The higher the number of new transactions, the faster the transactions are confirmed.
 The blocks assure a ``maximum confirmation time''.
@@
 
-There is a trade-off which must be considered in the weight of the transactions: the higher the weight, the better to prevent spams, but the worse to microtransactions.
+There is a trade-off which must be considered in the weight of the transactions: the higher the weight, the better to prevent spam, but the worse to microtransactions.
 So, trying to fulfill both necessities, the weight will be a function of the transaction's size (in bytes) and total amount:
@@
 
-In Bitcoin, the value of the fee is calculate as the difference between the transaction's outputs and the inputs.
+In Bitcoin, the value of the fee is calculated as the difference between the transaction's outputs and the inputs.
 For instance, a transaction with inputs summing 8,000 and outputs summing 7,000 is paying 1,000 of fees.

### 2018-05-08 Chapters/sdm-applications.tex
SDM Reinforcement learning citations and minor fixes.

@@
 What is incredible about these victories is that AlphaZero has almost no knowledge about chess game and has learned all its movement playing against itself for 4 hours.
-Basically, it knows only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm.
+Basically, it knows only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm \citep{silver2017mastering, silver2017masteringchess}.
 
@@
 How each type of reward should be handled depends on each algorithm.
+For further information, see \citet{sutton1998reinforcement}.
 
@@
 
-The weights of the rewards were chosen to prevent losing. Thus the weight of the negative reward was 5, while the weight of the positive reward was 2 and the neutral was 1.
+The weights of the rewards were chosen to prevent losing.
+Thus the weight of the negative reward was 5, while the weight of the positive reward was 2 and the neutral was 1.
+This choice are in accordance to Prospect Theory, which is a behavioral economic theory and states that losses hurt more than gains feel good \citep{kahneman2013prospect}.
 

### 2018-05-08 Chapters/Chapter02.tex
SDM Reinforcement learning citations and minor fixes.
sdm

@@
 \bigskip
-
-% TODO GARY KLEIN !!!
 
@@
 
-Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
+Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human long-term memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
 \begin{enumerate}
@@
 For instance, in 1998, \citet{lecun1998gradient} have developed algorithms which achieve from 88\% through a linear classifier, to 99.7\% through a convolutional net.
-For a review of algorithms' performance in the MNIST database, see \cite{mnist}.
+For a review of algorithms' performance in the MNIST database, see \cite{deng2012mnist}.
 

### 2018-05-10 defense-presentation.tex
Add defense presentation.

\documentclass{beamer}

\usetheme{Boadilla}

\title{Essays in Computational Management Science}
\author{Marcelo Salhab Brogliato}
\institute{EBAPE / FGV}
\date{\today}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Outline}
\begin{enumerate}[I]
\item Hathor: An alternative towards a scalable cryptocurrency
\item An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics
\item Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users
\end{enumerate}
\end{frame}

\part{Hathor: An alternative towards a scalable cryptocurrency}
\begin{frame}
\frametitle{Hathor: An alternative towards a scalable cryptocurrency}

The main contributions are:
\begin{enumerate}[i]
\item Mathematical analysis of Bitcoin
\item Proposal of Hathor
\item Analysis of Hathor
\end{enumerate}
\end{frame}

\part{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}
\begin{frame}
\frametitle{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}

The main contributions are:
\begin{enumerate}[i]
\item A validated, open-source, cross-platform framework
\item Identification and modeling of an autocorrelation
\item Noise filtering application
\item Classification application
\item Reinforcement learning application
\end{enumerate}
\end{frame}

\part{Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users}
\begin{frame}
\frametitle{Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users}

The main contributions are:
\begin{enumerate}[i]
\item Add a rejection parameter to the model
\item Estimate the parameters using only the number of active users
\item The possibility of reducing the number of active users (instead of only increase)
\end{enumerate}
\end{frame}

\end{document}

### 2018-05-10 Chapters/Chapter01.tex
Hathor: more results in Hathor's analysis.

@@
 This interruption significantly increases the overall performance of the simulation.
-If the accumulated weight is an important metric, the whole network may be updated in specific times to get a measurement (instead of every new transaction).
+If the accumulated weight of the whole DAG is an important metric, the whole DAG may be updated in specific times to get a measurement (instead of every new transaction).
 
@@
 This allows the simulation of many different scenarios, such as increasing the number of miners, a sudden increase in the number of transactions per second, a sudden decrease in the number of miners, and so forth.
+
+It is also possible to create metrics, which sample a statistic every $\Delta t$ seconds.
+There are two metrics available: (i) TipsMetric, which stores the number of tips at a given simulation time, and (ii) UtterlyAcceptanceMetric, which finds the new utterly accepted transactions and store how long it took.
+
+To find the network validated transactions, I run a breadth-first search (bfs) for each tip.
+Transactions visited in all searches are being confirmed by all tips and, by definition, are network validated transactions.
 
@@
 \end{figure}
+
+\section{Confirmation time}
 
@@
 Finally, even with both blocks and transactions, Hathor's blocks are similar to Bitcoin's blocks, and they share the same math.
-To confirm that, see Figure {fig:hathor-tbb}, where the red curve is Bitcoin's distribution of time between blocks and the blue histogram is Hathor's time between blocks.
+To confirm that, see Figure \ref{fig:hathor-tbb}, where the red curve is Bitcoin's distribution of time between blocks and the blue histogram is Hathor's time between blocks.
 I also made several tests adding and removing miners to check the difficulty adjustment and it worked properly.
@@
 \end{figure}
+
+\section{Load changing}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{./images01/new/many-loadings.pdf}
+
+\caption{DAG visualization when the loading is changed over time. \label{fig:hathor-load-changing}}
+\end{figure}
+
+\section{Number of tips}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Only low load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/tips_aggregate2.png}}
+
+\subfloat[All load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/tips_aggregate.png}}
+
+\caption{Histogram of the number of tips for different load scenarios. As expected, the number of tips increases with $\lambda_\text{TX}$. \label{fig:hathor-tips}}
+\end{figure}
+
+\section{Network validated transactions}
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Only low load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/nv_low.png}}
+
+\subfloat[All load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/nv_all.png}}
+
+\caption{Histogram of the time it takes for a transaction to be network validated. \label{fig:hathor-tips}}
+\end{figure}
+
+\section{Transaction's classification}
 

### 2018-05-10 Chapters/hathor-architecture.tex
Hathor: more results in Hathor's analysis.

@@
 For simplicity, the pending transactions encompass both in-progress and unconfirmed transactions.
+
+Another transaction classification concerns its validation by the network.
+A transaction is said to be network validated if there are confirming paths from all tips to the transactions, i.e., the whole network has validated that the transaction is valid.
+It is important to notice that a transaction may be network validated but still pending, or even be confirmed but not network validated.
 
@@
 %\section{Scripts \& Contracts}
+%\section{Bandwidth consumption}

### 2018-05-10 Chapters/Chapter02.tex
refs.

@@
 
+from call admission control \citep{kwon_atm_1998, hee-yong_kwon_atm_1997}, to behavior-based robotics \citep{Rajesh1998, mendes2008robot, jockel_sparse_2009}, to noise filtering \citep{meng_modified_2009} and many others
+
+There has recently been a surge of hardware implementations to minimize energy expenditure \citep{kang_energy-efficient_2015, kang_-memory_2016, li_hyperdimensional_2016, montagna_pulp-hd:_2018, salahuddin_energy_nodate}.
+
+see the recent work on modern machine learning and reinforcement learning \citep{aguiar_sdm-go:_2013, wu_kanerva_2018}
+
 Suddenly, you feel the danger.
@@
 
-Sparse Distributed Memory (SDM) \citep{Kanerva1988} (see also \citep{kanerva_sparse_1993, denning_sparse_1989} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
-This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from the subtlest of clues.
+Sparse Distributed Memory (SDM) \citep{Kanerva1988} (see also \citep{denning_sparse_1989,  flynn_sparse_1989, kanerva_sparse_1993, kanerva_parallel_1985, goos_binary_1996, kanerva_hyperdimensional_2009,  keeler_comparison_1988, rwcp_fully_1997, marinaro_spatter_1994, sahlgren_permutations_nodate, uesaka_foundations_2001}) is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+This model is used in all sort of applications due to its incredible ability to closely reflect the human capacity to remember past experiences from the subtlest of clues.
+Applications range from call admission control \citep{kwon_atm_1998, hee-yong_kwon_atm_1997}, to behavior-based robotics \citep{Rajesh1998, mendes2008robot, jockel_sparse_2009}, to noise filtering \citep{meng_modified_2009}, etc.
 

### 2018-05-10 Chapters/Chapter01.tex

@@
 
-\section{Load changing}
+\clearpage
+
+\begin{figure}[!htb]
+\centering\includegraphics[width=\textwidth]{./images01/sim/hathor-2.pdf}
+\caption{Visualization of a Hathor's graph with transactions and blocks.
+Red boxes are blocks; green circles are confirmed transactions; white circles are in-progress transactions; yellow circles are unconfirmed transactions (tips); and grey circles are transactions solving the proof-of-work which have not been propagated yet.
+The arrows show the confirmation chain.
+Block's arrows are in bold.
+\label{fig:hathor-dag-big}}
+\end{figure}
+
+\clearpage
+
+\section{Visualizing the network}
+
+Visualizing the network is not simple because the number of transactions and blocks is high, thus, arranging them and their edges is a non-trivial task.
+Therefore, most of the visualization are just part of the DAG which shows a window of time.
+
+To a better visualization of a Hathor's network, I run a simulation classifying the transactions in either confirmed, in-progress, or unconfirmed (tips).
+I also showed the transactions solving the proof-of-work which had not been propagated yet.
+See Figure \ref{fig:hathor-dag-big} and notice the chain of blocks inside the DAG.
+Confirmed transactions are in green circles, while in-progress transactions are in white circles and tips are in yellow circles.
+The blocks are in red boxes form a chain inside the DAG.
+Finally, the new transactions which are solving the proof-of-work and had not been propagated are in grey dashed circles.
+
+As new transactions have to chose two previous transactions to confirm, and just after they start to work in the proof-of-work, two new transactions eventually may chose the same tip because they do not know each other yet.
+So, if new transactions are coming in a low pace, the width of the swarm is small because the number of new transactions simultaneously solving the proof-of-work is also small.
+But, when new transactions are coming in a high volume, the width of the swarm increases because new transactions are choosing the same tip over and over.
+
+In summary, the width of the swarm depends on the number of transactions per second. The greater the number of new transactions per second, the larger the width of the swarm.
+
+To visualize the change in the width of the swarm, I run a simulation with a constant number of transactions per second, which was increased only in a specific window of time.
+The result can be seen in Figure \ref{fig:hathor-load-changing}, where the number of transactions per second suddenly increases, and the width increases until it reaches a stable value.
+Then, the number of transactions per second decreases, and the width returns to the previous value.
 
@@
 
+The number of tips at a given time also depends on the number of new transactions per second.
+The relationship between them was empirically measured in several simulations with different $\lambda_\text{TX}$.
+
+Analyzing Figure \ref{fig:hathor-tips}, it is easy to notice that both the average and the standard deviation of the number of tips increase when $\lambda_\text{TX}$ increases.
+According to \citet{tangle2016}, the average has the following equation:
+
+$$\frac{\lambda_\text{TX}}{\log(2)} \frac{2^{w_\text{TX}}}{H}$$
+
+In our simulation, $H = 100,000$ and $w_\text{TX} = 17$, so, the average number of tips should be equal to $1.89 \cdot \lambda_\text{TX}$.
+This model works best for low values of $\lambda_\text{TX}$, but diverges a little for higher values.
+For instance, if $\lambda_\text{TX} = 32$, the equation predicts $60.48$ tips on average yet we can check that the average is around $55$ tips.
+
+When $\lambda_\text{TX} \Rightarrow 0$, the number of tips goes towards one.
+The explanation is that there will be only one new transaction solving the proof-of-work per turn.
+So, new transactions will always confirm two tips, reducing the number of tips by one---each transaction confirms two tips and create one new one, hence the balance is -1.
+As it is impossible to have less to have less than one tip, it converges to one.
+
+When the number of tips is one, new transactions must chose a tip and another in-progress transaction. It should only happen in very low load scenarios, like during the launch of the network itself.
+
 \begin{figure}[!htb]
 \centering
-\subfloat[Only low load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/tips_aggregate2.png}}
+\subfloat[All load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/tips_aggregate.png}}
 
-\subfloat[All load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/tips_aggregate.png}}
+\subfloat[Zoom in the left side of (a)]{\includegraphics[width=\textwidth]{./images01/new2/tips_aggregate2.png}}
 
@@
 \caption{Histogram of the time it takes for a transaction to be network validated. \label{fig:hathor-tips}}
-\end{figure}
-
-\section{Transaction's classification}
-
-Afterward, to a better visualization of a Hathor's network, I run a simulation classifying the transactions in either confirmed, in-progress, or unconfirmed (tips).
-I also showed the transactions solving the proof-of-work which had not been propagated yet.
-See Figure \ref{fig:hathor-dag-big} and notice the chain of blocks inside the DAG.
-
-\begin{figure}[!htb]
-\centering\includegraphics[width=\textwidth]{./images01/sim/hathor-2.pdf}
-\caption{Visualization of a Hathor's graph with transactions and blocks.
-Red boxes are blocks; green circles are confirmed transactions; white circles are in-progress transactions; yellow circles are unconfirmed transactions (tips); and grey circles are transactions solving the proof-of-work which have not been propagated yet.
-The arrows show the confirmation chain.
-Block's arrows are in bold.
-\label{fig:hathor-dag-big}}
 \end{figure}

### 2018-05-10 Chapters/hathor-architecture.tex

@@
 Thus, at the beginning, there will be no tips, and the number will increase until it converges to its stable quantity.
-The process is in transient state until it reaches the stability.
-Then, it reaches the steady state.
-Notice that the number of tips also change in steady state, but its average does not change.
+The process is in transient state until it reaches stability.
+Then, it is in the steady state.
+Notice that the number of tips also change in steady state, but its average does not.
 It just floats around the average.
@@
 
-In order to estimate the number of tips in the steady state, let's consider the process in which the rate of inward tips is $\lambda_\text{TX}$, while the rate of outward tips is $\rho$.
-Let $K$ be the number of tips at a given time.
-Thus, after $\Delta t$ seconds, $\Delta \text{tips} = \lambda_\text{TX} \Delta t - \rho K \Delta t$.
-In the steady state, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
+\citet{tangle2016} has modeled the stochatic process of the total number of tips. It proves that the average number of tips, in the steady state, is:
 
-Finally, the rate of outward tips is twice the rate in which devices solve the proof-of-work of transactions. Let $H$ be the devices' hash rate, then, $\rho = 2 \cdot H \cdot 2^{-w_\text{TX}}$.
-%Twice?! It is a number between 1 and 2, but none of them.
-
-Therefore, in the steady state,
-
-$$K = \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H}$$
+$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{\log(2)H}$$
 
 %$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$
-For further information about the stochastic process of the total number of tips, see \citet{tangle2016}.
 

### 2018-05-11 Chapters/Chapter01.tex
Hathor: discussing the results.

@@
 In our simulation, $H = 100,000$ and $w_\text{TX} = 17$, so, the average number of tips should be equal to $1.89 \cdot \lambda_\text{TX}$.
-This model works best for low values of $\lambda_\text{TX}$, but diverges a little for higher values.
-For instance, if $\lambda_\text{TX} = 32$, the equation predicts $60.48$ tips on average yet we can check that the average is around $55$ tips.
+This model works best for low values of $\lambda_\text{TX}$ and diverges a little for higher values.
+For instance, if $\lambda_\text{TX} = 32$ tx/s, this equation predicts $60.48$ tips on average yet we can check in the empirical result that the average is around $55$ tips.
 
-When $\lambda_\text{TX} \Rightarrow 0$, the number of tips goes towards one.
+When $\lambda_\text{TX} \rightarrow 0$, the number of tips goes towards one.
 The explanation is that there will be only one new transaction solving the proof-of-work per turn.
@@
 
-When the number of tips is one, new transactions must chose a tip and another in-progress transaction. It should only happen in very low load scenarios, like during the launch of the network itself.
+When there is only one tip, new transactions must chose this only tip and another in-progress transaction.
+It should only happen in very low load scenarios, like during the launch of the network itself.
 
@@
 
+When all tips are confirming a transaction directly or indirectly, it is said that the transaction is network validated.
+It means that the whole network has checked the transactions and agrees that it is valid.
+
+When a transaction is network validated, all new transactions and blocks will confirm that transaction. Thus, its aggregated weight will increase as fast as possible.
+
+Let $\lambda_\text{TX}$ be the number of new transactions per second.
+If $\lambda_\text{TX}$ is constant, it means that, on average, there will be $\lambda_\text{TX} \Delta t$ new transactions after $\Delta t$ seconds (because the number of new transactions after $\Delta t$ seconds follows a Poisson distribution).
+All these transactions will confirm the network validated transactions.
+Hence, the number of transactions confirming a network validated transactions grows linearly.
+This result was also predicted by \citet[p.14]{tangle2016}.
+
+Suppose a transaction has just become network validated.
+Let $\text{acc}_0$ be its accumulated weight when it became network validated, $\eta$ be the average time between blocks, $w_\text{TX}$ be the average transaction weight, and $w_\text{BLK}$ be the average block weight.
+Then,
+
+\begin{align*}
+\text{acc}(\Delta t)
+&= \log_2 \left(2^{\text{acc}_0} + \lambda_\text{TX} \Delta t \cdot 2^{w_\text{TX}} + \left\lfloor \frac{\Delta t}{\eta} \right\rfloor 2^{w_\text{BLK}} \right) \\
+&= \text{acc}_0 + \log_2 \left(1 + \lambda_\text{TX} \Delta t \cdot 2^{w_\text{TX} - \text{acc}_0} + \left\lfloor \frac{\Delta t}{\eta} \right\rfloor 2^{w_\text{BLK} - \text{acc}_0} \right) \\
+&\simeq \text{acc}_0 + \log_2 \left(1 + \lambda_\text{TX} \Delta t \cdot 2^{w_\text{TX} - \text{acc}_0} + \frac{\Delta t}{\eta} 2^{w_\text{BLK} - \text{acc}_0} \right) \\
+&= \text{acc}_0 + \log_2 \left(1 + \lambda_\text{TX} \Delta t \cdot 2^{w_\text{TX} - \text{acc}_0} + \Delta t \cdot \eta^{-1} 2^{w_\text{BLK} - \text{acc}_0} \right) \\
+&= \text{acc}_0 + \log_2 \left[1 + \Delta t \cdot 2^{-\text{acc}_0} \cdot ( \lambda_\text{TX} 2^{w_\text{TX}} + \eta^{-1} 2^{w_\text{BLK}} ) \right]
+\end{align*}
+
+Therefore, after being network validated, the accumulated weight of a transaction grows logarithmically.
+
+In Figure \ref{fig:hathor-network-validated}, we can see how long it takes for a transaction to be network validated in different scenarios.
+The time is quite the same for $\lambda_\text{TX}$ below than one, but it changes for higher values.
+
+It is interesting to notice that there is a trade-off when $\lambda_\text{TX}$ increases.
+On the one hand, new transactions grow the DAG and accelerate the network validation, whereas, on the other hand, both the number of tips and the width of the swarm increases, dispercing this acceleration.
+The results show that, in fact, the time to be network validated increases with $\lambda_\text{TX}$.
+
+Anyway, for up to 32 tx/s (and $\eta=128$), it is reasonable to state that most transactions will be network validated after 35 seconds.
+
 \begin{figure}[!htb]
 \centering
+\subfloat[All load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/nv_all.png}}
+
 \subfloat[Only low load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/nv_low.png}}
 
-\subfloat[All load scenarios]{\includegraphics[width=\textwidth]{./images01/new2/nv_all.png}}
-
-\caption{Histogram of the time it takes for a transaction to be network validated. \label{fig:hathor-tips}}
+\caption{Histogram of the time it takes for a transaction to be network validated.
+A transaction is said to be network validated when all tips are confirming it directly or indirectly.
+\label{fig:hathor-network-validated}}
 \end{figure}

### 2018-05-11 Chapters/hathor-architecture.tex
Hathor: discussing the results.

@@
 
+A simple idea to get to this equation is through flow analysis: In steady state, we may say that the number of new tips must equal the number of tips being confirmed in a given time window.
+Thus, in order to estimate the number of tips in the steady state, let's consider the process in which the rate of inward tips is $\lambda_\text{TX}$, whereas the rate of outward tips is $\rho$.
+
+Let $K$ be the number of tips at a given time.
+Thus, after $\Delta t$ seconds, $\Delta \text{tips} = \lambda_\text{TX} \Delta t - \rho K \Delta t$.
+In the steady state, $\Delta \text{tips} = 0$, hence $K = \frac{\lambda_\text{TX}}{\rho}$.
+
+The rate of outward tips is proportional to the rate in which devices solve the proof-of-work of transactions.
+Let $H$ be the devices' hash rate, then, $\rho = \alpha \cdot H \cdot 2^{-w_\text{TX}}$, where $\alpha$ is the coefficient of proportionality.
+
+Finally,
+
+$$K = \frac{\lambda_\text{TX}}{\alpha} \frac{2^{w_\text{TX}}}{H}$$
+
+If new transactions would always confirm two tips, then $\alpha$ would equal 2.
+But, it may confirm one or even zero tips because concurrent new transactions may confirm their tips first.
+Thus, $0 < \alpha \le 2$.
+According to the equation found in \citet{tangle2016}, $\alpha = \log(2) = 0.693147$.
+
 %$$\frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{2H} \le K \le \frac{\lambda_\text{TX} \cdot 2^{w_\text{TX}}}{H}$$

### 2018-05-11 Chapters/Chapter02.tex
sdm done.

@@
 \bigskip
-
-from call admission control \citep{kwon_atm_1998, hee-yong_kwon_atm_1997}, to behavior-based robotics \citep{Rajesh1998, mendes2008robot, jockel_sparse_2009}, to noise filtering \citep{meng_modified_2009} and many others
-
-There has recently been a surge of hardware implementations to minimize energy expenditure \citep{kang_energy-efficient_2015, kang_-memory_2016, li_hyperdimensional_2016, montagna_pulp-hd:_2018, salahuddin_energy_nodate}.
-
-see the recent work on modern machine learning and reinforcement learning \citep{aguiar_sdm-go:_2013, wu_kanerva_2018}
 
@@
 Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.
-Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}:
+Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}\footnote{There has recently been a surge of hardware implementations to minimize energy expenditure and parallelize processes \citep{kang_energy-efficient_2015, kang_-memory_2016, li_hyperdimensional_2016, montagna_pulp-hd:_2018, salahuddin_energy_nodate}.
+While these implementations are extremely interesting, they do not afford the flexibility to experiment that software does.}:
 

### 2018-05-11 Chapters/hathor-bitcoin-math.tex
Changes in Hathor's bitcoin math, acknowledgments, and bass model.

@@
 
-\section{Fork analysis}
+%\section{Fork analysis}
 
@@
 
-When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork.
+%When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork.
 It is said to be a soft-fork when the rules are backward compatible, and a hard-fork when the rules are not backward compatible.
@@
 
-Suppose that the miners' are split in two groups, $G_1$ and $G_2$, with different rules.
+%Suppose that the miners' are split in two groups, $G_1$ and $G_2$, with different rules.
 Let's say $H_1$ and $H_2$ are their hash rate, respectively.
@@
 
-Scenario (i) is easy to analyze, as the network would split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.
+%Scenario (i) is easy to analyze, as the network would split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.
 
-Scenario (ii) is more trick.
+%Scenario (ii) is more trick.
 As $G_2$ accepts $G_1$ blocks, their hashrate will matter.
@@
 
-\begin{theorem}
-	When $H_1 > H_2$, the
-\end{theorem}
+%\begin{theorem}
+%	When $H_1 > H_2$, the
+%\end{theorem}
 

### 2018-05-11 FrontBackmatter/Acknowledgments.tex
Changes in Hathor's bitcoin math, acknowledgments, and bass model.

@@
 
-Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love. Especially, for understanding my absence.
+Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love.
+Especially, for understanding my absence and my unusual working times.
+Her loyalty and commitment make me the luckiest man alive.
+I love you.
 
@@
 His ideas have profoundly influenced my thoughts and my life since we have met.
-I must not forget his wife Paula, who became a dear friend and has always supported me as well, even when I was working in her living room until the sunrise.
+His optimism and trust in my abilities were undying.
+I must not forget his wife Paula, who became a dear friend and has always supported me as well, even when I was working in her living room until sunrise.
 Thank you for this amazing journey, my friends.
 
-Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
-I will never forget our all-you-can-eat adventures in Bros restaurant, the snacks in Catarina, and the drinks in Bigode bar.
+Thanks to my friends Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, and Andréia Sodré, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
+I will never forget our all-you-can-eat adventures in Bros restaurant, the snacks in Catarina, and the drinks in Bigode's bar.
 You have become close friends who I will carry for life.
 
-Thanks to my long-time partner and dear friend Claudio Abreu, who always encouraged me to take risks and to improve myself. He is the most fearless person I have ever met.
+Thanks to my long-time business partner and friend Claudio Abreu, who share a deep delight for engineering and has taught me so much about life, family, and business.
+He has always put faith in me and cheered for my victories.
+Thank you, my dear friend.
 
-Thanks to EBAPE / FGV for their great faculty, staff, and infrastructure.
+I thank my dear buddy and coauthor Daniel Chada for sharing the passion for research, always inspiring and amusing me with his astonishing ideas.
+Our long discussions have definitely shaped my researching abilities.
+Thank you, bro!
+
+Thanks to EBAPE / FGV for their finest faculty, staff, and infrastructure.
 A special thanks to Celene Silva Melo for her empathy and support besides all the help with paperwork, funding, and many other things.
-Another special thanks to doctor Luiz Antônio Joia, who trusted me in the first place and guided my first steps into the world of academy.
+Another special thanks to professor Luiz Antônio Joia, who trusted me in the first place and guided my first steps into the world of academy.
+
+Last, but far from least, I thank Kallen Givigi for being the ``Psych-os'' lab's protector and defender.
+She has always been there solving all sort of things (even personal matters) and whipping our lab into shape.
+Always kind and smiling, she gently asked for more focus and less mess.
+Your request is my command, my lady.
 

### 2018-05-11 Chapters/sdm-applications.tex

@@
 What is incredible about these victories is that AlphaZero has almost no knowledge about chess game and has learned all its movement playing against itself for 4 hours.
-Basically, it knows only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm \citep{silver2017mastering, silver2017masteringchess}.
+Basically, it knows only the valid movements and had to learn everything from scratch, which it did using a reinforcement learning algorithm\footnote{See also the recent work on modern machine learning and reinforcement learning \citep{aguiar_sdm-go:_2013, wu_kanerva_2018}} \citep{silver2017mastering, silver2017masteringchess}.
 

### 2018-05-11 Chapters/Chapter01.tex
Hathor: conclusion and minor fixes.

@@
 \chapter{Analysis of Bitcoin}
+\label{ch:hathor-bitcoin-math}
 \input{./Chapters/hathor-bitcoin-math.tex}
@@
 Today, Bitcoin network can barely handle 8 transactions per second without increasing the unconfirmed transaction list to hundreds of thousands --- several transactions take days to be confirmed.
-In order to increase Bitcoin's capacity, its community has first proposed and implemented segregated witness, which was not enough.
+In order to increase Bitcoin's capacity, its community has first proposed and implemented segregated witness, which improved scalability yet was not enough.
 Finally, they proposed the lightning network, which is in development and should be available in the next months.
@@
 It is the opposite of Bitcoin because the network benefits from high volume scenarios.
+As I have shown, Hathor seems to solve the scalability problem present in Blockchain-based cryptocurrencies.
 
-As the transactions also have a proof-of-work, it becomes harder to performan a spam attack.
-The attacker would spend a considerable amount of computational resources to solve the proof-of-work of all transactions.
-The amount of work depends on the transaction's weight.
+As the transactions also have a proof-of-work, it becomes harder to perform a spam attack.
+The attacker would spend a considerable amount of computational resources to solve the proof-of-work of every transaction, and the amount of work depends on the transaction's weight parameter.
 Future work may explore automatic adjustments in transaction's weight to improve spam prevention.
-For instance, the network can detect a high number of new transactions coming and increase the transaction's weight for a while.
-Or else, the transaction's weight may be a function of the time between an output and its spending transaction, so, transferring the same tokens over and over in a small window of time would require a higher weight.
+For instance, the network can detect a higher number of new transactions coming and increase the transaction's weight for a while.
+Or else, the transaction's weight may be a function of the time between an output being spent and its spending transaction, so, transferring the same tokens over and over in a small window of time would require more work.
+Anyway, the transaction's weight seems to tackle the spam issue.
+The new challenge is to set a proper transaction's weight which would prevent spam without impairing IoT devices.
+
+The last, but not least, challenge is the hashpower centralization.
+Although Bitcoin seems to have the most decentralized network among cryptocurrencies, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power \citep{gencer2018decentralization}.
+Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior.
+Hathor's architecture splits the hashpower among miners and users.
+Even if miners have more individual hashpower than users, because they would have rigs with appropriate cooling and energy supply, I believe their aggregate hashpower will not surpass users' aggregate hashpower when millions of devices are generating transactions.
+Future IoT devices may even come with an application-specific integrated circuit (asic) designed to solve Hathor's proof-of-work without spending too much battery.
+Future work may check common IoT processors' hashpower, which would allows us to estimate how many devices would be necessary to surpass miners' hashpower.
+
+Even though I have proposed to update block's weight every 24 hours (or 675 blocks), this was an arbitrary number.
+Future work may explore whether it would be feasible to continuously update block's weight, or what would be the optimal number of blocks between each update.
+I believe that the challenge of a continuous update approach would be preventing outdated nodes to discard valid blocks when two or more blocks were being propagated through Hathor's network.
+Maybe a solution would be allowing a range of block's weight instead of a single value, but future work would have to check whether this can be exploited by attackers.
+
+I also presented a mathematical analysis of Blockchain, going though mining, hashpower change, orphan blocks, and double-spending attacks.
+Most of the presented results may directly be applied to Hathor's blocks, since their foundations are the same.
+As \citet{tangle2016} has already analyzed Tangle, I have just applied their results with a few extensions.
+
+Future work may also further analyze other possibilities of attack in Hathor's network, such as malicious device not using random tip selection.
+Another major challenge affecting all cryptocurrencies is disk space use.
+How would one wipe out part of the blocks and transactions without putting security in risk?
+At first, all blocks and transactions are required to check whether anyone (including computer viruses) has tampered with transactions which had already been validated and stored in disk.
 

### 2018-05-11 Chapters/hathor-architecture.tex
Hathor: conclusion and minor fixes.

@@
 In Bitcoin, it is well-known that one should wait at least ``six confirmations'' before accepting a transaction.
-This Bitcoin's criteria is based on some math presented in Satoshi's seminal work \citep{nakamoto2008bitcoin}.
+This Bitcoin's criteria is based on some results presented in Satoshi's seminal work \citep{nakamoto2008bitcoin} and derived here in more detail in Chapter \ref{ch:hathor-bitcoin-math}.
 Adopting six confirmations is the same as demanding from attackers a minimal effort of six times the network's hash rate to successfully double spend those tokens.
@@
 
-For the blocks, $\eta = 128$, thus, $\mathbf{E}(T) = 128 \text{ seconds}$, and $\mathbf{V}(T) = 16,384$.
-The symmetrical confidence interval with $\alpha = 10\%$ is $[6.56, 383.45]$, i.e., 90\% of the cases the distance between blocks will be between 6 seconds and just under 7 minutes.
+For the blocks, $\eta = 128$, thus, $\mathbf{E}(T) = \eta = 128 \text{ seconds}$, and $\mathbf{V}(T) = \eta^2 = 16,384$, where $T$ is the random variable of the time between two consecutive blocks.
+The symmetrical confidence interval of $T$ with $\alpha = 10\%$ is $[6.56, 383.45]$, i.e., 90\% of the cases the distance between blocks will be between 6 seconds and just under 7 minutes.
 
@@
 
-This weight will be updated every 2,016 blocks to take into consideration the change of $H$.
+This weight will be updated every 675 blocks (which should happen every 24 hours) to take into consideration the change of $H$.
 First, $H$ will be estimated using the fact that $\mathbf{E}(Y_{2016}) = \frac{2016}{\lambda} = 2016 \eta = \frac{2016}{pH} = \frac{2016 \cdot 2^w}{H}$.

### 2018-05-11 FrontBackmatter/Acknowledgments.tex
Hathor: conclusion and minor fixes.

@@
 Her loyalty and commitment make me the luckiest man alive.
-I love you.
+I love you, Xu.
 
@@
 
-Thanks to my friends Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, and Andréia Sodré, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
+Thanks to my friends Daniel Chada, Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, and Andréia Sodré, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.
 I will never forget our all-you-can-eat adventures in Bros restaurant, the snacks in Catarina, and the drinks in Bigode's bar.
-You have become close friends who I will carry for life.
+You have become very close friends who I will carry for life.
 

### 2018-05-11 Chapters/hathor-bitcoin-math.tex
Hathor: finished \o/

@@
+% !TEX root = ../partial-blockchain.tex
+
 The primary objective of this chapter is to increase the understanding of the Bitcoin through mathematical tools.
@@
 
+\subsection{Hash rate suddenly changing}
+
 Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
@@
 
-It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a \Rightarrow \int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
+It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a$, which yields $\int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
 $$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
@@
 
-Thus,
+Let $\mathcal{L}$ be the Laplace Transform. Thus,
 \begin{align*}
@@
 When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied.
-In fact, $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.
+In fact, $b \rightarrow +\infty$ yields $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.
 
@@
 
-For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$.
+For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01$, which yields $b > 0.000163690$ and $n < 10.1818$.
 So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected.
@@
 
-The function $\mathbf{E}(T)|_{t_1}^{t_2}$ may be used to a piecewise linear analysis of any hash rate change.
-Let's analyze the hash doubling in one week.
-Then, $u(1 \text{week}) = u(604800) = 2$, thus $a^2 = \frac{1}{604800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$.
-Let's sample the interval $[0, 1 \text{week}]$ every hour, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.
+\subsection{Comparison of the models}
 
-For each point $t_i$, let $g_i(t) = [H(t-t_i) - H(t-604800)] u(t) + 2 H(t-604800)$, then $\mathbf{E}(T) = \mathbf{E}(T)|_{t_i}^{t_{i+1}} + 1 - e^{\frac{t}{\eta}}$. The result is presented in Figure .
+In order to compare the hash rate change models, namely (i) suddenly change, (ii) smoothly change, and (iii) linear change, I have applied each of them to the same scenarios.
+In the first scenario, the hashrate will double in the next week, whereas, in the second scenario, it will halve in the next week.
+
+In both the smoothly change model and the linear change model, I could have calculated the average time between blocks during the whole week.
+It would not give us much information, because the time between models would be increasing (or reducing) more and more as the days goes by.
+And we are really interested in the average time between blocks throughout the days, and not the average of the whole week.
+
+Thus, I have analyzed a piecewise hash rate change, i.e., I have calculated the average time between blocks for each hour throughout the week.
+First, I split the whole week into $24 \cdot 7$ intervals, $(t_0, t_1, t_2, \dots, t_168)$, where $t_i = 3600i$.
+Then, I calculated the average for each interval $(t_k, t_{k+1})$.
+Let $H_k^0$ and $H_k^1$ be the initial and final hash rate of the $(t_k, t_{k+1})$ interval.
+So, I also ensured the continuity of the hash rate between consecutive intervals, i.e., $H_k^1 = H_{k+1}^0$.
+
+I compared both the smoothly change model and the linear change model with the suddenly change model.
+The difference between them is negligible.
+Let $\epsilon$ be the maximum absolute error between the models, than $\epsilon < 0.8$ and $\epsilon/H < 0.2\%$, for all intervals.
+The maximum absolute error between the linear and the suddenly change models can be seen in Figure \ref{fig:bitcoin-H-change-error}.
+
+\begin{figure}[!htb]
+\centering \includegraphics[width=0.7\textwidth]{./images01/H-max-abs-error.png}
+
+\caption{Maximum absolute error between the linear and the suddently change models. \label{fig:bitcoin-H-change-error}}
+\end{figure}
+
+Therefore, we may conclude that it is reasonable to approximate the average time between blocks using only the suddenly change model in each interval of one hour.
+
+The average time between blocks throughout the days can be seen in Figure \ref{fig:bitcoin-H-change}.
+It was calculated using the suddenly change model with the hash rate changing linearly during the week.
+
+\begin{figure}[!htb]
+\centering
+\subfloat[Doubling the hash rate]{\includegraphics[width=\textwidth]{./images01/H-double-week.pdf}}
+
+\subfloat[Halving the hash rate]{\includegraphics[width=\textwidth]{./images01/H-halves-week.pdf}}
+
+\caption{The average time between blocks when the hash rate changes over time. \label{fig:bitcoin-H-change}}
+\end{figure}
 
@@
 
-\begin{figure}[ht]
-\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
-\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
-\end{figure}
+%\begin{figure}[ht]
+%\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
+%\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
+%\end{figure}
 
@@
 
-\citet{karame2012two}
-
-There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.
+There are many possible ways to attack the Bitcoin \citep{karame2012two, heilman2015eclipse, bahack2013theoretical, bonneau2016buy, neudecker2015simulation}.
+In this section, we are interested in a particular attack: the double spending attack.
 
@@
 
-Analyze when $k=2$ (for instance, Binance requires only two confirmations for deposits with less than $1,000$).
+% TODO Analyze when $k=2$ (for instance, Binance requires only two confirmations for deposits with less than $1,000$).
 
@@
 If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block.
+
 Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$.
-The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$.
-Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$.
-The probabilities form a simple geometric series, and so they are exponentially decreasing.
-We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite.
-In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\eta$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$.
-Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
-When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.
+In this case, the average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$.
 
-The average waiting time of a transaction to be confirmed, when $m < s$, is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.
+When $s > m$, the probabilities $\pi_n$ form a simple geometric series with common ratio smaller than one, which means the probabilities are exponentially decreasing.
+Since $\pi_n \rightarrow 0$ when $n \rightarrow \infty$, we may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite.
 
-When $m \ll s$, $z_s \rightarrow 1 + 1/m$.
-So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$.
-In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.
+When $s \le m$, the system is unstable, which means the unconfirmed transactions pool size keeps growing towards infinity.
+In this case, the system is not capable of processing the demand for a long period of time.
 
-When $\lambda_{\text{TX}} \rightarrow s$, $z_s \rightarrow 1$. So, $\mathbf{E}(\pi) \rightarrow +\infty$.
+Using the fact that $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $\lambda_{\text{blocks}} = 1/\eta$, the stability condition $s > m$ is reached when $\lambda_{\text{TX}} < s/\eta$.
 
-Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
+In the Bitcoin's network, the average number of transactions per block is $s=2,250$, so, the system is stable when $\lambda_{\text{TX}} < 2,250/600 = 3.75$ tx/s.
+Therefore, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
+When $\lambda_{\text{TX}} > 3.75$ tx/s, the unconfirmed transaction pool starts to grow indefinitely.
+
+When the system is stable, the average waiting time of a transaction to be confirmed is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.
+
+$m \ll s$ yields $z_s \rightarrow 1 + 1/m$.
+Thus, the average number of unconfirmed transactions $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$.
+In the Bitcoin's network, $m \ll s$ is reached when $\lambda_{\text{TX}} \ll 3.75$ tx/s.
+In other words, when the number of new transactions per second is way smaller than 3.75 tx/s, the average waiting time of a transaction to be confirmed is 600 seconds, which means, on average, all transactions will be confirmed in the next block.
+
+But, $\lambda_{\text{TX}} \rightarrow s/\eta$ yields $z_s \rightarrow 1$. Hence, $\mathbf{E}(\pi) \rightarrow +\infty$, which means the system is going towards instability.
+
+Therefore, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
+
+For instance, in order to be a stable system and process 15 transactions per second, each block would have to confirm, on average, 9,000 transactions. Bitcoin's network is really far from this point.
 

### 2018-05-11 Chapters/Chapter01.tex
small preciosismis...

@@
 
-Bitcoin's underlying technology blockchain has been called by many as a major invention, even comparable to the invention of the internet.
+Bitcoin's underlying technology, blockchain, has been called by many as a major invention, even comparable to the invention of the internet.
 But it is unlikely that Bitcoin and blockchain have achieved the final or most optimal design for a secure and scalable electronic transaction system.
@@
 The last, but not least, challenge is the hashpower centralization.
-Although Bitcoin seems to have the most decentralized network among cryptocurrencies, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power \citep{gencer2018decentralization}.
+Although Bitcoin seems to have the most decentralized network among cryptocurrencies, there are few miners and mining pools which \emph{together} control over 50\% of the network’s computing (hash)power \citep{gencer2018decentralization}.
 Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior.

### 2018-05-21 Chapters/Chapter01.tex
Typos typos...

@@
 As there are no miners, there are also no fees, which is a major incentive to newcomers.
-Tokens may be freely transfered without any loses, even for tiny amounts, enabling micropayments.
+Tokens may be freely transferred without any losses, even for tiny amounts, enabling micropayments.
 For example, online workers paid per hour may receive their payments every hour, reducing the default risk.
@@
 
-Simulator's random variables simulator are all sampled from their distributions.
+Simulator's random variables are all sampled from their distributions.
 The time between two transactions is sampled from an exponential distribution with $\lambda_\text{TX}$.
@@
 
-First, two simulations in the extreme cases: (i) no miners, (ii) no transactions.
+First, two simulations in the extreme cases: (i) no transactions, (ii) no miners.
 The first should have precisely the same behavior as Bitcoin's blocks with one block after the other forming a long chain.
@@
 So, new transactions will always confirm two tips, reducing the number of tips by one---each transaction confirms two tips and create one new one, hence the balance is -1.
-As it is impossible to have less to have less than one tip, it converges to one.
+As it is impossible to have less than one tip, it converges to one.
 
@@
 In Figure \ref{fig:hathor-network-validated}, we can see how long it takes for a transaction to be network validated in different scenarios.
-The time is quite the same for $\lambda_\text{TX}$ below than one, but it changes for higher values.
+The time is quite the same for $\lambda_\text{TX}$ less than one, but it changes for higher values.
 
 It is interesting to notice that there is a trade-off when $\lambda_\text{TX}$ increases.
-On the one hand, new transactions grow the DAG and accelerate the network validation, whereas, on the other hand, both the number of tips and the width of the swarm increases, dispercing this acceleration.
+On the one hand, new transactions grow the DAG and accelerate the network validation, whereas, on the other hand, both the number of tips and the width of the swarm increases, dispersing this acceleration.
 The results show that, in fact, the time to be network validated increases with $\lambda_\text{TX}$.

### 2018-05-21 Chapters/Chapter02.tex
Typos and Jamil's comments.

@@
 There is obviously a demand for use of SDM; but each group has been tied to their own \emph{ad-hoc} needs, and there has not been the emergence of a community centered on a tool.
-It is our belief that a tool such as standard open-source framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Juypyter notebooks.
+It is our belief that a tool such as standard open-source framework could bring orders of magnitude more researchers and attention if they were able to use the model, at zero cost, with an easy to use high-level language such as Python, in an intuitive platform such as Jupyter notebooks.
 Neuroscientists interested in long-term memory storage should not have to worry about high-bandwidth vector parallel computation.
@@
 \chapter{Conclusion}
+
+% Jamil: De maneira geral, achei muito ensaístico e pouco "paper". Me parece que vai precisar de retrabalho para submissão se encarar um reviewer mais tradicional. Eu gostei, no entanto.
 

### 2018-05-21 Chapters/hathor-architecture.tex
Minor grammar mistakes.

@@
 
-This work introduces the Hathor's architecture, which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
+This work introduces Hathor's architecture, which lies between Bitcoin's and Iota's and may be a solution to scaling, centralization, and spam issues.
 

### 2018-06-14 defense-presentation.tex
Defense presentation (draft)

@@
 \documentclass{beamer}
+
+\PassOptionsToPackage{utf8}{inputenc}
+    \usepackage{inputenc}
+
+\usepackage{graphicx}
 
@@
 
+\setlength{\parskip}{\baselineskip}
+
 \begin{document}
@@
 \titlepage
+\end{frame}
+
+\begin{frame}
+\frametitle{Introduction}
+
+Modern management and high technology interact in multiple, profound, ways. A whole new ecosystem seems to have emerged within computing and business.
+
+The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI, and cognitive technology \emph{as an executive}, not as a research scientist or specialized engineer.
+
+Professor Andrew Ng tells students at Stanford's Graduate School of Business that ``AI is the new electricity'', as his hyperbolic way to emphasize the potential transformational power of the technology.
+\end{frame}
+
+\begin{frame}
+\frametitle{Introduction}
+
+Both iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created.
+
 \end{frame}
@@
 
+The problems: lack of scalability, centralization, and spam attacks.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Cryptocurrency issues}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{./images-defense/bitcoin-congestion.png}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Contributions}
+
 The main contributions are:
@@
 \end{enumerate}
+\end{frame}
+
+\begin{frame}
+\frametitle{Bitcoin}
+
+The primary problem for creating digital money is how to prevent double spending.
+
+As the money is digital, and copies can be made \emph{ad nauseam}, what can prevent counterfeiting? What would prevent users from sending copies of the same money to two (or more) people?
+
+The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the 30-year literature on e-cash.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Blockchain (1)}
+
+Each attempt to find a new block is equivalent to a random sample from a discrete uniform distribution in the interval $[0, 2^{256}-1]$.
+A new block is only found when the sample is less than a given number $A$.
+
+The mining process is a sequence of failed attempts followed by a successful attempt.
+Let $X$ be the number of attempts until finding a new block.
+Then, $X$ follows a geometric distribution with probability $p = \frac{A}{2^{256}}$.
+
+Let $H$ be the network hashpower, i.e., the number of attempts per second that the network is able to calculate. Thus, the time to find a new block (or the time between blocks) is $T = X/H$.
+
+As $H$ changes over time, the given number $A$ is adjusted so that the network would find a block every $\eta$ seconds, which leads to $P(T \le t) = 1 - \left(1 - \frac{1}{\eta H} \right)^{tH}$.
+\end{frame}
+
+\begin{frame}
+\frametitle{Bitcoin's hashpower history}
+\begin{figure}
+\includegraphics[width=\textwidth]{./images-defense/bitcoin-H-history.png}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Blockchain (2)}
+
+\begin{theorem}
+    For $H$ large enough, the time between blocks may be approximated by an exponential distribution, i.e., $P(T \le t) = 1 - e^{t/\eta}$.
+\end{theorem}
+
+\begin{theorem}
+	The maximum absolute error when approximating the time between blocks by an exponential distribution is $e / H$.
+\end{theorem}
+
+So, for a maximum absolute error of $0.01 \% = 10^{-4}$, the network hashpower $H$ must be at least 26 kH/s.
+
+For comparison, Bitcoin's hashpower is around 35,000,000 TH/s = 35,000,000,000,000,000 kH/s = $35 \cdot 10^{18} H/s$.
+In this case, the maximum absolute error is $7.7 \cdot 10^{-20} = 0.0000000000000000000776$.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Blockchain (3)}
+
+Let $Y_n = \sum_{i=1}^n T_i$ be the total time to find $n$ consecutive blocks. From probability literature, $Y_n$ follows an Erlang distribution.
+
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{./images01/time-6-blocks.pdf}
+\caption{PDF of $Y_6$, i.e., probability of finding 6 consecutive blocks.}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Attack in the Bitcoin network}
+
+\begin{theorem}
+	Let $\beta$ be the percentage of the hashpower controlled by the attacker, and $\gamma$ the percentage of the hashpower without the attacker. Let $p = \frac{\gamma}{\beta + \gamma}$. Then,
+$$
+\mathbf{P}(\text{successful attack}) =
+\begin{cases}
+	1 - \sum_{s=0}^{k-1} \binom{k+s-1}{s} \left( (1-p)^s p^k - (1-p)^k p^s \right) \text{, $p \geq 0.5$} \\
+	1 \text{, $\gamma < \beta$}.\\
+\end{cases}
+$$
+\end{theorem}
+
+For $k=6$, $p=0.9$, $\mathbf{P}(\text{successful attack}) = 0.0005914121600000266$.
+
+For $k=6$, $p=0.7$, $\mathbf{P}(\text{successful attack}) = 0.15644958192000014$.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Attack in the Bitcoin network}
+\begin{figure}
+\includegraphics[width=0.75\textwidth]{./images01/fig-bitcoin-attack.png}
+\caption{Probability of a successful attack according to the network's hash rate of the attacker ($\beta$).}
+\end{figure}
 \end{frame}
@@
 \item A validated, open-source, cross-platform framework
-\item Identification and modeling of an autocorrelation
+\item Identification and modeling of the autocorrelation in the counters
 \item Noise filtering application
@@
 
+\begin{frame}
+\frametitle{Conclusion}
+\end{frame}
+
 \end{document}

### 2018-06-14 defense-presentation.tex
Defense presentation updated

@@
 \usepackage{graphicx}
+\usepackage{subfig}
+\usepackage[export]{adjustbox}
 
@@
 \begin{frame}
-\frametitle{Introduction}
-
-Both iPhone (and competitors) and Bitcoin (and competitors) have created giant platforms on top of which immense wealth has been created.
-
-\end{frame}
-
-\begin{frame}
 \frametitle{Outline}
@@
 
-The problems: lack of scalability, centralization, and spam attacks.
+The primary problem for creating digital money is how to prevent double spending.
+
+As the money is digital, and copies can be made \emph{ad nauseam}, what can prevent counterfeiting? What would prevent users from sending copies of the same money to two (or more) people?
+
+The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the 30-year literature on e-cash.
 
@@
 \begin{frame}
-\frametitle{Cryptocurrency issues}
+\frametitle{Bitcoin}
 
 \begin{figure}
-\includegraphics[width=\textwidth]{./images-defense/bitcoin-congestion.png}
+\includegraphics[width=0.9\textwidth]{./images-defense/bitcoin-blockchain.png}
 \end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Problems of Bitcoin}
+
+Lack of scalability (up to 3.7 tx/s for 1MB per block)? Spam attacks?
+
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{./images-defense/bitcoin-congestion.png}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Iota}
+
+\begin{figure}
+\centering\includegraphics[width=0.8\textwidth]{./images01/fig-tangle-example.pdf}
+\caption{White nodes represent transactions that have been confirmed at least once.
+Green circles represent unconfirmed transactions (tips).
+Gray and dashed nodes are the transactions currently solving the proof-of-work in order to be propagated.\label{fig-tangle-example}}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Problems of Iota}
+
+Iota does not work in low load scenarios. It uses a central coordinator to prevent attacks.
+
+They does not know what is the minimum number of transactions per seconds so that they can disable the central coordinator.
 
@@
 \item Mathematical analysis of Bitcoin
-\item Proposal of Hathor
+\item Proposal of Hathor, \small{an architecture which seems to work in both low and high load scenarios.}
 \item Analysis of Hathor
@@
 \begin{frame}
-\frametitle{Bitcoin}
+\frametitle{Hathor's architecture}
 
-The primary problem for creating digital money is how to prevent double spending.
-
-As the money is digital, and copies can be made \emph{ad nauseam}, what can prevent counterfeiting? What would prevent users from sending copies of the same money to two (or more) people?
-
-The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the 30-year literature on e-cash.
+\begin{figure}
+\centering\includegraphics[width=0.45\textwidth]{./images01/sim/hathor-2.pdf}
+\end{figure}
 
@@
 \begin{frame}
-\frametitle{Blockchain (1)}
+\frametitle{Mining the next block}
 
@@
 \begin{frame}
-\frametitle{Bitcoin's hashpower history}
-\begin{figure}
-\includegraphics[width=\textwidth]{./images-defense/bitcoin-H-history.png}
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{Blockchain (2)}
+\frametitle{Mining the next block (2)}
 
@@
 \begin{frame}
-\frametitle{Blockchain (3)}
+\frametitle{Bitcoin's hashpower history}
+\begin{figure}
+\includegraphics[width=\textwidth]{./images-defense/bitcoin-H-history.png}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Mining consecutive blocks}
 
@@
 \end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{The double-spending attack}
+
+In the double spending attack, the attacker's send some funds to the victim, let's say a merchant.
+
+They wait for $k$ confirmations of the transaction, and the victim delivers the good or the service to the attacker.
+
+Then, the attacker mine enough blocks with a conflicting transaction, double spending the funds which was sent to the victim.
+
+If the attacker is successful, the original transaction will be \textit{erased} and the victim will be left with no funds at all.
 
@@
 
+\begin{frame}
+\frametitle{Analysis of Hathor: boundary cases}
+
+\begin{figure}
+\centering
+\subfloat[No miners]{\includegraphics[width=0.5\textwidth]{./images01/sim/no_miners.pdf}}
+\subfloat[No transactions]{\includegraphics[width=0.5\textwidth]{./images01/sim/only_miners.pdf}}
+\caption{Visualization of a Hathor's graph in two particular cases: (a) no miners, (b) no transactions.}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Analysis of Hathor: confirmation time}
+\begin{figure}
+\centering
+\subfloat[Low load]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-low-load.png}}
+\subfloat[Mid load]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-mid-load.png}}
+
+\caption{Confirmation time in two scenarios: (a) low load, (b) mid load. The red curve is the distribuion of the time to find six blocks in Bitcoin (which follows an Erlang distribution).}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Analysis of Hathor: confirmation time (2)}
+\begin{figure}
+\includegraphics[width=0.75\textwidth]{./images01/sim/tct-many-loads-2.png}
+
+\caption{Confirmation time in many scenarios, moving from a low load ($\lambda_\text{TX} = 0.015625$) to a high load ($\lambda_\text{TX} = 2$). \label{fig:hathor-tct-many}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Analysis of Hathor: load changing over time}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{./images01/new/many-loadings.pdf}
+\caption{DAG visualization when the loading is changed over time.}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Analysis of Hathor: number of unconfirmed transactions}
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{./images01/new2/tips_aggregate.png}
+\caption{Histogram of the number of tips for different load scenarios.}
+\end{figure}
+\end{frame}
+
 \part{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}
@@
 
+SDM is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+
+This model is used in all sort of applications due to its ability to closely reflect the human capacity to remember past experiences.
+
+This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{SDM: How does it work?}
+
+\begin{figure}
+\centering\includegraphics[scale=0.75]{./images02/p_circle_r.pdf}
+\caption{Activated addresses inside access radius $r$ around the center address.}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{SDM: How does it work?}
+\begin{figure}
+\centering\includegraphics[scale=0.75]{./images02/p1_inter_p2.pdf}
+\caption{Shared addresses between the target datum $\eta$ and the cue $\eta_{x}$.}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{SDM: How does it work?}
+\begin{figure}
+\centering\includegraphics[scale=0.75]{./images02/p1_p2_iter_read.pdf}
+\caption{In this example, four iterative readings were required to converge from $\eta_{x}$ to $\eta$.}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{SDM: How does it work?}
+\begin{figure}
+\centering\includegraphics[scale=0.75]{./images02/p1_after_write.pdf}
+\caption{Hard-locations pointing, approximately, to the target bitstring.}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Contributions}
+
 The main contributions are:
 \begin{enumerate}[i]
-\item A validated, open-source, cross-platform framework
+\item A validated, open-source, cross-platform framework in which researchers are able to replicate someone else's results
 \item Identification and modeling of the autocorrelation in the counters
+\item Loss of neurons resilience
 \item Noise filtering application
@@
 \end{enumerate}
+\end{frame}
+
+\begin{frame}
+\frametitle{Loss of neurons resilience}
+\begin{figure}
+\centering\includegraphics[width=0.7\textwidth]{./images02/new-images/sdm-neuron-death-500k.png}
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
+Even when 50\% of neurons are dead, SDM recall is barely affected, which is an impressive result and matches with some clinical results of children submitted to hemispherectomy.
+\label{fig:sdm-neuron-death-500k}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Loss of neurons resilience}
+\begin{figure}
+\centering\includegraphics[width=0.7\textwidth]{./images02/new-images/sdm-neuron-death-1m.png}
+\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
+The more neurons are lost, the smaller the critical distance, i.e., the worse the SDM recall.
+\label{fig:sdm-neuron-death-1m}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Critical distance \& Autocorrelation}
+\begin{figure}
+\subfloat{\includegraphics[width=0.46\textwidth,valign=t]{./images02/kanerva-table-7-2-original.png}}
+\subfloat{\includegraphics[width=0.54\textwidth,valign=t]{./images02/sdm-10000w-table-7-2-6-samples.png}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Critical distance \& Autocorrelation}
+\begin{theorem}
+\emph{Each dimension has a small pull bias, which can be measured by}
+$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}$
+\end{theorem}
+
+\begin{theorem}
+The autocorrelation vanishes when $n \rightarrow \infty$, i.e., $\lim_{n \rightarrow \infty} P(x_i = y_i | d(x, y) \le r) = 1/2$.
+\end{theorem}
+
+When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = 0.552905498137$.
+\end{frame}
+
+\begin{frame}
+\frametitle{Critical distance \& Autocorrelation}
+\begin{figure}
+\includegraphics[width=0.75\textwidth]{./images02/autocorrelation/read-random-bs.png}
+\caption{The histogram was obtained through simulation. The red curve is the theoretical normal distribution.}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Critical distance \& Autocorrelation}
+The analytical model which does not consider autocorrelation predicts a critical distance of 209 bits.
+
+\begin{figure}
+\centering\includegraphics[width=0.5\textwidth]{./images02/figure-73-eq-zoom.png}
+\caption{Zoom-in around $d=209$.}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Classification application: training}
+\begin{figure}
+\centering\includegraphics[width=0.6\textwidth]{./images02/classification/groups.png}
+\caption{One noisy image for each of the 62 classification groups.
+\label{fig-classification-groups}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Classification application: confusing cases}
+\begin{figure}
+  \centering
+  \subfloat[``i'', ``l'', and ``r'' with 20\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/ilr-high-noise.png}}
+
+  \subfloat[``i'', ``l'', and ``r'' with 5\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/ilr-low-noise.png}}
+
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Classification application: confusing cases}
+\begin{figure}
+  \centering
+  \subfloat[``c'', ``d'', and ``o'' with 20\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/cdo-high-noise.png}}
+
+  \subfloat[``c'', ``d'', and ``o'' with 5\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/cdo-low-noise.png}}
+
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Classification application: confusing cases}
+\begin{figure}
+  \centering
+  \subfloat[``G'', ``O'', and ``Q'' with 20\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/GOQ-high-noise.png}}
+
+  \subfloat[``G'', ``O'', and ``Q'' with 5\% noise.]{\includegraphics[width=\textwidth]{./images02/classification/GOQ-low-noise.png}}
+
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Classification application: results}
+Generally speaking, the classification algorithm made some mistakes in the 20\% noise scenario, and just a few mistakes in the 5\% noise scenario.
+
+When the training set has 1,000 samples instead of 100 samples of each letter, the algorithm correctly classify almost all letters in both 20\% and 5\% noise scenarios, except for letters ``e'' and ``o''.
+
+It could not differentiate letters ``e'' and ``o'' because their differences are smaller than the critical distance.
+It would have to be trained more and more times in order to learn to differentiate them (because the critical distance decreases when more items are written into the memory).
+\end{frame}
+
+\begin{frame}
+\frametitle{Noise filtering: results}
+\begin{figure}
+\captionsetup[subfigure]{labelformat=empty}
+\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/labelB-20noise.png}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/labelC-20noise.png}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/labelD-20noise.png}}
+
+\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/label8-20noise.png}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Reinforcement learning: TicTacToe}
+\begin{figure}
+\centering\includegraphics[width=\textwidth]{./images-defense/ttt-example.png}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Reinforcement learning: Results}
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{./images02/ttt/results-smart.png}
+\caption{Results playing against the smart player. Each cycle was made of 100 games for training, and then 100 games for measuring statistics.
+\label{fig:ttt-results-smart}
+}
+\end{figure}
 \end{frame}
@@
 
-\begin{frame}
-\frametitle{Conclusion}
-\end{frame}
-
 \end{document}

### 2018-07-18 Chapters/sdm2.tex
properties of the space

@@
 That is, $n$-sized bitstrings correspond to nodes, and edges exist between nodes $iff$ they flip a single bit.
-Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community.
+Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community: it is a bipartite graph with chromatic number 2, and it is planar only if $n\le3$.
+The proofs, though elementary, illuminate some of its properties:  vertices connect bitstrings with an even number of 1's with bitstrings with an odd number of 1's, therefore (i) it is \emph{bipartite} and (ii) it \emph{may be colored with two colors}.
+As for planarity, the case $n=3$ is the largest planar one.
+Imagine a square inside a larger square, then connect the corresponding edges in both squares to obtain a planar drawing.
+$Q_4$ or larger demands three such squares, with the smallest and largest connected but not passing through the middle square.
+
 A good survey is provided by \citet{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.
@@
 One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, our 3-dimensional space.
+
+A major difference is that in the 3-d world, there exist very far points, while in $Q_n$ the largest possible distance is $n$.
+Another one is that the number of directions in $Q_n$ is enormous:  at \emph{every point} there are $n$ directions to choose from.
 

### 2018-07-24 Chapters/Chapter02.tex
small changes
properties of the space
Fix Section ??, Figure ??, and Appendix XYZ pointed out by Kanerva's review.

--- 2018-05-11 Chapters/Chapter02.tex
@@
 We will see tests of a generalized reading operation proposed by Physics Professor Paulo Murilo (personal communication).
-We will see what happens when neurons --- and
-all their information --- is simply and suddenly lost.
+We will see what happens when neurons --- and all their information --- is simply and suddenly lost.
 We will see whether information-theory can improve some of Kanerva's ideas.
@@
 What these results seem to imply is not only that the memory is extremely robust, but that a very small expected number of neurons activated in each access radius would preserve the most valueable information.
-This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{Magic numbers}.
+This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{section:Magic numbers}.
 Is there a thing such as excessive robustness, and, if so, when does robustness become waste?\footnote{Consider, for instance, error-correcting codes in Information Theory or the analogous computer ECC memory in Electrical Engineering: tradeoffs between robustness and waste must be considered at each design decision; Why should it be different in the modeling of human memory?}
@@
 \section{Magic numbers}
+\label{section:Magic numbers}
 
@@
 What are the `best' benchmarks possible?
-
-% Post-defense TODO: Use the robustness of neurons as a guide.  A more quantitative study of this loss of neurons... between d=0 and d=critical_distance
-% Post-defense TODO: magic numbers
-% Post-defense TODO: mu - m sigma, in which the expected number of neurons is, say, 256 (instead of >1000).
-% Post-defense TODO: study this.
-The marginal increase of 10% of Hard locations increases by how many percentage points the critical distance?
-In fact, is there a way to study this algebrically?
-That would be so cool!
 
@@
 When reading, they may compose the counters and give a more precise classification.
-
-% Post-Defense TODO:  SDM creating Voronoi Tesselations of concepts, with different sized areas for each.  This should solve i versus l.
 

### 2018-07-24 Chapters/sdm-autocorrelation.tex
Fix Section ??, Figure ??, and Appendix XYZ pointed out by Kanerva's review.
Eric's suggestions

@@
 Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).
-When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905$.
 Each bit of a hard location does indeed have a small pull bias.
@@
 In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has reduced significantly as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
-In fact, in this case, using Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = 0.5168761116022098$.
+In fact, in this case, using Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = 0.516876$.
 
@@
 
-Using the empirical variance of $\sigma^2 = 27838.3029124$, we calculate $P(wrong) = 0.22037771219874325$.
+Using the empirical variance of $\sigma^2 = 27838.3029$, we calculate $P(wrong) = 0.220377$.
 

### 2018-08-02 Chapters/Chapter02.tex
Proofs Qn

@@
 Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human long-term memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
+
 \begin{enumerate}
@@
 Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.
-Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}\footnote{There has recently been a surge of hardware implementations to minimize energy expenditure and parallelize processes \citep{kang_energy-efficient_2015, kang_-memory_2016, li_hyperdimensional_2016, montagna_pulp-hd:_2018, salahuddin_energy_nodate}.
+Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}\footnote{There has recently been a plethora of hardware implementations to minimize energy expenditure and parallelize processes \citep{kang_energy-efficient_2015, kang_-memory_2016, li_hyperdimensional_2016, montagna_pulp-hd:_2018, salahuddin_energy_nodate}.
 While these implementations are extremely interesting, they do not afford the flexibility to experiment that software does.}:
@@
 What these results seem to imply is not only that the memory is extremely robust, but that a very small expected number of neurons activated in each access radius would preserve the most valueable information.
-This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{section:Magic numbers}.
+This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{Magic numbers}.
 Is there a thing such as excessive robustness, and, if so, when does robustness become waste?\footnote{Consider, for instance, error-correcting codes in Information Theory or the analogous computer ECC memory in Electrical Engineering: tradeoffs between robustness and waste must be considered at each design decision; Why should it be different in the modeling of human memory?}
@@
 \section{Magic numbers}
-\label{section:Magic numbers}
 

### 2018-08-02 Chapters/sdm2.tex
minor
minor2
Proofs Qn

@@
 
-Sparse Distributed Memory (SDM) is a mathematical model developed and suggested as a theory of human memory by Finish Scientist Pentti \citet{Kanerva1988}.
+Sparse Distributed Memory (SDM) is a mathematical model developed and suggested as a theory of human memory by Finnish Scientist Pentti \citet{Kanerva1988}.
 It introduces many interesting mathematical properties of $n$-dimensional binary space that, in a memory model, seem to be remarkably psychologically plausible.
@@
 As for planarity, the case $n=3$ is the largest planar one.
-Imagine a square inside a larger square, then connect the corresponding edges in both squares to obtain a planar drawing.
-$Q_4$ or larger demands three such squares, with the smallest and largest connected but not passing through the middle square.
-
 A good survey is provided by \citet{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.

### 2018-08-02 Chapters/Introduction.tex
papers TODO
sdm ready?

@@
 
-Bitcoin is the first and most famous cryptocurrency, used worldwide, with a highly volatile market cap, as of this writing, of \$ 192 bi.
+Bitcoin is the first and most famous cryptocurrency, used worldwide, with a highly volatile market cap, as of this writing, of \$ 192 B.
 Even so, it faces serious scalability challenges; such as serious quality of service and network congestion when the number of transactions per second is high, and an increase in the transaction fees and uncertain delays in transactions' confirmations.
@@
 
-\emph{Iota} is a second cryptocurrency that, instead of using a blockchain, proposes the use of a ``tangle' architecture': a different way to register the currency ledger across thousands of computers.
+\emph{Iota} is a second cryptocurrency that, instead of using a blockchain, proposes the use of a ``tangle architecture'': a different way to register the currency ledger across thousands of computers.
 Although it has not been confirmed in practice yet, its architecture seems to be significantly more scalable than Bitcoin's blockchain.
 As we will see, the problem here is exactly the opposite of Bitcoin's.
-Iota needs a minimum of transactions per seconds in order to work properly.
+Iota needs a minimum number of transactions per second in order to work properly.
 
-Our analysis suggests an architecture for a distributed currency which is inspired in both Bitcoin's blockchain and Iota's tangle in order to solve the scalability problems.
+Our analysis suggests an architecture for a distributed currency which is inspired by both Bitcoin's blockchain and Iota's tangle in order to solve the scalability problems.
 While Bitcoin's network saturates when it hits a certain number of transactions per second, Iota's does not work properly with less than a certain number of transactions per second.
@@
 
-A second topic that may have an outsized influence on business and that we will be taking a closer look is a model of artificial intelligence --- and of human decision making.
+A second topic that may have an outsized influence on business and that we will be taking a closer look at is a model of artificial intelligence --- and of human decision making.
 
@@
 Beyond a formalization of this process of memory recall, Sparse Distributed Memory will offer us a plausible, both psychologically and neuroscientifically, path towards artificial intelligence.
-We'll need to take a turn here.
 
@@
 
+% Jamil: Se quiser citar algo, tem os papers do Solow (1956, 1957) ou o Advanced Macroeconomics do Romer (2016).
 Technology has been one of the underlying engines behind economic growth.
@@
 Spotify learns which kind of music users like the most and suggests playlists.
-Banks prevent fraud classifying which patterns seem to be erratic towards their customers' previous behavior.
+Banks prevent fraud by classifying which patterns seem to be erratic towards their customers' previous behavior.
 
+% Jamil: Não gosto de parágrafos soltos de uma frase em textos acadêmicos. Questão de estilo, mas comentando por comentar.
 It is gradually becoming impossible to imagine a world without artificial intelligence.
 
-Behind artificial intelligence systems, there is pattern recognition: The capacity to match information from new data with information which has already been seen and is stored in memory.
+Behind artificial intelligence systems, there is pattern recognition: the capacity to match information from new data with information which has already been seen and is stored in memory.
 It may be used in classification, face recognition, character recognition, and so forth.
-Even if AI cannot yet solve numerous hard problems --- such as the Turing Test, Bongard Problems or simply understanding when `Lawyers are sharks' \citep{french1990subcognition, french2000turing, linhares2000glimpse, french2001co} ---, it is clear that the technology must be taken seriously.
+Even if AI cannot yet solve numerous hard problems --- such as the Turing Test, Bongard Problems or simply understanding when `Lawyers are sharks' \citep{french1990subcognition, french2000turing, linhares2000glimpse, french2001co} --- it is clear that the technology must be taken seriously.
 
@@
 \section{Diffusion of innovation}
+
+% Jamil: Convém expandir um pouco mais isso aqui. Qualidade muito discrepante com as outras subseções do capítulo.
 

### 2018-08-05 Chapters/Introduction.tex
imac

--- 2018-04-17 Chapters/Introduction.tex
@@
 In the history of computing business, the most interesting level of analysis seems to be that of the \emph{platform}.
-It is where things can get built on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
+It is where things can go on top of, and communities emerge, and standards fight against each other, and fortunes are built or lost.
 It is, in a sense, a major part of the Big Drama of our moment in history.

### 2018-08-05 Chapters/sdm-autocorrelation.tex
imac

--- 2018-04-22 Chapters/sdm-autocorrelation.tex
@@
 \begin{theorem}
-\emph{Each dimension has a small pull bias, which can be measured by}
+\emph{Each dimension $i$ has a small pull bias, which can be measured by}
 \label{T1}
@@
 Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).
-When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+In our case of $n=1,000$ and $r=451$, for example, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
 Each bit of a hard location does indeed have a small pull bias.
@@
 
-As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 0.5 ($B(n, 0.5)$).
+As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 0.5, $B(n, 0.5)$.
 For $n$ sufficiently large, the Binomial distribution can be approximated by a Normal distribution, i.e., $B(n, 0.5) \rightarrow \mathcal{N}(\mu = n/2, \sigma^2 = n/4)$.
@@
 
-The result comes directly from applyting the equation above in $P(x_i = y_i | d(x, y) \le r)$.
+The result comes directly from applying the equation above in $P(x_i = y_i | d(x, y) \le r)$.
 
@@
 
-Another way to proof is to divide into two cases:
+Another way to prove is to divide into two cases:
 
@@
 Let $h$ be the number of activated hard locations.
-As the probability of activating a specific hard location is constant, $h \sim \text{Binomial}(H, p_1)$.
+As the probability of activating a specific hard location is a constant $h \sim \text{Binomial}(H, p_1)$.
 Thus, $\mathbf{E}[h] = \mu_h = Hp_1$ and $\mathbf{V}[h] = \sigma^2_h = Hp_1(1-p_1)$, where $p_1 = 2^{-n} \sum_{k=0}^{r} \binom{n}{k}$.
@@
 \begin{proof}
-As $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.
-
 By the central limit theorem, $Z$ is normally distributed.
@@
 \end{proof}
+As, in our case, $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.
 
@@
 
-The previous theorems show that there is bias in the counters. Let us analyze the $i$th counter of a hard location.
+The previous theorems show that there is bias in the counters.
+In this section we show that there is a small counter bias, as a function of the bitmatch in dimension $i$.
+If the Hard location's address bit is set to 1, the mean will be above zero.
+And the mean will be below zero if the address bit is set to 0.
 
+Let us analyze the $i$th counter of a hard location.
 Let $s$ be the number of bitstrings written into memory (in our case, $s=10,000$) and $\text{addr}_i$ be the $i$th bit of the hard location's address.
@@
 
-Let $Y_i$ be the number of bitmatches in the $i$ of a hard location's address after $s$ written bitstrings. Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
+Let us introduce another random variable.
+Let $Y_i$ be the number of bitmatches in the $i$th dimension of a hard location's address after $s$ written bitstrings.
+Then, $Y_i = \sum_{k=1}^{\theta} X_k$.
 
@@
 
-Applying the law of total variance, $\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
+Applying the law of total variance, \\
+$\mathbf{V}[Y] = \mathbf{E}[\mathbf{V}[Y|\theta]] + \mathbf{V}[\mathbf{E}[Y|\theta]] = \mathbf{E}[\theta p (1-p)] + \mathbf{V}[p \theta] = p(1-p) \mathbf{E}[\theta] + p^2 \mathbf{V}[\theta] = p(1-p) \mu_\theta + p^2 \sigma^2_\theta$.
 \end{proof}
@@
 
-In summary,
+In summary, what we learn is that
 
@@
 For ``random at x'', $p=0.5$, so $\mu = 0$ and $\sigma^2 = 10.7185$.
-See Figure \ref{fig:sdm-corr-counters}.
+The slight bias above or below 0 can be seen in Figure \ref{fig:sdm-corr-counters}.
 
@@
 \begin{align}
-P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 1 - \mathcal{N}.\text{cdf}(0)
+P(\text{cnt}_i \ge 0 | \text{addr}_i = 1) = P(\text{cnt}_i \le 0 | \text{addr}_i = 0) = 1 - \mathcal{N}.\text{cdf}(0)
 \end{align}
 
-For ``random written at x'', $p=0.5$ implies $\mu_\text{cnt} = 0$, which implies $P(\text{cnt}_i > 0 | \text{addr}_i = 1) = P(\text{cnt}_i < 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
+For ``random written at x'', $p=0.5$ implies $\mu_\text{cnt} = 0$, which implies $P(\text{cnt}_i \ge 0 | \text{addr}_i = 1) = P(\text{cnt}_i \le 0 | \text{addr}_i = 0) = 0.5$, independently of the parameters because they will only affect the variance and the normal distribution is symmetrical around the average.
 

### 2018-08-05 Chapters/Chapter01.tex

@@
 
-The concept of transferring money using cryptography as an underlying technology was shortly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}.
+The concept of transferring money using cryptography as an underlying technology was briefly presented in 1983 by \citet{chaum1983blind} and was deepened in a theoretical paper in 1985 \citep{chaum1985security}.
 However, it was only in 1988 that \citet{chaum1988untraceable} created the term \emph{electronic cash} and also proposed a basic and practical scheme which yielded untraceability yet allowed to trace double spendings.
@@
 
+% Eric: why does it must scale?
 The network must scale to support hundreds of transactions per second, while its capacity is around only eight transactions per second.
@@
 
-Bitcoin seems to have the most decentralized network among the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power (for details, see \citet{gencer2018decentralization}).
+Bitcoin seems to have the most decentralized network among the cryptocurrencies, even so, there are a few miners and mining pools which together control over 50\% of the network’s computing (hash)power (for details, see \citet{gencer2018decentralization}).
 Hence, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior.
@@
 One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network.
-This spam problem has happened several times and affects Bitcoin's trustworthy.
-\citet{bitcoinspam2017} reports a possible spam attack lasting at least contiguous 18 months.
+This spam problem has happened several times and affects Bitcoin's trustworthiness.
+\citet{bitcoinspam2017} reports a possible spam attack lasting at least a contiguous 18 months.
 
@@
 
-After the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created \citep{coinmarketcap}.
+Since the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created \citep{coinmarketcap}.
 In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain.
-Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the Directed Acyclic Graph (DAG) model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}.
+Some cryptocurrencies differ a lot from Bitcoin, like the ones which use the Directed Acyclic Graph (DAG) model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}.
 We are especially interested in one of them: Iota.
 
-Iota uses a DAG model, called tangle, which has a different design than Bitcoin's blockchain.
+Iota uses a DAG model, called Tangle, which has a different design than Bitcoin's blockchain.
 It has neither mining nor confirmation blocks and transaction fees.
@@
 It will be explained later in details.} and is used to confirm other transactions, forming a directed acyclic graph of transactions.
-Thus, a transaction is said to be confirmed when there is enough proof-of-work from the transactions confirming it directly or indirectly.
-There is no other way to confirm transactions but generating new transactions.
+Thus, a transaction is said to be confirmed when there is enough proof-of-work from the transactions confirming it either directly or indirectly.
+There is no other way to confirm transactions other than generating new transactions.
 
@@
 Therefore, theoretically, it scales to any large number of transactions per second.
-The scaling problem of tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work.
-While Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator which works as a trustworthy node \citep{iotacoordinator}.
+The scaling problem of Tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transactions per second; otherwise, the transactions are not confirmed, and the cryptocurrency does not work.
+When Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator that works as a trustworthy node \citep{iotacoordinator}.
 
 Every transaction confirmed by the central coordinator is assumed to be valid and cannot be reverted.
-The remaining of the network can verify a confirmation through the central coordinator's digital signature.
+The remainder of the network can verify a confirmation through the central coordinator's digital signature.
 The coordinator will not be necessary anymore when the number of transactions per second reaches a minimum value, but Iota's developers cannot say precisely what is this minimum value.
-This just elucidates that the tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severely affect Iota's trustworthiness).
+This just elucidates that the Tangle does not seem to work properly under a low volume of transactions (and fluctuations in the number of transactions per second may severely affect Iota's trustworthiness).
 
-The present work intends to propose and analyze a new architecture, named Hathor, which lies between Bitcoin and Iota and may be a viable solution to both scaling, centralization, and spam problems.
-We also present a mathematical analysis of Bitcoin's architecture.
+The present work intends to propose and analyze a new architecture, named Hathor, which is situated on a spectrum between Bitcoin and Iota and may be a viable solution to scaling, centralization, and spam problems.
+We also present a mathematical analysis of the Bitcoin's architecture.
+
+% TODO Eric: Why Bitcoin? Not Hahor?
 
@@
 In \citeauthor{nakamoto2008bitcoin}'s (2009) seminal paper, there is no distinction between bitcoin and blockchain.
-They are just one thing which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money on the internet, in which all users can agree as to whom is entitled to which funds.
+They are just considered parts of a single entity which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money on the internet, in which all users can agree as to whom is entitled to which funds.
 
@@
 Each new block confirms that all previous blocks are valid and have not been tampered with.
-The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network.
+The mechanism that assures the immutability is the proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proofs-of-work faster than all of the remaining machines of the network.
 The network agrees that work should be done in the block at the longest chain in the blockchain.
@@
 
-Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given threshold $A$.
+Bitcoin's blockchain uses as proof-of-work the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given threshold $A$.
 As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}.
-Thus, if the given number is $A = 2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero).
+Thus, if the given threshold is $A = 2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero).
 But if the given number is $A = 2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero).
@@
 Hence, it is not safe to accept a transaction when it is confirmed by only one block.
-It is the idea behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.
+This idea is behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.
 
@@
 A practical example of a network disagreement is the increase of the block size.
-No group with more than 50\% of the network's hash power agreed into increasing the maximum block size to increase the number of transactions confirmed by a block and thus increasing the network's capacity.
+In previous attempts to do so, no group with more than 50\% of the network's hash power agreed to increasing the maximum block size to increase the number of transactions confirmed by a block with the goal of thus increasing the network's capacity.
 Hence, the capacity remains the same, and the community has been discussing the issue in search of a consensus.
@@
 If one loses their private key, unfortunately, access to their funds will be lost forever.
-The public key may be used to a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.
+The public key may be used on a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.
 
@@
 Each transaction's output has a script that is executed by the miners to check whether one has or has not permission to spend that output.
-In other words, whether one has the ownership of that output.
+In other words, the script verifies whether one has the ownership of that output.
 In order to execute these scripts, the miners also need some data.
@@
 
-The output's scripts usually checks whether the public key is valid and whether the digital signature was signed by the private key associated to that public key.
-Although there are only 3 commonly used scripts, one may create a custom script using the Bitcoin's script language \footnote{Your courageous author once tried to make a transaction with custom script to try to double spend a deposit to an exchange, only to learn through this intrepid adventure that Bitcoin allows only 3 script patterns and the others are treated as invalid.}.
+The output's scripts usually check whether the public key is valid and whether the digital signature was signed by the private key associated with that public key.
+Although there are only 3 commonly used scripts, one may create a custom script using the Bitcoin's script language \footnote{Your courageous author once tried to make a transaction with custom script to try to double spend a deposit to an exchange, only to learn through this intrepid adventure that in practice Bitcoin allows only 3 script patterns and the others are treated as invalid.}.
 
-The input contains the data which prove that the sender is the owner of the referred outputs, i.e., the input which must be accepted by the output scripts being spent.
+The input contains the data which proves that the sender is the owner of the referred outputs, i.e., the input which must be accepted by the output scripts being spent.
 Usually, each input has the public key of the sender and a digital signature.
 
-Those users accustomed to block explorers may have been misled by the transaction information that these websites provide.
+Those users accustomed to online block explorer websites may have been misled by the transaction information that these websites provide.
 For example, suppose a miner receives a transaction with ``one input from address A1''.
@@
 
-Iota's underlying technology is Tangle, which has DAG-based architecture with a whole different approach to confirmations.
+Iota's underlying technology is Tangle, which has a DAG-based architecture with a wholly different approach to confirmations.
 It proposes that there is no need for a block to confirm transactions, as transactions can confirm themselves.
-Here, each transaction has its own proof-of-work, named weight, and they must confirm two other previous transactions.
+Here, each transaction has its own proof-of-work, named weight, and must confirm two other previous transactions.
 In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph (DAG), as in Fig.
@@
 
-Like Blockchain, Tangle is another technology to store immutable data and may be the underlying technology to different applications, such as cryptocurrencies, digital contracts (Ethereum-like), digital notaries, and so forth.
+Like Blockchain, Tangle is another technology to store immutable data and may be the underlying technology for different applications, such as cryptocurrencies, digital contracts (like Ethereum), digital notaries, and so forth.
 
-Transactions may be either confirmed or unconfirmed.
-The confirmed transactions have been already confirmed by at least one more transaction.
-It does not mean they are already irreversible and protected against a double spend attack --- it just means at least one transaction has done some work to confirm it.
+Transactions always exist in one of two states: confirmed or unconfirmed.
+The confirmed transactions have been already confirmed by at least one other transaction.
+Being confirmed does not mean that a transaction is already irreversible and protected against a double spending attack --- it just means at least one transaction has done some work to confirm it.
 The unconfirmed transactions are called tips and they are eager to be confirmed.
-Usually a new transaction selects two tips to confirm, but this rule may not be followed.
+Usually a new transaction selects two tips to confirm, but this rule is not enforced.
 
-Different from Bitcoin, transactions does not have scripts to check whether one may spend the tokens.
-Instead, it uses a fixed Winternitz hash-based digital signature \citep{dods2005hash}, i.e., whoever correctly signs the transaction may tranfer the tokens.
-It also supports multi-signature scheme (see \cite{iotamultisign}).
-Not allowing scripts is a disadvantage when compared to Bitcoin because future applications are limited and changing it would require a major modification in transaction format.
+In contrast to Bitcoin, transactions do not have scripts to check whether one may spend the tokens.
+Instead, Iota uses a fixed Winternitz hash-based digital signature \citep{dods2005hash}, i.e., whoever correctly signs the transaction may transfer the tokens.
+It also supports a multi-signature scheme (see \cite{iotamultisign}).
+Not allowing scripts is a disadvantage when compared to Bitcoin because future applications are limited and changing Iota would require a major modification in transaction format.
 
-At the beginning, the digital signature algorithm relied on the Curl hash function, which is a ternary hash function designed by Iota's developers.
-This hash function was replaced by the Kerl hash function after \cite{heilman2017iota} has found a critical vulnerability which enabled practical signature forgery attacks.
-The Kerl hash function is a variation of SHA-3 also designed by Iota's developers \citep{iotakerl}.
-Like the Curl hash function, Kerl has not been deeply studied by cryptography researchers and may have critical vulnerabilities.
+At the beginning of Iota's development, the digital signature algorithm relied on the Curl hash function, which is a ternary hash function designed by Iota's developers.
+This hash function was replaced by the Kerl hash function after \citet{heilman2017iota} has found a critical vulnerability that enabled practical signature forgery attacks.
+The Kerl hash function, also designed by Iota's developers \citep{iotakerl}, is a variation of SHA-3.
+Like the Curl hash function, Kerl has not been deeply studied by cryptography researchers and may have critical vulnerabilities that remain unknown to the public.
 
-Iota's tokens are pre-mined, which means they have been issued in the genesis transaction and no more tokens will ever be issued.
+Iota's tokens are pre-mined, which means they have all been issued in the genesis transaction and no more tokens will ever be issued.
 This means that there are no miners in Iota's network and only the users keep the network alive through their new transactions.
 This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it.
-As users perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away, making it even harder to reach the minimum required to work properly \citep{debo2010prices}.
-Even if the minimum has been reached, if the number of transactions per second plummet, the time to confirm the most recent transactions will increase significantly.
-This possibly poses a serious risk to the network stability and users' trustworthy.
+As users' perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away, making it even harder to reach the minimum required to work properly \citep{debo2010prices}.
+Even if the minimum has been reached, if the number of transactions per second plummets, the time to confirm the most recent transactions will increase significantly.
+This possibly poses a serious risk to the network stability and users' trust.
 
@@
 Tokens may be freely transferred without any losses, even for tiny amounts, enabling micropayments.
-For example, online workers paid per hour may receive their payments every hour, reducing the default risk.
-This allows untrusted parties to work together reducing the risk for both.
-The contractor always wants to pay in the end, while the workers always wants to get paid in the beginning.
+For example, online workers paid per hour may receive their payments every hour, reducing their default risk.
+This allows untrusted parties to work together, with reduced risk for both.
+The contractor always wants to pay at the end of a job, while the workers always want to get paid in the beginning.
 
-The Internet-Of-Things (IoT) technologies, which is a network of smart devices that are connected to the internet and exchange data, also benefits from not having fees.
+The Internet-Of-Things (IoT) technologies, which comprise a network of smart devices that are connected to the internet and exchange data, also benefits from not having fees.
 Imagine your refrigerator doing your groceries and automatically paying using your Iota tokens, or your electric car automatically paying for recharge.
@@
 
-Theoretically, Iota's network benefits from high volume of transactions.
+Theoretically, Iota's network benefits from a high volume of transactions.
 The more transactions are coming, the faster previous transactions are confirmed.
-This is a major feature of Tangle which contrasts with Bitcoin's difficulty to scale.
-On the other hand, low volume of transactions is a primary problem for Iota, since it would take too long to confirm previous transactions.
-As it is a known problem, Iota has created a central coordinator which works as a trustworthy node, clearing transactions.
+This is a major benefit of Tangle, which contrasts with Bitcoin's difficulty to scale.
+On the other hand, low transaction volume is a primary problem for Iota, since it would take too long to confirm previous transactions.
+As when few transactions arrive is a known problem, Iota has created a central coordinator which works as a trustworthy node, clearing transactions.
 In other words, every transaction directly or indirectly confirmed by the central coordinator is assumed to be cleared.
-Iota's developers claims that the coordinator will be turned off when the network outgrows a minimum (unknown) size.
+Iota's developers claim that the coordinator will be turned off when the network outgrows a minimum \emph{unknown} size.
 
@@
 In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7.
-It may be on purpose, or maybe transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction.
-The selection algorithm seems to be important to protect the network against double spend attacks.
+It may have been on purpose, or perhaps transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction.
+The selection algorithm seems to be critical in protecting the network against double spend attacks.
 For further information, see \cite{tangle2016}.

### 2018-08-05 Chapters/sdm-autocorrelation.tex

@@
 Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).
-In our case of $n=1,000$ and $r=451$, for example, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905498137$.
+When $n=1,000$ and $r=451$, $P(x_i = y_i | d(x, y) \le r) = p = 0.552905$.
 Each bit of a hard location does indeed have a small pull bias.

### 2018-08-05 defense-presentation.tex

@@
 \usetheme{Boadilla}
+%\usetheme{Berlin}
 
@@
 
-\begin{frame}
-\frametitle{Introduction}
-
-Modern management and high technology interact in multiple, profound, ways. A whole new ecosystem seems to have emerged within computing and business.
-
-The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI, and cognitive technology \emph{as an executive}, not as a research scientist or specialized engineer.
-
-Professor Andrew Ng tells students at Stanford's Graduate School of Business that ``AI is the new electricity'', as his hyperbolic way to emphasize the potential transformational power of the technology.
-\end{frame}
+%\begin{frame}
+%\frametitle{Introduction}
+%
+%Modern management and high technology interact in multiple, profound, ways. A whole new ecosystem seems to have emerged within computing and business.
+%
+%The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI, and cognitive technology \emph{as an executive}, not as a research scientist or specialized engineer.
+%
+%Professor Andrew Ng tells students at Stanford's Graduate School of Business that ``AI is the new electricity'', as his hyperbolic way to emphasize the potential transformational power of the technology.
+%\end{frame}
 
@@
 
-\part{Hathor: An alternative towards a scalable cryptocurrency}
+\section{Hathor: An alternative towards a scalable cryptocurrency}
+\begin{frame}
+\sectionpage
+\end{frame}
+
 \begin{frame}
@@
 
-The primary problem for creating digital money is how to prevent double spending.
+The primary problem for creating digital money is \textbf{how to prevent double spending}.
 
@@
 
-The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the 30-year literature on e-cash.
+The \textbf{no central point of trust} and \textbf{predictable money supply} together with a \textbf{clever solution to the double-spending problem} is what separates Bitcoin from the 30-year literature on e-cash.
 
@@
 \begin{frame}
-\frametitle{Bitcoin}
+\frametitle{Bitcoin \& Blockchain}
 
@@
 
-Lack of scalability (up to 3.7 tx/s for 1MB per block)? Spam attacks?
+Lack of scalability (up to 3.75 tx/s for 1MB per block)? Spam attacks?
 
@@
 \begin{frame}
-\frametitle{Iota}
+\frametitle{Iota \& Tangle}
 
@@
 
-Iota does not work in low load scenarios. It uses a central coordinator to prevent attacks.
+\textbf{Iota is not safe in low load scenarios.} It must use a central coordinator to prevent attacks.
 
-They does not know what is the minimum number of transactions per seconds so that they can disable the central coordinator.
+They plan to turn off the central coordinator, but \textbf{they do not know} what is the minimum number of transactions per seconds so that it is safe to do it.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Hathor's architecture}
+
+\begin{figure}
+\centering\includegraphics[width=0.45\textwidth]{./images01/sim/hathor-2.pdf}
+\end{figure}
 
@@
 \item Mathematical analysis of Bitcoin
-\item Proposal of Hathor, \small{an architecture which seems to work in both low and high load scenarios.}
+\item Proposal of Hathor, \emph{a novel architecture} which seems to work in both low and high load scenarios.
 \item Analysis of Hathor
 \end{enumerate}
-\end{frame}
-
-\begin{frame}
-\frametitle{Hathor's architecture}
-
-\begin{figure}
-\centering\includegraphics[width=0.45\textwidth]{./images01/sim/hathor-2.pdf}
-\end{figure}
-
 \end{frame}
@@
 
-In the double spending attack, the attacker's send some funds to the victim, let's say a merchant.
+In the double spending attack, \textbf{the attacker's send some funds to the victim}, let's say a merchant.
 
-They wait for $k$ confirmations of the transaction, and the victim delivers the good or the service to the attacker.
+They wait for $k$ confirmations of the transaction, and \textbf{the victim delivers the good or the service to the attacker}.
 
-Then, the attacker mine enough blocks with a conflicting transaction, double spending the funds which was sent to the victim.
+Then, the attacker mine enough blocks \textbf{with a conflicting transaction}, double spending the funds which was sent to the victim.
 
-If the attacker is successful, the original transaction will be \textit{erased} and the victim will be left with no funds at all.
+If the attacker is successful, the original transaction will be \textit{erased} and \textbf{the victim will be left with no funds at all}.
 
@@
 \begin{frame}
-\frametitle{Analysis of Hathor: boundary cases}
+\frametitle{Analysis of Hathor}
+
+\begin{figure}
+\centering\includegraphics[width=0.45\textwidth]{./images01/sim/hathor-2.pdf}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Analysis of Hathor: Boundary cases}
 
@@
 \begin{frame}
-\frametitle{Analysis of Hathor: confirmation time}
+\frametitle{Analysis of Hathor: Confirmation time}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Analysis of Hathor: confirmation time (2)}
+\frametitle{Analysis of Hathor: Confirmation time (2)}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Analysis of Hathor: load changing over time}
+\frametitle{Analysis of Hathor: Load changing over time}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Analysis of Hathor: number of unconfirmed transactions}
+\frametitle{Analysis of Hathor: Number of unconfirmed transactions}
 \begin{figure}
@@
 
-\part{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}
+%=============================================
+%=============================================
+%=============================================
+
+\section{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}
+\begin{frame}
+\sectionpage
+\end{frame}
+
 \begin{frame}
@@
 
-SDM is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+SDM is a mathematical \textbf{model of long-term memory} that has a number of neuroscientific and psychologically plausible dynamics.
 
-This model is used in all sort of applications due to its ability to closely reflect the human capacity to remember past experiences.
+This model is used in all sort of applications due to its ability to closely reflect \textbf{the human capacity to remember past experiences}.
 
-This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.
+This flexibility into \textbf{mapping one situation in another} is an essential human feature which is hard to replicate in computers.
 
@@
 \begin{frame}
-\frametitle{Classification application: training}
+\frametitle{Classification application: Training}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Classification application: confusing cases}
+\frametitle{Classification application: Confusing cases}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Classification application: confusing cases}
+\frametitle{Classification application: Confusing cases (2)}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Classification application: confusing cases}
+\frametitle{Classification application: Confusing cases (3)}
 \begin{figure}
@@
 \begin{frame}
-\frametitle{Classification application: results}
+\frametitle{Classification application: Results}
 Generally speaking, the classification algorithm made some mistakes in the 20\% noise scenario, and just a few mistakes in the 5\% noise scenario.
 
-When the training set has 1,000 samples instead of 100 samples of each letter, the algorithm correctly classify almost all letters in both 20\% and 5\% noise scenarios, except for letters ``e'' and ``o''.
+When the training set has 1,000 samples (instead of 100) of each letter, the algorithm correctly classify almost all letters in both 20\% and 5\% noise scenarios, except for letters ``e'' and ``o''.
 
-It could not differentiate letters ``e'' and ``o'' because their differences are smaller than the critical distance.
-It would have to be trained more and more times in order to learn to differentiate them (because the critical distance decreases when more items are written into the memory).
+It seems that the algorithm could not differentiate letters ``e'' and ``o'' because their differences are \textbf{smaller than the critical distance}.
+I believe it would just have to be trained more and more in order to learn to differentiate them (because the critical distance decreases when more items are written into the memory).
 \end{frame}
@@
 \begin{frame}
-\frametitle{Noise filtering: results}
+\frametitle{Noise filtering: Interference}
 \begin{figure}
-\captionsetup[subfigure]{labelformat=empty}
-\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/labelB-20noise.png}}
+\includegraphics[height=0.15\textheight]{./images02/filter/case2-b-10noise.png}
+\end{figure}
+\begin{figure}
+\includegraphics[height=0.15\textheight]{./images02/filter/case2-c-10noise.png}
+\end{figure}
+\begin{figure}
+\includegraphics[height=0.15\textheight]{./images02/filter/case2-d-10noise.png}
+\end{figure}
+\begin{figure}
+\includegraphics[height=0.15\textheight]{./images02/filter/case2-8-10noise.png}
+\end{figure}
+\end{frame}
 
-\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/labelC-20noise.png}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/labelD-20noise.png}}
-
-\subfloat[]{\includegraphics[width=0.5\textwidth]{./images02/filter/label8-20noise.png}}
+\begin{frame}
+\frametitle{Noise filtering: After using labels}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{./images02/filter/labelB-20noise.png}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{./images02/filter/labelC-20noise.png}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{./images02/filter/labelD-20noise.png}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{./images02/filter/label8-20noise.png}
 \end{figure}
@@
 
-\part{Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users}
+%=============================================
+%=============================================
+%=============================================
+
+\section{Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users}
+\begin{frame}
+\sectionpage
+\end{frame}
+
 \begin{frame}
@@
 
+Based on the Bass Model, which is in the Top 10 Most Influential Papers in 50 years of the Management Science journal.
+
+The Bass model assumes that there are two types of people: \textbf{the innovators} and \textbf{the imitators}---a simplification of Rogers' famous five classes (innovators, early adopters, early majority, late majority, and laggards).
+
+The Bass model would be a good choice to forecast the total number of users.
+But most companies deem the total number of users as confidential, and \textbf{they only publish the number of active users}.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{The extended model}
+\begin{align*}
+\text{\textbf{Bass model:}} \\
+\frac{f(t)}{1-F(t)} &= p + qF(t), F(0)=0 \\
+\\
+\text{\textbf{Proposed extension:}} \\
+\frac{f(t)}{1-F(t)} &= p + qF(t) - wR(t), F(0)=0, R(0)=0
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Contributions}
 The main contributions are:
@@
 \item Add a rejection parameter to the model
-\item Estimate the parameters using only the number of active users
-\item The possibility of reducing the number of active users (instead of only increase)
+\item Estimate the parameters using only the number of active users \\(instead of the total number of users)
+\item The possibility of reducing the number of active users \\(instead of only increasing)
 \end{enumerate}
@@
 
+\begin{frame}
+\frametitle{Models of rejection}
+\begin{align*}
+	\text{Model 1: }& r(t) = \nu f(t) \\
+    \\
+	\text{Model 2: }& \frac{r(t)}{1-R(t)} = \nu f(t) \\
+    \\
+	\text{Model 3: }& r(t) = \nu [F(t) - R(t)] \\
+    \\
+	\text{Model 4: }& \frac{r(t)}{1-R(t)} = \nu [F(t) - R(t)]
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Model 2}
+\begin{figure}
+	\centering
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images03/fb-model2-1.eps}}
+	\caption{Fit of Model 2 with Facebook's active users dataset.\label{fig:model2fit}
+		{$mF(t)$ is the total users, $mR(t)$ is the inactive users, and $mA(t)$ is the active users.
+The unit of these functions are thousands of people.
+The parameters are $m=1,497.50$, $p=0.000331$, $q=0.100088$, $w=0.140595$, and $\nu=0.187188$.
+The goodness of fit are $R^2=99.84\%$ and BIC=10,566.52.}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Model 3}
+\begin{figure}
+	\centering
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images03/fb-model3-1.eps}}
+	\caption{Fit of Model 3 with Facebook's active users dataset.\label{fig:model3fit}
+		{$mF(t)$ is the total users, $mR(t)$ is the inactive users, and $mA(t)$ is the active users.
+The unit of these functions are thousands of people.
+The parameters are $m=1,967.64$, $p=0.000184$, $q=0.097867$, $w=0.330511$, and $\nu=0.006912$.
+The goodness of fit are $R^2=99.83\%$ and BIC=11,485.68}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Results: Model 4}
+\begin{figure}
+	\centering
+	{\includegraphics*[scale=0.9,width=0.65\textwidth]{images03/fb-model4-1.eps}}
+	\caption{Fit of Model 4 with Facebook's active users dataset.\label{fig:model4fit}
+		{$mF(t)$ is the total users, $mR(t)$ is the inactive users, and $mA(t)$ is the active users.
+The unit of these functions are thousands of people.
+The parameters are $m=1,854.85$, $p=0.000183$, $q=0.099738$, $w=0.334454$, and $\nu=0.007007$.
+The goodness of fit are $R^2=99.84\%$ and BIC=10,724.55}}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\titlepage
+\end{frame}
+
 \end{document}

### 2018-08-19 Chapters/sdm-autocorrelation.tex
Figure autocorrelation

--- 2018-07-24 Chapters/sdm-autocorrelation.tex
@@
 Each bit of a hard location does indeed have a small pull bias.
-What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0.
+What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0---an intuition developed in Figure \ref{fig:bitmatchbias}.
+
+\begin{figure}[!h]
+\centering
+
+\includegraphics[width=\textwidth]{images02/autocorrelation/autocorrelation_n=1000_closeview=False.eps}
+
+\caption{The threshold size $r$ of hard-locations bring about the autocorrelation of Theorem \ref{T1}.
+Consider, for instance, the extremes \{0, $n$\}.
+Let us start with $n$ (in our case, $n=1000$ dimensions): given the information that $d(x,y)\le n$, the probability of a bitmatch in dimension $i$ is $1/2$; as $\forall x,y \in \{0, 1\}^N, d(x,y)\le n$.
+At the other extreme, consider that we have the information that $d(x,y)\le 0$: in this case $x=y$ and the probability of a bitmatch, in any dimension, is 1.
+The autocorrelation hence drops monotonically until convergence at $1/2$ as the distance grows.
+Numerically, our results converge to precisely $1/2$ only after $d(x,y)\ge 607$ for $n=1000$; $d(x,y)\ge 5332$ for $n=10000$, and $d(x,y)\ge 83$ for $n=100$.
+This will be seen in Lemma \ref{thm:sdm-access-radius}.
+\label{fig:bitmatchbias}}
+\end{figure}
 

### 2018-08-19 Chapters/sdm-applications.tex

@@
 Then, we have written the classification group bitstring into the bitstring associated to each noisy image, i.e., write(bs\_image, bs\_label).
-For a complete image training set, see Appendix XYZ.
 
@@
 It has hit all images for 44 out of 62 groups and made at least one miss for the other 18 groups.
-The misses may be seen in details in Figure \ref{fig-classificiation-high-noise-misses}.
+The misses may be seen in details in Figure \ref{fig-classification-high-noise-misses}.
 
@@
 The reward bitstring may be a positive, negative, or neutral bitstring.
-The positive reward bitstring is randomly generated and then only the bits related to the action will be set to 1 (see Figure \ref{fig:ttt-positive-reward-bitstring}).
+The positive reward bitstring is randomly generated and then only the bits related to the action will be set to 1 (see Figure \ref{fig:ttt-positive-reward-bitstrings}).
 The negative reward bitstring is also randomly generated and then only the bits related to the action will be set to 0.
@@
 \label{fig:ttt-positive-reward-bitstrings}}
+
 \end{figure}

### 2018-08-22 Chapters/hathor-architecture.tex

@@
 There may be any number of outputs provided that they comply with the number of newly generated tokens.
+Differently from Bitcoin's blocks, Hathor's blocks do not contain transactions inside it.
+They only confirm two transactions and another block.
 Each block collects all fees from all transactions confirmed by it which have not been confirmed by another block before.
@@
 Unless a block confirms either directly or indirectly an invalid transaction, every block is valid.
-There is no need to left a block behind since its proof-of-work is increasing transactions' accumulated weight.
+There is no need to leave a block behind since its proof-of-work is increasing transactions' accumulated weight.
 

### 2018-08-22 autocorrelationPaper/IEEEtran/to-include.tex

--- 2018-08-22 Chapters/sdm-autocorrelation.tex
@@
-% !TEX root = ../partial-sdm.tex
-
-\section{A deviation from the equator distance?}
-
-Kanerva writes\footnote{Email thread `SDM: A puzzling issue and an invitation', started March 16th 2018, in which we discussed the aforementioned discrepancy.
-To think that some centuries ago, all scientific publishing was the exchange of such letters.}:
-
-\begin{quote}
-    You have done an incredibly thorough analysis of SDM.
-I like the puzzle in your message and believe that your simulations are correct and to be learned from.
-So what to make of the difference compared to my Figure 7.3 (and your Figure \ref{fig:kanerva-figure-73-calculated})?
-I think the difference comes from my not having accounted fully for the effect of the other 9,999 vectors that are stored in the memory.
-You say in it\\
-
-   ``Our results show that the theoretical prediction is not accurate.
-There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).'' \\
-
-   I think that is correct.
-It also brings to mind a comment Louis Jaeckel made when we worked at NASA Ames.
-He pointed out that autoassociative storage (each vector is stored with itself as the address) introduces autocorrelation that my formula for Figure 7.2 did not take into account.
-When we read from memory, each stored vector exerts a pull toward itself, which also means that each bit of a retrieved vector is slightly biased toward the same bit of the read address, regardless of the read address.
-We never worked out the math.
-\end{quote}
-
-This is an important observation.
-A hard location is activated because it shares many dimensions with the items read from or written onto it.
-Imagine the `counter's eye view':  each individual counter `likes' to write on its own corresponding bit-address value more than it likes the opposite; as each hard-location has a say in its own area --- and nowhere else.
-
-%\footnote{If you are a cell I then I have some serious piece of advice to you.
-I would strongly suggest avoiding the neighborhood of a neuron.
-Neurons have serious authoritarian tendencies, bossing around --- generally electrifying --- the hopeless innocent cells around them just because.}
-
-Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
-Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
-Let us call these shared bitvalues a \emph{bitmatch in dimension $i$}.
-
-So, what is the probability of bitmatches given that we know the access radius $r$ between the address and a hard-location?
-\bigskip
-
-\begin{theorem}
-\emph{Each dimension $i$ has a small pull bias, which can be measured by}
-\label{T1}
-$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}.$
-\end{theorem}
-
-\begin{proof}
-    The left-hand expression $P(x_i = y_i | d(x, y) \le r)$ computes the probability of a bitmatch in $i$, given that we know that $x$ and $y$ are in the access radius defined by $r$, i.e., $d(x, y)\le r$.
-
-    Applying the law of total probability to the left-hand expression we obtain
-
-    \begin{align}
-    \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) P(d(x, y) = k | d(x, y) \le r)
-    \end{align}
-
-    We also know that
-
-    \begin{align}
-    P(x_i = y_i | d(x, y) = k) &= \frac{n-k}{n} \\
-    P(d(x, y) = k | d(x, y) \le r) &= \frac{\binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
-    \end{align}
-
-    Hence,
-
-    \begin{align}
-    P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \frac{n-k}{n} \binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
-    \end{align}
-
-    Finally, the combinatorial identity
-
-    \begin{align}
-    \frac{n-k}{n} \binom{n}{k} = \frac{(n-k)}{n} \frac{n!}{(n-k)! k!} = \frac{(n-1)!}{k! (n-1-k)!} = \binom{n-1}{k}
-    \end{align}
-
-    closes the theorem.
-
-\end{proof}
-
 Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).

### 2018-08-22 autocorrelationPaper/IEEEtran/to-include.tex
dangerous commit

@@
+Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
+Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
+Let us call these shared bitvalues a \emph{bitmatch in dimension $i$}.
+
+So, what is the probability of bitmatches given that we know the access radius $r$ between the address and a hard-location?
+\bigskip
+
+\begin{theorem}
+\emph{Each dimension $i$ has a small pull bias, which can be measured by}
+\label{T1}
+$P(x_i = y_i | d(x, y) \le r) = \dfrac{\sum_{k=0}^{r} \binom{n-1}{k}}{\sum_{k=0}^{r} \binom{n}{k}}.$
+\end{theorem}
+
+\begin{proof}
+    The left-hand expression $P(x_i = y_i | d(x, y) \le r)$ computes the probability of a bitmatch in $i$, given that we know that $x$ and $y$ are in the access radius defined by $r$, i.e., $d(x, y)\le r$.
+
+    Applying the law of total probability to the left-hand expression we obtain
+
+    \begin{align}
+    \sum_{k=0}^{r} P(x_i = y_i | d(x, y) = k \le r) P(d(x, y) = k | d(x, y) \le r)
+    \end{align}
+
+    We also know that
+
+    \begin{align}
+    P(x_i = y_i | d(x, y) = k) &= \frac{n-k}{n} \\
+    P(d(x, y) = k | d(x, y) \le r) &= \frac{\binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
+    \end{align}
+
+    Hence,
+
+    \begin{align}
+    P(x_i = y_i | d(x, y) \le r) = \frac{\sum_{k=0}^{r} \frac{n-k}{n} \binom{n}{k}}{\sum_{j=0}^{r} \binom{n}{j}}
+    \end{align}
+
+    Finally, the combinatorial identity
+
+    \begin{align}
+    \frac{n-k}{n} \binom{n}{k} = \frac{(n-k)}{n} \frac{n!}{(n-k)! k!} = \frac{(n-1)!}{k! (n-1-k)!} = \binom{n-1}{k}
+    \end{align}
+
+    closes the theorem.
+
+\end{proof}
+
 Theorem \ref{T1} is valid for both ``x written at x'' (autoassociative memory) and ``random written at x'' (heteroassociative memory).

### 2018-08-22 autocorrelationPaper/to-include.tex
moving folder

--- 2018-08-22 autocorrelationPaper/IEEEtran/to-include.tex
@@
-Let $x$ and $y$ be random bitstrings and $n$ be the number of dimensions in the memory; let $x_i$ and $y_i$ be the $i$-th bit of $x$ and $y$, respectively; and $d(x, y)$ be the Hamming distance.
-Whilst the probability of a shared bit-value between same dimension-bits in two random addresses is $1/2$, an address only activates hard-locations close to it.
-Let us call these shared bitvalues a \emph{bitmatch in dimension $i$}.
+\begin{figure}[h]
+\centering
+\subfloat[1 sample for each distance $x$ \label{fig:sdm-10000w-table-7-2-1sample} ]{\includegraphics[width=0.5\textwidth]{/././images02/sdm-10000w-table-7-2.png}}
+\subfloat[6 samples for each distance $x$ \label{fig:sdm-10000w-table-7-2-6samples} ]{\includegraphics[width=0.5\textwidth]{/././images02/sdm-10000w-table-7-2-6-samples.png}}
 
-So, what is the probability of bitmatches given that we know the access radius $r$ between the address and a hard-location?
+\caption{Results generated by the framework diverging from Kanerva's original Figure 7.3.
+Here we had a 1,000 bit, 1,000,000 hard location SDM with 10,000 random bitstrings written into it, which was also Kanerva's configuration.
+\label{fig:sdm-10000w-table-7-2}}
+\end{figure}
+
 \bigskip

### 2018-08-22 paperIEEE.tex
wtf no graphics?

%% LyX 2.3.0 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.

\RequirePackage{fix-cm} % fix some latex issues see: http://tug.org/texmf-dist/doc/latex/base/fixltx2e.pdf

\documentclass[twocolumn,journal]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{calc}
\usepackage[unicode=true,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=1,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{pdftitle={Your Title},
 pdfauthor={Your Name},
 pdfpagelayout=OneColumn, pdfnewwindow=true, pdfstartview=XYZ, plainpages=false}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
% protect \markboth against an old bug reintroduced in babel >= 3.8g
\let\oldforeign@language\foreign@language
\DeclareRobustCommand{\foreign@language}[1]{%
  \lowercase{\oldforeign@language{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% for subfigures/subtables
\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem*{cor}{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\DeclareMathOperator\erfc{erfc}
\DeclareMathOperator\erf{erf}

\makeatother

\begin{document}

\title{Autocorrelation in Sparse Distributed Memory}

\author{Marcelo S. Brogliato,~\IEEEmembership{Member,~IEEE,} and Alexandre
Linhares, \IEEEmembership{Member,~IEEE,}\thanks{Dr. Brogliato is with sdm.ai and with Behavioral and Decision Sciences,
EBAPE, Fundação Getulio Vargas, Rio de Janeiro, Brazil, e-mail: \protect\href{http://msbrogli@sdm.ai}{msbrogli@sdm.ai}.}\thanks{Dr. Linhares is with sdm.ai and with Behavioral and Decision Sciences,
EBAPE, Fundação Getulio Vargas, Rio de Janeiro, Brazil, e-mail: \protect\href{http://linhares@sdm.ai}{linhares@sdm.ai}.}}

\markboth{Working Paper}{Brogliato and Linhares: Autocorrelation of SDM}

\IEEEpubid{0000\textendash 0000/00\$00.00~\copyright~2012 IEEE}
\maketitle
\begin{abstract}
Sparse Distributed Memory (SDM) is a neuroscientific and psychologically
plausible model of human memory. In this paper we present an anomaly
between its previously theoretized behavior and its actual behavior,
and demonstrate that this anomaly is due to the autocorrelation involved
in the model. Our findings have systemwide implications for those
willing to implement and apply SDM in computational intelligence settings.
\end{abstract}

\begin{IEEEkeywords}
Sparse Distributed Memory, Autocorrelation, Neural Networks, Memory,
Systemwide properties.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle{}

\section{Introduction}

\IEEEPARstart{S}{parse distributed memory} has been shown to be
fodastic.

In this paragraph we explain what SDM is.

In this other paragraph we explain what the equator distance is. Then,
this introductory section breaks and moves on to the main findings
of this paper.

\begin{figure}[!htb]
\centering\includegraphics{./images02/calculated-table-72.png}
\caption{Kanerva's original Figure 7.3 generated using the equations from \citet{brogliato2014sparse}.
\label{fig:kanerva-figure-73-calculated}
}
\end{figure}

\input{./Chapters/sdm-autocorrelation.tex}

\section{Conclusions}

bla bla

\appendices{}

\section*{Acknowlegment}

The authors would like to thank Pentti Kanerva, Eric Paul Nichols,
José Ricardo de Almeida Torreão, Moacyr Alvim Horta Barbosa da Silva,
Flavio Codeço Coelho, Horacio Hideki Yanasse, and Paulo Murilo Castro
de Oliveira for their careful, comprehensive, reviews of the first
author\textquoteright s theses.

\bibliographystyle{plain}
\bibliography{mybib02}

\begin{IEEEbiography}[{\fbox{\begin{minipage}[t][1.25in]{1in}%
Replace this box by an image with a width of 1\,in and a height of
1.25\,in!%
\end{minipage}}}]{Your Name}
 All about you and the what your interests are.
\end{IEEEbiography}

\begin{IEEEbiographynophoto}{Coauthor}
Same again for the co-author, but without photo
\end{IEEEbiographynophoto}

\end{document}

### 2018-08-22 Chapters/sdm-autocorrelation.tex
wtf no graphics?
some fucking graphics finally appear

@@
-% !TEX root = ../partial-sdm.tex
+% !TEX root = ../paperIEEE.tex
 
@@
 
-\includegraphics[width=\textwidth]{images02/autocorrelation/autocorrelation_n=1000_closeview=False.eps}
+\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/autocorrelation_n=1000_closeview=False.eps}
 
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{./images02/sdm-10000bits-10000w-table-7-2.png}
+  \includegraphics[width=0.5\textwidth]{./images02/sdm-10000bits-10000w-table-7-2.png}
 
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{./images02/autocorrelation/same-bit-histogram.png}
+  \includegraphics[width=0.5\textwidth]{./images02/autocorrelation/same-bit-histogram.png}
 
@@
   \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
+
   \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
@@
   \subfloat[``random at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/random_at_x_counters.png}}
+
   \subfloat[``x at x'']{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_counters.png}}
@@
   \subfloat[Equation \ref{eqn:sdm-eta1-addr1}  ($\text{addr}_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr1.png}}
+
   \subfloat[Equation \ref{eqn:sdm-eta1-addr0} ($\text{addr}_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1_addr0.png}}
@@
   \subfloat[Equation \ref{eqn:sdm-eta1}  ($\eta_k=1)$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta1.png}}
+
   \subfloat[Equation \ref{eqn:sdm-eta0} ($\eta_k=0$)]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-counters-eta0.png}}
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/read-random-bs.png}
+  \includegraphics[width=0.5\textwidth]{./images02/autocorrelation/read-random-bs.png}
 
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{./images02/autocorrelation/single-read-dist.png}
+  \includegraphics[width=0.5\textwidth]{./images02/autocorrelation/single-read-dist.png}
 

### 2018-08-22 paperIEEE.tex
some fucking graphics finally appear

@@
+% Preview source code
+
 %% LyX 2.3.0 created this file.  For more info, see http://www.lyx.org/.
 %% Do not edit unless you really know what you are doing.
-
-\RequirePackage{fix-cm} % fix some latex issues see: http://tug.org/texmf-dist/doc/latex/base/fixltx2e.pdf
-
 \documentclass[twocolumn,journal]{IEEEtran}
@@
 \usepackage{calc}
+\usepackage{graphicx}
 \usepackage[unicode=true,
@@
 
+% !TEX root = ../partial-sdm.tex
+
+\@ifundefined{showcaptionsetup}{}{%
+ \PassOptionsToPackage{caption=false}{subfig}}
+\usepackage{subfig}
 \makeatother
@@
 \begin{IEEEkeywords}
-Sparse Distributed Memory, Autocorrelation, Neural Networks, Memory,
+Sparse Distributed Memory, Autocorrelation, Theoretical Neuroscience, Memory,
 Systemwide properties.
@@
 
-In this other paragraph we explain what the equator distance is. Then,
-this introductory section breaks and moves on to the main findings
+\begin{figure}[h!]
+  \centering
+  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
+
+  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
+
+  \caption{The value of the counters after $s=10,000$ writes shows the autocorrelation in the counters in autoassociative memories (``x at x'').
+The histogram was obtained through simulation.
+The red curve is the theoretical normal distribution found in equations (\ref{cntaddr1}) and (\ref{cntaddr0}).}
+  \label{fig:DELETE-sdm-corr-counters}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+
+\includegraphics[width=\columnwidth]{./images02/autocorrelation/autocorrelation_n=1000_closeview=False.eps}
+
+\caption{The threshold size $r$ of hard-locations bring about the autocorrelation of Theorem \ref{T1}.
+Consider, for instance, the extremes \{0, $n$\}.
+Let us start with $n$ (in our case, $n=1000$ dimensions): given the information that $d(x,y)\le n$, the probability of a bitmatch in dimension $i$ is $1/2$; as $\forall x,y \in \{0, 1\}^N, d(x,y)\le n$.
+At the other extreme, consider that we have the information that $d(x,y)\le 0$: in this case $x=y$ and the probability of a bitmatch, in any dimension, is 1.
+The autocorrelation hence drops monotonically until convergence at $1/2$ as the distance grows.
+Numerically, our results converge to precisely $1/2$ only after $d(x,y)\ge 607$ for $n=1000$; $d(x,y)\ge 5332$ for $n=10000$, and $d(x,y)\ge 83$ for $n=100$.
+This will be seen in Lemma \ref{thm:sdm-access-radius}.
+\label{fig:bitmatchbias}}
+\end{figure}
+
+In this other paragraph we explain what the equator distance is.
+
+Then, this introductory section breaks and moves on to the main findings
 of this paper.
 
-\begin{figure}[!htb]
-\centering\includegraphics{./images02/calculated-table-72.png}
-\caption{Kanerva's original Figure 7.3 generated using the equations from \citet{brogliato2014sparse}.
-\label{fig:kanerva-figure-73-calculated}
-}
-\end{figure}
-
-\input{./Chapters/sdm-autocorrelation.tex}
+\input{Chapters/sdm-autocorrelation.tex}
 

### 2018-08-22 delete.py
??

@@ -63,6 +63,11 @@ cbar.add_lines(CS2)
 
 plt.figure()
 
+λ=1.0
+Ωᵟ = 1
+λ ∈ sdm.ai(Ωᵟ)
+
+
 # Now make a contour plot with the levels specified,
 # and with the colormap generated automatically from a list
 # of colors.

### 2018-08-22 Chapters/hathor-bitcoin-math.tex
Merge branch 'master' of https://github.com/Alex-Linhares/msbrogli-phd

@@
 
-The primary objective of this chapter is to increase the understanding of the Bitcoin through mathematical tools.
+The primary objective of this chapter is to increase the understanding of Bitcoin through mathematical tools.
 
@@
 
-In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average.
-Suppose that the Bitcoin's network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network.
+In the Bitcoin protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average.
+Suppose that the Bitcoin network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network.
 The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block.
@@
 The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\eta H} \right)^{tH}$.
-But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.
+But, as the Bitcoin network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.
 
@@
 During 2017, the worst daily 90\% percentile was 21 seconds.
-Notice that both results may not be contradictory because Bitcoin network is continuously evolving.
+Notice that both results may not be contradictory because the Bitcoin network is continuously evolving.
 
@@
 % Análise: Mudança no network's hash rate antes de atualizar o $A$.
+
+Bitcoin network is open and miners can come and go as they wish, and their movements affect the network's total hash rate.
+From the theorems developed so far, we can predict how the network behaves in a constant hash rate scenario.
+But how a change in hash rate affects the network is still to be explored, and that is the goal of this section.
 
@@
 
-Let's say that the network's total hash rate has suddenly multiplied by $\alpha$.
+The most simple model for a change in the network's total hash rate is to consider a sudden shift as if it has jumped to a new value without a continuous transition to it.
+
+Let's say that the network's total hash rate has been suddenly multiplied by $\alpha$.
 So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\eta}$.
@@
 
+Now let's model a change in the network's total hash rate using a continuous and smoothly function.
+In this model, the hash rate is multiplied by a function that continuously goes from $1$ towards a new value $a$ over time.
+The smoothness of the transition --- or how fast it changes --- is also parameterized.
+
 Let $u(t) = \frac{1+abx}{1+bx}$.
@@
 It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a$, which yields $\int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
-$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
+$$F_T(t) = \mathbf{P}(T \leq t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
+
+Let $f_T(t) = \frac{d}{dt} F_T(t)$ be the probability density function, then:
+
 $$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
@@
 
-Let $\mathcal{L}$ be the Laplace Transform. Thus,
+To understand how the change in the hash rate affects the time between blocks, we will calculate the average time between blocks, i.e., $\mathbf{E}[T]$.
+For this, let's use the well-known result that the expected value of a random variable is equal to minus the derivative of the Laplace transformation of its probability density function when $s=0$.
+
+Let $\mathcal{L}$ be the Laplace Transform operator, then $\mathbf{E}[T] = - \frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0}$.
+First, we will calculate $\mathcal{L}\{F_T(t)\}$, and then we obtain $\mathcal{L}\{f_T(t)\}$ using $\mathcal{L}\{F_T(t)\}$.
+Thus,
+
 \begin{align*}
@@
 
-Finally, as $\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0}$,
+Finally, as $\mathbf{E}[T] = - \frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0}$,
 $$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n!
@@
 
-In order to compare the hash rate change models, namely (i) suddenly change, (ii) smoothly change, and (iii) linear change, I have applied each of them to the same scenarios.
+In order to compare the hash rate change models, namely (i) suddenly changing, (ii) smoothly changing, and (iii) linearly changing, I have applied each of them to the same scenarios.
 In the first scenario, the hashrate will double in the next week, whereas, in the second scenario, it will halve in the next week.
 
-In both the smoothly change model and the linear change model, I could have calculated the average time between blocks during the whole week.
-It would not give us much information, because the time between models would be increasing (or reducing) more and more as the days goes by.
-And we are really interested in the average time between blocks throughout the days, and not the average of the whole week.
+In both the smoothly change model and the linear change model, I could have calculated each model's average time between blocks during one week.
+But, it would not give us much information, because the estimated average time between models would be increasing (or decreasing) more and more as the days goes by.
+And we are really interested in the average time between blocks throughout the days, and not the average of one week.
 
@@
 
-I compared both the smoothly change model and the linear change model with the suddenly change model.
+I compared both the smoothly change model and the linear change model with the suddenly changing model.
 The difference between them is negligible.
 Let $\epsilon$ be the maximum absolute error between the models, than $\epsilon < 0.8$ and $\epsilon/H < 0.2\%$, for all intervals.
-The maximum absolute error between the linear and the suddenly change models can be seen in Figure \ref{fig:bitcoin-H-change-error}.
+The maximum absolute error between the linear and the suddenly changing models can be seen in Figure \ref{fig:bitcoin-H-change-error}.
 
@@
 
-Therefore, we may conclude that it is reasonable to approximate the average time between blocks using only the suddenly change model in each interval of one hour.
+Therefore, we may conclude that it is reasonable to approximate the average time between blocks using only the suddenly changing model in each interval of one hour.
 
 The average time between blocks throughout the days can be seen in Figure \ref{fig:bitcoin-H-change}.
-It was calculated using the suddenly change model with the hash rate changing linearly during the week.
+It was calculated using the suddenly changing model with the hash rate changing linearly during the week.
 
@@
 
-There are many possible ways to attack the Bitcoin \citep{karame2012two, heilman2015eclipse, bahack2013theoretical, bonneau2016buy, neudecker2015simulation}.
+There are many possible ways to attack the Bitcoin network \citep{karame2012two, heilman2015eclipse, bahack2013theoretical, bonneau2016buy, neudecker2015simulation}.
 In this section, we are interested in a particular attack: the double spending attack.
@@
 
-In the Bitcoin's network, the average number of transactions per block is $s=2,250$, so, the system is stable when $\lambda_{\text{TX}} < 2,250/600 = 3.75$ tx/s.
-Therefore, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle.
+In the Bitcoin network, the average number of transactions per block is $s=2,250$, so, the system is stable when $\lambda_{\text{TX}} < 2,250/600 = 3.75$ tx/s.
+Therefore, $3.75$ is the maximum number of new transactions per second that the Bitcoin network may handle.
 When $\lambda_{\text{TX}} > 3.75$ tx/s, the unconfirmed transaction pool starts to grow indefinitely.
@@
 Thus, the average number of unconfirmed transactions $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$.
-In the Bitcoin's network, $m \ll s$ is reached when $\lambda_{\text{TX}} \ll 3.75$ tx/s.
+In the Bitcoin network, $m \ll s$ is reached when $\lambda_{\text{TX}} \ll 3.75$ tx/s.
 In other words, when the number of new transactions per second is way smaller than 3.75 tx/s, the average waiting time of a transaction to be confirmed is 600 seconds, which means, on average, all transactions will be confirmed in the next block.
@@
 
-Therefore, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
+Therefore, we conclude that the Bitcoin network capacity is $\lambda_{\text{blocks}} = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
 

### 2018-08-23 paperIEEE.tex
Intro; outline; etc.

@@
 \usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage{calc}
@@
  pdfpagelayout=OneColumn, pdfnewwindow=true, pdfstartview=XYZ, plainpages=false}
+
+ \makeatletter
+ \def\endthebibliography{%
+   \def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}%
+   \endlist
+ }
+ \makeatother
+
+ \usepackage[round]{natbib}
 
@@
 
+\setcitestyle{square}
+
 \begin{document}
@@
 \maketitle
+
 \begin{abstract}
@@
 
-\IEEEPARstart{S}{parse distributed memory} has been shown to be
-fodastic.
+\IEEEPARstart{S}{parse distributed memory} (SDM) \citep{Kanerva1988} (see also \citep{denning_sparse_1989,  flynn_sparse_1989, kanerva_sparse_1993, kanerva_parallel_1985, goos_binary_1996, kanerva_hyperdimensional_2009,  keeler_comparison_1988, rwcp_fully_1997, marinaro_spatter_1994, sahlgren_permutations_nodate, uesaka_foundations_2001}) is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics.
+This model is used in all sort of applications due to its incredible ability to closely reflect the human capacity to remember past experiences from the subtlest of clues.
+Applications range from call admission control \citep{kwon_atm_1998, hee-yong_kwon_atm_1997}, to behavior-based robotics \citep{Rajesh1998, mendes2008robot, jockel_sparse_2009}, to noise filtering \citep{meng_modified_2009}, etc.
 
-In this paragraph we explain what SDM is.
+\subsection{Desiderata for a theory of memory}
+
+SDM reads like a \emph{desiderata} of a theory for human long-term memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
+
+\begin{enumerate}
+    \item Why are most concepts orthogonal, unrelated to each other?
+    \item Why is there Miller's magic number, i.e., we can't hold too many things in mind at once?
+    \item Why do we at times instantly recall an experience; other times we can't recall anything at all; and still other times we get into this strange tip-of-the-tongue situation...
+it is clearly `there'...
+but where?
+(and `what' is `there'?\footnote{Psychologists have documented interesting properties about this state:  it happens, for instance, around every 1-week cycle; and it happens mostly over proper names, subjects have access to the first letter and to the number of syllables better than chance, etc.
+Also interesting is their clever way of triggerring the state: through definitions of rarely-used words.
+For some psychological studies about the tip-of-tongue, see \citet{meyer_tip---tongue_1992, brown_review_1991, brown_tip_1966}.}.)
+    \item How does this recall process work?  What is remembering?
+    \item Why do neurons die and we still remember most everything?
+    \item What do neurons actually do?  What is their primary function?
+\end{enumerate}
+
+These are some of the questions touched by SDM.  It is a theory of long-term memory that constructs the process of recall from a very microscopic level of neuron-firings.
 
@@
   \centering
-  \subfloat[$\text{addr}_i=1$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr1.png}}
+  \subfloat[$Q_3$]{\includegraphics[width=0.4\columnwidth]{./images02/new-images/qn3.png}}
+  \subfloat[$Q_7$]{\includegraphics[width=0.4\columnwidth]{./images02/new-images/qn7.png}}
 
-  \subfloat[$\text{addr}_i=0$]{\includegraphics[width=0.5\textwidth]{./images02/autocorrelation/x_at_x_addr0.png}}
+  \subfloat[$Q_{10}$]{\includegraphics[width=\columnwidth]{./images02/new-images/qn10.png}}
+  \caption{Kanerva models the space of \emph{possible incoming neural signals} as an $n$-dimensional \emph{hypercube}, or $Q_n$.
+Here we have $Q_n$, for $n \in$ \{3, 7, 10\}.
+Each node corresponds to a bitstring in $\{0,1\}^n$, and two nodes are linked iff the bitstrings differ by a single dimension.
+A number of observations can be made here.
+First, the number of nodes grows as $O(2^n)$; which makes the space rapidly intractable.
+Another interesting observation is that most of the space lies `at the core' of the hypercube, at a distance of around $n/2$ from any given vantage point.\label{hypercubes}}
+\end{figure}
 
-  \caption{The value of the counters after $s=10,000$ writes shows the autocorrelation in the counters in autoassociative memories (``x at x'').
-The histogram was obtained through simulation.
-The red curve is the theoretical normal distribution found in equations (\ref{cntaddr1}) and (\ref{cntaddr0}).}
-  \label{fig:DELETE-sdm-corr-counters}
+\subsection{Review of the model}
+
+IN SDM, the data --- and address space on which it is stored --- are represented by large sequences of bits, called \emph{bitstrings}.
+The \emph{Hamming distance} provides comparisons between bitstrings and is used as a metric for the system.
+The Hamming distance is defined for two bitstrings of equal length as the number of positions in which bits differ.
+For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
+
+The space studied by Kanerva is also called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}.
+For a fixed $n \in \mathbb{Z}$, the graph $G = (V, E)$, in which $v \in V$ iff there is a bijective function $b: V\to \{0,1\}^n$ and $(v_i, v_j) \in E$ iff $H(b(v_i), b(v_j))=1$, where $H$ is the Hamming distance.
+That is, $n$-sized bitstrings correspond to nodes, and edges exist between nodes $iff$ they flip a single bit.
+Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community: it is a bipartite graph with chromatic number 2, and it is planar only if $n\le3$.
+The proofs, though elementary, illuminate some of its properties:  vertices connect bitstrings with an even number of 1's with bitstrings with an odd number of 1's, therefore (i) it is \emph{bipartite} and (ii) it \emph{may be colored with two colors}.
+As for planarity, the case $n=3$ is the largest planar one.
+A good survey is provided by \citep{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.
+
+One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, our 3-dimensional space.
+
+A major difference is that in the 3-d world, there exist very far points, while in $Q_n$ the largest possible distance is $n$.
+Another one is that the number of directions in $Q_n$ is enormous:  at \emph{every point} there are $n$ directions to choose from.
+
+Though both follow the triangle inequality ($d(A, C) \le d(A, B) + d(B, C)$), which in 3-d Euclidean distance may be loosely interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{ and } d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
+Hence, it makes no sense to say that $A$ is also close to $C$.
+
+There are numerous, beautiful, counter-intuitive notions involved in this space. This difference in intuition may trick even experienced researchers when analyzing some situations.
+
+Unlike traditional memory used by computers, SDM performs read and write operations in a multitude of addresses, also called neurons.
+That is, the data is not written, or it is not read in a single address spot, but in many addresses.
+These are called activated addresses, or activated neurons.
+
+The activation of addresses takes place according to their distances from the datum.
+Suppose one is writing datum $\eta$ at address $\xi$, then all addresses inside a circle with center $\xi$ and radius $r$ are activated.
+So, $\eta$ will be stored in all these activated addresses, which are around address $\xi$, such as in Figure \ref{fig-addresses-inside-access-radius}.
+An address $\xi'$ is inside the circle if its Hamming distance to the center $\xi$ is less than or equal to the radius $r$, i.e.
+$distance(\xi,\xi')\leq r$.
+
+\begin{figure}[!htb]
+\centering\includegraphics[scale=0.75]{./images02/p_circle_r.pdf}
+
+\caption{Activated addresses inside access \protect \\
+radius $r$ around the center address.\label{fig-addresses-inside-access-radius}}
 \end{figure}
+
+Every write or read in SDM memory activates a number of addresses with close distance.
+The data is written in these activated addresses or read from them.
+These issues will be addressed in due detail further on, but a major difference from a traditional computer memory is that the data are always stored and retrieved in a multitude of addresses.
+This way SDM memory has robustness against loss of addresses (e.g., death of a neuron).
+
+In traditional memory, each datum is stored in an address and every lookup of a specific datum requires a search through the memory.
+In spite of computer scientists having developed beautiful algorithms to perform fast searches, almost all of them do a precise search.
+That is, if you have an imprecise clue of what you need, these algorithms will simply fail.
+
+In SDM, the data space is the same as the address space, which amounts to a vectorial, binary space, that is, a $\{0,1\}^{n}$ space.
+This way, the addresses where the data will be written are the same as the data themselves.
+For example, the datum $\eta=00101_{b}\in\{0,1\}^{5}$ will be written to the address $\xi=\eta=00101_{b}$.
+If one chooses a radius of 1, the SDM will activate all addresses one bit away or less from the center address.
+So, the datum $00101_{b}$ will be written to the addresses $00101_{b}$, $10101_{b}$, $01101_{b}$, $00001_{b}$, $00111_{b}$, and $00100_{b}$.
+
+In this case, when one needs to retrieve the data, one could have an imprecise cue at most one bit away from $\eta$, since all addresses one bit away have $\eta$ stored in themselves.
+Extending this train of thought for larger dimensions and radius, exponential numbers of addresses are activated and one can see why SDM is a distributed memory.
+
+When reading a cue $\eta_{x}$ that is $x$ bits away from $\eta$, the cue shares many addresses with $\eta$.
+The number of shared addresses decreases as the cue's distance to $\eta$ increases, in other words, as $x$ increases.
+This is shown in Figure \ref{fig-shared-addresses}.
+The target datum $\eta$ was written in all shared addresses, thus they will bias the read output in the direction of $\eta$.
+If the cue is sufficiently near the target datum $\eta$, the read output will be closer to $\eta$ than $\eta_{x}$ was.
+Repeating the read operation increasingly gets results closer to $\eta$, until it is precisely the same.
+So, it may be necessary to perform more than one read operation to converge to the target data $\eta$.
+
+\begin{figure}[!htb]
+\centering\includegraphics[scale=0.75]{./images02/p1_inter_p2.pdf}
+
+\caption{Shared addresses between the \protect \\
+target datum $\eta$ and the cue $\eta_{x}$. \label{fig-shared-addresses}}
+\end{figure}
+
+The addresses of the $\{0,1\}^{n}$ space grow exponentially with the number of dimensions $n$, i.e., $N=2^{n}$.
+For $n=100$ we have $N\approx10^{30}$, which is incredibly large when related to computer memory.
+Furthermore, \citep{Kanerva1988} suggests $n$ between 100 and 10,000.
+To solve the feasibility problem of implementing this memory, Kanerva made a random sample of $\{0,1\}^{n}$, in his work, having $N'$ elements.
+All these addresses in the sample are called hard locations.
+Other elements of $\{0,1\}^{n}$, not in $N'$, are called virtual neurons.
+This is represented in Figure \ref{fig-hardlocations}.
+All properties of reading and write operations presented before remain valid but limited to hard locations.
+Kanerva suggests taking a sample of about one million hard locations.
+
+Using this sample of binary space, our data space does not exist completely.
+That is, the binary space has $2^{n}$ addresses, but the memory is far away from having these addresses available.
+In fact, only a fraction of this vectorial space is actually instantiated.
+Following Kanerva's suggestion of one million hard locations, for $n=100$, only $100\cdot10^{6}/2^{100}=7\cdot10^{-23}$ percent of the whole space exists, and for $n=1,000$ only $100\cdot10^{6}/2^{1000}=7\cdot10^{-294}$ percent.
+
+Kanerva also suggests the selection of a radius that will activate, on average, one-thousandth of the sample, which is 1,000 hard locations for a sample of one million addresses.
+In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$.
+We think that a 256-dimensional memory may be important because it presents conformity to Miller's magic number \citep{Linhares2011}.
+
+\begin{figure}[!htb]
+\centering\includegraphics[scale=0.75]{./images02/hardlocations.pdf}
+
+\caption{Hard-locations randomly sampled from binary space.\label{fig-hardlocations}}
+\end{figure}
+
+Since a cue $\eta_{x}$ near the target bitstring $\eta$ shares many hard locations with $\eta$, SDM can retrieve data from imprecise cues.
+Despite this feature, it is very important to know how imprecise this cue could be while still giving accurate results.
+What is the maximum distance from our cue to the original data that still retrieves the right answer?
+An interesting approach is to perform a read operation with a cue $\eta_{x}$, that is $x$ bits away from the target $\eta$.
+Then measure the distance from the read output and $\eta$.
+If this distance is smaller than $x$ we are converging.
+Convergence is simple to handle, just read again and again, until it converges to the target $\eta$.
+If this distance is greater than $x$ we are diverging.
+Finally, if this distance equals $x$ we are in a tip-of-the-tongue process.
+A tip-of-the-tongue psychologically happens when you know that you know, but you can't say what exactly it is.
+In SDM mathematical model, a tip-of-the-tongue process takes infinite time to converge.
+\citet{Kanerva1988} called this $x$ distance, where the read's output averages $x$, the critical distance.
+Intuitively, it is the distance from which smaller distances converge and greater distances diverge.
+In Figure \ref{fig-p1-p2-iterative-read}, the circle has radius equal to the critical distance and every $\eta_{x}$ inside the circle should converge.
+The figure also shows convergence in four readings.
+
+\begin{figure}[!htb]
+\centering\includegraphics[scale=0.75]{./images02/p1_p2_iter_read.pdf}
+
+\caption{In this example, four iterative readings were\protect \\
+required to converge from $\eta_{x}$ to $\eta$.\label{fig-p1-p2-iterative-read}}
+\end{figure}
+
+The $\{0,1\}^{n}$ space has $N=2^{n}$ locations from which we instantiate $N'$ samples.
+Each location in our sample is called a hard location.
+On these hard locations, we do operations of read and write.
+One of the insights of SDM is exactly the way we read and write: using data as addresses in a distributed fashion.
+Each datum $\eta$ is written in every activated hard location inside the access radius centered on the address, that equals datum, $\xi=\eta$.
+Kanerva suggested using an access radius $r$ having about one-thousandth of $N'$.
+As an imprecise cue $\eta_{x}$ shares hard locations with the target bitstring $\eta,$ it is possible to retrieve $\eta$ correctly.
+(Actually, probably more than one read is necessary to retrieve exactly $\eta.)$.
+Moreover, if some neurons are lost, only a fraction of the datum is lost and it is possible that the memory can still retrieve the right datum.
+
+A random bitstring is generated with equal probability of $0$'s and $1$'s for each bit.
+One can readily see that the average distance between two random bitstrings follows the binomial distribution with mean $n/2$ and standard deviation $\sqrt{n/4}$.
+For a large $n$, most of the space lies close to the mean and has fewer shared hard locations.
+As two bitstrings with distance far from $n/2$ are very improbable, \citet{Kanerva1988} defined that two bitstrings are orthogonal when their distance is $n/2$.
+
+The write operation needs to store, for each dimension bit which happened more ($0$'s or $1$'s).
+This way, each hard location has $n$ counters, one for each dimension.
+The counter is incremented for each bit $1$ and decremented for each bit $0$.
+Thus, if the counter is positive, there have been more $1$'s than $0$'s, if the counter is negative, there have been more $0$'s than $1$'s, and if the counter is zero, there have been an equal number of $1$'s and $0$'s.
+Table \ref{tab:write operation} shows an example of a write operation being performed in a 7-dimensional memory.
+
+\begin{table}
+\begin{tabular}{c|c|c|c|c|c|c|c|}
+\cline{2-8}
+$\eta$ & 0 & 1 & 1 & 0 & 1 & 0 & 0\tabularnewline
+\cline{2-8}
+$\xi_{\textit{before}}$ & 6 & -3 & 12 & -1 & 0 & 2 & 4\tabularnewline
+\cline{2-8}
+\multicolumn{1}{c}{} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}}\tabularnewline
+\cline{2-8}
+$\xi_{\textit{\textit{after}}}$ & \textbf{5} & \textbf{-2} & \textbf{13} & \textbf{-2} & \textbf{1} & \textbf{1} & \textbf{3}\tabularnewline
+\cline{2-8}
+\end{tabular}
+
+\caption{Write operation example in a 7-dimensional memory of data $\eta$
+being written to $\xi$, one of the activated addresses.\label{tab:write operation}}
+
+\end{table}
+
+The read is performed polling each activated hard location and statistically choosing the most written bit for each dimension.
+It consists of adding all $n$ counters from the activated hard locations and, for each bit, choosing bit 1 if the counter is positive, choosing bit 0 if the counter is negative, and randomly choose bit 0 or 1 if the counter is zero.
+
+\section{Neurons as pointers}
+
+One interesting view is that neurons in SDM work like pointers.
+As we write bitstrings in memory, the hard locations' counters are updated and some bits are flipped.
+Thus, the activated hard locations do not necessarily point individually to the bitstring that activated it, but together they point correctly.
+In other words, the read operation depends on many hard locations to be successful.
+This effect is represented in Figure \ref{fig-p1-pointers}: where all hard locations inside the circle are activated and they, individually, do not point to $\eta$.
+But, like vectors, adding them up points to $\eta$.
+If another datum $\nu$ is written into the memory near $\eta$, the shared hard locations will have information from both of them and would not point to either.
+All hard locations outside of the circle are also pointing somewhere (possibly other data points).
+This is not shown, however, in order to keep the picture clean and easily understandable.
+
+\begin{figure}[!htb]
+\centering\includegraphics[scale=0.75]{./images02/p1_after_write.pdf}
+
+\caption{Hard-locations pointing, approximately, to the target bitstring.\label{fig-p1-pointers}}
+\end{figure}
+
+\section{Concepts}
+
+Although Kanerva does not mention concepts directly in his book \citep{Kanerva1988}, the author's interpretation is that each bitstring may be mapped to a concept.
+Thus, unrelated concepts are orthogonal and concepts could be linked through a bitstring near both of them.
+For example, ``beauty'' and ``woman'' have distance $n/2$, but a bitstring that means ``beautiful woman'' could have distance $n/4$ to both of them.
+As a bitstring with distance $n/4$ is very improbable, it is linking those concepts together.
+\citet{Linhares2011} approached this concept via ``chunking through averaging''.
+
+Due to the distribution of hard locations between two random bitstrings, the vast majority of concepts is orthogonal to all others.
+Consider a non-scientific survey during a cognitive science seminar, where students asked to mention ideas unrelated to the course brought up terms like birthdays, boots, dinosaurs, fever, executive order, x-rays, and so on.
+Not only are the items unrelated to cognitive science, the topic of the seminar, but they are also unrelated to each other.
+
+For any two memory items, one can readily find a stream of thought relating two such items (``Darwin gave dinosaurs the boot''; ``she ran a fever on her birthday''; ``isn't it time for the Supreme Court to x-ray that executive order?'', ...
+and so forth).
+Robert French presents an intriguing example in which one suddenly creates a representation linking the otherwise unrelated concepts of ``coffee cups'' and ``old elephants'' \citep{French1997}.
+
+This mapping from concepts to bitstrings brings us two central questions: (i) Suppose we have a bitstring that is linking two major concepts.
+How do we know which concepts are linked together?
+(ii) From a concept bitstring, how can we list all concepts that are somehow linked to it?
+This second question is called the problem of spreading activation.
+
+\section{Read operation}
+
+In his work, Kanerva proposed and analyzed a read algorithm called here Kanerva's read.
+His read takes all activated hard locations counters and sum them.
+The resulting bitstring has bit $1$ where the result is positive, bit $0$ where the result is negative, and a random bit where the result is zero.
+In a word, each bit is chosen according to all written bitstrings in all hard locations, being equal to the bit more appeared.
+Table \ref{tab:kanerva-read} shows an example of Kanerva's read result bitstring.
 
@@
 
-\section{Conclusions}
+\section{Conclusion}
 
-bla bla
+\subsection{Effects on system-wide properties}
+The skeptical reader may ask: `Beyond theory; what are the consequences of this work?  How does it apply, if at all, to building systems that incorporate SDM?'
 
-\appendices{}
+Our conclusion is that this work is of special relevance as it brings about some \emph{systemwide properties} that models need to take into account.
+The first decision choice involved in building a system with SDM regards the size of $n$, the number of dimensions of the memory.
+Here, complementing the work of XXXXXXXX, we show that as $n$ grows, autocorrelation tends to $0$.
+But recall that as $n$ grows, the size of the critical distance also tends to $0$.
+It is therefore crucial to select a value for $n$ in which the model will retain its convergence properties, with the understanding of the possible effects brought about by autocorrelation.
+Systemwide properties such as these affect the entire behavior of the model and, as such, should always be taken into account by theoreticians and modelers alike.
+
+\subsection{Future Research}
+
+BPs, System1 vs System2, HRRs, etc.
 

### 2018-08-23 paperIEEE.tex
small

@@
 A good survey is provided by \citep{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.
-
-One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, our 3-dimensional space.
-
-A major difference is that in the 3-d world, there exist very far points, while in $Q_n$ the largest possible distance is $n$.
-Another one is that the number of directions in $Q_n$ is enormous:  at \emph{every point} there are $n$ directions to choose from.
-
-Though both follow the triangle inequality ($d(A, C) \le d(A, B) + d(B, C)$), which in 3-d Euclidean distance may be loosely interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{ and } d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings.
-Hence, it makes no sense to say that $A$ is also close to $C$.
-
-There are numerous, beautiful, counter-intuitive notions involved in this space. This difference in intuition may trick even experienced researchers when analyzing some situations.
 

### 2018-09-08 paperIEEE.tex
IEEE paper

@@
 
-\subsection{Desiderata for a theory of memory}
-
-SDM reads like a \emph{desiderata} of a theory for human long-term memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
-
-\begin{enumerate}
-    \item Why are most concepts orthogonal, unrelated to each other?
-    \item Why is there Miller's magic number, i.e., we can't hold too many things in mind at once?
-    \item Why do we at times instantly recall an experience; other times we can't recall anything at all; and still other times we get into this strange tip-of-the-tongue situation...
-it is clearly `there'...
-but where?
-(and `what' is `there'?\footnote{Psychologists have documented interesting properties about this state:  it happens, for instance, around every 1-week cycle; and it happens mostly over proper names, subjects have access to the first letter and to the number of syllables better than chance, etc.
-Also interesting is their clever way of triggerring the state: through definitions of rarely-used words.
-For some psychological studies about the tip-of-tongue, see \citet{meyer_tip---tongue_1992, brown_review_1991, brown_tip_1966}.}.)
-    \item How does this recall process work?  What is remembering?
-    \item Why do neurons die and we still remember most everything?
-    \item What do neurons actually do?  What is their primary function?
-\end{enumerate}
-
-These are some of the questions touched by SDM.  It is a theory of long-term memory that constructs the process of recall from a very microscopic level of neuron-firings.
-
 \begin{figure}[h!]
@@
 
-IN SDM, the data --- and address space on which it is stored --- are represented by large sequences of bits, called \emph{bitstrings}.
+IN SDM, the data --- and address space on which it is stored --- are represented by large \emph{bitstrings}.
 The \emph{Hamming distance} provides comparisons between bitstrings and is used as a metric for the system.
-The Hamming distance is defined for two bitstrings of equal length as the number of positions in which bits differ.
-For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2.
-
-The space studied by Kanerva is also called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}.
+This space is also called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}.
 For a fixed $n \in \mathbb{Z}$, the graph $G = (V, E)$, in which $v \in V$ iff there is a bijective function $b: V\to \{0,1\}^n$ and $(v_i, v_j) \in E$ iff $H(b(v_i), b(v_j))=1$, where $H$ is the Hamming distance.
@@
 Thus, if the counter is positive, there have been more $1$'s than $0$'s, if the counter is negative, there have been more $0$'s than $1$'s, and if the counter is zero, there have been an equal number of $1$'s and $0$'s.
-Table \ref{tab:write operation} shows an example of a write operation being performed in a 7-dimensional memory.
-
-\begin{table}
-\begin{tabular}{c|c|c|c|c|c|c|c|}
-\cline{2-8}
-$\eta$ & 0 & 1 & 1 & 0 & 1 & 0 & 0\tabularnewline
-\cline{2-8}
-$\xi_{\textit{before}}$ & 6 & -3 & 12 & -1 & 0 & 2 & 4\tabularnewline
-\cline{2-8}
-\multicolumn{1}{c}{} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ +1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}} & \multicolumn{1}{c}{\textcolor{red}{\small{}$\Downarrow$ -1}}\tabularnewline
-\cline{2-8}
-$\xi_{\textit{\textit{after}}}$ & \textbf{5} & \textbf{-2} & \textbf{13} & \textbf{-2} & \textbf{1} & \textbf{1} & \textbf{3}\tabularnewline
-\cline{2-8}
-\end{tabular}
-
-\caption{Write operation example in a 7-dimensional memory of data $\eta$
-being written to $\xi$, one of the activated addresses.\label{tab:write operation}}
-
-\end{table}
 
@@
 
-\section{Concepts}
-
-Although Kanerva does not mention concepts directly in his book \citep{Kanerva1988}, the author's interpretation is that each bitstring may be mapped to a concept.
-Thus, unrelated concepts are orthogonal and concepts could be linked through a bitstring near both of them.
-For example, ``beauty'' and ``woman'' have distance $n/2$, but a bitstring that means ``beautiful woman'' could have distance $n/4$ to both of them.
-As a bitstring with distance $n/4$ is very improbable, it is linking those concepts together.
-\citet{Linhares2011} approached this concept via ``chunking through averaging''.
-
-Due to the distribution of hard locations between two random bitstrings, the vast majority of concepts is orthogonal to all others.
-Consider a non-scientific survey during a cognitive science seminar, where students asked to mention ideas unrelated to the course brought up terms like birthdays, boots, dinosaurs, fever, executive order, x-rays, and so on.
-Not only are the items unrelated to cognitive science, the topic of the seminar, but they are also unrelated to each other.
-
-For any two memory items, one can readily find a stream of thought relating two such items (``Darwin gave dinosaurs the boot''; ``she ran a fever on her birthday''; ``isn't it time for the Supreme Court to x-ray that executive order?'', ...
-and so forth).
-Robert French presents an intriguing example in which one suddenly creates a representation linking the otherwise unrelated concepts of ``coffee cups'' and ``old elephants'' \citep{French1997}.
-
-This mapping from concepts to bitstrings brings us two central questions: (i) Suppose we have a bitstring that is linking two major concepts.
-How do we know which concepts are linked together?
-(ii) From a concept bitstring, how can we list all concepts that are somehow linked to it?
-This second question is called the problem of spreading activation.
-
 \section{Read operation}
@@
 Table \ref{tab:kanerva-read} shows an example of Kanerva's read result bitstring.
-
-\begin{figure}[h!]
-\centering
-
-\includegraphics[width=\columnwidth]{./images02/autocorrelation/autocorrelation_n=1000_closeview=False.eps}
-
-\caption{The threshold size $r$ of hard-locations bring about the autocorrelation of Theorem \ref{T1}.
-Consider, for instance, the extremes \{0, $n$\}.
-Let us start with $n$ (in our case, $n=1000$ dimensions): given the information that $d(x,y)\le n$, the probability of a bitmatch in dimension $i$ is $1/2$; as $\forall x,y \in \{0, 1\}^N, d(x,y)\le n$.
-At the other extreme, consider that we have the information that $d(x,y)\le 0$: in this case $x=y$ and the probability of a bitmatch, in any dimension, is 1.
-The autocorrelation hence drops monotonically until convergence at $1/2$ as the distance grows.
-Numerically, our results converge to precisely $1/2$ only after $d(x,y)\ge 607$ for $n=1000$; $d(x,y)\ge 5332$ for $n=10000$, and $d(x,y)\ge 83$ for $n=100$.
-This will be seen in Lemma \ref{thm:sdm-access-radius}.
-\label{fig:bitmatchbias}}
-\end{figure}
 
@@
 
+\subsection{Availability of the Code}
+
+OpenCL
+python (show example with figure)
+Technical Documentation
+
+Open-source Book
+
+another article with a larger overview of the developed system so far (Reinforcement learning, critical distance, noise filtering and prototype theory, and pattern recognition.)
+This --> transparency --> replicability -> incremental improvement --> parallel explorations by different teams around the world.
+
 \section*{Acknowlegment}
@@
 José Ricardo de Almeida Torreão, Moacyr Alvim Horta Barbosa da Silva,
-Flavio Codeço Coelho, Horacio Hideki Yanasse, and Paulo Murilo Castro
-de Oliveira for their careful, comprehensive, reviews of the first
-author\textquoteright s theses.
+Flavio Codeço Coelho and Paulo Murilo Castro de Oliveira for their careful, comprehensive, reviews of the first author\textquoteright s theses.
 

### 2018-09-20 paperIEEE.tex
minor

@@
 
- \usepackage[round]{natbib}
+ \usepackage[round, numbers]{natbib}
 
@@
 
-OpenCL
+OpenCL for the partial binomail sum (cite `Inference-Based Similarity Search in Randomized Montgomery Domains for Privacy-Preserving Biometric Identification')
 python (show example with figure)
@@
 
-The authors would like to thank Pentti Kanerva, Eric Paul Nichols,
-José Ricardo de Almeida Torreão, Moacyr Alvim Horta Barbosa da Silva,
-Flavio Codeço Coelho and Paulo Murilo Castro de Oliveira for their careful, comprehensive, reviews of the first author\textquoteright s theses.
+The authors would like to thank Pentti Kanerva, Eric Paul Nichols, José Ricardo de Almeida Torreão, Moacyr Alvim Horta Barbosa da Silva, Flavio Codeço Coelho and Paulo Murilo Castro de Oliveira for their careful, comprehensive, reviews of the first author\textquoteright s theses.
 
@@
 1.25\,in!%
-\end{minipage}}}]{Your Name}
- All about you and the what your interests are.
+\end{minipage}}}]{Marcelo Salhab Brogliato} received the Ph.D. degree from Getulio Vargas Foundation in 2018. Vialink, sdm.ai,
 \end{IEEEbiography}
 
-\begin{IEEEbiographynophoto}{Coauthor}
-Same again for the co-author, but without photo
-\end{IEEEbiographynophoto}
+\begin{IEEEbiography}[{\fbox{\begin{minipage}[t][1.25in]{1in}%
+Replace this box by an image with a width of 1\,in and a height of
+1.25\,in!%
+\end{minipage}}}]{Alexandre Linhares} received the Ph.D.
+degree from the National Institute of Space Research, 2001.
+He is an Associate Professor at the Center of Behavioral Research, EBAPE / Getulio Vargas Foundation.
+He has published circa 30 papers in journals such as \emph{Artificial Intelligence, Biological Cybernetics, Cognitive Science, IEEE Transactions on Evolutionary Computation, Information Sciences, Frontiers in Psychology,} etc.
+\end{IEEEbiography}
 

### 2018-11-09 Chapters/sdm-autocorrelation.tex
Intro; outline; etc.
IEEE paper
minor
Alignment respecting column borders

@@
 \section{A deviation from the equator distance?}
+
+\citet{Kanerva1994}
 
@@
     \begin{align}
-    \frac{n-k}{n} \binom{n}{k} = \frac{(n-k)}{n} \frac{n!}{(n-k)! k!} = \frac{(n-1)!}{k! (n-1-k)!} = \binom{n-1}{k}
+    \frac{n-k}{n} \binom{n}{k} = \binom{n-1}{k}
     \end{align}
@@
 \begin{align}
-P(\text{cnt}_i \ge 0 | \text{addr}_i = 1) = P(\text{cnt}_i \le 0 | \text{addr}_i = 0) = 1 - \mathcal{N}.\text{cdf}(0)
+P(\text{cnt}_i \ge 0 | \text{addr}_i = 1) = P(\text{cnt}_i \le 0 | \text{addr}_i = 0) \\
+ = 1 - \mathcal{N}.\text{cdf}(0)
 \end{align}
@@
 \begin{align}
-\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \mathcal{N}(\mu = \mu_\text{cnt} \mu_Z, \sigma^2 = \sigma_\text{cnt}^2 \mu_Z + \mu_\text{cnt}^2 \sigma^2_Z) \label{eqn:sdm-eta1-addr1} \\
-\sum_{k=1}^{h-Z} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] &\sim \mathcal{N}(\mu = -\mu_\text{cnt} (1-p) \mu_h, \sigma^2 = \sigma^2_\text{cnt} (1-p) \mu_h + \mu_\text{cnt}^2 \sigma^2_{h-Z}) \label{eqn:sdm-eta1-addr0}
+\begin{split}
+\sum_{k=1}^{Z} \left[ \text{cnt}_k | \text{addr}_k=\eta_1 \right] &\sim \\
+& \mathcal{N}(\mu = \mu_\text{cnt} \mu_Z, \sigma^2 = \sigma_\text{cnt}^2 \mu_Z + \mu_\text{cnt}^2 \sigma^2_Z) \label{eqn:sdm-eta1-addr1}
+\end{split}
+\end{align}
+, and, respectively:
+
+\begin{align}
+\begin{split}
+& \sum_{k=1}^{h-Z} \left[ \text{cnt}_k | \text{addr}_k \ne \eta_1 \right] \sim \\
+& \mathcal{N}(\mu = -\mu_\text{cnt} (1-p) \mu_h, \sigma^2 = \sigma^2_\text{cnt} (1-p) \mu_h + \mu_\text{cnt}^2 \sigma^2_{h-Z}) \label{eqn:sdm-eta1-addr0}
+\end{split}
 \end{align}
@@
 
-\begin{align}
-P(wrong) &= P(\text{acc}_i < 0 | \eta_i = 1) \cdot P(\eta_i = 1) + P(\text{acc}_i > 0 | \eta_i = 0) \cdot P(\eta_i = 0) \\
-    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{1-\mathcal{N}_{\eta_i=0}.\text{cdf}(0)}{2} \\
-    &= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} \\
-    &= \mathcal{N}_{\eta_i=1}.\text{cdf}(0)
-\end{align}
+\begin{align*}
+P(wrong) = P(\text{acc}_i < 0 | \eta_i = 1) \cdot P(\eta_i = 1) {}& \\
++P(\text{acc}_i > 0 | \eta_i = 0) \cdot P(\eta_i = 0) \\
+= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{1-\mathcal{N}_{\eta_i=0}.\text{cdf}(0)}{2} \\
+= \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} + \frac{\mathcal{N}_{\eta_i=1}.\text{cdf}(0)}{2} \\
+= \mathcal{N}_{\eta_i=1}.\text{cdf}(0)
+\end{align*}
 

### 2018-11-09 paperIEEE.tex
Alignment respecting column borders
Photo 2

@@
  {hyperref}
+
 \hypersetup{pdftitle={Your Title},
@@
  \usepackage[round, numbers]{natbib}
+ \bibliographystyle{unsrtnat}
 
@@
 
-\markboth{Working Paper}{Brogliato and Linhares: Autocorrelation of SDM}
+\markboth{Working Paper}{Brogliato and Linhares: Autocorrelation in SDM}
 
@@
 
-\begin{IEEEbiography}[{\fbox{\begin{minipage}[t][1.25in]{1in}%
-Replace this box by an image with a width of 1\,in and a height of
-1.25\,in!%
-\end{minipage}}}]{Alexandre Linhares} received the Ph.D.
+%\begin{IEEEbiography}[{\fbox{\begin{minipage}[t][1.25in]{1in}{Linhares.jpg}
+
+\begin{IEEEbiography}[
+  {
+  \includegraphics[width=1in,height=1.25in,clip]{./Chapters/Linhares.jpg}
+  }
+  ]
+{Alexandre Linhares} received the Ph.D.
 degree from the National Institute of Space Research, 2001.
 He is an Associate Professor at the Center of Behavioral Research, EBAPE / Getulio Vargas Foundation.
-He has published circa 30 papers in journals such as \emph{Artificial Intelligence, Biological Cybernetics, Cognitive Science, IEEE Transactions on Evolutionary Computation, Information Sciences, Frontiers in Psychology,} etc.
+He has published ~30 papers in journals such as \emph{Artificial Intelligence, Biological Cybernetics, Cognitive Science, IEEE Transactions on Evolutionary Computation, Information Sciences, Frontiers in Human Neuroscience,} etc.
 \end{IEEEbiography}

### 2019-06-30 paperIEEE.tex
macbookpro compiling

@@
 \DeclareMathOperator\erf{erf}
-
-% !TEX root = ../partial-sdm.tex
 
@@
 He is an Associate Professor at the Center of Behavioral Research, EBAPE / Getulio Vargas Foundation.
-He has published ~30 papers in journals such as \emph{Artificial Intelligence, Biological Cybernetics, Cognitive Science, IEEE Transactions on Evolutionary Computation, Information Sciences, Frontiers in Human Neuroscience,} etc.
+He has published ~30 papers in journals such as \emph{Artificial Intelligence, Applied Intelligence, Biological Cybernetics, Cognitive Science, Cognitive Systems Research, Computers & Operations Research, IEEE Transactions on Evolutionary Computation, Information Sciences, Frontiers in Human Neuroscience,} among others.
 \end{IEEEbiography}

### 2019-07-02 Chapters/sdm-autocorrelation.tex
macbookpro compiling
footnotes

@@
 
-\citet{Kanerva1994}
+Explain equator distance\footnote{explain equator distance}.
 
@@
 Neurons have serious authoritarian tendencies, bossing around --- generally electrifying --- the hopeless innocent cells around them just because.}
+
+\subsection{Bitmatches at $i$.}
 
@@
 Each bit of a hard location does indeed have a small pull bias.
-What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0---an intuition developed in Figure \ref{fig:bitmatchbias}.
+What is meant by this is that each particular dimension has a small preference toward positive values if its address bit is set to 1, and negative values if set to 0 --- an intuition developed in Figure \ref{fig:bitmatchbias}.
 
-\begin{figure}[!h]
+TRANSITION\footnote{Transition to what's coming next.}
+
+\begin{figure}[!ht]
 \centering
@@
 \begin{lemma}
-Let $r$ be the access radius given that $f$ percent of the hard locations are activated. Then, $\lim_{n \rightarrow \infty} r/n = 1/2$.
+Let $r$ be the access radius given that $f$ percent of the hard locations are activated.
+Then, $\lim_{n \rightarrow \infty} r/n = 1/2$\footnote{AS WRITTEN, this cannot be right.
+This is saying that, as the number of dimensions explodes, o raio sobre o número de dimensões converge p/ 1/2?
+Não, converge para zero.
+Temos que re-escrever.}.
 \label{thm:sdm-access-radius}
@@
 
-As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 0.5, $B(n, 0.5)$.
-For $n$ sufficiently large, the Binomial distribution can be approximated by a Normal distribution, i.e., $B(n, 0.5) \rightarrow \mathcal{N}(\mu = n/2, \sigma^2 = n/4)$.
+As the bits of the hard locations' addresses are randomly chosen, the distance between two hard locations follow a Binomial distribution with $n$ samples and probability 1/2, $B(n, 1/2)$.
+For $n$ sufficiently large, the Binomial distribution can be approximated by a Normal distribution, i.e., $B(n, 1/2) \rightarrow \mathcal{N}(\mu = n/2, \sigma^2 = n/4)$.
 
-Let $\Phi(x)$ be the cdf of the standard normal distribution. Let $z = \frac{r - n/2}{\sqrt{n}/2}$. Thus, $P(d(x, y) \le r) = \Phi(z)$. As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$.
+Let $\Phi(x)$ be the cdf of the standard normal distribution.
+Let $z = \frac{r - n/2}{\sqrt{n}/2}$.
+\footnote{$z$ is, hence, measuring the number of standard deviations from the average in the Gaussian, as $r$ varies.}  Thus, $P(d(x, y) \le r) = \Phi(z)$.
+As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$\footnote{WTF?
+$f$ isn't a percentage here, is it?}.
 
@@
 \end{proof}
+
+so what?\footnote{WHAT DOES THIS MEAN?}
+
+what now?\footnote{AND WHAT HAPPENS NOW?}
 
@@
 
-Another way to prove is to divide into two cases:
-
-Suppose that $n$ is an even integer, then,
+\bigskip
+Another way to prove is to divide into two cases:  Suppose that $n$ is an even integer, then,
 
@@
 
-In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has reduced significantly as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
+In Figure \ref{fig:sdm-10kbits-figure-7-3}, with $n=10,000$ and $r=4,845$, we can notice that the autocorrelation has reduced significantly---as predicted by Theorem \ref{thm:sdm-autocorrelation-convergence}.
 In fact, in this case, using Lemma \ref{thm:sdm-T1-n-large}, $P(x_i = y_i | d(x, y) \le r) = 0.516876$.
@@
 
-So far we have looked only at a single pair of bitstrings, the probability of a single bitmatch between bitstrings within the access radius distance.
+So far we have looked only at a single pair of bitstrings: the probability of a single bitmatch between bitstrings within the access radius distance.
 Now let us consider the number of activated hard locations exhibiting this bitmatch.
@@
 Then, $Z = \sum_{i=1}^{h} X_i$, where $X_i \sim \text{Bernoulli}(p)$, where $p = P(x_i = y_i | d(x, y) \le r)$.
+\footnote{TRANSITION!!!}
 
@@
 As, in our case, $P(973 < h < 1170) = 0.997$, by the central limit theorem, $Z$ may be approximated by a normal distribution.
-
-See Figure \ref{fig:sdm-same-bit-histogram} for a comparison between the theoretical model and a simulation.
+Figure \ref{fig:sdm-same-bit-histogram} provides a comparison between the theoretical model and a simulation.
 

### 2019-07-17 paperIEEE.tex
footnotes
Papaer
wtf

@@
 
-\section{Neurons as pointers}
+\subsection{Neurons as pointers}
 
@@
 
-\section{Read operation}
+\subsection{Read operation}
 
@@
 In a word, each bit is chosen according to all written bitstrings in all hard locations, being equal to the bit more appeared.
-Table \ref{tab:kanerva-read} shows an example of Kanerva's read result bitstring.
 
@@
 The first decision choice involved in building a system with SDM regards the size of $n$, the number of dimensions of the memory.
-Here, complementing the work of XXXXXXXX, we show that as $n$ grows, autocorrelation tends to $0$.
+Here, complementing the work of XXXXXXXX\footnote{check this}, we show that as $n$ grows, autocorrelation tends to $0$.
 But recall that as $n$ grows, the size of the critical distance also tends to $0$.
@@
   {
-  \includegraphics[width=1in,height=1.25in,clip]{./Chapters/Linhares.jpg}
+  \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./Chapters/Linhares.jpg}
   }

### 2019-07-18 Chapters/sdm-autocorrelation.tex
Papaer
wtf
quase draft

@@
 
-TRANSITION\footnote{Transition to what's coming next.}
-
 \begin{figure}[!ht]
@@
 Consider, for instance, the extremes \{0, $n$\}.
-Let us start with $n$ (in our case, $n=1000$ dimensions): given the information that $d(x,y)\le n$, the probability of a bitmatch in dimension $i$ is $1/2$; as $\forall x,y \in \{0, 1\}^N, d(x,y)\le n$.
-At the other extreme, consider that we have the information that $d(x,y)\le 0$: in this case $x=y$ and the probability of a bitmatch, in any dimension, is 1.
-The autocorrelation hence drops monotonically until convergence at $1/2$ as the distance grows.
-Numerically, our results converge to precisely $1/2$ only after $d(x,y)\ge 607$ for $n=1000$; $d(x,y)\ge 5332$ for $n=10000$, and $d(x,y)\ge 83$ for $n=100$.
-This will be seen in Lemma \ref{thm:sdm-access-radius}.
+Let us start with $n$ (in our case, $n=1000$ dimensions): given the information that $d(x,y)\le n$, the probability of a bitmatch in dimension $i$ is $p=1/2$; as $\forall x,y \in \{0, 1\}^N, d(x,y)\le n$.
+At the other extreme, consider that we have the information that $d(x,y)\le 0$: in this case $x=y$ and the probability of a bitmatch, in any dimension, is $p=1$.
+The autocorrelation hence drops monotonically until convergence at $p=1/2$ as the distance grows.
 \label{fig:bitmatchbias}}
@@
 
+A natural question is what happens to this pull bias as $n$ grows?  Let $f = P(d(x,y)\le r) = 2^{-n}\sum_{k=0}^{r}{\binom{n}{k}}$.  Then, we have:
+
 \begin{lemma}
-Let $r$ be the access radius given that $f$ percent of the hard locations are activated.
-Then, $\lim_{n \rightarrow \infty} r/n = 1/2$\footnote{AS WRITTEN, this cannot be right.
-This is saying that, as the number of dimensions explodes, o raio sobre o número de dimensões converge p/ 1/2?
-Não, converge para zero.
-Temos que re-escrever.}.
+Let $r$ be the access radius given that $f$ is the ratio of the hard locations activated. Then, $\lim_{n \rightarrow \infty} r/n = 1/2$.
 \label{thm:sdm-access-radius}
@@
 Let $z = \frac{r - n/2}{\sqrt{n}/2}$.
-\footnote{$z$ is, hence, measuring the number of standard deviations from the average in the Gaussian, as $r$ varies.}  Thus, $P(d(x, y) \le r) = \Phi(z)$.
-As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$\footnote{WTF?
-$f$ isn't a percentage here, is it?}.
+$z$ is, hence, measuring the number of standard deviations from the average in the Gaussian, as $r$ varies\footnote{MARCELO, CHEQUE ESTA ULTIMA FRASE PLEASE?}.
+Thus, $P(d(x, y) \le r) = \Phi(z)$.
+As $f = P(d(x, y) \le r)$, then, $f = \Phi(z)$ for our fixed $r$}.
 
@@
 
-so what?\footnote{WHAT DOES THIS MEAN?}
-
-what now?\footnote{AND WHAT HAPPENS NOW?}
+As seen on Fig.
+\ref{fig:bitmatchbias}, the bias disappears as $r$ grows.
+Numerically, our results converge to precisely $p=1/2$ only after $d(x,y)\ge 83$ for $n=100$, $d(x,y)\ge 607$ for $n=1000$; $d(x,y)\ge 5332$ for $n=10000$.
+This can be seen in Lemma \ref{thm:sdm-access-radius}.
 
@@
 In our case, $\left[ \text{acc}_i|\eta_i=1 \right] \sim \mathcal{N}(\mu = 128.62, \sigma^2 = 12865.69)$, and $\left[ \text{acc}_i|\eta_i=0 \right] \sim \mathcal{N}(\mu = -128.62, \sigma^2 = 12865.69)$.
-See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} has propagated to these images.
+See Figure \ref{fig:sdm-read} --- we can notice that the variance issue from Figure \ref{fig:sdm-read-sums} propagates directly here.
 
@@
 
-In order to check this probability, I have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculate the distance from the result of a single read.
-As the $P(wrong) = 0.22037$, I expected to get an average distance of 220.37 with a standard deviation of 13.10.
+In order to check this probability, we have run a simulation reading from 1,000 random bitstrings (which have never been written into memory) and calculated the distance from the result of a single read.
+As the $P(wrong) = 0.22037$, we expected to get an average distance of 220.37 with a standard deviation of 13.10.
 See Figure \ref{fig:sdm-read-random-bs} for the comparison between the simulated and the theoretical outcomes.
@@
   \includegraphics[width=0.5\textwidth]{./images02/autocorrelation/single-read-dist.png}
-
   \caption{New distance after a single read operation in a bitstring $\eta_d$, which is $d$ bits away from $\eta$.
@@
 Notice that when $d \ge 520$, the intersection between $\eta$ and $\eta_d$ is zero, which means there is only random bitstrings written into the activated hard locations.
-The distance 220 equals $1000 \cdot 0.220$ which is the probability find in Figure \ref{fig:sdm-read-random-bs}.}
+The distance 220 equals $1000 \cdot 0.220$ which is the probability found in Figure \ref{fig:sdm-read-random-bs}.}
   \label{fig:sdm-single-read-dist}

### 2019-07-18 paperIEEE.tex
quase draft

@@
 
-\caption{Activated addresses inside access \protect \\
-radius $r$ around the center address.\label{fig-addresses-inside-access-radius}}
+\caption{A planar projection of the space illustrates some of the mechanisms.  Activated addresses inside access radius $r$ around the center address.\label{fig-addresses-inside-access-radius}}
 \end{figure}
@@
 A tip-of-the-tongue psychologically happens when you know that you know, but you can't say what exactly it is.
-In SDM mathematical model, a tip-of-the-tongue process takes infinite time to converge.
-\citet{Kanerva1988} called this $x$ distance, where the read's output averages $x$, the critical distance.
+In SDM's mathematical model, a tip-of-the-tongue process takes infinite time to converge.
+\citet{Kanerva1988} called the distance in which the read's output averages $x$, the critical distance.
 Intuitively, it is the distance from which smaller distances converge and greater distances diverge.
@@
 
+It was previously believed that reading at a distance $d=n/2$ would retrieve items at the equator distance, but this is not necessarily true.
+We have shown the subtle, but previously unnoticed, effect of autocorrelation in sparse distributed memory.
+We have also derived excellent analytical approximations of the system's behavior.
+In this concluding section, some comments i) on the implications of this work, on ii) Future research, and iii) on the availability of the code are offered.
+
 \subsection{Effects on system-wide properties}
@@
 The first decision choice involved in building a system with SDM regards the size of $n$, the number of dimensions of the memory.
-Here, complementing the work of XXXXXXXX\footnote{check this}, we show that as $n$ grows, autocorrelation tends to $0$.
+Here, complementing the work of \cite{sjodin_improving_1995, sjodin_convergence_1998} in Jaeckel's SDM design\cite{jaeckel1989class}, we show that as $n$ grows, autocorrelation tends to $0$.
 But recall that as $n$ grows, the size of the critical distance also tends to $0$.
@@
 BPs, System1 vs System2, HRRs, etc.
+
+\begin{table}[ht]
+\begin{tabular}{|p{3.5cm}|p{4cm}|}  %{|c|c|}
+\hline
+System 1	& System 2\tabularnewline
+\hline
+\hline
+Unconscious Reasoning	& Conscious Reasoning\tabularnewline
+\hline
+Implicit	& Explicit\tabularnewline
+\hline
+Automatic	& Controlled\tabularnewline
+\hline
+Low Effort	& High Effort\tabularnewline
+\hline
+Large Capacity	& Small Capacity\tabularnewline
+\hline
+Rapid	& Slow\tabularnewline
+\hline
+Default Process	& Inhibitory\tabularnewline
+\hline
+Associative	& Rule-Based\tabularnewline
+\hline
+Contextualized	& Abstract\tabularnewline
+\hline
+Domain Specific	& Domain General\tabularnewline
+\hline
+Evolutionarily Old	& Evolutionarily Recent\tabularnewline
+\hline
+Nonverbal	& Linked to language\tabularnewline
+\hline
+Includes recognition, perception, orientation	& Includes rule following, comparisons, weighing of options\tabularnewline
+\hline
+Modular Cognition	& Fluid Intelligence\tabularnewline
+\hline
+Independent of working memory	& Limited by working memory capacity\tabularnewline
+\hline
+Non-Logical	& Logical\tabularnewline
+\hline
+Parallel	& Serial\tabularnewline
+\hline
+\end{tabular}
+
+\caption{Dual process theories.}
+
+\label{table:twosys}
+
+\end{table}
+
+\begin{figure}[ht!]
+  \centering
+
+  \subfloat[Bongard problem 71]{\includegraphics[width=0.35\textwidth]{./images02/Bongard-problemp071.png}}
+
+  \subfloat[Bongard problem 72]{\includegraphics[width=0.35\textwidth]{./images02/Bongard-problemp072.png}}
+
+  \subfloat[Bongard problem 73]{\includegraphics[width=0.35\textwidth]{./images02/Bongard-problemp073.png}}
+
+  \subfloat[Bongard problem 74]{\includegraphics[width=0.35\textwidth]{./images02/Bongard-problemp074.png}}
+
+%  \subfloat[Bongard problem 75]{\includegraphics[width=0.35\textwidth]{./images02/Bongard-problemp075.png}}
+
+%  \subfloat[Bongard problem 76]{\includegraphics[width=0.35\textwidth]{./images02/Bongard-problemp076.png}}
+
+  \caption{Bongard problems 71 - 74, courtesy of Dr.
+Harry Foundalis\cite{foundalis_phaeaco:_2006, douglas1980godel, bongard_recognition_1968, Linhares2000}.
+What distinguishes the boxes on the left hand side to those on the right hand side?
+This is the quintessential `system-2' type of problem that current lies beyond AI's capabilities.
+The reader should note that here, for each class, $n=6$, a hard task for Machine Learning.}
+
+  \label{fig:bongard-problems-74}
+\end{figure}
 
@@
   }
-  ]
-{Alexandre Linhares} received the Ph.D.
+  ]{Alexandre Linhares} received the Ph.D.
 degree from the National Institute of Space Research, 2001.

### 2019-09-23 Chapters/sdm-autocorrelation.tex
init-paper-softwareX

@@
 Let $\Phi(x)$ be the cdf of the standard normal distribution.
-Let $z = \frac{r - n/2}{\sqrt{n}/2}$.
-$z$ is, hence, measuring the number of standard deviations from the average in the Gaussian, as $r$ varies\footnote{MARCELO, CHEQUE ESTA ULTIMA FRASE PLEASE?}.
+If $z = \frac{r - n/2}{\sqrt{n}/2}$, $z$ is, hence, measuring the number of standard deviations from the average in the Gaussian\footnote{MARCELO, CHEQUE ESTA ULTIMA FRASE PLEASE?}.
 Thus, $P(d(x, y) \le r) = \Phi(z)$.

### 2019-09-23 SoftwareX/SoftwareX_article_template.tex
init-paper-softwareX

--- 2018-08-02 Chapters/Chapter02.tex
@@
-% !TEX root = ../partial-sdm.tex
+%%
+%% Copyright 2007, 2008, 2009 Elsevier Ltd
+%%
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%%
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%%
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
 
-\chapter{Introduction}
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%% SP 2008/03/01
 
-\bigskip
+\documentclass[preprint,12pt, a4paper]{elsarticle}
 
-\begin{flushright}{\slshape
-    {How is memory gradually built up during one's conscious, or\\
-    even unconscious, life and thought?  My guess is that everything \\
-    we experience is classified and registered on very many parallel \\
-    channels in different locations} \\ \medskip
-    --- Stanislaw Ulam}
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
 
-    \bigskip
-    \bigskip
-    {\slshape Pentti Kanerva's memory model was a revelation for me: \\
-it was the very first piece of research I had ever come across\\
-that made me feel I could glimpse the distant goal of understanding \\
-how the brain works as a whole. It gave me a concrete sense for\\
-how familiar mental phenomena could be thought of as distributed\\
-patterns of micro-events, thanks to beautiful mathematics. \\ \medskip
---- Douglas Hofstadter}
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
 
-\end{flushright}
-\bigskip
-\bigskip
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
 
-Suddenly, you feel the danger.
-You tighten your grip on the child's hand and step up your walking.
-In these large cities, walking on a dark alley was never recommended, but somehow that's what you have done --- and now everything you can think of is to get back to safety.
-You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exactly is causing your fear.
-One just feels it.
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
 
-We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind; a general lack of activity; that very feeling that something isn't right...
-without knowing precisely what isn't right, or even what `right' means --- recalling past experiences from memory and thus generating the feeling.
-Our memory is able to make a parallel between previous experiences and the clues.
-Although one has never been in the exactly same situation, one's brain involuntarily, through analogy with previous experiences, recognizes the possibility of danger.
-The whole process happens so fast it feels like it only took a single indivisible unit of psychological time.
-A single cycle, to use the computing word.
+\usepackage{float}
+\restylefloat{table}
 
-As much as we can rationalize our ability to understand the why of a feeling like this, it seems unlikely that we will be able to nail all reasons down into rules.
-Let us compare this to what a child might do.
-While an adult may respond by discreetly stepping up the walking speed, a child might open her eyes and be surprised and ask, indiscretely, questions about the walking speed...
-Because children do not have the accumulated experience of adults in dealing with large metropolian areas, their behavior differs markedly from ours.
+\journal{SoftwareX}
+
+\begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+\title{Sparse Distributed Memory}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \address[label1]{}
+%% \address[label2]{}
+
+\author{Marcelo Salhab Brogliato}
+\author{Alexandre Linhares}
+
+\address{Getulio Vargas Foundation}
+
+\begin{abstract}
+%% Text of abstract
+Ca. 100 words
+
+\end{abstract}
+
+\begin{keyword}
+%% keywords here, in the form: keyword \sep keyword
+Theoretical Neuroscience \sep Artificial Intelligence \sep Machine Learning \sep Memory \sep Cognitive Science
+
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+\end{keyword}
+
+\end{frontmatter}
+
+\section*{Required Metadata}
+\label{}
+ı
+\section*{Current code version}
+\label{}
+
+Ancillary data table required for subversion of the codebase. Kindly replace examples in right column with the correct information about your current code, and leave the left column as it is.
+
+\begin{table}[H]
+\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
+\hline
+\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this column} \\
+\hline
+C1 & Current code version & v.1.6 \\
+\hline
+C2 & Permanent link to code/repository used for this code version & $https://github.com/msbrogli/sdm-framework/releases/tag/v1.6$ \\
+\hline
+C3 & Code Ocean compute capsule & ????????? For example: $https://codeocean.com/2017/07/30/neurospeech-colon-an-open-source-software-for-parkinson-apos-s-speech-analysis/code$\\
+\hline
+C4 & Legal Code License   & GPL-2.0 \\
+\hline
+C5 & Code versioning system used & git \\
+\hline
+C6 & Software code languages, tools, and services used & C, python, OpenCL, Docker \\
+\hline
+C7 & Compilation requirements, operating environments & python (anaconda), , libbsd \\
+\hline
+C8 & If available Link to developer documentation/manual & For example: $https://buildmedia.readthedocs.org/media/pdf/sdm-framework/stable/sdm-framework.pdf$ \\
+\hline
+C9 & Support email for questions & linhares@sdm.ai \\
+\hline
+\end{tabular}
+\caption{Code metadata (mandatory)}
+\label{}
+\end{table}
+
+\linenumbers
+
+%% main text
+
+The permanent link to code/repository or the zip archive should include the following requirements:
+
+README.txt and LICENSE.txt.
+
+Source code in a src/ directory, not the root of the repository.
+
+Tag corresponding with the version of the software that is reviewed.
+
+Documentation in the repository in a docs/ directory, and/or READMEs, as appropriate.
+
+\section{Motivation and significance}
+\label{}
 
@@
 This model is used in all sort of applications due to its incredible ability to closely reflect the human capacity to remember past experiences from the subtlest of clues.
-Applications range from call admission control \citep{kwon_atm_1998, hee-yong_kwon_atm_1997}, to behavior-based robotics \citep{Rajesh1998, mendes2008robot, jockel_sparse_2009}, to noise filtering \citep{meng_modified_2009}, etc.
-
-This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.
-
-\section{Desiderata for a theory of memory}
-
-% TODO Venn Diagram Psychology, Neuroscience, Cognitive modelling
-% TODO Graph between these disciplines and how they might interact.
-
-Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human long-term memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
+Applications range from call admission control \citep{kwon_atm_1998, hee-yong_kwon_atm_1997}, to behavior-based robotics \citep{Rajesh1998, mendes2008robot, jockel_sparse_2009}, to noise filtering \citep{meng_modified_2009}, among others.
+To understand the breadth of topics that SDM encompasses, consider the following questions:
 
@@
     \item Why do we at times instantly recall an experience; other times we can't recall anything at all; and still other times we get into this strange tip-of-the-tongue situation...
-it is clearly `there'...
-but where?
-(and `what' is `there'?\footnote{Psychologists have documented interesting properties about this state:  it happens, for instance, around every 1-week cycle; and it happens mostly over proper names, subjects have access to the first letter and to the number of syllables better than chance, etc.
-Also interesting is their clever way of triggerring the state: through definitions of rarely-used words.
-For some psychological studies about the tip-of-tongue, see \citet{meyer_tip---tongue_1992, brown_review_1991, brown_tip_1966}.}.)
+the memory is clearly `there'...
+but remains innacessible.
     \item How does this recall process work?  What is remembering?
@@
 
-These are some of the questions touched by SDM.
-It is a theory of long-term memory that constructs the process of recall from a very microscopic level of neuron-firings.
-While there are numerous groups working on related research, it seems that most are simply bypassing each other; starting from scratch instead of building on top of what already exists.
-
-\section{The wasted effort of duplicated, ad hoc, work}
-
-As expected for such an important theory, SDM has been applied in many different fields, like pattern recognition \citep{norman2003modeling, rao1995natural}, noise reduction \citep{Meng2009}, handwriting recognition \citep{fan1997genetic}, robot automation \citep{Rajesh1998, mendes2008robot}, and so forth.
-\citet{Linhares2011} has argued that SDM respects the limits of short-term memory \citep{Miller1955, Cowan2001}.
-Kanerva's book has over 1,000 citations in Google Scholar.
-Despite this, there is not a reference implementation which would allow one to replicate the results published in a paper, to check the source code for details, and to improve it.
-Thus, even though intriguing results have been achieved using SDM, it requires counter-productive, duplicate effort from researchers to build on top of previous work.
-
-Our motivation is our own effort to run our models.
-As there is no reference implementation, we had to develop our own and run several simulations to ensure that our implementation was correct and bug-free.
-Thus, we had to deviate from our primary goal --- which was to test our hypothesis and explore the `ideas space' --- and to focus on the implementation itself.
-Furthermore, new members of our research group had to go through different source code developed by former members, in different languages.
-Consider the implementations of SDM present --- as of this writing --- on its wikipedia entry\citep{noauthor_sparse_2018}\footnote{There has recently been a plethora of hardware implementations to minimize energy expenditure and parallelize processes \citep{kang_energy-efficient_2015, kang_-memory_2016, li_hyperdimensional_2016, montagna_pulp-hd:_2018, salahuddin_energy_nodate}.
-While these implementations are extremely interesting, they do not afford the flexibility to experiment that software does.}:
+While these implementations are extremely interesting, they do not afford the flexibility to experiment that software does:
 
@@
 
-This time, however, it will be running on a computer:   anomalies between expected theory and numerical results will appear; implicit, hidden, assumptions will emerge; and reproducible experiments will be conducted.
+\section{Software description}
+\label{}
 
-\chapter{Notation}
+Describedddddd the software in as much as is necessary to establish a vocabulary needed to explain its impact.
 
-\begin{tabular}{cp{\textwidth}}
-  $n$ & Number of dimensions, i.e., $n=1,000$. \\
-  $N$ & Size of the binary space, $|\{0, 1\}^n| = 2^n$. \\
-  $N'$ & Number of hard locations samples from $\{0, 1\}^n$. Its typical value is 1,000,000, as suggested by \citet{Kanerva1988}. \\
-  $H$ & Same as $N'$. \\
-  $r$ & Access radius, i.e., when $n=1,000$ and $N'=1,000,000$, its typical value is $451$. This value is calculated to activate, on average, one-thousandth of $N'$. \\
-  $\eta$ & A bitstring, usually a datum. \\
-  $\eta_x$ & A clue $x$ bits away from $\eta$, i.e., $\text{dist}(\eta, \eta_x) = x$. \\
-  $\xi$ & A bitstring, usually an address. \\
-  $\text{dist}(x, y)$ & Hamming distance between $x$ and $y$. \\
-  $\text{d}(x, y)$ & Same as $\text{dist}(x, y)$.
-\end{tabular}\\
+\subsection{Software Architecture}
+\label{}
 
-\chapter{Sparse Distributed Memory}
-\input{./Chapters/sdm2.tex}
+Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.
 
-\chapter{Framework Architecture}
-\input{./Chapters/sdm-architecture.tex}
+\subsection{Software Functionalities}
+\label{}
 
-\chapter{Results (i): Performance}
-\input{./Chapters/sdm-performance.tex}
+Present the major functionalities of the software.
 
-\chapter{Results (ii): Framework validation}
-\input{./Chapters/sdm-validation.tex}
+\subsection{Sample code snippets analysis (optional)}
+\label{}
 
-\chapter{Results (iii): Loss of neurons}
-In SDM, the data is written distributed among millions of hard locations, which theoretically gives SDM robustness against loss of neurons.
-In other words, SDM should keep converging correctly even when some neurons are dead.
-The question is: how robust it really is?
-How many neurons may die before it starts to forget things?
-These questions have never been addressed before.
+\section{Illustrative Examples}
+\label{}
 
-Looking for answers to these questions, we run simulations in which we kept killing some neurons and checking whether SDM remained converging to a given bitstring or not.
-In these simulations, 10,000 random bitstrings were written to a 1,000-bit SDM with 1,000,000 hard locations, and we choose one of them as our target.
-As the bitstrings were all written exactly once, we may generalize the results.
-The code is available in the ``Resetting hard locations'' notebook \citep{sdmframework}.
+Provide at least one illustrative example to demonstrate the major functions.
 
-As neurons are hard locations in SDM, when we say that a neuron has been killed, we mean that its counters have been zeroed and a new random bitstring address has been assigned.
-During our simulations, no other bitstring has been written after the 10,000.
-Consequently, as their counters will remain zero, it is exactly like ignoring the dead hard locations in the subsequent reading operations.
+Optional: you may include one explanatory video that will appear next to your article, in the right hand side panel.
+(Please upload any video as a single supplementary file with your article.
+Only one MP4 formatted, with 50MB maximum size, video is possible per article.
+Recommended video dimensions are 640 x 480 at a maximum of 30 frames/second.
+Prior to submission please test and validate your .mp4 file at $ http://elsevier-apps.sciverse.com/GadgetVideoPodcastPlayerWeb/verification$.
+This tool will display your video exactly in the same way as it will appear on ScienceDirect.).
 
-\begin{figure}[!p]
-\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-200k.png}
-\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
-It shows that a loss of 200,000 neurons, 20\% of the total, does not seem to affect SDM whatsoever.
-\label{fig:sdm-neuron-death-200k}}
-\end{figure}
+\section{Impact}
+\label{}
 
-In Figure \ref{fig:sdm-neuron-death-200k}, we can notice that SDM is robust up to 200,000 neuron deaths which are 20\% of all hard locations.
-Its robustness is astonishing.
-In fact, SDM begins to be significantly affected by the loss of neurons after 600,000 neuron deaths (Figure \ref{fig:sdm-neuron-death-1m}) and obviously forgets everything when all neurons are dead.
+\textbf{This is the main section of the article and the reviewers weight the description here appropriately}
 
-\begin{figure}[!p]
-\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-1m.png}
-\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
-The more neurons are lost, the smaller the critical distance, i.e., the worse the SDM recall.
-\label{fig:sdm-neuron-death-1m}}
-\end{figure}
+Indicate in what way new research questions can be pursued as a result of the software (if any).
 
-It is interesting that 500,000 neuron deaths have a minor effect on SDM's recall capability (see Figure \ref{fig:sdm-neuron-death-500k}).
-It is analogous to do a hemispherectomy in a person and, after the procedure, the person being able to recall and learn almost just like before.
-In fact, there are clinical reports of children submitted to hemispherectomy who live an almost normal life with minor function problems.
+Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
 
-\begin{figure}[!p]
-\centering\includegraphics[width=\textwidth]{./images02/new-images/sdm-neuron-death-500k.png}
-\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
-Even when 50\% of neurons are dead, SDM recall is barely affected, which is an impressive result and matches with some clinical results of children submitted to hemispherectomy.
-\label{fig:sdm-neuron-death-500k}}
-\end{figure}
+Indicate in what way the software has changed the daily practice of its users (if so).
 
-An important observation is that around 800,000 neuron deaths (80\% of all neurons) the critical distance becomes small, i.e., SDM recall capacity is hugely diminished.
-After 900,000 neuron deaths, the critical distance is zero, and everything has been lost.
+Indicate how widespread the use of the software is within and outside the intended user group.
 
-Although there is some decrease in SDM recall after 600,000 neuron deaths, it is curious that there is a sudden change between 900,000 (90\%) and 1,000,000 (100\%).
-In Figure \ref{fig:sdm-neuron-death-details} we can see the details of this non-linear change.
-Notice that after 950,000 even the exact clue $\eta_0$ does not converge to $\eta$.
+Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).
 
-\begin{figure}[!p]
-\centering\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death.png}
-\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=1,000$ and $H=1,000,000$.
-\label{fig:sdm-neuron-death-details}}
-\end{figure}
+\section{Conclusions}
+\label{}
 
-We run the same simulation for a 256-bit SDM with 1,000,000 hard locations.
-The results were even more surprising, as the 256-bit SDM seems to be more robust to loss of neurons than the 1,000-bit SDM (see Figure \ref{fig:sdm-neuron-death-256bits}).
-Notice that the loss of 50\% of neurons barely affected the 256-bit SDM...
-and it even remained functional when facing an enormous loss of 90\% of neurons!
-What these results seem to imply is not only that the memory is extremely robust, but that a very small expected number of neurons activated in each access radius would preserve the most valueable information.
-This may have implications as to the number of standard deviations to define the access radius, as we will discuss in Section \ref{Magic numbers}.
-Is there a thing such as excessive robustness, and, if so, when does robustness become waste?\footnote{Consider, for instance, error-correcting codes in Information Theory or the analogous computer ECC memory in Electrical Engineering: tradeoffs between robustness and waste must be considered at each design decision; Why should it be different in the modeling of human memory?}
+Set out the conclusion of this original software publication.
 
-\begin{figure}[!p]
-\centering
-\subfloat[Up to 500,000 neuron deaths ]{\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death-256-500k.png}}
+\section{Conflict of Interest}
+Please select the appropriate text:
 
-\subfloat[From 600,000 to 1,000,000 neuron deaths ]{\includegraphics[width=\textwidth]{images02/new-images/sdm-neuron-death-256-1m.png}}
+Potential conflict of interest exists:
+We wish to draw the attention of the Editor to the following facts, which may be considered as potential conflicts of interest, and to significant financial contributions to this work.
+The nature of potential conflict of interest is described below: [Describe conflict of interest]
 
-\caption{This graph shows the SDM's robustness against loss of neurons in a SDM with $n=256$ and $H=1,000,000$.
-\label{fig:sdm-neuron-death-256bits}}
-\end{figure}
+No conflict of interest exists:
+We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
 
-\chapter{Results (iv): Critical distance}
-\input{./Chapters/sdm-critical-distance.tex}
+\section*{Acknowledgements}
+%\label{}
 
-\chapter{Results (v): Generalized read operation}
+Optionally thank people and institutes you need to acknowledge.
 
-Dr.
-Murilo observed that the models of Kanerva's read ($z=1$) and Chada's read ($z=0$) were simple variations of a generalized read with an exponent $z$, which suggests experimenting with different values.
-Mathematically, let $A$ be the set of the counters of the activated hard location, and $c_i$ be the counter of the $i$th bit.
-Then,
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
 
-$$
-s_i = \sum_{c \in A} \frac{c_i}{|c_i|} |c_i|^z
-$$
+%% \section{}
+%% \label{}
 
-The sum of $|c_i|^z$ turns the intermediate values from integers to floating point numbers. Thus, we have developed a specific read operation which stored the intermediate values in double variables.
+%% References:
+%% If you have bibdatabase file and want bibtex to generate the
+%% bibitems, please use
+%%
 
-The results, however, have not yielded performance improvements.
-Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows an evident deterioration, with a smaller critical distance and faster divergence at large-distance reads.
-This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
 
-We understand that the critical distance is an important parameter of SDM.
-The bigger the critical distance, the better, because SDM is able to converge even with farther clues.
-For $z>1$, the bigger the $z$, the smaller the critical distance.
-For $z = 6$, the critical distance almost reaches zero.
+\begin{thebibliography}
+\bibliographystyle{elsarticle-num}
+\bibliography{../mybib02}
 
-It is interesting that Kanerva has proposed $z=1$ without realizing the generalized reading. Even so, he proposed the $z$ with the highest critical distance.
+\end{thebibliography}
 
-\begin{figure}[h!]
-  \centering
-  \subfloat[SDM behavior when $z \in \{0.1, 0.2, 0.3, 0.4, 0.5, 1\}$ ]{\includegraphics[width=\textwidth]{./images02/new-images/iter_z_01-05.png}}
+\section*{Current executable software version}
+\label{}
 
-  \subfloat[SDM behavior when $z \in \{1.5, 3, 4.5, 6\}$ ]{\includegraphics[width=\textwidth]{./images02/new-images/z_15_3_45_6.png}}
+Ancillary data table required for sub version of the executable software: (x.1, x.2 etc.) kindly replace examples in right column with the correct information about your executables, and leave the left column as it is.
 
-  \caption{(a) and (b) show the behavior of a single read.
-As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.
-Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.
-It seems that Kanerva may not have considered that a single read would only `clean' a small number of dimensions \emph{after the critical distance}.
-What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthogonal distance 500 after a number of iterative readings.}
-  \label{fig:murillo-generalization-experiments}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=\textwidth]{./images02/new-images/z_0_1.png}}
-
-  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=\textwidth]{./images02/new-images/iter_z_all_2.png}}
-
-  \caption{(a) and (b) show the behavior of Figure \ref{fig:murillo-generalization-experiments}, now executed with 6-iterative reads.
-What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthogonal distance 500 after a number of iterative readings.}
-  \label{fig:murillo-generalization-experiments-6reads}
-\end{figure}
-
-%\chapter{Results (v): Performance}
-%\input{./Chapters/sdm-performance.tex}
-
-% Application
-\input{./Chapters/sdm-applications.tex}
-
-\chapter{Results (ix): Information-theoretical write operation}
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/info-theory/weights.png}}
-
-  \subfloat[$w(d)$ for the desired range.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region.png}}
-
-  \subfloat[Stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/info-theory/weights-region-stepwise.png}}
-
-  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
-(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
-  \label{fig:info-theory-hypothesis}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Write process weighted by the amount of information contained in the distance between the written bitstring and each hard location \label{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73.png}}
-
-  \subfloat[Zoom in Figure \ref{fig:info-theory-figure73} ]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-figure73-zoom.png}}
-
-  \subfloat[Behavior of weighted write operation according to the distance from the center and the number of items previously stored in the memory]{\includegraphics[width=0.7\textwidth]{./images02/info-theory/shannon-heatmap.png}}
-
-  \caption{Behavior of the critical distance under the information-theoretic weighted write operation when $n=1,000$, $H=1,000,000$ and $r=451$.}
-  \label{fig:info-theory-experiments}
-\end{figure}
-
-My advisor, Alexandre Linhares, has proposed another write operation: an information-theoretical weighted write.
-In it, the sum of the counter's value is weighted based on the distance between each hard location's address and the reading address.
-The logic behind it is to vary the importance of each hard location inside the circle.
-It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.
-Closer hard locations have lower probabilities and therefore should encode more information.
-
-Note that this is not the first time in which a weighted function has been applied to writing in SDM --- \citet{hely1997new} suggest a rather complex spreading model based on floating point signals in the interval [0.05, 1.0] --- they were, however, only able to test their model with 1,000 hard locations.
-
-Consider the following.
-Information Theory \citep{cover2012elements} let us compute the precise amount of information in an event when given its probability $p$, through the measure of \emph{self-information}:
-
-$$
-I(p)= -log_2(p)
-$$
-
-Now, given any two $n$-sized bitstrings, the probability of their Hamming distance being exactly $d$ is given by $P(X=d)= 2^{-n} \binom{n}{d}$, and the probability of it being at most $d$ is:
-
-$$
-P(X\leq d)= 2^{-n} {\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}}
-$$
-
-But we must consider that not all hard locations are activated in each write operation, which changes our probability function. Thus, let $r$ be the access radius then:
-
-\begin{align*}
-P(X = d | X \leq r) &= \frac{P((X = d) \cap (X \leq r))}{P(X \leq r)} \\
-    &= \frac{P(X = d)}{P(X \leq r)}, \quad \text{as $d \leq r$} \\
-    &= \frac{ 2^{-n} \binom{n}{d} }{ 2^{-n} \sum_{i=0}^{r} \binom{n}{i} } \\
-    &= \frac{\binom{n}{d}}{\sum_{i=0}^{r} \binom{n}{i}}, \quad d \leq r
-\end{align*}
-
-And the probability of it being at most $d$ is:
-
-$$
-P(X \leq d | X \leq r) = \frac{\sum_{i=0}^{d} \binom{n}{i}}{\sum_{i=0}^{r} \binom{n}{i}}, \quad d \leq r
-$$
-
-As expected, $P(X \leq d | X \leq r) = 1$ when $d = r$.
-
-Hence the weighted write would, on each hard location, sum (or subtract) using the following weights, as seen in Figure \ref{fig:info-theory-hypothesis}:
-
-$$
-w(d) = -\log_2 \left( P(X = d | X \leq r) \right) = - \log_2 \binom{n}{d} + \log_2 \sum_{i=0}^{r} \binom{n}{i}, \quad d \leq r
-$$
-
-The initial results of this \emph{Shannon write} operation can be seen in Figure \ref{fig:info-theory-experiments} and seem promising.
-It seems that, when $n=1,000$, $H=1,000,000$, $r=451$, and 10,000 written random bitstrings, the critical distance increased from around 221 to around 250.
-This increase may be interpreted as an improvement in SDM, because it would converge to the correct bitstring even for farther bitstrings.
-Note that 29 additional bits imply an attractor area $2^{29}$ times larger than the original.
-This Shannon write may affect memory capacity \citep{chou_capacity_1989, chou_capacity_1988, sjodin_improving_1995} --- possibly increasing it.
-Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, the shape of the attractor may have asymmetries.
-Whereas these are just some initial tests, the idea seems meritorious so far.
-As for future research, we will execute all tests in the thesis and compare this Shannon write with the original Kanerva model.
-
-%It is easy to interpret this weight through a binary tree approach.  How many binary questions would be needed to precisely define a bitstring inside the access radius and exactly $d$ bits away?
-
-%Another possibility would be to use the sum of all distances closer (and less likely) locations within the weighting function $w(d)$,
-
-%$w(d) = -\log_2 \left( 2^{-n} \displaystyle\sum_{i=0}^{d}{\binom{n}{i}} \right) = n - \log_2 \displaystyle\sum_{i=0}^{d}{\binom{n}{i}}$. \\
-
-%This can be seen in \ref{fig:info-theory-sum-hypothesis}.
-
-%and, consequently,
-
-%$p(H\geq n-d)=2^{-n}{\displaystyle\sum_{i=n-d}^{n}{\binom{n}{i}}  }$, \\
-
-%$p(d+1 \leq H \leq n-d-1)=2^n - 2^{1-n}{\displaystyle\sum_{i=0}^{d}{\binom{n}{i}}, \forall d<n/2}$. \\
-
-%\begin{figure}[h!]
-%  \centering
-%  \subfloat[$w_1(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}
-%
-%  \subfloat[$w_1(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}
-%
-%  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}
-
-%  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius.
-(a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
-%  \label{fig:info-theory-hypothesis}
-%\end{figure}
-
-%\begin{figure}[h!]
-%  \centering
-%  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}
-%
-%  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}
-%
-%  \subfloat[stepwise $\left \lfloor{w_2(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}
-%
-%  \caption{SOON TO BE DEPRECATED.
-Shannon write operation:  Computing the sum of low-likelihood signals.
-(a) entirety of the space; (b) region of interest; (c) Fast integer computation through a stepwise function.
-}
-%  \label{fig:info-theory-sum-hypothesis}
-%\end{figure}
-%
-
-\chapter{Conclusion}
-
-\begin{flushright}{\slshape
-
-    \begin{quote}
-    The most exciting phrase to hear in science, the one that heralds new discoveries, is not `Eureka!' (I found it!) but ``That's funny...'' \\
-    — Isaac Asimov
-    \end{quote}
-
-\bigskip
-\bigskip
-
-}
-\end{flushright}
-
-Sparse Distributed Memory is a viable model of human memory, yet it does require researchers to (re-)implement a number of parallel algorithms in different architectures.
-
-We provide a new, open-source, cross-platform, highly parallel framework in which researchers may be able to create hypotheses and test them computationally with minimal effort.
-The framework is well-documented for public release at this time (http://sdm-framework.readthedocs.io), it has already served as the backbone of Chada's Ph.D.
-thesis \citep{chada2016you}.
-The single-line command ``pip install sdm'' will install the framework on posix-like systems, and single-line commands will let users test the framework, generate some of the figures from Kanerva's theoretical predictions in their own machines, and --- if interested enough ---, test their own theories and improve the framework, and the benchmarks used to evaluate the framework, in open-source fashion.
-It is our belief that such work is a necessary component towards accelerating research in this promising field.
-
-Here are interesting questions that have been considered during this work, but have had to be left for future research.
-
-\section{Another `funny thing'...}
-
-Over the course of this work we looked into a `funny thing' that had appeared in my master's thesis:  the graphs of convergence did not reflect the original expectations put forth by Kanerva, and neither my own previous calculations.
-These numerical results pointed us back to the math and to some interesting random variables not studied before.
-
-Another funny thing that has appeared in my master's thesis was that the critical distance collapsed when the network was exclusively writing `random-at-x'.
-This is something I would like to look into but have not had the time to conduct a proper study yet.
-
-\section{Magic numbers}
-
-Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.
-More recently, he suggested the use of 10,000 dimensions, and on personal discussions suggested that this should be a minimum; as he has been concerned in latent semantic analysis and seems to be the proper scale in that application.
-
-Each parameter set choice like this will lead to particular numbers --- many of them emergent ---, such as the access radius size, critical distance, memory capacity, and so forth.
-For example, the number of dimensions $n$ leads to the number of standard deviations available on the space: $2 \sqrt{n}$.
-One then must choose a proper multiple $m\times\sigma$: Kanerva (and most in the literature) have been using $m=3$, as $\mu-3\sigma$ selects approximately $1/1000$ of the space.
-These constants will determine the size of the access radius.
-Then one must choose the number of hard locations, which will determine how many hard-locations are activated on average; and so on and so forth...
-
-One intriguing question here is:  is there a `better' number of dimensions and hard locations?
-If so, can such numbers be better studied algebraically or numerically?
-How should these parameters be compared?
-What are the tradeoffs that should be considered?
-What are the `best' benchmarks possible?
-
-% \subsection{Classification with context using sequences --- for words instead of only letters}
-
-\section{Symmetrical, rapidly accessible, hard locations}
-
-A hypercube with n dimensions can be divided by two hypercubes with $n-1$ dimensions.
-Is there an algorithm that separates the area of each hard location in such a form that there exists a function mapping each bitstring in $\{0,1\}^n$ to the set of hard locations it `belongs to'?
-Though this would break Kanerva's assumption of a randomly yet uniformly distributed set of hard locations --- for a perfectly symmetrical set of hard locations ---, there could be large performance gains if such a mapping function from a bitstring to its corresponding set of nearest hard locations exists.
-Note that others have attempted heuristic approaches to this\footnote{\citet{Kanerva2009} has recently mentioned that `The computationally most efficient implementation of it, by \citet{Karlsson95afast}, is equivalent to the RAM-based WISARD of  \citet{aleksander1982computer}.' Note, however, that Karlsson's model may introduce non-uniform asymmetries in the space, and seems to require a staggering $O(2^n)$ hard-locations.}.
-
-Consider the hypercube with $n$ dimensions.
-We want to select a subset of its vertices with cardinality $2^{20}$ that is symmetrically distributed over the space.
-Afterward, $\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.
-
-A reduction from measuring the distance to $2^{20}$ hard locations to a computation of $2^{10}$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.
-At large scales of computing, the very ability to perform some experiments is a function of sheer performance.
-The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}.
-A little more on this will follow below.
-
-\section{``i'' versus ``l''}
-
-The classification algorithm had some problems classifying the patterns ``i'' and ``l'', due to the low distance between them.
-In this case, SDM could not discriminate the differences --- it has only considered the big picture.
-Although this behavior is close to how humans see things, we also have the ability to zoom and focus on the details, clearly discerning letter ``i'' from letter ``l''.
-
-I have run the classification algorithm under the MNIST database of handwritten digits \citep{deng2012mnist}.
-First, SDM has been trained with the 60,000 training images, and then it classified the 10,000 testing images.
-In these initial tests, the memory has given the correct classification for 79.22\% of the images, which is inferior to the specialized algorithms.
-For instance, in 1998, \citet{lecun1998gradient} have developed algorithms which achieve from 88\% through a linear classifier, to 99.7\% through a convolutional net.
-For a review of algorithms' performance in the MNIST database, see \cite{deng2012mnist}.
-
-Looking into the reason behind images incorrectly classified, I have found that the issue is very related to the ``i'' versus ``l'' issue.
-Some handwritten digits are very close to others, and a ``2'' or a ``7'' may look like a ``1'', for instance.
-So, how can we solve this issue without using anything specific to images?
-Machine learning algorithms use specific techniques to improve performance.
-I would like to unveil a solution psychologically closer to how we behave --- even if that eventually leads to lower performance.
-
-An unexplored idea is to use multiple SDMs which communicate.
-A first SDM would write the whole picture, just like we have done.
-Another SDM would write specific regions of the image, just like our eye focusing on specific regions.
-When reading, they may compose the counters and give a more precise classification.
-
-\section{Deep learning, multiple SDMs --- and the incredible animal behavior of Dr. Linhares}
-
-There is the intention of studying what, if any, capabilities do multiple SDMs have...
-yet it is probably advisable to leave the following comments as originally made by my advisor, Dr.
-Linhares.
-To quote:
-
-\begin{table}[ht]
-\begin{tabular}{|p{4cm}|p{7cm}|}  %{|c|c|}
+\begin{table}[!h]
+\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
 \hline
-System 1	& System 2\tabularnewline
+\textbf{Nr.} & \textbf{(Executable) software metadata description} & \textbf{Please fill in this column} \\
 \hline
+S1 & Current software version & For example 1.1, 2.4 etc. \\
 \hline
-Unconscious Reasoning	& Conscious Reasoning\tabularnewline
+S2 & Permanent link to executables of this version  & For example: $https://github.com/combogenomics/$ $DuctApe/releases/tag/DuctApe-0.16.4$ \\
 \hline
-Implicit	& Explicit\tabularnewline
+S3 & Legal Software License & List one of the approved licenses \\
 \hline
-Automatic	& Controlled\tabularnewline
+S4 & Computing platforms/Operating Systems & For example Android, BSD, iOS, Linux, OS X, Microsoft Windows, Unix-like , IBM z/OS, distributed/web based etc. \\
 \hline
-Low Effort	& High Effort\tabularnewline
+S5 & Installation requirements \& dependencies & \\
 \hline
-Large Capacity	& Small Capacity\tabularnewline
+S6 & If available, link to user manual - if formally published include a reference to the publication in the reference list & For example: $http://mozart.github.io/documentation/$ \\
 \hline
-Rapid	& Slow\tabularnewline
-\hline
-Default Process	& Inhibitory\tabularnewline
-\hline
-Associative	& Rule-Based\tabularnewline
-\hline
-Contextualized	& Abstract\tabularnewline
-\hline
-Domain Specific	& Domain General\tabularnewline
-\hline
-Evolutionarily Old	& Evolutionarily Recent\tabularnewline
-\hline
-Nonverbal	& Linked to language\tabularnewline
-\hline
-Includes recognition, perception, orientation	& Includes rule following, comparisons, weighing of options\tabularnewline
-\hline
-Modular Cognition	& Fluid Intelligence\tabularnewline
-\hline
-Independent of working memory	& Limited by working memory capacity\tabularnewline
-\hline
-Non-Logical	& Logical\tabularnewline
-\hline
-Parallel	& Serial\tabularnewline
+S7 & Support email for questions & \\
 \hline
 \end{tabular}
-
-\caption{Dual process theories.}
-
-\end{table}\label{table:twosys}
-
-\begin{figure}[h!]
-  \centering
-
-  \subfloat[Bongard problem 71]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp071.png}}
-
-  \subfloat[Bongard problem 72]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp072.png}}
-
-  \subfloat[Bongard problem 73]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp073.png}}
-
-  \caption{Bongard problems 71 - 73, courtesy of noted Bongardologist, Dr. Harry Foundalis.  What distinguishes the boxes on the left hand side to those on the right hand side?}
-  \label{fig:bongard-problems-71}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-
-  \subfloat[Bongard problem 74]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp074.png}}
-
-  \subfloat[Bongard problem 75]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp075.png}}
-
-  \subfloat[Bongard problem 76]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp076.png}}
-
-  \label{fig:bongard-problems-74}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-
-  \subfloat[Bongard problem 77]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp077.png}}
-
-  \subfloat[Bongard problem 78]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp078.png}}
-
-  \subfloat[Bongard problem 79]{\includegraphics[width=0.7\textwidth]{./images02/Bongard-problemp079.png}}
-
-  \label{fig:bongard-problems-77}
-\end{figure}
-
-\begin{quote}
-
-    Those deep neural networks...
-alphaGo...
-alphazero...
-very impressive, and to me very right and proper in a sense and also very wrong and misguided in another.
-I have not seriously studied the topic.
-Eric Nichols explained it to me in a napkin, and I watched some youtube videos and browsed the books.
-So what follows, like all great science, is completely based on gossip and hearsay, prejudice and anecdote.
-Or at least that's how I believe all great \emph{hunches} over what is promising and what is a dead-end come from; that aesthetic sense that forms over time and seems to guide us towards what feels beautiful and correct and away from what seems foolish and wasteful...
-
-    So here is my hunch:  \emph{SDM is System 1 and System 1 is SDM} (e.g., as in Table \ref{table:twosys}).
-I haven't seen anyone say this out loud, but Kanerva has pretty much nailed system 1 --- I feel he is at least 90\% right.
-
-    You know those animals that create those burrows underground, creatures like the prairie dogs?
-There, underground, they build all sorts of passageways from here to where they want to go; and in the process they discover all that \emph{can't be passageway}.
-What can't be passageway --- what can't be corridor --- will never be used as passageway, by definition.
-And most of the world can't be passageway.
-Think about the animal behavior: endless dead-end nooks and cranies yet the beast won't give up!
-To me it's all very similar to what these systems seem to be doing...
-it seems they are crafting and molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...
-You have these monstrous networks creating layers upon layers in which they differentiate and distribute the errors throughout all the nooks and cranies of, say, a domain like the game of Go.
-Then, after some biblical number-crunching, it seems that there may be a very subtle way...
-however tortuous, however tight, however improbable, however unconfortable.
-
-    And these machines are exhibiting intelligence.
-The victories in the game of Go seem to have shown that alphaGo did have vague, non-explicit, non-formal, but very strong intuitions about what the nooks and cranies of the Go universe looks like.
-So here is what I like and I find beautiful:  These gigantic clusters of inter-related micro impressions that the system has acquired and can relate to.
-I imagine multiple SDMs exchanging information; gossiping to each other about all the dead-ends out there and trying to find some tiny tortuous pathway, and it feels to me that that's a good and proper way to move forward in research.
-
-    But it feels like a mistake to say that that is thinking.
-It's not.
-It is to me a great model of someone getting so attuned into an area that they are able to simply flow through tasks; in the positive psychology sense of the word flow.
-It's very powerful and very intelligent:  but it is a form of intelligence without thinking.
-Intelligence that can't make leaps from one arena to another; Intelligence that can't work on Bongard problems.
-So what I would fear as a dead-end in research would be to have another AI winter...
-decades of massive, massive number-crunching and machine learning of this type...
-in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figure \ref{fig:bongard-problems-71}).
-Bongard problems demand high-gossip between system 1 and system 2; they demand a high dose of imagination and mental imagery manipulation; they demand discrimination and demand clusterization, among many other things that have been spelled out before by \citet{bongard_recognition_1968, douglas1980godel, Linhares2000} and the landmark, deep, study being \citet{foundalis_phaeaco:_2006}.
-
-    Cognitive Science has to study the \emph{plausible} interaction between system 1 and system 2.
-We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa).
-We need both system 1 and system 2 to be plausible, in the psychological sense and in the neuroscientific sense.
-AlphaGo, for instance, has this wonderful model of the game, reflecting many of the properties of system 1...
-but then falls back into silliness as soon as it turns to the other side of the problem:  it uses montecarlo methods in very efficient fashion to probe some 70,000 thousands of positions per second and go on to to the next probing of this wonderful system 1...
-
-    It works; it is impressive; it is intelligent; but it is not thinking; and it hasn't thought us much about our cognition.
-
-    Hofstadter has nailed the interactions and the content...
-analogies, fluid concepts, things seen through a different perspective, abstract roles, variations on a theme, etc.
-Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.
-Isn't it time to work on System 2 models that are counterparts to SDM --- in fact, isn't it obvious that that's our glaring omission?
-Attaching something as beautiful as SDM to, say, some centralized hyper-efficient montecarlo algorithm is nothing short of heresy against all the gods of science for which the villain shall be put to death without clemency or mercy.
-Seems like the only reasonable path going forward...
-
-\end{quote}
-
-\section{Not a real Conclusion}
-
-% Preview source code for paragraph 8
-
-\begin{table}[ht]
-\begin{tabular}{|p{4cm}|p{7cm}|}  %{|c|c|}
-\hline
-Project & Observation\tabularnewline
-\hline
-\hline
-Source-code  & C, OpenCL, Python (NVIDIA CUDA, FPGAs desirable in the future)\tabularnewline
-\hline
-Step-by-step notebook studies & replication and dissemination of studies\tabularnewline
-\hline
-Documentation of framework & 100+ pages currently\tabularnewline
-\hline
-Open-Source Creative Commons Book & Under consideration (to be created from this thesis)\tabularnewline
-\hline
-Collection of SDM links, results, resources & modeled after `Awesome Machine Learning'\citep{misiti_awesome-machine-learning:_2018} and `Awesome Artificial Intelligence'\citep{lewis_awesome-artificial-intelligence:_2018} lists, community
-managed\tabularnewline
-\hline
-Slides from the Jupyter Notebooks & online course, currently in planning phase\tabularnewline
-\hline
-Videos following the Course Slides  & online course, currently in planning phase\tabularnewline
-\hline
-\end{tabular}
-
-\caption{Desiderata of open-source, creative-common deliverables, longer-term.
-Some of these items are mature for wide dissemination, whereas others are in planning phase.
-For an online SDM course, for instance, we might partner with a Ph.D.
-program in the Teaching of Mathematics and Physics, such as \citet{im/ufrj_programa_nodate}, who are intent on developing and measuring the effects of new educational tools.}
-
+\caption{Software metadata (optional)}
+\label{}
 \end{table}
 
-Let us revisit, in these concluding thoughts, the emphasis employed over speed of computation.
-At first sight, that might seem like a typical objective of efficiency in computer science.
-But we are not only interested in the computer science effects here --- the ambition is different.
-More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  communities form around open-source code, specialists in a particular arena magnify the capabilities of the code, debates and meetings are regularly held, and deep divisive disagreements lead to community splits in multiple potential directions of exploration.
-In a very real sense, given a shared code to work and experiment on, we can see farther.
-
-Beyond speed, I have also strived for \emph{ease of use} and \emph{replicable studies}.
-All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested.
-Most importantly, they also serve as a tutorial for learning about the framework's design, and perhaps as a teaching tool for SDM itself.
-I have generated a Docker image, which makes it even easier to explore the framework by delegating heavy computational work to cloud services.
-After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.
-We invite the reader to take a look and explore a little bit.
-
-The overarching intention here is to not only provide a starting point, but to provide a documented Framework in which SDM research can be conducted.
-Consider having the ability to compare the results of a new (‘forked’) model to the previous `best' (under a particular benchmark set).
-For example, some of the benchmarks that we plan to develop in future research are: how fast is convergence through iterative reading?
-How large is the attractor of the critical distance?
-How well does the system filter noise?
-How well does the system work under the supervised learning task?
-And other authors may be able to improve this benchmark set themselves, as is usual in open source development.
-It is perhaps this facility of ease to build on top of previous work that seems most exciting at this stage.
-
-Consider the misunderstanding concerning the SDM read operation:  Dr.
-Stan Franklin describes Kanerva's read operation in a way that each hard location, at each dimension, provides only a single bit of information to the read operation (instead of Kanerva's full counter).
-We have referred to this modified read operation as Chada read\footnote{Legend has it that my friend \text{\&} colleague, Dr.
-Daniel de Magalhães Chada, along with Linhares, did not consult and re-check with Kanerva's book and only discovered the discrepancy in code and ideas a couple of years afterward.}.
-Having an open, testable, codebase reduces the possibilities of such misunderstandings in the long run.
-Indeed, a high-quality codebase seems to have become a scientific community's form of unequivocally standing behind a consensus.
-For example, the journal \emph{Nature} analyzed the top-100 cited papers in history, to find:
-
-\begin{quote}
-...
-some surprises, not least that it takes a staggering 12,119 citations to rank in the top 100 — and that many of the world’s most famous papers do not make the cut.
-A few that do, such as the first observation of carbon nanotubes (number 36) are indeed classic discoveries.
-But the vast majority describe experimental methods or \emph{software that have become essential in their fields.
-[...] The list reveals just how powerfully research has been affected by computation and the analysis of large data sets.} \\
-\hfill --- \citet{van2014top}, emphasis mine.
-\end{quote}
-
-It is no coincidence that scientific journals such as \emph{BMC Neuroscience}, or the \emph{Journal of Machine Learning Research} have specific sections on open-source software.
-The journal \emph{Neurocomputing} states, bluntly: “software is scientific method by machine”.
-
-Of course, for the skeptical reader who may consider software a less worthy pursuit, there is also new work here.
-The mathematics of the model has been shown to be correct numerically (with a single, small, anomaly); we have shown how to execute unsupervised learning with nothing besides operations original to the SDM; we have studied the generalized Murilo read; we have seen noise filtering; the death of neurons; how information-theory may be of use; and finally, we have reproduced numerous of the original propositions put forth by Kanerva.
-The emphasis might have been on the \emph{breadth of topics}, in detriment of depth here or there.
-The work on, say, reinforcement learning, is most definitely not the last word we will see on the subject, but a challenge left for readers to contemplate.
-But this is due to our research group's enthusiasm for the topic; we do indeed believe that SDM is --- if not correct --- extremely close to a full scientific understanding of human long-term memory.
-If so, it is such a monumental achievement that we want readers to be able to see all of what we see and imagine the vastness of possibilities.
-
-Ralph Waldo Emerson once said: \emph{do not go where the path may lead.
-Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.
-It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.
-Some essays completely shut the door close at the end; this one intends to leave it wide open.
-As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.
-
-\chapter{Appendix}
-%\section{Generating Kanerva's table 7.3}
-
-%\includepdf[
-%  pages=-,
-%  pagecommand={\pagestyle{headings}},
-%  %addtotoc={1,section,1,Quilling Shapes,sec:shapes}
-%]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}
-
-\chapter*{List of Jupyter notebooks}
-
-\begin{enumerate}
-  \item Calculate critical distance.ipynb
-  \item Calculate radius for any SDM.ipynb
-  \item Classification Test 1.ipynb
-  \item Classification Test 2.ipynb
-  \item Classification Test 3-Copy1.ipynb
-  \item Classification Test 3.ipynb
-  \item Critical Distance - New.ipynb
-  \item Critical Distance.ipynb
-  \item Distance between bitstrings.ipynb
-  \item Distances of activated hard-locations.ipynb
-  \item Generates a new SDM.ipynb
-  \item Intersection of activated hard-locations.ipynb
-  \item Kanerva-Table-7 - Experiment.ipynb
-  \item Kanerva-Table-7.3 - Generic Read.ipynb
-  \item Kanerva-Table-7.3 - Multiple dimensions.ipynb
-  \item Kanerva-Table-7.3 - Weighted table write.ipynb
-  \item Kanerva-Table-7.3.ipynb
-  \item Kanerva's Figure 1.2.ipynb
-  \item Noise filter 2.ipynb
-  \item Noise filter A-Z 0-1 Generator.ipynb
-  \item Noise filter with labels.ipynb
-  \item Noise filter.ipynb
-  \item Number of activated hard-locations.ipynb
-  \item Performance test.ipynb
-  \item Reseting hard-locations.ipynb
-  \item Sequences (Kanerva Ch 8).ipynb
-  \item Sequences in one SDM (Kanerva Ch 8).ipynb
-  \item TicTacToe.ipynb
-  \item Weighted operations using information.ipynb
-\end{enumerate}
+\end{document}
+\endinput
+%%
+%% End of file `SoftwareX_article_template.tex'.

### 2019-09-23 paperIEEE.tex
init-paper-softwareX

@@
 The first decision choice involved in building a system with SDM regards the size of $n$, the number of dimensions of the memory.
-Here, complementing the work of \cite{sjodin_improving_1995, sjodin_convergence_1998} in Jaeckel's SDM design\cite{jaeckel1989class}, we show that as $n$ grows, autocorrelation tends to $0$.
+Here, complementing the work of \cite{sjodin_improving_1995, sjodin_convergence_1998} in Jaeckel's SDM design \cite{jaeckel1989class}, we show that as $n$ grows, autocorrelation tends to $0$.
 But recall that as $n$ grows, the size of the critical distance also tends to $0$.
@@
 This is the quintessential `system-2' type of problem that current lies beyond AI's capabilities.
-The reader should note that here, for each class, $n=6$, a hard task for Machine Learning.}
+The reader should note that here, for each class, $n=6$: a hard task for Machine Learning.}
 
@@
 This --> transparency --> replicability -> incremental improvement --> parallel explorations by different teams around the world.
+
+of the immense power of Artificial Intelligence.
 
